### 2015-09-22 imaging_demo/ms.tex
Create first ms.tex file with brief outline
From MNRAS template:
===================

(*) Add mnras.bst and mnras.cls to git repository
(*) Create changelog.txt and add to git repository
(*) Create epic.bib bibliography file and add to git repository
(*) Create a ms.tex file with a skeleton structure using
mnras_template.tex and add to git repository
(*) Latex compile the ms.tex file to create ms.bbl, ms.aux and ms.pdf.
Add these to git repo

% mnras_template.tex
%
% LaTeX template for creating an MNRAS paper
%
% v3.0 released 14 May 2015
% (version numbers match those of mnras.cls)
%
% Copyright (C) Royal Astronomical Society 2015
% Authors:
% Keith T. Smith (Royal Astronomical Society)

% Change log
%
% v3.0 May 2015
%    Renamed to match the new package name
%    Version number matches mnras.cls
%    A few minor tweaks to wording
% v1.0 September 2013
%    Beta testing only - never publicly released
%    First version: a simple (ish) template for creating an MNRAS paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Basic setup. Most papers should leave these options alone.
\documentclass[a4paper,fleqn,usenatbib]{mnras}

% MNRAS is set in Times font. If you don't have this installed (most LaTeX
% installations will be fine) or prefer the old Computer Modern fonts, comment
% out the following line
\usepackage{newtxtext,newtxmath}
% Depending on your LaTeX fonts installation, you might get better results with one of these:
%\usepackage{mathptmx}
%\usepackage{txfonts}

% Use vector fonts, so it zooms properly in on-screen viewing software
% Don't change these lines unless you know what you are doing
\usepackage[T1]{fontenc}
\usepackage{ae,aecompl}

%%%%% AUTHORS - PLACE YOUR OWN PACKAGES HERE %%%%%

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%

% Please keep new commands to a minimum, and use \newcommand not \def to avoid
% overwriting existing commands. Example:
%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%% TITLE PAGE %%%%%%%%%%%%%%%%%%%

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[E-field Parallel Imaging Correlator]{A Generic and Efficient E-field Parallel Imaging Correlator for Next-Generation Radio Telescopes}

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[Thyagarajan et al.]{
Nithyanandan Thyagarajan,$^{1}$\thanks{E-mail: t\_nithyanandan@asu.edu}
Adam P. Beardsley,$^{1}$
Judd D. Bowman$^{1}$
\newauthor
and Miguel F. Morales$^{2}$
\\
% List of institutions
$^{1}$Arizona State University, School of Earth and Space Exploration, Tempe, AZ 85287, USA\\
$^{2}$University of Washington, Department of Physics, Seattle, WA 98195, USA\\
}

% These dates will be filled out by the publisher
\date{Accepted XXX. Received YYY; in original form ZZZ}

% Enter the current year, for the copyright statements etc.
\pubyear{2015}

% Don't change these lines
\begin{document}
\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\maketitle

% Abstract of the paper
\begin{abstract}
Abstract here (250 words)
\end{abstract}

% Select between one and six entries from the list of approved keywords.
% Don't make up new ones.
\begin{keywords}
instrumentation: interferometers -- techniques: image processing -- techniques: interferometric
\end{keywords}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%% BODY OF PAPER %%%%%%%%%%%%%%%%%%

\section{Introduction}

Motivate MOFF from a technical and scientific standpoint. Refer to \citet{del07,mor09,mor11}.

\section{Mathematical Framework}

Refresh the math equivalence between MOFF and FX.

\section{Software Implementation}

Discuss implementation and make code available for public.

Calibration

\section{Verification}

Show examples using simulations

Discuss PSF differences due to slight differences arising out of gridding

Apply it on LWA data

\section{Analysis and Feasibility}
\subsection{Scaling Relations: MOFF vs. FX}
\subsection{Scaling Up}
\subsection{Case Study}
% \subsection{Figures and tables}

% Figures and tables should be placed at logical positions in the text. Don't
% worry about the exact layout, which will be handled by the publishers.

% Figures are referred to as e.g. Fig.~\ref{fig:example_figure}, and tables as
% e.g. Table~\ref{tab:example_table}.

% % Example figure
% \begin{figure}
% 	% To include a figure from a file named example.*
% 	% Allowable file formats are eps or ps if compiling using latex
% 	% or pdf, png, jpg if compiling using pdflatex
% 	\includegraphics[width=\columnwidth]{example}
%     \caption{This is an example figure. Captions appear below each figure.
% 	Give enough detail for the reader to understand what they're looking at,
% 	but leave detailed discussion to the main body of the text.}
%     \label{fig:example_figure}
% \end{figure}

% % Example table
% \begin{table}
% 	\centering
% 	\caption{This is an example table. Captions appear above each table.
% 	Remember to define the quantities, symbols and units used.}
% 	\label{tab:example_table}
% 	\begin{tabular}{lccr} % four columns, alignment for each
% 		\hline
% 		A & B & C & D\\
% 		\hline
% 		1 & 2 & 3 & 4\\
% 		2 & 4 & 6 & 8\\
% 		3 & 5 & 7 & 9\\
% 		\hline
% 	\end{tabular}
% \end{table}

\section{Conclusions}
\section*{Acknowledgements}

The Acknowledgements section is not numbered. Here you can thank helpful
colleagues, acknowledge funding agencies, telescopes and facilities used etc.
Try to keep it short.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%% REFERENCES %%%%%%%%%%%%%%%%%%

% The best way to enter references is to use BibTeX:

\bibliographystyle{mnras}
\bibliography{epic} % if your bibtex file is called example.bib

% Alternatively you could enter them by hand, like this:
% This method is tedious and prone to error if you have lots of references
% \begin{thebibliography}{99}
% \bibitem[\protect\citeauthoryear{Author}{2012}]{Author2012}
% Author A.~N., 2013, Journal of Improbable Astronomy, 1, 1
% \bibitem[\protect\citeauthoryear{Others}{2013}]{Others2013}
% Others S., 2012, Journal of Interesting Stuff, 17, 198
% \end{thebibliography}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%% APPENDICES %%%%%%%%%%%%%%%%%%%%%

\appendix

\section{Some extra material}

If you want to present additional material which would interrupt the flow of the main paper,
it can be placed in an Appendix which appears after the list of references.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Don't change these lines
\bsp	% typesetting comment
\label{lastpage}
\end{document}

% End of mnras_template.tex

### 2015-10-13 cal_demo/ms.tex
initial commit of cal paper, moving shared files up one directory

--- 2015-09-22 imaging_demo/ms.tex
@@
 % Basic setup. Most papers should leave these options alone.
-\documentclass[a4paper,fleqn,usenatbib]{mnras}
+\documentclass[a4paper,fleqn,usenatbib]{../mnras}
 
@@
 % Keep the title short and informative.
-\title[E-field Parallel Imaging Correlator]{A Generic and Efficient E-field Parallel Imaging Correlator for Next-Generation Radio Telescopes}
+\title[E-field Parallel Imaging Calibration]{An Efficient E-field Parallel Imaging Calibration Algorithm for Next-Generation Radio Telescopes}
 
@@
 % If you need two or more lines of authors, add an extra line using \newauthor
-\author[Thyagarajan et al.]{
-Nithyanandan Thyagarajan,$^{1}$\thanks{E-mail: t\_nithyanandan@asu.edu}
-Adam P. Beardsley,$^{1}$
+\author[Beardsley et al.]{
+Adam P. Beardsley,$^{1}$\thanks{E-mail: Adam.Beardsley@asu.edu}
+Nithyanandan Thyagarajan,$^{1}$
 Judd D. Bowman$^{1}$
@@
 
-Motivate MOFF from a technical and scientific standpoint. Refer to \citet{del07,mor09,mor11}.
+Motivate problem of calibrating direct imagers. Reference Nithyas paper. \citep{mor11}
 
@@
 
-Refresh the math equivalence between MOFF and FX.
+Introduce notation, describe measurement equation.
 
-\section{Software Implementation}
+\section{Simulation}
 
-Discuss implementation and make code available for public.
-
-Calibration
+Describe simulation
 
@@
 
-Show examples using simulations
+Connect to either cramer-rao or FX solutions in some way
 
-Discuss PSF differences due to slight differences arising out of gridding
-
-Apply it on LWA data
-
-\section{Analysis and Feasibility}
-\subsection{Scaling Relations: MOFF vs. FX}
-\subsection{Scaling Up}
-\subsection{Case Study}
 % \subsection{Figures and tables}
@@
 
-\bibliographystyle{mnras}
-\bibliography{epic} % if your bibtex file is called example.bib
+\bibliographystyle{../mnras}
+\bibliography{../epic} % if your bibtex file is called example.bib
 

### 2015-10-16 imaging_demo/ms.tex

@@
 %\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
+\newcommand{\Nant}{N_{\text{ant}}} 
+\newcommand{\Npix}{N_{\text{pix}}}
 
@@
 
+Radio astronomy is entering an era in which interferometers of hundreds to thousands of individual antennas are needed to achieve desired survey speeds.
+Nowhere is this more apparent than at radio frequencies below 1.4 GHz.
+The study of the history of hydrogen gas throughout the universe's evolution is pushing technology development towards arrays of low-cost antennas with large fields of view and densely packed apertures.
+Similarly, the search for transient objects and regular monitoring of the time-dependent sky is driving instruments in the same direction.
+A number of new telescopes are under development around the world based on this new paradigm, including the Murchison Widefield Array (MWA, \citealt{tin13}), the Precision Array for Probing the Epoch of Reionization (PAPER, \citealt{par10}), the Hydrogen Epoch of Reionization Array (HERA, \citealt{pob14}), the LOw Frequency ARray (LOFAR, \citealt{dev09}), the Canadian Hydrogen Intensity Mapping Experiment (CHIME, \citealt{ban14}), the Long Wavelength Array (LWA, \citealt{ell13}), and the low frequency component of the Square Kilometer Array Low Frequency Aperture Array (SKA-Low \citealt{mel13}).
+
+This paradigm shift requires a fundamentally new approach to the design of digital correlators \citep{lon00}.
+Modern correlators calculate the cross-power correlation between all antenna pairs in many narrow frequencies, forming \emph{visibilities}, the traditional fundamental measurement of radio interferometers.
+The computational requirements for a modern FX correlator scale with the number of antenna pairs, or the square of the number of antennas $\sim \Nant^2$ \citep{bun04}.
+For this reason traditional correlators have difficulty scaling to thousands of antennas.
+As an example, the full HERA correlator for 352 dishes with 200 MHz of bandwidth requires 212 trillion complex multiplies and adds per second (TMACS).
+Future arrays with thousands of collecting elements will require orders of magnitude more computation, making the correlator the dominant cost.
+
+For certain classes of radio arrays there is an alternative to the FX correlator that can lower the computational burden by directly performing a spatial fast Fourier transform (FFT) on the electric fields measured by each antenna in the array at each time step, removing the cross-correlation step.
+This relieves the computational scaling from the harsh $\Nant^2$ to the more gentle envelope of $\Npix\log\Npix$, where $\Npix$ is the number of pixels in the Fourier transform \citep[e.g.][]{mor11,teg09,teg10}.
+This architecture is often referred to as a ``direct imaging" correlator because it eliminates the intermediary cross-correlation data products of the FX and earlier lag correlators, but instead directly forms images from the electric field measurements.
+
+Direct imaging correlators have begun to be explored on deployed arrays including the Basic Element for SKA Training II (BLAST-2) array \citep{fos14}, the Omniscope \citep{zhe14}, and an earlier incarnation at higher frequencies with the intent of pulsar timing \citep{oto94, dai00}.
+However, each of these examples make assumptions about the redundancy of the array layout, and require the collecting elements are identical.
+On the other hand, the MOFF algorithm achieves the same $\Npix \log \Npix$ computational scaling without placing any restriction on antenna placement, can accommodate non-identical beam patterns, and is a provably optimal mapping \citep{mor11}.
+This algorithm uses the antenna beam patterns to grid the electric field measurements to a regular grid in the software holography/A-transpose fashion \citep{mor09, bha08, teg97} before performing the spatial FFT.
+This process has been shown to theoretically produce a data product identical to images produced from the traditional FX correlator.
+
+Here we present the first software implementation of the MOFF correlator, and announce the public release of the E-field Parallel Imaging Correlator (EPIC) code.
+We begin with a technical description of the algorithm in \S \ref{sec:math}, then discuss our particular implementation in \S \ref{sec:software}.
+We then verify the output data quality from our code in \S \ref{sec:verify} by presenting simulated images from both the EPIC correlator and comparing to a simulated FX correlator.
+We also demonstrate the performance with real-world data from the LWA.
+In \S \ref{sec:analysis} we analyze the scaling relationships of the algorithm.
+We identify specific array design classes where the EPIC correlator is computationally more efficient than the FX algorithm.
+We conclude and discuss future research prospects in \S\ref{sec:conclusions}
+
 Motivate MOFF from a technical and scientific standpoint. Refer to \citet{del07,mor09,mor11}.
 
-\section{Mathematical Framework}
+\section{Mathematical Framework}\label{sec:math}
 
@@
 
-\section{Software Implementation}
+\section{Software Implementation}\label{sec:software}
 
@@
 
-\section{Verification}
+\section{Verification}\label{sec:verify}
 
@@
 
-\section{Analysis and Feasibility}
+\section{Analysis and Feasibility}\label{sec:analysis}
 \subsection{Scaling Relations: MOFF vs. FX}
@@
 
-\section{Conclusions}
+\section{Conclusions}\label{sec:conclusions}
 \section*{Acknowledgements}

### 2015-10-22 imaging_demo/ms.tex
Add content under sections, update bibliography
From 011cd95 (master):
=====================

In epic.bib,

(*) Add references van Cittert (1934), Zernike (1938), Thompson
et al.(2001), Maneewongvatana and Mount (1999)
(*) Rearrange order of references

In ms.tex,

(*) Start adding content under section "Mathematical Framework"
(*) Start adding content under section "Software Implementation"
(*) Start adding content under section "Verification"

@@
 \newcommand{\Npix}{N_{\text{pix}}}
+\newcommand{\dif}{\mathrm{d}}
 
@@
 Similarly, the search for transient objects and regular monitoring of the time-dependent sky is driving instruments in the same direction.
-A number of new telescopes are under development around the world based on this new paradigm, including the Murchison Widefield Array (MWA, \citealt{tin13}), the Precision Array for Probing the Epoch of Reionization (PAPER, \citealt{par10}), the Hydrogen Epoch of Reionization Array (HERA, \citealt{pob14}), the LOw Frequency ARray (LOFAR, \citealt{dev09}), the Canadian Hydrogen Intensity Mapping Experiment (CHIME, \citealt{ban14}), the Long Wavelength Array (LWA, \citealt{ell13}), and the low frequency component of the Square Kilometer Array Low Frequency Aperture Array (SKA-Low \citealt{mel13}).
+A number of new telescopes are under development around the world based on this new paradigm, including the Murchison Widefield Array (MWA, \citealt{tin13}), the Precision Array for Probing the Epoch of Reionization (PAPER, \citealt{par10}), the Hydrogen Epoch of Reionization Array (HERA\footnote{http://reionization.org}), the LOw Frequency ARray (LOFAR, \citealt{dev09}), the Canadian Hydrogen Intensity Mapping Experiment (CHIME, \citealt{ban14}), the Long Wavelength Array (LWA, \citealt{ell13}), and the low frequency component of the Square Kilometer Array Low Frequency Aperture Array (SKA-Low \citealt{mel13}).
 
@@
 
-Direct imaging correlators have begun to be explored on deployed arrays including the Basic Element for SKA Training II (BLAST-2) array \citep{fos14}, the Omniscope \citep{zhe14}, and an earlier incarnation at higher frequencies with the intent of pulsar timing \citep{oto94, dai00}.
+Direct imaging correlators have begun to be explored on deployed arrays including the Basic Element for SKA Training II (BEST-2) array \citep{fos14}, the Omniscope \citep{zhe14}, and an earlier incarnation at higher frequencies with the intent of pulsar timing \citep{oto94, dai00}.
 However, each of these examples make assumptions about the redundancy of the array layout, and require the collecting elements are identical.
 On the other hand, the MOFF algorithm achieves the same $\Npix \log \Npix$ computational scaling without placing any restriction on antenna placement, can accommodate non-identical beam patterns, and is a provably optimal mapping \citep{mor11}.
-This algorithm uses the antenna beam patterns to grid the electric field measurements to a regular grid in the software holography/A-transpose fashion \citep{mor09, bha08, teg97} before performing the spatial FFT.
+This algorithm uses the antenna beam patterns to grid the electric field measurements to a regular grid in the software holography/A-transpose fashion \citep{mor09,bha08,teg97b} before performing the spatial FFT.
 This process has been shown to theoretically produce a data product identical to images produced from the traditional FX correlator.
@@
 We identify specific array design classes where the EPIC correlator is computationally more efficient than the FX algorithm.
-We conclude and discuss future research prospects in \S\ref{sec:conclusions}
+We conclude and discuss future research prospects in \S\ref{sec:conclusions}.
 
-Motivate MOFF from a technical and scientific standpoint. Refer to \citet{del07,mor09,mor11}.
+% Motivate MOFF from a technical and scientific standpoint. Refer to \citet{del07,mor09,mor11}.
 
@@
 
-Refresh the math equivalence between MOFF and FX.
+We provide a quick summary of the mathematical equivalence of the MOFF and FX correlators detailed in \citet{mor11}.
+
+Electric fields from astrophysical sources, $E(\hat{\mathbf{s}})$, in the sky coordinate system denoted by unit vector $\hat{\mathbf{s}}$ propagate towards the observer as:
+\begin{align}
+  \widetilde{E}(\mathbf{r}) &= \int E(\hat{\mathbf{s}})\,e^{-i2\pi\mathbf{r}\cdot\hat{\mathbf{s}}}\,\dif^2\hat{\mathbf{s}},
+\end{align}
+where, $\mathbf{r}$ denotes the observer's coordinate system and $\widetilde{E}(\mathbf{r})$ is the propagated electric field.
+Thus the propagated electric field is a linear superposition of the electric fields emanating from astronomical sources with appropriate complex phases.
+It can also be described as a Fourier transform of the electric fields in the sky coordinates.
+
+An antenna, $a$, measures a phased sum of these propagated electric fields over its effective collecting area with an additive receiver noise:
+\begin{align}\label{eqn:measured-E-field}
+  \widetilde{E}_a &= \int \widetilde{W}_a(\mathbf{r}-\mathbf{r}_a)\,\widetilde{E}(\mathbf{r})\,\dif^2\mathbf{r} + \widetilde{n}_a \\
+                  &= \int \widetilde{W}_a(\mathbf{r}-\mathbf{r}_a) \left[ \int E(\hat{\mathbf{s}})\,e^{-i2\pi\mathbf{r}\cdot\hat{\mathbf{s}}}\,\dif^2\hat{\mathbf{s}} \right] \dif^2\mathbf{r} + \widetilde{n}_a \\
+                  &= \int \delta(\mathbf{r}-\mathbf{r}_a) \left[ \int {W}_a(\hat{\mathbf{s}})\,E(\hat{\mathbf{s}})\,e^{-i2\pi\mathbf{r}\cdot\hat{\mathbf{s}}}\,\dif^2\hat{\mathbf{s}} \right] \dif^2\mathbf{r} + \widetilde{n}_a
+\end{align}
+where, $\widetilde{W}_a(\mathbf{r})$ is the aperture electric field illumination pattern of the antenna and its Fourier transform, $W_a(\hat{\mathbf{s}})$ is the directional antenna voltage response.
+
+Interferometers measure {\it visibility} -- the degree of coherence -- between electric fields measured by a pair of antennas \citep{van34,zer38,tho01}.
+Visibility, $\widetilde{V}_p$, can be written as:
+
+\begin{align}
+  \widetilde{V}_p &= \left\langle \widetilde{E}_a\widetilde{E}_b^\star \right\rangle \label{eqn:cc-vis}\\
+                  &= \int\delta(\mathbf{u}-\mathbf{u}_p)\left[\int e^{-i2\pi\mathbf{u}.\hat{\mathbf{s}}}\,B(\hat{\mathbf{s}})\,I(\hat{\mathbf{s}})\,\dif^2\hat{\mathbf{s}}\right] \dif^2\mathbf{u} + n_p,
+\end{align}
+where, $I(\hat{\mathbf{s}})$ is the sky brightness, $B(\hat{\mathbf{s}})$ is the antenna power response, $\mathbf{u}$ denotes the coordinates of the $uv$-measurement plane, $\mathbf{u}_p\equiv \mathbf{r}_a-\mathbf{r}_b$ is the antenna spacing vector corresponding to the antenna pair $p$, and $n_p$ is the receiver noise.
+This signifies that the visibility ($\widetilde{V}_p$) measured between a pair of antennas ($p$) is obtained by the multiplying the sky brightness $I(\hat{\mathbf{s}})$ by the antenna power response $B(\hat{\mathbf{s}})$ and Fourier transform from the directional coordinates ($\hat{\mathbf{s}}$) to $uv$ coordinates, which are then sampled at the locations of the antenna spacings (or baselines), namely, $\mathbf{u}_p$ using the sampling function $\delta(\mathbf{u}-\mathbf{u}_p)$ and added to the receiver noise $n_p$.
+This can be equivalently re-written as:
+\begin{align}\label{eqn:software-holography}
+  \widetilde{V}_p &= \int\delta(\mathbf{u}^\prime-\mathbf{u}_p)\left[\int \tilde{B}(\mathbf{u}^\prime-\mathbf{u}) \right. \nonumber\\
+  &\left. \qquad \times \left[\int e^{-i2\pi\mathbf{u}.\hat{\mathbf{s}}}\,I(\hat{\mathbf{s}})\,\dif^2\hat{\mathbf{s}}\right]\dif^2\mathbf{u}\vphantom{\int}\right] \dif^2\mathbf{u}^\prime + n_p,
+\end{align}
+where, $\tilde{B}(\mathbf{u})$ denotes the $uv$-space antenna power response obtained by a Fourier transform of $B(\hat{\mathbf{s}})$.
+Effectively, the multiplication in image space by $B(\hat{\mathbf{s}})$ has been replaced by a convolution with $\tilde{B}(\mathbf{u})$ in $uv$-space.
+This is the software holographic equivalent of traditional FX correlator output.
+Following the matrix notation of \citet{mor11}, the above equation can be expressed as:
+\begin{align}
+  \mathbf{m}(\mathbf{v}) &= \widetilde{B}(\mathbf{v},\mathbf{u})\,\mathbf{F}(\mathbf{u},\hat{\mathbf{s}})\,I(\hat{\mathbf{s}}) + \mathbf{n}(\mathbf{v}),
+\end{align}
+where, the sky brightness $I(\hat{\mathbf{s}})$ is Fourier transformed using $\mathbf{F}(\mathbf{u},\hat{\mathbf{s}})$ and the resultant spatial coherence function is weighted and summed using the antenna power response, $\widetilde{B}(\mathbf{v},\mathbf{u})$ in $uv$-space sampled at the baseline location to obtain the measured visibilities:
+\begin{align}
+  \mathbf{m}(\mathbf{v}) &= \left\langle \widetilde{E}(\mathbf{a})\ast\widetilde{E}(\mathbf{a}^\prime)\right\rangle_t. \label{eqn:matrix-cc-vis}
+\end{align}
+$\mathbf{m}(\mathbf{v})$ denotes visibilities measured by cross-correlating measured antenna electric fields over all possible pairs of $\mathbf{a}$ and $\mathbf{a}^\prime$.
+It is the same as equation~\ref{eqn:cc-vis} written in matrix notation.
+
+Using the optimal map-making formalism \citep{teg97a,teg97b}, a software holography image is formed using \citep{mor09}:
+\begin{align}
+  I^\prime(\hat{\mathbf{s}}) &= \mathbf{F}^\textrm{T}(\hat{\mathbf{s}},\mathbf{u})\,\widetilde{B}^{\,\textrm{T}}(\mathbf{u},\mathbf{v})\,\mathbf{N}^{-1}\,\mathbf{m}(\mathbf{v}) \label{eqn:dirty-image-FX}
+\end{align}
+where the measured visibilities are weighted by the inverse of the system noise, following a gridding process using the holographic antenna power response as the gridding kernel followed by a Fourier transform to create an image $I^\prime(\hat{\mathbf{s}})$ which is the optimal estimate of the true image $I(\hat{\mathbf{s}})$.
+
+It may be noted that the antenna power response in $uv$-plane can be written as:
+\begin{align}\label{eqn:antenna-power-breakup}
+  \tilde{B}_p(\mathbf{u}) &= \widetilde{W}_a(\mathbf{u}) \ast \widetilde{W}_b^\star(\mathbf{u})
+\end{align}
+where, $W_a(\mathbf{u})$ and $W_b(\mathbf{u})$ denote the individual antenna voltage responses.
+Thus $\tilde{B}_p(\mathbf{u})$ is a cross-correlation of the individual antenna voltage responses.
+Similarly, it can be shown that
+\begin{align}
+  \left\langle E(\hat{\mathbf{s}})\,E^\star(\hat{\mathbf{s}}^\prime)\right\rangle &= \left\langle E(\hat{\mathbf{s}})\,E^\star(\hat{\mathbf{s}}^\prime) \delta(\hat{\mathbf{s}}-\hat{\mathbf{s}}^\prime) \right\rangle \nonumber\\
+  &= \left\langle \left|E(\hat{\mathbf{s}})\right|^2 \right\rangle = I(\hat{\mathbf{s}}) \label{eqn:sky-intensity}
+\end{align}
+
+From equations~\ref{eqn:measured-E-field}, \ref{eqn:software-holography}, \ref{eqn:matrix-cc-vis}, \ref{eqn:antenna-power-breakup} and \ref{eqn:sky-intensity}, the optimal image in equation \ref{eqn:dirty-image-FX} can be expressed equivalently as:
+\begin{align}
+  I^\prime(\hat{\mathbf{s}}) &= \left\langle \left|\,\mathbf{F}^\textrm{T}(\hat{\mathbf{s}},\mathbf{r})\,\widetilde{\mathbf{W}}^\textrm{T}(\mathbf{r},\mathbf{a})\,\widetilde{\mathbf{N}}(\mathbf{a},\mathbf{a})\,\widetilde{E}(\mathbf{a})\,\right|^2\right\rangle_t.
+\label{eqn:dirty-image-MOFF}
+\end{align}
+The term inside the angular brackets before squaring has a very similar form as that in equation~\ref{eqn:dirty-image-FX}.
+It signifies that the measured antenna electric fields are weighted by the antenna noise, weighted and gridded by the antenna aperture kernel, Fourier transformed and finally squared to obtain the same image estimated that would have been obtained using equation~\ref{eqn:dirty-image-FX}.
+
+Equations~\ref{eqn:dirty-image-MOFF} and \ref{eqn:dirty-image-FX} are identical to each other obtained as a consequence of the multiplication-convolution theorem of Fourier transforms.
+In other words, equation~\ref{eqn:dirty-image-FX} is obtained by a Fourier transform of the gridded spatial correlation of electric fields measured by antennas while equation~\ref{eqn:dirty-image-MOFF} is obtained by squaring the Fourier transform of the gridded electric fields measured by the antennas.
+The time-scale of averaging in equations~\ref{eqn:cc-vis} and \ref{eqn:dirty-image-MOFF} is set by the coherence time-scale of the electric fields or the science requirement (e.g., timescale of variability or transients).
+
+There are some important differences between the two techniques:
+\begin{enumerate}
+  \item The time-averaging cannot be performed on a stochastic measurement but only on its statistical properties.
+In FX imaging, the visibilities measured between antenna pairs represent spatial correlations which can be time-averaged following which the they gridded and imaged.
+However, in MOFF imaging both antenna and gridded electric fields are stochastic and therefore have to be imaged before time-averaging.
+  \item In FX imaging, electric fields measured on antennas are not correlated with themselves and hence lack zero spacing measurements.
+In contrast, in MOFF imaging, since the gridded electric fields are imaged and squared, they contain information from auto-correlated electric fields at zero spacing.
+Hence, they have to be subtracted from the images.
+\end{enumerate}
 
@@
 
-Discuss implementation and make code available for public.
+We have implemented the MOFF imaging technique in our ``E-field Parallel Imaging Correlator'' -- a highly parallelized Object Oriented Python package,\footnote{EPIC package can be accessed at https://github.com/nithyanandan/EPIC} now publicly available.
+Besides implementing the MOFF imaging algorithm it also includes FX imaging using software holography technique and a simulator for generating electric fields from a sky model.
 
-Calibration
+Figure~\ref{fig:MOFF-flowchart} shows the flowchart for MOFF imaging.
+The propagated electric fields are shown on the left at different time stamps, $t_1\ldots t_\textrm{M}$.
+At each time stamp, the electric fields measured by antennas are denoted by $E_1(t)\ldots E_\textrm{N}(t)$.
+The F-engine performs a temporal Fourier transform on the electric field time-series to obtain electric field spectra $E_1(f)\ldots E_\textrm{N}(f)$ ($\widetilde{E}_a$ in matrix notation) for each of the antennas.
+Each of the complex antenna gains are calibrated to correct the corresponding electric field spectra.
+These calibrated electric fields are gridded using an antenna-based gridding convolution function following which it is spatially Fourier transformed and squared to obtain images for every time stamp.
+These images are then time-averaged to obtain the accumulated image $I(f)$ ($I(\hat{\mathbf{s}})$ in matrix notation).
+\begin{figure}
+  \includegraphics[width=\columnwidth]{MOFF_flowchart.eps}
+  \caption{A flowchart of MOFF imaging in EPIC.
+The propagated electric fields shown on the left are measured as time-series $E_1(t)\ldots E_\textrm{N}(t)$ by the antennas which are then Fourier transformed by the F-engine to produce electric field spectra $E_1(f)\ldots E_\textrm{N}(f)$.
+They are calibrated and gridded.
+The gridded electric fields $E_\textrm{g}(f)$ from each time series are imaged to produce an images $I_1(f)\ldots I_\textrm{N}(f)$.
+These images are time-averaged to obtain the final image $I(f)$.}
+  \label{fig:MOFF-flowchart}
+\end{figure}
+
+Figure~\ref{fig:FX-flowchart} shows the flowchart for software holographic imaging from a FX correlator.
+The antenna-based F-engine is identical to that in the MOFF processing.
+The electric field spectra each antenna are then cross-multiplied in the X-engine with those from all other antennas to obtain the visibilities $V_\textrm{ij}(f)$ ($\mathbf{m}(\mathbf{v})$ in matrix notation).
+They are calibrated and time-averaged to obtain $\langle V_\textrm{ij}(f)\rangle$ which are then gridded and imaged to obtain the image $I(f)$.
+The $I(f)$ obtained from both techniques are identical as explained in \S\ref{sec:math}.
+\begin{figure}
+  \includegraphics[width=\columnwidth]{FX_flowchart.eps}
+  \caption{A flowchart of FX imaging in EPIC.
+The FX process flow shares the F-engine with the MOFF process.
+Following the F-engine, the electric fields pass through the X-engine to obtain visibilities $V_\textrm{ij}(f)$ which are calibrated and time-averaged.
+Then they are gridded to obtain the gridded visibilities $V_\textrm{g}(f)$ which are then imaged to obtain the image $I(f)$.}
+  \label{fig:FX-flowchart}
+\end{figure}
+
+Here, we discuss the components of these architectures in detail.
+
+\subsection{Temporal Fourier transform}\label{sec:F-engine}
+
+This module is common to the MOFF and FX imaging techniques.
+Time samples of electric fields measured by the antenna and digitized by the A/D converter is Fourier transformed to generate electric field spectra.
+This step can be parallelized by antennas as shown in Figure~\ref{fig:f-engine}.
+The output is then fed to either MOFF and FX imaging pipelines.
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{F-engine.eps}
+  \caption{Block diagram of a F-engine. The electric field data streams from antennas are Fourier transformed in parallel to generate electric field spectra.}
+  \label{fig:f-engine}
+\end{figure}
+
+\subsection{Antenna-to-Grid Mapping}
+
+A grid is generated on the coordinate system in which antenna locations are specified with a grid spacing that is at most $\lambda_\textrm{min}/2$ even at the highest frequency to ensure there is no aliasing even from regions of the sky far away from the field of view.
+The number of locations on the grid is restricted to be a power of 2.
+
+The gridding kernel in the simplest case is given by the antenna aperture illumination function, $\widetilde{B}(\mathbf{r}-\mathbf{r}_a)$, which can specified either by a functional form or as a table of values against locations around the antennas.
+A nearest neighbor mapping from all antenna footprints to grid locations is created using an efficient k-d tree algorithm \citep{man99}.
+There is no restriction here that the aperture illumination function has to be identical across antennas.
+
+In the most general case, this gridding kernel could contain information on the $w$-projection effect, and even other time-dependent ionospheric effects.
+For a stationary antenna array in the absence of any time-dependent effects, this mapping has to be determined only once in the antenna array coordinate frame.
+The antenna-to-grid mapping matrix, $\mathbf{M}(\mathbf{r},\mathbf{e})$ is described as a transformation matrix from the space of measured electric fields ($\mathbf{e}$) to the antenna array grid denoted by $\mathbf{r}$.
+Since each antenna occupies a footprint typically the size of its aperture, $\mathbf{M}(\mathbf{r},\mathbf{e})$ which is generally of size $N_\textrm{grid}\times N_\textrm{a}$, reduces to a sparse block-diagonal matrix with only $N_\textrm{a}$ blocks and roughly $N_\textrm{k}$ non-zero entries per block.
+Figure~\ref{fig:a2g-mapping} illustrates the antenna-to-grid mapping matrix and the grid containing the mapped aperture footprints of the antennas.
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{a2g_mapping.eps}
+  \caption{Block diagram of an antenna-to-grid mapping.
+A sparse block-diagonal matrix of total size $N_\textrm{grid}\times N_\textrm{a}$ is created where each block contains roughly the number of pixels covered by the respective kernel.
+The antenna aperture illumination kernels do not have to be identical to each other.
+A discrete set of arbitrarily placed antennas are now placed onto a regular grid.}
+  \label{fig:a2g-mapping}
+\end{figure}
+
+\subsection{Calibration}
+
+\subsection{Gridding Convolution}
+
+The antenna array aperture illumination over the entire grid, $\widetilde{W}(\mathbf{r})$, is obtained by a projection of the individual antenna aperture illuminations:
+\begin{align}
+  \widetilde{W}(\mathbf{r}) &= \sum_a \widetilde{W}_a(\mathbf{r}-\mathbf{r}_a) \\
+                            &= \mathbf{M}(\mathbf{r},\mathbf{e})\,\mathcal{I}(\mathbf{e}),
+\end{align}
+where, $\mathcal{I}(\mathbf{e})$ is a row of ones.
+This is achieved using efficient multiplication with the sparse matrix created in the antenna-to-grid mapping process.
+Unless $\widetilde{W}(\mathbf{r})$ includes time-dependent effects of the ionosphere or the instrument, it needs to be computed just once for the entire observation.
+However, the gridding of electric fields must be computed at every readout of the electric field spectra.
+Thus,
+\begin{align}
+  \widetilde{E}(\mathbf{r}) &= \mathbf{M}(\mathbf{r},\mathbf{e})\,E(\mathbf{e}),
+\end{align}
+where, $E(\mathbf{e})$ denotes the spectra of measured antenna electric fields.
+
+In practice the grid is set to be a power of 2 such that the grid spacing is at most $\lambda/2$ even at the highest frequency in the observing band.
+
+\subsection{Spatial Fourier Transform}
+
+Before the spatial Fourier transform, the gridded electric fields are padded with zeros in order to match the grid size and angular size of each image pixel that would have been obtained with software holography of output from an FX correlator.
+
+In MOFF imaging, these are spatially Fourier transformed followed by a squaring operation at every timestamp for every frequency channel.
+In FX imaging, the spatial Fourier transform is performed only once per integration timescale and does not include a squaring operation.
+
+\subsection{Time-averaging}
+
+In MOFF imaging, the measured antenna electric fields and the corresponding holographic electric field images are zero-mean stochastic quantities.
+Hence, they cannot be time-averaged to reduce noise.
+The statistical quantity stable with time in this case are the square of the holographic electric field images.
+Thus, squared images have to be formed at every instant of time before averaging as indicated in equation~\ref{eqn:dirty-image-MOFF}.
+
+In contrast, visibilities measured by an antenna are statistically stable within an integration time interval.
+Hence, they are averaged after calibration as shown in equation~\ref{eqn:cc-vis}.
+It is advantageous to average them in visibilities before imaging because visibilities represent a compact representation of the information in images.
+Hence it is more computationally efficient to average antenna pair visibilities rather than images.
+Since this averaging has been performed already on the visibilities over an integration timescale, the imaging step has to be performed only once per integration cycle.
+FX imaging holds this advantage as long as the square of the number of antennas is smaller than the number of pixels on the image.
 
 \section{Verification}\label{sec:verify}
+
+We use the EPIC simulator to generate electric field samples from a sky model.
+In our example, we use 16 frequency channels each of width $\delta f=$100~kHz, 10 point sources of random flux densities at random locations.
+The number of timestamps in one integration cycle was kept at four where each A/D timeseries is $1/\delta f=10\,\mu$s long.
 
@@
 \section{Analysis and Feasibility}\label{sec:analysis}
+
 \subsection{Scaling Relations: MOFF vs. FX}
@@
 \subsection{Case Study}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{MOFF_FX_crossover_baseline_n-antennas_rho_fov_gridding.eps}
+  \caption{Current and instruments planned for future in parameter space of baseline length and number of antennas with MOFF and FX.}
+  \label{fig:parameter-space-bll-nant-instruments}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{MOFF_FX_computations_fov_gridding_annotated.eps}
+  \caption{Current and instruments planned for future in parameter space of number of complex multiplies and adds with MOFF and FX.}
+  \label{fig:parameter-space-computations-instruments}
+\end{figure}
+
 % \subsection{Figures and tables}

### 2015-10-22 imaging_demo/ms.tex
broke up paragraphs to about 80 character lines

@@
 
-Radio astronomy is entering an era in which interferometers of hundreds to thousands of individual antennas are needed to achieve desired survey speeds.
-Nowhere is this more apparent than at radio frequencies below 1.4 GHz.
-The study of the history of hydrogen gas throughout the universe's evolution is pushing technology development towards arrays of low-cost antennas with large fields of view and densely packed apertures.
-Similarly, the search for transient objects and regular monitoring of the time-dependent sky is driving instruments in the same direction.
-A number of new telescopes are under development around the world based on this new paradigm, including the Murchison Widefield Array (MWA, \citealt{tin13}), the Precision Array for Probing the Epoch of Reionization (PAPER, \citealt{par10}), the Hydrogen Epoch of Reionization Array (HERA\footnote{http://reionization.org}), the LOw Frequency ARray (LOFAR, \citealt{dev09}), the Canadian Hydrogen Intensity Mapping Experiment (CHIME, \citealt{ban14}), the Long Wavelength Array (LWA, \citealt{ell13}), and the low frequency component of the Square Kilometer Array Low Frequency Aperture Array (SKA-Low \citealt{mel13}).
+Radio astronomy is entering an era in which interferometers of hundreds to
+thousands of individual antennas are needed to achieve desired survey speeds.
+Nowhere is this more apparent than at radio frequencies below 1.4 GHz. The study
+of the history of hydrogen gas throughout the universe's evolution is pushing
+technology development towards arrays of low-cost antennas with large fields of
+view and densely packed apertures. Similarly, the search for transient objects
+and regular monitoring of the time-dependent sky is driving instruments in the
+same direction. A number of new telescopes are under development around the world
+based on this new paradigm, including the Murchison Widefield Array (MWA,
+\citealt{tin13}), the Precision Array for Probing the Epoch of Reionization
+(PAPER, \citealt{par10}), the Hydrogen Epoch of Reionization Array (HERA
+\footnote{http://reionization.org}), the LOw Frequency ARray (LOFAR,
+\citealt{dev09}), the Canadian Hydrogen Intensity Mapping Experiment (CHIME,
+\citealt{ban14}), the Long Wavelength Array (LWA, \citealt{ell13}), and the low
+frequency component of the Square Kilometer Array Low Frequency Aperture Array
+(SKA-Low \citealt{mel13}).
 
-This paradigm shift requires a fundamentally new approach to the design of digital correlators \citep{lon00}.
-Modern correlators calculate the cross-power correlation between all antenna pairs in many narrow frequencies, forming \emph{visibilities}, the traditional fundamental measurement of radio interferometers.
-The computational requirements for a modern FX correlator scale with the number of antenna pairs, or the square of the number of antennas $\sim \Nant^2$ \citep{bun04}.
-For this reason traditional correlators have difficulty scaling to thousands of antennas.
-As an example, the full HERA correlator for 352 dishes with 200 MHz of bandwidth requires 212 trillion complex multiplies and adds per second (TMACS).
-Future arrays with thousands of collecting elements will require orders of magnitude more computation, making the correlator the dominant cost.
+This paradigm shift requires a fundamentally new approach to the design of
+digital correlators \citep{lon00}. Modern correlators calculate the cross-power
+correlation between all antenna pairs in many narrow frequencies, forming
+\emph{visibilities}, the traditional fundamental measurement of radio
+interferometers. The computational requirements for a modern FX correlator scale
+with the number of antenna pairs, or the square of the number of antennas $\sim
+\Nant^2$ \citep{bun04}. For this reason traditional correlators have difficulty
+scaling to thousands of antennas. As an example, the full HERA correlator for
+352 dishes with 200 MHz of bandwidth requires 212 trillion complex multiplies
+and adds per second (TMACS). Future arrays with thousands of collecting elements
+will require orders of magnitude more computation, making the correlator the
+dominant cost.
 
-For certain classes of radio arrays there is an alternative to the FX correlator that can lower the computational burden by directly performing a spatial fast Fourier transform (FFT) on the electric fields measured by each antenna in the array at each time step, removing the cross-correlation step.
-This relieves the computational scaling from the harsh $\Nant^2$ to the more gentle envelope of $\Npix\log\Npix$, where $\Npix$ is the number of pixels in the Fourier transform \citep[e.g.][]{mor11,teg09,teg10}.
-This architecture is often referred to as a ``direct imaging" correlator because it eliminates the intermediary cross-correlation data products of the FX and earlier lag correlators, but instead directly forms images from the electric field measurements.
+For certain classes of radio arrays there is an alternative to the FX correlator
+that can lower the computational burden by directly performing a spatial fast
+Fourier transform (FFT) on the electric fields measured by each antenna in the
+array at each time step, removing the cross-correlation step. This relieves the
+computational scaling from the harsh $\Nant^2$ to the more gentle envelope of
+$\Npix\log\Npix$, where $\Npix$ is the number of pixels in the Fourier transform
+\citep[e.g.][]{mor11,teg09,teg10}. This architecture is often referred to as a
+``direct imaging" correlator because it eliminates the intermediary cross-
+correlation data products of the FX and earlier lag correlators, but instead
+directly forms images from the electric field measurements.
 
-Direct imaging correlators have begun to be explored on deployed arrays including the Basic Element for SKA Training II (BEST-2) array \citep{fos14}, the Omniscope \citep{zhe14}, and an earlier incarnation at higher frequencies with the intent of pulsar timing \citep{oto94, dai00}.
-However, each of these examples make assumptions about the redundancy of the array layout, and require the collecting elements are identical.
-On the other hand, the MOFF algorithm achieves the same $\Npix \log \Npix$ computational scaling without placing any restriction on antenna placement, can accommodate non-identical beam patterns, and is a provably optimal mapping \citep{mor11}.
-This algorithm uses the antenna beam patterns to grid the electric field measurements to a regular grid in the software holography/A-transpose fashion \citep{mor09,bha08,teg97b} before performing the spatial FFT.
-This process has been shown to theoretically produce a data product identical to images produced from the traditional FX correlator.
+Direct imaging correlators have begun to be explored on deployed arrays including
+the Basic Element for SKA Training II (BEST-2) array \citep{fos14}, the Omniscope
+\citep{zhe14}, and an earlier incarnation at higher frequencies with the intent
+of pulsar timing \citep{oto94, dai00}. However, each of these examples make
+assumptions about the redundancy of the array layout, and require the collecting
+elements are identical. On the other hand, the MOFF algorithm achieves the same
+$\Npix \log \Npix$ computational scaling without placing any restriction on
+antenna placement, can accommodate non-identical beam patterns, and is a provably
+optimal mapping \citep{mor11}. This algorithm uses the antenna beam patterns to
+grid the electric field measurements to a regular grid in the software
+holography/A-transpose fashion \citep{mor09,bha08,teg97b} before performing the
+spatial FFT. This process has been shown to theoretically produce a data product
+identical to images produced from the traditional FX correlator.
 
-Here we present the first software implementation of the MOFF correlator, and announce the public release of the E-field Parallel Imaging Correlator (EPIC) code.
-We begin with a technical description of the algorithm in \S \ref{sec:math}, then discuss our particular implementation in \S \ref{sec:software}.
-We then verify the output data quality from our code in \S \ref{sec:verify} by presenting simulated images from both the EPIC correlator and comparing to a simulated FX correlator.
-We also demonstrate the performance with real-world data from the LWA.
-In \S \ref{sec:analysis} we analyze the scaling relationships of the algorithm.
-We identify specific array design classes where the EPIC correlator is computationally more efficient than the FX algorithm.
-We conclude and discuss future research prospects in \S\ref{sec:conclusions}.
+Here we present the first software implementation of the MOFF correlator, and
+announce the public release of the E-field Parallel Imaging Correlator (EPIC)
+code. We begin with a technical description of the algorithm in \S
+\ref{sec:math}, then discuss our particular implementation in \S
+\ref{sec:software}. We then verify the output data quality from our code in \S
+\ref{sec:verify} by presenting simulated images from both the EPIC correlator
+and comparing to a simulated FX correlator. We also demonstrate the performance
+with real-world data from the LWA. In \S \ref{sec:analysis} we analyze the
+scaling relationships of the algorithm. We identify specific array design classes
+where the EPIC correlator is computationally more efficient than the FX
+algorithm. We conclude and discuss future research prospects in \S
+\ref{sec:conclusions}.
 
@@
 
-We provide a quick summary of the mathematical equivalence of the MOFF and FX correlators detailed in \citet{mor11}.
+We provide a quick summary of the mathematical equivalence of the MOFF and FX
+correlators detailed in \citet{mor11}.
 
-Electric fields from astrophysical sources, $E(\hat{\mathbf{s}})$, in the sky coordinate system denoted by unit vector $\hat{\mathbf{s}}$ propagate towards the observer as:
+Electric fields from astrophysical sources, $E(\hat{\mathbf{s}})$, in the sky
+coordinate system denoted by unit vector $\hat{\mathbf{s}}$ propagate towards
+the observer as:
 \begin{align}
@@
 \end{align}
-where, $\mathbf{r}$ denotes the observer's coordinate system and $\widetilde{E}(\mathbf{r})$ is the propagated electric field.
-Thus the propagated electric field is a linear superposition of the electric fields emanating from astronomical sources with appropriate complex phases.
-It can also be described as a Fourier transform of the electric fields in the sky coordinates.
+where, $\mathbf{r}$ denotes the observer's coordinate system and $\widetilde{E}
+(\mathbf{r})$ is the propagated electric field. Thus the propagated electric
+field is a linear superposition of the electric fields emanating from
+astronomical sources with appropriate complex phases. It can also be described
+as a Fourier transform of the electric fields in the sky coordinates.
 
-An antenna, $a$, measures a phased sum of these propagated electric fields over its effective collecting area with an additive receiver noise:
+An antenna, $a$, measures a phased sum of these propagated electric fields over
+its effective collecting area with an additive receiver noise:
 \begin{align}\label{eqn:measured-E-field}
@@
 \end{align}
-where, $\widetilde{W}_a(\mathbf{r})$ is the aperture electric field illumination pattern of the antenna and its Fourier transform, $W_a(\hat{\mathbf{s}})$ is the directional antenna voltage response.
+where, $\widetilde{W}_a(\mathbf{r})$ is the aperture electric field illumination
+pattern of the antenna and its Fourier transform, $W_a(\hat{\mathbf{s}})$ is the
+directional antenna voltage response.
 
-Interferometers measure {\it visibility} -- the degree of coherence -- between electric fields measured by a pair of antennas \citep{van34,zer38,tho01}.
+Interferometers measure {\it visibility} -- the degree of coherence -- between
+electric fields measured by a pair of antennas \citep{van34,zer38,tho01}.
 Visibility, $\widetilde{V}_p$, can be written as:
@@
 \end{align}
-where, $I(\hat{\mathbf{s}})$ is the sky brightness, $B(\hat{\mathbf{s}})$ is the antenna power response, $\mathbf{u}$ denotes the coordinates of the $uv$-measurement plane, $\mathbf{u}_p\equiv \mathbf{r}_a-\mathbf{r}_b$ is the antenna spacing vector corresponding to the antenna pair $p$, and $n_p$ is the receiver noise.
-This signifies that the visibility ($\widetilde{V}_p$) measured between a pair of antennas ($p$) is obtained by the multiplying the sky brightness $I(\hat{\mathbf{s}})$ by the antenna power response $B(\hat{\mathbf{s}})$ and Fourier transform from the directional coordinates ($\hat{\mathbf{s}}$) to $uv$ coordinates, which are then sampled at the locations of the antenna spacings (or baselines), namely, $\mathbf{u}_p$ using the sampling function $\delta(\mathbf{u}-\mathbf{u}_p)$ and added to the receiver noise $n_p$.
-This can be equivalently re-written as:
+where, $I(\hat{\mathbf{s}})$ is the sky brightness, $B(\hat{\mathbf{s}})$ is the
+antenna power response, $\mathbf{u}$ denotes the coordinates of the $uv$-
+measurement plane, $\mathbf{u}_p\equiv \mathbf{r}_a-\mathbf{r}_b$ is the antenna
+spacing vector corresponding to the antenna pair $p$, and $n_p$ is the receiver
+noise. This signifies that the visibility ($\widetilde{V}_p$) measured between a
+pair of antennas ($p$) is obtained by the multiplying the sky brightness
+$I(\hat{\mathbf{s}})$ by the antenna power response $B(\hat{\mathbf{s}})$ and
+Fourier transform from the directional coordinates ($\hat{\mathbf{s}}$) to $uv$
+coordinates, which are then sampled at the locations of the antenna spacings (or
+baselines), namely, $\mathbf{u}_p$ using the sampling function
+$\delta(\mathbf{u}-\mathbf{u}_p)$ and added to the receiver noise $n_p$. This can
+be equivalently re-written as:
 \begin{align}\label{eqn:software-holography}
@@
 \end{align}
-where, $\tilde{B}(\mathbf{u})$ denotes the $uv$-space antenna power response obtained by a Fourier transform of $B(\hat{\mathbf{s}})$.
-Effectively, the multiplication in image space by $B(\hat{\mathbf{s}})$ has been replaced by a convolution with $\tilde{B}(\mathbf{u})$ in $uv$-space.
-This is the software holographic equivalent of traditional FX correlator output.
-Following the matrix notation of \citet{mor11}, the above equation can be expressed as:
+where, $\tilde{B}(\mathbf{u})$ denotes the $uv$-space antenna power response
+obtained by a Fourier transform of $B(\hat{\mathbf{s}})$. Effectively, the
+multiplication in image space by $B(\hat{\mathbf{s}})$ has been replaced by a
+convolution with $\tilde{B}(\mathbf{u})$ in $uv$-space. This is the software
+holographic equivalent of traditional FX correlator output. Following the matrix
+notation of \citet{mor11}, the above equation can be expressed as:
 \begin{align}
@@
 \end{align}
-where, the sky brightness $I(\hat{\mathbf{s}})$ is Fourier transformed using $\mathbf{F}(\mathbf{u},\hat{\mathbf{s}})$ and the resultant spatial coherence function is weighted and summed using the antenna power response, $\widetilde{B}(\mathbf{v},\mathbf{u})$ in $uv$-space sampled at the baseline location to obtain the measured visibilities:
+where, the sky brightness $I(\hat{\mathbf{s}})$ is Fourier transformed using
+$\mathbf{F}(\mathbf{u},\hat{\mathbf{s}})$ and the resultant spatial coherence
+function is weighted and summed using the antenna power response,
+$\widetilde{B}(\mathbf{v},\mathbf{u})$ in $uv$-space sampled at the baseline
+location to obtain the measured visibilities:
 \begin{align}
@@
 \end{align}
-$\mathbf{m}(\mathbf{v})$ denotes visibilities measured by cross-correlating measured antenna electric fields over all possible pairs of $\mathbf{a}$ and $\mathbf{a}^\prime$.
-It is the same as equation~\ref{eqn:cc-vis} written in matrix notation.
+$\mathbf{m}(\mathbf{v})$ denotes visibilities measured by cross-correlating
+measured antenna electric fields over all possible pairs of $\mathbf{a}$ and
+$\mathbf{a}^\prime$. It is the same as equation~\ref{eqn:cc-vis} written in
+matrix notation.
 
-Using the optimal map-making formalism \citep{teg97a,teg97b}, a software holography image is formed using \citep{mor09}:
+Using the optimal map-making formalism \citep{teg97a,teg97b}, a software
+holography image is formed using \citep{mor09}:
 \begin{align}
@@
 \end{align}
-where the measured visibilities are weighted by the inverse of the system noise, following a gridding process using the holographic antenna power response as the gridding kernel followed by a Fourier transform to create an image $I^\prime(\hat{\mathbf{s}})$ which is the optimal estimate of the true image $I(\hat{\mathbf{s}})$.
+where the measured visibilities are weighted by the inverse of the system noise,
+following a gridding process using the holographic antenna power response as the
+gridding kernel followed by a Fourier transform to create an image
+$I^\prime(\hat{\mathbf{s}})$ which is the optimal estimate of the true image
+$I(\hat{\mathbf{s}})$.
 
@@
 \end{align}
-where, $W_a(\mathbf{u})$ and $W_b(\mathbf{u})$ denote the individual antenna voltage responses.
-Thus $\tilde{B}_p(\mathbf{u})$ is a cross-correlation of the individual antenna voltage responses.
-Similarly, it can be shown that
+where, $W_a(\mathbf{u})$ and $W_b(\mathbf{u})$ denote the individual antenna
+voltage responses. Thus $\tilde{B}_p(\mathbf{u})$ is a cross-correlation of the
+individual antenna voltage responses. Similarly, it can be shown that 
 \begin{align}
@@
 
-From equations~\ref{eqn:measured-E-field}, \ref{eqn:software-holography}, \ref{eqn:matrix-cc-vis}, \ref{eqn:antenna-power-breakup} and \ref{eqn:sky-intensity}, the optimal image in equation \ref{eqn:dirty-image-FX} can be expressed equivalently as:
+From equations~\ref{eqn:measured-E-field}, \ref{eqn:software-holography},
+\ref{eqn:matrix-cc-vis}, \ref{eqn:antenna-power-breakup} and
+\ref{eqn:sky-intensity}, the optimal image in equation \ref{eqn:dirty-image-FX}
+can be expressed equivalently as:
 \begin{align}
@@
 \end{align}
-The term inside the angular brackets before squaring has a very similar form as that in equation~\ref{eqn:dirty-image-FX}.
-It signifies that the measured antenna electric fields are weighted by the antenna noise, weighted and gridded by the antenna aperture kernel, Fourier transformed and finally squared to obtain the same image estimated that would have been obtained using equation~\ref{eqn:dirty-image-FX}.
+The term inside the angular brackets before squaring has a very similar form as
+that in equation~\ref{eqn:dirty-image-FX}. It signifies that the measured antenna
+electric fields are weighted by the antenna noise, weighted and gridded by the
+antenna aperture kernel, Fourier transformed and finally squared to obtain the
+same image estimated that would have been obtained using equation~
+\ref{eqn:dirty-image-FX}.
 
-Equations~\ref{eqn:dirty-image-MOFF} and \ref{eqn:dirty-image-FX} are identical to each other obtained as a consequence of the multiplication-convolution theorem of Fourier transforms.
-In other words, equation~\ref{eqn:dirty-image-FX} is obtained by a Fourier transform of the gridded spatial correlation of electric fields measured by antennas while equation~\ref{eqn:dirty-image-MOFF} is obtained by squaring the Fourier transform of the gridded electric fields measured by the antennas.
-The time-scale of averaging in equations~\ref{eqn:cc-vis} and \ref{eqn:dirty-image-MOFF} is set by the coherence time-scale of the electric fields or the science requirement (e.g., timescale of variability or transients).
+Equations~\ref{eqn:dirty-image-MOFF} and \ref{eqn:dirty-image-FX} are identical
+to each other obtained as a consequence of the multiplication-convolution theorem
+of Fourier transforms. In other words, equation~\ref{eqn:dirty-image-FX} is
+obtained by a Fourier transform of the gridded spatial correlation of electric
+fields measured by antennas while equation~\ref{eqn:dirty-image-MOFF} is obtained
+by squaring the Fourier transform of the gridded electric fields measured by the
+antennas. The time-scale of averaging in equations~\ref{eqn:cc-vis} and
+\ref{eqn:dirty-image-MOFF} is set by the coherence time-scale of the electric
+fields or the science requirement (e.g., timescale of variability or transients).
 
@@
 \begin{enumerate}
-  \item The time-averaging cannot be performed on a stochastic measurement but only on its statistical properties.
-In FX imaging, the visibilities measured between antenna pairs represent spatial correlations which can be time-averaged following which the they gridded and imaged.
-However, in MOFF imaging both antenna and gridded electric fields are stochastic and therefore have to be imaged before time-averaging.
-  \item In FX imaging, electric fields measured on antennas are not correlated with themselves and hence lack zero spacing measurements.
-In contrast, in MOFF imaging, since the gridded electric fields are imaged and squared, they contain information from auto-correlated electric fields at zero spacing.
+\item The time-averaging cannot be performed on a stochastic measurement but
+  only on its statistical properties. In FX imaging, the visibilities measured
+  between antenna pairs represent spatial correlations which can be time-averaged
+  following which the they gridded and imaged. However, in MOFF imaging both
+  antenna and gridded electric fields are stochastic and therefore have to be
+  imaged before time-averaging. 
+\item In FX imaging, electric fields measured on antennas are not correlated with
+  themselves and hence lack zero spacing measurements. In contrast, in MOFF
+  imaging, since the gridded electric fields are imaged and squared, they
+  contain information from auto-correlated electric fields at zero spacing.
   Hence, they have to be subtracted from the images.
@@
 
-We have implemented the MOFF imaging technique in our ``E-field Parallel Imaging Correlator'' -- a highly parallelized Object Oriented Python package,\footnote{EPIC package can be accessed at https://github.com/nithyanandan/EPIC} now publicly available.
-Besides implementing the MOFF imaging algorithm it also includes FX imaging using software holography technique and a simulator for generating electric fields from a sky model.
+We have implemented the MOFF imaging technique in our ``E-field Parallel Imaging
+Correlator'' -- a highly parallelized Object Oriented Python package,
+\footnote{EPIC package can be accessed at https://github.com/nithyanandan/EPIC}
+now publicly available. Besides implementing the MOFF imaging algorithm it also
+includes FX imaging using software holography technique and a simulator for
+generating electric fields from a sky model.
 
-Figure~\ref{fig:MOFF-flowchart} shows the flowchart for MOFF imaging.
-The propagated electric fields are shown on the left at different time stamps, $t_1\ldots t_\textrm{M}$.
-At each time stamp, the electric fields measured by antennas are denoted by $E_1(t)\ldots E_\textrm{N}(t)$.
-The F-engine performs a temporal Fourier transform on the electric field time-series to obtain electric field spectra $E_1(f)\ldots E_\textrm{N}(f)$ ($\widetilde{E}_a$ in matrix notation) for each of the antennas.
-Each of the complex antenna gains are calibrated to correct the corresponding electric field spectra.
-These calibrated electric fields are gridded using an antenna-based gridding convolution function following which it is spatially Fourier transformed and squared to obtain images for every time stamp.
-These images are then time-averaged to obtain the accumulated image $I(f)$ ($I(\hat{\mathbf{s}})$ in matrix notation).
+Figure~\ref{fig:MOFF-flowchart} shows the flowchart for MOFF imaging. The
+propagated electric fields are shown on the left at different time stamps,
+$t_1\ldots t_\textrm{M}$. At each time stamp, the electric fields measured by
+antennas are denoted by $E_1(t)\ldots E_\textrm{N}(t)$. The F-engine performs a
+temporal Fourier transform on the electric field time-series to obtain electric
+field spectra $E_1(f)\ldots E_\textrm{N}(f)$ ($\widetilde{E}_a$ in matrix
+notation) for each of the antennas. Each of the complex antenna gains are
+calibrated to correct the corresponding electric field spectra. These calibrated
+electric fields are gridded using an antenna-based gridding convolution function
+following which it is spatially Fourier transformed and squared to obtain images
+for every time stamp. These images are then time-averaged to obtain the
+accumulated image $I(f)$ ($I(\hat{\mathbf{s}})$ in matrix notation).
 \begin{figure}
   \includegraphics[width=\columnwidth]{MOFF_flowchart.eps}
-  \caption{A flowchart of MOFF imaging in EPIC.
-The propagated electric fields shown on the left are measured as time-series $E_1(t)\ldots E_\textrm{N}(t)$ by the antennas which are then Fourier transformed by the F-engine to produce electric field spectra $E_1(f)\ldots E_\textrm{N}(f)$.
-They are calibrated and gridded.
-The gridded electric fields $E_\textrm{g}(f)$ from each time series are imaged to produce an images $I_1(f)\ldots I_\textrm{N}(f)$.
-These images are time-averaged to obtain the final image $I(f)$.}
+  \caption{A flowchart of MOFF imaging in EPIC. The propagated electric fields
+    shown on the left are measured as time-series $E_1(t)\ldots E_\textrm{N}(t)$
+    by the antennas which are then Fourier transformed by the F-engine to produce
+    electric field spectra $E_1(f)\ldots E_\textrm{N}(f)$. They are calibrated and
+    gridded. The gridded electric fields $E_\textrm{g}(f)$ from each time series
+    are imaged to produce an images $I_1(f)\ldots I_\textrm{N}(f)$. These images
+    are time-averaged to obtain the final image $I(f)$.}
   \label{fig:MOFF-flowchart}
@@
 
-Figure~\ref{fig:FX-flowchart} shows the flowchart for software holographic imaging from a FX correlator.
-The antenna-based F-engine is identical to that in the MOFF processing.
-The electric field spectra each antenna are then cross-multiplied in the X-engine with those from all other antennas to obtain the visibilities $V_\textrm{ij}(f)$ ($\mathbf{m}(\mathbf{v})$ in matrix notation).
-They are calibrated and time-averaged to obtain $\langle V_\textrm{ij}(f)\rangle$ which are then gridded and imaged to obtain the image $I(f)$.
-The $I(f)$ obtained from both techniques are identical as explained in \S\ref{sec:math}.
+Figure~\ref{fig:FX-flowchart} shows the flowchart for software holographic
+imaging from a FX correlator. The antenna-based F-engine is identical to that in
+the MOFF processing. The electric field spectra each antenna are then cross-
+multiplied in the X-engine with those from all other antennas to obtain the
+visibilities $V_\textrm{ij}(f)$ ($\mathbf{m}(\mathbf{v})$ in matrix notation).
+They are calibrated and time-averaged to obtain $\langle V_\textrm{ij}(f)\rangle$
+which are then gridded and imaged to obtain the image $I(f)$. The $I(f)$ obtained
+from both techniques are identical as explained in \S\ref{sec:math}.
 \begin{figure}
   \includegraphics[width=\columnwidth]{FX_flowchart.eps}
-  \caption{A flowchart of FX imaging in EPIC.
-The FX process flow shares the F-engine with the MOFF process.
-Following the F-engine, the electric fields pass through the X-engine to obtain visibilities $V_\textrm{ij}(f)$ which are calibrated and time-averaged.
-Then they are gridded to obtain the gridded visibilities $V_\textrm{g}(f)$ which are then imaged to obtain the image $I(f)$.}
+  \caption{A flowchart of FX imaging in EPIC. The FX process flow shares the
+    F-engine with the MOFF process. Following the F-engine, the electric fields
+    pass through the X-engine to obtain visibilities $V_\textrm{ij}(f)$ which are
+    calibrated and time-averaged. Then they are gridded to obtain the gridded
+    visibilities $V_\textrm{g}(f)$ which are then imaged to obtain the image
+    $I(f)$.}
   \label{fig:FX-flowchart}
@@
 
-This module is common to the MOFF and FX imaging techniques.
-Time samples of electric fields measured by the antenna and digitized by the A/D converter is Fourier transformed to generate electric field spectra.
-This step can be parallelized by antennas as shown in Figure~\ref{fig:f-engine}.
-The output is then fed to either MOFF and FX imaging pipelines.
+This module is common to the MOFF and FX imaging techniques. Time samples of
+electric fields measured by the antenna and digitized by the A/D converter is
+Fourier transformed to generate electric field spectra. This step can be
+parallelized by antennas as shown in Figure~\ref{fig:f-engine}. The output is
+then fed to either MOFF and FX imaging pipelines.
 
@@
   \includegraphics[width=\columnwidth]{F-engine.eps}
-  \caption{Block diagram of a F-engine. The electric field data streams from antennas are Fourier transformed in parallel to generate electric field spectra.}
+  \caption{Block diagram of a F-engine. The electric field data streams from
+    antennas are Fourier transformed in parallel to generate electric field
+    spectra.}
   \label{fig:f-engine}
@@
 
-A grid is generated on the coordinate system in which antenna locations are specified with a grid spacing that is at most $\lambda_\textrm{min}/2$ even at the highest frequency to ensure there is no aliasing even from regions of the sky far away from the field of view.
-The number of locations on the grid is restricted to be a power of 2.
+A grid is generated on the coordinate system in which antenna locations are
+specified with a grid spacing that is at most $\lambda_\textrm{min}/2$ even at the
+highest frequency to ensure there is no aliasing even from regions of the sky far
+away from the field of view. The number of locations on the grid is restricted to
+be a power of 2.
 
-The gridding kernel in the simplest case is given by the antenna aperture illumination function, $\widetilde{B}(\mathbf{r}-\mathbf{r}_a)$, which can specified either by a functional form or as a table of values against locations around the antennas.
-A nearest neighbor mapping from all antenna footprints to grid locations is created using an efficient k-d tree algorithm \citep{man99}.
-There is no restriction here that the aperture illumination function has to be identical across antennas.
+The gridding kernel in the simplest case is given by the antenna aperture
+illumination function, $\widetilde{B}(\mathbf{r}-\mathbf{r}_a)$, which can
+specified either by a functional form or as a table of values against locations
+around the antennas. A nearest neighbor mapping from all antenna footprints to
+grid locations is created using an efficient k-d tree algorithm \citep{man99}.
+There is no restriction here that the aperture illumination function has to be
+identical across antennas.
 
-In the most general case, this gridding kernel could contain information on the $w$-projection effect, and even other time-dependent ionospheric effects.
-For a stationary antenna array in the absence of any time-dependent effects, this mapping has to be determined only once in the antenna array coordinate frame.
-The antenna-to-grid mapping matrix, $\mathbf{M}(\mathbf{r},\mathbf{e})$ is described as a transformation matrix from the space of measured electric fields ($\mathbf{e}$) to the antenna array grid denoted by $\mathbf{r}$.
-Since each antenna occupies a footprint typically the size of its aperture, $\mathbf{M}(\mathbf{r},\mathbf{e})$ which is generally of size $N_\textrm{grid}\times N_\textrm{a}$, reduces to a sparse block-diagonal matrix with only $N_\textrm{a}$ blocks and roughly $N_\textrm{k}$ non-zero entries per block.
-Figure~\ref{fig:a2g-mapping} illustrates the antenna-to-grid mapping matrix and the grid containing the mapped aperture footprints of the antennas.
+In the most general case, this gridding kernel could contain information on the
+$w$-projection effect, and even other time-dependent ionospheric effects. For a
+stationary antenna array in the absence of any time-dependent effects, this
+mapping has to be determined only once in the antenna array coordinate frame. The
+antenna-to-grid mapping matrix, $\mathbf{M}(\mathbf{r},\mathbf{e})$ is described
+as a transformation matrix from the space of measured electric fields
+($\mathbf{e}$) to the antenna array grid denoted by $\mathbf{r}$. Since each
+antenna occupies a footprint typically the size of its aperture,
+$\mathbf{M}(\mathbf{r},\mathbf{e})$ which is generally of size
+$N_\textrm{grid}\times N_\textrm{a}$, reduces to a sparse block-diagonal matrix
+with only $N_\textrm{a}$ blocks and roughly $N_\textrm{k}$ non-zero entries per
+block. Figure~\ref{fig:a2g-mapping} illustrates the antenna-to-grid mapping
+matrix and the grid containing the mapped aperture footprints of the antennas.
 
@@
   \includegraphics[width=\columnwidth]{a2g_mapping.eps}
-  \caption{Block diagram of an antenna-to-grid mapping.
-A sparse block-diagonal matrix of total size $N_\textrm{grid}\times N_\textrm{a}$ is created where each block contains roughly the number of pixels covered by the respective kernel.
-The antenna aperture illumination kernels do not have to be identical to each other.
-A discrete set of arbitrarily placed antennas are now placed onto a regular grid.}
+  \caption{Block diagram of an antenna-to-grid mapping. A sparse block-diagonal
+    matrix of total size $N_\textrm{grid}\times N_\textrm{a}$ is created where each
+    block contains roughly the number of pixels covered by the respective kernel.
+    The antenna aperture illumination kernels do not have to be identical to each
+    other. A discrete set of arbitrarily placed antennas are now placed onto a
+    regular grid.}
   \label{fig:a2g-mapping}
@@
 
-The antenna array aperture illumination over the entire grid, $\widetilde{W}(\mathbf{r})$, is obtained by a projection of the individual antenna aperture illuminations:
+The antenna array aperture illumination over the entire grid,
+$\widetilde{W}(\mathbf{r})$, is obtained by a projection of the individual
+antenna aperture illuminations:
 \begin{align}
@@
 \end{align}
-where, $\mathcal{I}(\mathbf{e})$ is a row of ones.
-This is achieved using efficient multiplication with the sparse matrix created in the antenna-to-grid mapping process.
-Unless $\widetilde{W}(\mathbf{r})$ includes time-dependent effects of the ionosphere or the instrument, it needs to be computed just once for the entire observation.
-However, the gridding of electric fields must be computed at every readout of the electric field spectra.
-Thus,
+where, $\mathcal{I}(\mathbf{e})$ is a row of ones. This is achieved using
+efficient multiplication with the sparse matrix created in the antenna-to-grid
+mapping process. Unless $\widetilde{W}(\mathbf{r})$ includes time-dependent
+effects of the ionosphere or the instrument, it needs to be computed just once
+for the entire observation. However, the gridding of electric fields must be
+computed at every readout of the electric field spectra. Thus,
 \begin{align}
@@
 
-In practice the grid is set to be a power of 2 such that the grid spacing is at most $\lambda/2$ even at the highest frequency in the observing band.
+In practice the grid is set to be a power of 2 such that the grid spacing is at
+most $\lambda/2$ even at the highest frequency in the observing band.
 
@@
 
-Before the spatial Fourier transform, the gridded electric fields are padded with zeros in order to match the grid size and angular size of each image pixel that would have been obtained with software holography of output from an FX correlator.
+Before the spatial Fourier transform, the gridded electric fields are padded with
+zeros in order to match the grid size and angular size of each image pixel that
+would have been obtained with software holography of output from an FX
+correlator.
 
-In MOFF imaging, these are spatially Fourier transformed followed by a squaring operation at every timestamp for every frequency channel.
-In FX imaging, the spatial Fourier transform is performed only once per integration timescale and does not include a squaring operation.
+In MOFF imaging, these are spatially Fourier transformed followed by a squaring
+operation at every timestamp for every frequency channel. In FX imaging, the
+spatial Fourier transform is performed only once per integration timescale and
+does not include a squaring operation.
 
@@
 
-In MOFF imaging, the measured antenna electric fields and the corresponding holographic electric field images are zero-mean stochastic quantities.
-Hence, they cannot be time-averaged to reduce noise.
-The statistical quantity stable with time in this case are the square of the holographic electric field images.
-Thus, squared images have to be formed at every instant of time before averaging as indicated in equation~\ref{eqn:dirty-image-MOFF}.
+In MOFF imaging, the measured antenna electric fields and the corresponding
+holographic electric field images are zero-mean stochastic quantities. Hence,
+they cannot be time-averaged to reduce noise. The statistical quantity stable
+with time in this case are the square of the holographic electric field images.
+Thus, squared images have to be formed at every instant of time before averaging
+as indicated in equation~\ref{eqn:dirty-image-MOFF}.
 
-In contrast, visibilities measured by an antenna are statistically stable within an integration time interval.
-Hence, they are averaged after calibration as shown in equation~\ref{eqn:cc-vis}.
-It is advantageous to average them in visibilities before imaging because visibilities represent a compact representation of the information in images.
-Hence it is more computationally efficient to average antenna pair visibilities rather than images.
-Since this averaging has been performed already on the visibilities over an integration timescale, the imaging step has to be performed only once per integration cycle.
-FX imaging holds this advantage as long as the square of the number of antennas is smaller than the number of pixels on the image.
+In contrast, visibilities measured by an antenna are statistically stable within
+an integration time interval. Hence, they are averaged after calibration as shown
+in equation~\ref{eqn:cc-vis}. It is advantageous to average them in visibilities
+before imaging because visibilities represent a compact representation of the
+information in images. Hence it is more computationally efficient to average
+antenna pair visibilities rather than images. Since this averaging has been
+performed already on the visibilities over an integration timescale, the imaging
+step has to be performed only once per integration cycle. FX imaging holds this
+advantage as long as the square of the number of antennas is smaller than the
+number of pixels on the image.
 
@@
 
-We use the EPIC simulator to generate electric field samples from a sky model.
-In our example, we use 16 frequency channels each of width $\delta f=$100~kHz, 10 point sources of random flux densities at random locations.
-The number of timestamps in one integration cycle was kept at four where each A/D timeseries is $1/\delta f=10\,\mu$s long.
+We use the EPIC simulator to generate electric field samples from a sky model. In
+our example, we use 16 frequency channels each of width $\delta f=$100~kHz, 10
+point sources of random flux densities at random locations. The number of
+timestamps in one integration cycle was kept at four where each A/D timeseries is
+$1/\delta f=10\,\mu$s long.
 
@@
   \includegraphics[width=\columnwidth]{MOFF_FX_crossover_baseline_n-antennas_rho_fov_gridding.eps}
-  \caption{Current and instruments planned for future in parameter space of baseline length and number of antennas with MOFF and FX.}
+  \caption{Current and instruments planned for future in parameter space of
+    baseline length and number of antennas with MOFF and FX.}
   \label{fig:parameter-space-bll-nant-instruments}
@@
   \includegraphics[width=\columnwidth]{MOFF_FX_computations_fov_gridding_annotated.eps}
-  \caption{Current and instruments planned for future in parameter space of number of complex multiplies and adds with MOFF and FX.}
+  \caption{Current and instruments planned for future in parameter space of
+    number of complex multiplies and adds with MOFF and FX.}
   \label{fig:parameter-space-computations-instruments}

### 2015-10-22 imaging_demo/ms.tex

@@
 \subsection{Calibration}
+Calibration of direct imaging correlator remains an unresolved matter. Contrary
+to the FX data flow, direct imagers mix the signals from all antennas before
+averaging and writing to disk. It is therefore essential to apply gain solutions 
+before the gridding step. Previous efforts have resorted to applying FX-generated
+calibration solutions \citep{zhe14, fos14}, or integrate a dedicated FX correlator 
+which periodically forms the full visibility matrix \citep{wij09, dev09}.
+
+In a companion paper to appear soon, we demonstrate a novel calibration 
+technique (EPICal) which leverages the data products formed by direct imaging 
+correlators to estimate antenna complex gains. This method correlates the antenna 
+electric field signals with an image pixel form the output of the correlator in the 
+feedback calibration fashion outlined in \citealt{mor11} (illustrated in Figure 
+\ref{fig:MOFF-flowchart} by the arrow leading from the imager to the calibration 
+block). Furthermore it allows for arbitrarily complex sky models, and following the 
+MOFF algorithm places no restriction on array layout, and accounts for non-identical 
+antenna beam patterns. Because only a single correlation is needed for each 
+antenna, the computation complexity scales only as $N_{\mathrm{ant}}$.
+
+The calibration module included in the EPIC repository allows for application of pre-
+determined calibration solutions, or can solve for the complex gains using the 
+EPICal algorithm.
 

### 2015-10-22 imaging_demo/ms.tex
tried my hand at reworking math framework

@@
 
-We provide a quick summary of the mathematical equivalence of the MOFF and FX
-correlators detailed in \citet{mor11}.
+We provide a brief summary of the amthematical equivalence of the MOFF and FX correlators detailed in \citet{mor11}.
+We first relate the dirty image produced from visibilities to the electric fields of astrophysical sources, then show that operations can be reordered to produce the same images at a lower computational cost.
 
 Electric fields from astrophysical sources, $E(\hat{\mathbf{s}})$, in the sky
-coordinate system denoted by unit vector $\hat{\mathbf{s}}$ propagate towards
-the observer as:
+coordinate system denoted by sine-projected unit vector $\hat{\mathbf{s}}$, 
+propagate towards the observer as:
 \begin{align}
@@
 \end{align}
-where, $\mathbf{r}$ denotes the observer's coordinate system and $\widetilde{E}
+where, $\mathbf{r}$ denotes the observer's location (measured in wavelengths relative to some arbitrary origin) and $\widetilde{E}
 (\mathbf{r})$ is the propagated electric field. Thus the propagated electric
@@
                   &= \int \widetilde{W}_a(\mathbf{r}-\mathbf{r}_a) \left[ \int E(\hat{\mathbf{s}})\,e^{-i2\pi\mathbf{r}\cdot\hat{\mathbf{s}}}\,\dif^2\hat{\mathbf{s}} \right] \dif^2\mathbf{r} + \widetilde{n}_a \\
-                  &= \int \delta(\mathbf{r}-\mathbf{r}_a) \left[ \int {W}_a(\hat{\mathbf{s}})\,E(\hat{\mathbf{s}})\,e^{-i2\pi\mathbf{r}\cdot\hat{\mathbf{s}}}\,\dif^2\hat{\mathbf{s}} \right] \dif^2\mathbf{r} + \widetilde{n}_a
+                  &= \int {W}_a(\hat{\mathbf{s}})\,E(\hat{\mathbf{s}})\,e^{-i2\pi\mathbf{r}_a\!\cdot\,\hat{\mathbf{s}}}\,\dif^2\hat{\mathbf{s}} + \widetilde{n}_a
 \end{align}
 where, $\widetilde{W}_a(\mathbf{r})$ is the aperture electric field illumination
-pattern of the antenna and its Fourier transform, $W_a(\hat{\mathbf{s}})$ is the
+pattern of the antenna and its Fourier transform, $W_a(\hat{\mathbf{s}})$, is the
 directional antenna voltage response.
 
-Interferometers measure {\it visibility} -- the degree of coherence -- between
+Interferometers measure {\it visibilities} -- the degree of coherence -- between
 electric fields measured by a pair of antennas \citep{van34,zer38,tho01}.
-Visibility, $\widetilde{V}_p$, can be written as:
-
+A visibility, $\widetilde{V}_p$, can be written as:
 \begin{align}
-  \widetilde{V}_p &= \left\langle \widetilde{E}_a\widetilde{E}_b^\star \right\rangle \label{eqn:cc-vis}\\
-                  &= \int\delta(\mathbf{u}-\mathbf{u}_p)\left[\int e^{-i2\pi\mathbf{u}.\hat{\mathbf{s}}}\,B(\hat{\mathbf{s}})\,I(\hat{\mathbf{s}})\,\dif^2\hat{\mathbf{s}}\right] \dif^2\mathbf{u} + n_p,
+  \widetilde{V}_p = &\left\langle \widetilde{E}_a\widetilde{E}_b^\star \right\rangle_t \label{eqn:cc-vis}\hfill\\
+  		= &\left\langle \left[ \int {W}_a(\hat{\mathbf{s}})\,E(\hat{\mathbf{s}})\,e^{-i2\pi\mathbf{r}_a\!\cdot\,\hat{\mathbf{s}}}\,\dif^2\hat{\mathbf{s}} + \widetilde{n}_a \right] \right. \nonumber\\
+		&\qquad \times \left.\left[ \int {W}^\star_b(\hat{\mathbf{s'}})\,E^\star (\hat{\mathbf{s'}})\,e^{i2\pi\mathbf{r}_b\!\cdot\,\hat{\mathbf{s'}}}\,\dif^2\hat{\mathbf{s'}} + \widetilde{n}^\star_b \right] \right\rangle_t \\
+                  =& \iint {W}_a(\hat{\mathbf{s}}) {W}^\star_b(\hat{\mathbf{s'}}) \left\langle E(\hat{\mathbf{s}}) E^\star(\hat{\mathbf{s'}}) \right\rangle_t e^{-i2\pi(\mathbf{r}_a\!\cdot\,\hat{\mathbf{s}}-\mathbf{r}_b\!\cdot\,\hat{\mathbf{s'}})}\,\dif^2\hat{\mathbf{s}}\,\dif^2\hat{\mathbf{s'}},
 \end{align}
-where, $I(\hat{\mathbf{s}})$ is the sky brightness, $B(\hat{\mathbf{s}})$ is the
-antenna power response, $\mathbf{u}$ denotes the coordinates of the $uv$-
-measurement plane, $\mathbf{u}_p\equiv \mathbf{r}_a-\mathbf{r}_b$ is the antenna
-spacing vector corresponding to the antenna pair $p$, and $n_p$ is the receiver
-noise. This signifies that the visibility ($\widetilde{V}_p$) measured between a
+where we have brought the time average into the integral under the assumption that the aperture illumination pattern does not change over the time-scale of the averaging.
+This expression can be further simplified with the sky brightness, $I(\hat{\mathbf{s}}) \delta(\hat{\mathbf{s}}-\hat{\mathbf{s'}})= \left\langle E(\hat{\mathbf{s}}) E^\star(\hat{\mathbf{s'}}) \right\rangle_t $, and defining the antenna pair sky power response function (or the primary beam), $B_p(\hat{\mathbf{s}})~\equiv~{W}_a(\hat{\mathbf{s}})~{W}^\star_b(\hat{\mathbf{s}})$.
+The result is the visibility expressed in terms of the sky brightness, the primary beam, and uncorrelated noise terms which we group into $\widetilde{n}_p$,
+\begin{equation}
+\widetilde{V}_p = \int e^{-i2\pi\mathbf{u}_p\!\cdot\,\hat{\mathbf{s}}}\,B_p(\hat{\mathbf{s}})\,I(\hat{\mathbf{s}})\,\dif^2\hat{\mathbf{s}} + \widetilde{n}_p,
+\end{equation}
+where the baseline coordinate $\mathbf{u}_p=\mathbf{r}_a-\mathbf{r}_b$ is the vector separation between the two antennas. This signifies that the visibility ($\widetilde{V}_p$) measured between a
 pair of antennas ($p$) is obtained by the multiplying the sky brightness
 $I(\hat{\mathbf{s}})$ by the antenna power response $B(\hat{\mathbf{s}})$ and
-Fourier transform from the directional coordinates ($\hat{\mathbf{s}}$) to $uv$
+Fourier transforming from the directional coordinates ($\hat{\mathbf{s}}$) to $uv$
 coordinates, which are then sampled at the locations of the antenna spacings (or
-baselines), namely, $\mathbf{u}_p$ using the sampling function
-$\delta(\mathbf{u}-\mathbf{u}_p)$ and added to the receiver noise $n_p$. This can
+baselines), namely, $\mathbf{u}_p$, and added to the receiver noise $n_p$.
+
+This can
 be equivalently re-written as:
 \begin{align}\label{eqn:software-holography}
-  \widetilde{V}_p &= \int\delta(\mathbf{u}^\prime-\mathbf{u}_p)\left[\int \tilde{B}(\mathbf{u}^\prime-\mathbf{u}) \right. \nonumber\\
-  &\left. \qquad \times \left[\int e^{-i2\pi\mathbf{u}.\hat{\mathbf{s}}}\,I(\hat{\mathbf{s}})\,\dif^2\hat{\mathbf{s}}\right]\dif^2\mathbf{u}\vphantom{\int}\right] \dif^2\mathbf{u}^\prime + n_p,
+  \widetilde{V}_p &= \int \tilde{B}(\mathbf{u}^\prime-\mathbf{u}) \times \left[\int e^{-i2\pi\mathbf{u}.\hat{\mathbf{s}}}\,I(\hat{\mathbf{s}})\,\dif^2\hat{\mathbf{s}}\right]\dif^2\mathbf{u}\vphantom{\int} + n_p,
 \end{align}
@@
 convolution with $\tilde{B}(\mathbf{u})$ in $uv$-space. This is the software
-holographic equivalent of traditional FX correlator output. Following the matrix
-notation of \citet{mor11}, the above equation can be expressed as:
+holographic equivalent of traditional FX correlator output.
+
+Following the matrix
+notation of \citet{mor11}, the above measurement equation can be expressed as:
 \begin{align}
@@
 \end{align}
-where, the sky brightness $I(\hat{\mathbf{s}})$ is Fourier transformed using
+where the sky brightness $I(\hat{\mathbf{s}})$ is Fourier transformed using
 $\mathbf{F}(\mathbf{u},\hat{\mathbf{s}})$ and the resultant spatial coherence
@@
 \begin{align}
-  \mathbf{m}(\mathbf{v}) &= \left\langle \widetilde{E}(\mathbf{a})\ast\widetilde{E}(\mathbf{a}^\prime)\right\rangle_t. \label{eqn:matrix-cc-vis}
+  \mathbf{m}(\mathbf{v}) &= \left\langle \widetilde{E}^\star(\mathbf{a})\widetilde{E}(\mathbf{a}^\prime)\right\rangle_t, \label{eqn:matrix-cc-vis}
 \end{align}
-$\mathbf{m}(\mathbf{v})$ denotes visibilities measured by cross-correlating
+where $\mathbf{m}(\mathbf{v})$ denotes visibilities measured by cross-correlating
 measured antenna electric fields over all possible pairs of $\mathbf{a}$ and
@@
 \begin{align}
-  I^\prime(\hat{\mathbf{s}}) &= \mathbf{F}^\textrm{T}(\hat{\mathbf{s}},\mathbf{u})\,\widetilde{B}^{\,\textrm{T}}(\mathbf{u},\mathbf{v})\,\mathbf{N}^{-1}\,\mathbf{m}(\mathbf{v}) \label{eqn:dirty-image-FX}
+  I^\prime(\hat{\mathbf{s}}) &= \mathbf{F}^\textrm{T}(\hat{\mathbf{s}},\mathbf{u})\,\widetilde{B}^{\,\textrm{T}}(\mathbf{u},\mathbf{v})\,\mathbf{N}^{-1}(\mathbf{v},\mathbf{v})\,\mathbf{m}(\mathbf{v}) \label{eqn:dirty-image-FX}
 \end{align}
 where the measured visibilities are weighted by the inverse of the system noise,
-following a gridding process using the holographic antenna power response as the
-gridding kernel followed by a Fourier transform to create an image
-$I^\prime(\hat{\mathbf{s}})$ which is the optimal estimate of the true image
-$I(\hat{\mathbf{s}})$.
+followed by a gridding process using the holographic antenna power response as the
+gridding kernel, followed by a Fourier transform to create an image
+$I^\prime(\hat{\mathbf{s}})$. This is the optimal estimate of the true image $I(\hat{\mathbf{s}})$ given the visibility measurements.
 
-It may be noted that the antenna power response in $uv$-plane can be written as:
-\begin{align}\label{eqn:antenna-power-breakup}
-  \tilde{B}_p(\mathbf{u}) &= \widetilde{W}_a(\mathbf{u}) \ast \widetilde{W}_b^\star(\mathbf{u})
-\end{align}
-where, $W_a(\mathbf{u})$ and $W_b(\mathbf{u})$ denote the individual antenna
-voltage responses. Thus $\tilde{B}_p(\mathbf{u})$ is a cross-correlation of the
-individual antenna voltage responses. Similarly, it can be shown that 
+The intermediate step of gridding with the antenna power response can be expressed as a convolution of a data vector generated by gridding the electric fields directly with the antenna illumination pattern.
 \begin{align}
-  \left\langle E(\hat{\mathbf{s}})\,E^\star(\hat{\mathbf{s}}^\prime)\right\rangle &= \left\langle E(\hat{\mathbf{s}})\,E^\star(\hat{\mathbf{s}}^\prime) \delta(\hat{\mathbf{s}}-\hat{\mathbf{s}}^\prime) \right\rangle \nonumber\\
-  &= \left\langle \left|E(\hat{\mathbf{s}})\right|^2 \right\rangle = I(\hat{\mathbf{s}}) \label{eqn:sky-intensity}
+&\widetilde{B}^{\,\textrm{T}}(\mathbf{u},\mathbf{v})\,\mathbf{N}^{-1}(\mathbf{v} ,\mathbf{v})\,\mathbf{m}(\mathbf{v})=\nonumber\\
+&\qquad\left\langle \left[\widetilde{\mathbf{W}}^\textrm{T}_a(\mathbf{r},\mathbf{a})\, \widetilde{\mathbf{N}}^{-1}\!(\mathbf{a},\mathbf{a})\, \widetilde{\mathbf{E}}(\mathbf{a})\right]
+\ast\left[\widetilde{\mathbf{W}}_a(\mathbf{r},\mathbf{a})\, \mathbf{N}^{-1}\!(\mathbf{a},\mathbf{a})\, \mathbf{E}^\star(\mathbf{a})\right]\right\rangle_t\label{eqn:e-field-conv}
 \end{align}
 
-From equations~\ref{eqn:measured-E-field}, \ref{eqn:software-holography},
-\ref{eqn:matrix-cc-vis}, \ref{eqn:antenna-power-breakup} and
-\ref{eqn:sky-intensity}, the optimal image in equation \ref{eqn:dirty-image-FX}
-can be expressed equivalently as:
+We can then use the multiplication-convolution theorem to move the convolution in Equation \ref{eqn:e-field-conv} to a square after the Fourier transform in Equation \ref{eqn:dirty-image-FX}.
 \begin{align}
-  I^\prime(\hat{\mathbf{s}}) &= \left\langle \left|\,\mathbf{F}^\textrm{T}(\hat{\mathbf{s}},\mathbf{r})\,\widetilde{\mathbf{W}}^\textrm{T}(\mathbf{r},\mathbf{a})\,\widetilde{\mathbf{N}}(\mathbf{a},\mathbf{a})\,\widetilde{E}(\mathbf{a})\,\right|^2\right\rangle_t.
+  I^\prime(\hat{\mathbf{s}}) &= \left\langle \left|\,\mathbf{F}^\textrm{T}(\hat{\mathbf{s}},\mathbf{r})\,\widetilde{\mathbf{W}}^\textrm{T}(\mathbf{r},\mathbf{a})\,\widetilde{\mathbf{N}}^{-1}(\mathbf{a},\mathbf{a})\,\widetilde{E}(\mathbf{a})\,\right|^2\right\rangle_t.
 \label{eqn:dirty-image-MOFF}
@@
 
-Equations~\ref{eqn:dirty-image-MOFF} and \ref{eqn:dirty-image-FX} are identical
-to each other obtained as a consequence of the multiplication-convolution theorem
-of Fourier transforms. In other words, equation~\ref{eqn:dirty-image-FX} is
-obtained by a Fourier transform of the gridded spatial correlation of electric
-fields measured by antennas while equation~\ref{eqn:dirty-image-MOFF} is obtained
-by squaring the Fourier transform of the gridded electric fields measured by the
-antennas. The time-scale of averaging in equations~\ref{eqn:cc-vis} and
-\ref{eqn:dirty-image-MOFF} is set by the coherence time-scale of the electric
-fields or the science requirement (e.g., timescale of variability or transients).
+Equation \ref{eqn:dirty-image-MOFF} is the optimal imaging equation used by the MOFF algorithm.
+While mathematically equivalent to Equation~\ref{eqn:dirty-image-FX}, squaring in image space rather than convolving in $uv$ space potentially saves orders of magnitude in computation.
 
@@
   between antenna pairs represent spatial correlations which can be time-averaged
-  following which the they gridded and imaged. However, in MOFF imaging both
-  antenna and gridded electric fields are stochastic and therefore have to be
-  imaged before time-averaging. 
-\item In FX imaging, electric fields measured on antennas are not correlated with
+  followed by gridding and imaging. However, in MOFF imaging both
+  antenna and gridded electric fields are stochastic and therefore must be
+  imaged and squared before time-averaging. 
+\item In FX imaging, electric fields measured by antennas are not correlated with
   themselves and hence lack zero spacing measurements. In contrast, in MOFF
@@
   contain information from auto-correlated electric fields at zero spacing.
-  Hence, they have to be subtracted from the images.
+  Hence, they must be subtracted from the images.
 \end{enumerate}
+
+%%%%%%%%%
+
+%We provide a brief summary of the mathematical equivalence of the MOFF and FX
+%correlators detailed in \citet{mor11}. 
+%
+%Electric fields from astrophysical sources, $E(\hat{\mathbf{s}})$, in the sky
+%coordinate system denoted by unit vector $\hat{\mathbf{s}}$ propagate towards
+%the observer as:
+%\begin{align}
+%  \widetilde{E}(\mathbf{r}) &= \int E(\hat{\mathbf{s}})\,e^{-i2\pi\mathbf{r}\cdot\hat{\mathbf{s}}}\,\dif^2\hat{\mathbf{s}},
+%\end{align}
+%where, $\mathbf{r}$ denotes the observer's coordinate system and $\widetilde{E}
+%(\mathbf{r})$ is the propagated electric field. Thus the propagated electric
+%field is a linear superposition of the electric fields emanating from
+%astronomical sources with appropriate complex phases. It can also be described
+%as a Fourier transform of the electric fields in the sky coordinates. 
+%
+%An antenna, $a$, measures a phased sum of these propagated electric fields over
+%its effective collecting area with an additive receiver noise:
+%\begin{align}\label{eqn:measured-E-field}
+%  \widetilde{E}_a &= \int \widetilde{W}_a(\mathbf{r}-\mathbf{r}_a)\,\widetilde{E}(\mathbf{r})\,\dif^2\mathbf{r} + \widetilde{n}_a \\
+%                  &= \int \widetilde{W}_a(\mathbf{r}-\mathbf{r}_a) \left[ \int E(\hat{\mathbf{s}})\,e^{-i2\pi\mathbf{r}\cdot\hat{\mathbf{s}}}\,\dif^2\hat{\mathbf{s}} \right] \dif^2\mathbf{r} + \widetilde{n}_a \\
+%                  &= \int \delta(\mathbf{r}-\mathbf{r}_a) \left[ \int {W}_a(\hat{\mathbf{s}})\,E(\hat{\mathbf{s}})\,e^{-i2\pi\mathbf{r}\cdot\hat{\mathbf{s}}}\,\dif^2\hat{\mathbf{s}} \right] \dif^2\mathbf{r} + \widetilde{n}_a
+%\end{align}
+%where, $\widetilde{W}_a(\mathbf{r})$ is the aperture electric field illumination
+%pattern of the antenna and its Fourier transform, $W_a(\hat{\mathbf{s}})$ is the
+%directional antenna voltage response.
+%
+%Interferometers measure {\it visibility} -- the degree of coherence -- between
+%electric fields measured by a pair of antennas \citep{van34,zer38,tho01}.
+%Visibility, $\widetilde{V}_p$, can be written as:
+%
+%\begin{align}
+%  \widetilde{V}_p &= \left\langle \widetilde{E}_a\widetilde{E}_b^\star \right\rangle \label{eqn:cc-vis}\\
+%                  &= \int\delta(\mathbf{u}-\mathbf{u}_p)\left[\int e^{-i2\pi\mathbf{u}.\hat{\mathbf{s}}}\,B(\hat{\mathbf{s}})\,I(\hat{\mathbf{s}})\,\dif^2\hat{\mathbf{s}}\right] \dif^2\mathbf{u} + n_p,
+%\end{align}
+%where, $I(\hat{\mathbf{s}})$ is the sky brightness, $B(\hat{\mathbf{s}})$ is the
+%antenna power response, $\mathbf{u}$ denotes the coordinates of the $uv$-
+%measurement plane, $\mathbf{u}_p\equiv \mathbf{r}_a-\mathbf{r}_b$ is the antenna
+%spacing vector corresponding to the antenna pair $p$, and $n_p$ is the receiver
+%noise. This signifies that the visibility ($\widetilde{V}_p$) measured between a
+%pair of antennas ($p$) is obtained by the multiplying the sky brightness
+%$I(\hat{\mathbf{s}})$ by the antenna power response $B(\hat{\mathbf{s}})$ and
+%Fourier transform from the directional coordinates ($\hat{\mathbf{s}}$) to $uv$
+%coordinates, which are then sampled at the locations of the antenna spacings (or
+%baselines), namely, $\mathbf{u}_p$ using the sampling function
+%$\delta(\mathbf{u}-\mathbf{u}_p)$ and added to the receiver noise $n_p$. This can
+%be equivalently re-written as:
+%\begin{align}\label{eqn:software-holography}
+%  \widetilde{V}_p &= \int\delta(\mathbf{u}^\prime-\mathbf{u}_p)\left[\int \tilde{B}(\mathbf{u}^\prime-\mathbf{u}) \right. \nonumber\\
+%  &\left. \qquad \times \left[\int e^{-i2\pi\mathbf{u}.\hat{\mathbf{s}}}\,I(\hat{\mathbf{s}})\,\dif^2\hat{\mathbf{s}}\right]\dif^2\mathbf{u}\vphantom{\int}\right] \dif^2\mathbf{u}^\prime + n_p,
+%\end{align}
+%where, $\tilde{B}(\mathbf{u})$ denotes the $uv$-space antenna power response
+%obtained by a Fourier transform of $B(\hat{\mathbf{s}})$. Effectively, the
+%multiplication in image space by $B(\hat{\mathbf{s}})$ has been replaced by a
+%convolution with $\tilde{B}(\mathbf{u})$ in $uv$-space. This is the software
+%holographic equivalent of traditional FX correlator output. Following the matrix
+%notation of \citet{mor11}, the above equation can be expressed as:
+%\begin{align}
+%  \mathbf{m}(\mathbf{v}) &= \widetilde{B}(\mathbf{v},\mathbf{u})\,\mathbf{F}(\mathbf{u},\hat{\mathbf{s}})\,I(\hat{\mathbf{s}}) + \mathbf{n}(\mathbf{v}),
+%\end{align}
+%where, the sky brightness $I(\hat{\mathbf{s}})$ is Fourier transformed using
+%$\mathbf{F}(\mathbf{u},\hat{\mathbf{s}})$ and the resultant spatial coherence
+%function is weighted and summed using the antenna power response,
+%$\widetilde{B}(\mathbf{v},\mathbf{u})$ in $uv$-space sampled at the baseline
+%location to obtain the measured visibilities:
+%\begin{align}
+%  \mathbf{m}(\mathbf{v}) &= \left\langle \widetilde{E}(\mathbf{a})\ast\widetilde{E}(\mathbf{a}^\prime)\right\rangle_t. \label{eqn:matrix-cc-vis}
+%\end{align}
+%$\mathbf{m}(\mathbf{v})$ denotes visibilities measured by cross-correlating
+%measured antenna electric fields over all possible pairs of $\mathbf{a}$ and
+%$\mathbf{a}^\prime$. It is the same as equation~\ref{eqn:cc-vis} written in
+%matrix notation.
+%
+%Using the optimal map-making formalism \citep{teg97a,teg97b}, a software
+%holography image is formed using \citep{mor09}:
+%\begin{align}
+%  I^\prime(\hat{\mathbf{s}}) &= \mathbf{F}^\textrm{T}(\hat{\mathbf{s}},\mathbf{u})\,\widetilde{B}^{\,\textrm{T}}(\mathbf{u},\mathbf{v})\,\mathbf{N}^{-1}\,\mathbf{m}(\mathbf{v}) \label{eqn:dirty-image-FX}
+%\end{align}
+%where the measured visibilities are weighted by the inverse of the system noise,
+%following a gridding process using the holographic antenna power response as the
+%gridding kernel followed by a Fourier transform to create an image
+%$I^\prime(\hat{\mathbf{s}})$ which is the optimal estimate of the true image
+%$I(\hat{\mathbf{s}})$.
+
+%It may be noted that the antenna power response in $uv$-plane can be written as:
+%\begin{align}\label{eqn:antenna-power-breakup}
+%  \tilde{B}_p(\mathbf{u}) &= \widetilde{W}_a(\mathbf{u}) \ast \widetilde{W}_b^\star(\mathbf{u})
+%\end{align}
+%where, $W_a(\mathbf{u})$ and $W_b(\mathbf{u})$ denote the individual antenna
+%voltage responses. Thus $\tilde{B}_p(\mathbf{u})$ is a cross-correlation of the
+%individual antenna voltage responses. Similarly, it can be shown that 
+%\begin{align}
+%  \left\langle E(\hat{\mathbf{s}})\,E^\star(\hat{\mathbf{s}}^\prime)\right\rangle &= \left\langle E(\hat{\mathbf{s}})\,E^\star(\hat{\mathbf{s}}^\prime) \delta(\hat{\mathbf{s}}-\hat{\mathbf{s}}^\prime) \right\rangle \nonumber\\
+%  &= \left\langle \left|E(\hat{\mathbf{s}})\right|^2 \right\rangle = I(\hat{\mathbf{s}}) \label{eqn:sky-intensity}
+%\end{align}
+%
+%From equations~\ref{eqn:measured-E-field}, \ref{eqn:software-holography},
+%\ref{eqn:matrix-cc-vis}, \ref{eqn:antenna-power-breakup} and
+%\ref{eqn:sky-intensity}, the optimal image in equation \ref{eqn:dirty-image-FX}
+%can be expressed equivalently as:
+%\begin{align}
+%  I^\prime(\hat{\mathbf{s}}) &= \left\langle \left|\,\mathbf{F}^\textrm{T}(\hat{\mathbf{s}},\mathbf{r})\,\widetilde{\mathbf{W}}^\textrm{T}(\mathbf{r},\mathbf{a})\,\widetilde{\mathbf{N}}(\mathbf{a},\mathbf{a})\,\widetilde{E}(\mathbf{a})\,\right|^2\right\rangle_t.
+\label{eqn:dirty-image-MOFF}
+%\end{align}
+%The term inside the angular brackets before squaring has a very similar form as
+%that in equation~\ref{eqn:dirty-image-FX}. It signifies that the measured antenna
+%electric fields are weighted by the antenna noise, weighted and gridded by the
+%antenna aperture kernel, Fourier transformed and finally squared to obtain the
+%same image estimated that would have been obtained using equation~
+%\ref{eqn:dirty-image-FX}.
+
+%Equations~\ref{eqn:dirty-image-MOFF} and \ref{eqn:dirty-image-FX} are identical
+%to each other obtained as a consequence of the multiplication-convolution theorem
+%of Fourier transforms. In other words, equation~\ref{eqn:dirty-image-FX} is
+%obtained by a Fourier transform of the gridded spatial correlation of electric
+%fields measured by antennas while equation~\ref{eqn:dirty-image-MOFF} is obtained
+%by squaring the Fourier transform of the gridded electric fields measured by the
+%antennas. The time-scale of averaging in equations~\ref{eqn:cc-vis} and
+%\ref{eqn:dirty-image-MOFF} is set by the coherence time-scale of the electric
+%fields or the science requirement (e.g., timescale of variability or transients).
+%
+%There are some important differences between the two techniques:
+%\begin{enumerate}
+%\item The time-averaging cannot be performed on a stochastic measurement but
+%  only on its statistical properties. In FX imaging, the visibilities measured
+%  between antenna pairs represent spatial correlations which can be time-averaged
+%  following which the they gridded and imaged. However, in MOFF imaging both
+%  antenna and gridded electric fields are stochastic and therefore have to be
+%  imaged before time-averaging. 
+%\item In FX imaging, electric fields measured on antennas are not correlated with
+%  themselves and hence lack zero spacing measurements. In contrast, in MOFF
+%  imaging, since the gridded electric fields are imaged and squared, they
+%  contain information from auto-correlated electric fields at zero spacing.
+%  Hence, they have to be subtracted from the images.
+%\end{enumerate}
 

### 2015-10-22 imaging_demo/ms.tex
forgot to break up lines

@@
 $\Npix \log \Npix$ computational scaling without placing any restriction on
-antenna placement, can accommodate non-identical beam patterns, and is a provably
-optimal mapping \citep{mor11}. This algorithm uses the antenna beam patterns to
-grid the electric field measurements to a regular grid in the software
+antenna placement, can accommodate non-identical beam patterns, and is 
+a provably optimal mapping \citep{mor11}. This algorithm uses the antenna beam 
+patterns to grid the electric field measurements to a regular grid in the software
 holography/A-transpose fashion \citep{mor09,bha08,teg97b} before performing the
@@
 
-We provide a brief summary of the amthematical equivalence of the MOFF and FX correlators detailed in \citet{mor11}.
-We first relate the dirty image produced from visibilities to the electric fields of astrophysical sources, then show that operations can be reordered to produce the same images at a lower computational cost.
+We provide a brief summary of the amthematical equivalence of the MOFF and 
+FX correlators detailed in \citet{mor11}. We first relate the dirty image produced 
+from visibilities to the electric fields of astrophysical sources, then show that 
+operations can be reordered to produce the same images at a lower 
+computational cost.
 
@@
 \end{align}
-where, $\mathbf{r}$ denotes the observer's location (measured in wavelengths relative to some arbitrary origin) and $\widetilde{E}
+where, $\mathbf{r}$ denotes the observer's location (measured in wavelengths 
+relative to some arbitrary origin) and $\widetilde{E}
 (\mathbf{r})$ is the propagated electric field. Thus the propagated electric
@@
 \end{align}
-where we have brought the time average into the integral under the assumption that the aperture illumination pattern does not change over the time-scale of the averaging.
-This expression can be further simplified with the sky brightness, $I(\hat{\mathbf{s}}) \delta(\hat{\mathbf{s}}-\hat{\mathbf{s'}})= \left\langle E(\hat{\mathbf{s}}) E^\star(\hat{\mathbf{s'}}) \right\rangle_t $, and defining the antenna pair sky power response function (or the primary beam), $B_p(\hat{\mathbf{s}})~\equiv~{W}_a(\hat{\mathbf{s}})~{W}^\star_b(\hat{\mathbf{s}})$.
-The result is the visibility expressed in terms of the sky brightness, the primary beam, and uncorrelated noise terms which we group into $\widetilde{n}_p$,
+where we have brought the time average into the integral under the assumption that 
+the aperture illumination pattern does not change over the time-scale of the 
+averaging. This expression can be further simplified with the sky brightness, 
+$I(\hat{\mathbf{s}}) \delta(\hat{\mathbf{s}}-\hat{\mathbf{s'}})= \left\langle 
+E(\hat{\mathbf{s}}) E^\star(\hat{\mathbf{s'}}) \right\rangle_t $, and defining the 
+antenna pair sky power response function (or the primary beam), 
+$B_p(\hat{\mathbf{s}})~\equiv~{W}_a(\hat{\mathbf{s}})~{W}^\star_b(\hat{\mathbf{s}})$. 
+The result is the visibility expressed in terms of the sky brightness, the primary 
+beam, and uncorrelated noise terms which we group into $\widetilde{n}_p$,
 \begin{equation}
@@
 \end{equation}
-where the baseline coordinate $\mathbf{u}_p=\mathbf{r}_a-\mathbf{r}_b$ is the vector separation between the two antennas. This signifies that the visibility ($\widetilde{V}_p$) measured between a
+where the baseline coordinate $\mathbf{u}_p=\mathbf{r}_a-\mathbf{r}_b$ is the 
+vector separation between the two antennas. This signifies that the visibility 
+($\widetilde{V}_p$) measured between a
 pair of antennas ($p$) is obtained by the multiplying the sky brightness
@@
 where the measured visibilities are weighted by the inverse of the system noise,
-followed by a gridding process using the holographic antenna power response as the
-gridding kernel, followed by a Fourier transform to create an image
-$I^\prime(\hat{\mathbf{s}})$. This is the optimal estimate of the true image $I(\hat{\mathbf{s}})$ given the visibility measurements.
+followed by a gridding process using the holographic antenna power response as 
+the gridding kernel, followed by a Fourier transform to create an image
+$I^\prime(\hat{\mathbf{s}})$. This is the optimal estimate of the true image 
+$I(\hat{\mathbf{s}})$ given the visibility measurements.
 
-The intermediate step of gridding with the antenna power response can be expressed as a convolution of a data vector generated by gridding the electric fields directly with the antenna illumination pattern.
+The intermediate step of gridding with the antenna power response can be 
+expressed as a convolution of a data vector generated by gridding the electric fields 
+directly with the antenna illumination pattern.
 \begin{align}
@@
 
-We can then use the multiplication-convolution theorem to move the convolution in Equation \ref{eqn:e-field-conv} to a square after the Fourier transform in Equation \ref{eqn:dirty-image-FX}.
+We can then use the multiplication-convolution theorem to move the convolution in 
+Equation \ref{eqn:e-field-conv} to a square after the Fourier transform in Equation 
+\ref{eqn:dirty-image-FX}.
 \begin{align}
@@
 
-Equation \ref{eqn:dirty-image-MOFF} is the optimal imaging equation used by the MOFF algorithm.
-While mathematically equivalent to Equation~\ref{eqn:dirty-image-FX}, squaring in image space rather than convolving in $uv$ space potentially saves orders of magnitude in computation.
+Equation \ref{eqn:dirty-image-MOFF} is the optimal imaging equation used by the 
+MOFF algorithm. While mathematically equivalent to Equation~
+\ref{eqn:dirty-image-FX}, squaring in image space rather than convolving in $uv$ 
+space potentially saves orders of magnitude in computation.
 

### 2015-10-23 imaging_demo/ms.tex
Add more text and figures in sections 3 and 4
From a3bda8c (master):
=====================

In epic.bib,

(*) Rearrange references alphabetically
(*) Add Beardsley et al. (2012) to references

In ms.tex,

(*) Add more text and figures in sections 3 and 4

--- 2015-10-22 imaging_demo/ms.tex
@@
 antenna aperture illuminations:
-\begin{align}
+\begin{align}\label{eqn:gridding-convolution}
   \widetilde{W}(\mathbf{r}) &= \sum_a \widetilde{W}_a(\mathbf{r}-\mathbf{r}_a) \\
@@
 
+\subsection{Dealing with antenna auto-correlations}\label{sec:rm-autocorr}
+
+The squaring operation under MOFF imaging in the image plane introduces antenna 
+auto-correlations around the zero spacing in the $uv$-plane which are absent in 
+traditional visibility-based imaging. In order for a robust comparison, these 
+auto-correlations have to be removed from both the $uv$ and image planes.
+
+The shape and extent of these auto-correlations can be estimated from the 
+antenna aperture illumination pattern which is a quantity already available 
+from the gridding step. Figure~\ref{fig:autocorr_wts_PB} shows the estimated 
+weights from antenna auto-correlations in the $uv$-plane (left) and the 
+corresponding response in the image plane (right). The latter is simply the 
+directional antenna power response.
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{autocorr_uvwts_pbeam.png}
+  \caption{The auto-correlation of weights of a square shaped antenna aperture
+    in the $uv$ plane (left) and the corresponding directional antenna power 
+    response on the sky (right) in coordinates specified by direction cosines. 
+    The antenna auto-correlation weights are normalized to a sum
+    of unity yielding a peak response of unity in the antenna's directional
+    power pattern on the sky. The color scale for the directional power 
+    pattern is logarithmic. The black circle indicates the sky horizon and
+    values beyond it are not physical and hence ignored.}
+  \label{fig:autocorr_wts_PB}
+\end{figure}
+
+We inverse Fourier transform the squared 
+images and beams back to the $uv$ plane and subtract the estimated 
+auto-correlation kernel scaled to the peak value centered at the zero spacing 
+pixel. The final averaged image is obtained by Fourier transforming the $uv$ 
+plane data and weights with the auto-correlations subtracted to the image plane. 
+These images are now comparable to those obtained from visibility-based imaging. 
+This step of removing auto-correlations is required to be performed only once 
+per integration timescale and does not add significant cost to the full operation.
+
 \section{Verification}\label{sec:verify}
@@
 We use the EPIC simulator to generate electric field samples from a sky model. In
-our example, we use 16 frequency channels each of width $\delta f=$100~kHz, 10
-point sources of random flux densities at random locations. The number of
-timestamps in one integration cycle was kept at four where each A/D timeseries is
-$1/\delta f=10\,\mu$s long.
+our simulations, we use 64 frequency channels each of width $\delta f = 40$~kHz, 
+10 point sources of flux densities 10~Jy at random locations. The number of
+timestamps integrated in one integration cycle was kept at eight where each A/D 
+timeseries is $1/\delta f=25\,\mu$s long. We use the MWA array layout 
+\citep{bea12} for demonstration. Only the inner 51 tiles within a square bounding
+box of 150~m on each side were used. We assumed all tiles are identical and have 
+a square shaped electric field illumination footprint 4.4~m on each side.
 
-Show examples using simulations
+Figure~\ref{fig:MOFF-FX-image} shows the dirty images (top) and synthesized beams
+(bottom) obtained with antenna-based MOFF and FX visibility-based imaging 
+algorithms packaged in EPIC. The antenna auto-correlations that correspond to 
+zero spacing have been removed from the MOFF image and the corresponding 
+synthesized beam. The sky positions of the simulated sources are indicated by 
+solid black circles. The reconstructed sky image has the simulated sources at the
+expected sky positions in either case. Both algorithms result in images and 
+synthesized beams that are well matched with each other. Their fluxes are 
+modulated by a multiplicative power pattern corresponding to that of a uniform
+square aperture.
 
-Discuss PSF differences due to slight differences arising out of gridding
+\begin{figure}
+  \includegraphics[width=\columnwidth]{MOFF_FX_image_comparison.eps}
+  \caption{Dirty images (top) and synthesized beams (bottom) obtained from 
+    simulated data using antenna-based MOFF algorithm (left) and FX 
+    visibility-based software holography (right). The solid black circles in the
+    top panels indicate the simulated source positions. The antenna 
+    auto-correlations at zero-spacing have been removed from the MOFF images. 
+    The images in either case reconstruct the sources at the right locations 
+    with the fluxes expected after multiplication by the antenna power pattern. 
+    The synthesized beams from the two algorithms are well matches in size and
+    shape. The overall modulation by the power pattern is seen clearly in both
+    images.}
+  \label{fig:MOFF-FX-image}
+\end{figure}
 
-Apply it on LWA data
+We investigate the two imaging algorithms for differences from the point of 
+view of the quality of their outputs. We begin by comparing the gridded 
+cross-correlation weights in the $uv$-plane. In MOFF imaging, weights 
+from antenna auto-correlations have been removed as described in 
+\S\ref{sec:rm-autocorr}.
+
+Figure~\ref{fig:MOFF-FX-uvwts} shows the cross-correlation weights obtained 
+with MOFF imaging (left) and visibility-based imaging (right). The first 
+notable difference is in the weights around zero spacing. Though both show a
+dominant void around zero-spacing, the void obtained with MOFF algorithm shows 
+many pixels with non-zero weights. In contrast, the zero-spacing void from 
+traditional imaging consists of predominantly zero-valued pixels. The gridding
+process in the former involves rounding the antenna footprint to the nearest 
+grid pixel. Depending on the exact location of the center of the antenna 
+relative to the grid, the grid pixels that receive contribution from an
+antenna may be a pixel narrower along one or both axes relative to that from
+another identical antenna but with a different center location relative to the
+grid. The resulting auto-correlation footprint will also not necessarily be 
+identical. Hence, using a single expected footprint for subtraction will
+typically leave some residuals behind as seen in the void region in the 
+weights obtained with MOFF imaging.
+
+These residuals can be mitigated by: 
+\begin{enumerate}
+  \item making the grid spacing finer which makes the rounding error less 
+    susceptible to the location of the antenna center relative to the grid, 
+    and
+  \item subtracting each auto-correlation of antenna weights separately by 
+    using the shape and extent of the footprint appropriate for that specific
+    antenna aperture. 
+\end{enumerate}
+The latter is the most general solution applicable especially in the case of 
+heterogeneous antenna arrays.
+
+The other notable difference is that weights outside the zero-spacing void in
+some regions are different from each other at the few percent level though the 
+sum of weights in these regions are identical. For instance, note the 
+difference in weights at $(u,v)\approx (2,-10)$. This is found to arise 
+because with MOFF imaging the antenna locations were rounded to the nearest 
+grid pixel whereas in visibility-based imaging the baseline locations 
+(difference between antenna locations) were rounded to the nearest grid pixel. 
+Since rounding is not a commutative operation (i.e. rounding of the difference 
+value is not necessarily equal to the difference of the rounded values), the 
+gridding operation in the two cases introduces rounding errors in the placement 
+of antenna aperture and cross-correlated aperture weights. Each antenna 
+aperture weight projected to the nearest grid location can be displaced by 
+$\sim 1$--2 pixels. This effect can be mitigated only by making the grid 
+spacing finer.
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{MOFF_FX_comparison_uvwts_test_aperture_zero_spacing_removed.png}
+  \caption{Dirty images (top) and synthesized beams (bottom) obtained from 
+    simulated data using antenna-based MOFF algorithm (left) and FX 
+    visibility-based software holography (right). The solid black circles in the
+    top panels indicate the simulated source positions. The antenna 
+    auto-correlations at zero-spacing have been removed from the MOFF images. 
+    The images in either case reconstruct the sources at the right locations 
+    with the fluxes expected after multiplication by the antenna power pattern. 
+    The synthesized beams from the two algorithms are well matches in size and
+    shape. The overall modulation by the power pattern is seen clearly in both
+    images.}
+  \label{fig:MOFF-FX-uvwts}
+\end{figure}
+
+We study the effect of the differences in gridded weights on the image plane. 
+The synthesized beam can be written as:
+\begin{align}\label{eqn:syn-beam}
+  B(\hat{\mathbf{s}}) &= \left| \sum_\mathbf{r} \widetilde{W}(\mathbf{r})\,e^{-i\frac{2\pi}{\lambda}(\mathbf{r}\cdot\hat{\mathbf{s}})}\right|^2, \\
+  \textrm{or}, \quad B(l,m) &= \left| \sum_{r_x,r_y} \widetilde{W}(r_x,r_y)\,e^{-i\frac{2\pi}{\lambda}(r_x l + r_y m)}\right|^2 
+\end{align}
+where, $\hat{\mathbf{s}}\equiv (l,m)$, $\mathbf{r}\equiv (r_x,r_y)$ and 
+$\widetilde{W}(\mathbf{r})$ is given by equation~\ref{eqn:gridding-convolution}.
+If the errors in the determination of $r_x$ and $r_y$ are denoted by 
+$\delta r_x$ and $\delta r_y$ respectively, the error on $B(l,m)$ can be 
+estimated using a simple propagation of errors, where,
+\begin{align}
+  \frac{\partial{B}(l,m)}{\partial{r_x}} &= e^{-i\frac{2\pi}{\lambda}\mathbf{r}\cdot\hat{\mathbf{s}}}\left[\frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_x}} - i\frac{2\pi l}{\lambda}\widetilde{W}(\mathbf{r})\right] \label{eqn:syn-beam-pdev1}\\
+  &\qquad\qquad \times \sum_{\mathbf{r}^\prime} \widetilde{W}^\star(\mathbf{r}^\prime)\,e^{+i\frac{2\pi}{\lambda}\mathbf{r}^\prime\cdot \hat{\mathbf{s}}} \\
+  &= e^{-i\frac{2\pi}{\lambda}\mathbf{r}\cdot\hat{\mathbf{s}}}\left[\frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_x}} - i\frac{2\pi l}{\lambda}\widetilde{W}(\mathbf{r})\right]\,W^\star(\hat{\mathbf{s}}), \\
+  \textrm{and},\frac{\partial{B}(l,m)}{\partial{r_y}} &= e^{-i\frac{2\pi}{\lambda}\mathbf{r}\cdot\hat{\mathbf{s}}}\left[\frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_x}} - i\frac{2\pi m}{\lambda}\widetilde{W}(\mathbf{r})\right]\,W^\star(\hat{\mathbf{s}}).
+\label{eqn:syn-beam-pdev2}
+\end{align}
+It must be noted that although we start with a squaring operation in 
+equation~\ref{eqn:syn-beam}, in equations~\ref{eqn:syn-beam-pdev1} through 
+\ref{eqn:syn-beam-pdev2}, the summations are performed only when 
+$\mathbf{r}\ne \mathbf{r}^\prime$ and therefore do not include the 
+auto-correlations of the antenna weights. The error expected in $B(l,m)$ is:
+\begin{align}
+  \left|\delta B(l,m)\right|^2 &= \sum_{r_x}\left|\frac{\partial{B}(l,m)}{\partial{r_x}}\right|^2 \delta r_x^2 + \sum_{r_y}\left|\frac{\partial{B}(l,m)}{\partial{r_y}}\right|^2 \delta r_y^2 \\
+  &= \delta r_x^2 \sum_{r_x}\left|\frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_x}} - i\frac{2\pi l}{\lambda}\widetilde{W}(\mathbf{r})\right|^2\,\left|W(\hat{\mathbf{s}})\right|^2 \nonumber\\
+  &\qquad + \delta r_y^2 \sum_{r_y}\left|\frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_y}} - i\frac{2\pi m}{\lambda}\widetilde{W}(\mathbf{r})\right|^2\,\left|W(\hat{\mathbf{s}})\right|^2
+\end{align}
+For the current simulations, 
+\begin{align}
+  \delta r_x = \delta r_y &\simeq 1, \\
+  \widetilde{W}(r_x,r_y) &= \{0, \textrm{constant}\}, \\
+  \textrm{and},\qquad \frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_x}} &= \frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_y}} = 0.
+\end{align}
+Hence,
+\begin{align}
+  \left|\delta B(l,m)\right|^2 &\simeq \left(\frac{2\pi}{\lambda}\right)^2\left(l^2+m^2\right)\,B(\hat{\mathbf{s}})\,\sum_\mathbf{r}\,\left|\widetilde{W}(\mathbf{r})\right|^2, \\
+\textrm{or},\qquad \left|\delta B(l,m)\right| &\simeq \left(\frac{2\pi}{\lambda}\right)\sqrt{l^2+m^2}\,\left|W(\hat{\mathbf{s}})\right|\,\left(\sum_\mathbf{r}\,\left|\widetilde{W}(\mathbf{r})\right|^2\right)^{1/2}, \label{eqn:syn-beam-error}
+\end{align}
+where, $B(\hat{\mathbf{s}}) \equiv \left|W(\hat{\mathbf{s}})\right|^2$. Thus, 
+the error on the synthesized beam is proportional to the antenna power pattern
+and increases linearly outward in radial bins of direction cosines.
+
+Figure~\ref{fig:psf-diff} shows the difference between the synthesized beams 
+obtained with the two methods. A difference map between the two synthesized 
+beams is shown on top. The amplitude of the difference appears to be 
+modulated by the directional voltage response of the antenna. At the bottom, 
+in radial bins, the rms of the synthesized beam (gray) and the rms of the 
+difference map (black) are plotted in percentage units relative to the peak 
+(to be read using the axis on the left side of the plot). The antenna power 
+pattern (red; to be read using the scale on the right) is plotted for reference.
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{diff_psf_MOFF-FX_test_aperture.png}
+  \caption{Map of difference between the synthesized beams obtained with the 
+    two methods (top) and radial statistics of the synthesized beams and 
+    their differences (bottom). The maximum difference is of the order of a 
+    few percent. The difference appears to be modulated in amplitude by the 
+    power pattern of the antenna. }
+  \label{fig:psf-diff}
+\end{figure}
+
+The synthesized beam rms is proportional to the antenna power pattern as 
+expected from a point spread function of uncorrected for the antenna power 
+pattern. A notable observation is that the rms of the difference of
+the synthesized beams seems proportional to the directional antenna voltage
+pattern and also increases gradually with distance from the center as seen 
+from the diverging amplitudes between the black and red curves as predicted
+in equation~\ref{eqn:syn-beam-error}. The rms of the difference appears to 
+be bounded by the actual rms of the synthesized beam indicating that these
+rounding errors in gridding do not affect the statistics of the images or
+the synthesized beams.
 

### 2015-10-23 imaging_demo/ms.tex
Remove usage of lambda in section 4
From a723a20 (master):
=====================

In ms.tex,

(*) Remove usage of lambda in section 4
Create subsections under section 4
From 9414f28 (master):
=====================

In ms.tex,

(*) Create subsections under section 4
fixed some typos and other small things.

@@
 now publicly available. Besides implementing the MOFF imaging algorithm it also
-includes FX imaging using software holography technique and a simulator for
+includes FX imaging using the software holography technique and a simulator for
 generating electric fields from a sky model.
@@
 electric fields are gridded using an antenna-based gridding convolution function
-following which it is spatially Fourier transformed and squared to obtain images
+after which it is spatially Fourier transformed and squared to obtain images
 for every time stamp. These images are then time-averaged to obtain the
@@
 imaging from a FX correlator. The antenna-based F-engine is identical to that in
-the MOFF processing. The electric field spectra each antenna are then cross-
+the MOFF processing. The electric field spectra from each antenna are then cross-
 multiplied in the X-engine with those from all other antennas to obtain the
@@
 illumination function, $\widetilde{B}(\mathbf{r}-\mathbf{r}_a)$, which can
-specified either by a functional form or as a table of values against locations
+be specified either by a functional form or as a table of values against locations
 around the antennas. A nearest neighbor mapping from all antenna footprints to
@@
 stationary antenna array in the absence of any time-dependent effects, this
-mapping has to be determined only once in the antenna array coordinate frame. The
+mapping must only be determined once in the antenna array coordinate frame. The
 antenna-to-grid mapping matrix, $\mathbf{M}(\mathbf{r},\mathbf{e})$ is described
@@
 antenna occupies a footprint typically the size of its aperture,
-$\mathbf{M}(\mathbf{r},\mathbf{e})$ which is generally of size
+$\mathbf{M}(\mathbf{r},\mathbf{e})$, which is generally of size
 $N_\textrm{grid}\times N_\textrm{a}$, reduces to a sparse block-diagonal matrix
@@
 \subsection{Calibration}
-Calibration of direct imaging correlator remains an unresolved matter. Contrary
+Calibration of direct imaging correlators remains an unresolved matter. Contrary
 to the FX data flow, direct imagers mix the signals from all antennas before
@@
 before the gridding step. Previous efforts have resorted to applying FX-generated
-calibration solutions \citep{zhe14, fos14}, or integrate a dedicated FX correlator 
+calibration solutions \citep{zhe14, fos14}, or integrating a dedicated FX correlator 
 which periodically forms the full visibility matrix \citep{wij09, dev09}.
@@
 \end{align}
-where, $E(\mathbf{e})$ denotes the spectra of measured antenna electric fields.
-
-In practice the grid is set to be a power of 2 such that the grid spacing is at
-most $\lambda/2$ even at the highest frequency in the observing band.
+where $E(\mathbf{e})$ denotes the spectra of measured antenna electric fields.
+% The sentence below was already stated earlier.
+%In practice the grid is set to be a power of 2 such that the grid spacing is at
+%most $\lambda/2$ even at the highest frequency in the observing band.
 
@@
 
+\subsection{Simulations}\label{sec:sim}
+
 We use the EPIC simulator to generate electric field samples from a sky model. In
@@
 \end{figure}
+
+\subsection{Differences in outputs}\label{sec:diff}
 
@@
 \begin{align}\label{eqn:syn-beam}
-  B(\hat{\mathbf{s}}) &= \left| \sum_\mathbf{r} \widetilde{W}(\mathbf{r})\,e^{-i\frac{2\pi}{\lambda}(\mathbf{r}\cdot\hat{\mathbf{s}})}\right|^2, \\
-  \textrm{or}, \quad B(l,m) &= \left| \sum_{r_x,r_y} \widetilde{W}(r_x,r_y)\,e^{-i\frac{2\pi}{\lambda}(r_x l + r_y m)}\right|^2 
+  B(\hat{\mathbf{s}}) &= \left| \sum_\mathbf{r} \widetilde{W}(\mathbf{r})\,e^{-i2\pi\mathbf{r}\cdot\hat{\mathbf{s}}}\right|^2, \\
+  \textrm{or}, \quad B(l,m) &= \left| \sum_{r_x,r_y} \widetilde{W}(r_x,r_y)\,e^{-i2\pi(r_x l + r_y m)}\right|^2 
 \end{align}
@@
 \begin{align}
-  \frac{\partial{B}(l,m)}{\partial{r_x}} &= e^{-i\frac{2\pi}{\lambda}\mathbf{r}\cdot\hat{\mathbf{s}}}\left[\frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_x}} - i\frac{2\pi l}{\lambda}\widetilde{W}(\mathbf{r})\right] \label{eqn:syn-beam-pdev1}\\
-  &\qquad\qquad \times \sum_{\mathbf{r}^\prime} \widetilde{W}^\star(\mathbf{r}^\prime)\,e^{+i\frac{2\pi}{\lambda}\mathbf{r}^\prime\cdot \hat{\mathbf{s}}} \\
-  &= e^{-i\frac{2\pi}{\lambda}\mathbf{r}\cdot\hat{\mathbf{s}}}\left[\frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_x}} - i\frac{2\pi l}{\lambda}\widetilde{W}(\mathbf{r})\right]\,W^\star(\hat{\mathbf{s}}), \\
-  \textrm{and},\frac{\partial{B}(l,m)}{\partial{r_y}} &= e^{-i\frac{2\pi}{\lambda}\mathbf{r}\cdot\hat{\mathbf{s}}}\left[\frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_x}} - i\frac{2\pi m}{\lambda}\widetilde{W}(\mathbf{r})\right]\,W^\star(\hat{\mathbf{s}}).
+  \frac{\partial{B}(l,m)}{\partial{r_x}} &= e^{-i2\pi\mathbf{r}\cdot\hat{\mathbf{s}}}\left[\frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_x}} - i 2\pi l\widetilde{W}(\mathbf{r})\right] \label{eqn:syn-beam-pdev1}\\
+  &\qquad\qquad \times \sum_{\mathbf{r}^\prime} \widetilde{W}^\star(\mathbf{r}^\prime)\,e^{+i2\pi\mathbf{r}^\prime\cdot \hat{\mathbf{s}}} \\
+  &= e^{-i2\pi\mathbf{r}\cdot\hat{\mathbf{s}}}\left[\frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_x}} - i2\pi l\widetilde{W}(\mathbf{r})\right]\,W^\star(\hat{\mathbf{s}}), \\
+  \textrm{and},\frac{\partial{B}(l,m)}{\partial{r_y}} &= e^{-i2\pi\mathbf{r}\cdot\hat{\mathbf{s}}}\left[\frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_x}} - i2\pi m\widetilde{W}(\mathbf{r})\right]\,W^\star(\hat{\mathbf{s}}).
 \label{eqn:syn-beam-pdev2}
@@
   \left|\delta B(l,m)\right|^2 &= \sum_{r_x}\left|\frac{\partial{B}(l,m)}{\partial{r_x}}\right|^2 \delta r_x^2 + \sum_{r_y}\left|\frac{\partial{B}(l,m)}{\partial{r_y}}\right|^2 \delta r_y^2 \\
-  &= \delta r_x^2 \sum_{r_x}\left|\frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_x}} - i\frac{2\pi l}{\lambda}\widetilde{W}(\mathbf{r})\right|^2\,\left|W(\hat{\mathbf{s}})\right|^2 \nonumber\\
-  &\qquad + \delta r_y^2 \sum_{r_y}\left|\frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_y}} - i\frac{2\pi m}{\lambda}\widetilde{W}(\mathbf{r})\right|^2\,\left|W(\hat{\mathbf{s}})\right|^2
+  &= \delta r_x^2 \sum_{r_x}\left|\frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_x}} - i2\pi l\widetilde{W}(\mathbf{r})\right|^2\,\left|W(\hat{\mathbf{s}})\right|^2 \nonumber\\
+  &\qquad + \delta r_y^2 \sum_{r_y}\left|\frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_y}} - i2\pi m\widetilde{W}(\mathbf{r})\right|^2\,\left|W(\hat{\mathbf{s}})\right|^2
 \end{align}
@@
 \begin{align}
-  \left|\delta B(l,m)\right|^2 &\simeq \left(\frac{2\pi}{\lambda}\right)^2\left(l^2+m^2\right)\,B(\hat{\mathbf{s}})\,\sum_\mathbf{r}\,\left|\widetilde{W}(\mathbf{r})\right|^2, \\
-\textrm{or},\qquad \left|\delta B(l,m)\right| &\simeq \left(\frac{2\pi}{\lambda}\right)\sqrt{l^2+m^2}\,\left|W(\hat{\mathbf{s}})\right|\,\left(\sum_\mathbf{r}\,\left|\widetilde{W}(\mathbf{r})\right|^2\right)^{1/2}, \label{eqn:syn-beam-error}
+  \left|\delta B(l,m)\right|^2 &\simeq 4\pi^2\left(l^2+m^2\right)\,B(\hat{\mathbf{s}})\,\sum_\mathbf{r}\,\left|\widetilde{W}(\mathbf{r})\right|^2, \\
+\textrm{or},\qquad \left|\delta B(l,m)\right| &\simeq 2\pi\sqrt{l^2+m^2}\,\left|W(\hat{\mathbf{s}})\right|\,\left(\sum_\mathbf{r}\,\left|\widetilde{W}(\mathbf{r})\right|^2\right)^{1/2}, \label{eqn:syn-beam-error}
 \end{align}
@@
 
+\subsection{Application to LWA data}\label{sec:LWA-data}
+
 \section{Analysis and Feasibility}\label{sec:analysis}
@@
   \includegraphics[width=\columnwidth]{MOFF_FX_computations_fov_gridding_annotated.eps}
-  \caption{Current and instruments planned for future in parameter space of
+  \caption{Current and planned instruments in parameter space of
     number of complex multiplies and adds with MOFF and FX.}

### 2015-10-23 imaging_demo/ms.tex
Add content under LWA data processing
From 3c587ae (master):
=====================

In ms.tex,

(*) Add content under LWA data processing

@@
 now publicly available. Besides implementing the MOFF imaging algorithm it also
-includes FX imaging using the software holography technique and a simulator for
+includes FX imaging using software holography technique and a simulator for
 generating electric fields from a sky model.
@@
 electric fields are gridded using an antenna-based gridding convolution function
-after which it is spatially Fourier transformed and squared to obtain images
+following which it is spatially Fourier transformed and squared to obtain images
 for every time stamp. These images are then time-averaged to obtain the
@@
 imaging from a FX correlator. The antenna-based F-engine is identical to that in
-the MOFF processing. The electric field spectra from each antenna are then cross-
+the MOFF processing. The electric field spectra each antenna are then cross-
 multiplied in the X-engine with those from all other antennas to obtain the
@@
 illumination function, $\widetilde{B}(\mathbf{r}-\mathbf{r}_a)$, which can
-be specified either by a functional form or as a table of values against locations
+specified either by a functional form or as a table of values against locations
 around the antennas. A nearest neighbor mapping from all antenna footprints to
@@
 stationary antenna array in the absence of any time-dependent effects, this
-mapping must only be determined once in the antenna array coordinate frame. The
+mapping has to be determined only once in the antenna array coordinate frame. The
 antenna-to-grid mapping matrix, $\mathbf{M}(\mathbf{r},\mathbf{e})$ is described
@@
 antenna occupies a footprint typically the size of its aperture,
-$\mathbf{M}(\mathbf{r},\mathbf{e})$, which is generally of size
+$\mathbf{M}(\mathbf{r},\mathbf{e})$ which is generally of size
 $N_\textrm{grid}\times N_\textrm{a}$, reduces to a sparse block-diagonal matrix
@@
 \subsection{Calibration}
-Calibration of direct imaging correlators remains an unresolved matter. Contrary
+Calibration of direct imaging correlator remains an unresolved matter. Contrary
 to the FX data flow, direct imagers mix the signals from all antennas before
@@
 before the gridding step. Previous efforts have resorted to applying FX-generated
-calibration solutions \citep{zhe14, fos14}, or integrating a dedicated FX correlator 
+calibration solutions \citep{zhe14, fos14}, or integrate a dedicated FX correlator 
 which periodically forms the full visibility matrix \citep{wij09, dev09}.
@@
 \end{align}
-where $E(\mathbf{e})$ denotes the spectra of measured antenna electric fields.
-% The sentence below was already stated earlier.
-%In practice the grid is set to be a power of 2 such that the grid spacing is at
-%most $\lambda/2$ even at the highest frequency in the observing band.
+where, $E(\mathbf{e})$ denotes the spectra of measured antenna electric fields.
+
+In practice the grid is set to be a power of 2 such that the grid spacing is at
+most $\lambda/2$ even at the highest frequency in the observing band.
 
@@
 
+We have analyzed narrow band data from the LWA in New Mexico. This data is  
+in LWA TBN format with 512 voltage time samples from 255 antennas within 
+roughly a diameter of 100~m. The data is centered at a frequency of 74.03~MHz,
+with a sample rate (equal to the bandwidth) of 100~kHz with 512 complex time 
+samples in a A/D writeout timescale of 5.12~ms, a frequency resolution of 
+195.3125~Hz and dual polarization. There are 391 such timestamps yielding a
+total duration of 2~s.
+
+We have matched the cable delays of all the antennas but yet to deploy the 
+calibration pipeline we have developed on this data. Therefore, in 
+subsequent discussion we assumed the data is calibrated.
+
+Figure~\ref{fig:LWA-image} shows the image produced with MOFF imaging 
+packaged in EPIC after averaging over the entire 2~s of data and the inner
+$\approx 80$\% of bandwidth (roughly 80~kHz). The image is shown in direction 
+cosine coordinates -- $l$ along x-axis and $m$ along y-axis. Due to lack of
+calibration the flux scale is arbitrary. Despite the lack of calibration,
+only with delay compensation, we see Cyg~A and Cas~A prominently at
+$(l,m)\approx (0, 0.125)$, and (0.4,0.55) respectively and Sgr~A faintly at
+(-0.5,-0.75), thus validating the functionality of the EPIC package.
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{LWA_sample_image_391_iterations.png}
+  \caption{Image from LWA TBN data obtained with MOFF imaging using EPIC 
+    package after averaging over 2~s and $\approx 80$~kHz. The x- and y-axes
+    denote direction cosines $l$ and $m$ respectively. The antenna voltages
+    are compensated for their respective delays. The flux scale is arbitrary
+    and not calibrated. Cyg~A is prominently visible near the center at
+    $(l,m)\approx (0, 0.125)$, Cas~A near (0.4,0.55) and Sgr~A near 
+    (-0.5,-0.75).}
+  \label{fig:LWA-image}
+\end{figure}
+
 \section{Analysis and Feasibility}\label{sec:analysis}
@@
   \includegraphics[width=\columnwidth]{MOFF_FX_computations_fov_gridding_annotated.eps}
-  \caption{Current and planned instruments in parameter space of
+  \caption{Current and instruments planned for future in parameter space of
     number of complex multiplies and adds with MOFF and FX.}

### 2015-10-23 imaging_demo/ms.tex
cleaning up some math notation

--- 2015-10-23 imaging_demo/ms.tex
@@
 
-Interferometers measure {\it visibilities} -- the degree of coherence -- between
+Interferometers measure {\it visibilities} -- the degree of coherence between
 electric fields measured by a pair of antennas \citep{van34,zer38,tho01}.
@@
 
-Following the matrix
-notation of \citet{mor11}, the above measurement equation can be expressed as:
+From here we adopt the matrix notation of \citet{mor11}, where vectors are 
+represented with single coordinates, and matrices are represented by two 
+coordinates denoting the spaces the operator transforms between. 
+In this notation, the above measurement equation can be expressed as:
 \begin{align}
-  \mathbf{m}(\mathbf{v}) &= \widetilde{B}(\mathbf{v},\mathbf{u})\,\mathbf{F}(\mathbf{u},\hat{\mathbf{s}})\,I(\hat{\mathbf{s}}) + \mathbf{n}(\mathbf{v}),
+  \mathbf{m}(\mathbf{v}) &= \widetilde{\mathbf{B}}(\mathbf{v},\mathbf{u})\,\mathbf{F}(\mathbf{u},\hat{\mathbf{s}})\,\mathbf{I}(\hat{\mathbf{s}}) + \mathbf{n}(\mathbf{v}),
 \end{align}
-where the sky brightness $I(\hat{\mathbf{s}})$ is Fourier transformed using
+where the sky brightness $\mathbf{I}(\hat{\mathbf{s}})$ is Fourier transformed using
 $\mathbf{F}(\mathbf{u},\hat{\mathbf{s}})$ and the resultant spatial coherence
 function is weighted and summed using the antenna power response,
-$\widetilde{B}(\mathbf{v},\mathbf{u})$ in $uv$-space sampled at the baseline
+$\widetilde{\mathbf{B}}(\mathbf{v},\mathbf{u})$ in $uv$-space sampled at the baseline
 location to obtain the measured visibilities:
 \begin{align}
-  \mathbf{m}(\mathbf{v}) &= \left\langle \widetilde{E}^\star(\mathbf{a})\widetilde{E}(\mathbf{a}^\prime)\right\rangle_t, \label{eqn:matrix-cc-vis}
+  \mathbf{m}(\mathbf{v}) &= \left\langle \widetilde{\mathbf{E}}^\star(\mathbf{a})\,\widetilde{\mathbf{E}}(\mathbf{a}^\prime)\right\rangle_t, \label{eqn:matrix-cc-vis}
 \end{align}
@@
 \begin{align}
-  I^\prime(\hat{\mathbf{s}}) &= \mathbf{F}^\textrm{T}(\hat{\mathbf{s}},\mathbf{u})\,\widetilde{B}^{\,\textrm{T}}(\mathbf{u},\mathbf{v})\,\mathbf{N}^{-1}(\mathbf{v},\mathbf{v})\,\mathbf{m}(\mathbf{v}) \label{eqn:dirty-image-FX}
+  \mathbf{I}^\prime(\hat{\mathbf{s}}) &= \mathbf{F}^\textrm{T}(\hat{\mathbf{s}},\mathbf{u})\,\widetilde{\mathbf{B}}^{\,\textrm{T}}(\mathbf{u},\mathbf{v})\,\mathbf{N}^{-1}(\mathbf{v},\mathbf{v})\,\mathbf{m}(\mathbf{v}) \label{eqn:dirty-image-FX}
 \end{align}
@@
 the gridding kernel, followed by a Fourier transform to create an image
-$I^\prime(\hat{\mathbf{s}})$. This is the optimal estimate of the true image 
-$I(\hat{\mathbf{s}})$ given the visibility measurements.
+$\mathbf{I}^\prime(\hat{\mathbf{s}})$. This is the optimal estimate of the true image 
+$\mathbf{I}(\hat{\mathbf{s}})$ given the visibility measurements.
 
@@
 \begin{align}
-&\widetilde{B}^{\,\textrm{T}}(\mathbf{u},\mathbf{v})\,\mathbf{N}^{-1}(\mathbf{v} ,\mathbf{v})\,\mathbf{m}(\mathbf{v})=\nonumber\\
+&\widetilde{\mathbf{B}}^{\,\textrm{T}}(\mathbf{u},\mathbf{v})\,\mathbf{N}^{-1}(\mathbf{v} ,\mathbf{v})\,\mathbf{m}(\mathbf{v})=\nonumber\\
 &\qquad\left\langle \left[\widetilde{\mathbf{W}}^\textrm{T}_a(\mathbf{r},\mathbf{a})\, \widetilde{\mathbf{N}}^{-1}\!(\mathbf{a},\mathbf{a})\, \widetilde{\mathbf{E}}(\mathbf{a})\right]
@@
 \begin{align}
-  I^\prime(\hat{\mathbf{s}}) &= \left\langle \left|\,\mathbf{F}^\textrm{T}(\hat{\mathbf{s}},\mathbf{r})\,\widetilde{\mathbf{W}}^\textrm{T}(\mathbf{r},\mathbf{a})\,\widetilde{\mathbf{N}}^{-1}(\mathbf{a},\mathbf{a})\,\widetilde{E}(\mathbf{a})\,\right|^2\right\rangle_t.
+  \mathbf{I}^\prime(\hat{\mathbf{s}}) &= \left\langle \left|\,\mathbf{F}^\textrm{T}(\hat{\mathbf{s}},\mathbf{r})\,\widetilde{\mathbf{W}}^\textrm{T}(\mathbf{r},\mathbf{a})\,\widetilde{\mathbf{N}}^{-1}(\mathbf{a},\mathbf{a})\,\widetilde{\mathbf{E}}(\mathbf{a})\,\right|^2\right\rangle_t.
 \label{eqn:dirty-image-MOFF}
@@
 temporal Fourier transform on the electric field time-series to obtain electric
-field spectra $E_1(f)\ldots E_\textrm{N}(f)$ ($\widetilde{E}_a$ in matrix
+field spectra $E_1(f)\ldots E_\textrm{N}(f)$ ($\widetilde{\mathbf{E}}(\mathbf{a})$ in matrix
 notation) for each of the antennas. Each of the complex antenna gains are
@@
 for every time stamp. These images are then time-averaged to obtain the
-accumulated image $I(f)$ ($I(\hat{\mathbf{s}})$ in matrix notation).
+accumulated image $I(f)$ ($\mathbf{I}(\hat{\mathbf{s}})$ in matrix notation).
 \begin{figure}
@@
 mapping must only be determined once in the antenna array coordinate frame. The
-antenna-to-grid mapping matrix, $\mathbf{M}(\mathbf{r},\mathbf{e})$ is described
-as a transformation matrix from the space of measured electric fields
-($\mathbf{e}$) to the antenna array grid denoted by $\mathbf{r}$. Since each
+antenna-to-grid mapping matrix, $\mathbf{M}(\mathbf{r},\mathbf{a})$ is described
+as a transformation matrix from the space of measured electric fields by the antennas
+($\mathbf{a}$) to the antenna array grid denoted by the coordinate $\mathbf{r}$. Since each
 antenna occupies a footprint typically the size of its aperture,
-$\mathbf{M}(\mathbf{r},\mathbf{e})$, which is generally of size
+$\mathbf{M}(\mathbf{r},\mathbf{a})$, which is generally of size
 $N_\textrm{grid}\times N_\textrm{a}$, reduces to a sparse block-diagonal matrix
@@
 The antenna array aperture illumination over the entire grid,
-$\widetilde{W}(\mathbf{r})$, is obtained by a projection of the individual
+$\widetilde{\mathbf{W}}(\mathbf{r})$, is obtained by a projection of the individual
 antenna aperture illuminations:
 \begin{align}\label{eqn:gridding-convolution}
-  \widetilde{W}(\mathbf{r}) &= \sum_a \widetilde{W}_a(\mathbf{r}-\mathbf{r}_a) \\
-                            &= \mathbf{M}(\mathbf{r},\mathbf{e})\,\mathcal{I}(\mathbf{e}),
+  \widetilde{\mathbf{W}}(\mathbf{r}) &= \sum_a \widetilde{\mathbf{W}}_a(\mathbf{r}-\mathbf{r}_a) \\
+                            &= \mathbf{M}(\mathbf{r},\mathbf{a})\,\mathcal{I}(\mathbf{a}),
 \end{align}
-where, $\mathcal{I}(\mathbf{e})$ is a row of ones. This is achieved using
+where, $\mathcal{I}(\mathbf{a})$ is a row of ones. This is achieved using
 efficient multiplication with the sparse matrix created in the antenna-to-grid
-mapping process. Unless $\widetilde{W}(\mathbf{r})$ includes time-dependent
+mapping process. Unless $\widetilde{\mathbf{W}}(\mathbf{r})$ includes time-dependent
 effects of the ionosphere or the instrument, it needs to be computed just once
 for the entire observation. However, the gridding of electric fields must be
-computed at every readout of the electric field spectra. Thus,
+computed at every readout of the electric field spectra.
 \begin{align}
-  \widetilde{E}(\mathbf{r}) &= \mathbf{M}(\mathbf{r},\mathbf{e})\,E(\mathbf{e}),
+  \widetilde{\mathbf{E}}(\mathbf{r}) &= \mathbf{M}(\mathbf{r},\mathbf{a})\,\widetilde{\mathbf{E}}(\mathbf{a}),
 \end{align}
-where $E(\mathbf{e})$ denotes the spectra of measured antenna electric fields.
 % The sentence below was already stated earlier.

### 2015-10-24 imaging_demo/ms.tex
Expand bibtex file and content under section 5
From 63a05b7 (master):
=====================

In epic.bib,

(*) Add references Cornwell et al.(2008),
Thyagarajan et al.(2015a), Thyagarajan et al.(2015b)

In ms.tex,

(*) Add content under section 5

@@
 averaging. This expression can be further simplified with the sky brightness, 
-$I(\hat{\mathbf{s}}) \delta(\hat{\mathbf{s}}-\hat{\mathbf{s'}})= \left\langle 
-E(\hat{\mathbf{s}}) E^\star(\hat{\mathbf{s'}}) \right\rangle_t $, and defining the 
+$I(\hat{\mathbf{s}})= \left\langle E(\hat{\mathbf{s}}) E^\star(\hat{\mathbf{s'}}) 
+\right\rangle_t  \delta(\hat{\mathbf{s}}-\hat{\mathbf{s'}})$, and defining the 
 antenna pair sky power response function (or the primary beam), 
@@
 for the entire observation. However, the gridding of electric fields must be
-computed at every readout of the electric field spectra.
-\begin{align}
-  \widetilde{\mathbf{E}}(\mathbf{r}) &= \mathbf{M}(\mathbf{r},\mathbf{a})\,\widetilde{\mathbf{E}}(\mathbf{a}),
+computed at every readout of the electric field spectra,
+\begin{align}
+  \widetilde{\mathbf{E}}(\mathbf{r}) &= \mathbf{M}(\mathbf{r},\mathbf{a})\,\widetilde{\mathbf{E}}(\mathbf{a}).
 \end{align}
@@
 \section{Analysis and Feasibility}\label{sec:analysis}
+
+We now investigate the feasibility of implementing the EPIC imager on current
+and future radio telescopes.
+
+We have profiled the core routines of EPIC line-by-line for various ranges of
+parameters such as antenna filling fraction, maximum baseline 
+length, bandwidth and frequency resolution, integration timescale, etc. for 
+HERA antenna layouts which are highly compact.
+
+Of all steps in the MOFF pipeline that are repeated for every writeout from 
+the F-engine, the slowest step even for dense HERA 
+layouts is found to be the spatial two-dimensional FFT in the imaging stage 
+relative to applying the sparse matrix gridding convolution, squaring or 
+time-averaging. For instance, even in the dense array layout scenario that 
+makes these stages perform the slowest, the gridding convolution, squaring and 
+time-averaging take up only $\simeq 40$\%, $\simeq 40$\% and $\simeq 15$\% 
+respectively of the time taken by the spatial Fourier transform. With sparser 
+arrays they will be even faster.
+
+In the visibility-based imaging, the predominant computational cost is at
+the X-engine requiring $N_\textrm{a}^2$ complex multiplications per channel 
+per correlator writeout timescale.
+
+In the following discussions, we will assume that the computational cost for
+the MOFF imaging is determined by the spatial Fourier transform while  
+that for visibility-based imaging comes from $N_\textrm{a}^2$ 
+cross-correlations. However, if non-linearities such 
+as non-coplanarity of baselines \citep{cor08} and wide-field phenomena like 
+the {\it pitchfork} effect \citep{thy15a,thy15b} are to be corrected for, 
+the antenna based illumination footprint can start becoming less compact
+in the measurement plane and can result in a costlier gridding process.
+
+The number of complex multiplications and additions in the spatial Fourier 
+transform implemented via Fast Fourier Transform \citep[FFT;][]{coo65} is 
+$\approx \beta N_\textrm{g}\log N_\textrm{g}$ where $N_\textrm{g}$ 
+is the number of pixels on the grid and $\beta$ is a constant that depends on 
+the implementation of twiddle FFT algorithms \citep{bri74}. In our study, we 
+set $\beta=5$, a value much more conservative than was indicated in 
+\citet{mor11}. We set the number of complex multiplications in the X-engine 
+in visibility based imaging to $N_\textrm{a}^2$.
+
+We consider antenna array layouts described by three quantities fundamental
+to radio interferometry, namely, maximum baseline length, number of antennas,
+and the size of each antenna.
+
+% Figure~\ref{fig:parameter-space-bll-nant-instruments} shows the
 

### 2015-10-24 imaging_demo/ms.tex
Add more content under section 5
From 5a9dd6e (master):
=====================

In ms.tex,

(*) Add more content under section 5

@@
 \usepackage{amssymb}	% Extra maths symbols
+\usepackage[flushleft]{threeparttable} % For table notes
 
@@
 
-Figure~\ref{fig:MOFF-flowchart} shows the flowchart for MOFF imaging. The
+Fig.~\ref{fig:MOFF-flowchart} shows the flowchart for MOFF imaging. The
 propagated electric fields are shown on the left at different time stamps,
@@
 
-Figure~\ref{fig:FX-flowchart} shows the flowchart for software holographic
+Fig.~\ref{fig:FX-flowchart} shows the flowchart for software holographic
 imaging from a FX correlator. The antenna-based F-engine is identical to that in
@@
 Fourier transformed to generate electric field spectra. This step can be
-parallelized by antennas as shown in Figure~\ref{fig:f-engine}. The output is
+parallelized by antennas as shown in Fig.~\ref{fig:f-engine}. The output is
 then fed to either MOFF and FX imaging pipelines.
@@
 with only $N_\textrm{a}$ blocks and roughly $N_\textrm{k}$ non-zero entries per
-block. Figure~\ref{fig:a2g-mapping} illustrates the antenna-to-grid mapping
+block. Fig.~\ref{fig:a2g-mapping} illustrates the antenna-to-grid mapping
 matrix and the grid containing the mapped aperture footprints of the antennas.
@@
 electric field signals with an image pixel form the output of the correlator in the 
-feedback calibration fashion outlined in \citealt{mor11} (illustrated in Figure 
+feedback calibration fashion outlined in \citealt{mor11} (illustrated in Fig. 
 \ref{fig:MOFF-flowchart} by the arrow leading from the imager to the calibration 
@@
 antenna aperture illumination pattern which is a quantity already available 
-from the gridding step. Figure~\ref{fig:autocorr_wts_PB} shows the estimated 
+from the gridding step. Fig.~\ref{fig:autocorr_wts_PB} shows the estimated 
 weights from antenna auto-correlations in the $uv$-plane (left) and the 
@@
 
-Figure~\ref{fig:MOFF-FX-image} shows the dirty images (top) and synthesized beams
+Fig.~\ref{fig:MOFF-FX-image} shows the dirty images (top) and synthesized beams
 (bottom) obtained with antenna-based MOFF and FX visibility-based imaging 
@@
 
-Figure~\ref{fig:MOFF-FX-uvwts} shows the cross-correlation weights obtained 
+Fig.~\ref{fig:MOFF-FX-uvwts} shows the cross-correlation weights obtained 
 with MOFF imaging (left) and visibility-based imaging (right). The first 
@@
 
-Figure~\ref{fig:psf-diff} shows the difference between the synthesized beams 
+Fig.~\ref{fig:psf-diff} shows the difference between the synthesized beams 
 obtained with the two methods. A difference map between the two synthesized 
@@
 
-Figure~\ref{fig:LWA-image} shows the image produced with MOFF imaging 
+Fig.~\ref{fig:LWA-image} shows the image produced with MOFF imaging 
 packaged in EPIC after averaging over the entire 2~s of data and the inner
@@
 
-We consider antenna array layouts described by three quantities fundamental
+We consider a variety of current and planned radio telescopes. Their antenna 
+layouts are summarized in Table~\ref{tab:antenna-layouts}. The size of the 
+layout gives the maximum baseline $b_\textrm{max}$. The grid spacing is 
+determined by the science goals of the experiment in general. For our 
+purpose, we assume a typical requirement that only the field of view of the 
+antenna is to be imaged. This sets the grid spacing to be equal to the size 
+of the antenna, $A_\textrm{a}$. Hence, 
+$N_\textrm{g}\simeq b_\textrm{max}^2/A_\textrm{a}$.
+
+\begin{table}
+  \centering
+  \caption{Radio telescopes and array layouts.}
+  \label{tab:antenna-layouts}
+  \begin{threeparttable}
+  \begin{tabular}{ccccc} 
+    \hline
+    Telescope & Core & Number & Antenna & Frequency \\
+              & size & of Antennas & size & \\
+    (key)     & $b_\textrm{max}$ (in m) & $N_\textrm{a}$ & $A_\textrm{a}$ (in m$^2$) & $f_0$ (in MHz) \\
+    \hline
+    MWA & 1400 & 112 & 16 & 150 \\
+    (MWA-112) \\ \\
+    MWA & 1400 & 496 & 16 & 150 \\
+    (MWA-496) \\ \\
+    LOFAR & 3500 & 24 & & 50 \\
+    (LOFAR-C) & & & & \\ \\
+    LWA1 & 100 & 256 & 10 & 50 \\
+    (LWA1) \\ \\
+    LWA-OV\tnote{a} & 200 & 256 & 10 & 50 \\
+    (LWA-OV) & & & & \\ \\
+    HERA & 70 & 19 & 154 & 150 \\
+    (HERA-19) \\ \\
+    HERA & 98 & 37 & 154 & 150 \\
+    (HERA-37) \\ \\
+    HERA & 294 & 331 & 154 & 150 \\
+    (HERA-331) \\ \\
+    HERA & 1330 & 6769\tnote{b} & 154 & 150 \\
+    (HERA-6769) \\ \\
+    SKA1-Low & 1000 & 750\tnote{c} & 962 & 150 \\
+    (SKA1-LC) \\ \\
+    SKA1-Low & 1000 & 192,000\tnote{c} & 962 & 150 \\
+    Dipoles\tnote{d} \\
+    (SKA1-LCD) \\
+    \hline
+  \end{tabular}
+  \begin{tablenotes}
+    \item[a] Owens Valley LWA
+    \item[b] Hypothetically chosen to have a total collecting area of 
+      1~km$^2$
+    \item[c] This is the number of elements expected to be in the core, 
+      three-fourths of the total number
+    \item[d] All dipoles are used as independent elements without station 
+      beamforming
+  \end{tablenotes}
+  \end{threeparttable}
+\end{table}
+\footnotetext{LWA at Ovens Valley}
+
+Fig.~\ref{fig:parameter-space-computations-instruments} shows the number of 
+complex operations per frequency channel per integration timescale. Telescopes 
+that fall to the left of this line indicate MOFF imaging is computationally
+more efficient than visibility based imaging. All HERA layouts except possibly
+HERA-19 are in a parameter space where MOFF imaging holds the advantage. The 
+solid line showing future trajectory of HERA like systems will be clearly 
+favoured by MOFF imaging. The gray shaded area is for a projected LWA 
+expansion and is also predominantly in the region favouring MOFF imaging. 
+It is bounded by the LWA1 and LWA-OV on the left and right 
+respectively. The current (see Table~\ref{tab:antenna-layouts} and
+an expanded layout with a four-fold increase in number of elements over a
+two-fold increase in $b_\textrm{max}$ provide the bounds at the bottom and top 
+respectively. Current instruments such as MWA and LOFAR lie in parameter space
+favouring visibility based imaging.
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{MOFF_FX_computations_fov_gridding_annotated.eps}
+  \caption{Current and planned instruments in parameter space of
+    number of complex multiplies and adds with MOFF and FX. The dashed line
+    is the boundary at which the number of operations with MOFF and visibility 
+    based imaging are equal. MOFF imaging is more efficient for telescopes 
+    occupying the left of this line and vice versa. All the HERA layouts except
+    possibly HERA-19 lie in the parameter space favoured by MOFF imaging. And so 
+    are SKA1-LC and SKA1-LCD. The solid black curve shows the projected 
+    trajectory of bigger close-packed hexagonal layouts similar to HERA. The 
+    gray shaded area denotes the projected trajectory of the LWA bounded by LWA1 
+    (left edge), LWA-OV (right edge), current layout (bottom) and a four-fold 
+    increase in the number of elements in a two-fold increase in the core size
+    (top). Current instruments such as MWA and LOFAR fall in a region favoured 
+    by visibility based imaging.}
+  \label{fig:parameter-space-computations-instruments}
+\end{figure}
+
+We now consider antenna array layouts described by three quantities essential
 to radio interferometry, namely, maximum baseline length, number of antennas,
@@
 
-% Figure~\ref{fig:parameter-space-bll-nant-instruments} shows the
-
-\subsection{Scaling Relations: MOFF vs. FX}
-\subsection{Scaling Up}
-\subsection{Case Study}
+Fig.~\ref{fig:parameter-space-bll-nant-instruments} shows the ratio of the
+number of computations required with visibility based imaging relative to 
+MOFF imaging and the boundaries where this ratio is unity. The color scale 
+refers to the ratio while using a 1~m antenna element. Pairs of black and
+white lines of the same line style correspond to a specific antenna size. For 
+a given antenna size (annotated on the plot), the white line denotes the 
+minimum baseline length that could achieve the closest packing as a function 
+of the number of antennas. Regions to the right of the white line for a 
+given antenna size imply a physically impossible scenario where antennas will 
+have to be packed overlapping with each other. For the same antenna size, the 
+black line denotes the boundary where the ratio of the number of computations 
+with either algorithm becomes unity. Regions to the right of the black lines 
+favour usage of the MOFF algorithm. Thus the wedge between the white and black 
+lines denotes the region in parameter space where MOFF algorithm holds a 
+definite advantage while visibility based algorithms will be favoured to the 
+left of the black lines. As antenna size increases the maximum number of 
+antennas for a dense packing as a function of baseline length decreases. Hence 
+the white lines shift leftward as antenna size increases. Similarly, with 
+increase in antenna size, $N_\textrm{g}$ also decreases when field of view 
+imaging is achieved with an increasing grid spacing equal to antenna size and
+hence lowers the amount of computations required with the MOFF algorithm. This
+shifts the black curves leftward. HERA-331, HERA-1027, LWA1, SKA1-LC and 
+SKA1-LCD are inside their respective wedges indicating they are in regions that
+favour the usage of MOFF imaging.
 
@@
   \caption{Current and instruments planned for future in parameter space of
-    baseline length and number of antennas with MOFF and FX.}
+    baseline length and number of antennas with MOFF and FX. Color scale shown 
+    is for the ratio of number of computations required by visibility based
+    imaging to that by the MOFF algorithm. Different line styles denote 
+    different antenna sizes annotated. For a given antenna size, the 
+    corresponding white line denotes the maximum number of antennas that can 
+    be packed inside various baseline lengths. The region to the right of the 
+    white lines for corresponding antenna size is physically disallowed. 
+    Regions to the left of the black lines of the same corresponding antenna 
+    size (or line style) favours usage of visibility based imaging. Region 
+    inside the wedge enclosed by the black and white lines favours using the 
+    MOFF algorithm for any given antenna size. These wedges shift leftward with 
+    increasing antenna size. HERA-331, HERA-1027, LWA1, SKA1-LC and SKA1-LCD
+    are inside their corresponding wedges that favour the usage of MOFF 
+    imaging.}
   \label{fig:parameter-space-bll-nant-instruments}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{MOFF_FX_computations_fov_gridding_annotated.eps}
-  \caption{Current and planned instruments in parameter space of
-    number of complex multiplies and adds with MOFF and FX.}
-  \label{fig:parameter-space-computations-instruments}
 \end{figure}

### 2015-10-24 imaging_demo/ms.tex
Add content under section 6
From 68b60bc (master):
=====================

In ms.tex,

(*) Add content under section 6

@@
 antenna-to-grid mapping matrix, $\mathbf{M}(\mathbf{r},\mathbf{a})$ is described
-as a transformation matrix from the space of measured electric fields by the antennas
-($\mathbf{a}$) to the antenna array grid denoted by the coordinate $\mathbf{r}$. Since each
-antenna occupies a footprint typically the size of its aperture,
-$\mathbf{M}(\mathbf{r},\mathbf{a})$, which is generally of size
+as a transformation matrix from the space of measured electric fields by the 
+antennas ($\mathbf{a}$) to the antenna array grid denoted by the coordinate 
+$\mathbf{r}$. Since each antenna occupies a footprint typically the size of its 
+aperture, $\mathbf{M}(\mathbf{r},\mathbf{a})$, which is generally of size
 $N_\textrm{grid}\times N_\textrm{a}$, reduces to a sparse block-diagonal matrix
 with only $N_\textrm{a}$ blocks and roughly $N_\textrm{k}$ non-zero entries per
-block. Fig.~\ref{fig:a2g-mapping} illustrates the antenna-to-grid mapping
+block. This sparse matrix is stored in a Compressed Sparse Row (CSR) format. 
+Fig.~\ref{fig:a2g-mapping} illustrates the antenna-to-grid mapping
 matrix and the grid containing the mapped aperture footprints of the antennas.
@@
 \end{align}
-where, $\mathcal{I}(\mathbf{a})$ is a row of ones. This is achieved using
+where, $\mathcal{I}(\mathbf{a})$ is a row of ones. This is achieved by
 efficient multiplication with the sparse matrix created in the antenna-to-grid
-mapping process. Unless $\widetilde{\mathbf{W}}(\mathbf{r})$ includes time-dependent
+mapping process using the sparse matrices module in Python's SciPy package. 
+Unless $\widetilde{\mathbf{W}}(\mathbf{r})$ includes time-dependent
 effects of the ionosphere or the instrument, it needs to be computed just once
@@
 
+% The synthesized beam can be written as:
+% \begin{align}\label{eqn:syn-beam}
+%   B(\hat{\mathbf{s}}) &= \left| \sum_\mathbf{r} \widetilde{W}(\mathbf{r})\,e^{-i2\pi\mathbf{r}\cdot\hat{\mathbf{s}}}\right|^2, \\
+%   \textrm{or}, \quad B(l,m) &= \left| \sum_{r_x,r_y} \widetilde{W}(r_x,r_y)\,e^{-i2\pi(r_x l + r_y m)}\right|^2 
+% \end{align}
+% where, $\hat{\mathbf{s}}\equiv (l,m)$, $\mathbf{r}\equiv (r_x,r_y)$ and 
+% $\widetilde{W}(\mathbf{r})$ is given by equation~\ref{eqn:gridding-convolution}.
+% If the errors in the determination of $r_x$ and $r_y$ are denoted by 
+% $\delta r_x$ and $\delta r_y$ respectively, the error on $B(l,m)$ can be 
+% estimated using a simple propagation of errors, where,
+% \begin{align}
+%   \frac{\partial{B}(l,m)}{\partial{r_x}} &= e^{-i2\pi\mathbf{r}\cdot\hat{\mathbf{s}}}\left[\frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_x}} - i 2\pi l\widetilde{W}(\mathbf{r})\right] \label{eqn:syn-beam-pdev1}\\
+%   &\qquad\qquad \times \sum_{\mathbf{r}^\prime} \widetilde{W}^\star(\mathbf{r}^\prime)\,e^{+i2\pi\mathbf{r}^\prime\cdot \hat{\mathbf{s}}} \\
+%   &= e^{-i2\pi\mathbf{r}\cdot\hat{\mathbf{s}}}\left[\frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_x}} - i2\pi l\widetilde{W}(\mathbf{r})\right]\,W^\star(\hat{\mathbf{s}}), \\
+%   \textrm{and},\frac{\partial{B}(l,m)}{\partial{r_y}} &= e^{-i2\pi\mathbf{r}\cdot\hat{\mathbf{s}}}\left[\frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_x}} - i2\pi m\widetilde{W}(\mathbf{r})\right]\,W^\star(\hat{\mathbf{s}}).
+\label{eqn:syn-beam-pdev2}
+% \end{align}
+% It must be noted that although we start with a squaring operation in 
+% equation~\ref{eqn:syn-beam}, in equations~\ref{eqn:syn-beam-pdev1} through 
+% \ref{eqn:syn-beam-pdev2}, the summations are performed only when 
+% $\mathbf{r}\ne \mathbf{r}^\prime$ and therefore do not include the 
+% auto-correlations of the antenna weights. The error expected in $B(l,m)$ is:
+% \begin{align}
+%   \left|\delta B(l,m)\right|^2 &= \sum_{r_x}\left|\frac{\partial{B}(l,m)}{\partial{r_x}}\right|^2 \delta r_x^2 + \sum_{r_y}\left|\frac{\partial{B}(l,m)}{\partial{r_y}}\right|^2 \delta r_y^2 \\
+%   &= \delta r_x^2 \sum_{r_x}\left|\frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_x}} - i2\pi l\widetilde{W}(\mathbf{r})\right|^2\,\left|W(\hat{\mathbf{s}})\right|^2 \nonumber\\
+%   &\qquad + \delta r_y^2 \sum_{r_y}\left|\frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_y}} - i2\pi m\widetilde{W}(\mathbf{r})\right|^2\,\left|W(\hat{\mathbf{s}})\right|^2
+% \end{align}
+% For the current simulations, 
+% \begin{align}
+%   \delta r_x = \delta r_y &\simeq 1, \\
+%   \widetilde{W}(r_x,r_y) &= \{0, \textrm{constant}\}, \\
+%   \textrm{and},\qquad \frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_x}} &= \frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_y}} = 0.
+% \end{align}
+% Hence,
+% \begin{align}
+%   \left|\delta B(l,m)\right|^2 &\simeq 4\pi^2\left(l^2+m^2\right)\,B(\hat{\mathbf{s}})\,\sum_\mathbf{r}\,\left|\widetilde{W}(\mathbf{r})\right|^2, \\
+% \textrm{or},\qquad \left|\delta B(l,m)\right| &\simeq 2\pi\sqrt{l^2+m^2}\,\left|W(\hat{\mathbf{s}})\right|\,\left(\sum_\mathbf{r}\,\left|\widetilde{W}(\mathbf{r})\right|^2\right)^{1/2}, \label{eqn:syn-beam-error}
+% \end{align}
+% where, $B(\hat{\mathbf{s}}) \equiv \left|W(\hat{\mathbf{s}})\right|^2$. Thus, 
+% the error on the synthesized beam is proportional to the antenna power pattern
+% and increases linearly outward in radial bins of direction cosines.
+
 We study the effect of the differences in gridded weights on the image plane. 
-The synthesized beam can be written as:
-\begin{align}\label{eqn:syn-beam}
-  B(\hat{\mathbf{s}}) &= \left| \sum_\mathbf{r} \widetilde{W}(\mathbf{r})\,e^{-i2\pi\mathbf{r}\cdot\hat{\mathbf{s}}}\right|^2, \\
-  \textrm{or}, \quad B(l,m) &= \left| \sum_{r_x,r_y} \widetilde{W}(r_x,r_y)\,e^{-i2\pi(r_x l + r_y m)}\right|^2 
-\end{align}
-where, $\hat{\mathbf{s}}\equiv (l,m)$, $\mathbf{r}\equiv (r_x,r_y)$ and 
-$\widetilde{W}(\mathbf{r})$ is given by equation~\ref{eqn:gridding-convolution}.
-If the errors in the determination of $r_x$ and $r_y$ are denoted by 
-$\delta r_x$ and $\delta r_y$ respectively, the error on $B(l,m)$ can be 
-estimated using a simple propagation of errors, where,
-\begin{align}
-  \frac{\partial{B}(l,m)}{\partial{r_x}} &= e^{-i2\pi\mathbf{r}\cdot\hat{\mathbf{s}}}\left[\frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_x}} - i 2\pi l\widetilde{W}(\mathbf{r})\right] \label{eqn:syn-beam-pdev1}\\
-  &\qquad\qquad \times \sum_{\mathbf{r}^\prime} \widetilde{W}^\star(\mathbf{r}^\prime)\,e^{+i2\pi\mathbf{r}^\prime\cdot \hat{\mathbf{s}}} \\
-  &= e^{-i2\pi\mathbf{r}\cdot\hat{\mathbf{s}}}\left[\frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_x}} - i2\pi l\widetilde{W}(\mathbf{r})\right]\,W^\star(\hat{\mathbf{s}}), \\
-  \textrm{and},\frac{\partial{B}(l,m)}{\partial{r_y}} &= e^{-i2\pi\mathbf{r}\cdot\hat{\mathbf{s}}}\left[\frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_x}} - i2\pi m\widetilde{W}(\mathbf{r})\right]\,W^\star(\hat{\mathbf{s}}).
-\label{eqn:syn-beam-pdev2}
-\end{align}
-It must be noted that although we start with a squaring operation in 
-equation~\ref{eqn:syn-beam}, in equations~\ref{eqn:syn-beam-pdev1} through 
-\ref{eqn:syn-beam-pdev2}, the summations are performed only when 
-$\mathbf{r}\ne \mathbf{r}^\prime$ and therefore do not include the 
-auto-correlations of the antenna weights. The error expected in $B(l,m)$ is:
-\begin{align}
-  \left|\delta B(l,m)\right|^2 &= \sum_{r_x}\left|\frac{\partial{B}(l,m)}{\partial{r_x}}\right|^2 \delta r_x^2 + \sum_{r_y}\left|\frac{\partial{B}(l,m)}{\partial{r_y}}\right|^2 \delta r_y^2 \\
-  &= \delta r_x^2 \sum_{r_x}\left|\frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_x}} - i2\pi l\widetilde{W}(\mathbf{r})\right|^2\,\left|W(\hat{\mathbf{s}})\right|^2 \nonumber\\
-  &\qquad + \delta r_y^2 \sum_{r_y}\left|\frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_y}} - i2\pi m\widetilde{W}(\mathbf{r})\right|^2\,\left|W(\hat{\mathbf{s}})\right|^2
-\end{align}
-For the current simulations, 
-\begin{align}
-  \delta r_x = \delta r_y &\simeq 1, \\
-  \widetilde{W}(r_x,r_y) &= \{0, \textrm{constant}\}, \\
-  \textrm{and},\qquad \frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_x}} &= \frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_y}} = 0.
-\end{align}
-Hence,
-\begin{align}
-  \left|\delta B(l,m)\right|^2 &\simeq 4\pi^2\left(l^2+m^2\right)\,B(\hat{\mathbf{s}})\,\sum_\mathbf{r}\,\left|\widetilde{W}(\mathbf{r})\right|^2, \\
-\textrm{or},\qquad \left|\delta B(l,m)\right| &\simeq 2\pi\sqrt{l^2+m^2}\,\left|W(\hat{\mathbf{s}})\right|\,\left(\sum_\mathbf{r}\,\left|\widetilde{W}(\mathbf{r})\right|^2\right)^{1/2}, \label{eqn:syn-beam-error}
-\end{align}
-where, $B(\hat{\mathbf{s}}) \equiv \left|W(\hat{\mathbf{s}})\right|^2$. Thus, 
-the error on the synthesized beam is proportional to the antenna power pattern
-and increases linearly outward in radial bins of direction cosines.
-
 Fig.~\ref{fig:psf-diff} shows the difference between the synthesized beams 
@@
 beams is shown on top. The amplitude of the difference appears to be 
-modulated by the directional voltage response of the antenna. At the bottom, 
+modulated by the directional power response of the antenna. At the bottom, 
 in radial bins, the rms of the synthesized beam (gray) and the rms of the 
@@
 The synthesized beam rms is proportional to the antenna power pattern as 
-expected from a point spread function of uncorrected for the antenna power 
-pattern. A notable observation is that the rms of the difference of
-the synthesized beams seems proportional to the directional antenna voltage
-pattern and also increases gradually with distance from the center as seen 
-from the diverging amplitudes between the black and red curves as predicted
-in equation~\ref{eqn:syn-beam-error}. The rms of the difference appears to 
-be bounded by the actual rms of the synthesized beam indicating that these
-rounding errors in gridding do not affect the statistics of the images or
-the synthesized beams.
+expected from a point spread function uncorrected for the antenna power 
+pattern. The amplitude of the rms of difference of the synthesized beams is 
+also modulated by the antenna power pattern. The rms of the difference is 
+definitely lesser than the rms of the synthesized beam in the central regions 
+up to $(l^2+m^2)^{1/2}\lesssim 0.3$. This implies that the beams are well 
+matched in the central regions. In the outer regions, their mismatch is 
+comparable to the rms of synthesized beams. This indicates the two 
+synthesized beams are not completely randomly different from each other in
+which case the rms of the difference would have been $\approx \sqrt{2}$
+higher than the rms of the each of the synthesized beams. This indicates 
+that while differences exist, large fractions of them are still well matched
+to each other.Thus the rounding errors in gridding do not affect the 
+statistics of the images or the synthesized beams.
 
@@
 \end{table}
-\footnotetext{LWA at Ovens Valley}
 
@@
 SKA1-LCD are inside their respective wedges indicating they are in regions that
-favour the usage of MOFF imaging.
+favour the usage of MOFF imaging. For fixed baseline length, regions favouring
+the MOFF algorithm tend to be towards large $N_\textrm{a}$ indicating large-N
+dense array layouts with smaller antenna elements are best suited for deploying 
+the MOFF algorithm.
 
@@
 \section{Conclusions}\label{sec:conclusions}
+
+As radio astronomy is entering a new era, advances in instrumentation have to
+be accompanied by equal advances in processing techniques to manage 
+computational resources. Many future radio telescopes such as the SKA, HERA and 
+LWA are headed towards the large-N dense array layout model for which 
+computational cost from traditional FX/XF correlator based architecture and 
+visibility based imaging starts rising steeply. We have provided the first 
+software demonstration of a general purpose imaging algorithm using our generic 
+and efficient EPIC software that is designed to bring this cost down from 
+$\mathcal{O}(N^2)$ to $\mathcal{O}(N\log N)$. Under the class of direct imaging 
+techniques, ours is one of the most generic -- neither does it place any 
+constraint on the array layout to be on a regular grid nor does it require the
+antenna array to be homogeneous.
+
+Our package, now publicly available, written in object oriented Python is 
+highly modularized and parallelizable. It includes an implementation of the 
+MOFF algorithm in addition to visibility based software holography imaging and 
+a data simulator for sky models. It has been successfully tested on simulated 
+as well as real LWA observations.
+
+The MOFF algorithm packaged with EPIC is already found to be most suitable 
+for many present and planned radio telescopes such as the LWA, HERA and SKA. 
+In general, MOFF is most suited to operate in the region of parameter space
+characterized by dense packing of a large number of antennas especially when 
+consisting of a large number of small antenna elements.
+
+A unique and significant advantage is the instantaneous availability of 
+calibrated time-domain images bundled together with the hardware frontend 
+such as the F-engine as an integrated module. Hence, it is a compelling 
+candidate for time-domain radio astronomy, e.g. search for and monitoring of 
+transients. Transient detection pipeline at the backend of EPIC can be 
+fine-tuned to target fast transients such as the Fast Radio Bursts 
+\citep[FRB;][]{tho13} on millisecond timescales at GHz frequencies or slow 
+transients from planetary and exoplanetary origins at frequencies around 
+100~MHz.
+
+Thus, EPIC with the MOFF algorithm packaged is uniquely poised to offer a
+substantial advantage to imaging with large-N dense arrays typical of 
+next-generation radio telescopes as well as push the frontiers of 
+time-domain astronomy to fill gaps in understanding the science behind 
+phenomena responsible for extreme transient events in the Universe.
+
+In the near future, we plan to upgrade our current Python implementation 
+of EPIC to a GPU based pipeline in order to operate on real-time data and 
+develop a transient trigger and monitor backend. In the meanwhile, we plan 
+to demonstrate the capability of EPIC to calibrate and image from heterogeneous 
+arrays and incorporate corrections for non-coplanarity of baselines and 
+direction-dependence of calibration.
+
 \section*{Acknowledgements}
 
-The Acknowledgements section is not numbered. Here you can thank helpful
-colleagues, acknowledge funding agencies, telescopes and facilities used etc.
-Try to keep it short.
+We thank Larry D'Addario, Gregg Hallinan and Harish Vedantham for their valuable 
+inputs.
 
@@
 
-\appendix
-
-\section{Some extra material}
-
-If you want to present additional material which would interrupt the flow of the main paper,
-it can be placed in an Appendix which appears after the list of references.
+% \appendix
+
+% \section{Some extra material}
+
+% If you want to present additional material which would interrupt the flow of the main paper,
+% it can be placed in an Appendix which appears after the list of references.
 

### 2015-10-24 imaging_demo/ms.tex
Add abstract
From 95fd8c8 (master):
=====================

In ms.tex,

(*) Add abstract

@@
 \begin{abstract}
-Abstract here (250 words)
+Modern radio telescopes are marching towards densely packed array layouts 
+consisting of a large number of antennas ($N$). Since the cost of traditional 
+correlators scales as $\mathcal{O}(N^2)$, there will be a steep cost for 
+realizing the full imaging potential of these powerful instruments. Through 
+our generic and efficient E-field Parallel Imaging Correlator (EPIC), we present 
+the first software demonstration of a generalized direct imaging algorithm 
+known by the name Modular Optimal Frequency Fourier (MOFF) imager. It takes 
+advantage of the multiplication-convolution theorem of Fourier transform. Not 
+only does it bring down the cost to $\mathcal{O}(N\log N)$ but can also image 
+from irregularly arranged antenna arrays. EPIC is highly modular and 
+parallelizable. It is implemented in object oriented Python and is publicly 
+available. We have verified the images produced to be mathematically identical 
+to those produced using traditional techniques. We have also validated our 
+implementation on data observed with the Long Wavelength Array. Antenna systems
+with a dense filling factor consisting of a large number of antennas such as 
+LWA, SKA, and HERA will gain a significant advantage by deploying EPIC. 
+Inherent availability of calibrated time-domain images on timescales roughly 
+equal to the writeout timescale of the digitizer will make it a prime 
+candidate for transient searches of primarily Fast Radio Bursts (FRB) as well 
+as slower planteary and exoplantenary phenomena. 
 \end{abstract}

### 2015-10-24 imaging_demo/ms.tex
fixing some typos

@@
 the first software demonstration of a generalized direct imaging algorithm 
-known by the name Modular Optimal Frequency Fourier (MOFF) imager. It takes 
-advantage of the multiplication-convolution theorem of Fourier transform. Not 
+known as the Modular Optimal Frequency Fourier (MOFF) imager. It takes 
+advantage of the multiplication-convolution theorem of Fourier transforms. Not 
 only does it bring down the cost to $\mathcal{O}(N\log N)$ but can also image 
-from irregularly arranged antenna arrays. EPIC is highly modular and 
+from irregularly arranged heterogeneous antenna arrays. EPIC is highly modular and 
 parallelizable. It is implemented in object oriented Python and is publicly 
@@
 
-Here, we discuss the components of these architectures in detail.
+Here we discuss the components of these architectures in detail.
 
@@
 auto-correlations around the zero spacing in the $uv$-plane which are absent in 
-traditional visibility-based imaging. In order for a robust comparison, these 
+traditional visibility-based imaging. In order to achieve a robust comparison, these 
 auto-correlations have to be removed from both the $uv$ and image planes.
@@
 expected from a point spread function uncorrected for the antenna power 
-pattern. The amplitude of the rms of difference of the synthesized beams is 
+pattern. The rms of differenced synthesized beams is 
 also modulated by the antenna power pattern. The rms of the difference is 
@@
 It is bounded by the LWA1 and LWA-OV on the left and right 
-respectively. The current (see Table~\ref{tab:antenna-layouts} and
+respectively. The current (see Table~\ref{tab:antenna-layouts}) and
 an expanded layout with a four-fold increase in number of elements over a

### 2015-10-26 imaging_demo/ms.tex
Incorporate many of the first round of suggestions
From a5f2570 (master):
=====================

(*) Add LWA_MOFF_bandavg_image_4_iterations_analytic_aperture.eps to
repository and remove LWA_sample_image_391_iterations.png from the
repository

In epic.bib,

(*) Add Bowman et al.(2013) to references

In ms.tex,

(*) Many many of the changes suggested by Judd and Adam
(*) Modify table

@@
 \begin{abstract}
-Modern radio telescopes are marching towards densely packed array layouts 
-consisting of a large number of antennas ($N$). Since the cost of traditional 
-correlators scales as $\mathcal{O}(N^2)$, there will be a steep cost for 
+Modern radio telescopes are favouring densely packed array layouts 
+consisting of a large number of antennas 
+($N_\textrm{a}\gtrsim 1000$--$\gtrsim 10000$). Since the cost of traditional 
+correlators scales as $\mathcal{O}(N_\textrm{a}^2)$, there will be a steep cost for 
 realizing the full imaging potential of these powerful instruments. Through 
@@
 advantage of the multiplication-convolution theorem of Fourier transforms. Not 
-only does it bring down the cost to $\mathcal{O}(N\log N)$ but can also image 
-from irregularly arranged heterogeneous antenna arrays. EPIC is highly modular and 
-parallelizable. It is implemented in object oriented Python and is publicly 
-available. We have verified the images produced to be mathematically identical 
-to those produced using traditional techniques. We have also validated our 
-implementation on data observed with the Long Wavelength Array. Antenna systems
-with a dense filling factor consisting of a large number of antennas such as 
-LWA, SKA, and HERA will gain a significant advantage by deploying EPIC. 
-Inherent availability of calibrated time-domain images on timescales roughly 
-equal to the writeout timescale of the digitizer will make it a prime 
-candidate for transient searches of primarily Fast Radio Bursts (FRB) as well 
-as slower planteary and exoplantenary phenomena. 
+only does it bring down the cost to $\mathcal{O}(N_\textrm{a}\log N_\textrm{a})$ 
+but can also image from irregularly arranged heterogeneous antenna arrays. 
+EPIC is highly modular and parallelizable. It is implemented in object oriented 
+Python and is publicly available. We have verified the images produced to be 
+mathematically identical to those produced using traditional techniques. We have 
+also validated our implementation on data observed with the Long Wavelength 
+Array. Antenna systems with a dense filling factor consisting of a large number 
+of antennas such as LWA, SKA, and HERA will gain a significant advantage by 
+deploying EPIC. Inherent availability of calibrated time-domain images on 
+timescales roughly equal to the writeout timescale of the digitizer will make 
+it a prime candidate for transient searches of primarily Fast Radio Bursts (FRB) 
+as well as slower planteary and exoplantenary phenomena. 
 \end{abstract}
@@
 same direction. A number of new telescopes are under development around the world
-based on this new paradigm, including the Murchison Widefield Array (MWA,
-\citealt{tin13}), the Precision Array for Probing the Epoch of Reionization
-(PAPER, \citealt{par10}), the Hydrogen Epoch of Reionization Array (HERA
-\footnote{http://reionization.org}), the LOw Frequency ARray (LOFAR,
-\citealt{dev09}), the Canadian Hydrogen Intensity Mapping Experiment (CHIME,
+based on this new paradigm, including the Murchison Widefield Array (MWA;
+\citealt{tin13,bow13}), the Precision Array for Probing the Epoch of Reionization
+(PAPER; \citealt{par10}), the Hydrogen Epoch of Reionization 
+Array\footnote{http://reionization.org} (HERA), the LOw Frequency ARray (LOFAR;
+\citealt{van13}), the Canadian Hydrogen Intensity Mapping Experiment (CHIME,
 \citealt{ban14}), the Long Wavelength Array (LWA, \citealt{ell13}), and the low
@@
   between antenna pairs represent spatial correlations which can be time-averaged
-  followed by gridding and imaging. However, in MOFF imaging both
-  antenna and gridded electric fields are stochastic and therefore must be
-  imaged and squared before time-averaging. 
+  followed by gridding and imaging. However, in MOFF imaging both antenna and 
+  gridded electric fields are stochastic and therefore must be imaged and squared 
+  before time-averaging. 
 \item In FX imaging, electric fields measured by antennas are not correlated with
   themselves and hence lack zero spacing measurements. In contrast, in MOFF
-  imaging, since the gridded electric fields are imaged and squared, they
-  contain information from auto-correlated electric fields at zero spacing.
-  Hence, they must be subtracted from the images.
+  imaging, since the gridded electric fields are imaged and squared, they retain 
+  information from auto-correlated electric fields at zero spacing and thus yield 
+  the true total power of the imaged field.
 \end{enumerate}
@@
 \subsection{Calibration}
-Calibration of direct imaging correlators remains an unresolved matter. Contrary
+Calibration of direct imaging correlators remains a challenge. Contrary
 to the FX data flow, direct imagers mix the signals from all antennas before
@@
 before the gridding step. Previous efforts have resorted to applying FX-generated
-calibration solutions \citep{zhe14, fos14}, or integrating a dedicated FX correlator 
-which periodically forms the full visibility matrix \citep{wij09, dev09}.
+calibration solutions \citep{zhe14,fos14}, or integrating a dedicated FX 
+correlator which periodically forms the full visibility matrix 
+\citep{wij09,dev09}.
 
@@
 in equation~\ref{eqn:cc-vis}. It is advantageous to average them in visibilities
-before imaging because visibilities represent a compact representation of the
-information in images. Hence it is more computationally efficient to average
-antenna pair visibilities rather than images. Since this averaging has been
-performed already on the visibilities over an integration timescale, the imaging
-step has to be performed only once per integration cycle. FX imaging holds this
-advantage as long as the square of the number of antennas is smaller than the
-number of pixels on the image.
+before imaging because the repeated cost of spatial FFT can be avoided. Since 
+this averaging has been performed already on the visibilities over an integration 
+timescale, the imaging step has to be performed only once per integration cycle. 
+% FX imaging holds this advantage as long as the square of the number of antennas 
+% is smaller than the number of pixels on the image.
 
@@
 auto-correlations around the zero spacing in the $uv$-plane which are absent in 
-traditional visibility-based imaging. In order to achieve a robust comparison, these 
-auto-correlations have to be removed from both the $uv$ and image planes.
+traditional visibility-based imaging. In order to facilitate a robust comparison, 
+these auto-correlations are removed from both the $uv$ and image planes, which 
+is otherwise not an essential part of the core algorithm. We describe below how 
+they can be removed from the $uv$ plane in a straightforward manner.
 
 The shape and extent of these auto-correlations can be estimated from the 
-antenna aperture illumination pattern which is a quantity already available 
-from the gridding step. Fig.~\ref{fig:autocorr_wts_PB} shows the estimated 
-weights from antenna auto-correlations in the $uv$-plane (left) and the 
-corresponding response in the image plane (right). The latter is simply the 
-directional antenna power response.
+antenna aperture illumination pattern. The aperture illumination patterns
+are already available from the gridding step. Fig.~\ref{fig:autocorr_wts_PB} 
+shows the estimated weights from antenna auto-correlations in the $uv$-plane 
+(left) and the corresponding response in the image plane (right). The latter 
+is simply the directional antenna power response.
 
@@
 \section{Verification}\label{sec:verify}
+
+In order to prove the accuracy of the EPIC code, we verify the images produced 
+through simulations. We simulate electric field streams from a model sky and 
+process the data through both the MOFF and visibility based imaging algorithms. 
+We then compare the output images to demonstrate their equivalence.
 
@@
 
-\subsection{Differences in outputs}\label{sec:diff}
+\subsection{Comparison of outputs}\label{sec:diff}
 
@@
 The latter is the most general solution applicable especially in the case of 
-heterogeneous antenna arrays.
+heterogeneous antenna arrays and is under active development for EPIC.
 
@@
 $\sim 1$--2 pixels. This effect can be mitigated only by making the grid 
-spacing finer.
+spacing finer at the expense of increased computational cost.
 
@@
     with the fluxes expected after multiplication by the antenna power pattern. 
-    The synthesized beams from the two algorithms are well matches in size and
+    The synthesized beams from the two algorithms are well matched in size and
     shape. The overall modulation by the power pattern is seen clearly in both
@@
 that while differences exist, large fractions of them are still well matched
-to each other.Thus the rounding errors in gridding do not affect the 
-statistics of the images or the synthesized beams.
+to each other even out to the horizon. Thus the rounding errors in gridding 
+do not affect the statistics of the images or the synthesized beams.
 
@@
 
-We have analyzed narrow band data from the LWA in New Mexico. This data is  
-in LWA TBN format with 512 voltage time samples from 255 antennas within 
-roughly a diameter of 100~m. The data is centered at a frequency of 74.03~MHz,
-with a sample rate (equal to the bandwidth) of 100~kHz with 512 complex time 
-samples in a A/D writeout timescale of 5.12~ms, a frequency resolution of 
-195.3125~Hz and dual polarization. There are 391 such timestamps yielding a
-total duration of 2~s.
-
-We have matched the cable delays of all the antennas but yet to deploy the 
-calibration pipeline we have developed on this data. Therefore, in 
-subsequent discussion we assumed the data is calibrated.
+Here we demonstrate our software using narrow band data from the LWA in New 
+Mexico. This data is in LWA narrow-band transient buffer (TBN) format with 512 
+voltage time samples from 255 antennas within roughly a diameter of 100~m. The 
+data is centered at a frequency of 74.03~MHz, with a sample rate (equal to the 
+bandwidth) of 100~kHz with 512 complex time samples in a A/D writeout timescale 
+of 5.12~ms, a frequency resolution of 195.3125~Hz and dual polarization. There 
+are 391 such timestamps yielding a total duration of 2~s.
+
+We corrected the cable delays, but otherwise assume the data is sufficiently 
+calibrated to image directly. A test of EPICal on this data will be conducted 
+in the future.
 
@@
 $\approx 80$\% of bandwidth (roughly 80~kHz). The image is shown in direction 
-cosine coordinates -- $l$ along x-axis and $m$ along y-axis. Due to lack of
-calibration the flux scale is arbitrary. Despite the lack of calibration,
-only with delay compensation, we see Cyg~A and Cas~A prominently at
-$(l,m)\approx (0, 0.125)$, and (0.4,0.55) respectively and Sgr~A faintly at
-(-0.5,-0.75), thus validating the functionality of the EPIC package.
+cosine coordinates -- $l$ along x-axis and $m$ along y-axis. The flux scale 
+is arbitrary. Even in this proof-of-concept demonstration, we see Cyg~A and 
+Cas~A prominently and Sgr~A faintly as annotated, thus validating the 
+functionality of the EPIC package.
 
 \begin{figure}
-  \includegraphics[width=\columnwidth]{LWA_sample_image_391_iterations.png}
+  \includegraphics[width=\columnwidth]{LWA_MOFF_bandavg_image_4_iterations_analytic_aperture.eps}
   \caption{Image from LWA TBN data obtained with MOFF imaging using EPIC 
@@
     are compensated for their respective delays. The flux scale is arbitrary
-    and not calibrated. Cyg~A is prominently visible near the center at
-    $(l,m)\approx (0, 0.125)$, Cas~A near (0.4,0.55) and Sgr~A near 
-    (-0.5,-0.75).}
+    and not calibrated. Locations of Cyg~A, Cas~A, Sgr~A, Her~A and Jupiter
+    are annotated.}
   \label{fig:LWA-image}
@@
 We have profiled the core routines of EPIC line-by-line for various ranges of 
-parameters such as antenna filling fraction, maximum baseline 
-length, bandwidth and frequency resolution, integration timescale, etc. for 
-HERA antenna layouts which are highly compact.
+parameters such as antenna filling fraction, maximum baseline length, bandwidth 
+and frequency resolution, integration timescale, etc. for HERA antenna layouts 
+which are highly compact. However, we note that in general, the hardware and 
+optimization of routines in place will determine the relative speeds of the 
+different stages in the pipeline.
 
 Of all steps in the MOFF pipeline that are repeated for every writeout from 
-the F-engine, the slowest step even for dense HERA 
-layouts is found to be the spatial two-dimensional FFT in the imaging stage 
-relative to applying the sparse matrix gridding convolution, squaring or 
-time-averaging. For instance, even in the dense array layout scenario that 
-makes these stages perform the slowest, the gridding convolution, squaring and 
-time-averaging take up only $\simeq 40$\%, $\simeq 40$\% and $\simeq 15$\% 
-respectively of the time taken by the spatial Fourier transform. With sparser 
-arrays they will be even faster.
+the F-engine, the slowest step even for dense HERA layouts is found to be the 
+spatial two-dimensional FFT in the imaging stage relative to applying the sparse 
+matrix gridding convolution, squaring or time-averaging. For instance, even in 
+the dense array layout scenario that makes these stages perform the slowest, the 
+gridding convolution, squaring and time-averaging take up only $\simeq 40$\%, 
+$\simeq 40$\% and $\simeq 15$\% respectively of the time taken by the spatial 
+Fourier transform. With sparser arrays the gridding process will be be even 
+faster.
 
@@
               & size & of Antennas & size & \\
-    (key)     & $b_\textrm{max}$ (in m) & $N_\textrm{a}$ & $A_\textrm{a}$ (in m$^2$) & $f_0$ (in MHz) \\
+              & $b_\textrm{max}$ (in m) & $N_\textrm{a}$ & $A_\textrm{a}$ (in m$^2$) & $f_0$ (in MHz) \\
     \hline
-    MWA & 1400 & 112 & 16 & 150 \\
-    (MWA-112) \\ \\
-    MWA & 1400 & 496 & 16 & 150 \\
-    (MWA-496) \\ \\
-    LOFAR & 3500 & 24 & & 50 \\
-    (LOFAR-C) & & & & \\ \\
+    MWA-112\tnote{a} & 1400 & 112 & 16 & 150 \\
+    MWA-240\tnote{a} & 1400 & 240 & 16 & 150 \\
+    MWA-496\tnote{a} & 1400 & 496 & 16 & 150 \\
+    MWA-112\tnote{a} & 1400 & 1008 & 16 & 150 \\
+    LOFAR-LC\tnote{b} & 3500 & 24 & 5809 & 50 \\
+    LOFAR-HC\tnote{b} & 3500 & 48 & 745 & 150 \\
     LWA1 & 100 & 256 & 10 & 50 \\
-    (LWA1) \\ \\
-    LWA-OV\tnote{a} & 200 & 256 & 10 & 50 \\
-    (LWA-OV) & & & & \\ \\
-    HERA & 70 & 19 & 154 & 150 \\
-    (HERA-19) \\ \\
-    HERA & 98 & 37 & 154 & 150 \\
-    (HERA-37) \\ \\
-    HERA & 294 & 331 & 154 & 150 \\
-    (HERA-331) \\ \\
-    HERA & 1330 & 6769\tnote{b} & 154 & 150 \\
-    (HERA-6769) \\ \\
-    SKA1-Low & 1000 & 750\tnote{c} & 962 & 150 \\
-    (SKA1-LC) \\ \\
-    SKA1-Low & 1000 & 192,000\tnote{c} & 962 & 150 \\
-    Dipoles\tnote{d} \\
-    (SKA1-LCD) \\
+    LWA-OV\tnote{c} & 200 & 256 & 10 & 50 \\
+    HERA-19 & 70 & 19 & 154 & 150 \\
+    HERA-37 & 98 & 37 & 154 & 150 \\
+    HERA-331 & 294 & 331 & 154 & 150 \\
+    HERA-6769\tnote{d} & 1330 & 6769 & 154 & 150 \\
+    SKA1-LC\tnote{e} & 1000 & 750 & 962 & 150 \\
+    SKA1-LCD\tnote{f} & 1000 & 192,000 & 962 & 150 \\
+    CHIME & 100 & 1280 & 8 & 600 \\
+    HIRAX & 200 & 1024 & 6 & 600 \\
     \hline
@@
   \begin{tablenotes}
-    \item[a] Owens Valley LWA
-    \item[b] Hypothetically chosen to have a total collecting area of 
+    \item[a] MWA-N denotes N tiles in the specified core diameter
+    \item[b] LC and HC denotes low band and high band stations inside the 
+      specified core diameter 
+    \item[c] Owens Valley LWA
+    \item[d] Hypothetically chosen to have a total collecting area of 
       1~km$^2$
-    \item[c] This is the number of elements expected to be in the core, 
-      three-fourths of the total number
-    \item[d] All dipoles are used as independent elements without station 
-      beamforming
+    \item[e] This is the number of beamformed stations expected to be in the 
+      core, roughly three-fourths of the total number
+    \item[f] All dipoles inside the core are used as independent elements 
+      without station beamforming
   \end{tablenotes}
@@
 dense array layouts with smaller antenna elements are best suited for deploying 
-the MOFF algorithm.
+EPIC.
 
@@
 
-% \subsection{Figures and tables}
-
-% Figures and tables should be placed at logical positions in the text. Don't
-% worry about the exact layout, which will be handled by the publishers.
-
-% Figures are referred to as e.g. Fig.~\ref{fig:example_figure}, and tables as
-% e.g. Table~\ref{tab:example_table}.
-
-% % Example figure
-% \begin{figure}
-% 	% To include a figure from a file named example.*
-% 	% Allowable file formats are eps or ps if compiling using latex
-% 	% or pdf, png, jpg if compiling using pdflatex
-% 	\includegraphics[width=\columnwidth]{example}
-%     \caption{This is an example figure. Captions appear below each figure.
-% 	Give enough detail for the reader to understand what they're looking at,
-% 	but leave detailed discussion to the main body of the text.}
-%     \label{fig:example_figure}
-% \end{figure}
-
-% % Example table
-% \begin{table}
-% 	\centering
-% 	\caption{This is an example table. Captions appear above each table.
-% 	Remember to define the quantities, symbols and units used.}
-% 	\label{tab:example_table}
-% 	\begin{tabular}{lccr} % four columns, alignment for each
-% 		\hline
-% 		A & B & C & D\\
-% 		\hline
-% 		1 & 2 & 3 & 4\\
-% 		2 & 4 & 6 & 8\\
-% 		3 & 5 & 7 & 9\\
-% 		\hline
-% 	\end{tabular}
-% \end{table}
-
 \section{Conclusions}\label{sec:conclusions}
@@
 
-We thank Larry D'Addario, Gregg Hallinan and Harish Vedantham for their valuable 
-inputs.
+We thank Greg Taylor for providing us the LWA data, and Larry D'Addario, 
+Gregg Hallinan and Harish Vedantham for their valuable inputs. Construction of 
+the LWA has been supported by the Office of Naval Research under Contract 
+N00014-07-C-0147. Support for operations and continuing development of the LWA1 
+is provided by the National Science Foundation under grant AST-1139974 of the 
+University Radio Observatory program.
 

### 2015-10-26 imaging_demo/ms.tex
Restructure section 3, describe software modules
From 7ed89ad (master):
=====================

(*) Add new figure EPIC-modules.eps to the repository

In ms.tex,

(*) Add a new figure illustrating software architecture and modules
(*) Restructure subsections in section 3
(*) Add subsection on software architecture
(*) Move removal of auto-correlations to section 4 under Verification

@@
 &\qquad\left\langle \left[\widetilde{\mathbf{W}}^\textrm{T}_a(\mathbf{r},\mathbf{a})\, \widetilde{\mathbf{N}}^{-1}\!(\mathbf{a},\mathbf{a})\, \widetilde{\mathbf{E}}(\mathbf{a})\right]
-\ast\left[\widetilde{\mathbf{W}}_a(\mathbf{r},\mathbf{a})\, \mathbf{N}^{-1}\!(\mathbf{a},\mathbf{a})\, \mathbf{E}^\star(\mathbf{a})\right]\right\rangle_t\label{eqn:e-field-conv}
+\ast\left[\widetilde{\mathbf{W}}_a(\mathbf{r},\mathbf{a})\, \mathbf{N}^{-1}\!(\mathbf{a},\mathbf{a})\, \widetilde{\mathbf{E}}^\star(\mathbf{a})\right]\right\rangle_t\label{eqn:e-field-conv}
 \end{align}
@@
 
+\subsection{Data Flow}
+
+EPIC can accept dual-polarization inputs and produce images of all four 
+cross-polarizations. Currently two data input formats exist for reading in the 
+electric field time samples measured by the antennas -- simulated electric 
+fields based on a sky model using the simulator packaged with EPIC; and LWA 
+data. Efforts to build interfaces for data from other telescopes are underway.
+
 Fig.~\ref{fig:MOFF-flowchart} shows the flowchart for MOFF imaging. The
@@
 temporal Fourier transform on the electric field time-series to obtain electric
-field spectra $E_1(f)\ldots E_\textrm{N}(f)$ ($\widetilde{\mathbf{E}}(\mathbf{a})$ in matrix
-notation) for each of the antennas. Each of the complex antenna gains are
-calibrated to correct the corresponding electric field spectra. These calibrated
-electric fields are gridded using an antenna-based gridding convolution function
-after which it is spatially Fourier transformed and squared to obtain images
-for every time stamp. These images are then time-averaged to obtain the
-accumulated image $I(f)$ ($\mathbf{I}(\hat{\mathbf{s}})$ in matrix notation).
+field spectra $E_1(f)\ldots E_\textrm{N}(f)$ 
+($\widetilde{\mathbf{E}}(\mathbf{a})$ in matrix notation) for each of the 
+antennas. Each of the complex antenna gains are calibrated to correct the 
+corresponding electric field spectra. These calibrated electric fields are 
+gridded using an antenna-based gridding convolution function after which it is 
+spatially Fourier transformed and squared to obtain images for every time stamp. 
+These images are then time-averaged to obtain the accumulated image 
+$I(f)$ ($\mathbf{I}(\hat{\mathbf{s}})$ in matrix notation).
 \begin{figure}
@@
 
-\subsection{Temporal Fourier transform}\label{sec:F-engine}
-
-This module is common to the MOFF and FX imaging techniques. Time samples of
-electric fields measured by the antenna and digitized by the A/D converter is
-Fourier transformed to generate electric field spectra. This step can be
-parallelized by antennas as shown in Fig.~\ref{fig:f-engine}. The output is
-then fed to either MOFF and FX imaging pipelines.
+\par\medskip
+\noindent {\bf Temporal Fourier transform}
+\par\medskip
+\noindent This module is common to the MOFF and FX imaging techniques. Time 
+samples of electric fields measured by the antenna and digitized by the A/D 
+converter is Fourier transformed to generate electric field spectra. This step 
+can be parallelized by antennas as shown in Fig.~\ref{fig:f-engine}. The output 
+is then fed to either MOFF and FX imaging pipelines.
 
@@
 
-\subsection{Antenna-to-Grid Mapping}
-
-A grid is generated on the coordinate system in which antenna locations are
-specified with a grid spacing that is at most $\lambda_\textrm{min}/2$ even at the
-highest frequency to ensure there is no aliasing even from regions of the sky far
-away from the field of view. The number of locations on the grid is restricted to
-be a power of 2.
+\par\medskip
+\noindent {\bf Antenna-to-Grid Mapping}
+\par\medskip
+\noindent A grid is generated on the coordinate system in which antenna 
+locations are specified with a grid spacing. The grid spacing can be controlled 
+by the user. By default, it is set to be at $\lambda_\textrm{min}/2$ even at the 
+highest frequency to ensure there is no aliasing even from regions of the sky 
+far away from the field of view. The number of locations on the grid is 
+restricted to be a power of 2.
 
@@
 illumination function, $\widetilde{B}(\mathbf{r}-\mathbf{r}_a)$, which can
-be specified either by a functional form or as a table of values against locations
-around the antennas. A nearest neighbor mapping from all antenna footprints to
-grid locations is created using an efficient k-d tree algorithm \citep{man99}.
-There is no restriction here that the aperture illumination function has to be
-identical across antennas.
+be specified either by a functional form or as a table of values against 
+locations around the antennas. A nearest neighbor mapping from all antenna 
+footprints to grid locations is created using an efficient k-d tree algorithm 
+\citep{man99}. There is no restriction here that the aperture illumination 
+function has to be identical across antennas.
 
@@
 
-\subsection{Calibration}
-Calibration of direct imaging correlators remains a challenge. Contrary
+\par\medskip
+\noindent {\bf Calibration}
+\par\medskip
+\noindent Calibration of direct imaging correlators remains a challenge. Contrary
 to the FX data flow, direct imagers mix the signals from all antennas before
@@
 correlators to estimate antenna complex gains. This method correlates the antenna 
-electric field signals with an image pixel form the output of the correlator in the 
-feedback calibration fashion outlined in \citealt{mor11} (illustrated in Fig. 
+electric field signals with an image pixel form the output of the correlator in 
+the feedback calibration fashion outlined in \citealt{mor11} (illustrated in Fig. 
 \ref{fig:MOFF-flowchart} by the arrow leading from the imager to the calibration 
-block). Furthermore it allows for arbitrarily complex sky models, and following the 
-MOFF algorithm places no restriction on array layout, and accounts for non-identical 
-antenna beam patterns. Because only a single correlation is needed for each 
-antenna, the computation complexity scales only as $N_{\mathrm{ant}}$.
-
-The calibration module included in the EPIC repository allows for application of pre-
-determined calibration solutions, or can solve for the complex gains using the 
-EPICal algorithm.
-
-\subsection{Gridding Convolution}
-
-The antenna array aperture illumination over the entire grid,
-$\widetilde{\mathbf{W}}(\mathbf{r})$, is obtained by a projection of the individual
-antenna aperture illuminations:
+block). Furthermore it allows for arbitrarily complex sky models, and following 
+the MOFF algorithm places no restriction on array layout, and accounts for 
+non-identical antenna beam patterns. Because only a single correlation is needed 
+for each antenna, the computation complexity scales only as $N_{\mathrm{ant}}$.
+
+The calibration module included in the EPIC repository allows for application of 
+pre-determined calibration solutions, or can solve for the complex gains using 
+the EPICal algorithm.
+
+\par\medskip
+\noindent {\bf Gridding Convolution}
+\par\medskip
+\noindent The antenna array aperture illumination over the entire grid,
+$\widetilde{\mathbf{W}}(\mathbf{r})$, is obtained by a projection of the 
+individual antenna aperture illuminations:
 \begin{align}\label{eqn:gridding-convolution}
@@
 
-\subsection{Spatial Fourier Transform}
-
-Before the spatial Fourier transform, the gridded electric fields are padded with
-zeros in order to match the grid size and angular size of each image pixel that
-would have been obtained with software holography of output from an FX
+\par\medskip
+\noindent {\bf Spatial Fourier Transform}
+\par\medskip
+\noindent Before the spatial Fourier transform, the gridded electric fields are 
+padded with zeros in order to match the grid size and angular size of each image 
+pixel that would have been obtained with software holography of output from an FX
 correlator.
@@
 
-\subsection{Time-averaging}
-
-In MOFF imaging, the measured antenna electric fields and the corresponding
-holographic electric field images are zero-mean stochastic quantities. Hence,
-they cannot be time-averaged to reduce noise. The statistical quantity stable
-with time in this case are the square of the holographic electric field images.
-Thus, squared images have to be formed at every instant of time before averaging
-as indicated in equation~\ref{eqn:dirty-image-MOFF}.
+\par\medskip
+\noindent {\bf Time-averaging}
+\par\medskip
+\noindent In MOFF imaging, the measured antenna electric fields and the 
+corresponding holographic electric field images are zero-mean stochastic 
+quantities. Hence, they cannot be time-averaged to reduce noise. The statistical 
+quantity stable with time in this case are the square of the holographic 
+electric field images. Thus, squared images have to be formed at every instant 
+of time before averaging as indicated in equation~\ref{eqn:dirty-image-MOFF}.
 
@@
 
+\subsection{Software Architecture}\label{sec:software-modules}
+
+EPIC is built using object oriented programming in Python and is built on
+carefully crafted modules which closely represent real-life entities in radio 
+interferometer arrays and observations. The essential modules along with their
+key attributes and methods are illustrated in Fig.~\ref{fig:software-modules}.
+These modules are described below.
+
+\begin{figure*}
+  \includegraphics[width=\linewidth]{EPIC-modules.eps}
+  \caption{Software architecture of EPIC with core modules, their essential
+    attributes and functions. The Antenna module forms the fundamental building 
+    block. It consists of electric field time-series and spectra and the 
+    F-engine that performs a temporal FFT to obtain electric field spectra from 
+    the time-series. The Interferometer module is made of a pair of Antenna 
+    modules. Its main function is the X-engine (FX or XF) to produce visibility 
+    spectra. The Antenna Array module is made of all individual Antenna modules
+    as its components and contains collective properties about the antenna 
+    subsystems. Its core function is the creation of antenna-to-grid mapping and 
+    that of gridded aperture weights and electric fields. The Interferometer 
+    Array module is very similar in principle to the Antenna Array module 
+    except it operates on cross-correlations and produces gridded visibilities.
+    The Image module takes gridded electric fields or visibilities and performs
+    a two-dimensional spatial FFT (and squares the intermediate image in case 
+    of the former) to produce output images.}
+  \label{fig:software-modules}
+\end{figure*}
+
+\subsubsection{Antenna module}
+
+The Antenna module is a fundamental building block upon which all the other 
+modules are built. There is one Antenna module per antenna each having 
+attributes -- the propagated electric field time-series, $E(t)$, and spectrum 
+$E(f)$ for both polarizations. The most important function inside this module 
+is the F-engine that Fourier transforms time-series electric field data into 
+spectrum.
+
+The other function (not shown in the figure) is to update the data as new data 
+keeps streaming in. This can also be parallelized. Another important attribute
+(not shown in figure) are antenna flags for each polarization appropriate for 
+the data stream being held by the Antenna module.
+
+\subsubsection{Interferometer module}
+
+The Interferometer module holds the attributes and functions pertaining to
+a pair of antennas and represents the cross-correlation information obtained
+from the pair. Its important attributes are the two Antenna modules. It also
+contains four cross-polarized visibility time-series and spectra.
+
+The most important function of the Interferometer module is the X-engine. 
+This is essentially a software analog of hardware correlators of real 
+telescope systems. The X-engine can be toggled between two states of operation,
+namely, the FX and XF modes. The FX mode obtains the electric field spectra, 
+$E(f)$ from the individual Antenna modules inside the Interferometer module and
+multiplies the two to obtain visibility spectra, $V(f)$. On the other hand, the
+XF mode cross-correlates the electric field time-series from its Antenna modules
+to obtain the visibilities as a function of lags, $V_t(t)$, which is then Fourier 
+transformed to obtain $V(f)$. Both modules can operate on dual-polarizations to
+obtain all four cross-polarizations.
+
+The other attributes (not shown in figure) are the flags applicable for each
+cross-polarization for the current data stream. Similar to the Antenna module,
+it has an update function that can update the visibilities $V_t(t)$ or $V(f)$
+directly rather than through the electric fields of its component antennas. 
+This functionality is to allow EPIC to operate while attached to the backend 
+of traditional correlator systems. Currently, this feature is not utilized 
+for purposes of this paper.
+
+The Interferometer module forms the fundamental unit for the InterferometerArray
+module (to be discussed below) and in general for visibility based correlator 
+and imaging systems.
+
+\subsubsection{Antenna Array module}
+
+The AntennaArray module consists of all the Antenna modules as its attributes
+and represents the collective properties of its component antennas. By virtue of 
+holding each antenna data independently in their respective modules the F-engine 
+for the entire array can be distributed to the F-engines of the component 
+antenna modules thus achieving a highly parallelized F-engine while emulating 
+real telescope systems.
+
+The most important attributes held by this module are the antenna aperture 
+illumination weights and electric fields projected on the grid using the 
+gridding convolution method described above and implemented by the gridding 
+function (denoted by grid()) in this module. Significant parts of the 
+antenna-to-grid mapping and gridding convolution are parallelizable 
+across antennas and frequencies.
+
+Individual antenna flags are carried over as additional weights to be applied
+to the gridded aperture illumination and electric fields. A series of data 
+streams can be stacked up to take advantage of the array optimization available 
+in Python. This module is also equipped to manage dual-polarization.
+
+\subsubsection{Interferometer Array module}
+
+Similar to the AntennaArray module, the InterferometerArray module consists of
+individual Interferometer modules. It can parallelize the correlator operations
+by distributing the X-operation over the X-engines of its component 
+Interferometer modules. The interferometer-to-grid mapping and gridding 
+convolution are very similar in nature to that of the Antenna Array module. 
+Flag-based grid weights, stacking and ability to handle all four 
+cross-polarizations are built into this module.
+
+\subsubsection{Image module}
+
+The Image module is built as a general purpose module that can switch between 
+operating on gridded electric fields or visibilities. At its heart, it consists
+of a two-dimensional FFT where the padding can be specified by the user to 
+control the resolution in the output images. In case of electric field based
+imaging, there is an additional step of squaring the holographic electric field
+images. This is packaged into the function annotated as imagr().
+
+Besides its core functions of spatial Fourier transform and squaring, it can 
+stack, accumulate and average images, and optionally remove the antenna
+auto-correlations centered around the zero-spacing pixel in the $uv$ plane. 
+It also handles all four cross-polarization products. Currently, it can 
+support writing data out in standard FITS format.
+
+\section{Verification}\label{sec:verify}
+
+In order to prove the accuracy of the EPIC code, we verify the images produced 
+through simulations. We simulate electric field streams from a model sky and 
+process the data through both the MOFF and visibility based imaging algorithms. 
+We then compare the output images to demonstrate their equivalence.
+
 \subsection{Dealing with antenna auto-correlations}\label{sec:rm-autocorr}
 
-The squaring operation under MOFF imaging in the image plane introduces antenna 
-auto-correlations around the zero spacing in the $uv$-plane which are absent in 
-traditional visibility-based imaging. In order to facilitate a robust comparison, 
-these auto-correlations are removed from both the $uv$ and image planes, which 
-is otherwise not an essential part of the core algorithm. We describe below how 
-they can be removed from the $uv$ plane in a straightforward manner.
+Before diving into the details of simulations and comparison with visibility 
+based imaging, we describe the elimination of a minor mathematical difference 
+between the two techniques. The squaring operation under MOFF imaging in the 
+image plane introduces antenna auto-correlations around the zero spacing in 
+the $uv$-plane which are absent in traditional visibility-based imaging. In 
+order to facilitate a robust comparison, these auto-correlations are removed 
+from both the $uv$ and image planes, which is otherwise not an essential part 
+of the core algorithm. We describe below how they can be removed from the $uv$ 
+plane in a straightforward manner.
 
@@
 
-We inverse Fourier transform the squared 
-images and beams back to the $uv$ plane and subtract the estimated 
-auto-correlation kernel scaled to the peak value centered at the zero spacing 
-pixel. The final averaged image is obtained by Fourier transforming the $uv$ 
-plane data and weights with the auto-correlations subtracted to the image plane. 
-These images are now comparable to those obtained from visibility-based imaging. 
-This step of removing auto-correlations is required to be performed only once 
-per integration timescale and does not add significant cost to the full operation.
-
-\section{Verification}\label{sec:verify}
-
-In order to prove the accuracy of the EPIC code, we verify the images produced 
-through simulations. We simulate electric field streams from a model sky and 
-process the data through both the MOFF and visibility based imaging algorithms. 
-We then compare the output images to demonstrate their equivalence.
+We inverse Fourier transform the squared images and beams back to the $uv$ 
+plane and subtract the estimated auto-correlation kernel scaled to the peak 
+value centered at the zero spacing pixel. The final averaged image is obtained 
+by Fourier transforming the $uv$ plane data and weights with the 
+auto-correlations subtracted to the image plane. These images are now comparable 
+to those obtained from visibility-based imaging. This step of removing 
+auto-correlations is optional and required to be performed only once per 
+integration timescale and does not add significant cost to the full operation.
 
@@
 In the visibility-based imaging, the predominant computational cost is at
-the X-engine requiring $N_\textrm{a}^2$ complex multiplications per channel 
-per correlator writeout timescale.
+the X-engine requiring $N_\textrm{a}(N_\textrm{a}-1)/2$ complex multiplications 
+per channel at every correlator writeout timescale.
 
@@
 the MOFF imaging is determined by the spatial Fourier transform while  
-that for visibility-based imaging comes from $N_\textrm{a}^2$ 
-cross-correlations. However, if non-linearities such 
-as non-coplanarity of baselines \citep{cor08} and wide-field phenomena like 
-the {\it pitchfork} effect \citep{thy15a,thy15b} are to be corrected for, 
-the antenna based illumination footprint can start becoming less compact
-in the measurement plane and can result in a costlier gridding process.
+that for visibility-based imaging comes from $N_\textrm{a}(N_\textrm{a}-1)/2$ 
+cross-correlations. However, if non-linearities such as non-coplanarity of 
+baselines \citep{cor08} and wide-field phenomena like the {\it pitchfork} 
+effect \citep{thy15a,thy15b} are to be corrected for, the antenna based 
+illumination footprint can start becoming less compact in the measurement 
+plane and can result in a costlier gridding process.
 
@@
 an expanded layout with a four-fold increase in number of elements over a
-two-fold increase in $b_\textrm{max}$ provide the bounds at the bottom and top 
+50\% increase in $b_\textrm{max}$ provide the bounds at the bottom and top 
 respectively. Current instruments such as MWA and LOFAR lie in parameter space
@@
     based imaging are equal. MOFF imaging is more efficient for telescopes 
-    occupying the left of this line and vice versa. All the HERA layouts except
-    possibly HERA-19 lie in the parameter space favoured by MOFF imaging. And so 
-    are SKA1-LC and SKA1-LCD. The solid black curve shows the projected 
-    trajectory of bigger close-packed hexagonal layouts similar to HERA. The 
-    gray shaded area denotes the projected trajectory of the LWA bounded by LWA1 
-    (left edge), LWA-OV (right edge), current layout (bottom) and a four-fold 
-    increase in the number of elements in a two-fold increase in the core size
-    (top). Current instruments such as MWA and LOFAR fall in a region favoured 
-    by visibility based imaging.}
+    occupying the left of this line and vice versa. CHIME, HIRAX and all the 
+    HERA layouts except HERA-19 and HERA-37 lie in the parameter space favoured 
+    by MOFF imaging. And so are SKA1-LC and SKA1-LCD. The solid black curve 
+    shows the projected trajectory of bigger close-packed hexagonal layouts 
+    similar to HERA. The gray shaded area denotes the projected trajectory of 
+    the LWA bounded by LWA1 (left edge), LWA-OV (right edge), current layout 
+    (bottom) and a four-fold increase in the number of elements within a 50\% 
+    increase in the core size (top). Current instruments such as MWA and LOFAR 
+    fall in a region favoured by visibility based imaging.}
   \label{fig:parameter-space-computations-instruments}

### 2015-10-26 imaging_demo/ms.tex
fixing more typos

@@
 
-We provide a brief summary of the amthematical equivalence of the MOFF and 
+We provide a brief summary of the mathematical equivalence of the MOFF and 
 FX correlators detailed in \citet{mor11}. We first relate the dirty image produced 
@@
 pair of antennas ($p$) is obtained by the multiplying the sky brightness
-$I(\hat{\mathbf{s}})$ by the antenna power response $B(\hat{\mathbf{s}})$ and
+$I(\hat{\mathbf{s}})$ by the antenna power response $B_p(\hat{\mathbf{s}})$ and
 Fourier transforming from the directional coordinates ($\hat{\mathbf{s}}$) to $uv$
@@
 \begin{align}
-  \mathbf{m}(\mathbf{v}) &= \left\langle \widetilde{\mathbf{E}}^\star(\mathbf{a})\,\widetilde{\mathbf{E}}(\mathbf{a}^\prime)\right\rangle_t, \label{eqn:matrix-cc-vis}
+  \mathbf{m}(\mathbf{v}) &= \left\langle \widetilde{\mathbf{E}}(\mathbf{a})\,\widetilde{\mathbf{E}}^\star(\mathbf{a}^\prime)\right\rangle_t, \label{eqn:matrix-cc-vis}
 \end{align}
@@
 We have implemented the MOFF imaging technique in our ``E-field Parallel Imaging
-Correlator'' -- a highly parallelized Object Oriented Python package,
-\footnote{EPIC package can be accessed at https://github.com/nithyanandan/EPIC}
+Correlator'' -- a highly parallelized Object Oriented Python 
+package,\footnote{EPIC package can be accessed at https://github.com/nithyanandan/EPIC}
 now publicly available. Besides implementing the MOFF imaging algorithm it also
@@
 correlators to estimate antenna complex gains. This method correlates the antenna 
-electric field signals with an image pixel form the output of the correlator in 
+electric field signals with an image pixel from the output of the correlator in 
 the feedback calibration fashion outlined in \citealt{mor11} (illustrated in Fig. 
@@
 non-identical antenna beam patterns. Because only a single correlation is needed 
-for each antenna, the computation complexity scales only as $N_{\mathrm{ant}}$.
+for each antenna, the computation complexity scales only as $N_{\mathrm{a}}$.
 
@@
 is the F-engine that Fourier transforms time-series electric field data into 
-spectrum.
+spectra.
 
@@
 $\approx 80$\% of bandwidth (roughly 80~kHz). The image is shown in direction 
-cosine coordinates -- $l$ along x-axis and $m$ along y-axis. The flux scale 
+cosine coordinates -- $l$ along the horizontal axis and $m$ along the vertical axis. The flux scale 
 is arbitrary. Even in this proof-of-concept demonstration, we see Cyg~A and 

### 2015-10-27 imaging_demo/ms.tex
Add new material in section 5, update figures
From 88cd7fb (master):
=====================

In ms.tex,

(*) Restructure section 5 into two subsections, one of which is a new
subsection on data throughput with the two algorithms
(*) Change figure 12
(*) Update figure 10
(*) Rearrange subsections in section 4
(*) Update figure 5
(*) Incorporate other changes suggested by Judd and Adam
(*) Minor changes to abstract, introduction and conclusions

@@
 %\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
-\newcommand{\Nant}{N_{\text{ant}}} 
+\newcommand{\Nant}{N_\textrm{a}}
+\newcommand{\Ngrid}{N_\textrm{g}} 
 \newcommand{\Npix}{N_{\text{pix}}}
@@
 consisting of a large number of antennas 
-($N_\textrm{a}\gtrsim 1000$--$\gtrsim 10000$). Since the cost of traditional 
-correlators scales as $\mathcal{O}(N_\textrm{a}^2)$, there will be a steep cost for 
+($\Nant\gtrsim 1000$--$\gtrsim 10000$). Since the complexity of traditional 
+correlators scales as $\mathcal{O}(\Nant^2)$, there will be a steep cost for 
 realizing the full imaging potential of these powerful instruments. Through 
@@
 advantage of the multiplication-convolution theorem of Fourier transforms. Not 
-only does it bring down the cost to $\mathcal{O}(N_\textrm{a}\log N_\textrm{a})$ 
+only does it bring down the cost to $\mathcal{O}(\Nant\log \Nant)$ 
 but can also image from irregularly arranged heterogeneous antenna arrays. 
@@
 deploying EPIC. Inherent availability of calibrated time-domain images on 
-timescales roughly equal to the writeout timescale of the digitizer will make 
-it a prime candidate for transient searches of primarily Fast Radio Bursts (FRB) 
-as well as slower planteary and exoplantenary phenomena. 
+timescales roughly equal to the writeout timescale of the digitizer and vastly 
+lower bandwidth relative to visibility based systems will make it a prime 
+candidate for transient searches of Fast Radio Bursts (FRB) as well as 
+planetary and exoplantenary phenomena. 
 \end{abstract}
@@
 and regular monitoring of the time-dependent sky is driving instruments in the
-same direction. A number of new telescopes are under development around the world
-based on this new paradigm, including the Murchison Widefield Array (MWA;
-\citealt{tin13,bow13}), the Precision Array for Probing the Epoch of Reionization
-(PAPER; \citealt{par10}), the Hydrogen Epoch of Reionization 
-Array\footnote{http://reionization.org} (HERA), the LOw Frequency ARray (LOFAR;
-\citealt{van13}), the Canadian Hydrogen Intensity Mapping Experiment (CHIME,
-\citealt{ban14}), the Long Wavelength Array (LWA, \citealt{ell13}), and the low
-frequency component of the Square Kilometer Array Low Frequency Aperture Array
-(SKA-Low \citealt{mel13}).
+same direction with the added requirement of fast read-outs. A number of new 
+telescopes are under development around the world based on this new paradigm, 
+including the Murchison Widefield Array (MWA;\citealt{tin13,bow13}), the 
+Precision Array for Probing the Epoch of Reionization (PAPER; \citealt{par10}), 
+the Hydrogen Epoch of Reionization Array\footnote{http://reionization.org} 
+(HERA), the LOw Frequency ARray (LOFAR;\citealt{van13}), the Canadian Hydrogen 
+Intensity Mapping Experiment (CHIME,\citealt{ban14}), the Long Wavelength 
+Array (LWA, \citealt{ell13}), and the low frequency component of the Square 
+Kilometer Array Low Frequency Aperture Array (SKA1-Low \citealt{mel13}).
 
@@
 computational scaling from the harsh $\Nant^2$ to the more gentle envelope of
-$\Npix\log\Npix$, where $\Npix$ is the number of pixels in the Fourier transform
-\citep[e.g.][]{mor11,teg09,teg10}. This architecture is often referred to as a
-``direct imaging" correlator because it eliminates the intermediary cross-
-correlation data products of the FX and earlier lag correlators, but instead
-directly forms images from the electric field measurements.
+$\Ngrid\log\Ngrid$, where $\Ngrid$ is the number of grid pixels in the Fourier 
+transform \citep[e.g.][]{mor11,teg09,teg10}. This architecture is often referred 
+to as a ``direct imaging" correlator because it eliminates the intermediary 
+cross-correlation data products of the FX and earlier lag correlators, but 
+instead directly forms images from the electric field measurements.
 
@@
 elements are identical. On the other hand, the MOFF algorithm achieves the same
-$\Npix \log \Npix$ computational scaling without placing any restriction on
+$\Ngrid \log \Ngrid$ computational scaling without placing any restriction on
 antenna placement, can accommodate non-identical beam patterns, and is 
@@
 announce the public release of the E-field Parallel Imaging Correlator (EPIC)
-code. We begin with a technical description of the algorithm in \S
-\ref{sec:math}, then discuss our particular implementation in \S
-\ref{sec:software}. We then verify the output data quality from our code in \S
-\ref{sec:verify} by presenting simulated images from both the EPIC correlator
-and comparing to a simulated FX correlator. We also demonstrate the performance
-with real-world data from the LWA. In \S \ref{sec:analysis} we analyze the
-scaling relationships of the algorithm. We identify specific array design classes
-where the EPIC correlator is computationally more efficient than the FX
-algorithm. We conclude and discuss future research prospects in \S
-\ref{sec:conclusions}.
-
-% Motivate MOFF from a technical and scientific standpoint. Refer to \citet{del07,mor09,mor11}.
+code. We begin with a technical description of the algorithm in \S\ref{sec:math}, 
+then discuss our particular implementation in \S\ref{sec:software}. We then 
+verify the output data quality from our code in \S\ref{sec:verify} by presenting 
+simulated images from both the EPIC correlator and comparing to a simulated FX 
+correlator. We also demonstrate the performance with real-world data from the 
+LWA. In \S \ref{sec:analysis} we analyze the scaling relationships of the 
+algorithm. We identify specific array design classes where the EPIC correlator 
+is computationally more efficient and in the field of transients, demands 
+significantly lesser bandwidth relative to visibility based approaches. We 
+conclude and discuss future research prospects in \S\ref{sec:conclusions}.
 
@@
 
-From here we adopt the matrix notation of \citet{mor11}, where vectors are 
+Hereafter, we adopt the matrix notation of \citet{mor11}, where vectors are 
 represented with single coordinates, and matrices are represented by two 
@@
 \item The time-averaging cannot be performed on a stochastic measurement but
-  only on its statistical properties. In FX imaging, the visibilities measured
-  between antenna pairs represent spatial correlations which can be time-averaged
-  followed by gridding and imaging. However, in MOFF imaging both antenna and 
-  gridded electric fields are stochastic and therefore must be imaged and squared 
-  before time-averaging. 
-\item In FX imaging, electric fields measured by antennas are not correlated with
-  themselves and hence lack zero spacing measurements. In contrast, in MOFF
-  imaging, since the gridded electric fields are imaged and squared, they retain 
-  information from auto-correlated electric fields at zero spacing and thus yield 
-  the true total power of the imaged field.
+  only on its statistical properties. In visibility based imaging, the 
+  visibilities measured between antenna pairs represent spatial correlations 
+  which can be time-averaged followed by gridding and imaging. However, in MOFF 
+  imaging both antenna and gridded electric fields are stochastic and therefore 
+  must be imaged and squared before time-averaging. 
+\item In visibility based imaging, electric fields measured by antennas are not 
+  correlated with themselves and hence lack zero spacing measurements. In 
+  contrast, in MOFF imaging, since the gridded electric fields are imaged and 
+  squared, they retain information from auto-correlated electric fields at zero 
+  spacing and thus yield the true total power of the imaged field.
 \end{enumerate}
@@
 now publicly available. Besides implementing the MOFF imaging algorithm it also
-includes FX imaging using the software holography technique and a simulator for
-generating electric fields from a sky model.
+includes visibility based imaging using the software holography technique and a 
+simulator for generating electric fields from a sky model.
 
@@
 
-Fig.~\ref{fig:FX-flowchart} shows the flowchart for software holographic
-imaging from a FX correlator. The antenna-based F-engine is identical to that in
-the MOFF processing. The electric field spectra from each antenna are then cross-
-multiplied in the X-engine with those from all other antennas to obtain the
-visibilities $V_\textrm{ij}(f)$ ($\mathbf{m}(\mathbf{v})$ in matrix notation).
-They are calibrated and time-averaged to obtain $\langle V_\textrm{ij}(f)\rangle$
-which are then gridded and imaged to obtain the image $I(f)$. The $I(f)$ obtained
-from both techniques are identical as explained in \S\ref{sec:math}.
+Fig.~\ref{fig:FX-flowchart} shows the flowchart for a visibility based software 
+holographic imaging from a FX correlator. The antenna-based F-engine is identical 
+to that in the MOFF processing. The electric field spectra from each antenna are 
+then cross-multiplied in the X-engine with those from all other antennas to 
+obtain the visibilities $V_\textrm{ij}(f)$ ($\mathbf{m}(\mathbf{v})$ in matrix 
+notation). They are calibrated and time-averaged to obtain 
+$\langle V_\textrm{ij}(f)\rangle$ which are then gridded and imaged to obtain the 
+image $I(f)$. The $I(f)$ obtained from both techniques are identical as explained 
+in \S\ref{sec:math}.
 \begin{figure}
   \includegraphics[width=\columnwidth]{FX_flowchart.eps}
-  \caption{A flowchart of FX imaging in EPIC. The FX process flow shares the
-    F-engine with the MOFF process. Following the F-engine, the electric fields
-    pass through the X-engine to obtain visibilities $V_\textrm{ij}(f)$ which are
-    calibrated and time-averaged. Then they are gridded to obtain the gridded
-    visibilities $V_\textrm{g}(f)$ which are then imaged to obtain the image
+  \caption{A flowchart of visibility based software holographic imaging in EPIC. 
+    The FX process flow shares the F-engine with the MOFF process. Following the 
+    F-engine, the electric fields pass through the X-engine to obtain 
+    visibilities $V_\textrm{ij}(f)$ which are calibrated and time-averaged. Then 
+    they are gridded to obtain the gridded visibilities $V_\textrm{g}(f)$ which 
+    are then Fourier transformed to obtain the image.
     $I(f)$.}
@@
 \par\medskip
-\noindent This module is common to the MOFF and FX imaging techniques. Time 
-samples of electric fields measured by the antenna and digitized by the A/D 
-converter is Fourier transformed to generate electric field spectra. This step 
-can be parallelized by antennas as shown in Fig.~\ref{fig:f-engine}. The output 
-is then fed to either MOFF and FX imaging pipelines.
+\noindent This module is common to the MOFF and visibility based imaging 
+techniques. Time samples of electric fields measured by the antenna and 
+digitized by the A/D converter is Fourier transformed to generate electric 
+field spectra. This step can be parallelized by antennas as shown in 
+Fig.~\ref{fig:f-engine}. The output is then fed to either MOFF or visibility 
+based imaging pipelines.
 
@@
 aperture, $\mathbf{M}(\mathbf{r},\mathbf{a})$, which is generally of size
-$N_\textrm{grid}\times N_\textrm{a}$, reduces to a sparse block-diagonal matrix
-with only $N_\textrm{a}$ blocks and roughly $N_\textrm{k}$ non-zero entries per
+$\Ngrid\times \Nant$, reduces to a sparse block-diagonal matrix
+with only $\Nant$ blocks and roughly $N_\textrm{k}$ non-zero entries per
 block. This sparse matrix is stored in a Compressed Sparse Row (CSR) format. 
@@
   \caption{Block diagram of an antenna-to-grid mapping. A sparse block-diagonal
-    matrix of total size $N_\textrm{grid}\times N_\textrm{a}$ is created where each
+    matrix of total size $\Ngrid\times \Nant$ is created where each
     block contains roughly the number of pixels covered by the respective kernel.
@@
 In MOFF imaging, these are spatially Fourier transformed followed by a squaring
-operation at every timestamp for every frequency channel. In FX imaging, the
-spatial Fourier transform is performed only once per integration timescale and
-does not include a squaring operation.
+operation at every timestamp for every frequency channel. In visibility based 
+imaging, the spatial Fourier transform is performed only once per integration 
+timescale and does not include a squaring operation.
 
@@
 timescale, the imaging step has to be performed only once per integration cycle.
-% FX imaging holds this advantage as long as the square of the number of antennas 
-% is smaller than the number of pixels on the image.
 
@@
   \caption{Software architecture of EPIC with core modules, their essential
-    attributes and functions. The Antenna module forms the fundamental building 
+    attributes and functions. The antenna module forms the fundamental building 
     block. It consists of electric field time-series and spectra and the 
     F-engine that performs a temporal FFT to obtain electric field spectra from 
-    the time-series. The Interferometer module is made of a pair of Antenna 
+    the time-series. The interferometer module is made of a pair of antenna 
     modules. Its main function is the X-engine (FX or XF) to produce visibility 
-    spectra. The Antenna Array module is made of all individual Antenna modules
+    spectra. The antenna array module is made of all individual antenna modules
     as its components and contains collective properties about the antenna 
     subsystems. Its core function is the creation of antenna-to-grid mapping and 
-    that of gridded aperture weights and electric fields. The Interferometer 
-    Array module is very similar in principle to the Antenna Array module 
+    that of gridded aperture weights and electric fields. The interferometer 
+    array module is very similar in principle to the antenna array module 
     except it operates on cross-correlations and produces gridded visibilities.
-    The Image module takes gridded electric fields or visibilities and performs
+    The image module takes gridded electric fields or visibilities and performs
     a two-dimensional spatial FFT (and squares the intermediate image in case 
-    of the former) to produce output images.}
+    of the former) to produce output images. Broadly, the MOFF algorithm is 
+    implemented by modules below the horizontal dashed line while the visibility 
+    based imaging uses modules above the line. The exact processing pathway 
+    implementing the MOFF algorithm is shown in bolded modules.}
   \label{fig:software-modules}
@@
 
-\subsubsection{Antenna module}
-
-The Antenna module is a fundamental building block upon which all the other 
-modules are built. There is one Antenna module per antenna each having 
+\subsubsection{Antenna Module}
+
+The antenna module is a fundamental building block upon which all the other 
+modules are built. There is one antenna module per antenna each having 
 attributes -- the propagated electric field time-series, $E(t)$, and spectrum 
@@
 (not shown in figure) are antenna flags for each polarization appropriate for 
-the data stream being held by the Antenna module.
-
-\subsubsection{Interferometer module}
-
-The Interferometer module holds the attributes and functions pertaining to
+the data stream being held by the module.
+
+\subsubsection{Interferometer Module}
+
+The interferometer module holds the attributes and functions pertaining to
 a pair of antennas and represents the cross-correlation information obtained
-from the pair. Its important attributes are the two Antenna modules. It also
+from the pair. Its primary attributes are the two antenna modules. It also
 contains four cross-polarized visibility time-series and spectra.
 
-The most important function of the Interferometer module is the X-engine. 
-This is essentially a software analog of hardware correlators of real 
-telescope systems. The X-engine can be toggled between two states of operation,
+The critical component of the interferometer module is the X-engine. This is 
+essentially a software analog of hardware correlators of real telescope 
+systems. The X-engine can be toggled between two states of operation, 
 namely, the FX and XF modes. The FX mode obtains the electric field spectra, 
-$E(f)$ from the individual Antenna modules inside the Interferometer module and
-multiplies the two to obtain visibility spectra, $V(f)$. On the other hand, the
-XF mode cross-correlates the electric field time-series from its Antenna modules
+$E(f)$ from the individual antenna modules inside this module and multiplies 
+the two to obtain visibility spectra, $V(f)$. On the other hand, the XF mode 
+cross-correlates the electric field time-series from its Antenna modules
 to obtain the visibilities as a function of lags, $V_t(t)$, which is then Fourier 
@@
 The other attributes (not shown in figure) are the flags applicable for each
-cross-polarization for the current data stream. Similar to the Antenna module,
+cross-polarization for the current data stream. Similar to the antenna module,
 it has an update function that can update the visibilities $V_t(t)$ or $V(f)$
@@
 This functionality is to allow EPIC to operate while attached to the backend 
-of traditional correlator systems. Currently, this feature is not utilized 
-for purposes of this paper.
-
-The Interferometer module forms the fundamental unit for the InterferometerArray
-module (to be discussed below) and in general for visibility based correlator 
-and imaging systems.
-
-\subsubsection{Antenna Array module}
-
-The AntennaArray module consists of all the Antenna modules as its attributes
+of traditional correlator systems. This feature is not utilized for purposes 
+of this paper.
+
+This module forms the fundamental unit for the interferometer array module 
+(to be discussed below) and in general for visibility based correlator and 
+imaging systems.
+
+\subsubsection{Antenna Array Module}
+
+The antenna array module consists of all the antenna modules as its attributes
 and represents the collective properties of its component antennas. By virtue of 
@@
 
-The most important attributes held by this module are the antenna aperture 
+The primary attributes held by this module are the antenna aperture 
 illumination weights and electric fields projected on the grid using the 
 gridding convolution method described above and implemented by the gridding 
-function (denoted by grid()) in this module. Significant parts of the 
-antenna-to-grid mapping and gridding convolution are parallelizable 
-across antennas and frequencies.
+function in this module. Significant parts of the antenna-to-grid mapping and 
+gridding convolution are parallelizable across antennas and frequencies.
 
@@
 
-\subsubsection{Interferometer Array module}
-
-Similar to the AntennaArray module, the InterferometerArray module consists of
-individual Interferometer modules. It can parallelize the correlator operations
+\subsubsection{Interferometer Array Module}
+
+Similar to the antenna array module, the interferometer array module consists of
+individual interferometer modules. It can parallelize the correlator operations
 by distributing the X-operation over the X-engines of its component 
-Interferometer modules. The interferometer-to-grid mapping and gridding 
-convolution are very similar in nature to that of the Antenna Array module. 
+interferometer modules. The interferometer-to-grid mapping and gridding 
+convolution are very similar in nature to that of the antenna array module. 
 Flag-based grid weights, stacking and ability to handle all four 
@@
 
-\subsubsection{Image module}
-
-The Image module is built as a general purpose module that can switch between 
+\subsubsection{Image Module}
+
+The image module is built as a general purpose module that can switch between 
 operating on gridded electric fields or visibilities. At its heart, it consists
-of a two-dimensional FFT where the padding can be specified by the user to 
-control the resolution in the output images. In case of electric field based
-imaging, there is an additional step of squaring the holographic electric field
-images. This is packaged into the function annotated as imagr().
+of a two-dimensional spatial FFT where the padding can be specified by the user 
+to control the resolution in the output images. In case of MOFF imaging, there 
+is an additional step of squaring the holographic electric field images.
 
@@
 
-In order to prove the accuracy of the EPIC code, we verify the images produced 
-through simulations. We simulate electric field streams from a model sky and 
-process the data through both the MOFF and visibility based imaging algorithms. 
-We then compare the output images to demonstrate their equivalence.
+In order to verify the accuracy of the EPIC code, we characterize the images 
+produced through simulations. We simulate electric field streams from a model 
+sky and process the data through both the MOFF and visibility based imaging 
+algorithms. We then compare the output images to demonstrate their equivalence.
+
+\subsection{Simulations}\label{sec:sim}
+
+We use the EPIC simulator to generate electric field samples from a sky model. In
+our simulations, we use 64 frequency channels each of width $\delta f = 40$~kHz, 
+10 point sources of flux densities 10~Jy at random locations. The number of
+timestamps integrated in one integration cycle was kept at eight where each A/D 
+timeseries is $1/\delta f=25\,\mu$s long. We use the MWA array layout 
+\citep{bea12} for demonstration. Only the inner 51 tiles within a square bounding
+box of 150~m on each side were used. We assumed all tiles are identical and have 
+a square shaped electric field illumination footprint 4.4~m on each side.
 
@@
 
-Before diving into the details of simulations and comparison with visibility 
-based imaging, we describe the elimination of a minor mathematical difference 
-between the two techniques. The squaring operation under MOFF imaging in the 
-image plane introduces antenna auto-correlations around the zero spacing in 
-the $uv$-plane which are absent in traditional visibility-based imaging. In 
-order to facilitate a robust comparison, these auto-correlations are removed 
-from both the $uv$ and image planes, which is otherwise not an essential part 
-of the core algorithm. We describe below how they can be removed from the $uv$ 
-plane in a straightforward manner.
+Before the outputs can be compared, we describe the elimination of a minor 
+mathematical difference between the two techniques. The squaring operation under 
+MOFF imaging in the image plane introduces antenna auto-correlations around the 
+zero spacing in the $uv$-plane which are absent in traditional visibility based 
+imaging. In order to facilitate a robust comparison between MOFF and visibility 
+based imaging techniques, these auto-correlations are removed from both the MOFF 
+algorithm output, which is otherwise not an essential part of the core algorithm. 
+We describe below how they are removed.
 
@@
 auto-correlations subtracted to the image plane. These images are now comparable 
-to those obtained from visibility-based imaging. This step of removing 
+to those obtained from visibility based imaging. This step of removing 
 auto-correlations is optional and required to be performed only once per 
@@
 
-\subsection{Simulations}\label{sec:sim}
-
-We use the EPIC simulator to generate electric field samples from a sky model. In
-our simulations, we use 64 frequency channels each of width $\delta f = 40$~kHz, 
-10 point sources of flux densities 10~Jy at random locations. The number of
-timestamps integrated in one integration cycle was kept at eight where each A/D 
-timeseries is $1/\delta f=25\,\mu$s long. We use the MWA array layout 
-\citep{bea12} for demonstration. Only the inner 51 tiles within a square bounding
-box of 150~m on each side were used. We assumed all tiles are identical and have 
-a square shaped electric field illumination footprint 4.4~m on each side.
+\subsection{Comparison of outputs}\label{sec:diff}
+
+We investigate the two imaging algorithms for differences from the point of 
+view of the quality of their outputs. We begin by comparing the gridded 
+cross-correlation weights in the $uv$-plane. In MOFF imaging, weights 
+from antenna auto-correlations have been removed as described in 
+\S\ref{sec:rm-autocorr}.
 
 Fig.~\ref{fig:MOFF-FX-image} shows the dirty images (top) and synthesized beams
-(bottom) obtained with antenna-based MOFF and FX visibility-based imaging 
+(bottom) obtained with antenna-based MOFF and FX visibility based imaging 
 algorithms packaged in EPIC. The antenna auto-correlations that correspond to 
@@
 
-\subsection{Comparison of outputs}\label{sec:diff}
-
-We investigate the two imaging algorithms for differences from the point of 
-view of the quality of their outputs. We begin by comparing the gridded 
-cross-correlation weights in the $uv$-plane. In MOFF imaging, weights 
-from antenna auto-correlations have been removed as described in 
-\S\ref{sec:rm-autocorr}.
-
 Fig.~\ref{fig:MOFF-FX-uvwts} shows the cross-correlation weights obtained 
-with MOFF imaging (left) and visibility-based imaging (right). The first 
+with MOFF imaging (left) and visibility based imaging (right). The first 
 notable difference is in the weights around zero spacing. Though both show a
@@
 because with MOFF imaging the antenna locations were rounded to the nearest 
-grid pixel whereas in visibility-based imaging the baseline locations 
+grid pixel whereas in visibility based imaging the baseline locations 
 (difference between antenna locations) were rounded to the nearest grid pixel. 
@@
 $\approx 80$\% of bandwidth (roughly 80~kHz). The image is shown in direction 
-cosine coordinates -- $l$ along the horizontal axis and $m$ along the vertical axis. The flux scale 
-is arbitrary. Even in this proof-of-concept demonstration, we see Cyg~A and 
-Cas~A prominently and Sgr~A faintly as annotated, thus validating the 
+cosine coordinates -- $l$ along the horizontal axis and $m$ along the vertical 
+axis. The flux scale is arbitrary. Even in this proof-of-concept demonstration, 
+we see Cyg~A and Cas~A prominently as annotated, thus validating the 
 functionality of the EPIC package.
@@
     are compensated for their respective delays. The flux scale is arbitrary
-    and not calibrated. Locations of Cyg~A, Cas~A, Sgr~A, Her~A and Jupiter
-    are annotated.}
+    and not calibrated. Locations of Cyg~A and Cas~A are annotated.}
   \label{fig:LWA-image}
@@
 and future radio telescopes.
+
+\subsection{Processing Volumes}
 
@@
 the dense array layout scenario that makes these stages perform the slowest, the 
-gridding convolution, squaring and time-averaging take up only $\simeq 40$\%, 
-$\simeq 40$\% and $\simeq 15$\% respectively of the time taken by the spatial 
-Fourier transform. With sparser arrays the gridding process will be be even 
-faster.
-
-In the visibility-based imaging, the predominant computational cost is at
-the X-engine requiring $N_\textrm{a}(N_\textrm{a}-1)/2$ complex multiplications 
+gridding convolution, squaring and time-averaging take up only $\lesssim 20$\%, 
+$\lesssim 20$\% and $\lesssim 5$\% respectively of the total time taken by all 
+these processes while the spatial Fourier transform takes up $\gtrsim 55\%$ of
+the total time. With sparser arrays the gridding process will be be even faster.
+
+In the visibility based imaging, the predominant computational cost is at
+the X-engine requiring $\Nant(\Nant-1)/2$ complex multiplications 
 per channel at every correlator writeout timescale.
@@
 the MOFF imaging is determined by the spatial Fourier transform while  
-that for visibility-based imaging comes from $N_\textrm{a}(N_\textrm{a}-1)/2$ 
+that for visibility based imaging comes from $\Nant(\Nant-1)/2$ 
 cross-correlations. However, if non-linearities such as non-coplanarity of 
@@
 transform implemented via Fast Fourier Transform \citep[FFT;][]{coo65} is 
-$\approx \beta N_\textrm{g}\log N_\textrm{g}$ where $N_\textrm{g}$ 
+$\approx \beta \Ngrid\log \Ngrid$ where $\Ngrid$ 
 is the number of pixels on the grid and $\beta$ is a constant that depends on 
 the implementation of twiddle FFT algorithms \citep{bri74}. In our study, we 
-set $\beta=5$, a value much more conservative than was indicated in 
-\citet{mor11}. We set the number of complex multiplications in the X-engine 
-in visibility based imaging to $N_\textrm{a}^2$.
+set $\beta=5$, a value\footnote{http://www.fftw.org/speed/method.html} much 
+more conservative than was indicated in \citet{mor11}. We set the number of 
+complex multiplications in the X-engine in visibility based imaging to 
+$\Nant^2$.
 
@@
 of the antenna, $A_\textrm{a}$. Hence, 
-$N_\textrm{g}\simeq b_\textrm{max}^2/A_\textrm{a}$.
+$\Ngrid\simeq b_\textrm{max}^2/A_\textrm{a}$.
 
@@
               & size & of Antennas & size & \\
-              & $b_\textrm{max}$ (in m) & $N_\textrm{a}$ & $A_\textrm{a}$ (in m$^2$) & $f_0$ (in MHz) \\
+              & $b_\textrm{max}$ (in m) & $\Nant$ & $A_\textrm{a}$ (in m$^2$) & $f_0$ (in MHz) \\
     \hline
@@
 
-Fig.~\ref{fig:parameter-space-bll-nant-instruments} shows the ratio of the
-number of computations required with visibility based imaging relative to 
-MOFF imaging and the boundaries where this ratio is unity. The color scale 
-refers to the ratio while using a 1~m antenna element. Pairs of black and
-white lines of the same line style correspond to a specific antenna size. For 
-a given antenna size (annotated on the plot), the white line denotes the 
-minimum baseline length that could achieve the closest packing as a function 
-of the number of antennas. Regions to the right of the white line for a 
-given antenna size imply a physically impossible scenario where antennas will 
-have to be packed overlapping with each other. For the same antenna size, the 
-black line denotes the boundary where the ratio of the number of computations 
-with either algorithm becomes unity. Regions to the right of the black lines 
-favour usage of the MOFF algorithm. Thus the wedge between the white and black 
-lines denotes the region in parameter space where MOFF algorithm holds a 
-definite advantage while visibility based algorithms will be favoured to the 
-left of the black lines. As antenna size increases the maximum number of 
-antennas for a dense packing as a function of baseline length decreases. Hence 
-the white lines shift leftward as antenna size increases. Similarly, with 
-increase in antenna size, $N_\textrm{g}$ also decreases when field of view 
-imaging is achieved with an increasing grid spacing equal to antenna size and
-hence lowers the amount of computations required with the MOFF algorithm. This
-shifts the black curves leftward. HERA-331, HERA-1027, LWA1, SKA1-LC and 
-SKA1-LCD are inside their respective wedges indicating they are in regions that
-favour the usage of MOFF imaging. For fixed baseline length, regions favouring
-the MOFF algorithm tend to be towards large $N_\textrm{a}$ indicating large-N
-dense array layouts with smaller antenna elements are best suited for deploying 
-EPIC.
+Fig.~\ref{fig:parameter-space-bll-nant-instruments} shows the boundaries where 
+the ratio of the number of computations required with visibility based imaging 
+relative to MOFF imaging is unity. The different colored lines correspond to 
+different antenna sizes (cyan - 1~m$^2$, blue - 7~m$^2$, purple - 16~m$^2$, 
+green - 28~m$^2$, orange - 150~m$^2$, red - 740~m$^2$, gray - 5900~m$^2$). Solid 
+line style in each color denotes the maximum number of antennas with the 
+corresponding antenna size that can be densely packed inside various baseline 
+lengths. The region to the right of the solid lines for corresponding antenna 
+size represents a scenario of overlapping antennas that is physically 
+impossible. Dashed lines of each color denote the boundary to the left of which 
+visibility based imaging is favoured for the corresponding antenna size. Region 
+inside the wedge enclosed by the solid and dashed lines favours using the MOFF 
+algorithm for the corresponding antenna size over visibility based imaging. 
+As antenna size increases the maximum number of antennas for a dense packing as 
+a function of baseline length decreases. Hence, the solid lines shift leftward 
+as antenna size increases. Similarly, with increase in antenna size, $\Ngrid$ 
+also decreases when field of view imaging is achieved with an increasing grid 
+spacing equal to antenna size and hence lowers the amount of computations 
+required with the MOFF algorithm. This shifts the dashed curves leftward. 
+The different antennas are color coded by roughly the class of antenna size 
+they fall into. Thus symbols of one color falling into a wedge of the same color 
+indicates MOFF imaging is favoured for those telescopes and vice versa. For 
+e.g., MOFF imaging is favoured in HERA-33 and HERA-6769 because they are inside 
+the wedge but not so in cases of HERA-19 and HERA-37. A majority of the 
+next-generation radio telescopes, namely, HERA-331 and its future expanded 
+versions, SKA1-LC, SKA1-LCD, HIRAX, CHIME, LWA1 and its future expansions 
+(LWA1-x2x1 and LWA1-x4x1.5 denote two-fold and four-fold increase in the number 
+of antennas within a core diameter that is 1 and 1.5 times the current size of 
+100~m respectively), and a couple of the expanded versions of LWA-OV (LWA-OVx2x1 
+and LWA-OVx4x1.5 denote two-fold and four-fold increase in the number of antennas
+within a core diameter that is 1 and 1.5 times the current size of 200~m 
+respectively) will fall in the regime where MOFF imaging will be desirable. For 
+a fixed baseline length, regions favouring the MOFF algorithm tend to be towards 
+large $\Nant$ indicating large-N dense array layouts with smaller antenna 
+elements are best suited for deploying EPIC.
 
 \begin{figure}
-  \includegraphics[width=\columnwidth]{MOFF_FX_crossover_baseline_n-antennas_rho_fov_gridding.eps}
-  \caption{Current and instruments planned for future in parameter space of
-    baseline length and number of antennas with MOFF and FX. Color scale shown 
-    is for the ratio of number of computations required by visibility based
-    imaging to that by the MOFF algorithm. Different line styles denote 
-    different antenna sizes annotated. For a given antenna size, the 
-    corresponding white line denotes the maximum number of antennas that can 
+  \includegraphics[width=\columnwidth]{MOFF_FX_crossover_baseline_n-antennas_rho_fov_gridding_legended.eps}
+  \caption{Current and planned instruments in parameter space of
+    baseline length and number of antennas with MOFF and FX. Line styles of 
+    different colors denote different classes of antenna sizes (cyan - 1~m$^2$, 
+    blue - 7~m$^2$, purple - 16~m$^2$, green - 28~m$^2$, orange - 150~m$^2$, 
+    red - 740~m$^2$, gray - 5900~m$^2$). Solid line style in each color denotes 
+    the maximum number of antennas with the corresponding antenna size that can 
     be packed inside various baseline lengths. The region to the right of the 
-    white lines for corresponding antenna size is physically disallowed. 
-    Regions to the left of the black lines of the same corresponding antenna 
-    size (or line style) favours usage of visibility based imaging. Region 
-    inside the wedge enclosed by the black and white lines favours using the 
-    MOFF algorithm for any given antenna size. These wedges shift leftward with 
-    increasing antenna size. HERA-331, HERA-1027, LWA1, SKA1-LC and SKA1-LCD
-    are inside their corresponding wedges that favour the usage of MOFF 
-    imaging.}
+    solid lines for corresponding antenna size is physically disallowed. Dashed
+    lines of each color denote the boundary to the left of which visibility
+    based imaging is favoured for the corresponding antenna size. Region inside 
+    the wedge enclosed by the solid and dashed lines favours using the MOFF 
+    algorithm for the corresponding antenna size. These wedges shift leftward 
+    with increasing antenna size. The different antennas are color coded by 
+    roughly the class of antenna size they fall into. Thus symbols of one 
+    color falling into a wedge of the same color indicates MOFF imaging is 
+    advantageous for those telescopes and vice versa. For e.g., MOFF imaging
+    is favoured in HERA-33 and HERA-6769 because they are inside the wedge 
+    but not so in cases of HERA-19 and HERA-37.}
   \label{fig:parameter-space-bll-nant-instruments}
 \end{figure}
+
+\subsection{Data Throughput}
+
+We elaborate on the data rates required with the MOFF and visibility based
+algorithms. This is particularly relevant in the context of radio transient 
+detection pipelines.
+
+Implementation of the MOFF algorithm with EPIC yields calibrated images on
+timescales of the output generated by the digitizer and is set by the inverse
+of the frequency channel width. These calibrated images are accumulated and 
+averaged to a certain timescale depending on science or hardware requirements,
+or when the sky has rotated significantly, whichever is lesser. In visibility
+based algorithms, the visibilities are accumulated and averaged to this 
+timescale before images are produced. Thus the data throughput (in samples per 
+second) with MOFF and 
+X-engine outputs are: 
+\begin{align}
+  r_\textrm{MOFF} &\sim \frac{4\Ngrid}{\Delta t} \left(\frac{\Delta B}{\Delta f}\right) \\
+  r_\textrm{X} &\sim 2\,\frac{\Nant(\Nant-1)/2}{\Delta t} \left(\frac{\Delta B}{\Delta f}\right),
+\end{align}
+where, factor 4 in the expression for $r_\textrm{MOFF}$ accounts for imaging
+after zero-padding the gridded electric fields, the leading factor of 2 in the
+expression for $r_\textrm{X}$ accounts for the real and imaginary parts of the
+complex visibilities, $\Delta B$ is the bandwidth, $\Delta f$ is the frequency 
+resolution, and $\Delta t$ is the timescale over which the transient phenomenon
+is sampled and the data (images or visibilities) are averaged to.
+
+Though a full understanding of the FRB phenomena is yet to emerge, there are 
+indications the timescales of FRB objects are $\Delta t \sim 1$--10~ms 
+\citep{tho13}. For a telescope like HERA, $\Delta B \simeq 100$~MHz, 
+$\Delta f \simeq 100$~kHz. For HERA-331, $\Nant=331$ and with a grid 
+spacing to image the field of view, $\Ngrid \simeq 441$ or 
+$\Ngrid \simeq 1024$ if $\Ngrid$ is preferred as a power of 2. 
+Using 8 bytes for each floating point sample, the throughputs are 
+$r_\textrm{MOFF} \lesssim 3$~GB/s and $t_\textrm{X} \simeq 81$~GB/s. For HERA-37,
+$r_\textrm{MOFF} \lesssim 190$~MB/s and $t_\textrm{X} \simeq 1$~GB/s. In such a 
+case, The X-engine throughput corresponds to $\simeq 3.5$~TB an hour. 
+Conversely, for the same data throughput, the MOFF algorithm can sample even 
+shorter timescales. Table~\ref{tab:data-rates} shows these data rates for some 
+of the current and planned telescopes for $\Delta t=10$~ms.
+
+\begin{table}
+  \centering
+  \caption{Data throughput for various telescopes with MOFF and X-engine 
+    outputs on timescales of $\Delta t=10$~ms assuming $\Delta B=100$~MHz and
+    $\Delta f=100$~kHz.}
+  \label{tab:data-rates}
+  \begin{threeparttable}
+  \begin{tabular}{ccc} 
+    \hline
+    Telescope\tnote{a} & $r_\textrm{MOFF}$\tnote{b} & $r_\textrm{X}$ \\
+              & (GB/s) & (GB/s) \\
+    \hline
+    LWA1 & $\simeq 3$ & $\simeq 48.5$ \\
+    LWA-OV & $\simeq 12$ & $\simeq 48.5$ \\
+    HERA-19 & $\lesssim 0.19$ & $\simeq 0.25$ \\
+    HERA-37 & $\lesssim 0.19$ & $\simeq 1$ \\
+    HERA-331 & $\lesssim 3$ & $\simeq 81$ \\
+    CHIME & $\lesssim 6.1$ & $\simeq 1220$ \\
+    \hline
+  \end{tabular}
+  \begin{tablenotes}
+    \item[a] Antenna layouts are listed in Table~\ref{tab:antenna-layouts}.
+    \item[b] $\Ngrid$ is usually greater than true value because of 
+      rounding to the next power of 2. Thus $r_\textrm{MOFF}$ estimates are
+      usually lesser than the conservative values listed here.
+  \end{tablenotes}
+  \end{threeparttable}
+\end{table}
+
+Thus even with conservative estimates, the MOFF algorithm provides very 
+economic data throughput for a majority of next generation radio telescopes
+with a dense layout. The most significant advantage is that calibrated images 
+are also available at no extra cost.
 
@@
 
-A unique and significant advantage is the instantaneous availability of 
-calibrated time-domain images bundled together with the hardware frontend 
-such as the F-engine as an integrated module. Hence, it is a compelling 
-candidate for time-domain radio astronomy, e.g. search for and monitoring of 
-transients. Transient detection pipeline at the backend of EPIC can be 
-fine-tuned to target fast transients such as the Fast Radio Bursts 
-\citep[FRB;][]{tho13} on millisecond timescales at GHz frequencies or slow 
-transients from planetary and exoplanetary origins at frequencies around 
-100~MHz.
+A unique advantage is the instantaneous availability of calibrated time-domain 
+images bundled together with the hardware frontend such as the F-engine as an 
+integrated module with significant savings in data throughput relative to a 
+X-engine based pipeline. Hence, it is a compelling candidate for time-domain 
+radio astronomy, e.g. search for and monitoring of transients. Transient 
+detection pipeline at the backend of EPIC can be fine-tuned to target fast 
+transients such as the Fast Radio Bursts \citep[FRB;][]{tho13} on millisecond 
+timescales at GHz frequencies or slow transients from planetary and 
+exoplanetary origins at frequencies around 100~MHz.
 

### 2015-10-28 imaging_demo/ms.tex
Improve explanation for image diff, update figures
From cadd827 (master):
=====================

(*) Add updated image comparison figure, uvwts comparison figure to
the repository and remove older versions

In ms.tex,

(*) Correct minor typos throughtout the paper
(*) Update figures 6, 7, 8, 9 and 12
(*) Recalculate throughput for FX data rates in case of transients
(*) Emphasize the whole point of EPIC
(*) Provide a unified explanation for differences seen with the two
imaging approaches

@@
 \begin{abstract}
-Modern radio telescopes are favouring densely packed array layouts 
-consisting of a large number of antennas 
-($\Nant\gtrsim 1000$--$\gtrsim 10000$). Since the complexity of traditional 
-correlators scales as $\mathcal{O}(\Nant^2)$, there will be a steep cost for 
-realizing the full imaging potential of these powerful instruments. Through 
-our generic and efficient E-field Parallel Imaging Correlator (EPIC), we present 
-the first software demonstration of a generalized direct imaging algorithm 
-known as the Modular Optimal Frequency Fourier (MOFF) imager. It takes 
+Modern radio telescopes are favoring densely packed array layouts consisting of 
+large numbers of antennas ($\Nant\gtrsim 1000$). Since the complexity of 
+traditional correlators scales as $\mathcal{O}(\Nant^2)$, there will be a steep 
+cost for realizing the full imaging potential of these powerful instruments. 
+Through our generic and efficient E-field Parallel Imaging Correlator (EPIC), 
+we present the first software demonstration of a generalized direct imaging 
+algorithm known as the Modular Optimal Frequency Fourier (MOFF) imager. It takes 
 advantage of the multiplication-convolution theorem of Fourier transforms. Not 
-only does it bring down the cost to $\mathcal{O}(\Nant\log \Nant)$ 
-but can also image from irregularly arranged heterogeneous antenna arrays. 
-EPIC is highly modular and parallelizable. It is implemented in object oriented 
-Python and is publicly available. We have verified the images produced to be 
-mathematically identical to those produced using traditional techniques. We have 
-also validated our implementation on data observed with the Long Wavelength 
-Array. Antenna systems with a dense filling factor consisting of a large number 
-of antennas such as LWA, SKA, and HERA will gain a significant advantage by 
-deploying EPIC. Inherent availability of calibrated time-domain images on 
-timescales roughly equal to the writeout timescale of the digitizer and vastly 
-lower bandwidth relative to visibility based systems will make it a prime 
-candidate for transient searches of Fast Radio Bursts (FRB) as well as 
-planetary and exoplantenary phenomena. 
+only does it bring down the cost for dense layouts to 
+$\mathcal{O}(\Nant\log_2\Nant)$ but can also image from irregularly arranged 
+heterogeneous antenna arrays. EPIC is highly modular and parallelizable, 
+implemented in object oriented Python, and publicly available. We have verified 
+the images produced to be equivalent to those produced using traditional 
+techniques. We have also validated our implementation on data observed with the 
+Long Wavelength Array (LWA). Antenna systems with a dense filling factor 
+consisting of a large number of antennas such as LWA, the Square Kilometre Array, 
+Hydrogen Epoch of Reionization Array, and Canadian Hydrogen Intensity Mapping 
+Experiment will gain significant advantage by deploying EPIC. Inherent 
+availability of calibrated time-domain images on timescales roughly equal to the 
+writeout timescale of the digitizer and vastly lower bandwidth relative to 
+visibility based systems will make it a prime candidate for transient searches 
+of Fast Radio Bursts (FRB) as well as planetary and exoplanetary phenomena. 
 \end{abstract}
@@
 Intensity Mapping Experiment (CHIME,\citealt{ban14}), the Long Wavelength 
-Array (LWA, \citealt{ell13}), and the low frequency component of the Square 
-Kilometer Array Low Frequency Aperture Array (SKA1-Low \citealt{mel13}).
+Array (LWA, \citealt{ell13}), and the low frequency Square Kilometer Array 
+(SKA1-Low \citealt{mel13}).
 
@@
 correlation between all antenna pairs in many narrow frequencies, forming
-\emph{visibilities}, the traditional fundamental measurement of radio
+\emph{visibilities}, the fundamental measurement of traditional radio
 interferometers. The computational requirements for a modern FX correlator scale
@@
 computational scaling from the harsh $\Nant^2$ to the more gentle envelope of
-$\Ngrid\log\Ngrid$, where $\Ngrid$ is the number of grid pixels in the Fourier 
+$\Ngrid\log_2\Ngrid$, where $\Ngrid$ is the number of pixels in the Fourier 
 transform \citep[e.g.][]{mor11,teg09,teg10}. This architecture is often referred 
@@
 elements are identical. On the other hand, the MOFF algorithm achieves the same
-$\Ngrid \log \Ngrid$ computational scaling without placing any restriction on
-antenna placement, can accommodate non-identical beam patterns, and is 
-a provably optimal mapping \citep{mor11}. This algorithm uses the antenna beam 
-patterns to grid the electric field measurements to a regular grid in the software
+$\Ngrid \log_2 \Ngrid$ computational scaling without placing any restriction on
+antenna placement, can accommodate non-identical beam patterns, and is provably 
+optimal \citep{mor11}. This algorithm uses the antenna beam patterns to 
+grid the electric field measurements to a regular grid in the software 
 holography/A-transpose fashion \citep{mor09,bha08,teg97b} before performing the
 spatial FFT. This process has been shown to theoretically produce a data product
-identical to images produced from the traditional FX correlator.
+identical to images produced from traditional visibility based techniques.
 
@@
 announce the public release of the E-field Parallel Imaging Correlator (EPIC)
-code. We begin with a technical description of the algorithm in \S\ref{sec:math}, 
+code. EPIC is a highly parallel, object oriented Python package that primarily 
+implements the MOFF imaging algorithm besides emulating real-life telescopes and 
+FX/XF correlators in software, and includes a visibility based imaging technique 
+for reference. It is intended to provide a development platform to test different 
+imaging approaches, characterize scaling relations and serve as a stepping stone 
+for real-life GPU/FPGA based implementation on telescopes.
+
+We begin with a technical description of the algorithm in \S\ref{sec:math}, 
 then discuss our particular implementation in \S\ref{sec:software}. We then 
@@
 the aperture illumination pattern does not change over the time-scale of the 
-averaging. This expression can be further simplified with the sky brightness, 
+averaging and $\star$ denotes complex conjugation. This expression can be further 
+simplified with the sky brightness, 
 $I(\hat{\mathbf{s}})= \left\langle E(\hat{\mathbf{s}}) E^\star(\hat{\mathbf{s'}}) 
 \right\rangle_t  \delta(\hat{\mathbf{s}}-\hat{\mathbf{s'}})$, and defining the 
-antenna pair sky power response function (or the primary beam), 
+antenna pair sky power response function (or the directional antenna power pattern), 
 $B_p(\hat{\mathbf{s}})~\equiv~{W}_a(\hat{\mathbf{s}})~{W}^\star_b(\hat{\mathbf{s}})$. 
-The result is the visibility expressed in terms of the sky brightness, the primary 
-beam, and uncorrelated noise terms which we group into $\widetilde{n}_p$,
+The result is the visibility expressed in terms of the sky brightness, the power 
+response, and uncorrelated noise terms which we group into $\widetilde{n}_p$.
 \begin{equation}
@@
 \end{equation}
-where the baseline coordinate $\mathbf{u}_p=\mathbf{r}_a-\mathbf{r}_b$ is the 
+where, the baseline coordinate $\mathbf{u}_p=\mathbf{r}_a-\mathbf{r}_b$ is the 
 vector separation between the two antennas. This signifies that the visibility 
@@
 
-This can
-be equivalently re-written as:
+This can be equivalently re-written as:
 \begin{align}\label{eqn:software-holography}
-  \widetilde{V}_p &= \int \tilde{B}(\mathbf{u}^\prime-\mathbf{u}) \times \left[\int e^{-i2\pi\mathbf{u}.\hat{\mathbf{s}}}\,I(\hat{\mathbf{s}})\,\dif^2\hat{\mathbf{s}}\right]\dif^2\mathbf{u}\vphantom{\int} + n_p,
+  \widetilde{V}_p &= \int \widetilde{B}(\mathbf{u}^\prime-\mathbf{u}) \left[\int e^{-i2\pi\mathbf{u}.\hat{\mathbf{s}}}\,I(\hat{\mathbf{s}})\,\dif^2\hat{\mathbf{s}}\right]\dif^2\mathbf{u}\vphantom{\int} + n_p,
 \end{align}
-where, $\tilde{B}(\mathbf{u})$ denotes the $uv$-space antenna power response
+where, $\widetilde{B}(\mathbf{u})$ denotes the $uv$-space antenna power response
 obtained by a Fourier transform of $B(\hat{\mathbf{s}})$. Effectively, the
 multiplication in image space by $B(\hat{\mathbf{s}})$ has been replaced by a
-convolution with $\tilde{B}(\mathbf{u})$ in $uv$-space. This is the software
+convolution with $\widetilde{B}(\mathbf{u})$ in $uv$-space. This is the software
 holographic equivalent of traditional FX correlator output.
@@
 gridded using an antenna-based gridding convolution function after which it is 
-spatially Fourier transformed and squared to obtain images for every time stamp. 
-These images are then time-averaged to obtain the accumulated image 
+spatially Fourier transformed and squared to obtain an image cube for every time 
+stamp. These images are then time-averaged to obtain the accumulated image 
 $I(f)$ ($\mathbf{I}(\hat{\mathbf{s}})$ in matrix notation).
@@
     shown on the left are measured as time-series $E_1(t)\ldots E_\textrm{N}(t)$
-    by the antennas which are then Fourier transformed by the F-engine to produce
-    electric field spectra $E_1(f)\ldots E_\textrm{N}(f)$. They are calibrated and
-    gridded. The gridded electric fields $E_\textrm{g}(f)$ from each time series
-    are imaged to produce an images $I_1(f)\ldots I_\textrm{N}(f)$. These images
-    are time-averaged to obtain the final image $I(f)$.}
+    by the antennas $\textrm{A}_1\ldots \textrm{A}_\textrm{N}$ which are then 
+    Fourier transformed by the F-engine to produce electric field spectra 
+    $E_1(f)\ldots E_\textrm{N}(f)$. They are calibrated and gridded. The gridded 
+    electric fields $E_\textrm{g}(f)$ from each time series are imaged to produce 
+    images $I_1(f)\ldots I_\textrm{M}(f)$. These images are time-averaged to 
+    obtain the final image $I(f)$.}
   \label{fig:MOFF-flowchart}
@@
 $\langle V_\textrm{ij}(f)\rangle$ which are then gridded and imaged to obtain the 
-image $I(f)$. The $I(f)$ obtained from both techniques are identical as explained 
-in \S\ref{sec:math}.
+image $I(f)$. The $I(f)$ obtained from both techniques are theoretically 
+identical as explained in \S\ref{sec:math}.
 \begin{figure}
@@
     The FX process flow shares the F-engine with the MOFF process. Following the 
-    F-engine, the electric fields pass through the X-engine to obtain 
+    F-engine, the electric fields pass through the X-engine to produce 
     visibilities $V_\textrm{ij}(f)$ which are calibrated and time-averaged. Then 
     they are gridded to obtain the gridded visibilities $V_\textrm{g}(f)$ which 
-    are then Fourier transformed to obtain the image.
-    $I(f)$.}
+    are then Fourier transformed to obtain the image, $I(f)$.}
   \label{fig:FX-flowchart}
@@
 Here we discuss the components of these architectures in detail.
-
-\par\medskip
-\noindent {\bf Temporal Fourier transform}
-\par\medskip
-\noindent This module is common to the MOFF and visibility based imaging 
-techniques. Time samples of electric fields measured by the antenna and 
-digitized by the A/D converter is Fourier transformed to generate electric 
-field spectra. This step can be parallelized by antennas as shown in 
-Fig.~\ref{fig:f-engine}. The output is then fed to either MOFF or visibility 
-based imaging pipelines.
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{F-engine.eps}
-  \caption{Block diagram of a F-engine. The electric field data streams from
-    antennas are Fourier transformed in parallel to generate electric field
-    spectra.}
-  \label{fig:f-engine}
-\end{figure}
 
@@
 
-In the most general case, this gridding kernel could contain information on the
-$w$-projection effect, and even other time-dependent ionospheric effects. For a
+In the most general case, this gridding kernel could contain information on
+$w$-projection effects, and other time-dependent ionospheric effects. For a
 stationary antenna array in the absence of any time-dependent effects, this
@@
 \par\medskip
+\noindent {\bf Temporal Fourier transform}
+\par\medskip
+\noindent This module is common to the MOFF and visibility based imaging 
+techniques. Time samples of electric fields measured by the antenna and 
+digitized by the A/D converter is Fourier transformed to generate electric 
+field spectra. This step can be parallelized across antennas as shown in 
+Fig.~\ref{fig:f-engine}. The output is then fed to either MOFF or visibility 
+based imaging pipelines.
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{F-engine.eps}
+  \caption{Block diagram of a F-engine. The electric field data streams from
+    antennas are Fourier transformed in parallel to generate electric field
+    spectra.}
+  \label{fig:f-engine}
+\end{figure}
+
+\par\medskip
 \noindent {\bf Calibration}
@@
 efficient multiplication with the sparse matrix created in the antenna-to-grid
-mapping process using the sparse matrices module in Python's SciPy package. 
+mapping process using the sparse matrices module in Python SciPy package. 
 Unless $\widetilde{\mathbf{W}}(\mathbf{r})$ includes time-dependent
@@
 In MOFF imaging, these are spatially Fourier transformed followed by a squaring
-operation at every timestamp for every frequency channel. In visibility based 
+operation at every time stamp for every frequency channel. In visibility based 
 imaging, the spatial Fourier transform is performed only once per integration 
@@
     as its components and contains collective properties about the antenna 
-    subsystems. Its core function is the creation of antenna-to-grid mapping and 
-    that of gridded aperture weights and electric fields. The interferometer 
-    array module is very similar in principle to the antenna array module 
-    except it operates on cross-correlations and produces gridded visibilities.
-    The image module takes gridded electric fields or visibilities and performs
-    a two-dimensional spatial FFT (and squares the intermediate image in case 
+    subsystems. Its core function is the creation of antenna-to-grid mapping, 
+    gridded aperture weights and electric fields. The interferometer array 
+    module is very similar in principle to the antenna array module except it 
+    operates on cross-correlations and produces gridded visibilities. The image 
+    module takes gridded electric fields or visibilities and performs a 
+    two-dimensional spatial FFT (and squares the intermediate image in case 
     of the former) to produce output images. Broadly, the MOFF algorithm is 
@@
 and represents the collective properties of its component antennas. By virtue of 
-holding each antenna data independently in their respective modules the F-engine 
+holding each antenna data independently in their respective modules, the F-engine 
 for the entire array can be distributed to the F-engines of the component 
@@
 produced through simulations. We simulate electric field streams from a model 
-sky and process the data through both the MOFF and visibility based imaging 
-algorithms. We then compare the output images to demonstrate their equivalence.
+sky and process the data through both the MOFF and a visibility based imaging 
+algorithm. We then compare the output images to demonstrate their equivalence.
 
@@
 10 point sources of flux densities 10~Jy at random locations. The number of
-timestamps integrated in one integration cycle was kept at eight where each A/D 
+time stamps integrated in one integration cycle was kept at eight where each A/D 
 timeseries is $1/\delta f=25\,\mu$s long. We use the MWA array layout 
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{MOFF_FX_image_comparison.eps}
-  \caption{Dirty images (top) and synthesized beams (bottom) obtained from 
-    simulated data using antenna-based MOFF algorithm (left) and FX 
-    visibility-based software holography (right). The solid black circles in the
-    top panels indicate the simulated source positions. The antenna 
-    auto-correlations at zero-spacing have been removed from the MOFF images. 
-    The images in either case reconstruct the sources at the right locations 
-    with the fluxes expected after multiplication by the antenna power pattern. 
-    The synthesized beams from the two algorithms are well matches in size and
-    shape. The overall modulation by the power pattern is seen clearly in both
-    images.}
-  \label{fig:MOFF-FX-image}
-\end{figure}
-
-Fig.~\ref{fig:MOFF-FX-uvwts} shows the cross-correlation weights obtained 
-with MOFF imaging (left) and visibility based imaging (right). The first 
-notable difference is in the weights around zero spacing. Though both show a
-dominant void around zero-spacing, the void obtained with MOFF algorithm shows 
-many pixels with non-zero weights. In contrast, the zero-spacing void from 
-traditional imaging consists of predominantly zero-valued pixels. The gridding
-process in the former involves rounding the antenna footprint to the nearest 
-grid pixel. Depending on the exact location of the center of the antenna 
-relative to the grid, the grid pixels that receive contribution from an
-antenna may be a pixel narrower along one or both axes relative to that from
-another identical antenna but with a different center location relative to the
-grid. The resulting auto-correlation footprint will also not necessarily be 
-identical. Hence, using a single expected footprint for subtraction will
-typically leave some residuals behind as seen in the void region in the 
-weights obtained with MOFF imaging.
-
-These residuals can be mitigated by: 
-\begin{enumerate}
-  \item making the grid spacing finer which makes the rounding error less 
-    susceptible to the location of the antenna center relative to the grid, 
-    and
-  \item subtracting each auto-correlation of antenna weights separately by 
-    using the shape and extent of the footprint appropriate for that specific
-    antenna aperture. 
-\end{enumerate}
-The latter is the most general solution applicable especially in the case of 
-heterogeneous antenna arrays and is under active development for EPIC.
-
-The other notable difference is that weights outside the zero-spacing void in
-some regions are different from each other at the few percent level though the 
-sum of weights in these regions are identical. For instance, note the 
-difference in weights at $(u,v)\approx (2,-10)$. This is found to arise 
-because with MOFF imaging the antenna locations were rounded to the nearest 
-grid pixel whereas in visibility based imaging the baseline locations 
-(difference between antenna locations) were rounded to the nearest grid pixel. 
-Since rounding is not a commutative operation (i.e. rounding of the difference 
-value is not necessarily equal to the difference of the rounded values), the 
-gridding operation in the two cases introduces rounding errors in the placement 
-of antenna aperture and cross-correlated aperture weights. Each antenna 
-aperture weight projected to the nearest grid location can be displaced by 
-$\sim 1$--2 pixels. This effect can be mitigated only by making the grid 
-spacing finer at the expense of increased computational cost.
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{MOFF_FX_comparison_uvwts_test_aperture_zero_spacing_removed.png}
+  \includegraphics[width=\columnwidth]{MOFF_FX_comparison_10_random_source_positions_4_iterations_test_aperture_zoomed.eps}
   \caption{Dirty images (top) and synthesized beams (bottom) obtained from 
@@
     images.}
+  \label{fig:MOFF-FX-image}
+\end{figure}
+
+Fig.~\ref{fig:MOFF-FX-uvwts} shows the correlated weights obtained with MOFF 
+imaging (left) and visibility based imaging (right). There are two notable 
+differences. The first is in the weights around zero spacing. Though both show a
+dominant void around zero-spacing, the void obtained with MOFF algorithm shows 
+many pixels with non-zero weights. In contrast, the zero-spacing void from 
+traditional imaging consists of predominantly zero-valued pixels. The second 
+difference is in the weights outside the zero-spacing void. In some regions they 
+are different from each other at the few percent level though the sum of weights 
+in these regions is the same. For instance, note the difference in weights at 
+$(u,v)\approx (2,-10)$. The reason for these differences is described below.
+
+The gridding step in MOFF imaging samples the antenna footprint (either in 
+analytic or lookup table formats) at the grid locations. Coverage of grid pixels 
+by an antenna footprint may be $\sim 1$ pixel narrower particularly at the 
+edge of the footprint along one or both directions relative to that from another 
+identical antenna but with a fractionally different location relative to the 
+grid. This depends on the exact location of the center of the antenna relative 
+to the grid and the coarseness of grid spacing. This first order loss of 
+precision of the sampled footprint propagates to second order ($\sim 2$ pixels) 
+upon correlation of the discretized weights. In other words, the correlated 
+weights may suffer further loss of precision in their sampled footprint after 
+correlation of two footprints each of which could be less precise to first 
+order. On the other hand, in visibility based imaging, a directly sampled $uv$ 
+plane antenna power response (which is theoretically identical to the correlation 
+of individual antenna footprints) centered on a baseline has a loss in precision 
+at most to first order. Thus, although in the limit of infinitesimally small 
+grid spacing they should be identical, the coarseness of grid spacing introduces 
+subtle differences between the two.
+
+These differences which are dependent on the coarseness of grid spacing can be 
+mitigated by making the grid spacing finer at the expense of increased 
+computational cost. Residuals centered around zero spacing can also be lowered 
+by subtracting each auto-correlation of antenna weights separately by using the 
+shape and extent of the sampled footprint appropriate for that specific antenna 
+aperture. This is a general solution applicable even in case of heterogeneous 
+antenna arrays and is under active development for EPIC.
+
+% ----
+
+% Fig.~\ref{fig:MOFF-FX-uvwts} shows the cross-correlation weights obtained 
+% with MOFF imaging (left) and visibility based imaging (right). The first 
+% notable difference is in the weights around zero spacing. Though both show a
+% dominant void around zero-spacing, the void obtained with MOFF algorithm shows 
+% many pixels with non-zero weights. In contrast, the zero-spacing void from 
+% traditional imaging consists of predominantly zero-valued pixels. The gridding
+% process in the former involves tying the antenna footprint (either in 
+% analytic or lookup table formats) to the nearest grid pixel. Depending on the 
+% exact location of the center of the antenna relative to the grid and the 
+% coarseness of grid spacing, grid pixels that receive contribution from an 
+% antenna may be a pixel narrower along one or both axes relative to that from 
+% another identical antenna but with a different center location relative to the 
+% grid. The resulting auto-correlation footprint will also not necessarily be 
+% identical. Hence, using a single expected footprint for subtraction will 
+% typically leave some residuals behind as seen in the void region in the weights 
+% obtained with MOFF imaging.
+
+% These residuals can be mitigated by: 
+% \begin{enumerate}
+%   \item making the grid spacing finer which makes the nearest neighbour 
+%     association between the antenna footprint and the grid less susceptible 
+%     (proportional to grid spacing) to the location of the antenna center 
+%     relative to the grid, and
+%   \item subtracting each auto-correlation of antenna weights separately by 
+%     using the shape and extent of the footprint appropriate for that specific
+%     antenna aperture. 
+% \end{enumerate}
+% The latter is the most general solution applicable especially in the case of 
+% heterogeneous antenna arrays and is under active development for EPIC.
+
+% The other notable difference is that weights outside the zero-spacing void in
+% some regions are different from each other at the few percent level though the 
+% sum of weights in these regions are identical. For instance, note the 
+% difference in weights at $(u,v)\approx (2,-10)$. This is found to arise 
+% because with MOFF imaging the antenna locations were tied to the nearest 
+% grid pixel whereas in visibility based imaging the baseline locations 
+% (difference between antenna locations) were tied to the nearest grid pixel. 
+% Since rounding is not a commutative operation (i.e. rounding of the difference 
+% value is not necessarily equal to the difference of the rounded values), the 
+% gridding operation in the two cases introduces rounding errors in the placement 
+% of antenna aperture and cross-correlated aperture weights. Each antenna 
+% aperture weight projected to the nearest grid location can be displaced by 
+% $\sim 1$--2 pixels. This effect can be mitigated only by making the grid 
+% spacing finer at the expense of increased computational cost.
+
+% ----
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{MOFF_FX_comparison_uvwts_test_aperture_zero_spacing_removed.eps}
+  \caption{Comparison of $uv$ plane antenna power response with MOFF (left) and 
+    visibility based imaging (right). The zero-spacing footprint from the $uv$ 
+    weights obtained with MOFF has been subtracted. Differences in the void 
+    centered around zero-spacing and other outer regions is due to loss of 
+    sampled antenna footprint precision due to coarseness of grid spacing 
+    further made severe by the correlation process.}
   \label{fig:MOFF-FX-uvwts}
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{diff_psf_MOFF-FX_test_aperture.png}
+  \includegraphics[width=\columnwidth]{diff_psf_MOFF-FX_test_aperture.eps}
   \caption{Map of difference between the synthesized beams obtained with the 
     two methods (top) and radial statistics of the synthesized beams and 
-    their differences (bottom). The maximum difference is of the order of a 
-    few percent. The difference appears to be modulated in amplitude by the 
-    power pattern of the antenna. }
+    their differences (bottom). The radial variations of rms of the synthesized 
+    beam (gray) and that of the rms of the difference (black) are shown as 
+    percentage of the peak synthesized beam. Radially averaged directional 
+    antenna power response in absolute scale is shown in red and is to be read 
+    with the scale on the right side of the axis. The maximum difference is of 
+    the order of a few percent. Amplitude of the rms of the difference is 
+    modulated by the power pattern of the antenna.}
   \label{fig:psf-diff}
@@
 The synthesized beam rms is proportional to the antenna power pattern as 
-expected from a point spread function uncorrected for the antenna power 
-pattern. The rms of differenced synthesized beams is 
-also modulated by the antenna power pattern. The rms of the difference is 
-definitely lesser than the rms of the synthesized beam in the central regions 
-up to $(l^2+m^2)^{1/2}\lesssim 0.3$. This implies that the beams are well 
-matched in the central regions. In the outer regions, their mismatch is 
-comparable to the rms of synthesized beams. This indicates the two 
-synthesized beams are not completely randomly different from each other in
-which case the rms of the difference would have been $\approx \sqrt{2}$
-higher than the rms of the each of the synthesized beams. This indicates 
-that while differences exist, large fractions of them are still well matched
-to each other even out to the horizon. Thus the rounding errors in gridding 
-do not affect the statistics of the images or the synthesized beams.
+expected from a point spread function uncorrected for the antenna power pattern. 
+The rms of differenced synthesized beams is also modulated by the antenna power 
+pattern. The rms of the difference is definitely lesser than the rms of the 
+synthesized beam in the central regions up to $(l^2+m^2)^{1/2}\lesssim 0.3$. This 
+implies that the beams are well matched in the central regions. In the outer 
+regions, their mismatch is comparable to the rms of synthesized beams. This 
+indicates the two synthesized beams are not completely randomly different from 
+each other in which case the rms of the difference would have been 
+$\approx \sqrt{2}$ higher than the rms of the each of the synthesized beams. 
+This indicates that while differences exist, large fractions of them are still 
+well matched to each other even out to the horizon. Thus the rounding errors in 
+gridding do not affect the statistics of the images or the synthesized beams.
 
@@
 
-Here we demonstrate our software using narrow band data from the LWA in New 
-Mexico. This data is in LWA narrow-band transient buffer (TBN) format with 512 
-voltage time samples from 255 antennas within roughly a diameter of 100~m. The 
-data is centered at a frequency of 74.03~MHz, with a sample rate (equal to the 
-bandwidth) of 100~kHz with 512 complex time samples in a A/D writeout timescale 
-of 5.12~ms, a frequency resolution of 195.3125~Hz and dual polarization. There 
-are 391 such timestamps yielding a total duration of 2~s.
+Here we demonstrate our software using narrow band data from the LWA station 
+in New Mexico. This data is in LWA narrow-band transient buffer (TBN) format with 
+512 voltage time samples from 255 antennas within roughly a diameter of 100~m. 
+The data is centered at a frequency of 74.03~MHz, with a sample rate (equal to 
+the bandwidth) of 100~kHz with 512 complex time samples in a A/D writeout 
+timescale of 5.12~ms, a frequency resolution of 195.3125~Hz and dual 
+polarization. There are 391 such time stamps yielding a total duration of 2~s.
 
@@
 Fig.~\ref{fig:LWA-image} shows the image produced with MOFF imaging 
-packaged in EPIC after averaging over the entire 2~s of data and the inner
+packaged in EPIC after averaging over 20~ms of data and the inner
 $\approx 80$\% of bandwidth (roughly 80~kHz). The image is shown in direction 
@@
   \caption{Image from LWA TBN data obtained with MOFF imaging using EPIC 
-    package after averaging over 2~s and $\approx 80$~kHz. The x- and y-axes
+    package after averaging over 20~ms and $\approx 80$~kHz. The x- and y-axes
     denote direction cosines $l$ and $m$ respectively. The antenna voltages
-    are compensated for their respective delays. The flux scale is arbitrary
-    and not calibrated. Locations of Cyg~A and Cas~A are annotated.}
+    are compensated for their respective delays. The flux scale is arbitrary.
+    Locations of Cyg~A and Cas~A are annotated.}
   \label{fig:LWA-image}
@@
 the MOFF imaging is determined by the spatial Fourier transform while  
-that for visibility based imaging comes from $\Nant(\Nant-1)/2$ 
-cross-correlations. However, if non-linearities such as non-coplanarity of 
-baselines \citep{cor08} and wide-field phenomena like the {\it pitchfork} 
-effect \citep{thy15a,thy15b} are to be corrected for, the antenna based 
-illumination footprint can start becoming less compact in the measurement 
-plane and can result in a costlier gridding process.
+that for visibility based imaging comes from the cross-correlations. However, 
+if non-linearities such as non-coplanarity of baselines \citep{cor08} and 
+wide-field phenomena like the {\it pitchfork} effect \citep{thy15a,thy15b} 
+are to be corrected for, the antenna based illumination footprint can start 
+becoming less compact in the measurement plane and can result in a costlier 
+gridding process.
 
@@
 transform implemented via Fast Fourier Transform \citep[FFT;][]{coo65} is 
-$\approx \beta \Ngrid\log \Ngrid$ where $\Ngrid$ 
-is the number of pixels on the grid and $\beta$ is a constant that depends on 
-the implementation of twiddle FFT algorithms \citep{bri74}. In our study, we 
-set $\beta=5$, a value\footnote{http://www.fftw.org/speed/method.html} much 
-more conservative than was indicated in \citet{mor11}. We set the number of 
-complex multiplications in the X-engine in visibility based imaging to 
-$\Nant^2$.
+$\approx \beta\,(4\Ngrid)\log_2(4\Ngrid)$ where $\Ngrid$ is the number of 
+pixels on the grid, the factor 4 accounts for increase in number of pixels as a 
+result of zero-padding before spatial Fourier transform, and $\beta$ is a 
+constant that depends on the implementation of twiddle FFT algorithms 
+\citep{bri74}. In our study, we set $\beta=5$, a 
+value\footnote{http://www.fftw.org/speed/method.html} much more conservative 
+than was indicated in \citet{mor11}. We set the number of complex 
+multiplications in the X-engine in visibility based imaging to 
+$\Nant(\Nant-1)/2$.
 
@@
     SKA1-LC\tnote{e} & 1000 & 750 & 962 & 150 \\
-    SKA1-LCD\tnote{f} & 1000 & 192,000 & 962 & 150 \\
+    SKA1-LCD\tnote{f} & 1000 & 192,000 & 2 & 150 \\
     CHIME & 100 & 1280 & 8 & 600 \\
-    HIRAX & 200 & 1024 & 6 & 600 \\
+    HIRAX\tnote{g} & 200 & 1024 & 6 & 600 \\
     \hline
@@
       without station beamforming
+    \item[g] Hydrogen Intensity mapping and Real-time Analysis eXperiment
   \end{tablenotes}
@@
 Fig.~\ref{fig:parameter-space-computations-instruments} shows the number of 
-complex operations per frequency channel per integration timescale. Telescopes 
-that fall to the left of this line indicate MOFF imaging is computationally
-more efficient than visibility based imaging. All HERA layouts except possibly
-HERA-19 are in a parameter space where MOFF imaging holds the advantage. The 
-solid line showing future trajectory of HERA like systems will be clearly 
-favoured by MOFF imaging. The gray shaded area is for a projected LWA 
-expansion and is also predominantly in the region favouring MOFF imaging. 
-It is bounded by the LWA1 and LWA-OV on the left and right 
-respectively. The current (see Table~\ref{tab:antenna-layouts}) and
-an expanded layout with a four-fold increase in number of elements over a
-50\% increase in $b_\textrm{max}$ provide the bounds at the bottom and top 
-respectively. Current instruments such as MWA and LOFAR lie in parameter space
-favouring visibility based imaging.
+complex operations per frequency channel per cross-polarization per integration 
+timescale. Telescopes that fall to the left of the solid line indicate MOFF 
+imaging is computationally more efficient than visibility based imaging. All HERA 
+layouts except HERA-19 and possibly HERA-37 are in a region of parameter space 
+where MOFF imaging holds the advantage. The dashed line showing future trajectory 
+of HERA like systems will be clearly favored by MOFF imaging. The dotted line is 
+similarly a hypothetical trajectory for the MWA with more tiles added inside the 
+same core diameter. The gray shaded area is for a projected LWA expansion and is 
+also predominantly in the region favoring MOFF imaging. It is bounded by the LWA1 
+and LWA-OV on the left and right respectively. The current (see 
+Table~\ref{tab:antenna-layouts}) and a hypothetical expanded layout with a 
+four-fold increase in number of elements over a 50\% increase in $b_\textrm{max}$ 
+provide the bounds at the bottom and top respectively. Current instruments such 
+as MWA and LOFAR lie in parameter space favoring visibility based imaging.
 
@@
   \caption{Current and planned instruments in parameter space of
-    number of complex multiplies and adds with MOFF and FX. The dashed line
+    number of complex multiplies and adds with MOFF and FX. The solid line
     is the boundary at which the number of operations with MOFF and visibility 
     based imaging are equal. MOFF imaging is more efficient for telescopes 
-    occupying the left of this line and vice versa. CHIME, HIRAX and all the 
-    HERA layouts except HERA-19 and HERA-37 lie in the parameter space favoured 
-    by MOFF imaging. And so are SKA1-LC and SKA1-LCD. The solid black curve 
-    shows the projected trajectory of bigger close-packed hexagonal layouts 
-    similar to HERA. The gray shaded area denotes the projected trajectory of 
-    the LWA bounded by LWA1 (left edge), LWA-OV (right edge), current layout 
-    (bottom) and a four-fold increase in the number of elements within a 50\% 
-    increase in the core size (top). Current instruments such as MWA and LOFAR 
-    fall in a region favoured by visibility based imaging.}
+    occupying the left of this line and vice versa. CHIME, HIRAX, SKA1-LC, 
+    SKA1-LCD and all the HERA layouts except HERA-19 and HERA-37 lie in the 
+    parameter space favored by MOFF imaging. The dashed line shows the projected 
+    trajectory of hypothetical expanded HERA layouts. The dotted line similarly
+    shows hypothetical expanded MWA layouts with more tiles added in the same
+    core. The gray shaded area denotes the projected trajectory of the LWA 
+    bounded by LWA1 (left edge), LWA-OV (right edge), current layout (bottom) 
+    and a four-fold increase in the number of elements within a 50\% increase in 
+    the core size (top). Current instruments such as MWA and LOFAR fall in a 
+    region favored by visibility based imaging.}
   \label{fig:parameter-space-computations-instruments}
@@
 
-We now consider antenna array layouts described by three quantities essential
-to radio interferometry, namely, maximum baseline length, number of antennas,
+We now consider antenna array layouts described by three essential quantities 
+in radio interferometry, namely, maximum baseline length, number of antennas,
 and the size of each antenna.
@@
 impossible. Dashed lines of each color denote the boundary to the left of which 
-visibility based imaging is favoured for the corresponding antenna size. Region 
-inside the wedge enclosed by the solid and dashed lines favours using the MOFF 
+visibility based imaging is favored for the corresponding antenna size. Region 
+inside the wedge enclosed by the solid and dashed lines favors using the MOFF 
 algorithm for the corresponding antenna size over visibility based imaging.
+
 As antenna size increases the maximum number of antennas for a dense packing as 
@@
 required with the MOFF algorithm. This shifts the dashed curves leftward.
+
 The different antennas are color coded by roughly the class of antenna size 
 they fall into. Thus symbols of one color falling into a wedge of the same color 
-indicates MOFF imaging is favoured for those telescopes and vice versa. For 
-e.g., MOFF imaging is favoured in HERA-33 and HERA-6769 because they are inside 
+indicates MOFF imaging is favored for those telescopes and vice versa. For 
+e.g., MOFF imaging is favored in HERA-33 and HERA-6769 because they are inside 
 the wedge but not so in cases of HERA-19 and HERA-37. A majority of the 
@@
 respectively) will fall in the regime where MOFF imaging will be desirable. For 
-a fixed baseline length, regions favouring the MOFF algorithm tend to be towards 
+a fixed baseline length, regions favoring the MOFF algorithm tend to be towards 
 large $\Nant$ indicating large-N dense array layouts with smaller antenna 
@@
     lines of each color denote the boundary to the left of which visibility
-    based imaging is favoured for the corresponding antenna size. Region inside 
-    the wedge enclosed by the solid and dashed lines favours using the MOFF 
+    based imaging is favored for the corresponding antenna size. Region inside 
+    the wedge enclosed by the solid and dashed lines favors using the MOFF 
     algorithm for the corresponding antenna size. These wedges shift leftward 
@@
     advantageous for those telescopes and vice versa. For e.g., MOFF imaging
-    is favoured in HERA-33 and HERA-6769 because they are inside the wedge 
+    is favored in HERA-331 and HERA-6769 because they are inside the wedge 
     but not so in cases of HERA-19 and HERA-37.}
@@
 algorithms. This is particularly relevant in the context of radio transient 
-detection pipelines.
+detection.
 
@@
 timescale before images are produced. Thus the data throughput (in samples per 
-second) with MOFF and 
-X-engine outputs are: 
+second) per cross-polarization with MOFF and X-engine outputs are: 
 \begin{align}
@@
 \end{align}
-where, factor 4 in the expression for $r_\textrm{MOFF}$ accounts for imaging
+where, the factor 4 in the expression for $r_\textrm{MOFF}$ accounts for imaging
 after zero-padding the gridded electric fields, the leading factor of 2 in the
@@
 spacing to image the field of view, $\Ngrid \simeq 441$ or 
-$\Ngrid \simeq 1024$ if $\Ngrid$ is preferred as a power of 2. 
-Using 8 bytes for each floating point sample, the throughputs are 
-$r_\textrm{MOFF} \lesssim 3$~GB/s and $t_\textrm{X} \simeq 81$~GB/s. For HERA-37,
-$r_\textrm{MOFF} \lesssim 190$~MB/s and $t_\textrm{X} \simeq 1$~GB/s. In such a 
-case, The X-engine throughput corresponds to $\simeq 3.5$~TB an hour. 
-Conversely, for the same data throughput, the MOFF algorithm can sample even 
-shorter timescales. Table~\ref{tab:data-rates} shows these data rates for some 
-of the current and planned telescopes for $\Delta t=10$~ms.
+$\Ngrid \simeq 1024$ if $\Ngrid$ is preferred as a power of 2 in either 
+direction. Using 8 bytes for each floating point sample in the MOFF images and 
+4 bytes each for real and imaginary parts of visibility samples, the throughputs 
+are $r_\textrm{MOFF} \lesssim 3$~GB/s and $r_\textrm{X} \simeq 41$~GB/s. For 
+HERA-37, $r_\textrm{MOFF} \lesssim 190$~MB/s and $r_\textrm{X} \simeq 0.5$~GB/s. 
+In such a case, The X-engine throughput corresponds to an extreme rate of 
+$\simeq 1.8$~TB an hour per cross-polarization. Conversely, for the same data 
+throughput, the MOFF algorithm can sample even shorter timescales.
+
+Table~\ref{tab:data-rates} shows these data rates for some of the current and 
+planned telescopes for $\Delta t=10$~ms. In almost all cases listed, even with 
+conservative estimates, the MOFF algorithm provides very economic data 
+throughput for a majority of next generation radio telescopes with a dense 
+layout. The most significant advantage is that calibrated images are also 
+available at no extra cost.
 
@@
   \centering
-  \caption{Data throughput for various telescopes with MOFF and X-engine 
-    outputs on timescales of $\Delta t=10$~ms assuming $\Delta B=100$~MHz and
-    $\Delta f=100$~kHz.}
+  \caption{Data throughput per cross-polarization for various telescopes with 
+    MOFF and X-engine outputs on timescales of $\Delta t=10$~ms assuming 
+    $\Delta B=100$~MHz and $\Delta f=100$~kHz.}
   \label{tab:data-rates}
@@
     Telescope\tnote{a} & $r_\textrm{MOFF}$\tnote{b} & $r_\textrm{X}$ \\
-              & (GB/s) & (GB/s) \\
+              & (GB/s)\tnote{c} & (GB/s)\tnote{c} \\
     \hline
-    LWA1 & $\simeq 3$ & $\simeq 48.5$ \\
-    LWA-OV & $\simeq 12$ & $\simeq 48.5$ \\
-    HERA-19 & $\lesssim 0.19$ & $\simeq 0.25$ \\
-    HERA-37 & $\lesssim 0.19$ & $\simeq 1$ \\
-    HERA-331 & $\lesssim 3$ & $\simeq 81$ \\
-    CHIME & $\lesssim 6.1$ & $\simeq 1220$ \\
+    LWA1 & $\simeq 3$ & $\simeq 24.3$ \\
+    LWA-OV & $\simeq 12$ & $\simeq 24.3$ \\
+    HERA-19 & $\lesssim 0.19$ & $\simeq 0.13$ \\
+    HERA-37 & $\lesssim 0.19$ & $\simeq 0.5$ \\
+    HERA-331 & $\lesssim 3$ & $\simeq 41$ \\
+    CHIME & $\lesssim 6.1$ & $\simeq 610$ \\
     \hline
@@
     \item[b] $\Ngrid$ is usually greater than true value because of 
-      rounding to the next power of 2. Thus $r_\textrm{MOFF}$ estimates are
-      usually lesser than the conservative values listed here.
+      rounding to the next power of 2 in either direction. Thus 
+      $r_\textrm{MOFF}$ is usually lesser than the conservative values 
+      listed here.
+    \item[c] We assume 8 bytes for each real sample from MOFF images, and
+      4 bytes each for real and imaginary parts of visibility samples.
   \end{tablenotes}
@@
 \end{table}
-
-Thus even with conservative estimates, the MOFF algorithm provides very 
-economic data throughput for a majority of next generation radio telescopes
-with a dense layout. The most significant advantage is that calibrated images 
-are also available at no extra cost.
 
@@
 and efficient EPIC software that is designed to bring this cost down from 
-$\mathcal{O}(N^2)$ to $\mathcal{O}(N\log N)$. Under the class of direct imaging 
+$\mathcal{O}(N^2)$ to $\mathcal{O}(N\log_2 N)$. Under the class of direct imaging 
 techniques, ours is one of the most generic -- neither does it place any 
@@
 
-Our package, now publicly available, written in object oriented Python is 
+Our EPIC package, now publicly available, written in object oriented Python, is 
 highly modularized and parallelizable. It includes an implementation of the 
 MOFF algorithm in addition to visibility based software holography imaging and 
-a data simulator for sky models. It has been successfully tested on simulated 
-as well as real LWA observations.
+a data simulator for sky models. It is designed to provide a development 
+platform to compare different imaging approaches and serve as a stepping stone 
+for real-life GPU/FPGA based implementation on telescopes. It has been 
+successfully tested on simulated as well as real LWA observations.
 
 The MOFF algorithm packaged with EPIC is already found to be most suitable 
-for many present and planned radio telescopes such as the LWA, HERA and SKA. 
-In general, MOFF is most suited to operate in the region of parameter space
-characterized by dense packing of a large number of antennas especially when 
-consisting of a large number of small antenna elements.
+for many present and planned radio telescopes such as the LWA, HERA, CHIME,
+HIRAX and SKA. In general, MOFF is most suited to operate in the region of 
+parameter space characterized by dense packing of a large number of antennas 
+especially when consisting of a large number of small antenna elements.
 
@@
 images bundled together with the hardware frontend such as the F-engine as an 
-integrated module with significant savings in data throughput relative to a 
-X-engine based pipeline. Hence, it is a compelling candidate for time-domain 
-radio astronomy, e.g. search for and monitoring of transients. Transient 
-detection pipeline at the backend of EPIC can be fine-tuned to target fast 
-transients such as the Fast Radio Bursts \citep[FRB;][]{tho13} on millisecond 
-timescales at GHz frequencies or slow transients from planetary and 
+integrated module. It is seen to have significant savings in data throughput 
+relative to a X-engine based pipeline. Hence, it is a compelling candidate for 
+time-domain radio astronomy, e.g. search for and monitoring of transients. 
+Transient detection pipeline at the backend of EPIC can be fine-tuned to target 
+fast transients such as the Fast Radio Bursts \citep[FRB;][]{tho13} on 
+millisecond timescales at GHz frequencies or slow transients from planetary and 
 exoplanetary origins at frequencies around 100~MHz.
@@
 
-\section*{Acknowledgements}
-
-We thank Greg Taylor for providing us the LWA data, and Larry D'Addario, 
-Gregg Hallinan and Harish Vedantham for their valuable inputs. Construction of 
-the LWA has been supported by the Office of Naval Research under Contract 
-N00014-07-C-0147. Support for operations and continuing development of the LWA1 
-is provided by the National Science Foundation under grant AST-1139974 of the 
-University Radio Observatory program.
+\section*{Acknowledgments}
+
+We thank Larry D'Addario, Gregg Hallinan, Joseph Lazio and Harish Vedantham for 
+their valuable inputs, and Greg Taylor for providing us with LWA data. 
+Construction of the LWA has been supported by the Office of Naval Research under 
+Contract N00014-07-C-0147. Support for operations and continuing development of 
+the LWA1 is provided by the National Science Foundation under grant AST-1139974 
+of the University Radio Observatory program.
 

### 2015-10-28 imaging_demo/ms.tex
Make it compatible with arXiv (submitted v1)
From 70146f3 (arxiv):
====================

In ms.tex,

(*) Change including .bib file to including .bbl file
(*) Remove commented lines
(*) Remove figure filename extensions
(*) Comment out \usepackage{newtxtext,newtxmath} to be compatible
with arXiv submission
(*) Change fontsize of first table to scriptsize to fit the column
width

@@
 % Basic setup. Most papers should leave these options alone.
-\documentclass[a4paper,fleqn,usenatbib]{../mnras}
+\documentclass[a4paper,fleqn,usenatbib]{mnras}
 
@@
 % out the following line
-\usepackage{newtxtext,newtxmath}
+% \usepackage{newtxtext,newtxmath}
 % Depending on your LaTeX fonts installation, you might get better results with one of these:
-%\usepackage{mathptmx}
-%\usepackage{txfonts}
+% \usepackage{mathptmx}
+% \usepackage{txfonts}
 
@@
 \end{keywords}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-%%%%%%%%%
-
-%We provide a brief summary of the mathematical equivalence of the MOFF and FX
-%correlators detailed in \citet{mor11}. 
-%
-%Electric fields from astrophysical sources, $E(\hat{\mathbf{s}})$, in the sky
-%coordinate system denoted by unit vector $\hat{\mathbf{s}}$ propagate towards
-%the observer as:
-%\begin{align}
-%  \widetilde{E}(\mathbf{r}) &= \int E(\hat{\mathbf{s}})\,e^{-i2\pi\mathbf{r}\cdot\hat{\mathbf{s}}}\,\dif^2\hat{\mathbf{s}},
-%\end{align}
-%where, $\mathbf{r}$ denotes the observer's coordinate system and $\widetilde{E}
-%(\mathbf{r})$ is the propagated electric field. Thus the propagated electric
-%field is a linear superposition of the electric fields emanating from
-%astronomical sources with appropriate complex phases. It can also be described
-%as a Fourier transform of the electric fields in the sky coordinates. 
-%
-%An antenna, $a$, measures a phased sum of these propagated electric fields over
-%its effective collecting area with an additive receiver noise:
-%\begin{align}\label{eqn:measured-E-field}
-%  \widetilde{E}_a &= \int \widetilde{W}_a(\mathbf{r}-\mathbf{r}_a)\,\widetilde{E}(\mathbf{r})\,\dif^2\mathbf{r} + \widetilde{n}_a \\
-%                  &= \int \widetilde{W}_a(\mathbf{r}-\mathbf{r}_a) \left[ \int E(\hat{\mathbf{s}})\,e^{-i2\pi\mathbf{r}\cdot\hat{\mathbf{s}}}\,\dif^2\hat{\mathbf{s}} \right] \dif^2\mathbf{r} + \widetilde{n}_a \\
-%                  &= \int \delta(\mathbf{r}-\mathbf{r}_a) \left[ \int {W}_a(\hat{\mathbf{s}})\,E(\hat{\mathbf{s}})\,e^{-i2\pi\mathbf{r}\cdot\hat{\mathbf{s}}}\,\dif^2\hat{\mathbf{s}} \right] \dif^2\mathbf{r} + \widetilde{n}_a
-%\end{align}
-%where, $\widetilde{W}_a(\mathbf{r})$ is the aperture electric field illumination
-%pattern of the antenna and its Fourier transform, $W_a(\hat{\mathbf{s}})$ is the
-%directional antenna voltage response.
-%
-%Interferometers measure {\it visibility} -- the degree of coherence -- between
-%electric fields measured by a pair of antennas \citep{van34,zer38,tho01}.
-%Visibility, $\widetilde{V}_p$, can be written as:
-%
-%\begin{align}
-%  \widetilde{V}_p &= \left\langle \widetilde{E}_a\widetilde{E}_b^\star \right\rangle \label{eqn:cc-vis}\\
-%                  &= \int\delta(\mathbf{u}-\mathbf{u}_p)\left[\int e^{-i2\pi\mathbf{u}.\hat{\mathbf{s}}}\,B(\hat{\mathbf{s}})\,I(\hat{\mathbf{s}})\,\dif^2\hat{\mathbf{s}}\right] \dif^2\mathbf{u} + n_p,
-%\end{align}
-%where, $I(\hat{\mathbf{s}})$ is the sky brightness, $B(\hat{\mathbf{s}})$ is the
-%antenna power response, $\mathbf{u}$ denotes the coordinates of the $uv$-
-%measurement plane, $\mathbf{u}_p\equiv \mathbf{r}_a-\mathbf{r}_b$ is the antenna
-%spacing vector corresponding to the antenna pair $p$, and $n_p$ is the receiver
-%noise. This signifies that the visibility ($\widetilde{V}_p$) measured between a
-%pair of antennas ($p$) is obtained by the multiplying the sky brightness
-%$I(\hat{\mathbf{s}})$ by the antenna power response $B(\hat{\mathbf{s}})$ and
-%Fourier transform from the directional coordinates ($\hat{\mathbf{s}}$) to $uv$
-%coordinates, which are then sampled at the locations of the antenna spacings (or
-%baselines), namely, $\mathbf{u}_p$ using the sampling function
-%$\delta(\mathbf{u}-\mathbf{u}_p)$ and added to the receiver noise $n_p$. This can
-%be equivalently re-written as:
-%\begin{align}\label{eqn:software-holography}
-%  \widetilde{V}_p &= \int\delta(\mathbf{u}^\prime-\mathbf{u}_p)\left[\int \tilde{B}(\mathbf{u}^\prime-\mathbf{u}) \right. \nonumber\\
-%  &\left. \qquad \times \left[\int e^{-i2\pi\mathbf{u}.\hat{\mathbf{s}}}\,I(\hat{\mathbf{s}})\,\dif^2\hat{\mathbf{s}}\right]\dif^2\mathbf{u}\vphantom{\int}\right] \dif^2\mathbf{u}^\prime + n_p,
-%\end{align}
-%where, $\tilde{B}(\mathbf{u})$ denotes the $uv$-space antenna power response
-%obtained by a Fourier transform of $B(\hat{\mathbf{s}})$. Effectively, the
-%multiplication in image space by $B(\hat{\mathbf{s}})$ has been replaced by a
-%convolution with $\tilde{B}(\mathbf{u})$ in $uv$-space. This is the software
-%holographic equivalent of traditional FX correlator output. Following the matrix
-%notation of \citet{mor11}, the above equation can be expressed as:
-%\begin{align}
-%  \mathbf{m}(\mathbf{v}) &= \widetilde{B}(\mathbf{v},\mathbf{u})\,\mathbf{F}(\mathbf{u},\hat{\mathbf{s}})\,I(\hat{\mathbf{s}}) + \mathbf{n}(\mathbf{v}),
-%\end{align}
-%where, the sky brightness $I(\hat{\mathbf{s}})$ is Fourier transformed using
-%$\mathbf{F}(\mathbf{u},\hat{\mathbf{s}})$ and the resultant spatial coherence
-%function is weighted and summed using the antenna power response,
-%$\widetilde{B}(\mathbf{v},\mathbf{u})$ in $uv$-space sampled at the baseline
-%location to obtain the measured visibilities:
-%\begin{align}
-%  \mathbf{m}(\mathbf{v}) &= \left\langle \widetilde{E}(\mathbf{a})\ast\widetilde{E}(\mathbf{a}^\prime)\right\rangle_t. \label{eqn:matrix-cc-vis}
-%\end{align}
-%$\mathbf{m}(\mathbf{v})$ denotes visibilities measured by cross-correlating
-%measured antenna electric fields over all possible pairs of $\mathbf{a}$ and
-%$\mathbf{a}^\prime$. It is the same as equation~\ref{eqn:cc-vis} written in
-%matrix notation.
-%
-%Using the optimal map-making formalism \citep{teg97a,teg97b}, a software
-%holography image is formed using \citep{mor09}:
-%\begin{align}
-%  I^\prime(\hat{\mathbf{s}}) &= \mathbf{F}^\textrm{T}(\hat{\mathbf{s}},\mathbf{u})\,\widetilde{B}^{\,\textrm{T}}(\mathbf{u},\mathbf{v})\,\mathbf{N}^{-1}\,\mathbf{m}(\mathbf{v}) \label{eqn:dirty-image-FX}
-%\end{align}
-%where the measured visibilities are weighted by the inverse of the system noise,
-%following a gridding process using the holographic antenna power response as the
-%gridding kernel followed by a Fourier transform to create an image
-%$I^\prime(\hat{\mathbf{s}})$ which is the optimal estimate of the true image
-%$I(\hat{\mathbf{s}})$.
-
-%It may be noted that the antenna power response in $uv$-plane can be written as:
-%\begin{align}\label{eqn:antenna-power-breakup}
-%  \tilde{B}_p(\mathbf{u}) &= \widetilde{W}_a(\mathbf{u}) \ast \widetilde{W}_b^\star(\mathbf{u})
-%\end{align}
-%where, $W_a(\mathbf{u})$ and $W_b(\mathbf{u})$ denote the individual antenna
-%voltage responses. Thus $\tilde{B}_p(\mathbf{u})$ is a cross-correlation of the
-%individual antenna voltage responses. Similarly, it can be shown that 
-%\begin{align}
-%  \left\langle E(\hat{\mathbf{s}})\,E^\star(\hat{\mathbf{s}}^\prime)\right\rangle &= \left\langle E(\hat{\mathbf{s}})\,E^\star(\hat{\mathbf{s}}^\prime) \delta(\hat{\mathbf{s}}-\hat{\mathbf{s}}^\prime) \right\rangle \nonumber\\
-%  &= \left\langle \left|E(\hat{\mathbf{s}})\right|^2 \right\rangle = I(\hat{\mathbf{s}}) \label{eqn:sky-intensity}
-%\end{align}
-%
-%From equations~\ref{eqn:measured-E-field}, \ref{eqn:software-holography},
-%\ref{eqn:matrix-cc-vis}, \ref{eqn:antenna-power-breakup} and
-%\ref{eqn:sky-intensity}, the optimal image in equation \ref{eqn:dirty-image-FX}
-%can be expressed equivalently as:
-%\begin{align}
-%  I^\prime(\hat{\mathbf{s}}) &= \left\langle \left|\,\mathbf{F}^\textrm{T}(\hat{\mathbf{s}},\mathbf{r})\,\widetilde{\mathbf{W}}^\textrm{T}(\mathbf{r},\mathbf{a})\,\widetilde{\mathbf{N}}(\mathbf{a},\mathbf{a})\,\widetilde{E}(\mathbf{a})\,\right|^2\right\rangle_t.
-\label{eqn:dirty-image-MOFF}
-%\end{align}
-%The term inside the angular brackets before squaring has a very similar form as
-%that in equation~\ref{eqn:dirty-image-FX}. It signifies that the measured antenna
-%electric fields are weighted by the antenna noise, weighted and gridded by the
-%antenna aperture kernel, Fourier transformed and finally squared to obtain the
-%same image estimated that would have been obtained using equation~
-%\ref{eqn:dirty-image-FX}.
-
-%Equations~\ref{eqn:dirty-image-MOFF} and \ref{eqn:dirty-image-FX} are identical
-%to each other obtained as a consequence of the multiplication-convolution theorem
-%of Fourier transforms. In other words, equation~\ref{eqn:dirty-image-FX} is
-%obtained by a Fourier transform of the gridded spatial correlation of electric
-%fields measured by antennas while equation~\ref{eqn:dirty-image-MOFF} is obtained
-%by squaring the Fourier transform of the gridded electric fields measured by the
-%antennas. The time-scale of averaging in equations~\ref{eqn:cc-vis} and
-%\ref{eqn:dirty-image-MOFF} is set by the coherence time-scale of the electric
-%fields or the science requirement (e.g., timescale of variability or transients).
-%
-%There are some important differences between the two techniques:
-%\begin{enumerate}
-%\item The time-averaging cannot be performed on a stochastic measurement but
-%  only on its statistical properties. In FX imaging, the visibilities measured
-%  between antenna pairs represent spatial correlations which can be time-averaged
-%  following which the they gridded and imaged. However, in MOFF imaging both
-%  antenna and gridded electric fields are stochastic and therefore have to be
-%  imaged before time-averaging. 
-%\item In FX imaging, electric fields measured on antennas are not correlated with
-%  themselves and hence lack zero spacing measurements. In contrast, in MOFF
-%  imaging, since the gridded electric fields are imaged and squared, they
-%  contain information from auto-correlated electric fields at zero spacing.
-%  Hence, they have to be subtracted from the images.
-%\end{enumerate}
-
 \section{Software Implementation}\label{sec:software}
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{MOFF_flowchart.eps}
+  \includegraphics[width=\columnwidth]{MOFF_flowchart}
+  % \includegraphics[width=\columnwidth]{MOFF_flowchart.eps}
   \caption{A flowchart of MOFF imaging in EPIC. The propagated electric fields
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{FX_flowchart.eps}
+  \includegraphics[width=\columnwidth]{FX_flowchart}
+  % \includegraphics[width=\columnwidth]{FX_flowchart.eps}
   \caption{A flowchart of visibility based software holographic imaging in EPIC. 
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{a2g_mapping.eps}
+  \includegraphics[width=\columnwidth]{a2g_mapping}
+  % \includegraphics[width=\columnwidth]{a2g_mapping.eps}
   \caption{Block diagram of an antenna-to-grid mapping. A sparse block-diagonal
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{F-engine.eps}
+  \includegraphics[width=\columnwidth]{F-engine}
+  % \includegraphics[width=\columnwidth]{F-engine.eps}
   \caption{Block diagram of a F-engine. The electric field data streams from
@@
 \end{align}
-% The sentence below was already stated earlier.
-%In practice the grid is set to be a power of 2 such that the grid spacing is at
-%most $\lambda/2$ even at the highest frequency in the observing band.
 
@@
 \begin{figure*}
-  \includegraphics[width=\linewidth]{EPIC-modules.eps}
+  \includegraphics[width=\linewidth]{EPIC-modules}
+  % \includegraphics[width=\linewidth]{EPIC-modules.eps}
   \caption{Software architecture of EPIC with core modules, their essential
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{autocorr_uvwts_pbeam.png}
+  \includegraphics[width=\columnwidth]{autocorr_uvwts_pbeam}
+  % \includegraphics[width=\columnwidth]{autocorr_uvwts_pbeam.png}
   \caption{The auto-correlation of weights of a square shaped antenna aperture
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{MOFF_FX_comparison_10_random_source_positions_4_iterations_test_aperture_zoomed.eps}
+  \includegraphics[width=\columnwidth]{MOFF_FX_comparison_10_random_source_positions_4_iterations_test_aperture_zoomed}
+  % \includegraphics[width=\columnwidth]{MOFF_FX_comparison_10_random_source_positions_4_iterations_test_aperture_zoomed.eps}
   \caption{Dirty images (top) and synthesized beams (bottom) obtained from 
@@
 
-% ----
-
-% Fig.~\ref{fig:MOFF-FX-uvwts} shows the cross-correlation weights obtained 
-% with MOFF imaging (left) and visibility based imaging (right). The first 
-% notable difference is in the weights around zero spacing. Though both show a
-% dominant void around zero-spacing, the void obtained with MOFF algorithm shows 
-% many pixels with non-zero weights. In contrast, the zero-spacing void from 
-% traditional imaging consists of predominantly zero-valued pixels. The gridding
-% process in the former involves tying the antenna footprint (either in 
-% analytic or lookup table formats) to the nearest grid pixel. Depending on the 
-% exact location of the center of the antenna relative to the grid and the 
-% coarseness of grid spacing, grid pixels that receive contribution from an 
-% antenna may be a pixel narrower along one or both axes relative to that from 
-% another identical antenna but with a different center location relative to the 
-% grid. The resulting auto-correlation footprint will also not necessarily be 
-% identical. Hence, using a single expected footprint for subtraction will 
-% typically leave some residuals behind as seen in the void region in the weights 
-% obtained with MOFF imaging.
-
-% These residuals can be mitigated by: 
-% \begin{enumerate}
-%   \item making the grid spacing finer which makes the nearest neighbour 
-%     association between the antenna footprint and the grid less susceptible 
-%     (proportional to grid spacing) to the location of the antenna center 
-%     relative to the grid, and
-%   \item subtracting each auto-correlation of antenna weights separately by 
-%     using the shape and extent of the footprint appropriate for that specific
-%     antenna aperture. 
-% \end{enumerate}
-% The latter is the most general solution applicable especially in the case of 
-% heterogeneous antenna arrays and is under active development for EPIC.
-
-% The other notable difference is that weights outside the zero-spacing void in
-% some regions are different from each other at the few percent level though the 
-% sum of weights in these regions are identical. For instance, note the 
-% difference in weights at $(u,v)\approx (2,-10)$. This is found to arise 
-% because with MOFF imaging the antenna locations were tied to the nearest 
-% grid pixel whereas in visibility based imaging the baseline locations 
-% (difference between antenna locations) were tied to the nearest grid pixel. 
-% Since rounding is not a commutative operation (i.e. rounding of the difference 
-% value is not necessarily equal to the difference of the rounded values), the 
-% gridding operation in the two cases introduces rounding errors in the placement 
-% of antenna aperture and cross-correlated aperture weights. Each antenna 
-% aperture weight projected to the nearest grid location can be displaced by 
-% $\sim 1$--2 pixels. This effect can be mitigated only by making the grid 
-% spacing finer at the expense of increased computational cost.
-
-% ----
-
 \begin{figure}
-  \includegraphics[width=\columnwidth]{MOFF_FX_comparison_uvwts_test_aperture_zero_spacing_removed.eps}
+  \includegraphics[width=\columnwidth]{MOFF_FX_comparison_uvwts_test_aperture_zero_spacing_removed}
+  % \includegraphics[width=\columnwidth]{MOFF_FX_comparison_uvwts_test_aperture_zero_spacing_removed.eps}
   \caption{Comparison of $uv$ plane antenna power response with MOFF (left) and 
@@
 \end{figure}
-
-% The synthesized beam can be written as:
-% \begin{align}\label{eqn:syn-beam}
-%   B(\hat{\mathbf{s}}) &= \left| \sum_\mathbf{r} \widetilde{W}(\mathbf{r})\,e^{-i2\pi\mathbf{r}\cdot\hat{\mathbf{s}}}\right|^2, \\
-%   \textrm{or}, \quad B(l,m) &= \left| \sum_{r_x,r_y} \widetilde{W}(r_x,r_y)\,e^{-i2\pi(r_x l + r_y m)}\right|^2 
-% \end{align}
-% where, $\hat{\mathbf{s}}\equiv (l,m)$, $\mathbf{r}\equiv (r_x,r_y)$ and 
-% $\widetilde{W}(\mathbf{r})$ is given by equation~\ref{eqn:gridding-convolution}.
-% If the errors in the determination of $r_x$ and $r_y$ are denoted by 
-% $\delta r_x$ and $\delta r_y$ respectively, the error on $B(l,m)$ can be 
-% estimated using a simple propagation of errors, where,
-% \begin{align}
-%   \frac{\partial{B}(l,m)}{\partial{r_x}} &= e^{-i2\pi\mathbf{r}\cdot\hat{\mathbf{s}}}\left[\frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_x}} - i 2\pi l\widetilde{W}(\mathbf{r})\right] \label{eqn:syn-beam-pdev1}\\
-%   &\qquad\qquad \times \sum_{\mathbf{r}^\prime} \widetilde{W}^\star(\mathbf{r}^\prime)\,e^{+i2\pi\mathbf{r}^\prime\cdot \hat{\mathbf{s}}} \\
-%   &= e^{-i2\pi\mathbf{r}\cdot\hat{\mathbf{s}}}\left[\frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_x}} - i2\pi l\widetilde{W}(\mathbf{r})\right]\,W^\star(\hat{\mathbf{s}}), \\
-%   \textrm{and},\frac{\partial{B}(l,m)}{\partial{r_y}} &= e^{-i2\pi\mathbf{r}\cdot\hat{\mathbf{s}}}\left[\frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_x}} - i2\pi m\widetilde{W}(\mathbf{r})\right]\,W^\star(\hat{\mathbf{s}}).
-\label{eqn:syn-beam-pdev2}
-% \end{align}
-% It must be noted that although we start with a squaring operation in 
-% equation~\ref{eqn:syn-beam}, in equations~\ref{eqn:syn-beam-pdev1} through 
-% \ref{eqn:syn-beam-pdev2}, the summations are performed only when 
-% $\mathbf{r}\ne \mathbf{r}^\prime$ and therefore do not include the 
-% auto-correlations of the antenna weights. The error expected in $B(l,m)$ is:
-% \begin{align}
-%   \left|\delta B(l,m)\right|^2 &= \sum_{r_x}\left|\frac{\partial{B}(l,m)}{\partial{r_x}}\right|^2 \delta r_x^2 + \sum_{r_y}\left|\frac{\partial{B}(l,m)}{\partial{r_y}}\right|^2 \delta r_y^2 \\
-%   &= \delta r_x^2 \sum_{r_x}\left|\frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_x}} - i2\pi l\widetilde{W}(\mathbf{r})\right|^2\,\left|W(\hat{\mathbf{s}})\right|^2 \nonumber\\
-%   &\qquad + \delta r_y^2 \sum_{r_y}\left|\frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_y}} - i2\pi m\widetilde{W}(\mathbf{r})\right|^2\,\left|W(\hat{\mathbf{s}})\right|^2
-% \end{align}
-% For the current simulations, 
-% \begin{align}
-%   \delta r_x = \delta r_y &\simeq 1, \\
-%   \widetilde{W}(r_x,r_y) &= \{0, \textrm{constant}\}, \\
-%   \textrm{and},\qquad \frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_x}} &= \frac{\partial{\widetilde{W}(r_x,r_y)}}{\partial{r_y}} = 0.
-% \end{align}
-% Hence,
-% \begin{align}
-%   \left|\delta B(l,m)\right|^2 &\simeq 4\pi^2\left(l^2+m^2\right)\,B(\hat{\mathbf{s}})\,\sum_\mathbf{r}\,\left|\widetilde{W}(\mathbf{r})\right|^2, \\
-% \textrm{or},\qquad \left|\delta B(l,m)\right| &\simeq 2\pi\sqrt{l^2+m^2}\,\left|W(\hat{\mathbf{s}})\right|\,\left(\sum_\mathbf{r}\,\left|\widetilde{W}(\mathbf{r})\right|^2\right)^{1/2}, \label{eqn:syn-beam-error}
-% \end{align}
-% where, $B(\hat{\mathbf{s}}) \equiv \left|W(\hat{\mathbf{s}})\right|^2$. Thus, 
-% the error on the synthesized beam is proportional to the antenna power pattern
-% and increases linearly outward in radial bins of direction cosines.
 
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{diff_psf_MOFF-FX_test_aperture.eps}
+  \includegraphics[width=\columnwidth]{diff_psf_MOFF-FX_test_aperture}
+  % \includegraphics[width=\columnwidth]{diff_psf_MOFF-FX_test_aperture.eps}
   \caption{Map of difference between the synthesized beams obtained with the 
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{LWA_MOFF_bandavg_image_4_iterations_analytic_aperture.eps}
+  \includegraphics[width=\columnwidth]{LWA_MOFF_bandavg_image_4_iterations_analytic_aperture}
+  % \includegraphics[width=\columnwidth]{LWA_MOFF_bandavg_image_4_iterations_analytic_aperture.eps}
   \caption{Image from LWA TBN data obtained with MOFF imaging using EPIC 
@@
 \begin{table}
+  \scriptsize
   \centering
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{MOFF_FX_computations_fov_gridding_annotated.eps}
+  \includegraphics[width=\columnwidth]{MOFF_FX_computations_fov_gridding_annotated}
+  % \includegraphics[width=\columnwidth]{MOFF_FX_computations_fov_gridding_annotated.eps}
   \caption{Current and planned instruments in parameter space of
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{MOFF_FX_crossover_baseline_n-antennas_rho_fov_gridding_legended.eps}
+  \includegraphics[width=\columnwidth]{MOFF_FX_crossover_baseline_n-antennas_rho_fov_gridding_legended}
+  % \includegraphics[width=\columnwidth]{MOFF_FX_crossover_baseline_n-antennas_rho_fov_gridding_legended.eps}
   \caption{Current and planned instruments in parameter space of
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 %%%%%%%%%%%%%%%%%%%% REFERENCES %%%%%%%%%%%%%%%%%%
 
-% The best way to enter references is to use BibTeX:
-
-\bibliographystyle{../mnras}
-\bibliography{../epic} % if your bibtex file is called example.bib
-
-% Alternatively you could enter them by hand, like this:
-% This method is tedious and prone to error if you have lots of references
-% \begin{thebibliography}{99}
-% \bibitem[\protect\citeauthoryear{Author}{2012}]{Author2012}
-% Author A.~N., 2013, Journal of Improbable Astronomy, 1, 1
-% \bibitem[\protect\citeauthoryear{Others}{2013}]{Others2013}
-% Others S., 2012, Journal of Interesting Stuff, 17, 198
-% \end{thebibliography}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\bibliographystyle{mnras}
+% \bibliography{../epic} 
+\input{ms.bbl}
 

### 2015-10-29 imaging_demo/ms.tex
Fix minor errors in previous version and resubmit
From f6809de (arxiv):
====================

In ms.tex,

(*) Mention calibration scales as number of calibrators as well as
number of antennas
(*) Correct the collecting area of HIRAX in the table
(*) Change caption of figure 9 to left and right instead of top and
bottom
(*) Update caption of figure 7

@@
 non-identical antenna beam patterns. Because only a single correlation is needed 
-for each antenna, the computation complexity scales only as $N_{\mathrm{a}}$.
+for each antenna, the computation complexity scales only as $\sim N_{\mathrm{a}}N_\textrm{c}$, where $N_\textrm{c}$ is the number of calibrators in the field.
 
@@
   \caption{Dirty images (top) and synthesized beams (bottom) obtained from 
-    simulated data using antenna-based MOFF algorithm (left) and FX 
-    visibility-based software holography (right). The solid black circles in the
+    simulated data using EPIC implementation of antenna-based MOFF algorithm 
+    (left) and visibility-based imaging (right). The solid black circles in the
     top panels indicate the simulated source positions. The antenna 
@@
   \caption{Map of difference between the synthesized beams obtained with the 
-    two methods (top) and radial statistics of the synthesized beams and 
-    their differences (bottom). The radial variations of rms of the synthesized 
+    two methods (left) and radial statistics of the synthesized beams and 
+    their differences (right). The radial variations of rms of the synthesized 
     beam (gray) and that of the rms of the difference (black) are shown as 
@@
     CHIME & 100 & 1280 & 8 & 600 \\
-    HIRAX\tnote{g} & 200 & 1024 & 6 & 600 \\
+    HIRAX\tnote{g} & 200 & 1024 & 28 & 600 \\
     \hline

### 2015-11-06 imaging_demo/ms.tex
Update abstract with minor suggestions from Adam
From 70146f3 (master):
=====================

In ms.tex,

(*) Update abstract with suggestions from Adam

--- 2015-10-28 imaging_demo/ms.tex
@@
 we present the first software demonstration of a generalized direct imaging 
-algorithm known as the Modular Optimal Frequency Fourier (MOFF) imager. It takes 
+algorithm, namely, the Modular Optimal Frequency Fourier (MOFF) imager. It takes 
 advantage of the multiplication-convolution theorem of Fourier transforms. Not 
@@
 the images produced to be equivalent to those produced using traditional 
-techniques. We have also validated our implementation on data observed with the 
-Long Wavelength Array (LWA). Antenna systems with a dense filling factor 
-consisting of a large number of antennas such as LWA, the Square Kilometre Array, 
-Hydrogen Epoch of Reionization Array, and Canadian Hydrogen Intensity Mapping 
-Experiment will gain significant advantage by deploying EPIC. Inherent 
-availability of calibrated time-domain images on timescales roughly equal to the 
-writeout timescale of the digitizer and vastly lower bandwidth relative to 
-visibility based systems will make it a prime candidate for transient searches 
-of Fast Radio Bursts (FRB) as well as planetary and exoplanetary phenomena. 
+techniques to within a precision determined by coarseness of gridding. We have 
+also validated our implementation on data observed with the Long Wavelength Array 
+(LWA). Antenna layouts with a dense filling factor consisting of a large number 
+of antennas such as LWA, the Square Kilometre Array, Hydrogen Epoch of 
+Reionization Array, and Canadian Hydrogen Intensity Mapping Experiment will gain 
+significant computational advantage by deploying EPIC. Inherent availability of 
+calibrated time-domain images on timescales roughly equal to the writeout 
+timescale of the digitizer and vastly lower I/O bandwidth relative to visibility 
+based systems will make it a prime candidate for transient searches of Fast Radio 
+Bursts (FRB) as well as planetary and exoplanetary phenomena. 
 \end{abstract}

### 2015-11-06 imaging_demo/ms.tex

--- 2015-10-29 imaging_demo/ms.tex
@@
 % out the following line
-% \usepackage{newtxtext,newtxmath}
+\usepackage{newtxtext,newtxmath}
 % Depending on your LaTeX fonts installation, you might get better results with one of these:
@@
 non-identical antenna beam patterns. Because only a single correlation is needed 
-for each antenna, the computation complexity scales only as $\sim N_{\mathrm{a}}N_\textrm{c}$, where $N_\textrm{c}$ is the number of calibrators in the field.
+for each antenna, the computation complexity scales only as 
+$\sim N_{\mathrm{a}}N_\textrm{c}$, where $N_\textrm{c}$ is the number of calibrators 
+in the field.
 
@@
 \bibliographystyle{mnras}
-% \bibliography{../epic} 
-\input{ms.bbl}
+\bibliography{../epic} 
+% \input{ms.bbl}
 

### 2015-11-06 imaging_demo/ms.tex
Update section 1 with changes suggested by Adam
From a01e819 (master):
=====================

In ms.tex,

(*) Update section 1 with changes suggested by Adam

@@
 technology development towards arrays of low-cost antennas with large fields of
-view and densely packed apertures. Similarly, the search for transient objects
+view and densely packed layouts. Similarly, the search for transient objects
 and regular monitoring of the time-dependent sky is driving instruments in the
 same direction with the added requirement of fast read-outs. A number of new 
-telescopes are under development around the world based on this new paradigm, 
-including the Murchison Widefield Array (MWA;\citealt{tin13,bow13}), the 
-Precision Array for Probing the Epoch of Reionization (PAPER; \citealt{par10}), 
-the Hydrogen Epoch of Reionization Array\footnote{http://reionization.org} 
-(HERA), the LOw Frequency ARray (LOFAR;\citealt{van13}), the Canadian Hydrogen 
-Intensity Mapping Experiment (CHIME,\citealt{ban14}), the Long Wavelength 
-Array (LWA, \citealt{ell13}), and the low frequency Square Kilometer Array 
-(SKA1-Low \citealt{mel13}).
+telescopes are being or were developed around the world based on this new 
+paradigm, the Hydrogen Epoch of Reionization 
+Array\footnote{http://reionization.org} (HERA), including the Murchison 
+Widefield Array (MWA;\citealt{tin13,bow13}), the Precision Array for Probing 
+the Epoch of Reionization (PAPER; \citealt{par10}), the LOw Frequency ARray 
+(LOFAR;\citealt{van13}), the Canadian Hydrogen Intensity Mapping Experiment 
+(CHIME,\citealt{ban14}), the Long Wavelength Array (LWA, \citealt{ell13}), and 
+the low frequency Square Kilometer Array (SKA1-Low \citealt{mel13}).
 
@@
 computational scaling from the harsh $\Nant^2$ to the more gentle envelope of
-$\Ngrid\log_2\Ngrid$, where $\Ngrid$ is the number of pixels in the Fourier 
-transform \citep[e.g.][]{mor11,teg09,teg10}. This architecture is often referred 
-to as a ``direct imaging" correlator because it eliminates the intermediary 
-cross-correlation data products of the FX and earlier lag correlators, but 
+$\sim\Ngrid\log_2\Ngrid$, where $\sim\Ngrid$ is the number of grid points in the 
+Fourier transform \citep[e.g.][]{mor11,teg09,teg10}. This architecture is often 
+referred to as a ``direct imaging'' correlator because it eliminates the 
+intermediate cross-correlation data products of the FX and XF correlators, but 
 instead directly forms images from the electric field measurements.
@@
 correlator. We also demonstrate the performance with real-world data from the 
-LWA. In \S \ref{sec:analysis} we analyze the scaling relationships of the 
-algorithm. We identify specific array design classes where the EPIC correlator 
-is computationally more efficient and in the field of transients, demands 
-significantly lesser bandwidth relative to visibility based approaches. We 
-conclude and discuss future research prospects in \S\ref{sec:conclusions}.
+LWA. In \S \ref{sec:analysis}, we explore the scalability of the algorithm in the 
+context of several array design choices. We identify specific array design 
+classes where the EPIC correlator -- is computationally more efficient; and in 
+the field of transients, demands significantly lesser I/O bandwidth relative to 
+visibility based approaches. We conclude and discuss future research prospects 
+in \S\ref{sec:conclusions}.
 

### 2015-11-06 imaging_demo/ms.tex
Update section 1 with changes suggested by Adam
From a01e819 (master):
=====================

In ms.tex,

(*) Update section 1 with changes suggested by Adam

@@
 context of several array design choices. We identify specific array design 
-classes where the EPIC correlator -- is computationally more efficient; and in 
-the field of transients, demands significantly lesser I/O bandwidth relative to 
+classes where the EPIC correlator is computationally more efficient and in the 
+field of transients, demands significantly lesser I/O bandwidth relative to 
 visibility based approaches. We conclude and discuss future research prospects 

### 2015-11-06 imaging_demo/ms.tex
Update section 2 with suggested changes
From b7d9d50 (review):
=====================

In ms.tex,

(*) Update section 2 with suggested changes

--- 2015-11-06 imaging_demo/ms.tex
@@
 \begin{align}
-  \widetilde{V}_p = &\left\langle \widetilde{E}_a\widetilde{E}_b^\star \right\rangle_t \label{eqn:cc-vis}\hfill\\
-  		= &\left\langle \left[ \int {W}_a(\hat{\mathbf{s}})\,E(\hat{\mathbf{s}})\,e^{-i2\pi\mathbf{r}_a\!\cdot\,\hat{\mathbf{s}}}\,\dif^2\hat{\mathbf{s}} + \widetilde{n}_a \right] \right. \nonumber\\
+  \widetilde{V}_p &= \left\langle \widetilde{E}_a\widetilde{E}_b^\star \right\rangle_t \label{eqn:cc-vis}\hfill\\
+  		&= \left\langle \left[ \int {W}_a(\hat{\mathbf{s}})\,E(\hat{\mathbf{s}})\,e^{-i2\pi\mathbf{r}_a\!\cdot\,\hat{\mathbf{s}}}\,\dif^2\hat{\mathbf{s}} + \widetilde{n}_a \right] \right. \nonumber\\
 		&\qquad \times \left.\left[ \int {W}^\star_b(\hat{\mathbf{s'}})\,E^\star (\hat{\mathbf{s'}})\,e^{i2\pi\mathbf{r}_b\!\cdot\,\hat{\mathbf{s'}}}\,\dif^2\hat{\mathbf{s'}} + \widetilde{n}^\star_b \right] \right\rangle_t \\
-                  =& \iint {W}_a(\hat{\mathbf{s}}) {W}^\star_b(\hat{\mathbf{s'}}) \left\langle E(\hat{\mathbf{s}}) E^\star(\hat{\mathbf{s'}}) \right\rangle_t e^{-i2\pi(\mathbf{r}_a\!\cdot\,\hat{\mathbf{s}}-\mathbf{r}_b\!\cdot\,\hat{\mathbf{s'}})}\,\dif^2\hat{\mathbf{s}}\,\dif^2\hat{\mathbf{s'}},
+                  &= \iint {W}_a(\hat{\mathbf{s}}) {W}^\star_b(\hat{\mathbf{s'}}) \left\langle E(\hat{\mathbf{s}}) E^\star(\hat{\mathbf{s'}}) \right\rangle_t e^{-i2\pi(\mathbf{r}_a\!\cdot\,\hat{\mathbf{s}}-\mathbf{r}_b\!\cdot\,\hat{\mathbf{s'}})}\,\dif^2\hat{\mathbf{s}}\,\dif^2\hat{\mathbf{s'}},
 \end{align}
@@
 \begin{align}\label{eqn:software-holography}
-  \widetilde{V}_p &= \int \widetilde{B}(\mathbf{u}^\prime-\mathbf{u}) \left[\int e^{-i2\pi\mathbf{u}.\hat{\mathbf{s}}}\,I(\hat{\mathbf{s}})\,\dif^2\hat{\mathbf{s}}\right]\dif^2\mathbf{u}\vphantom{\int} + n_p,
+  \widetilde{V}_p &= \int \widetilde{B}(\mathbf{u}^\prime-\mathbf{u}) \left[\int e^{-i2\pi\mathbf{u}.\hat{\mathbf{s}}}\,I(\hat{\mathbf{s}})\,\dif^2\hat{\mathbf{s}}\right]\dif^2\mathbf{u} + n_p,
 \end{align}
@@
 convolution with $\widetilde{B}(\mathbf{u})$ in $uv$-space. This is the software
-holographic equivalent of traditional FX correlator output.
+holographic equivalent of a traditional FX correlator output.
 
@@
 the gridding kernel, followed by a Fourier transform to create an image
-$\mathbf{I}^\prime(\hat{\mathbf{s}})$. This is the optimal estimate of the true image 
-$\mathbf{I}(\hat{\mathbf{s}})$ given the visibility measurements.
+$\mathbf{I}^\prime(\hat{\mathbf{s}})$. This is the optimal estimate of the true 
+image $\mathbf{I}(\hat{\mathbf{s}})$ given the visibility measurements.
 
 The intermediate step of gridding with the antenna power response can be 
-expressed as a convolution of a data vector generated by gridding the electric fields 
-directly with the antenna illumination pattern.
+expressed as a convolution of a data vector generated by gridding the electric 
+fields directly with the antenna illumination pattern.
 \begin{align}
-&\widetilde{\mathbf{B}}^{\,\textrm{T}}(\mathbf{u},\mathbf{v})\,\mathbf{N}^{-1}(\mathbf{v} ,\mathbf{v})\,\mathbf{m}(\mathbf{v})=\nonumber\\
-&\qquad\left\langle \left[\widetilde{\mathbf{W}}^\textrm{T}_a(\mathbf{r},\mathbf{a})\, \widetilde{\mathbf{N}}^{-1}\!(\mathbf{a},\mathbf{a})\, \widetilde{\mathbf{E}}(\mathbf{a})\right]
-\ast\left[\widetilde{\mathbf{W}}_a(\mathbf{r},\mathbf{a})\, \mathbf{N}^{-1}\!(\mathbf{a},\mathbf{a})\, \widetilde{\mathbf{E}}^\star(\mathbf{a})\right]\right\rangle_t\label{eqn:e-field-conv}
+\widetilde{\mathbf{B}}^{\,\textrm{T}}(\mathbf{u},\mathbf{v})\,\mathbf{N}^{-1}(\mathbf{v} ,\mathbf{v})\,\mathbf{m}(\mathbf{v}) &= \left\langle \left[\widetilde{\mathbf{W}}^\textrm{T}_a(\mathbf{r},\mathbf{a})\, \widetilde{\mathbf{N}}^{-1}\!(\mathbf{a},\mathbf{a})\, \widetilde{\mathbf{E}}(\mathbf{a})\right]\right.
+\nonumber\\
+&\quad\ast\left.\left[\widetilde{\mathbf{W}}_a(\mathbf{r},\mathbf{a})\, \mathbf{N}^{-1}\!(\mathbf{a},\mathbf{a})\, \widetilde{\mathbf{E}}^\star(\mathbf{a})\right]\right\rangle_t\label{eqn:e-field-conv}
 \end{align}
@@
 We can then use the multiplication-convolution theorem to move the convolution in 
-Equation \ref{eqn:e-field-conv} to a square after the Fourier transform in Equation 
+Equation \ref{eqn:e-field-conv} to a square after the Fourier transform in 
+Equation 
 \ref{eqn:dirty-image-FX}.
@@
 antenna aperture kernel, Fourier transformed and finally squared to obtain the
-same image estimated that would have been obtained using equation~
+same image estimate that would have been obtained using equation~
 \ref{eqn:dirty-image-FX}.

### 2015-11-06 imaging_demo/ms.tex
Update section 3 with suggested changes
From 3b460d9 (review):
=====================

In ms.tex,

(*) Update section 3 with suggested changes

@@
 EPIC can accept dual-polarization inputs and produce images of all four 
-cross-polarizations. Currently two data input formats exist for reading in the 
+instrumental cross-polarizations. Currently two data input formats exist for 
+reading in the 
 electric field time samples measured by the antennas -- simulated electric 
@@
 then cross-multiplied in the X-engine with those from all other antennas to 
-obtain the visibilities $V_\textrm{ij}(f)$ ($\mathbf{m}(\mathbf{v})$ in matrix 
+obtain the visibilities $V_\textrm{p}(f)$ ($\mathbf{m}(\mathbf{v})$ in matrix 
 notation). They are calibrated and time-averaged to obtain 
-$\langle V_\textrm{ij}(f)\rangle$ which are then gridded and imaged to obtain the 
+$\langle V_\textrm{p}(f)\rangle$ which are then gridded and imaged to obtain the 
 image $I(f)$. The $I(f)$ obtained from both techniques are theoretically 
@@
     F-engine, the electric fields pass through the X-engine to produce 
-    visibilities $V_\textrm{ij}(f)$ which are calibrated and time-averaged. Then 
+    visibilities $V_\textrm{p}(f)$ which are calibrated and time-averaged. Then 
     they are gridded to obtain the gridded visibilities $V_\textrm{g}(f)$ which 
@@
 locations are specified with a grid spacing. The grid spacing can be controlled 
-by the user. By default, it is set to be at $\lambda_\textrm{min}/2$ even at the 
-highest frequency to ensure there is no aliasing even from regions of the sky 
+by the user. By default, it is set to be at $\lambda/2$ even at the 
+lowest wavelength to ensure there is no aliasing even from regions of the sky 
 far away from the field of view. The number of locations on the grid is 
-restricted to be a power of 2.
+restricted to be a power of 2 for efficient use of FFT.
 
@@
 with only $\Nant$ blocks and roughly $N_\textrm{k}$ non-zero entries per
-block. This sparse matrix is stored in a Compressed Sparse Row (CSR) format. 
-Fig.~\ref{fig:a2g-mapping} illustrates the antenna-to-grid mapping
+block, where $N_\textrm{k}$ is the number of grid points that fall inside an
+antenna's footprint. This sparse matrix is stored in a Compressed Sparse Row 
+(CSR) format. Fig.~\ref{fig:a2g-mapping} illustrates the antenna-to-grid mapping
 matrix and the grid containing the mapped aperture footprints of the antennas.
@@
 the MOFF algorithm places no restriction on array layout, and accounts for 
-non-identical antenna beam patterns. Because only a single correlation is needed 
-for each antenna, the computation complexity scales only as 
-$\sim N_{\mathrm{a}}N_\textrm{c}$, where $N_\textrm{c}$ is the number of calibrators 
-in the field.
-
-The calibration module included in the EPIC repository allows for application of 
-pre-determined calibration solutions, or can solve for the complex gains using 
-the EPICal algorithm.
+non-identical antenna beam patterns. Direction dependent calibration can be 
+achieved by correlating antenna signals with output pixels in the direction of 
+$N_\textrm{c}$ calibration sources, then fitting for a functional model of the 
+sky. Since antennas are only correlated with calibrator pixels, the 
+computational complexity scales as $\sim N_\textrm{a} N_\textrm{c}$. % And since 
+% only a single correlation is needed for each antenna, the computation complexity 
+% scales only as $\sim N_{\mathrm{a}}N_\textrm{c}$, where $N_\textrm{c}$ is the 
+% number of calibrators in the field.
+
+The calibration module included in EPIC allows for application of pre-determined 
+calibration solutions, or can solve for the complex gains using the EPICal 
+algorithm.
 
@@
 padded with zeros in order to match the grid size and angular size of each image 
-pixel that would have been obtained with software holography of output from an FX
-correlator.
+pixel that would have been obtained with the software holography output from an 
+FX correlator.
 
@@
 The other function (not shown in the figure) is to update the data as new data 
-keeps streaming in. This can also be parallelized. Another important attribute
+streams in. This can also be parallelized. Another important attribute
 (not shown in figure) are antenna flags for each polarization appropriate for 
@@
 from the pair. Its primary attributes are the two antenna modules. It also
-contains four cross-polarized visibility time-series and spectra.
+contains four cross-polarized visibility time-series (even for the FX 
+correlator for diagnostic purposes) and spectra.
 
@@
 auto-correlations centered around the zero-spacing pixel in the $uv$ plane. 
-It also handles all four cross-polarization products. Currently, it can 
-support writing data out in standard FITS format.
+It also handles all four cross-polarization products. Currently, it supports 
+writing data out in standard FITS format.
 

### 2015-11-06 imaging_demo/ms.tex
Update sections 4, 5 and 6 with suggested changes
From 137f5a4 (review):
=====================

(*) Update figure 12 (layouts in parameter space)

In ms.tex,

(*) Update sections 4, 5 and 6 with suggested changes

@@
 imaging. In order to facilitate a robust comparison between MOFF and visibility 
-based imaging techniques, these auto-correlations are removed from both the MOFF 
+based imaging techniques, these auto-correlations are removed from the MOFF 
 algorithm output, which is otherwise not an essential part of the core algorithm. 
@@
 We investigate the two imaging algorithms for differences from the point of 
-view of the quality of their outputs. We begin by comparing the gridded 
-cross-correlation weights in the $uv$-plane. In MOFF imaging, weights 
-from antenna auto-correlations have been removed as described in 
-\S\ref{sec:rm-autocorr}.
+view of the quality of their outputs. We begin by comparing the images produced 
+with the two approaches.
 
@@
 algorithms packaged in EPIC. The antenna auto-correlations that correspond to 
-zero spacing have been removed from the MOFF image and the corresponding 
-synthesized beam. The sky positions of the simulated sources are indicated by 
-solid black circles. The reconstructed sky image has the simulated sources at the
-expected sky positions in either case. Both algorithms result in images and 
-synthesized beams that are well matched with each other. Their fluxes are 
-modulated by a multiplicative power pattern corresponding to that of a uniform
-square aperture.
+zero spacing have been removed from the correlated weights and data in the 
+$uv$ plane, the MOFF image and the corresponding synthesized beam as 
+described in \S\ref{sec:rm-autocorr}. The sky positions of the simulated sources 
+are indicated by solid black circles. The reconstructed sky image has the 
+simulated sources at the expected sky positions in either case. Both algorithms 
+result in images and synthesized beams that are well matched with each other. 
+Their fluxes are modulated by a multiplicative power pattern corresponding to 
+that of a uniform square aperture.
 
@@
 
+Now we investigate the gridded cross-correlation weights in the $uv$-plane. 
 Fig.~\ref{fig:MOFF-FX-uvwts} shows the correlated weights obtained with MOFF 
@@
 Here we demonstrate our software using narrow band data from the LWA station 
-in New Mexico. This data is in LWA narrow-band transient buffer (TBN) format with 
-512 voltage time samples from 255 antennas within roughly a diameter of 100~m. 
-The data is centered at a frequency of 74.03~MHz, with a sample rate (equal to 
-the bandwidth) of 100~kHz with 512 complex time samples in a A/D writeout 
-timescale of 5.12~ms, a frequency resolution of 195.3125~Hz and dual 
-polarization. There are 391 such time stamps yielding a total duration of 2~s.
+in New Mexico. This data is in LWA narrow-band transient buffer (TBN) format 
+from 255 antennas within roughly a diameter of 100~m. The data is centered at a 
+frequency of 74.03~MHz, with a sample rate (equal to the bandwidth) of 100~kHz 
+with 512 complex time samples per antenna in a A/D writeout timescale of 
+5.12~ms, a frequency resolution of 195.3125~Hz and dual polarization. There are 
+391 such writeouts (or time stamps; each contains 512 time samples at 100~kHz 
+sampling ) yielding a total duration of 2~s.
 
@@
 Fig.~\ref{fig:LWA-image} shows the image produced with MOFF imaging 
-packaged in EPIC after averaging over 20~ms of data and the inner
-$\approx 80$\% of bandwidth (roughly 80~kHz). The image is shown in direction 
-cosine coordinates -- $l$ along the horizontal axis and $m$ along the vertical 
-axis. The flux scale is arbitrary. Even in this proof-of-concept demonstration, 
-we see Cyg~A and Cas~A prominently as annotated, thus validating the 
-functionality of the EPIC package.
+packaged in EPIC after averaging over $\approx$~20~ms (four writeouts) of data 
+and the inner $\approx 80$\% of bandwidth (roughly 80~kHz). The image is shown 
+in direction cosine coordinates -- $l$ along the horizontal axis and $m$ along 
+the vertical axis. The flux scale is arbitrary. Even in this proof-of-concept 
+demonstration, we see Cyg~A and Cas~A prominently as annotated, thus validating 
+the functionality of the EPIC package.
 
@@
 Of all steps in the MOFF pipeline that are repeated for every writeout from 
-the F-engine, the slowest step even for dense HERA layouts is found to be the 
-spatial two-dimensional FFT in the imaging stage relative to applying the sparse 
-matrix gridding convolution, squaring or time-averaging. For instance, even in 
-the dense array layout scenario that makes these stages perform the slowest, the 
-gridding convolution, squaring and time-averaging take up only $\lesssim 20$\%, 
-$\lesssim 20$\% and $\lesssim 5$\% respectively of the total time taken by all 
-these processes while the spatial Fourier transform takes up $\gtrsim 55\%$ of
-the total time. With sparser arrays the gridding process will be be even faster.
-
-In the visibility based imaging, the predominant computational cost is at
-the X-engine requiring $\Nant(\Nant-1)/2$ complex multiplications 
-per channel at every correlator writeout timescale.
+the F-engine, the most expensive step even for dense HERA layouts is found to be 
+the spatial two-dimensional FFT in the imaging stage relative to applying the 
+sparse matrix gridding convolution, squaring or time-averaging. For instance, 
+even in the conservative dense array layout scenario that makes these other 
+stages even more expensive, the gridding convolution, squaring and time-averaging 
+take up only $\lesssim 20$\%, $\lesssim 20$\% and $\lesssim 5$\% respectively of 
+the total processing time while the spatial Fourier transform takes up 
+$\gtrsim 55\%$ of the total time. With sparser arrays the gridding process will 
+be be even faster.
+
+In visibility based imaging, the predominant computational cost is at the 
+X-engine requiring $\Nant(\Nant-1)/2$ complex multiplications per channel at 
+every A/D writeout timescale.
 
@@
 imaging is computationally more efficient than visibility based imaging. All HERA 
-layouts except HERA-19 and possibly HERA-37 are in a region of parameter space 
+layouts except HERA-19 and HERA-37 are in a region of parameter space 
 where MOFF imaging holds the advantage. The dashed line showing future trajectory 
@@
 indicates MOFF imaging is favored for those telescopes and vice versa. For 
-e.g., MOFF imaging is favored in HERA-33 and HERA-6769 because they are inside 
+e.g., MOFF imaging is favored in HERA-331 and HERA-6769 because they are inside 
 the wedge but not so in cases of HERA-19 and HERA-37. A majority of the 
 next-generation radio telescopes, namely, HERA-331 and its future expanded 
-versions, SKA1-LC, SKA1-LCD, HIRAX, CHIME, LWA1 and its future expansions 
-(LWA1-x2x1 and LWA1-x4x1.5 denote two-fold and four-fold increase in the number 
+versions, SKA1-LC, SKA1-LCD, HIRAX, and CHIME will fall in the regime where MOFF 
+imaging will be desirable. LWA1 and LWA-OV are already very close to the dividing 
+line. Their hypothetical expansions,\footnote{LWA1-x2x1 and LWA1-x4x1.5, and 
+LWA-OVx2x1 and LWA-OVx4x1.5 denote two-fold and four-fold increase in the number 
 of antennas within a core diameter that is 1 and 1.5 times the current size of 
-100~m respectively), and a couple of the expanded versions of LWA-OV (LWA-OVx2x1 
-and LWA-OVx4x1.5 denote two-fold and four-fold increase in the number of antennas
-within a core diameter that is 1 and 1.5 times the current size of 200~m 
-respectively) will fall in the regime where MOFF imaging will be desirable. For 
-a fixed baseline length, regions favoring the MOFF algorithm tend to be towards 
-large $\Nant$ indicating large-N dense array layouts with smaller antenna 
-elements are best suited for deploying EPIC.
+100~m and 200~m respectively for LWA1 and LWA-OV.} will be in the computational 
+regime favored by MOFF. For a fixed baseline length, regions favoring the MOFF 
+algorithm tend to be towards large $\Nant$ indicating large-N dense array 
+layouts with smaller antenna elements are best suited for deploying EPIC.
 
@@
     color falling into a wedge of the same color indicates MOFF imaging is 
-    advantageous for those telescopes and vice versa. For e.g., MOFF imaging
+    advantageous for those telescopes and vice versa. For example, MOFF imaging
     is favored in HERA-331 and HERA-6769 because they are inside the wedge 
@@
 and efficient EPIC software that is designed to bring this cost down from 
-$\mathcal{O}(N^2)$ to $\mathcal{O}(N\log_2 N)$. Under the class of direct imaging 
+$\mathcal{O}(\Nant^2)$ to $\mathcal{O}(\Ngrid\log_2 \Ngrid)$. Under the class of 
+direct imaging 
 techniques, ours is one of the most generic -- neither does it place any 
@@
 for real-life GPU/FPGA based implementation on telescopes. It has been 
-successfully tested on simulated as well as real LWA observations.
+successfully tested on simulated MWA observations as well as real LWA 
+observations.
 
@@
 
-A unique advantage is the instantaneous availability of calibrated time-domain 
-images bundled together with the hardware frontend such as the F-engine as an 
-integrated module. It is seen to have significant savings in data throughput 
-relative to a X-engine based pipeline. Hence, it is a compelling candidate for 
-time-domain radio astronomy, e.g. search for and monitoring of transients. 
-Transient detection pipeline at the backend of EPIC can be fine-tuned to target 
-fast transients such as the Fast Radio Bursts \citep[FRB;][]{tho13} on 
-millisecond timescales at GHz frequencies or slow transients from planetary and 
-exoplanetary origins at frequencies around 100~MHz.
+It is seen to have significant savings in data throughput relative to a 
+X-engine based pipeline. A unique advantage is the instantaneous availability 
+of calibrated time-domain images at no extra cost. Hence, it is a compelling 
+candidate for time-domain radio astronomy, e.g. search for and monitoring of 
+transients. Potentially, it could allow on-chip processing thus lowering even 
+further the relatively low I/O bandwidth shown in table~\ref{tab:data-rates} 
+except when a transient event is detected. Transient detection pipelines at 
+the backend of EPIC can be fine-tuned to target fast transients such as the 
+Fast Radio Bursts \citep[FRB;][]{tho13} on millisecond timescales at GHz 
+frequencies or slow transients from planetary and exoplanetary origins at 
+frequencies around 100~MHz.
 

### 2015-11-07 imaging_demo/ms.tex
Minor changes throughout, update figures
From 1e9f59c (review):
=====================

Update MOFF and FX flowcharts, F-engine

In ms.tex,

(*) Minor changes throughout

@@
 telescopes are being or were developed around the world based on this new 
-paradigm, the Hydrogen Epoch of Reionization 
-Array\footnote{http://reionization.org} (HERA), including the Murchison 
-Widefield Array (MWA;\citealt{tin13,bow13}), the Precision Array for Probing 
-the Epoch of Reionization (PAPER; \citealt{par10}), the LOw Frequency ARray 
+paradigm, including the Hydrogen Epoch of Reionization 
+Array\footnote{http://reionization.org} (HERA), the Murchison Widefield Array 
+(MWA;\citealt{tin13,bow13}), the Precision Array for Probing the Epoch of 
+Reionization (PAPER; \citealt{par10}), the LOw Frequency ARray 
 (LOFAR;\citealt{van13}), the Canadian Hydrogen Intensity Mapping Experiment 
@@
 computational scaling from the harsh $\Nant^2$ to the more gentle envelope of
-$\sim\Ngrid\log_2\Ngrid$, where $\sim\Ngrid$ is the number of grid points in the 
+$\sim\Ngrid\log_2\Ngrid$, where $\Ngrid$ is the number of grid points in the 
 Fourier transform \citep[e.g.][]{mor11,teg09,teg10}. This architecture is often 
@@
 Correlator'' -- a highly parallelized Object Oriented Python 
-package,\footnote{EPIC package can be accessed at https://github.com/nithyanandan/EPIC}
-now publicly available. Besides implementing the MOFF imaging algorithm it also
-includes visibility based imaging using the software holography technique and a 
-simulator for generating electric fields from a sky model.
+package,\footnote{The E-field Parallel Imaging Correlator (EPIC) package can be 
+accessed at https://github.com/nithyanandan/EPIC} now publicly available. Besides 
+implementing the MOFF imaging algorithm it also includes visibility based imaging 
+using the software holography technique and a simulator for generating electric 
+fields from a sky model.
 
@@
 temporal Fourier transform on the electric field time-series to obtain electric
-field spectra $E_1(f)\ldots E_\textrm{N}(f)$ 
+field spectra $\widetilde{E}_1(f)\ldots \widetilde{E}_\textrm{N}(f)$ 
 ($\widetilde{\mathbf{E}}(\mathbf{a})$ in matrix notation) for each of the 
@@
     Fourier transformed by the F-engine to produce electric field spectra 
-    $E_1(f)\ldots E_\textrm{N}(f)$. They are calibrated and gridded. The gridded 
-    electric fields $E_\textrm{g}(f)$ from each time series are imaged to produce 
-    images $I_1(f)\ldots I_\textrm{M}(f)$. These images are time-averaged to 
-    obtain the final image $I(f)$.}
+    $\widetilde{E}_1(f)\ldots \widetilde{E}_\textrm{N}(f)$. They are calibrated 
+    and gridded. The gridded electric fields $\widetilde{E}_\textrm{g}(f)$ from 
+    each time series are imaged to produce images 
+    $I_1(f)\ldots I_\textrm{M}(f)$. These images are time-averaged to obtain the 
+    final image $I(f)$.}
   \label{fig:MOFF-flowchart}
@@
 then cross-multiplied in the X-engine with those from all other antennas to 
-obtain the visibilities $V_\textrm{p}(f)$ ($\mathbf{m}(\mathbf{v})$ in matrix 
-notation). They are calibrated and time-averaged to obtain 
-$\langle V_\textrm{p}(f)\rangle$ which are then gridded and imaged to obtain the 
-image $I(f)$. The $I(f)$ obtained from both techniques are theoretically 
-identical as explained in \S\ref{sec:math}.
+obtain the visibilities $\widetilde{V}_\textrm{p}(f)$ ($\mathbf{m}(\mathbf{v})$ 
+in matrix notation). They are calibrated and time-averaged to obtain 
+$\langle \widetilde{V}_\textrm{p}(f)\rangle$ which are then gridded and imaged to 
+obtain the image $I(f)$. The $I(f)$ obtained from both techniques are 
+theoretically identical as explained in \S\ref{sec:math}.
 \begin{figure}
@@
     F-engine, the electric fields pass through the X-engine to produce 
-    visibilities $V_\textrm{p}(f)$ which are calibrated and time-averaged. Then 
-    they are gridded to obtain the gridded visibilities $V_\textrm{g}(f)$ which 
-    are then Fourier transformed to obtain the image, $I(f)$.}
+    visibilities $\widetilde{V}_\textrm{p}(f)$ which are calibrated and 
+    time-averaged. Then they are gridded to obtain the gridded visibilities 
+    $\widetilde{V}_\textrm{g}(f)$ which are then Fourier transformed to obtain 
+    the image, $I(f)$.}
   \label{fig:FX-flowchart}
@@
 locations are specified with a grid spacing. The grid spacing can be controlled 
-by the user. By default, it is set to be at $\lambda/2$ even at the 
+by the user. By default, it is set to be $\le\lambda/2$ even at the 
 lowest wavelength to ensure there is no aliasing even from regions of the sky 
@@
 attributes -- the propagated electric field time-series, $E(t)$, and spectrum 
-$E(f)$ for both polarizations. The most important function inside this module 
-is the F-engine that Fourier transforms time-series electric field data into 
-spectra.
+$\widetilde{E}(f)$ for both polarizations. The most important function inside 
+this module is the F-engine that Fourier transforms time-series electric field 
+data into spectra.
 
@@
 namely, the FX and XF modes. The FX mode obtains the electric field spectra, 
-$E(f)$ from the individual antenna modules inside this module and multiplies 
-the two to obtain visibility spectra, $V(f)$. On the other hand, the XF mode 
-cross-correlates the electric field time-series from its Antenna modules
-to obtain the visibilities as a function of lags, $V_t(t)$, which is then Fourier 
-transformed to obtain $V(f)$. Both modules can operate on dual-polarizations to
-obtain all four cross-polarizations.
+$\widetilde{E}(f)$ from the individual antenna modules inside this module and 
+multiplies the two to obtain visibility spectra, $\widetilde{V}(f)$. On the 
+other hand, the XF mode cross-correlates the electric field time-series from 
+its Antenna modules to obtain the visibilities as a function of lags, 
+$\widetilde{V}_t(t)$, which is then Fourier transformed to obtain 
+$\widetilde{V}(f)$. Both modules can operate on dual-polarizations to obtain 
+all four cross-polarizations.
 
@@
 cross-polarization for the current data stream. Similar to the antenna module,
-it has an update function that can update the visibilities $V_t(t)$ or $V(f)$
-directly rather than through the electric fields of its component antennas. 
-This functionality is to allow EPIC to operate while attached to the backend 
-of traditional correlator systems. This feature is not utilized for purposes 
-of this paper.
+it has an update function that can update the visibilities $\widetilde{V}_t(t)$ 
+or $\widetilde{V}(f)$ directly rather than through the electric fields of its 
+component antennas. This functionality is to allow EPIC to operate while 
+attached to the backend of traditional correlator systems. This feature is not 
+utilized for purposes of this paper.
 
@@
 391 such writeouts (or time stamps; each contains 512 time samples at 100~kHz 
-sampling ) yielding a total duration of 2~s.
+sampling) yielding a total duration of 2~s.
 
@@
 demonstration, we see Cyg~A and Cas~A prominently as annotated, thus validating 
-the functionality of the EPIC package.
+the functionality of EPIC.
 
@@
 
-We elaborate on the data rates required with the MOFF and visibility based
+We elaborate on the I/O data rates required with the MOFF and visibility based
 algorithms. This is particularly relevant in the context of radio transient 
@@
 transients. Potentially, it could allow on-chip processing thus lowering even 
-further the relatively low I/O bandwidth shown in table~\ref{tab:data-rates} 
-except when a transient event is detected. Transient detection pipelines at 
-the backend of EPIC can be fine-tuned to target fast transients such as the 
-Fast Radio Bursts \citep[FRB;][]{tho13} on millisecond timescales at GHz 
-frequencies or slow transients from planetary and exoplanetary origins at 
-frequencies around 100~MHz.
+further the already relatively low I/O bandwidth shown in 
+table~\ref{tab:data-rates} except when a transient event is detected. Transient 
+detection pipelines at the backend of EPIC can be fine-tuned to target fast 
+transients such as the Fast Radio Bursts \citep[FRB;][]{tho13} on millisecond 
+timescales at GHz frequencies or slow transients from planetary and exoplanetary 
+origins at frequencies around 100~MHz.
 

### 2015-11-07 imaging_demo/ms.tex
Move software architecture section to appendix
From cb584b3 (review):
=====================

In ms.tex,

(*) Move section 3.2 (software architecture) to appendix
(*) Remove section 3.1 heading "Data flow"
(*) Add a sentence at end of section 3 to refer the interested reader
to the appendix
(*) Update abstract
(*) Few minor changes

@@
 \begin{abstract}
-Modern radio telescopes are favoring densely packed array layouts consisting of 
+Modern radio telescopes are favoring densely packed array layouts with 
 large numbers of antennas ($\Nant\gtrsim 1000$). Since the complexity of 
@@
 $\mathcal{O}(\Nant\log_2\Nant)$ but can also image from irregularly arranged 
-heterogeneous antenna arrays. EPIC is highly modular and parallelizable, 
-implemented in object oriented Python, and publicly available. We have verified 
-the images produced to be equivalent to those produced using traditional 
-techniques to within a precision determined by coarseness of gridding. We have 
-also validated our implementation on data observed with the Long Wavelength Array 
-(LWA). Antenna layouts with a dense filling factor consisting of a large number 
-of antennas such as LWA, the Square Kilometre Array, Hydrogen Epoch of 
-Reionization Array, and Canadian Hydrogen Intensity Mapping Experiment will gain 
-significant computational advantage by deploying EPIC. Inherent availability of 
-calibrated time-domain images on timescales roughly equal to the writeout 
-timescale of the digitizer and vastly lower I/O bandwidth relative to visibility 
-based systems will make it a prime candidate for transient searches of Fast Radio 
-Bursts (FRB) as well as planetary and exoplanetary phenomena. 
+heterogeneous antenna. EPIC is highly modular and parallelizable, implemented in 
+object oriented Python, and publicly available. We have verified the images 
+produced to be equivalent to those produced using traditional techniques to 
+within a precision determined by coarseness of gridding. We have also validated 
+our implementation on data observed with the Long Wavelength Array (LWA). Antenna 
+layouts with a dense filling factor consisting of a large number of antennas such 
+as LWA, the Square Kilometre Array, Hydrogen Epoch of Reionization Array, and 
+Canadian Hydrogen Intensity Mapping Experiment will gain significant 
+computational advantage by deploying EPIC. Inherent availability of calibrated 
+time-domain images on digitizer writeout timescales and vastly lower I/O 
+bandwidth relative to visibility based systems will make it a prime candidate 
+for transient searches of Fast Radio Bursts (FRB) as well as planetary and 
+exoplanetary phenomena. 
 \end{abstract}
@@
 For certain classes of radio arrays there is an alternative to the FX correlator
-that can lower the computational burden by directly performing a spatial fast
-Fourier transform (FFT) on the electric fields measured by each antenna in the
-array at each time step, removing the cross-correlation step. This relieves the
-computational scaling from the harsh $\Nant^2$ to the more gentle envelope of
-$\sim\Ngrid\log_2\Ngrid$, where $\Ngrid$ is the number of grid points in the 
-Fourier transform \citep[e.g.][]{mor11,teg09,teg10}. This architecture is often 
-referred to as a ``direct imaging'' correlator because it eliminates the 
-intermediate cross-correlation data products of the FX and XF correlators, but 
-instead directly forms images from the electric field measurements.
+that can lower the computational burden by directly performing a spatial Fast
+Fourier Transform \citep[FFT;][]{coo65} on the electric fields measured by each 
+antenna in the array at each time step, removing the cross-correlation step. 
+This relieves the computational scaling from the harsh $\Nant^2$ to the more 
+gentle envelope of $\sim\Ngrid\log_2\Ngrid$, where $\Ngrid$ is the number of 
+grid points in the Fourier transform \citep[e.g.][]{mor11,teg09,teg10}. This 
+architecture is often referred to as a ``direct imaging'' correlator because it 
+eliminates the intermediate cross-correlation data products of the FX and XF 
+correlators, but instead directly forms images from the electric field 
+measurements.
 
@@
 
-\subsection{Data Flow}
+% \subsection{Data Flow}
 
@@
 
-\subsection{Software Architecture}\label{sec:software-modules}
-
-EPIC is built using object oriented programming in Python and is built on
-carefully crafted modules which closely represent real-life entities in radio 
-interferometer arrays and observations. The essential modules along with their
-key attributes and methods are illustrated in Fig.~\ref{fig:software-modules}.
-These modules are described below.
-
-\begin{figure*}
-  \includegraphics[width=\linewidth]{EPIC-modules}
-  % \includegraphics[width=\linewidth]{EPIC-modules.eps}
-  \caption{Software architecture of EPIC with core modules, their essential
-    attributes and functions. The antenna module forms the fundamental building 
-    block. It consists of electric field time-series and spectra and the 
-    F-engine that performs a temporal FFT to obtain electric field spectra from 
-    the time-series. The interferometer module is made of a pair of antenna 
-    modules. Its main function is the X-engine (FX or XF) to produce visibility 
-    spectra. The antenna array module is made of all individual antenna modules
-    as its components and contains collective properties about the antenna 
-    subsystems. Its core function is the creation of antenna-to-grid mapping, 
-    gridded aperture weights and electric fields. The interferometer array 
-    module is very similar in principle to the antenna array module except it 
-    operates on cross-correlations and produces gridded visibilities. The image 
-    module takes gridded electric fields or visibilities and performs a 
-    two-dimensional spatial FFT (and squares the intermediate image in case 
-    of the former) to produce output images. Broadly, the MOFF algorithm is 
-    implemented by modules below the horizontal dashed line while the visibility 
-    based imaging uses modules above the line. The exact processing pathway 
-    implementing the MOFF algorithm is shown in bolded modules.}
-  \label{fig:software-modules}
-\end{figure*}
-
-\subsubsection{Antenna Module}
-
-The antenna module is a fundamental building block upon which all the other 
-modules are built. There is one antenna module per antenna each having 
-attributes -- the propagated electric field time-series, $E(t)$, and spectrum 
-$\widetilde{E}(f)$ for both polarizations. The most important function inside 
-this module is the F-engine that Fourier transforms time-series electric field 
-data into spectra.
-
-The other function (not shown in the figure) is to update the data as new data 
-streams in. This can also be parallelized. Another important attribute
-(not shown in figure) are antenna flags for each polarization appropriate for 
-the data stream being held by the module.
-
-\subsubsection{Interferometer Module}
-
-The interferometer module holds the attributes and functions pertaining to
-a pair of antennas and represents the cross-correlation information obtained
-from the pair. Its primary attributes are the two antenna modules. It also
-contains four cross-polarized visibility time-series (even for the FX 
-correlator for diagnostic purposes) and spectra.
-
-The critical component of the interferometer module is the X-engine. This is 
-essentially a software analog of hardware correlators of real telescope 
-systems. The X-engine can be toggled between two states of operation, 
-namely, the FX and XF modes. The FX mode obtains the electric field spectra, 
-$\widetilde{E}(f)$ from the individual antenna modules inside this module and 
-multiplies the two to obtain visibility spectra, $\widetilde{V}(f)$. On the 
-other hand, the XF mode cross-correlates the electric field time-series from 
-its Antenna modules to obtain the visibilities as a function of lags, 
-$\widetilde{V}_t(t)$, which is then Fourier transformed to obtain 
-$\widetilde{V}(f)$. Both modules can operate on dual-polarizations to obtain 
-all four cross-polarizations.
-
-The other attributes (not shown in figure) are the flags applicable for each
-cross-polarization for the current data stream. Similar to the antenna module,
-it has an update function that can update the visibilities $\widetilde{V}_t(t)$ 
-or $\widetilde{V}(f)$ directly rather than through the electric fields of its 
-component antennas. This functionality is to allow EPIC to operate while 
-attached to the backend of traditional correlator systems. This feature is not 
-utilized for purposes of this paper.
-
-This module forms the fundamental unit for the interferometer array module 
-(to be discussed below) and in general for visibility based correlator and 
-imaging systems.
-
-\subsubsection{Antenna Array Module}
-
-The antenna array module consists of all the antenna modules as its attributes
-and represents the collective properties of its component antennas. By virtue of 
-holding each antenna data independently in their respective modules, the F-engine 
-for the entire array can be distributed to the F-engines of the component 
-antenna modules thus achieving a highly parallelized F-engine while emulating 
-real telescope systems.
-
-The primary attributes held by this module are the antenna aperture 
-illumination weights and electric fields projected on the grid using the 
-gridding convolution method described above and implemented by the gridding 
-function in this module. Significant parts of the antenna-to-grid mapping and 
-gridding convolution are parallelizable across antennas and frequencies.
-
-Individual antenna flags are carried over as additional weights to be applied
-to the gridded aperture illumination and electric fields. A series of data 
-streams can be stacked up to take advantage of the array optimization available 
-in Python. This module is also equipped to manage dual-polarization.
-
-\subsubsection{Interferometer Array Module}
-
-Similar to the antenna array module, the interferometer array module consists of
-individual interferometer modules. It can parallelize the correlator operations
-by distributing the X-operation over the X-engines of its component 
-interferometer modules. The interferometer-to-grid mapping and gridding 
-convolution are very similar in nature to that of the antenna array module. 
-Flag-based grid weights, stacking and ability to handle all four 
-cross-polarizations are built into this module.
-
-\subsubsection{Image Module}
-
-The image module is built as a general purpose module that can switch between 
-operating on gridded electric fields or visibilities. At its heart, it consists
-of a two-dimensional spatial FFT where the padding can be specified by the user 
-to control the resolution in the output images. In case of MOFF imaging, there 
-is an additional step of squaring the holographic electric field images.
-
-Besides its core functions of spatial Fourier transform and squaring, it can 
-stack, accumulate and average images, and optionally remove the antenna
-auto-correlations centered around the zero-spacing pixel in the $uv$ plane. 
-It also handles all four cross-polarization products. Currently, it supports 
-writing data out in standard FITS format.
+A high level software architecture of the EPIC package is described in the 
+appendix \S\ref{sec:software-modules} for the interested reader.
 
@@
 The number of complex multiplications and additions in the spatial Fourier 
-transform implemented via Fast Fourier Transform \citep[FFT;][]{coo65} is 
-$\approx \beta\,(4\Ngrid)\log_2(4\Ngrid)$ where $\Ngrid$ is the number of 
-pixels on the grid, the factor 4 accounts for increase in number of pixels as a 
-result of zero-padding before spatial Fourier transform, and $\beta$ is a 
-constant that depends on the implementation of twiddle FFT algorithms 
-\citep{bri74}. In our study, we set $\beta=5$, a 
+transform implemented via FFT is $\approx \beta\,(4\Ngrid)\log_2(4\Ngrid)$ 
+where $\Ngrid$ is the number of pixels on the grid, the factor 4 accounts for 
+increase in number of pixels as a result of zero-padding before spatial Fourier 
+transform, and $\beta$ is a constant that depends on the implementation of 
+twiddle FFT algorithms \citep{bri74}. In our study, we set $\beta=5$, a 
 value\footnote{http://www.fftw.org/speed/method.html} much more conservative 
@@
 
-% \appendix
+\appendix
+
+\section{Software Architecture}\label{sec:software-modules}
+
+EPIC is built using object oriented programming in Python and is built on
+carefully crafted modules which closely represent real-life entities in radio 
+interferometer arrays and observations. The essential modules along with their
+key attributes and methods are illustrated in Fig.~\ref{fig:software-modules}.
+These modules are described below.
+
+\begin{figure*}
+  \includegraphics[width=\linewidth]{EPIC-modules}
+  % \includegraphics[width=\linewidth]{EPIC-modules.eps}
+  \caption{Software architecture of EPIC with core modules, their essential
+    attributes and functions. The antenna module forms the fundamental building 
+    block. It consists of electric field time-series and spectra and the 
+    F-engine that performs a temporal FFT to obtain electric field spectra from 
+    the time-series. The interferometer module is made of a pair of antenna 
+    modules. Its main function is the X-engine (FX or XF) to produce visibility 
+    spectra. The antenna array module is made of all individual antenna modules
+    as its components and contains collective properties about the antenna 
+    subsystems. Its core function is the creation of antenna-to-grid mapping, 
+    gridded aperture weights and electric fields. The interferometer array 
+    module is very similar in principle to the antenna array module except it 
+    operates on cross-correlations and produces gridded visibilities. The image 
+    module takes gridded electric fields or visibilities and performs a 
+    two-dimensional spatial FFT (and squares the intermediate image in case 
+    of the former) to produce output images. Broadly, the MOFF algorithm is 
+    implemented by modules below the horizontal dashed line while the visibility 
+    based imaging uses modules above the line. The exact processing pathway 
+    implementing the MOFF algorithm is shown in bolded modules.}
+  \label{fig:software-modules}
+\end{figure*}
+
+\subsection{Antenna Module}
+
+The antenna module is a fundamental building block upon which all the other 
+modules are built. There is one antenna module per antenna each having 
+attributes -- the propagated electric field time-series, $E(t)$, and spectrum 
+$\widetilde{E}(f)$ for both polarizations. The most important function inside 
+this module is the F-engine that Fourier transforms time-series electric field 
+data into spectra.
+
+The other function (not shown in the figure) is to update the data as new data 
+streams in. This can also be parallelized. Another important attribute
+(not shown in figure) are antenna flags for each polarization appropriate for 
+the data stream being held by the module.
+
+\subsection{Interferometer Module}
+
+The interferometer module holds the attributes and functions pertaining to
+a pair of antennas and represents the cross-correlation information obtained
+from the pair. Its primary attributes are the two antenna modules. It also
+contains four cross-polarized visibility time-series (even for the FX 
+correlator for diagnostic purposes) and spectra.
+
+The critical component of the interferometer module is the X-engine. This is 
+essentially a software analog of hardware correlators of real telescope 
+systems. The X-engine can be toggled between two states of operation, 
+namely, the FX and XF modes. The FX mode obtains the electric field spectra, 
+$\widetilde{E}(f)$ from the individual antenna modules inside this module and 
+multiplies the two to obtain visibility spectra, $\widetilde{V}(f)$. On the 
+other hand, the XF mode cross-correlates the electric field time-series from 
+its Antenna modules to obtain the visibilities as a function of lags, 
+$\widetilde{V}_t(t)$, which is then Fourier transformed to obtain 
+$\widetilde{V}(f)$. Both modules can operate on dual-polarizations to obtain 
+all four cross-polarizations.
+
+The other attributes (not shown in figure) are the flags applicable for each
+cross-polarization for the current data stream. Similar to the antenna module,
+it has an update function that can update the visibilities $\widetilde{V}_t(t)$ 
+or $\widetilde{V}(f)$ directly rather than through the electric fields of its 
+component antennas. This functionality is to allow EPIC to operate while 
+attached to the backend of traditional correlator systems. This feature is not 
+utilized for purposes of this paper.
+
+This module forms the fundamental unit for the interferometer array module 
+(to be discussed below) and in general for visibility based correlator and 
+imaging systems.
+
+\subsection{Antenna Array Module}
+
+The antenna array module consists of all the antenna modules as its attributes
+and represents the collective properties of its component antennas. By virtue of 
+holding each antenna data independently in their respective modules, the F-engine 
+for the entire array can be distributed to the F-engines of the component 
+antenna modules thus achieving a highly parallelized F-engine while emulating 
+real telescope systems.
+
+The primary attributes held by this module are the antenna aperture 
+illumination weights and electric fields projected on the grid using the 
+gridding convolution method described above and implemented by the gridding 
+function in this module. Significant parts of the antenna-to-grid mapping and 
+gridding convolution are parallelizable across antennas and frequencies.
+
+Individual antenna flags are carried over as additional weights to be applied
+to the gridded aperture illumination and electric fields. A series of data 
+streams can be stacked up to take advantage of the array optimization available 
+in Python. This module is also equipped to manage dual-polarization.
+
+\subsection{Interferometer Array Module}
+
+Similar to the antenna array module, the interferometer array module consists of
+individual interferometer modules. It can parallelize the correlator operations
+by distributing the X-operation over the X-engines of its component 
+interferometer modules. The interferometer-to-grid mapping and gridding 
+convolution are very similar in nature to that of the antenna array module. 
+Flag-based grid weights, stacking and ability to handle all four 
+cross-polarizations are built into this module.
+
+\subsection{Image Module}
+
+The image module is built as a general purpose module that can switch between 
+operating on gridded electric fields or visibilities. At its heart, it consists
+of a two-dimensional spatial FFT where the padding can be specified by the user 
+to control the resolution in the output images. In case of MOFF imaging, there 
+is an additional step of squaring the holographic electric field images.
+
+Besides its core functions of spatial Fourier transform and squaring, it can 
+stack, accumulate and average images, and optionally remove the antenna
+auto-correlations centered around the zero-spacing pixel in the $uv$ plane. 
+It also handles all four cross-polarization products. Currently, it supports 
+writing data out in standard FITS format.
 

### 2015-11-08 imaging_demo/ms.tex
Make manuscript MNRAS-compatible
From 44f7935 (master):
=====================

(*) Rename figures to figure*.eps in git

In ms.tex,

(*) Rename figures to figure*.eps
(*) Use MNRAS style file and epic.bib in the current directory

@@
 % Basic setup. Most papers should leave these options alone.
-\documentclass[a4paper,fleqn,usenatbib]{../mnras}
+% \documentclass[a4paper,fleqn,usenatbib]{../mnras}
+\documentclass[a4paper,fleqn,usenatbib]{mnras}
 
@@
 \begin{abstract}
-Modern radio telescopes are favoring densely packed array layouts with 
+Modern radio telescopes are favouring densely packed array layouts with 
 large numbers of antennas ($\Nant\gtrsim 1000$). Since the complexity of 
@@
 computational advantage by deploying EPIC. Inherent availability of calibrated 
-time-domain images on digitizer writeout timescales and vastly lower I/O 
-bandwidth relative to visibility based systems will make it a prime candidate 
+time-domain images on digitizer writeout time-scales and vastly lower I/O 
+bandwidth relative to visibility-based systems will make it a prime candidate 
 for transient searches of Fast Radio Bursts (FRB) as well as planetary and 
@@
 spatial FFT. This process has been shown to theoretically produce a data product
-identical to images produced from traditional visibility based techniques.
+identical to images produced from traditional visibility-based techniques.
 
@@
 implements the MOFF imaging algorithm besides emulating real-life telescopes and 
-FX/XF correlators in software, and includes a visibility based imaging technique 
+FX/XF correlators in software, and includes a visibility-based imaging technique 
 for reference. It is intended to provide a development platform to test different 
 imaging approaches, characterize scaling relations and serve as a stepping stone 
-for real-life GPU/FPGA based implementation on telescopes.
+for real-life GPU/FPGA-based implementation on telescopes.
 
@@
 the field of transients, demands significantly lesser I/O bandwidth relative to 
-visibility based approaches. We conclude and discuss future research prospects 
+visibility-based approaches. We conclude and discuss future research prospects 
 in \S\ref{sec:conclusions}.
@@
 \item The time-averaging cannot be performed on a stochastic measurement but
-  only on its statistical properties. In visibility based imaging, the 
+  only on its statistical properties. In visibility-based imaging, the 
   visibilities measured between antenna pairs represent spatial correlations 
@@
   must be imaged and squared before time-averaging. 
-\item In visibility based imaging, electric fields measured by antennas are not 
+\item In visibility-based imaging, electric fields measured by antennas are not 
   correlated with themselves and hence lack zero spacing measurements. In 
@@
 accessed at https://github.com/nithyanandan/EPIC} now publicly available. Besides 
-implementing the MOFF imaging algorithm it also includes visibility based imaging 
+implementing the MOFF imaging algorithm it also includes visibility-based imaging 
 using the software holography technique and a simulator for generating electric 
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{MOFF_flowchart}
+  \includegraphics[width=\columnwidth]{figure1}
+  % \includegraphics[width=\columnwidth]{MOFF_flowchart}
   % \includegraphics[width=\columnwidth]{MOFF_flowchart.eps}
@@
 
-Fig.~\ref{fig:FX-flowchart} shows the flowchart for a visibility based software 
+Fig.~\ref{fig:FX-flowchart} shows the flowchart for a visibility-based software 
 holographic imaging from a FX correlator. The antenna-based F-engine is identical 
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{FX_flowchart}
+  \includegraphics[width=\columnwidth]{figure2}
+  % \includegraphics[width=\columnwidth]{FX_flowchart}
   % \includegraphics[width=\columnwidth]{FX_flowchart.eps}
-  \caption{A flowchart of visibility based software holographic imaging in EPIC. 
+  \caption{A flowchart of visibility-based software holographic imaging in EPIC. 
     The FX process flow shares the F-engine with the MOFF process. Following the 
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{a2g_mapping}
+  \includegraphics[width=\columnwidth]{figure3}
+  % \includegraphics[width=\columnwidth]{a2g_mapping}
   % \includegraphics[width=\columnwidth]{a2g_mapping.eps}
@@
     The antenna aperture illumination kernels do not have to be identical to each
-    other. A discrete set of arbitrarily placed antennas are now placed onto a
+    other. A discrete set of arbitrarily placed antennas are now placed on to a
     regular grid.}
@@
 \par\medskip
-\noindent This module is common to the MOFF and visibility based imaging 
+\noindent This module is common to the MOFF and visibility-based imaging 
 techniques. Time samples of electric fields measured by the antenna and 
@@
 field spectra. This step can be parallelized across antennas as shown in 
-Fig.~\ref{fig:f-engine}. The output is then fed to either MOFF or visibility 
-based imaging pipelines.
+Fig.~\ref{fig:f-engine}. The output is then fed to either MOFF or 
+visibility-based imaging pipelines.
 
 \begin{figure}
-  \includegraphics[width=\columnwidth]{F-engine}
+  \includegraphics[width=\columnwidth]{figure4}
+  % \includegraphics[width=\columnwidth]{F-engine}
   % \includegraphics[width=\columnwidth]{F-engine.eps}
@@
 electric field signals with an image pixel from the output of the correlator in 
-the feedback calibration fashion outlined in \citealt{mor11} (illustrated in Fig. 
-\ref{fig:MOFF-flowchart} by the arrow leading from the imager to the calibration 
-block). Furthermore it allows for arbitrarily complex sky models, and following 
-the MOFF algorithm places no restriction on array layout, and accounts for 
-non-identical antenna beam patterns. Direction dependent calibration can be 
-achieved by correlating antenna signals with output pixels in the direction of 
-$N_\textrm{c}$ calibration sources, then fitting for a functional model of the 
-sky. Since antennas are only correlated with calibrator pixels, the 
-computational complexity scales as $\sim N_\textrm{a} N_\textrm{c}$. % And since 
-% only a single correlation is needed for each antenna, the computation complexity 
-% scales only as $\sim N_{\mathrm{a}}N_\textrm{c}$, where $N_\textrm{c}$ is the 
-% number of calibrators in the field.
+the feedback calibration fashion outlined in \citealt{mor11} (illustrated in 
+Fig.~\ref{fig:MOFF-flowchart} by the arrow leading from the imager to the 
+calibration block). Furthermore it allows for arbitrarily complex sky models, 
+and following the MOFF algorithm places no restriction on array layout, and 
+accounts for non-identical antenna beam patterns. Direction dependent calibration 
+can be achieved by correlating antenna signals with output pixels in the 
+direction of $N_\textrm{c}$ calibration sources, then fitting for a functional 
+model of the sky. Since antennas are only correlated with calibrator pixels, the 
+computational complexity scales as $\sim N_\textrm{a} N_\textrm{c}$.
 
@@
 In MOFF imaging, these are spatially Fourier transformed followed by a squaring
-operation at every time stamp for every frequency channel. In visibility based 
+operation at every time stamp for every frequency channel. In visibility-based 
 imaging, the spatial Fourier transform is performed only once per integration 
-timescale and does not include a squaring operation.
+time-scale and does not include a squaring operation.
 
@@
 this averaging has been performed already on the visibilities over an integration 
-timescale, the imaging step has to be performed only once per integration cycle.
+time-scale, the imaging step has to be performed only once per integration cycle.
 
@@
 produced through simulations. We simulate electric field streams from a model 
-sky and process the data through both the MOFF and a visibility based imaging 
+sky and process the data through both the MOFF and a visibility-based imaging 
 algorithm. We then compare the output images to demonstrate their equivalence.
@@
 
-We use the EPIC simulator to generate electric field samples from a sky model. In
-our simulations, we use 64 frequency channels each of width $\delta f = 40$~kHz, 
-10 point sources of flux densities 10~Jy at random locations. The number of
-time stamps integrated in one integration cycle was kept at eight where each A/D 
-timeseries is $1/\delta f=25\,\mu$s long. We use the MWA array layout 
-\citep{bea12} for demonstration. Only the inner 51 tiles within a square bounding
-box of 150~m on each side were used. We assumed all tiles are identical and have 
-a square shaped electric field illumination footprint 4.4~m on each side.
-
-\subsection{Dealing with antenna auto-correlations}\label{sec:rm-autocorr}
+We use the EPIC simulator to generate stochastic electric field samples from a 
+sky model consisting of 10 point sources of flux densities 10~Jy each at random 
+locations. In our simulations, we use 64 frequency channels each of width 
+$\Delta f = 40$~kHz. The number of time stamps integrated in one integration 
+cycle was kept at eight where each A/D timeseries is $1/\Delta f=25\,\mu$s long. 
+We use the MWA array layout \citep{bea12} for demonstration. Only the inner 51 
+tiles within a square bounding box of 150~m on each side were used. We assumed 
+all tiles are identical and have a square shaped electric field illumination 
+footprint 4.4~m on each side. Besides the stochastic sky noise present in the 
+simulated electric fields, no noise from the instrument is added.
+
+\subsection{Antenna auto-correlations}\label{sec:rm-autocorr}
 
@@
 MOFF imaging in the image plane introduces antenna auto-correlations around the 
-zero spacing in the $uv$-plane which are absent in traditional visibility based 
-imaging. In order to facilitate a robust comparison between MOFF and visibility 
-based imaging techniques, these auto-correlations are removed from the MOFF 
-algorithm output, which is otherwise not an essential part of the core algorithm. 
-We describe below how they are removed.
+zero spacing in the $uv$-plane which are absent in traditional visibility-based 
+imaging. In order to facilitate a robust comparison between MOFF and 
+visibility-based imaging techniques, these auto-correlations are removed from 
+the MOFF algorithm output, which is otherwise not an essential part of the core 
+algorithm. We describe below how they are removed.
 
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{autocorr_uvwts_pbeam}
+  \includegraphics[width=\columnwidth]{figure5}
+  % \includegraphics[width=\columnwidth]{autocorr_uvwts_pbeam}
   % \includegraphics[width=\columnwidth]{autocorr_uvwts_pbeam.png}
@@
 plane and subtract the estimated auto-correlation kernel scaled to the peak 
-value centered at the zero spacing pixel. The final averaged image is obtained 
+value centreed at the zero spacing pixel. The final averaged image is obtained 
 by Fourier transforming the $uv$ plane data and weights with the 
 auto-correlations subtracted to the image plane. These images are now comparable 
-to those obtained from visibility based imaging. This step of removing 
+to those obtained from visibility-based imaging. This step of removing 
 auto-correlations is optional and required to be performed only once per 
-integration timescale and does not add significant cost to the full operation.
+integration time-scale and does not add significant cost to the full operation.
 
@@
 Fig.~\ref{fig:MOFF-FX-image} shows the dirty images (top) and synthesized beams
-(bottom) obtained with antenna-based MOFF and FX visibility based imaging 
+(bottom) obtained with antenna-based MOFF and FX visibility-based imaging 
 algorithms packaged in EPIC. The antenna auto-correlations that correspond to 
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{MOFF_FX_comparison_10_random_source_positions_4_iterations_test_aperture_zoomed}
+  \includegraphics[width=\columnwidth]{figure6}
+  % \includegraphics[width=\columnwidth]{MOFF_FX_comparison_10_random_source_positions_4_iterations_test_aperture_zoomed}
   % \includegraphics[width=\columnwidth]{MOFF_FX_comparison_10_random_source_positions_4_iterations_test_aperture_zoomed.eps}
@@
 Fig.~\ref{fig:MOFF-FX-uvwts} shows the correlated weights obtained with MOFF 
-imaging (left) and visibility based imaging (right). There are two notable 
+imaging (left) and visibility-based imaging (right). There are two notable 
 differences. The first is in the weights around zero spacing. Though both show a
@@
 identical antenna but with a fractionally different location relative to the 
-grid. This depends on the exact location of the center of the antenna relative 
+grid. This depends on the exact location of the centre of the antenna relative 
 to the grid and the coarseness of grid spacing. This first order loss of 
@@
 correlation of two footprints each of which could be less precise to first 
-order. On the other hand, in visibility based imaging, a directly sampled $uv$ 
+order. On the other hand, in visibility-based imaging, a directly sampled $uv$ 
 plane antenna power response (which is theoretically identical to the correlation 
-of individual antenna footprints) centered on a baseline has a loss in precision 
+of individual antenna footprints) centreed on a baseline has a loss in precision 
 at most to first order. Thus, although in the limit of infinitesimally small 
@@
 mitigated by making the grid spacing finer at the expense of increased 
-computational cost. Residuals centered around zero spacing can also be lowered 
+computational cost. Residuals centreed around zero spacing can also be lowered 
 by subtracting each auto-correlation of antenna weights separately by using the 
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{MOFF_FX_comparison_uvwts_test_aperture_zero_spacing_removed}
+  \includegraphics[width=\columnwidth]{figure7}
+  % \includegraphics[width=\columnwidth]{MOFF_FX_comparison_uvwts_test_aperture_zero_spacing_removed}
   % \includegraphics[width=\columnwidth]{MOFF_FX_comparison_uvwts_test_aperture_zero_spacing_removed.eps}
   \caption{Comparison of $uv$ plane antenna power response with MOFF (left) and 
-    visibility based imaging (right). The zero-spacing footprint from the $uv$ 
+    visibility-based imaging (right). The zero-spacing footprint from the $uv$ 
     weights obtained with MOFF has been subtracted. Differences in the void 
-    centered around zero-spacing and other outer regions is due to loss of 
+    centreed around zero-spacing and other outer regions is due to loss of 
     sampled antenna footprint precision due to coarseness of grid spacing 
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{diff_psf_MOFF-FX_test_aperture}
+  \includegraphics[width=\columnwidth]{figure8}
+  % \includegraphics[width=\columnwidth]{diff_psf_MOFF-FX_test_aperture}
   % \includegraphics[width=\columnwidth]{diff_psf_MOFF-FX_test_aperture.eps}
@@
 in New Mexico. This data is in LWA narrow-band transient buffer (TBN) format 
-from 255 antennas within roughly a diameter of 100~m. The data is centered at a 
+from 255 antennas within roughly a diameter of 100~m. The data is centreed at a 
 frequency of 74.03~MHz, with a sample rate (equal to the bandwidth) of 100~kHz 
-with 512 complex time samples per antenna in a A/D writeout timescale of 
+with 512 complex time samples per antenna in a A/D writeout time-scale of 
 5.12~ms, a frequency resolution of 195.3125~Hz and dual polarization. There are 
@@
 packaged in EPIC after averaging over $\approx$~20~ms (four writeouts) of data 
-and the inner $\approx 80$\% of bandwidth (roughly 80~kHz). The image is shown 
-in direction cosine coordinates -- $l$ along the horizontal axis and $m$ along 
-the vertical axis. The flux scale is arbitrary. Even in this proof-of-concept 
-demonstration, we see Cyg~A and Cas~A prominently as annotated, thus validating 
-the functionality of EPIC.
+and the inner $\approx 80$ per cent of bandwidth (roughly 80~kHz). The image is 
+shown in direction cosine coordinates -- $l$ along the horizontal axis and $m$ 
+along the vertical axis. The flux scale is arbitrary. Even in this 
+proof-of-concept demonstration, we see Cyg~A and Cas~A prominently as annotated, 
+thus validating the functionality of EPIC.
 
 \begin{figure}
-  \includegraphics[width=\columnwidth]{LWA_MOFF_bandavg_image_4_iterations_analytic_aperture}
+  \includegraphics[width=\columnwidth]{figure9}
+  % \includegraphics[width=\columnwidth]{LWA_MOFF_bandavg_image_4_iterations_analytic_aperture}
   % \includegraphics[width=\columnwidth]{LWA_MOFF_bandavg_image_4_iterations_analytic_aperture.eps}
@@
 parameters such as antenna filling fraction, maximum baseline length, bandwidth 
-and frequency resolution, integration timescale, etc. for HERA antenna layouts 
+and frequency resolution, integration time-scale, etc. for HERA antenna layouts 
 which are highly compact. However, we note that in general, the hardware and 
@@
 stages even more expensive, the gridding convolution, squaring and time-averaging 
-take up only $\lesssim 20$\%, $\lesssim 20$\% and $\lesssim 5$\% respectively of 
-the total processing time while the spatial Fourier transform takes up 
-$\gtrsim 55\%$ of the total time. With sparser arrays the gridding process will 
-be be even faster.
-
-In visibility based imaging, the predominant computational cost is at the 
+take up only $\lesssim 20$, $\lesssim 20$ and $\lesssim 5$ per cent respectively 
+of the total processing time while the spatial Fourier transform takes up 
+$\gtrsim 55$ per cent of the total time. With sparser arrays the gridding process 
+will be be even faster.
+
+In visibility-based imaging, the predominant computational cost is at the 
 X-engine requiring $\Nant(\Nant-1)/2$ complex multiplications per channel at 
-every A/D writeout timescale.
+every A/D writeout time-scale.
 
@@
 the MOFF imaging is determined by the spatial Fourier transform while  
-that for visibility based imaging comes from the cross-correlations. However, 
+that for visibility-based imaging comes from the cross-correlations. However, 
 if non-linearities such as non-coplanarity of baselines \citep{cor08} and 
 wide-field phenomena like the {\it pitchfork} effect \citep{thy15a,thy15b} 
-are to be corrected for, the antenna based illumination footprint can start 
+are to be corrected for, the antenna-based illumination footprint can start 
 becoming less compact in the measurement plane and can result in a costlier 
@@
 than was indicated in \citet{mor11}. We set the number of complex 
-multiplications in the X-engine in visibility based imaging to 
+multiplications in the X-engine in visibility-based imaging to 
 $\Nant(\Nant-1)/2$.
@@
 complex operations per frequency channel per cross-polarization per integration 
-timescale. Telescopes that fall to the left of the solid line indicate MOFF 
-imaging is computationally more efficient than visibility based imaging. All HERA 
+time-scale. Telescopes that fall to the left of the solid line indicate MOFF 
+imaging is computationally more efficient than visibility-based imaging. All HERA 
 layouts except HERA-19 and HERA-37 are in a region of parameter space 
 where MOFF imaging holds the advantage. The dashed line showing future trajectory 
-of HERA like systems will be clearly favored by MOFF imaging. The dotted line is 
+of HERA like systems will be clearly favoured by MOFF imaging. The dotted line is 
 similarly a hypothetical trajectory for the MWA with more tiles added inside the 
 same core diameter. The gray shaded area is for a projected LWA expansion and is 
-also predominantly in the region favoring MOFF imaging. It is bounded by the LWA1 
+also predominantly in the region favouring MOFF imaging. It is bounded by the LWA1 
 and LWA-OV on the left and right respectively. The current (see 
 Table~\ref{tab:antenna-layouts}) and a hypothetical expanded layout with a 
-four-fold increase in number of elements over a 50\% increase in $b_\textrm{max}$ 
-provide the bounds at the bottom and top respectively. Current instruments such 
-as MWA and LOFAR lie in parameter space favoring visibility based imaging.
+four-fold increase in number of elements over a 50 per cent increase in 
+$b_\textrm{max}$ provide the bounds at the bottom and top respectively. Current 
+instruments such as MWA and LOFAR lie in parameter space favouring 
+visibility-based imaging.
 
 \begin{figure}
-  \includegraphics[width=\columnwidth]{MOFF_FX_computations_fov_gridding_annotated}
+  \includegraphics[width=\columnwidth]{figure10}
+  % \includegraphics[width=\columnwidth]{MOFF_FX_computations_fov_gridding_annotated}
   % \includegraphics[width=\columnwidth]{MOFF_FX_computations_fov_gridding_annotated.eps}
@@
     number of complex multiplies and adds with MOFF and FX. The solid line
-    is the boundary at which the number of operations with MOFF and visibility 
-    based imaging are equal. MOFF imaging is more efficient for telescopes 
-    occupying the left of this line and vice versa. CHIME, HIRAX, SKA1-LC, 
-    SKA1-LCD and all the HERA layouts except HERA-19 and HERA-37 lie in the 
-    parameter space favored by MOFF imaging. The dashed line shows the projected 
-    trajectory of hypothetical expanded HERA layouts. The dotted line similarly
-    shows hypothetical expanded MWA layouts with more tiles added in the same
-    core. The gray shaded area denotes the projected trajectory of the LWA 
-    bounded by LWA1 (left edge), LWA-OV (right edge), current layout (bottom) 
-    and a four-fold increase in the number of elements within a 50\% increase in 
-    the core size (top). Current instruments such as MWA and LOFAR fall in a 
-    region favored by visibility based imaging.}
+    is the boundary at which the number of operations with MOFF and 
+    visibility-based imaging are equal. MOFF imaging is more efficient for 
+    telescopes occupying the left of this line and vice versa. CHIME, HIRAX, 
+    SKA1-LC, SKA1-LCD and all the HERA layouts except HERA-19 and HERA-37 lie 
+    in the parameter space favoured by MOFF imaging. The dashed line shows the 
+    projected trajectory of hypothetical expanded HERA layouts. The dotted line 
+    similarly shows hypothetical expanded MWA layouts with more tiles added in 
+    the same core. The gray shaded area denotes the projected trajectory of the 
+    LWA bounded by LWA1 (left edge), LWA-OV (right edge), current layout 
+    (bottom) and a four-fold increase in the number of elements within a 50 per 
+    cent increase in the core size (top). Current instruments such as MWA and 
+    LOFAR fall in a region favoured by visibility-based imaging.}
   \label{fig:parameter-space-computations-instruments}
@@
 Fig.~\ref{fig:parameter-space-bll-nant-instruments} shows the boundaries where 
-the ratio of the number of computations required with visibility based imaging 
+the ratio of the number of computations required with visibility-based imaging 
 relative to MOFF imaging is unity. The different colored lines correspond to 
@@
 impossible. Dashed lines of each color denote the boundary to the left of which 
-visibility based imaging is favored for the corresponding antenna size. Region 
-inside the wedge enclosed by the solid and dashed lines favors using the MOFF 
-algorithm for the corresponding antenna size over visibility based imaging.
+visibility-based imaging is favoured for the corresponding antenna size. Region 
+inside the wedge enclosed by the solid and dashed lines favours using the MOFF 
+algorithm for the corresponding antenna size over visibility-based imaging.
 
@@
 they fall into. Thus symbols of one color falling into a wedge of the same color 
-indicates MOFF imaging is favored for those telescopes and vice versa. For 
-e.g., MOFF imaging is favored in HERA-331 and HERA-6769 because they are inside 
+indicates MOFF imaging is favoured for those telescopes and vice versa. For 
+e.g., MOFF imaging is favoured in HERA-331 and HERA-6769 because they are inside 
 the wedge but not so in cases of HERA-19 and HERA-37. A majority of the 
@@
 100~m and 200~m respectively for LWA1 and LWA-OV.} will be in the computational 
-regime favored by MOFF. For a fixed baseline length, regions favoring the MOFF 
+regime favoured by MOFF. For a fixed baseline length, regions favouring the MOFF 
 algorithm tend to be towards large $\Nant$ indicating large-N dense array 
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{MOFF_FX_crossover_baseline_n-antennas_rho_fov_gridding_legended}
+  \includegraphics[width=\columnwidth]{figure11}
+  % \includegraphics[width=\columnwidth]{MOFF_FX_crossover_baseline_n-antennas_rho_fov_gridding_legended}
   % \includegraphics[width=\columnwidth]{MOFF_FX_crossover_baseline_n-antennas_rho_fov_gridding_legended.eps}
@@
     solid lines for corresponding antenna size is physically disallowed. Dashed
-    lines of each color denote the boundary to the left of which visibility
-    based imaging is favored for the corresponding antenna size. Region inside 
-    the wedge enclosed by the solid and dashed lines favors using the MOFF 
-    algorithm for the corresponding antenna size. These wedges shift leftward 
-    with increasing antenna size. The different antennas are color coded by 
-    roughly the class of antenna size they fall into. Thus symbols of one 
-    color falling into a wedge of the same color indicates MOFF imaging is 
+    lines of each color denote the boundary to the left of which 
+    visibility-based imaging is favoured for the corresponding antenna size. 
+    Region inside the wedge enclosed by the solid and dashed lines favours using 
+    the MOFF algorithm for the corresponding antenna size. These wedges shift 
+    leftward with increasing antenna size. The different antennas are color 
+    coded by roughly the class of antenna size they fall into. Thus symbols of 
+    one color falling into a wedge of the same color indicates MOFF imaging is 
     advantageous for those telescopes and vice versa. For example, MOFF imaging
-    is favored in HERA-331 and HERA-6769 because they are inside the wedge 
+    is favoured in HERA-331 and HERA-6769 because they are inside the wedge 
     but not so in cases of HERA-19 and HERA-37.}
@@
 
-We elaborate on the I/O data rates required with the MOFF and visibility based
+We elaborate on the I/O data rates required with the MOFF and visibility-based
 algorithms. This is particularly relevant in the context of radio transient 
@@
 Implementation of the MOFF algorithm with EPIC yields calibrated images on
-timescales of the output generated by the digitizer and is set by the inverse
+time-scales of the output generated by the digitizer and is set by the inverse
 of the frequency channel width. These calibrated images are accumulated and 
-averaged to a certain timescale depending on science or hardware requirements,
-or when the sky has rotated significantly, whichever is lesser. In visibility
-based algorithms, the visibilities are accumulated and averaged to this 
-timescale before images are produced. Thus the data throughput (in samples per 
-second) per cross-polarization with MOFF and X-engine outputs are: 
+averaged to a certain time-scale depending on science or hardware requirements,
+or when the sky has rotated significantly, whichever is lesser. In 
+visibility-based algorithms, the visibilities are accumulated and averaged to 
+this time-scale before images are produced. Thus the data throughput (in samples 
+per second) per cross-polarization with MOFF and X-engine outputs are: 
 \begin{align}
@@
 complex visibilities, $\Delta B$ is the bandwidth, $\Delta f$ is the frequency 
-resolution, and $\Delta t$ is the timescale over which the transient phenomenon
+resolution, and $\Delta t$ is the time-scale over which the transient phenomenon
 is sampled and the data (images or visibilities) are averaged to.
@@
 Though a full understanding of the FRB phenomena is yet to emerge, there are 
-indications the timescales of FRB objects are $\Delta t \sim 1$--10~ms 
+indications the time-scales of FRB objects are $\Delta t \sim 1$--10~ms 
 \citep{tho13}. For a telescope like HERA, $\Delta B \simeq 100$~MHz, 
@@
 4 bytes each for real and imaginary parts of visibility samples, the throughputs 
-are $r_\textrm{MOFF} \lesssim 3$~GB/s and $r_\textrm{X} \simeq 41$~GB/s. For 
-HERA-37, $r_\textrm{MOFF} \lesssim 190$~MB/s and $r_\textrm{X} \simeq 0.5$~GB/s. 
-In such a case, The X-engine throughput corresponds to an extreme rate of 
-$\simeq 1.8$~TB an hour per cross-polarization. Conversely, for the same data 
-throughput, the MOFF algorithm can sample even shorter timescales.
+are $r_\textrm{MOFF} \lesssim 3$~GB~s$^{-1}$ and 
+$r_\textrm{X} \simeq 41$~GB~s$^{-1}$. For HERA-37, 
+$r_\textrm{MOFF} \lesssim 190$~MB~s$^{-1}$ and 
+$r_\textrm{X} \simeq 0.5$~GB~s$^{-1}$. In such a case, The X-engine throughput 
+corresponds to an extreme rate of $\simeq 1.8$~TB an hour per 
+cross-polarization. Conversely, for the same data throughput, the MOFF algorithm 
+can sample even shorter time-scales.
 
@@
   \caption{Data throughput per cross-polarization for various telescopes with 
-    MOFF and X-engine outputs on timescales of $\Delta t=10$~ms assuming 
+    MOFF and X-engine outputs on time-scales of $\Delta t=10$~ms assuming 
     $\Delta B=100$~MHz and $\Delta f=100$~kHz.}
@@
     Telescope\tnote{a} & $r_\textrm{MOFF}$\tnote{b} & $r_\textrm{X}$ \\
-              & (GB/s)\tnote{c} & (GB/s)\tnote{c} \\
+              & (GB~s$^{-1}$)\tnote{c} & (GB~s$^{-1}$)\tnote{c} \\
     \hline
@@
 LWA are headed towards the large-N dense array layout model for which 
-computational cost from traditional FX/XF correlator based architecture and 
-visibility based imaging starts rising steeply. We have provided the first 
+computational cost from traditional FX/XF correlator-based architecture and 
+visibility-based imaging starts rising steeply. We have provided the first 
 software demonstration of a general purpose imaging algorithm using our generic 
@@
 highly modularized and parallelizable. It includes an implementation of the 
-MOFF algorithm in addition to visibility based software holography imaging and 
+MOFF algorithm in addition to visibility-based software holography imaging and 
 a data simulator for sky models. It is designed to provide a development 
 platform to compare different imaging approaches and serve as a stepping stone 
-for real-life GPU/FPGA based implementation on telescopes. It has been 
+for real-life GPU/FPGA-based implementation on telescopes. It has been 
 successfully tested on simulated MWA observations as well as real LWA 
@@
 further the already relatively low I/O bandwidth shown in 
-table~\ref{tab:data-rates} except when a transient event is detected. Transient 
+Table~\ref{tab:data-rates} except when a transient event is detected. Transient 
 detection pipelines at the backend of EPIC can be fine-tuned to target fast 
 transients such as the Fast Radio Bursts \citep[FRB;][]{tho13} on millisecond 
-timescales at GHz frequencies or slow transients from planetary and exoplanetary 
+time-scales at GHz frequencies or slow transients from planetary and exoplanetary 
 origins at frequencies around 100~MHz.
@@
 In the near future, we plan to upgrade our current Python implementation 
-of EPIC to a GPU based pipeline in order to operate on real-time data and 
+of EPIC to a GPU-based pipeline in order to operate on real-time data and 
 develop a transient trigger and monitor backend. In the meanwhile, we plan 
@@
 
-\section*{Acknowledgments}
+\section*{Acknowledgements}
 
 We thank Larry D'Addario, Gregg Hallinan, Joseph Lazio and Harish Vedantham for 
-their valuable inputs, and Greg Taylor for providing us with LWA data. 
+their valuable inputs, and Greg Taylor for providing us with LWA data. This work 
+has been supported by the National Science Foundation through award AST-1206552. 
 Construction of the LWA has been supported by the Office of Naval Research under 
@@
 \bibliographystyle{mnras}
-\bibliography{../epic} 
+\bibliography{epic} 
+% \bibliography{../epic} 
 % \input{ms.bbl}
@@
 \begin{figure*}
-  \includegraphics[width=\linewidth]{EPIC-modules}
+  \includegraphics[width=\linewidth]{figureA1}
+  % \includegraphics[width=\linewidth]{EPIC-modules}
   % \includegraphics[width=\linewidth]{EPIC-modules.eps}
@@
     of the former) to produce output images. Broadly, the MOFF algorithm is 
-    implemented by modules below the horizontal dashed line while the visibility 
-    based imaging uses modules above the line. The exact processing pathway 
-    implementing the MOFF algorithm is shown in bolded modules.}
+    implemented by modules below the horizontal dashed line while the 
+    visibility-based imaging uses modules above the line. The exact processing 
+    pathway implementing the MOFF algorithm is shown in bolded modules.}
   \label{fig:software-modules}
@@
 The other function (not shown in the figure) is to update the data as new data 
-streams in. This can also be parallelized. Another important attribute
-(not shown in figure) are antenna flags for each polarization appropriate for 
-the data stream being held by the module.
+streams in. This can also be parallelized. Another important attribute consists 
+of antenna flags (not shown in the figure) for each polarization appropriate 
+for the data stream being held by the module.
 
@@
 
-The other attributes (not shown in figure) are the flags applicable for each
+The other attributes (not shown in the figure) are the flags applicable for each
 cross-polarization for the current data stream. Similar to the antenna module,
@@
 This module forms the fundamental unit for the interferometer array module 
-(to be discussed below) and in general for visibility based correlator and 
+(to be discussed below) and in general for visibility-based correlator and 
 imaging systems.
@@
 stack, accumulate and average images, and optionally remove the antenna
-auto-correlations centered around the zero-spacing pixel in the $uv$ plane. 
+auto-correlations centreed around the zero-spacing pixel in the $uv$ plane. 
 It also handles all four cross-polarization products. Currently, it supports 

### 2015-11-08 imaging_demo/ms.tex
Avoid Type 3 fonts in all figures for MNRAS specs
From dca9220 (mnras):
====================

(*) Remake all figures to not use Type 3 fonts to be
MNRAS-compatible

In ms.tex,

(*) Use updated figures
(*) Change use of commented F-engine figure to parallel_F_engine

@@
 \usepackage{ae,aecompl}
+% \pdfminorversion=5
 
@@
 $I(f)$ ($\mathbf{I}(\hat{\mathbf{s}})$ in matrix notation).
+
 \begin{figure}
@@
 theoretically identical as explained in \S\ref{sec:math}.
+
 \begin{figure}
@@
   \includegraphics[width=\columnwidth]{figure4}
-  % \includegraphics[width=\columnwidth]{F-engine}
-  % \includegraphics[width=\columnwidth]{F-engine.eps}
+  % \includegraphics[width=\columnwidth]{parallel_F_engine}
+  % \includegraphics[width=\columnwidth]{parallel_F_engine.eps}
   \caption{Block diagram of a F-engine. The electric field data streams from
@@
 line. Their hypothetical expansions,\footnote{LWA1-x2x1 and LWA1-x4x1.5, and 
-LWA-OVx2x1 and LWA-OVx4x1.5 denote two-fold and four-fold increase in the number 
+LWA-OVx2x1 and LWA-OVx4x1.5 denote two-fold and four-fold increase in number 
 of antennas within a core diameter that is 1 and 1.5 times the current size of 

### 2015-11-10 cal_demo/ms.tex
started the math section of cal paper
cleaned up cal tex file

@@
-% mnras_template.tex
-%
-% LaTeX template for creating an MNRAS paper
-%
-% v3.0 released 14 May 2015
-% (version numbers match those of mnras.cls)
-%
-% Copyright (C) Royal Astronomical Society 2015
-% Authors:
-% Keith T. Smith (Royal Astronomical Society)
-
-% Change log
-%
-% v3.0 May 2015
-%    Renamed to match the new package name
-%    Version number matches mnras.cls
-%    A few minor tweaks to wording
-% v1.0 September 2013
-%    Beta testing only - never publicly released
-%    First version: a simple (ish) template for creating an MNRAS paper
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % Basic setup. Most papers should leave these options alone.
@@
 
-% MNRAS is set in Times font. If you don't have this installed (most LaTeX
-% installations will be fine) or prefer the old Computer Modern fonts, comment
-% out the following line
 \usepackage{newtxtext,newtxmath}
-% Depending on your LaTeX fonts installation, you might get better results with one of these:
-%\usepackage{mathptmx}
-%\usepackage{txfonts}
 
@@
 
-%%%%% AUTHORS - PLACE YOUR OWN PACKAGES HERE %%%%%
+%%%%% PACKAGES %%%%%
 
@@
 \usepackage{amssymb}	% Extra maths symbols
+\usepackage{color}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%% CUSTOM COMMANDS %%%%%
 
-%%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%
-
-% Please keep new commands to a minimum, and use \newcommand not \def to avoid
-% overwriting existing commands. Example:
-%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newcommand{\s}{\hat{\mathbf{s}}} % s-hat for sine-projected direction
+\newcommand{\spix}{\hat{\mathbf{s}}_{\mathrm{pix}}}
+\newcommand{\Cna}{\mathcal{C}^{(n)}_a}
+\newcommand{\ri}{\mathbf{r}_i}
+\newcommand{\ra}{\mathbf{r}_a}
+\newcommand{\rb}{\mathbf{r}_b}
+\newcommand{\Nant}{N_{\mathrm{ant}}}
+\newcommand{\beamr}{\widetilde{W}}
+\newcommand{\beamtheta}{W}
+\newcommand{\Er}[1]{\widetilde{E}_{#1}}
+\newcommand{\Erest}[1]{\hat{\widetilde{E}}_{#1}}
+\newcommand{\V}{\widetilde{V}}
 
@@
 
-% Select between one and six entries from the list of approved keywords.
-% Don't make up new ones.
 \begin{keywords}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 %%%%%%%%%%%%%%%%% BODY OF PAPER %%%%%%%%%%%%%%%%%%
@@
 \section{Introduction}
+The field of radio astronomy is undergoing a paradigm shift toward interferometers consisting of hundreds to thousands of small, widefield antennas.
+This shift is necessary to match the survey speeds required for both precision cosmology experiments, and searches for fast radio transients.
+Many arrays with this design are already built or under construction including \cite{XXX}.
 
@@
 Introduce notation, describe measurement equation.
+
+\begin{equation}
+\mathcal{C}^n(a,\spix,f) \equiv \left<E_a(f) E'^*(\spix,f)\right>_t
+\end{equation}
+$\mathcal{C}^n(a,\spix,f) \rightarrow \Cna$
+
+\begin{equation}
+\Er{a} = g_a E^T_a
+\end{equation}
+\begin{align}
+\hat{E}(\spix) & = \frac{1}{\Nant} \sum_i e^{2\pi i \spix \cdot \ri} \Erest{i} \\
+& = \frac{1}{\Nant} \sum_i e^{2\pi i \spix \cdot \ri} \sum_b \beamr_b(\ri-\rb) \Erest{b} \\
+& = \frac{1}{\Nant} \sum_i e^{2\pi i \spix \cdot \ri} \sum_b \beamr_b(\ri-\rb) h^{(n)}_b g_b\Er{b}^T
+\end{align}
+We can then play a trick to transform the beam.
+\begin{align}
+\hat{E}(\spix) & = \frac{1}{\Nant} \sum_b h^{(n)}_b g_b\Er{b}^T e^{2\pi i \spix \cdot \rb}\sum_i \beamr_b(\ri-\rb)e^{2\pi i \spix \cdot (\ri-\rb)}\\
+& = \frac{1}{\Nant} \sum_b h^{(n)}_b g_b\Er{b}^T e^{2\pi i \spix \cdot \rb}\beamtheta_b(\spix)
+\end{align}
+
+Plugging in, we get,
+\begin{align}
+\Cna & = \left<\Er{a} \hat{E}^*(\spix)\right>_t \\ 
+& = \left<g_a \Er{a}^T \frac{1}{\Nant} \sum_b h^{*(n)}_b g^*_b\Er{b}^{*T} e^{-2\pi i \spix \cdot \rb} \beamtheta^*_b(\spix)\right>_t \\
+\end{align}
+Pulling time independent pieces out.
+\begin{align}
+\Cna & = \frac{g_a}{\Nant} \sum_b h^{*(n)}_b g^*_b \beamtheta^*_b(\spix) e^{-2\pi i \spix \cdot \rb} \left<\Er{a}^T \Er{b}^{*T} \right>_t \\
+& = \frac{g_a}{\Nant} \sum_b h^{*(n)}_b g^*_b \beamtheta^*_b(\spix) e^{-2\pi i \spix \cdot \rb} \V^T_{ab}
+\end{align}
+Solve for $g^{(n+1)}_a$.
+\begin{equation}
+g^{(n+1)}_a = \Cna \Nant \left[ \sum_b h^{*(n)}_b g^{*(n)}_b \beamtheta^*_b(\spix) e^{-2\pi i \spix \cdot \rb} \V^T_{ab} \right]^{-1}
+\end{equation}
+
+In the case where $h_b = 1/g_b$, this simplifies slightly.
+\begin{equation}
+g^{(n+1)}_a = \Cna \Nant \left[ \sum_b \beamtheta^*_b(\spix) e^{-2\pi i \spix \cdot \rb} \V^T_{ab} \right]^{-1}
+\end{equation}
 
@@
 
-% \subsection{Figures and tables}
-
-% Figures and tables should be placed at logical positions in the text. Don't
-% worry about the exact layout, which will be handled by the publishers.
-
-% Figures are referred to as e.g. Fig.~\ref{fig:example_figure}, and tables as
-% e.g. Table~\ref{tab:example_table}.
-
-% % Example figure
-% \begin{figure}
-% 	% To include a figure from a file named example.*
-% 	% Allowable file formats are eps or ps if compiling using latex
-% 	% or pdf, png, jpg if compiling using pdflatex
-% 	\includegraphics[width=\columnwidth]{example}
-%     \caption{This is an example figure. Captions appear below each figure.
-% 	Give enough detail for the reader to understand what they're looking at,
-% 	but leave detailed discussion to the main body of the text.}
-%     \label{fig:example_figure}
-% \end{figure}
-
-% % Example table
-% \begin{table}
-% 	\centering
-% 	\caption{This is an example table. Captions appear above each table.
-% 	Remember to define the quantities, symbols and units used.}
-% 	\label{tab:example_table}
-% 	\begin{tabular}{lccr} % four columns, alignment for each
-% 		\hline
-% 		A & B & C & D\\
-% 		\hline
-% 		1 & 2 & 3 & 4\\
-% 		2 & 4 & 6 & 8\\
-% 		3 & 5 & 7 & 9\\
-% 		\hline
-% 	\end{tabular}
-% \end{table}
-
 \section{Conclusions}
 \section*{Acknowledgements}
-
-The Acknowledgements section is not numbered. Here you can thank helpful
-colleagues, acknowledge funding agencies, telescopes and facilities used etc.
-Try to keep it short.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+This work has been supported by the National Science Foundation through award AST-1206552.
 
@@
 \bibliography{../epic} % if your bibtex file is called example.bib
-
-% Alternatively you could enter them by hand, like this:
-% This method is tedious and prone to error if you have lots of references
-% \begin{thebibliography}{99}
-% \bibitem[\protect\citeauthoryear{Author}{2012}]{Author2012}
-% Author A.~N., 2013, Journal of Improbable Astronomy, 1, 1
-% \bibitem[\protect\citeauthoryear{Others}{2013}]{Others2013}
-% Others S., 2012, Journal of Interesting Stuff, 17, 198
-% \end{thebibliography}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2015-11-12 cal_demo/ms.tex
wrote intro for cal paper, added thy15 to bib file

@@
 
-\newcommand{\s}{\hat{\mathbf{s}}} % s-hat for sine-projected direction
-\newcommand{\spix}{\hat{\mathbf{s}}_{\mathrm{pix}}}
-\newcommand{\Cna}{\mathcal{C}^{(n)}_a}
-\newcommand{\ri}{\mathbf{r}_i}
-\newcommand{\ra}{\mathbf{r}_a}
-\newcommand{\rb}{\mathbf{r}_b}
-\newcommand{\Nant}{N_{\mathrm{ant}}}
-\newcommand{\beamr}{\widetilde{W}}
-\newcommand{\beamtheta}{W}
-\newcommand{\Er}[1]{\widetilde{E}_{#1}}
-\newcommand{\Erest}[1]{\hat{\widetilde{E}}_{#1}}
-\newcommand{\V}{\widetilde{V}}
+\newcommand{\Nant}{\ensuremath{N_{\mathrm{ant}}}}
+\newcommand{\Ng}{\ensuremath{N_{\mathrm{g}}}}
+\newcommand{\s}{\ensuremath{\hat{\mathbf{s}}}} % s-hat for sine-projected direction
+\newcommand{\spix}{\ensuremath{\hat{\mathbf{s}}_{\mathrm{pix}}}}
+\newcommand{\Cna}{\ensuremath{\mathcal{C}^{(n)}_a}}
+\newcommand{\ri}{\ensuremath{\mathbf{r}_i}}
+\newcommand{\ra}{\ensuremath{\mathbf{r}_a}}
+\newcommand{\rb}{\ensuremath{\mathbf{r}_b}}
+\newcommand{\beamr}{\ensuremath{\widetilde{W}}}
+\newcommand{\beamtheta}{\ensuremath{W}}
+\newcommand{\Er}[1]{\ensuremath{\widetilde{E}_{#1}}}
+\newcommand{\Erest}[1]{\ensuremath{\hat{\widetilde{E}}_{#1}}}
+\newcommand{\V}{\ensuremath{\widetilde{V}}}
 
@@
 \section{Introduction}
-The field of radio astronomy is undergoing a paradigm shift toward interferometers consisting of hundreds to thousands of small, widefield antennas.
-This shift is necessary to match the survey speeds required for both precision cosmology experiments, and searches for fast radio transients.
-Many arrays with this design are already built or under construction including \cite{XXX}.
+In order to satisfy the survey speeds required for precision cosmology as well as searches for fast radio transients, radio astronomy is undergoing a paradigm shift toward interferometers consisting of hundreds to thousands of small, widefield antennas.
+Many arrays with this design are already built or under construction including the Hydrogen Epoch of Reionization Array\footnote{http://reionization.org} (HERA), the Murchison Widefield Array (MWA;\citealt{tin13,bow13}), the Precision Array for Probing the Epoch of Reionization (PAPER; \citealt{par10}), the LOw Frequency ARray (LOFAR;\citealt{van13}), the Canadian Hydrogen Intensity Mapping Experiment (CHIME,\citealt{ban14}), the Long Wavelength Array (LWA, \citealt{ell13}), and the low frequency Square Kilometer Array (SKA1-Low \citealt{mel13}).
 
-Motivate problem of calibrating direct imagers. Reference Nithyas paper. \citep{mor11}
+Traditional radio correlators cross-multiply the voltage signals from all pairs of antennas, and the computation scales as the number of antennas squared, $\mathcal{O}(\Nant^2)$ \citep{bun04}.
+As the number of elements in future arrays grows, the computational cost will become prohibitively expensive, and exploring efficient correlator schemes is essential to enable next generation instruments \citep{lon00}.
+Meanwhile, radio transient monitoring requires access to high time and frequency resolution data.
+For example, Fast Radio Bursts (FRBs) are highly unexplored at low frequencies (< 1 GHz), but are expected to occur on timescales $\Delta t \sim$ 1--10~ms \citep{tho13}.
+Recording the full visibility matrix for $\Nant \gtrsim 10^3$ arrays at this timescale leads to extremely high data write rates.
 
-\section{Mathematical Framework}
+Direct imaging correlators are a new variety of radio correlator which aim to alleviate both the computational strain of forming $\Nant^2$ correlations and the high data throughput associated with short timescale science.
+This is done by performing a spatial fast Fourier transform (FFT) to image the antenna voltages, then squaring and averaging in time.
+This process scales as $\mathcal{O}(\Ng \log_2 \Ng)$, where \Ng~is the number of grid points in the FFT \citep{mor11, teg09, teg10}.
+For certain classes of telescopes, significantly those envisioned for next generation cosmology experiments, this scaling is a large improvement over the $\Nant^2$ scaling of traditional methods.
+Furthermore, because images are generated online, the native output bandwidth will be lowered (assuming $\Ng < \Nant^2$), and has the potential to be lowered even further with online transient processing.
+
+A handful of prototype direct imaging correlators have been tested on arrays including the Basic Element for SKA Training II (BEST-2) array \citep{fos14}, the Omniscope \citep{zhe14}, and an earlier pulsar timing experiment at GHz frequencies \citep{oto94, dai00}.
+Each of these are examples of so-called FFT correlators -- a subclass of direct imaging correlators which rely on identical antennas with restricted placement, which allows the FFT to be performed without gridding.
+We recently released the E-field Parallel Imaging Correlator \citep[EPIC;][]{thy15c}, which is a software implementation of the Modular Optimal Frequency Fourier \citep[MOFF;][]{mor11} imaging algorithm.
+This architecture leverages the software holography/A-transpose framework to grid electric field data streams before performing the spatial FFT, allowing for an optimal map without placing constraints on array layout or requiring identical antennas \citep{mor09,bha08,teg97a}.
+
+A challenge common to all direct imaging algorithms is calibration of the antenna gains.
+Traditionally, pair-wise visibilities are written to disk and used to calibrate offline.
+However, a direct imaging correlator mixes the signals from all antennas before averaging and writing to disk, making calibration a requirement at the front end.
+Previous solutions have involved applying calibration solutions generated from a parallel FX correlator \citep{zhe14, fos14}, or integrating a dedicated FX correlator which periodically formed the full visibility matrix to solve for gains \citep{wij09,dev09}.
+While these solutions were sufficient to enable the exploration of FFT correlators and beamformers, they will not scale to future arrays with $\Nant \gtrsim 10^3$.
+
+Here we present the E-field Parallel Imaging Calibration (EPICal) algorithm -- a novel solution to the calibration problem, which can be integrated into direct imaging correlators and scales only as the number of antennas, $\mathcal{O}(\Nant)$.
+This method uses a correlation of the uncalibrated antenna signal stream with an output image pixel from the backend of the correlator to solve for the complex gains of the antennas.
+In the limit of a simple sky, the algorithm reduces to the self-cal algorithm, and even in a complex limit can be shown to be equivalent to visibility-based solutions.
+An example implementation of the algorithm is available with the EPIC software package\footnote{http://github.com/nithyanandan/EPIC}
+We establish the mathematical framework and derive the calibration algorithm in \S \ref{sec:math}.
+We then demonstrate the algorithm in simulations in \S \ref{sec:sim}, and apply to a sample LWA data set in \S \ref{sec:data}.
+Then we discuss the noise properties of the resulting gain solutions in \S \ref{sec:noise}.
+Finally we conclude and discuss potential extensions to the algorithm in \S \ref{sec:discussion}.
+
+\section{Mathematical Framework}\label{sec:math}
+We begin by establishing the mathematical framework for the calibration problem.
+We derive the calibration solutions for the MOFF algorithm (adopting the notation of \citealt{thy15c}), but note the result is easily extended to FFT correlator algorithms by removing the gridding step.
 
@@
 
-\section{Simulation}
+\section{Simulation}\label{sec:sim}
 
@@
 
-\section{Verification}
+\section{Application to LWA data}\label{sec:data}
+Apply to LWA data.
+
+\section{Noise analysis}\label{sec:noise}
 
@@
 
-\section{Conclusions}
+Assume we did form visibilities with some integration time.
+Assuming the noise on each visibility, $\sigma_{ab}$, is independent, we can write the likelihood function of measuring $V_{ab}$ given the true value, the gains, and the noise.
+\begin{equation}
+\mathcal{L}(V_{ab};\mathbf{g}) = \frac{1}{2\pi \sigma_{ab}^2}\exp\left[-\frac{\left|V_{ab} - g_a g_b^* V_{ab}^T\right|^2}{2\sigma_{ab}^2}\right]
+\end{equation}
+Then the likelihood of the set of all visibilities will be,
+\begin{equation}
+\mathcal{L}(\mathbf{V};\mathbf{g}) = \prod_a \prod_{b > a} \mathcal{L}(V_{ab};\mathbf{g})
+\end{equation}
+The Fischer information matrix for the set of gain parameters is
+\begin{equation}
+\mathbf{F}^g_{ij} = \left<\left. \frac{\partial \ln \mathcal{L}(\mathbf{V};\mathbf{g})}{\partial g_i} \frac{\partial \ln \mathcal{L}(\mathbf{V};\mathbf{g})}{\partial g_j}\right|_\mathbf{g}\right>.
+\end{equation}
+We next evaluate the derivate of the log-likelihood.
+\begin{equation}
+\frac{\partial \ln \mathcal{L}(\mathbf{V};\mathbf{g})}{\partial g_i} = 
+\sum_{a \ne i} \frac{g_a^* V_{ai}^{T*} (V_{ai} - g_a g_i^* V_{ai}^T)}{2 \sigma_{ai}^2}
+\end{equation}
+
+\textcolor{red}{This part of the argument needs serious work!} In order to find the Cram\'er-Rao lower bound on the variance of the complex parameter, $g_i$, we consider the term of the Fischer matrix where the first derivative is taken with respect to $g_i$, and the second with $g_i^*$.
+The result is
+\begin{align}
+\mathbf{F}^g_{ii} = & \left<\left[ \sum_{a \ne i} \frac{g_a^* V_{ai}^{T*} (V_{ai} - g_a g_i^* V_{ai}^T)}{2 \sigma_{ai}^2} \right] \right. \times \nonumber \\
+& \left. \left[ \sum_{b \ne i} \frac{g_b V_{bi}^{T} (V_{bi}^* - g_b^* g_i V_{bi}^{T*})}{2 \sigma_{bi}^2} \right] \right> \nonumber \\
+= &\sum_{a\ne i} \sum_{b \ne i} \frac{g_a^* g_b V_{ai}^{T*} V_{bi}^T}{4 \sigma_{ai}^2 \sigma_{bi}^2} \left< V_{ai}V_{bi}^* - g_i g_b^* V_{ai}V_{bi}^{T*}\nonumber \right.\\
+&\left. - g_a g_i^* V_{ai}^T V_{bi}^* + g_a g_b^* |g_i|^2 V_{ai}^T V_{bi}^{T*}\right>
+\label{eq:Fischer_before_expect}
+\end{align}
+
+The expected values are easy to evaluate. Each visibility will average to the ``true" value times the respective gains. The term with two visibilities will include a noise term.
+\begin{equation}
+\left<V_{ai}V_{bi}^*\right> = |g_i|^2 g_a g_b^* V_{ai}^T V_{bi}^{T*} + \sigma_{ai}^2 \delta_{ab}
+\end{equation}
+Here $\delta_{ab}$ is the Kronecker delta selecting the term where $a=b$ and the noise correlates. Plugging in the expectation values, equation \ref{eq:Fischer_before_expect} simplifies greatly to
+\begin{equation}
+\mathbf{F}^g_{ii} = \sum_{a \ne i} \frac{\left|g_a V_{ia}^T \right|^2}{4\sigma_{ai}^2}
+\end{equation}
+
+Finally we relate our result to the theoretical best uncertainty we can place on our unknown gain parameter using the Cram\'er-Rao lower bound.
+\begin{equation}\label{eq:cramer_rao}
+\sigma_{g_i}^2 \ge \left[ \mathbf{F}^g_{ii} \right]^{-1} = \left[ \sum_{a \ne i} \frac{\left|g_a V_{ia}^T \right|^2}{4\sigma_{ai}^2} \right]^{-1}
+\end{equation}
+
+\section{Discussion}\label{sec:discussion}
+
 \section*{Acknowledgements}

### 2015-11-17 cal_demo/ms.tex
more text to cal paper, added schematic figure

@@
 \newcommand{\spix}{\ensuremath{\hat{\mathbf{s}}_{\mathrm{pix}}}}
-\newcommand{\Cna}{\ensuremath{\mathcal{C}^{(n)}_a}}
+\newcommand{\Cna}{\ensuremath{\mathcal{C}^{(n)}_{a,\spix}}}
 \newcommand{\ri}{\ensuremath{\mathbf{r}_i}}
@@
 \newcommand{\V}{\ensuremath{\widetilde{V}}}
+\newcommand{\dif}{\mathrm{d}}
 
@@
 
-Introduce notation, describe measurement equation.
+The electric field incident on the ground, $\Er{}(\mathbf{r},f,t)$, is related to the sky electric field, $E(\hat{\mathbf{s}},f,t)$, through a Fourier transform.
+\begin{equation}
+\Er{}(\mathbf{r},f,t) = \int E(\hat{\mathbf{s}},f,t) e^{-2\pi i \mathbf{r}\cdot \hat{\mathbf{s}}}\, \dif^2 \hat{\mathbf{s}}
+\end{equation}
+Here $\hat{\mathbf{s}}$ denotes the sine-projected unit vector for the sky angle, $\mathbf{r}$ is the observer's location (measured in wavelengths relative to an arbitrary origin), and $f$ and $t$ denote the frequency and time dependence, respectively.
+We will encounter several quantities which we attempt to estimate.
+We distinguish the ``true" values with a superscript $T$, while the estimates are denoted with a hat.
+We define the true antenna signal as a convolution of the antenna voltage pattern, $\beamr$, with the electric field on the ground.
+\begin{equation}
+\Er{a}^T(f,t) \equiv \int \beamr_a(\mathbf{r}-\ra) \Er{}(\mathbf{r},f,t) \, \dif^2 \mathbf{r}
+\end{equation}
+The subscript $a$ labels the antenna, and $\mathbf{r}_a$ is the location of antenna $a$.
 
+We next model the measured, uncalibrated electric field as a multiplicative complex gain and an additive noise term applied to the true antenna electric field. 
 \begin{equation}
-\mathcal{C}^n(a,\spix,f) \equiv \left<E_a(f) E'^*(\spix,f)\right>_t
+\Er{a}(f,t) = g^T_a(f,t) \Er{a}^T(f,t) + \widetilde{n}_a(f,t)
 \end{equation}
-$\mathcal{C}^n(a,\spix,f) \rightarrow \Cna$
+Note that this quantity is neither a true or estimated value.
+The noise term is strictly receiver noise -- noise introduced by the instrument.
+Any sky noise is implicitly included in the time dependence of the sky electric field.
+As the noise of modern low frequency arrays is heavily dominated by sky noise, we will neglect $\widetilde{n}_a$ for now, but will inspect its effects at the end of this section.
 
+The goal of our calibration algorithm will be to estimate the antenna gains.
+We will assume the gains have no time dependence within the timescale of finding our solutions.
+\textcolor{red}{cite someone about gain stability}.
+Furthermore, we will treat each frequency channel independently, and drop the $f$ to simplify notation.
+
+The MOFF algorithm next calls for a calibration.
+For now we will assume we have formed an estimate of the gains after $n$ iterations of a calibration loop, and can apply these gains.
+Typically calibration amounts to dividing the measured fields by the complex gain, or multiplying by the complex conjugate of the gain.
+We will abstain from this choice for now, and instead use a multiplicative factor $h^{(n)}_a$ to represent the application of our $n^{\mathrm{th}}$-loop  estimate of the gain for antenna $a$.
 \begin{equation}
-\Er{a} = g_a E^T_a
+\hat{\Er{a}} = h^{(n)}_a \Er{a}
 \end{equation}
-\begin{align}
-\hat{E}(\spix) & = \frac{1}{\Nant} \sum_i e^{2\pi i \spix \cdot \ri} \Erest{i} \\
-& = \frac{1}{\Nant} \sum_i e^{2\pi i \spix \cdot \ri} \sum_b \beamr_b(\ri-\rb) \Erest{b} \\
-& = \frac{1}{\Nant} \sum_i e^{2\pi i \spix \cdot \ri} \sum_b \beamr_b(\ri-\rb) h^{(n)}_b g_b\Er{b}^T
-\end{align}
-We can then play a trick to transform the beam.
-\begin{align}
-\hat{E}(\spix) & = \frac{1}{\Nant} \sum_b h^{(n)}_b g_b\Er{b}^T e^{2\pi i \spix \cdot \rb}\sum_i \beamr_b(\ri-\rb)e^{2\pi i \spix \cdot (\ri-\rb)}\\
-& = \frac{1}{\Nant} \sum_b h^{(n)}_b g_b\Er{b}^T e^{2\pi i \spix \cdot \rb}\beamtheta_b(\spix)
+
+A dirty image is next formed by gridding the calibrated fields with the antenna voltage pattern, Fourier transforming, squaring, and averaging in time.
+The estimated value for a pixel, \spix, can be expressed as
+\begin{equation}
+\hat{I}(\spix) = \left<\left| \frac{1}{\Nant}\sum_i e^{2\pi i \mathbf{r}\cdot \spix} \sum_a \beamr_a(\ri - \ra) h^{(n)}_a g^T_a \Er{a}^T(t) \right|^2 \right>_t.
+\end{equation}
+This is the final output of the MOFF correlator. However, for calibration purposes we will be interested in the electric field image just prior to squaring and averaging.
+\begin{equation}
+\hat{E}(\spix,t) = \frac{1}{\Nant} \sum_i e^{2\pi i \ri \cdot \spix} \sum_a \beamr_a(\ri - \ra) h^{(n)}_a g^T_a \Er{a}^T(t)
+\end{equation}
+We can simplify this expression by exchanging the sums to transform the beam term into sky coordinates.
+\begin{align}\label{eq:epix}
+\hat{E}(\spix,t) & = \frac{1}{\Nant} \sum_a h^{(n)}_a g^T_a\Er{a}^T e^{2\pi i \spix \cdot \ra}\sum_i \beamr_b(\ri-\ra)e^{2\pi i \spix \cdot (\ri-\ra)} \nonumber\\
+& = \frac{1}{\Nant} \sum_a h^{(n)}_a g^T_a\Er{a}^T e^{2\pi i \spix \cdot \ra}\beamtheta_a(\spix)
 \end{align}
 
-Plugging in, we get,
-\begin{align}
-\Cna & = \left<\Er{a} \hat{E}^*(\spix)\right>_t \\ 
-& = \left<g_a \Er{a}^T \frac{1}{\Nant} \sum_b h^{*(n)}_b g^*_b\Er{b}^{*T} e^{-2\pi i \spix \cdot \rb} \beamtheta^*_b(\spix)\right>_t \\
+Next we move toward the feedback calibration outlined in \citealt{mor11}.
+The calibration loop described there assumed a simple sky of a single point source, but we aim for a more generalized solution for arbitrarily complex sky models.
+As such the quantity we wish to inspect is a correlation of the uncalibrated antenna signals input to the correlator with the holographic electric field image output from the correlator.
+We define the quantity,
+\begin{equation}\label{eq:Cna}
+\Cna \equiv \left<\Er{a}(t) \hat{E}^*(\spix,t)\right>_t,
+\end{equation}
+where the superscript $n$ again represents the quantity formed in the $n^\mathrm{th}$ calibration loop.
+Naturally, because the right hand side contains the output of the correlator, this will be an iterative process.
+The following will hold for any chosen pixel, \spix, though we will see it is advantageous to choose a pixel which contains a bright source.
+
+Plugging equation \ref{eq:epix} into equation \ref{eq:Cna}, we find,
+\begin{align}\label{eq:cna}
+\Cna & = \left<g^T_a \Er{a}^T(t) \frac{1}{\Nant} \sum_b h^{*(n)}_b g^{T*}_b\Er{b}^{T*}(t) e^{-2\pi i \spix \cdot \rb} \beamtheta^*_b(\spix)\right>_t \nonumber \\
+& = \frac{g^T_a}{\Nant} \sum_b h^{*(n)}_b g^{T*}_b \beamtheta^*_b(\spix) e^{-2\pi i \spix \cdot \rb} \left<\Er{a}^T \Er{b}^{*T} \right>_t \nonumber \\
+& = \frac{g^T_a}{\Nant} \sum_b h^{*(n)}_b g^{T*}_b \beamtheta^*_b(\spix) e^{-2\pi i \spix \cdot \rb} \V^T_{ab}
 \end{align}
-Pulling time independent pieces out.
-\begin{align}
-\Cna & = \frac{g_a}{\Nant} \sum_b h^{*(n)}_b g^*_b \beamtheta^*_b(\spix) e^{-2\pi i \spix \cdot \rb} \left<\Er{a}^T \Er{b}^{*T} \right>_t \\
-& = \frac{g_a}{\Nant} \sum_b h^{*(n)}_b g^*_b \beamtheta^*_b(\spix) e^{-2\pi i \spix \cdot \rb} \V^T_{ab}
-\end{align}
-Solve for $g^{(n+1)}_a$.
-\begin{equation}
-g^{(n+1)}_a = \Cna \Nant \left[ \sum_b h^{*(n)}_b g^{*(n)}_b \beamtheta^*_b(\spix) e^{-2\pi i \spix \cdot \rb} \V^T_{ab} \right]^{-1}
+where in the second step we group time-dependent terms, and in the third we define the true visibilities as the correlation between true antenna electric field measurements.
+It is easy to see from here that the net effect of including the receiver noise term, $\widetilde{n}$
+
+Finally, we estimate our solution by assuming our current estimate of the gains, $\hat{g}^{(n)}_b$ is approximately correct and substitute into the sum for the true gains.
+We also require model visibilities formed from sky and primary beam models in place of true visibilities.
+\begin{equation}\label{eq:cal_solution}
+\hat{g}^{(n+1)}_a = \Cna \Nant \left[ \sum_b h^{*(n)}_b \hat{g}^{*(n)}_b \beamtheta^*_b(\spix) e^{-2\pi i \spix \cdot \rb} \V^T_{ab} \right]^{-1}
 \end{equation}
-
+This equation is our prescription for estimating the antenna gains of a direct imaging array.
+The approach is iterative in nature, and requires a sky model.
+However, the sky model can be precomputed offline, and the online computation complexity scales only as $\mathcal{O}(\Nant)$ as we form a $\Cna$ for each antenna.
 In the case where $h_b = 1/g_b$, this simplifies slightly.
@@
 
+We note an encouraging analogy to the solutions found in \citealt{mit08} (their equation~11).
+In the case where \spix\, is the central pixel of the image, we see \Cna\, is simply the sum of visibilities involving antenna $a$ in equation~\ref{eq:cna}.
+This is equivalent to the numerator equation~11 in \citealt{mit08}, in the single polarization case.
+The denominator then is the sum over modeled visibilities, as in our equation~\ref{eq:cal_solution} (save bookkeeping to track Fourier transform conventions and gain applications).
+The major difference in our solution is the inclusion of auto-correlations in both sums.
+If desired, a deployed system could adapt equation~\ref{eq:cal_solution} to remove the auto-correlations.
+This would require generating these correlations, which again only scale as $\mathcal{O}(\Nant)$, and are generally useful for array diagnostics.
+
+We thus confirm what was shown for the simple-sky case in \citealt{mor11} -- our calibration solutions are equivalent to those derived from visibility data.
+We will pursue this further in section~\ref{sec:noise}.
+
+We show schematically the process of calibrating a direct imaging correlator in figure~\ref{fig:schematic}.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figures/schematic.pdf}
+\caption{Caption goes here}
+\label{fig:schematic}
+\end{center}
+\end{figure}
+
 \section{Simulation}\label{sec:sim}
-
-Describe simulation
+We first demonstrate our calibration method through a controlled simulation.
 
 \section{Application to LWA data}\label{sec:data}
+Reference for Cyg-A: \citealt{coh07}. Ref for Cas-A: \citealt{kas07}.
 Apply to LWA data.

### 2015-11-18 cal_demo/ms.tex
tweaking wording, adding text

@@
 \newcommand{\spix}{\ensuremath{\hat{\mathbf{s}}_{\mathrm{pix}}}}
-\newcommand{\Cna}{\ensuremath{\mathcal{C}^{(n)}_{a,\spix}}}
+\newcommand{\Cna}[1][n]{\ensuremath{\mathcal{C}^{(#1)}_{a,\spix}}}
 \newcommand{\ri}{\ensuremath{\mathbf{r}_i}}
@@
 \newcommand{\dif}{\mathrm{d}}
+\newcommand{\caliter}{50}
+\newcommand{\damp}{\ensuremath{\Delta}}
+\newcommand{\itr}{20}
 
@@
 We next model the measured, uncalibrated electric field as a multiplicative complex gain and an additive noise term applied to the true antenna electric field. 
-\begin{equation}
+\begin{equation}\label{eq:apply_gain}
 \Er{a}(f,t) = g^T_a(f,t) \Er{a}^T(f,t) + \widetilde{n}_a(f,t)
@@
 \begin{equation}
-g^{(n+1)}_a = \Cna \Nant \left[ \sum_b \beamtheta^*_b(\spix) e^{-2\pi i \spix \cdot \rb} \V^T_{ab} \right]^{-1}
+\hat{g}^{(n+1)}_a = \Cna \Nant \left[ \sum_b \beamtheta^*_b(\spix) e^{-2\pi i \spix \cdot \rb} \V^T_{ab} \right]^{-1}
 \end{equation}
@@
 We note an encouraging analogy to the solutions found in \citealt{mit08} (their equation~11).
-In the case where \spix\, is the central pixel of the image, we see \Cna\, is simply the sum of visibilities involving antenna $a$ in equation~\ref{eq:cna}.
+In the case where \spix~is the central pixel of the image, we see \Cna\, is simply the sum of visibilities involving antenna $a$ in equation~\ref{eq:cna}.
 This is equivalent to the numerator equation~11 in \citealt{mit08}, in the single polarization case.
@@
 We show schematically the process of calibrating a direct imaging correlator in figure~\ref{fig:schematic}.
+Computationally expensive steps that must be performed ``on-chip" are shown inside the gray box.
+The uncalibrated antenna signals are tapped out after the F-engine and correlated against the output image pixel of interest.
+The correlated values are then passed off-chip to estimate the gains using equation~\ref{eq:cal_solution}, and additional fitting if desired.
+The gains are then passed back to the correlator to update the calibration for subsequent data streams.
 
@@
 \includegraphics[width=\columnwidth]{figures/schematic.pdf}
-\caption{Caption goes here}
+\caption{\textcolor{red}{Caption goes here}}
 \label{fig:schematic}
@@
 We first demonstrate our calibration method through a controlled simulation.
+The simulation software used is included in the EPIC package, and we follow a procedure based on that described in \citealt{thy15c}, using the same toy-model sky with 10 point sources of flux density 10 Jy at random location.
+We assume perfect knowledge of the sky to build our model visibilities.
+
+We use 64 frequency channels each 40 kHz, and each timeseries is 25~$\mu$s long.
+However, because each channel is independent in the calibration loop we only show the central channel here.
+Using the MWA array layout \citep{bea12}, we select the inner 51 antennas within a square bounding box of 150~m.
+The antenna voltage pattern used is a 4.4~m square tophat on the ground.
+
+In order to show our ability to recover unknown complex antenna gains, we create a set of random numbers with amplitude $1\pm0.25$, and completely random phase.
+These are our ``true gains", and we apply them to the frequency-domain simulated antenna electric fields as in equation~\ref{eq:apply_gain}.
+Our analysis is blind to these values until the end of the process to check accuracy.
+We start with an initial guess of unity gains.
+
+We next process \caliter~time stamps, creating both a short-term integrated image and \Cna[0] correlations.
+Unless otherwise specified, the software automatically selects the brightest source in the sky model to choose the correlation pixel.
+We use our correlation values to update the gains.
+We found that simply using equation~\ref{eq:cal_solution} resulted in large oscillations in the solutions, which eventually converged.
+To mitigate this we introduce a tunable damping factor, $0< \damp \leq 1$, which is used to attenuate the gain update, effectively giving the solutions memory of previous iterations.
+Through experimentation we found $\Delta=0.65$ resulted in the quickest convergence for this simulation.
+
+We continue the calibration loop by updating our gain estimates, dividing the data by the gains, forming images and \Cna correlations to form the next estimate.
+This process was repeated \itr~times.
+In figure~\ref{fig:before_image} we show the integrated image created using our initial guess gains and the first \caliter~time steps.
+Because the gain phases were completely random the image is
 

### 2016-01-22 cal_demo/ms.tex
Adding lots of cal demo text and figures

--- 2015-11-17 cal_demo/ms.tex
@@
 
-\newcommand{\Nant}{\ensuremath{N_{\mathrm{ant}}}}
+\newcommand{\Nant}{\ensuremath{N_{\mathrm{a}}}}
 \newcommand{\Ng}{\ensuremath{N_{\mathrm{g}}}}
 \newcommand{\s}{\ensuremath{\hat{\mathbf{s}}}} % s-hat for sine-projected direction
-\newcommand{\spix}{\ensuremath{\hat{\mathbf{s}}_{\mathrm{pix}}}}
-\newcommand{\Cna}{\ensuremath{\mathcal{C}^{(n)}_{a,\spix}}}
+\newcommand{\spix}{\ensuremath{\hat{\mathbf{s}}_{0}}}
+\newcommand{\Cna}[1][n]{\ensuremath{\mathcal{C}^{(#1)}_{a,\spix}}}
 \newcommand{\ri}{\ensuremath{\mathbf{r}_i}}
@@
 \newcommand{\Er}[1]{\ensuremath{\widetilde{E}_{#1}}}
-\newcommand{\Erest}[1]{\ensuremath{\hat{\widetilde{E}}_{#1}}}
+\newcommand{\Erest}[1]{\ensuremath{\widetilde{E}'_{#1}}}
+\newcommand{\Ethetaest}{\ensuremath{E'}}
 \newcommand{\V}{\ensuremath{\widetilde{V}}}
 \newcommand{\dif}{\mathrm{d}}
+\newcommand{\caliter}{400}
+\newcommand{\damp}{\ensuremath{\gamma}}
+\newcommand{\itr}{20}
 
@@
 In order to satisfy the survey speeds required for precision cosmology as well as searches for fast radio transients, radio astronomy is undergoing a paradigm shift toward interferometers consisting of hundreds to thousands of small, widefield antennas.
-Many arrays with this design are already built or under construction including the Hydrogen Epoch of Reionization Array\footnote{http://reionization.org} (HERA), the Murchison Widefield Array (MWA;\citealt{tin13,bow13}), the Precision Array for Probing the Epoch of Reionization (PAPER; \citealt{par10}), the LOw Frequency ARray (LOFAR;\citealt{van13}), the Canadian Hydrogen Intensity Mapping Experiment (CHIME,\citealt{ban14}), the Long Wavelength Array (LWA, \citealt{ell13}), and the low frequency Square Kilometer Array (SKA1-Low \citealt{mel13}).
+Many arrays with this design are already built or under construction including the Hydrogen Epoch of Reionization Array\footnote{http://reionization.org} (HERA), the Murchison Widefield Array (MWA;\citealt{tin13,bow13}), the Precision Array for Probing the Epoch of Reionization (PAPER; \citealt{par10}), the LOw Frequency ARray (LOFAR; \citealt{van13}), the Canadian Hydrogen Intensity Mapping Experiment (CHIME,\citealt{ban14}), the Long Wavelength Array (LWA, \citealt{ell13}), and the low frequency Square Kilometer Array (SKA1-Low \citealt{mel13}).
 
@@
 Meanwhile, radio transient monitoring requires access to high time and frequency resolution data.
-For example, Fast Radio Bursts (FRBs) are highly unexplored at low frequencies (< 1 GHz), but are expected to occur on timescales $\Delta t \sim$ 1--10~ms \citep{tho13}.
+For example, fast radio bursts (FRBs) are highly unexplored at low frequencies (< 1 GHz), but are expected to occur on timescales $\Delta t \sim$ 1--10~ms \citep{tho13}.
 Recording the full visibility matrix for $\Nant \gtrsim 10^3$ arrays at this timescale leads to extremely high data write rates.
@@
 This method uses a correlation of the uncalibrated antenna signal stream with an output image pixel from the backend of the correlator to solve for the complex gains of the antennas.
-In the limit of a simple sky, the algorithm reduces to the self-cal algorithm, and even in a complex limit can be shown to be equivalent to visibility-based solutions.
-An example implementation of the algorithm is available with the EPIC software package\footnote{http://github.com/nithyanandan/EPIC}
+Because the calibration must be applied before gridding and imaging, our solution requires an iterative approach where the data from one time series is used to update the gains which are applied to the following time series.
+An example implementation of the algorithm is available with the EPIC software package\footnote{http://github.com/nithyanandan/EPIC}.
+
 We establish the mathematical framework and derive the calibration algorithm in \S \ref{sec:math}.
@@
 We will encounter several quantities which we attempt to estimate.
-We distinguish the ``true" values with a superscript $T$, while the estimates are denoted with a hat.
+We distinguish the ``true" values with a superscript $T$, while the estimates are denoted with a prime.
 We define the true antenna signal as a convolution of the antenna voltage pattern, $\beamr$, with the electric field on the ground.
@@
 The MOFF algorithm next calls for a calibration.
-For now we will assume we have formed an estimate of the gains after $n$ iterations of a calibration loop, and can apply these gains.
-Typically calibration amounts to dividing the measured fields by the complex gain, or multiplying by the complex conjugate of the gain.
-We will abstain from this choice for now, and instead use a multiplicative factor $h^{(n)}_a$ to represent the application of our $n^{\mathrm{th}}$-loop  estimate of the gain for antenna $a$.
+We will assume we have formed an estimate of the gains after $n$ iterations of a calibration loop, and derive an updated $n+1$ estimate.
+Typically calibration amounts to dividing the measured fields by the gain (in the case of sky noise dominance), or multiplying by the complex conjugate of the gain (receiver noise dominance).
+We will abstain from this choice for now, and instead use a multiplicative factor $h^{(n)}_a$ to represent the application of our $n^{\mathrm{th}}$-loop estimate of the gain for antenna $a$:
 \begin{equation}
-\hat{\Er{a}} = h^{(n)}_a \Er{a}
+\Erest{a} = h^{(n)}_a \Er{a}
+\end{equation}
+where,
+\begin{equation}
+h^{(n)}_a=\begin{cases}
+1/g^{(n)}_a & \mbox{sky noise dominated} \\ 
+g^{*(n)}_a & \mbox{receiver noise dominated}.
+\end{cases}
 \end{equation}
@@
 A dirty image is next formed by gridding the calibrated fields with the antenna voltage pattern, Fourier transforming, squaring, and averaging in time.
-The estimated value for a pixel, \spix, can be expressed as
+The estimated value for a pixel, $\s_i$, can be expressed as
 \begin{equation}
-\hat{I}(\spix) = \left<\left| \frac{1}{\Nant}\sum_i e^{2\pi i \mathbf{r}\cdot \spix} \sum_a \beamr_a(\ri - \ra) h^{(n)}_a g^T_a \Er{a}^T(t) \right|^2 \right>_t.
+I'(\s_i) = \left<\left| \frac{1}{\Nant}\sum_i e^{2\pi i \mathbf{r}\cdot \s_i} \sum_a \beamr_a(\ri - \ra) h^{(n)}_a g^T_a \Er{a}^T(t) \right|^2 \right>_t.
 \end{equation}
@@
 \begin{equation}
-\hat{E}(\spix,t) = \frac{1}{\Nant} \sum_i e^{2\pi i \ri \cdot \spix} \sum_a \beamr_a(\ri - \ra) h^{(n)}_a g^T_a \Er{a}^T(t)
+\Ethetaest (\s_i,t) = \frac{1}{\Nant} \sum_i e^{2\pi i \ri \cdot \s_i} \sum_a \beamr_a(\ri - \ra) h^{(n)}_a g^T_a \Er{a}^T(t)
 \end{equation}
@@
 \begin{align}\label{eq:epix}
-\hat{E}(\spix,t) & = \frac{1}{\Nant} \sum_a h^{(n)}_a g^T_a\Er{a}^T e^{2\pi i \spix \cdot \ra}\sum_i \beamr_b(\ri-\ra)e^{2\pi i \spix \cdot (\ri-\ra)} \nonumber\\
-& = \frac{1}{\Nant} \sum_a h^{(n)}_a g^T_a\Er{a}^T e^{2\pi i \spix \cdot \ra}\beamtheta_a(\spix)
+\Ethetaest(\s_i,t) & = \frac{1}{\Nant} \sum_a h^{(n)}_a g^T_a\Er{a}^T(t) e^{2\pi i \s_i \cdot \ra}\sum_i \beamr_b(\ri-\ra)e^{2\pi i \s_i \cdot (\ri-\ra)} \nonumber\\
+& = \frac{1}{\Nant} \sum_a h^{(n)}_a g^T_a\Er{a}^T(t) e^{2\pi i \s_i \cdot \ra}\beamtheta_a(\s_i)
 \end{align}
@@
 \begin{equation}\label{eq:Cna}
-\Cna \equiv \left<\Er{a}(t) \hat{E}^*(\spix,t)\right>_t,
+\Cna \equiv \left<\Er{a}(t) E'^*(\spix,t)\right>_t,
 \end{equation}
-where the superscript $n$ again represents the quantity formed in the $n^\mathrm{th}$ calibration loop.
-Naturally, because the right hand side contains the output of the correlator, this will be an iterative process.
-The following will hold for any chosen pixel, \spix, though we will see it is advantageous to choose a pixel which contains a bright source.
+where the superscript $n$ again represents the quantity formed in the $n^\mathrm{th}$ calibration loop, and \spix\, is the pixel center nearest a bright calibrator of interest.
+The following will hold for any chosen pixel, \spix, though it is advantageous to choose a pixel which contains a bright source to achieve a high signal to noise.
 
@@
 where in the second step we group time-dependent terms, and in the third we define the true visibilities as the correlation between true antenna electric field measurements.
-It is easy to see from here that the net effect of including the receiver noise term, $\widetilde{n}$
+It is easy to see from here that the net effect of including the receiver noise term, $\widetilde{n}_a(f,t)$, will result in added noise on the true visibilities.
+In principle this can include any noise correlations between antennas.
+The implementation included in the EPIC software package restricts the noise term to include only the auto-correlation noise bias, and assumes baseline dependent noise is zero mean.
 
-Finally, we estimate our solution by assuming our current estimate of the gains, $\hat{g}^{(n)}_b$ is approximately correct and substitute into the sum for the true gains.
+Finally, we find an update to our gain solution by assuming our current estimate of the gains, $g^{(n)}_b$, is approximately correct and substitute into the sum for the true gains.
 We also require model visibilities formed from sky and primary beam models in place of true visibilities.
 \begin{equation}\label{eq:cal_solution}
-\hat{g}^{(n+1)}_a = \Cna \Nant \left[ \sum_b h^{*(n)}_b \hat{g}^{*(n)}_b \beamtheta^*_b(\spix) e^{-2\pi i \spix \cdot \rb} \V^T_{ab} \right]^{-1}
+g'^{(n+1)}_a = \Cna \Nant \left[ \sum_b h^{*(n)}_b g^{*(n)}_b \beamtheta^*_b(\spix) e^{-2\pi i \spix \cdot \rb} \V^T_{ab} \right]^{-1}
 \end{equation}
@@
 However, the sky model can be precomputed offline, and the online computation complexity scales only as $\mathcal{O}(\Nant)$ as we form a $\Cna$ for each antenna.
-In the case where $h_b = 1/g_b$, this simplifies slightly.
-\begin{equation}
-g^{(n+1)}_a = \Cna \Nant \left[ \sum_b \beamtheta^*_b(\spix) e^{-2\pi i \spix \cdot \rb} \V^T_{ab} \right]^{-1}
+In the case where $h_b = 1/g_b$ (typical calibration procedure for sky-noise dominated systems), this simplifies slightly.
+\begin{equation}\label{eq:cal_solution_simple}
+g'^{(n+1)}_a = \Cna \Nant \left[ \sum_b \beamtheta^*_b(\spix) e^{-2\pi i \spix \cdot \rb} \V^T_{ab} \right]^{-1}
 \end{equation}
 
-We note an encouraging analogy to the solutions found in \citealt{mit08} (their equation~11).
-In the case where \spix\, is the central pixel of the image, we see \Cna\, is simply the sum of visibilities involving antenna $a$ in equation~\ref{eq:cna}.
-This is equivalent to the numerator equation~11 in \citealt{mit08}, in the single polarization case.
-The denominator then is the sum over modeled visibilities, as in our equation~\ref{eq:cal_solution} (save bookkeeping to track Fourier transform conventions and gain applications).
-The major difference in our solution is the inclusion of auto-correlations in both sums.
-If desired, a deployed system could adapt equation~\ref{eq:cal_solution} to remove the auto-correlations.
-This would require generating these correlations, which again only scale as $\mathcal{O}(\Nant)$, and are generally useful for array diagnostics.
-
-We thus confirm what was shown for the simple-sky case in \citealt{mor11} -- our calibration solutions are equivalent to those derived from visibility data.
-We will pursue this further in section~\ref{sec:noise}.
+While testing we found equation~\ref{eq:cal_solution} resulted in oscillatory gain solutions as it was iterated, as is often the case in iterative minimization methods.
+To mitigate this we introduce a damping factor, $0 \leq \damp <1$, which is used to attenuate the gain update, effectively giving the solutions memory of previous iterations.
+\begin{equation}
+g^{(n+1)}_a = (1-\damp) g'^{(n+1)}_a + \damp g^{(n)}_a
+\end{equation}
+We found that while equation~\ref{eq:cal_solution} does indeed converge on good solutions, the process is faster by tuning the damping factor.
+We note the small difference between $g^{(n+1)}_a$ and $g'^{(n+1)}_a$, where the prime indicates our current best estimate of the true gain, while no prime is actually used in the iterative calibration loop.
+Once the loop converges the difference is a longer effective integration for the non-primed version (lower thermal noise).
 
 We show schematically the process of calibrating a direct imaging correlator in figure~\ref{fig:schematic}.
+Computationally expensive steps that must be performed ``on-chip" are shown inside the gray box.
+The uncalibrated antenna signals are tapped out after the F-engine and correlated against the output image pixel of interest.
+The correlated values are then passed off-chip to estimate the gains using equation~\ref{eq:cal_solution}, and additional fitting if desired.
+The gains are then passed back to the correlator to update the calibration for subsequent data streams.
 
@@
 
+An important feature to note is that, like the MOFF-generated images themselves, equations~\ref{eq:cal_solution} and~\ref{eq:cal_solution_simple} include the antenna auto-correlations (the sum is over \emph{all} $b$, not excluding $a$).
+It can be difficult to perfectly model the noise bias from auto-correlations, which can often times be far brighter than the visibilities themselves.
+It can therefore be beneficial to subtract this term directly from \Cna, and exclude the $b=a$ term in the sum.
+\begin{equation}
+\Cna \rightarrow \Cna - \frac{1}{\Nant} h^{*(n)}_a\beamtheta^*_a e^{-2\pi i \spix \cdot \ra} \left<|\Er{a}|^2\right>_t
+\end{equation}
+This requires generating these correlations, which again only scale as $\mathcal{O}(\Nant)$, and are generally useful for array diagnostics.
+
+We conclude this section by connecting our calibration expression to that found in a visibility framework.
+In the limit of a single bright calibrating source at phase center, we can greatly simplify equations~\ref{eq:cna} and~\ref{eq:cal_solution}.
+We will assume the beams are normalized such that $\beamtheta(0)=1$.
+We can further drop the exponential phase terms because $\spix=0$.
+We then absorb the true gains and gain corrections into the true visibilities in equation~\ref{eq:cna} to express as a sum of measured visibilities.
+\begin{equation}
+\mathcal{C}^{(n)}_{a,0} \rightarrow \frac{1}{\Nant}\sum_b h^{*(n)}_b \V_{ab}
+\end{equation}
+
+We next plug this expression into equation~\ref{eq:cal_solution} to find our simplified calibration solution for a single bright point source.
+Because our sky is a single bright point source, the model visibilities are simply the flux of the source, $S_{\mathrm{src}}$.
+\begin{equation}
+g'^{(n+1)}_a \rightarrow \left[\sum_b h^{*(n)}_b \V_{ab}\right] \times \left[S_{\mathrm{src}}\sum_b h^{*(n)}_b g^{*(n)}_b\right]^{-1}
+\end{equation}
+This is simply a gain-weighted sum of the measured visibilities over the flux of the source, which is indeed the limiting result from a visibility approach, for example seen in \citealt{mit08}.
+The ability to recover the equivalent expression despite not actually forming the visibilities is a result of the fact that only sums over visibilities come into the FX solution, as was described in \citealt{mor11}.
+We have confirmed the limiting case equivalence here, and will explore the more general case in more detail in \S~\ref{sec:noise}.
+
 \section{Simulation}\label{sec:sim}
 We first demonstrate our calibration method through a controlled simulation.
+A complex gain is created for each antenna with random phase and amplitude, which is used to corrupt the simulated data stream, then we attempt to recover the gains using our calibration routine.
+The simulation software used is included in the EPIC package.
+
+Our simulated signal consists of 10 random point sources with flux densities 0.5~Jy~$\lesssim S \lesssim 1$~Jy.
+For an antenna array we use the inner 51 antennas of the MWA layout \citep{bea12}, within a bounding box of 150~m.
+The antenna voltage pattern used is a 4.4~m square tophat on the ground.
+Because our algorithm treats frequency channels independently, we simulate only one channel.
+For context we treat this channel as a single 40 kHz, meaning each subsequent timestep is separated by 25~$\mu$s.
+
+For our unknown gains, we create a set of random numbers with amplitude $1\pm0.25$, and completely random phase.
+These are our ``true gains", and we apply them to the frequency-domain simulated antenna electric fields as in equation~\ref{eq:apply_gain}.
+Our analysis is blind to these values until the end of the process to check accuracy.
+The gain estimates are initialized with unity, $g^{(0)}_a=1$.
+
+We next process and image \caliter~time steps (10~ms).
+We also form the correlations, \Cna[0], used in our calibration loop.
+The pixel used for the correlation is the source with the largest apparent flux (intrinsic flux attenuated by the primary beam).
+These correlation values are used to update the gain estimates, which in turn are used to calibrate the following \caliter~time steps.
+Through experimentation we found a damping factor of $\damp=0.35$ resulted in the quickest convergence in this simulation.
+
+The calibration loop continues by updating the gain estimates every \caliter~time steps.
+The phases of our gain estimates are shown in figure~\ref{fig:sim_phase} for 20 such iterations.
+The phase error plotted is the phase relative to the true gain for each antenna (various colored lines).
+One antenna was used as a reference to fix the absolute phase, so has zero phase error.
+The other 50 antennas are shown to have error spanning $2\pi$ initially, and after about 10 iterations lock into a solution, settling down to noise levels around iteration 15 (0.15~s).
+We stop the simulation when the updated gains trace the thermal noise of the simulated sources, which can be seen by the coherence of the 50 antenna gains after iteration 15.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figures/cal_paper_sim_phase.pdf}
+\caption{Phase error of gain estimates as a function of iteration for simulated calibration.
+The gains were initialized with random phases, but the calibration loop was able to recover the correct phases after about 15 iterations.
+Each line represents an antenna in our 51 MWA antenna sample.
+}
+\label{fig:sim_phase}
+\end{center}
+\end{figure}
+
+The estimated gain amplitudes for the simulations are shown in figure~\ref{fig:sim_amp}.
+The quantity plotted is the magnitude of the estimated gains over the true gains, $\left|g^{(n)}_a/g^T_a\right|$, which places all antennas on the same scale.
+We can see the amplitudes converge toward their true values around the same time as the phases (iteration $\sim$15).
+At the beginning of calibration we can see the value of the damping factor.
+At $n=0$, a couple of gains are shown to have abnormally high amplitude estimates, notably one about 3.3 times its true value (red line).
+These unbalanced high estimates caused the entire set of gains to be under estimated at $n=1$, even with a damping factor of 0.35.
+By $n=5$ the unbalanced amplitudes have been damped out and the calibration continues.
+Without the damping factor, the oscillation seen in the first couple iterations would have been significantly larger and taken much longer to fade out.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figures/cal_paper_sim_amps.pdf}
+\caption{\textcolor{red}{caption!}}
+\label{fig:sim_amp}
+\end{center}
+\end{figure}
+
+Images created at the beginning of calibration and at the end are shown in figure~\ref{fig:sim_images}.
+Each image is 10~ms integration, corresponding to all snapshot images created with a given set of gain estimates.
+The top panel shows the image produced with our initialized unity gains.
+Because the phases are completely random, the image is essentially noise with the primary beam evident.
+After 20 iterations, the image is far more clear, shown in the bottom panel.
+Each of the ten simulated sources are clearly visible, indicated with red circles.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figures/cal_paper_sim_image_before.pdf}
+\includegraphics[width=\columnwidth]{figures/cal_paper_sim_image_after.pdf}
+\caption{\textcolor{red}{caption!}}
+\label{fig:sim_images}
+\end{center}
+\end{figure}
 
@@
 Apply to LWA data.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figures/cal_paper_data_phases.pdf}
+\caption{\textcolor{red}{caption!}}
+\label{fig:data_phase}
+\end{center}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figures/cal_paper_data_amps.pdf}
+\caption{\textcolor{red}{caption!}}
+\label{fig:data_amp}
+\end{center}
+\end{figure}
+
+\begin{figure*}
+\begin{center}
+\includegraphics[width=0.3\linewidth]{figures/cal_paper_data_image_before.pdf}
+\includegraphics[width=0.3\linewidth]{figures/cal_paper_data_image_after.pdf}
+\includegraphics[width=0.3\linewidth]{figures/cal_paper_gsm_beam_weighted.pdf}
+\caption{\textcolor{red}{caption!}}
+\label{fig:data_images}
+\end{center}
+\end{figure*}
 

### 2016-01-25 cal_demo/ms.tex
Adding data text to cal demo

@@
 \section{Application to LWA data}\label{sec:data}
-Reference for Cyg-A: \citealt{coh07}. Ref for Cas-A: \citealt{kas07}.
-Apply to LWA data.
+We next demonstrate our calibration algorithm using an observation from the LWA station in New Mexico.
+The data is from the LWA narrow-band transient buffer (TBN), with time ordered voltage data from 255 antennas within a core radius of 100~m.
+The central frequency is 74.03 MHz, and a bandwidth of 100 kHz and writeout timescale of 5.12 ms (frequency channel resolution of 195.3125 Hz).
+For this demonstration we limit ourselves to a single polarization.
+
+After correcting for geometric cable delays, the instrument is naturally well calibrated, as was seen in the demonstration of the EPIC imager in \citealt{thy15c}.
+However, we will aim to improve yet on this calibration using our algorithm.
+
+We proceed by forming model visibilities.
+We model only two bright objects as point sources: Cyg A with flux 16611.68 Jy \citep{coh07}; and Cas A with flux 17693.9 Jy \citep{kas07}.
+Because the raw data is attenuated by the primary beam of the instrument, we also account for this in our model using beam values consistent with \cite{hic12}.
+
+We made several choices while studying the behavior of the LWA data to improve our calibration.
+We inspected the amplitudes of the voltages to be roughly consistent with average gain amplitudes of 0.25, so we initialized our gains at this level to allow the calibration to converge quickly.
+We also found a boost in signal to noise is achieved easily by averaging frequency channels and assuming the gains are constant within a sub-band.
+Here we average solutions across 150 channels, or about 29 kHz.
+With a fractional bandwidth $B/f_0 = 3.9 \times 10^{-4}$, we felt safe assuming a smooth bandpass.
+Because there is still a fair amount of noise in the data, a damping factor of $\gamma = 0.7$ was adopted.
+
+Finally, we found that seven antennas\footnote{LWA antenna IDs 48, 85, 124, 148, 203, 217, and 244 were flagged.} produced unstable gain solutions, and in fact corrupted the ensemble.
+We therefore flagged these antennas for our analysis, resulting in a total of 248 antennas to calibrate.
+In each calibration loop, we form our feedback correlations and update our gain estimates over 10 timestamps (51.2 ms).
+We iterate the loop 30 times for a total of 1.536 seconds.
+The results of this experiment are shown in figures~\ref{fig:data_phase}~--~\ref{fig:data_images}.
+
+Figure~\ref{fig:data_phase} shows the phase of our gain estimates over our 30 calibration iterations, again with each colored line representing a different antenna.
+Given the quality of uncalibrated image demonstrated in \cite{thy15c}, it is a bit surprising to see the phase variation in our solutions.
+However, the phases are relatively flat after about 15 iterations (modulo noise), and exhibit a central ``trunk" where the majority of phases are congregated.
+This behavior is suggestive that while the uncalibrated voltages were able to produce a viable image, the minor changes from our solutions will focus the image and improve the quality.
+The actual location of the ``trunk" (slightly negative) is simply determined by the reference antenna chosen to have identically zero phase, but happens to be slightly more positive than the bulk of antennas.
+We also note that some phases appear to exceed $\pm \pi$ radians, this is due to the unwrapping done for plotting clarity.
 
@@
 
+The gain amplitudes as a function of calibration iteration are shown in figure~\ref{fig:data_amp}.
+There is a fairly wide range in gain amplitudes (from 0.12 to 0.59).
+However, this is not surprising due to the non-uniformity in the cables from the antennas to the receivers.
+Again, the amplitudes are noisy by relatively flat.
+
 \begin{figure}
@@
 
+Figure~\ref{fig:data_images} shows the improvement in the images due to our calibration.
+The left panel shows the uncalibrated image integrated over 51.2 ms, 29 kHz.
+Cyg~A is prominent near the center of the image, and Cas~A is also clearly visible in the upper right.
+The middle panel shows the image produced after calibration with identical integration time and bandwidth.
+The sidelobes throughout the image are significantly suppressed, and the galactic plane is much more evident.
+We also note that the feature just to the right of Cyg~A is dimmer in the calibrated image, better matching the expected flux from the global sky model (GSM, \citealt{deo08}).
+We show the GSM modulated by two factors of the LWA primary beam in the right panel of figure~\ref{fig:data_images} for reference.
+
 \begin{figure*}
 \begin{center}
-\includegraphics[width=0.3\linewidth]{figures/cal_paper_data_image_before.pdf}
-\includegraphics[width=0.3\linewidth]{figures/cal_paper_data_image_after.pdf}
-\includegraphics[width=0.3\linewidth]{figures/cal_paper_gsm_beam_weighted.pdf}
+\includegraphics[width=0.33\linewidth]{figures/cal_paper_data_image_before.pdf}
+\includegraphics[width=0.33\linewidth]{figures/cal_paper_data_image_after.pdf}
+\includegraphics[width=0.33\linewidth]{figures/cal_paper_gsm_beam_weighted.pdf}
 \caption{\textcolor{red}{caption!}}
@@
 \end{figure*}
+
+To compare the image qualities quantitatively we compute the dynamic range, defined as the peak of the image over the noise level.
+We estimate the noise level as the median of the absolute deviation of the image.
+With this metric we find the calibrated image to have a 55\% dynamic range improvement over the uncalibrated image.
 

### 2016-01-25 cal_demo/ms.tex
setting up noise analysis section of cal demo

@@
 
-Connect to either cramer-rao or FX solutions in some way
+Here we study the properties of our calibration solutions with respect to noise in the system.
+In order to compare to solutions from visibility-based calibration, we first derive the theoretical lower bound of the error in these solutions, given a set of measured visibilities, $\mathbf{V}$.
+We will do this through the Fischer information formalism and employ the Cram\'er-Rao lower bound
 
@@
 \end{equation}
-Then the likelihood of the set of all visibilities will be,
+Then the likelihood of the set of all visibilities is the product of all individual likelihoods.
 \begin{equation}
@@
 \begin{equation}
-\mathbf{F}^g_{ij} = \left<\left. \frac{\partial \ln \mathcal{L}(\mathbf{V};\mathbf{g})}{\partial g_i} \frac{\partial \ln \mathcal{L}(\mathbf{V};\mathbf{g})}{\partial g_j}\right|_\mathbf{g}\right>.
+\mathbf{F}^g_{ij} = \left<\left. \frac{\partial \ln \mathcal{L}(\mathbf{V};\mathbf{g})}{\partial g_i} \frac{\partial \ln \mathcal{L}(\mathbf{V};\mathbf{g})}{\partial g_j}\right|_\mathbf{g}\right>,
 \end{equation}
+where the expectation is taken over all possible visibility measurements.
 We next evaluate the derivate of the log-likelihood.
@@
 
-The expected values are easy to evaluate. Each visibility will average to the ``true" value times the respective gains. The term with two visibilities will include a noise term.
+The expected values are easy to evaluate.
+Each visibility will average to the ``true" value times the respective gains.
+The term with two visibilities will include a noise bias on autocorrelation terms.
 \begin{equation}
@@
 \end{equation}
+
+We next run a set of EPICal simulations to compare the noisy results to the theoretical lower bound given a set of equivalent visibilities.
+Each simulation is run with a simple sky of a single point source with flux 1 Jy.
+We add noise to the system by adding a Gaussian-random complex value to each channelized electric field antenna measurement as in equation~\ref{eq:apply_gain}.
+The ``true" gains in our simulation are unity, as well as our initial estimate.
+This emulates a system where the initially volatile gain correction has been suppressed and only noise remains to corrupt our estimation, allowing us to run many simulations on a shorter timescale.
+
+We vary two parameters in our simulations: the amount of noise added to our measurements, and the length of integration in each calibration loop.
 
@@
 
-\appendix
+%\appendix
 
-\section{Some extra material}
+%\section{Some extra material}
 
-If you want to present additional material which would interrupt the flow of the main paper,
-it can be placed in an Appendix which appears after the list of references.
+%If you want to present additional material which would interrupt the flow of the main paper,
+%it can be placed in an Appendix which appears after the list of references.
 

### 2016-01-26 cal_demo/ms.tex
adding some captions to figures

@@
 \includegraphics[width=\columnwidth]{figures/schematic.pdf}
-\caption{\textcolor{red}{Caption goes here}}
+\caption{The general data flow of the MOFF correlator, with a feedback calibration loop.
+A pixel from the (unsquared) image is tapped out and correlated against the input antenna electric field signals to form $\Cna$ coefficients, which are used to estimate the antenna gains.
+These gain estimates are then fed back into the correlator to be applied to subsequent data, and the process is repeated.
+The gray box shows operations which must be done at high speed (before averaging in time), which white boxes show operations which can be performed ``off-chip".}
 \label{fig:schematic}
@@
 \includegraphics[width=\columnwidth]{figures/cal_paper_sim_amps.pdf}
-\caption{\textcolor{red}{caption!}}
+\caption{Gain estimate amplitudes as a function of iteration for simulated calibration.
+Again, each line represents an antenna in the 51 MWA antenna sample.
+The gain estimates were initialized randomly, while the true values were unity.
+After about 15 iterations we see the calibration loop has settled around the correct values, with only noise remaining.}
 \label{fig:sim_amp}
@@
 \includegraphics[width=\columnwidth]{figures/cal_paper_sim_image_after.pdf}
-\caption{\textcolor{red}{caption!}}
+\caption{Images formed during simulated calibration.
+\emph{Top:} An image generated from a single frequency channel and 10 ms integration after the gain estimates are randomized.
+As expected with random phases, the image is completely noisy, with the shape of the primary beam evident.
+\emph{Bottom:} An image formed after calibration, again with a single frequency channel and 10 ms integration.
+The ten simulated (and modeled) point sources are easily visible, and highlighted with red circles.
+}
 \label{fig:sim_images}
@@
 \includegraphics[width=\columnwidth]{figures/cal_paper_data_phases.pdf}
-\caption{\textcolor{red}{caption!}}
+\caption{Gain phase solutions as a function of calibration iteration for an LWA TBN observation.
+The gain estimates are initialized with zero phase, but quickly span a $2\pi$ range, and settle into relatively flat, albeit noisy, solutions.
+The majority of phases congregate near zero, which is not surprising given the fairly good quality image produced from uncalibrated data.
+}
 \label{fig:data_phase}
@@
 \includegraphics[width=\columnwidth]{figures/cal_paper_data_amps.pdf}
-\caption{\textcolor{red}{caption!}}
+\caption{Gain amplitudes as a function of calibration iteration for an LWA TBN observation.
+The gains estimates were initialized with amplitude 0.25 after inspection of the electric field values compared to our model sky.
+The solutions are noisy, but flat.
+The range in amplitudes is due to the non-uniformity of cables between the LWA antennas and receivers.
+}
 \label{fig:data_amp}
@@
 \includegraphics[width=0.33\linewidth]{figures/cal_paper_gsm_beam_weighted.pdf}
-\caption{\textcolor{red}{caption!}}
+\caption{Images produced before (left) and after (middle) calibrating LWA data.
+These images were produces with 29 kHz bandwidth and 51.2 ms integration.
+The calibrated image shows significant reduction in sidelobe rumble throughout, while retaining the prominent Cyg~A and Cas~A sources.
+The galactic plane is also substantially more evident after calibration.
+For reference, the GSM is shown in the same coordinates and weighted by two factors of the LWA primary beam in the right panel.
+}
 \label{fig:data_images}
@@
 
-\textcolor{red}{This part of the argument needs serious work!} In order to find the Cram\'er-Rao lower bound on the variance of the complex parameter, $g_i$, we consider the term of the Fischer matrix where the first derivative is taken with respect to $g_i$, and the second with $g_i^*$.
+In order to find the Cram\'er-Rao lower bound on the variance of the complex parameter, $g_i$, we consider the term of the Fischer matrix where the first derivative is taken with respect to $g_i$, and the second with $g_i^*$.
 The result is

### 2016-02-01 cal_demo/ms.tex

@@
 % Keep the title short and informative.
-\title[E-field Parallel Imaging Calibration]{An Efficient E-field Parallel Imaging Calibration Algorithm for Next-Generation Radio Telescopes}
+\title[EPICal]{The E-field Parallel Imaging Calibration Algorithm for Next-Generation Radio Telescopes}
 
@@
 \begin{abstract}
-Abstract here (250 words)
+We present the E-field Parallel Imaging Calibration (EPICal) algorithm, which addresses the need for a real-time calibration method for direct imaging radio astronomy correlators.
+Direct imaging involves a spatial fast Fourier transform of antenna voltages, alleviating the harsh $\mathcal{O}(\Nant^2)$ computational scaling to a more gentle $\mathcal{O}(\Nant \log_2 \Nant)$, which can save orders of magnitude in computation cost for next generation arrays consisting of hundreds to thousands of antennas.
+However, because signals are mixed in the correlator, gain correction must be applied on the front end.
+We develop the EPICal algorithm to form gain solutions in real time without ever forming visibilities.
+This method scales as the number of antennas, and produces results equivalent to those from visibilities.
+Through simulations and application to Long Wavelength Array data we show this algorithm is a feasible solution for next generation instruments.
 \end{abstract}
@@
 In order to satisfy the survey speeds required for precision cosmology as well as searches for fast radio transients, radio astronomy is undergoing a paradigm shift toward interferometers consisting of hundreds to thousands of small, widefield antennas.
-Many arrays with this design are already built or under construction including the Hydrogen Epoch of Reionization Array\footnote{http://reionization.org} (HERA), the Murchison Widefield Array (MWA;\citealt{tin13,bow13}), the Precision Array for Probing the Epoch of Reionization (PAPER; \citealt{par10}), the LOw Frequency ARray (LOFAR; \citealt{van13}), the Canadian Hydrogen Intensity Mapping Experiment (CHIME,\citealt{ban14}), the Long Wavelength Array (LWA, \citealt{ell13}), and the low frequency Square Kilometer Array (SKA1-Low \citealt{mel13}).
+Many arrays with this design are already built or under construction including the Hydrogen Epoch of Reionization Array\footnote{http://reionization.org} (HERA), the Murchison Widefield Array (MWA; \citealt{tin13,bow13}), the Precision Array for Probing the Epoch of Reionization (PAPER; \citealt{par10}), the LOw Frequency ARray (LOFAR; \citealt{van13}), the Canadian Hydrogen Intensity Mapping Experiment (CHIME,\citealt{ban14}), the Long Wavelength Array (LWA, \citealt{ell13}), and the low frequency Square Kilometer Array (SKA1-Low \citealt{mel13}).
 
@@
 We will assume the gains have no time dependence within the timescale of finding our solutions.
-\textcolor{red}{cite someone about gain stability}.
 Furthermore, we will treat each frequency channel independently, and drop the $f$ to simplify notation.
@@
 In order to compare to solutions from visibility-based calibration, we first derive the theoretical lower bound of the error in these solutions, given a set of measured visibilities, $\mathbf{V}$.
-We will do this through the Fischer information formalism and employ the Cram\'er-Rao lower bound
+Motivated by the performance analysis seen in \cite{kaz13}, we will do this through the Fischer information formalism and employ the Cram\'er-Rao lower bound.
 
@@
 
-Finally we relate our result to the theoretical best uncertainty we can place on our unknown gain parameter using the Cram\'er-Rao lower bound.
+Finally we relate our result to the theoretical best uncertainty we can place on our unknown gain parameter using the Cram\'er-Rao lower bound \cite{kay93}.
 \begin{equation}\label{eq:cramer_rao}
@@
 
-We next run a set of EPICal simulations to compare the noisy results to the theoretical lower bound given a set of equivalent visibilities.
+We next run a set of EPICal simulations to compare the noisy results to the theoretical lower bound if we had formed an analogous set of visibilities.
 Each simulation is run with a simple sky of a single point source with flux 1 Jy.
@@
 
-We vary two parameters in our simulations: the amount of noise added to our measurements, and the length of integration in each calibration loop.
+We vary two parameters in our simulations: the variance of the complex random noise added to our electric field measurements ($\sigma_0$), and the length of integration in each calibration loop, $t_{\mathrm{cal}}$.
+The equivalent visibility variance in equation~\ref{eq:cramer_rao} is given by
+\begin{equation}
+\sigma_{ai}^2 = \frac{\sigma_0^2}{B t_{\mathrm{eff}}},
+\end{equation}
+where $B$ is the bandwidth of the simulation, and $t_{\mathrm{eff}}$ is the effective integration time based on the length of our calibration loop and the damping factor.
+\begin{equation}
+t_{\mathrm{eff}} = t_{\mathrm{cal}} \times \frac{1-\damp^2}{(1-\damp)^2}
+\end{equation}
+
+We demonstrate the trend of the gain errors with respect to these parameters in figure~\ref{fig:noise}.
+For each $\sigma_0$ and $t_{\mathrm{cal}}$ we run the simulation for seven iterations of the calibration loop to allow the noise to fully corrupt our estimates.
+We then compute the variance of the resulting gains across the antennas.
+Each color line represents a different amount of input noise, with the dashed line showing the measured variance and the solid line showing the theoretical lower limit from equation~\ref{eq:cramer_rao}.
+The agreement between the dashed and solid lines demonstrates that not only do we reach an uncertainty consistent with visibility-based solutions, but that we actually reach the theoretical lower limit based on the amount of information contained in an equivalent set of visibility measurements.
+
+\begin{figure}
+\begin{center}
+\includegraphics[draft]{figures/noise.pdf}
+\caption{\textcolor{red}{caption!}
+}
+\label{fig:noise}
+\end{center}
+\end{figure}
 
 \section{Discussion}\label{sec:discussion}
+Through simulations and application to real data, we have shown that the EPICal algorithm is a viable solution for calibrating direct imaging arrays in real time.
+The computation necessary only scales with the number of antenna elements, making it a sub-dominant cost factor when designing the correlator.
+This strategy will enable fast read-out and arrays with many thousands of antennas, which will be necessary for future radio transient and cosmology experiments respectively.
+
+There remain many avenues of future work before EPICal can be deployed on an instrument.
+Here we name a few potential considerations for further study.
+These topics can be thought of as extensions to the white ``estimate gains" box in figure~\ref{fig:schematic}, which can be performed at a much lower cadence than the actual correlations.
+
+\textbf{Fitting gain models.} With some knowledge of the instrumental bandpass, the signal-to-noise of the gain solutions can be greatly improved by fitting a model to the per-frequency solutions derived here.
+We used this method at a rudimentary level in our demonstration to LWA data by assuming the gains were constant over a narrow bandwidth.
+One could easily improve on this by extending the bandwidth and fitting for a low order polynomial in phase and amplitude.
+Additionally, with knowledge of the poly-phase filters applied to the data stream, we could include channels closer to the edge of the band.
+
+\textbf{Improving sky model.} In this work we used a perfect sky model (for simulation), or a very simple sky model (for LWA data).
+In principle the images produced by the EPIC correlator can be used to improve the sky model used in calibration -- similar to a major loop in self-calibration.
+This can be especially useful for compact, widefield arrays which require a model of both compact and diffuse sources over a large patch of sky.
+In addition to requiring the sky model, the calibration also requires this model to be attenuated by the primary beam, which can be difficult to measure at the precision necessary \citep[e.g.][]{neb15,vir14,thy15b}.
+However, the direct imaging correlator provides exactly the model necessary in real time, and can be iterated over to improve these images and the gain solutions.
+
+\textbf{Dynamic parameters.} In our controlled experiments we fine tuned a number of parameters based on our testing (e.g.
+damping factor, integration time, frequency averaging).
+A deployed system will require robust determination of these parameters to operate continuously.
+The specifics will be heavily dependent on the stability of the instrument, the frequency of observation, and the sky.
+For example, a stable instrument may be able to use short integrations to determine a rough estimate of the gains before switching to much longer integration (on order seconds to minutes) to highly increase signal to noise.
+At low frequencies or high sky resolution, the dynamics of the ionosphere are important, and will likely drive the limit of time integration allowed.
+
+\textbf{Direction dependent gains.} Even below the timescale of a dynamic ionosphere, a widefield instrument may require direction dependent gains to account for a varying thickness ionospheric sheet.
+The holographic mapping framework used by the MOFF algorithm, and in turn the EPIC and EPICal software can naturally account for direction dependence through and updated model of the antenna voltage pattern.
+This can be achieved by first forming complex gain estimates toward multiple directions on the sky by selecting several pixels and independently following the EPICal algorithm for each.
+These solutions can then be used to fit a beam model on the sky (analogous to the gain model fitting mentioned above), and updating the antenna voltage pattern.
+This updated pattern would then feed into the gridding step of the correlator, allowing the imager to convolve the signals with the effective beam pattern in the ground plane.
+Because this method requires multiple pixel correlations, the scaling now becomes $\mathcal{O}(\Nant N_{\mathrm{pix}})$, which will typically be much lower than the $\mathcal{O}(\Ng \log_2 \Ng)$ of the correlator itself.
+
+The work here will serve as a foundation for further development.
+We have shown that the EPICal algorithm produces reliable calibration solutions, and have identified several aspects to increase the scope.
+With next generation instruments in the planning and development stages, EPICal is poised to greatly contribute to their success.
 
@@
 This work has been supported by the National Science Foundation through award AST-1206552. 
+We thank Danny Jacobs for his valuable inputs, and Greg Taylor for providing us with LWA data. 
+Construction of the LWA has been supported by the Office of Naval Research under 
+Contract N00014-07-C-0147. Support for operations and continuing development of 
+the LWA1 is provided by the National Science Foundation under grant AST-1139974 
+of the University Radio Observatory program.
 

### 2016-02-16 cal_demo/ms.tex
rewriting noise section of cal paper. Adding noise figure

@@
 
-\section{Noise analysis}\label{sec:noise}
+\section{Noise trends}\label{sec:noise}
+Here we study the properties of our calibration solutions with respect to noise in the system, and compare to solutions found from visibilities.
+For this comparison we run a suite of simulations varying the receiver noise and integration times, while forming $\Cna$ correlations for EPICal gain solutions, and simultaneously forming visibilities from the same E-field streams to find visibility-based gain solutions.
 
-Here we study the properties of our calibration solutions with respect to noise in the system.
+We use a simulated sky consisting of a 5 Jy calibrator source, and 10 other random sources with apparent flux densities 0.2 to 0.5 Jy (total sky power of about 8 Jy).
+While we generate these other sources randomly, we keep them fixed for each run.
+Again, we use the MWA core layout (51 antennas), and a 4.4~m square tophat antenna voltage pattern.
+We simulate 64 frequency channels.
+Because our calibration loop is frequency independent, each channel can be treated as a separate trial of the simulation, and is used to better estimate the statistics.
+
+To simulate receiver noise, we add a gaussian distributed complex random number to each antenna electric field measurement at each time stamp according to equation~\ref{eq:apply_gain}.
+The level of the receiver noise, $\sigma_r = \left<\left|\widetilde{n}_a(f,t)\right|^2\right>$, is varied in different simulation runs.
+
+We initialize our gain estimates with the true values (unity), and allow the estimate to be corrupted by the noise through one iteration of the calibration loop.
+In order to allow the noise to fully corrupt the gains we use a small damping factor, $\damp=0.05$.
+Using a factor of zero leads to non-convergent results.
+We therefore use the same damping factor for the visibility-based solutions for a fair comparison.
+We then observe the error in the gain estimates introduced by the noise.
+\begin{equation}\label{eq:gain_error}
+\sigma_g = \left[\frac{1}{N_f \Nant} \sum_f \sum_a \frac{\left|g'_a(f)-g^T_a(f)\right|^2}{\left|g^T_a(f)\right|^2}\right]^{1/2}
+\end{equation}
+
+We simultaneously form simulated visibilities by correlating all pairs of antenna E-field measurements (including the receiver noise).
+At the end of the calibration loop we find the visibility-based gain estimates by minimizing
+\begin{equation}\label{eq:vis_cal}
+\chi^2 = \sum_a\sum_{b\ne a} \left|\V_{ab}-g_a g_b^* \V^T_{ab}\right|^2.
+\end{equation}
+The error of the visibility-based gains is estimated in the same way as in equation~\ref{eq:gain_error}.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figures/cal_paper_noise_med_sky.pdf}
+\caption{Gain estimate errors as a function of integration time and receiver noise (line color).
+The solid lines represent the error on the EPICal derived gain estimates, while the dashed lines represent the error on the visibility based estimates.
+On average the EPICal error is 1.5\% higher than visibilities, and at worst it was 3.5\% higher.
+}
+\label{fig:noise}
+\end{center}
+\end{figure}
+
+%\section{Noise analysis}\label{sec:noise_old}
+%\textcolor{red}{This section will most likely disappear in favor of the Noise Trends section.}
+%
+%Here we study the properties of our calibration solutions with respect to noise in the system.
 In order to compare to solutions from visibility-based calibration, we first derive the theoretical lower bound of the error in these solutions, given a set of measured visibilities, $\mathbf{V}$.
 Motivated by the performance analysis seen in \cite{kaz13}, we will do this through the Fischer information formalism and employ the Cram\'er-Rao lower bound.
-
-Assume we did form visibilities with some integration time.
+%
+%Assume we did form visibilities with some integration time.
 Assuming the noise on each visibility, $\sigma_{ab}$, is independent, we can write the likelihood function of measuring $V_{ab}$ given the true value, the gains, and the noise.
-\begin{equation}
-\mathcal{L}(V_{ab};\mathbf{g}) = \frac{1}{2\pi \sigma_{ab}^2}\exp\left[-\frac{\left|V_{ab} - g_a g_b^* V_{ab}^T\right|^2}{2\sigma_{ab}^2}\right]
-\end{equation}
-Then the likelihood of the set of all visibilities is the product of all individual likelihoods.
-\begin{equation}
-\mathcal{L}(\mathbf{V};\mathbf{g}) = \prod_a \prod_{b > a} \mathcal{L}(V_{ab};\mathbf{g})
-\end{equation}
-The Fischer information matrix for the set of gain parameters is
-\begin{equation}
-\mathbf{F}^g_{ij} = \left<\left. \frac{\partial \ln \mathcal{L}(\mathbf{V};\mathbf{g})}{\partial g_i} \frac{\partial \ln \mathcal{L}(\mathbf{V};\mathbf{g})}{\partial g_j}\right|_\mathbf{g}\right>,
-\end{equation}
-where the expectation is taken over all possible visibility measurements.
-We next evaluate the derivate of the log-likelihood.
-\begin{equation}
-\frac{\partial \ln \mathcal{L}(\mathbf{V};\mathbf{g})}{\partial g_i} = 
-\sum_{a \ne i} \frac{g_a^* V_{ai}^{T*} (V_{ai} - g_a g_i^* V_{ai}^T)}{2 \sigma_{ai}^2}
-\end{equation}
-
-In order to find the Cram\'er-Rao lower bound on the variance of the complex parameter, $g_i$, we consider the term of the Fischer matrix where the first derivative is taken with respect to $g_i$, and the second with $g_i^*$.
+%\begin{equation}
+%\mathcal{L}(V_{ab};\mathbf{g}) = \frac{1}{2\pi \sigma_{ab}^2}\exp\left[-\frac{\left|V_{ab} - g_a g_b^* V_{ab}^T\right|^2}{2\sigma_{ab}^2}\right]
+%\end{equation}
+%Then the likelihood of the set of all visibilities is the product of all individual likelihoods.
+%\begin{equation}
+%\mathcal{L}(\mathbf{V};\mathbf{g}) = \prod_a \prod_{b > a} \mathcal{L}(V_{ab};\mathbf{g})
+%\end{equation}
+%The Fischer information matrix for the set of gain parameters is
+%\begin{equation}
+%\mathbf{F}^g_{ij} = \left<\left. \frac{\partial \ln \mathcal{L}(\mathbf{V};\mathbf{g})}{\partial g_i} \frac{\partial \ln \mathcal{L}(\mathbf{V};\mathbf{g})}{\partial g_j}\right|_\mathbf{g}\right>,
+%\end{equation}
+%where the expectation is taken over all possible visibility measurements.
+%We next evaluate the derivate of the log-likelihood.
+%\begin{equation}
+%\frac{\partial \ln \mathcal{L}(\mathbf{V};\mathbf{g})}{\partial g_i} = 
+%\sum_{a \ne i} \frac{g_a^* V_{ai}^{T*} (V_{ai} - g_a g_i^* V_{ai}^T)}{2 \sigma_{ai}^2}
+%\end{equation}
+%
+%In order to find the Cram\'er-Rao lower bound on the variance of the complex parameter, $g_i$, we consider the term of the Fischer matrix where the first derivative is taken with respect to $g_i$, and the second with $g_i^*$.
 The result is
-\begin{align}
-\mathbf{F}^g_{ii} = & \left<\left[ \sum_{a \ne i} \frac{g_a^* V_{ai}^{T*} (V_{ai} - g_a g_i^* V_{ai}^T)}{2 \sigma_{ai}^2} \right] \right. \times \nonumber \\
-& \left. \left[ \sum_{b \ne i} \frac{g_b V_{bi}^{T} (V_{bi}^* - g_b^* g_i V_{bi}^{T*})}{2 \sigma_{bi}^2} \right] \right> \nonumber \\
-= &\sum_{a\ne i} \sum_{b \ne i} \frac{g_a^* g_b V_{ai}^{T*} V_{bi}^T}{4 \sigma_{ai}^2 \sigma_{bi}^2} \left< V_{ai}V_{bi}^* - g_i g_b^* V_{ai}V_{bi}^{T*}\nonumber \right.\\
-&\left. - g_a g_i^* V_{ai}^T V_{bi}^* + g_a g_b^* |g_i|^2 V_{ai}^T V_{bi}^{T*}\right>
-\label{eq:Fischer_before_expect}
-\end{align}
-
-The expected values are easy to evaluate.
+%\begin{align}
+%\mathbf{F}^g_{ii} = & \left<\left[ \sum_{a \ne i} \frac{g_a^* V_{ai}^{T*} (V_{ai} - g_a g_i^* V_{ai}^T)}{2 \sigma_{ai}^2} \right] \right. \times \nonumber \\
+%& \left. \left[ \sum_{b \ne i} \frac{g_b V_{bi}^{T} (V_{bi}^* - g_b^* g_i V_{bi}^{T*})}{2 \sigma_{bi}^2} \right] \right> \nonumber \\
+%= &\sum_{a\ne i} \sum_{b \ne i} \frac{g_a^* g_b V_{ai}^{T*} V_{bi}^T}{4 \sigma_{ai}^2 \sigma_{bi}^2} \left< V_{ai}V_{bi}^* - g_i g_b^* V_{ai}V_{bi}^{T*}\nonumber \right.\\
+%&\left. - g_a g_i^* V_{ai}^T V_{bi}^* + g_a g_b^* |g_i|^2 V_{ai}^T V_{bi}^{T*}\right>
+%\label{eq:Fischer_before_expect}
+%\end{align}
+%
+%The expected values are easy to evaluate.
 Each visibility will average to the ``true" value times the respective gains.
 The term with two visibilities will include a noise bias on autocorrelation terms.
-\begin{equation}
-\left<V_{ai}V_{bi}^*\right> = |g_i|^2 g_a g_b^* V_{ai}^T V_{bi}^{T*} + \sigma_{ai}^2 \delta_{ab}
-\end{equation}
-Here $\delta_{ab}$ is the Kronecker delta selecting the term where $a=b$ and the noise correlates. Plugging in the expectation values, equation \ref{eq:Fischer_before_expect} simplifies greatly to
-\begin{equation}
-\mathbf{F}^g_{ii} = \sum_{a \ne i} \frac{\left|g_a V_{ia}^T \right|^2}{4\sigma_{ai}^2}
-\end{equation}
-
-Finally we relate our result to the theoretical best uncertainty we can place on our unknown gain parameter using the Cram\'er-Rao lower bound \cite{kay93}.
-\begin{equation}\label{eq:cramer_rao}
-\sigma_{g_i}^2 \ge \left[ \mathbf{F}^g_{ii} \right]^{-1} = \left[ \sum_{a \ne i} \frac{\left|g_a V_{ia}^T \right|^2}{4\sigma_{ai}^2} \right]^{-1}
-\end{equation}
-
-We next run a set of EPICal simulations to compare the noisy results to the theoretical lower bound if we had formed an analogous set of visibilities.
+%\begin{equation}
+%\left<V_{ai}V_{bi}^*\right> = |g_i|^2 g_a g_b^* V_{ai}^T V_{bi}^{T*} + \sigma_{ai}^2 \delta_{ab}
+%\end{equation}
+%Here $\delta_{ab}$ is the Kronecker delta selecting the term where $a=b$ and the noise correlates. Plugging in the expectation values, equation \ref{eq:Fischer_before_expect} simplifies greatly to
+%\begin{equation}
+%\mathbf{F}^g_{ii} = \sum_{a \ne i} \frac{\left|g_a V_{ia}^T \right|^2}{4\sigma_{ai}^2}
+%\end{equation}
+%
+%Finally we relate our result to the theoretical best uncertainty we can place on our unknown gain parameter using the Cram\'er-Rao lower bound \cite{kay93}.
+%\begin{equation}\label{eq:cramer_rao}
+%\sigma_{g_i}^2 \ge \left[ \mathbf{F}^g_{ii} \right]^{-1} = \left[ \sum_{a \ne i} \frac{\left|g_a V_{ia}^T \right|^2}{4\sigma_{ai}^2} \right]^{-1}
+%\end{equation}
+%
+%We next run a set of EPICal simulations to compare the noisy results to the theoretical lower bound if we had formed an analogous set of visibilities.
 Each simulation is run with a simple sky of a single point source with flux 1 Jy.
@@
 This emulates a system where the initially volatile gain correction has been suppressed and only noise remains to corrupt our estimation, allowing us to run many simulations on a shorter timescale.
-
-We vary two parameters in our simulations: the variance of the complex random noise added to our electric field measurements ($\sigma_0$), and the length of integration in each calibration loop, $t_{\mathrm{cal}}$.
+%
+%We vary two parameters in our simulations: the variance of the complex random noise added to our electric field measurements ($\sigma_0$), and the length of integration in each calibration loop, $t_{\mathrm{cal}}$.
 The equivalent visibility variance in equation~\ref{eq:cramer_rao} is given by
-\begin{equation}
-\sigma_{ai}^2 = \frac{\sigma_0^2}{B t_{\mathrm{eff}}},
-\end{equation}
-where $B$ is the bandwidth of the simulation, and $t_{\mathrm{eff}}$ is the effective integration time based on the length of our calibration loop and the damping factor.
-\begin{equation}
-t_{\mathrm{eff}} = t_{\mathrm{cal}} \times \frac{1-\damp^2}{(1-\damp)^2}
-\end{equation}
-
-We demonstrate the trend of the gain errors with respect to these parameters in figure~\ref{fig:noise}.
+%\begin{equation}
+%\sigma_{ai}^2 = \frac{\sigma_0^2}{B t_{\mathrm{eff}}},
+%\end{equation}
+%where $B$ is the bandwidth of the simulation, and $t_{\mathrm{eff}}$ is the effective integration time based on the length of our calibration loop and the damping factor.
+%\begin{equation}
+%t_{\mathrm{eff}} = t_{\mathrm{cal}} \times \frac{1+\damp}{1-\damp}
+%\end{equation}
+%
+%We demonstrate the trend of the gain errors with respect to these parameters in figure~\ref{fig:noise}.
 For each $\sigma_0$ and $t_{\mathrm{cal}}$ we run the simulation for seven iterations of the calibration loop to allow the noise to fully corrupt our estimates.
@@
 The agreement between the dashed and solid lines demonstrates that not only do we reach an uncertainty consistent with visibility-based solutions, but that we actually reach the theoretical lower limit based on the amount of information contained in an equivalent set of visibility measurements.
-
-\begin{figure}
-\begin{center}
-\includegraphics[draft]{figures/noise.pdf}
-\caption{\textcolor{red}{caption!}
-}
-\label{fig:noise}
-\end{center}
-\end{figure}
+%
+%
 

### 2016-02-16 imaging_demo/ms.tex
Merge mnras branch on to arxiv branch
From b8e3bc0 (arxiv) + 2add9d8 (mnras):
======================================

Merge mnras branch on to arxiv branch

--- 2015-11-08 imaging_demo/ms.tex
@@
 % Basic setup. Most papers should leave these options alone.
-% \documentclass[a4paper,fleqn,usenatbib]{../mnras}
-\documentclass[a4paper,fleqn,usenatbib]{mnras}
+\documentclass[a4paper,fleqn,usenatbib]{../mnras}
 
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figure1}
-  % \includegraphics[width=\columnwidth]{MOFF_flowchart}
+  \includegraphics[width=\columnwidth]{MOFF_flowchart}
   % \includegraphics[width=\columnwidth]{MOFF_flowchart.eps}
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figure2}
-  % \includegraphics[width=\columnwidth]{FX_flowchart}
+  \includegraphics[width=\columnwidth]{FX_flowchart}
   % \includegraphics[width=\columnwidth]{FX_flowchart.eps}
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figure3}
-  % \includegraphics[width=\columnwidth]{a2g_mapping}
+  \includegraphics[width=\columnwidth]{a2g_mapping}
   % \includegraphics[width=\columnwidth]{a2g_mapping.eps}
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figure4}
-  % \includegraphics[width=\columnwidth]{F-engine}
+  \includegraphics[width=\columnwidth]{F-engine}
   % \includegraphics[width=\columnwidth]{F-engine.eps}
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figure5}
-  % \includegraphics[width=\columnwidth]{autocorr_uvwts_pbeam}
+  \includegraphics[width=\columnwidth]{autocorr_uvwts_pbeam}
   % \includegraphics[width=\columnwidth]{autocorr_uvwts_pbeam.png}
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figure6}
-  % \includegraphics[width=\columnwidth]{MOFF_FX_comparison_10_random_source_positions_4_iterations_test_aperture_zoomed}
+  \includegraphics[width=\columnwidth]{MOFF_FX_comparison_10_random_source_positions_4_iterations_test_aperture_zoomed}
   % \includegraphics[width=\columnwidth]{MOFF_FX_comparison_10_random_source_positions_4_iterations_test_aperture_zoomed.eps}
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figure7}
-  % \includegraphics[width=\columnwidth]{MOFF_FX_comparison_uvwts_test_aperture_zero_spacing_removed}
+  \includegraphics[width=\columnwidth]{MOFF_FX_comparison_uvwts_test_aperture_zero_spacing_removed}
   % \includegraphics[width=\columnwidth]{MOFF_FX_comparison_uvwts_test_aperture_zero_spacing_removed.eps}
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figure8}
-  % \includegraphics[width=\columnwidth]{diff_psf_MOFF-FX_test_aperture}
+  \includegraphics[width=\columnwidth]{diff_psf_MOFF-FX_test_aperture}
   % \includegraphics[width=\columnwidth]{diff_psf_MOFF-FX_test_aperture.eps}
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figure9}
-  % \includegraphics[width=\columnwidth]{LWA_MOFF_bandavg_image_4_iterations_analytic_aperture}
+  \includegraphics[width=\columnwidth]{LWA_MOFF_bandavg_image_4_iterations_analytic_aperture}
   % \includegraphics[width=\columnwidth]{LWA_MOFF_bandavg_image_4_iterations_analytic_aperture.eps}
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figure10}
-  % \includegraphics[width=\columnwidth]{MOFF_FX_computations_fov_gridding_annotated}
+  \includegraphics[width=\columnwidth]{MOFF_FX_computations_fov_gridding_annotated}
   % \includegraphics[width=\columnwidth]{MOFF_FX_computations_fov_gridding_annotated.eps}
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figure11}
-  % \includegraphics[width=\columnwidth]{MOFF_FX_crossover_baseline_n-antennas_rho_fov_gridding_legended}
+  \includegraphics[width=\columnwidth]{MOFF_FX_crossover_baseline_n-antennas_rho_fov_gridding_legended}
   % \includegraphics[width=\columnwidth]{MOFF_FX_crossover_baseline_n-antennas_rho_fov_gridding_legended.eps}
@@
 \bibliographystyle{mnras}
-\bibliography{epic} 
-% \bibliography{../epic} 
+\bibliography{../epic} 
 % \input{ms.bbl}
@@
 \begin{figure*}
-  \includegraphics[width=\linewidth]{figureA1}
-  % \includegraphics[width=\linewidth]{EPIC-modules}
+  \includegraphics[width=\linewidth]{EPIC-modules}
   % \includegraphics[width=\linewidth]{EPIC-modules.eps}

### 2016-02-17 cal_demo/ms.tex
updating cal paper to address judds comments

@@
 % Keep the title short and informative.
-\title[EPICal]{The E-field Parallel Imaging Calibration Algorithm for Next-Generation Radio Telescopes}
+\title[EPICal]{An efficient feedback calibration algorithm for direct imaging radio telescopes
+%The E-field Parallel Imaging Calibration Algorithm for Next-Generation Radio Telescopes
+}
 
@@
 We will assume we have formed an estimate of the gains after $n$ iterations of a calibration loop, and derive an updated $n+1$ estimate.
-Typically calibration amounts to dividing the measured fields by the gain (in the case of sky noise dominance), or multiplying by the complex conjugate of the gain (receiver noise dominance).
+How the calibration gain is applied can depend on the dominant source of the noise in the measurement.
+If the noise is dominated by sky fluctuations, the noise level is affected by the antenna gains so dividing by the complex gains will flatten the noise across antennas.
+However, if the noise is dominated in the receiver (after the complex gain factor of equation~\ref{eq:apply_gain}), a lower amplitude gain results in lower signal to noise for that antenna.
+Therefore, one should multiply by the conjugate of the gain to down weight appropriately while removing the phase acquired.
 We will abstain from this choice for now, and instead use a multiplicative factor $h^{(n)}_a$ to represent the application of our $n^{\mathrm{th}}$-loop estimate of the gain for antenna $a$:
@@
 where in the second step we group time-dependent terms, and in the third we define the true visibilities as the correlation between true antenna electric field measurements.
-It is easy to see from here that the net effect of including the receiver noise term, $\widetilde{n}_a(f,t)$, will result in added noise on the true visibilities.
-In principle this can include any noise correlations between antennas.
-The implementation included in the EPIC software package restricts the noise term to include only the auto-correlation noise bias, and assumes baseline dependent noise is zero mean.
+It is easy to see from here that the net effect of including the receiver noise term, $\widetilde{n}_a(f,t)$, will result in added noise on the true visibilities, including a bias on the auto-correlation terms.
+In principle a bias can result on any visibility from any noise correlations between antennas, but the implementation included in the EPIC software package restricts the noise term to include only the auto-correlation noise bias, and assumes baseline dependent noise is zero mean.
 
@@
 \end{equation}
-We found that while equation~\ref{eq:cal_solution} does indeed converge on good solutions, the process is faster by tuning the damping factor.
-We note the small difference between $g^{(n+1)}_a$ and $g'^{(n+1)}_a$, where the prime indicates our current best estimate of the true gain, while no prime is actually used in the iterative calibration loop.
-Once the loop converges the difference is a longer effective integration for the non-primed version (lower thermal noise).
+We found that while equation~\ref{eq:cal_solution} does indeed converge on good solutions, the process is made faster by tuning the damping factor.
+While $g'^{(n+1)}_a$ is the best estimate of the gains after the $n^{\textrm{th}}$ iteration, the damped version, $g^{(n+1)}_a$, is actually used in the iterative calibration loop.
+Once the loop converges the difference is a longer effective integration for the damped version (lower thermal noise).
 
@@
 We made several choices while studying the behavior of the LWA data to improve our calibration.
-We inspected the amplitudes of the voltages to be roughly consistent with average gain amplitudes of 0.25, so we initialized our gains at this level to allow the calibration to converge quickly.
+Through our previous imaging work, we noted that the flux scale of uncalibrated images was consistent with average gain amplitudes of 0.25.
+To allow the calibration to converge quickly we initialized our gain estimates at this level.
 We also found a boost in signal to noise is achieved easily by averaging frequency channels and assuming the gains are constant within a sub-band.
@@
 The actual location of the ``trunk" (slightly negative) is simply determined by the reference antenna chosen to have identically zero phase, but happens to be slightly more positive than the bulk of antennas.
-We also note that some phases appear to exceed $\pm \pi$ radians, this is due to the unwrapping done for plotting clarity.
+For plotting clarity, we unwrapped phases resulting in phases that appear to exceed $\pm \pi$.
 
@@
 However, this is not surprising due to the non-uniformity in the cables from the antennas to the receivers.
-Again, the amplitudes are noisy by relatively flat.
+Again, the amplitudes are noisy but relatively flat.
 
@@
 
-We use a simulated sky consisting of a 5 Jy calibrator source, and 10 other random sources with apparent flux densities 0.2 to 0.5 Jy (total sky power of about 8 Jy).
-While we generate these other sources randomly, we keep them fixed for each run.
-Again, we use the MWA core layout (51 antennas), and a 4.4~m square tophat antenna voltage pattern.
-We simulate 64 frequency channels.
+We use a simulated sky consisting of a 5 Jy calibrator source, and 10 other random sources with apparent flux densities 0.2 to 0.5 Jy (total sky power $\approx$~8 Jy).
+We generated these sources randomly, but kept them fixed for each run.
+As in section~\ref{sec:sim}, we use the MWA core layout (51 antennas), and a 4.4~m square tophat antenna voltage pattern.
+We simulate 64 40~kHz frequency channels.
 Because our calibration loop is frequency independent, each channel can be treated as a separate trial of the simulation, and is used to better estimate the statistics.
@@
 \end{figure}
+
+\textcolor{red}{Complex sky (23.2 Jy total sky power) -- EPICal 20.2\% higher error. Difference was floating point precision for the simple sky (single point source)}
 

### 2016-02-17 cal_demo/ms.tex
More text in the cal noise analysis

@@
 We develop the EPICal algorithm to form gain solutions in real time without ever forming visibilities.
-This method scales as the number of antennas, and produces results equivalent to those from visibilities.
+This method scales as the number of antennas, and produces results comparable to those from visibilities.
 Through simulations and application to Long Wavelength Array data we show this algorithm is a feasible solution for next generation instruments.
@@
 We initialize our gain estimates with the true values (unity), and allow the estimate to be corrupted by the noise through one iteration of the calibration loop.
-In order to allow the noise to fully corrupt the gains we use a small damping factor, $\damp=0.05$.
-Using a factor of zero leads to non-convergent results.
-We therefore use the same damping factor for the visibility-based solutions for a fair comparison.
+Ideally we would use a damping factor of zero to allow for the noise to fully corrupt the gains.
+However, we found that a zero damping factor resulted in non-convergent results, so we adopted a factor, $\damp=0.05$.
+For a fair comparison, we use the same factor for the visibility-based solutions.
 We then observe the error in the gain estimates introduced by the noise.
@@
 
+The results of our simulations are shown in figure~\ref{fig:noise}.
+The gain errors from our EPICal method is shown with solid lines, each color representing different levels of receiver noise.
+Each data point represents a separate simulation run (independent realization of the noise).
+The general trends are consistent with expectations -- decreasing as the square root of integration time, and increasing with receiver noise.
+The lowest receiver noise level (1.0 Jy) is significantly below the sky noise contribution (about 8 Jy), so the black line is not as low, compared to the red, as one would expect from receiver noise consideration alone.
+
 \begin{figure}
@@
 
-\textcolor{red}{Complex sky (23.2 Jy total sky power) -- EPICal 20.2\% higher error. Difference was floating point precision for the simple sky (single point source)}
+The dashed lines in figure~\ref{fig:noise} represent the errors from the visibility-based calibration.
+Each data point corresponds to the same simulation run as the corresponding EPICal point (same realization of noise).
+Generally we see the visibility-based calibration achieves slightly lower noise solutions.
+On average the EPICal errors are 1.5\% higher, and at the worst is 3.5\% higher.
+
+We repeated this process for two extremes -- a simple sky model of a single point source, and a more complicated sky with 50 sources.
+In the simple sky case, we found the EPICal and visibility gain estimates to be identical to floating point precision, as expected from our discussion in section~\ref{sec:math}.
+In the more complicated sky case (about 23 Jy total sky power), we saw the EPICal gains exhibit errors 20\% higher than the visibility derived gains.
+The trend of higher EPICal errors with more complicated skies is due to the single correlated pixel not containing the full signal to noise of the data.
+This reaffirms our intuition that using a pixel with a bright source is ideal, but will still not match visibilities for complicated skies.
+We suggest additional strategies for improving this signal to noise in the following section.
 

### 2016-02-17 cal_demo/ms.tex
Further tweaking cal paper before sending to coauthors

@@
 % Enter the current year, for the copyright statements etc.
-\pubyear{2015}
+\pubyear{2016}
 
@@
 Next we move toward the feedback calibration outlined in \citealt{mor11}.
-The calibration loop described there assumed a simple sky of a single point source, but we aim for a more generalized solution for arbitrarily complex sky models.
-As such the quantity we wish to inspect is a correlation of the uncalibrated antenna signals input to the correlator with the holographic electric field image output from the correlator.
-We define the quantity,
+There it was suggested to form a correlation of the uncalibrated antenna measurements with an image pixel from the output of the correlator.
+This is a statistically stationary quantity, which can be related to the sum of visibilities involving the antenna used in the correlation -- exactly the sum needed to calibrate a simple sky of a single point source.
+However, we aim for a more generalized solution for arbitrarily complex sky models.
+We therefore study the full expression for the antenna-pixel correlation,
 \begin{equation}\label{eq:Cna}
@@
 The middle panel shows the image produced after calibration with identical integration time and bandwidth.
-The sidelobes throughout the image are significantly suppressed, and the galactic plane is much more evident.
+The sidelobes throughout the image are significantly suppressed, and the galactic plane is much more evident, despite only modeling Cyg~A and Cas~A.
 We also note that the feature just to the right of Cyg~A is dimmer in the calibrated image, better matching the expected flux from the global sky model (GSM, \citealt{deo08}).
@@
 For reference, the GSM is shown in the same coordinates and weighted by two factors of the LWA primary beam in the right panel.
+While much of the GSM is visible in the calibrated image, we note that the model used to calibrate was only the two bright sources, Cyg~A and Cas~A.
 }
@@
 Here we study the properties of our calibration solutions with respect to noise in the system, and compare to solutions found from visibilities.
-For this comparison we run a suite of simulations varying the receiver noise and integration times, while forming $\Cna$ correlations for EPICal gain solutions, and simultaneously forming visibilities from the same E-field streams to find visibility-based gain solutions.
+For this comparison we run a suite of simulations varying the receiver noise and integration times, while forming antenna-pixel correlations, $\Cna$, for EPICal gain solutions, and simultaneously forming visibilities from the same E-field streams to find visibility-based gain solutions.
 
@@
 We simulate 64 40~kHz frequency channels.
-Because our calibration loop is frequency independent, each channel can be treated as a separate trial of the simulation, and is used to better estimate the statistics.
+Because our calibration loop treats frequency channels independently, each channel can be treated as a separate trial of the simulation, and is used to better estimate the statistics.
 
@@
 
-We initialize our gain estimates with the true values (unity), and allow the estimate to be corrupted by the noise through one iteration of the calibration loop.
+We initialize our gain estimates with the true values (${g^{(0)}_a=g^T_a=1}$), and allow the estimate to be corrupted by the noise through one iteration of the calibration loop.
 Ideally we would use a damping factor of zero to allow for the noise to fully corrupt the gains.
-However, we found that a zero damping factor resulted in non-convergent results, so we adopted a factor, $\damp=0.05$.
+However, we found that a zero damping factor resulted in non-convergent results, so we adopted a small factor, $\damp=0.05$.
 For a fair comparison, we use the same factor for the visibility-based solutions.
@@
 In addition to requiring the sky model, the calibration also requires this model to be attenuated by the primary beam, which can be difficult to measure at the precision necessary \citep[e.g.][]{neb15,vir14,thy15b}.
-However, the direct imaging correlator provides exactly the model necessary in real time, and can be iterated over to improve these images and the gain solutions.
+However, the direct imaging correlator provides exactly the model necessary in real time, and can be iterated over to improve the images and gain solutions.
 
@@
 For example, a stable instrument may be able to use short integrations to determine a rough estimate of the gains before switching to much longer integration (on order seconds to minutes) to highly increase signal to noise.
-At low frequencies or high sky resolution, the dynamics of the ionosphere are important, and will likely drive the limit of time integration allowed.
+At low frequencies or high imaging resolution, the dynamics of the ionosphere are important, and will likely drive the limit of time integration allowed.
+
+\textbf{Multiple pixel correlations.} As pointed out in section~\ref{sec:noise}, for a sky-noise dominated system, the signal to noise in the calibrator pixel dominates the error on the EPICal generated gain estimates.
+Besides the methods described above to increase signal to noise, one could also use correlations of multiple image pixels to incorporate a higher fraction of the total sky noise.
+Of course this increases the computational cost to a scaling of $\mathcal{O}(\Nant N_{\mathrm{pix}})$, which will typically be much lower than the $\mathcal{O}(\Ng \log_2 \Ng)$ of the correlator itself.
 
 \textbf{Direction dependent gains.} Even below the timescale of a dynamic ionosphere, a widefield instrument may require direction dependent gains to account for a varying thickness ionospheric sheet.
-The holographic mapping framework used by the MOFF algorithm, and in turn the EPIC and EPICal software can naturally account for direction dependence through and updated model of the antenna voltage pattern.
-This can be achieved by first forming complex gain estimates toward multiple directions on the sky by selecting several pixels and independently following the EPICal algorithm for each.
+The holographic mapping framework used by the MOFF algorithm, and in turn the EPIC and EPICal software can naturally account for direction dependence through an updated model of the antenna voltage pattern.
+This can be achieved by using multiple pixel correlations to independently solve for gains in each pixel direction.
 These solutions can then be used to fit a beam model on the sky (analogous to the gain model fitting mentioned above), and updating the antenna voltage pattern.
 This updated pattern would then feed into the gridding step of the correlator, allowing the imager to convolve the signals with the effective beam pattern in the ground plane.
-Because this method requires multiple pixel correlations, the scaling now becomes $\mathcal{O}(\Nant N_{\mathrm{pix}})$, which will typically be much lower than the $\mathcal{O}(\Ng \log_2 \Ng)$ of the correlator itself.
 

### 2016-02-22 cal_demo/ms.tex
One comment in section 1
(*) Need to point exactly why MOFF needs calibration right up front
and how the situation is different for a visibility-based calibration

@@
 Traditionally, pair-wise visibilities are written to disk and used to calibrate offline.
-However, a direct imaging correlator mixes the signals from all antennas before averaging and writing to disk, making calibration a requirement at the front end.
+However, a direct imaging correlator mixes the signals from all antennas before averaging and writing to disk, making calibration a requirement at the front end ({\bf need to make this distinction more clear by pointing where exactly calibration happens in a X-based system and MOFF}).
 Previous solutions have involved applying calibration solutions generated from a parallel FX correlator \citep{zhe14, fos14}, or integrating a dedicated FX correlator which periodically formed the full visibility matrix to solve for gains \citep{wij09,dev09}.

### 2016-02-22 cal_demo/ms.tex
Add comments to section 2
(*) Comments are in bold face

@@
 The MOFF algorithm next calls for a calibration.
-We will assume we have formed an estimate of the gains after $n$ iterations of a calibration loop, and derive an updated $n+1$ estimate.
+We will assume we have formed an estimate of the gains after $n$ iterations of a calibration loop, and derive an updated {\bf estimate for the ($n+1$)-{\it th} iteration}.
 How the calibration gain is applied can depend on the dominant source of the noise in the measurement.
@@
 However, if the noise is dominated in the receiver (after the complex gain factor of equation~\ref{eq:apply_gain}), a lower amplitude gain results in lower signal to noise for that antenna.
-Therefore, one should multiply by the conjugate of the gain to down weight appropriately while removing the phase acquired.
+Therefore, one should multiply by the conjugate of the gain to down weight {\bf it} appropriately while removing the phase acquired.
 We will abstain from this choice for now, and instead use a multiplicative factor $h^{(n)}_a$ to represent the application of our $n^{\mathrm{th}}$-loop estimate of the gain for antenna $a$:
@@
 h^{(n)}_a=\begin{cases}
-1/g^{(n)}_a & \mbox{sky noise dominated} \\ 
-g^{*(n)}_a & \mbox{receiver noise dominated}.
+1/g^{(n)}_a, & \mbox{sky noise dominated} \\ 
+g^{*(n)}_a, & \mbox{receiver noise dominated}.
 \end{cases}
@@
 \begin{align}\label{eq:epix}
-\Ethetaest(\s_i,t) & = \frac{1}{\Nant} \sum_a h^{(n)}_a g^T_a\Er{a}^T(t) e^{2\pi i \s_i \cdot \ra}\sum_i \beamr_b(\ri-\ra)e^{2\pi i \s_i \cdot (\ri-\ra)} \nonumber\\
+\Ethetaest(\s_i,t) & = \frac{1}{\Nant} \sum_a h^{(n)}_a g^T_a\Er{a}^T(t) e^{2\pi i \s_i \cdot \ra} \nonumber \\ 
+  & \qquad\qquad \sum_i \beamr_b(\ri-\ra)e^{2\pi i \s_i \cdot (\ri-\ra)} \nonumber\\
 & = \frac{1}{\Nant} \sum_a h^{(n)}_a g^T_a\Er{a}^T(t) e^{2\pi i \s_i \cdot \ra}\beamtheta_a(\s_i)
@@
 It is easy to see from here that the net effect of including the receiver noise term, $\widetilde{n}_a(f,t)$, will result in added noise on the true visibilities, including a bias on the auto-correlation terms.
-In principle a bias can result on any visibility from any noise correlations between antennas, but the implementation included in the EPIC software package restricts the noise term to include only the auto-correlation noise bias, and assumes baseline dependent noise is zero mean.
+In principle a bias can result on any visibility from any noise correlations between antennas, but the implementation included in the EPIC software package restricts the noise term to include only the auto-correlation noise bias, and assumes baseline dependent noise {\bf has} zero mean.
 
@@
 While $g'^{(n+1)}_a$ is the best estimate of the gains after the $n^{\textrm{th}}$ iteration, the damped version, $g^{(n+1)}_a$, is actually used in the iterative calibration loop.
-Once the loop converges the difference is a longer effective integration for the damped version (lower thermal noise).
+Once the loop converges the difference is a longer effective integration for the damped version (lower thermal noise) ({\bf the last sentence is too condensed that it is not clear what you are trying to say}).
 
@@
 Computationally expensive steps that must be performed ``on-chip" are shown inside the gray box.
-The uncalibrated antenna signals are tapped out after the F-engine and correlated against the output image pixel of interest.
+The uncalibrated antenna signals are tapped out after the F-engine and correlated against the output image pixel of interest ({\bf i.e., equation~\ref{eq:Cna} is ``on-chip''}).
 The correlated values are then passed off-chip to estimate the gains using equation~\ref{eq:cal_solution}, and additional fitting if desired.
@@
 These gain estimates are then fed back into the correlator to be applied to subsequent data, and the process is repeated.
-The gray box shows operations which must be done at high speed (before averaging in time), which white boxes show operations which can be performed ``off-chip".}
+The gray box shows operations which must be done at high speed (before averaging in time), which white boxes show operations which can be performed ``off-chip".
+({\bf I have some suggestions to tweak this figure.})}
 \label{fig:schematic}
@@
 This requires generating these correlations, which again only scale as $\mathcal{O}(\Nant)$, and are generally useful for array diagnostics.
+({\bf What happened to the $g_a^T$ term outside the sum?
+And why did you drop $\hat{\mathbf{s}}_0$ in $W_a^*$}?)
 
@@
 \end{equation}
+({\bf It is not clear what happened to the $g$ terms.})
 

### 2016-02-23 cal_demo/ms.tex
Add comments to section 3

@@
 
-For our unknown gains, we create a set of random numbers with amplitude $1\pm0.25$, and completely random phase.
+For our unknown gains, we create a set of random numbers with amplitude $1\pm0.25$ ({\bf replace with amplitude in range [0.75, 1.25]}), and completely random phase.
 These are our ``true gains", and we apply them to the frequency-domain simulated antenna electric fields as in equation~\ref{eq:apply_gain}.
@@
 Images created at the beginning of calibration and at the end are shown in figure~\ref{fig:sim_images}.
-Each image is 10~ms integration, corresponding to all snapshot images created with a given set of gain estimates.
+Each image is {\bf obtained over} 10~ms integration, corresponding to all snapshot images created with a given set of gain estimates.
 The top panel shows the image produced with our initialized unity gains.

### 2016-02-23 cal_demo/ms.tex
Add comments in section 4

@@
 The data is from the LWA narrow-band transient buffer (TBN), with time ordered voltage data from 255 antennas within a core radius of 100~m.
-The central frequency is 74.03 MHz, and a bandwidth of 100 kHz and writeout timescale of 5.12 ms (frequency channel resolution of 195.3125 Hz).
+The central frequency is 74.03 MHz, {\bf with} a bandwidth of 100 kHz and writeout timescale of 5.12 ms (frequency channel resolution of 195.3125 Hz).
 For this demonstration we limit ourselves to a single polarization.
@@
 After correcting for geometric cable delays, the instrument is naturally well calibrated, as was seen in the demonstration of the EPIC imager in \citealt{thy15c}.
-However, we will aim to improve yet on this calibration using our algorithm.
+However, we will aim to improve {\bf further} on this calibration using our algorithm.
 
@@
 Here we average solutions across 150 channels, or about 29 kHz.
-With a fractional bandwidth $B/f_0 = 3.9 \times 10^{-4}$, we felt safe assuming a smooth bandpass.
-Because there is still a fair amount of noise in the data, a damping factor of $\gamma = 0.7$ was adopted.
+With a fractional bandwidth $B/f_0 = 3.9 \times 10^{-4} \ll 1$, {\bf we assume a smooth bandpass across the band}.
+% Because there is still a fair amount of noise in the data,
+A damping factor of $\gamma = 0.7$ was adopted.
 
-Finally, we found that seven antennas\footnote{LWA antenna IDs 48, 85, 124, 148, 203, 217, and 244 were flagged.} produced unstable gain solutions, and in fact corrupted the ensemble.
-We therefore flagged these antennas for our analysis, resulting in a total of 248 antennas to calibrate.
-In each calibration loop, we form our feedback correlations and update our gain estimates over 10 timestamps (51.2 ms).
+Finally, we found that seven antennas\footnote{LWA antenna IDs 48, 85, 124, 148, 203, 217, and 244 were flagged.} produced unstable gain solutions, and in fact corrupted the ensemble ({\bf specifically what is the ensemble exactly?}).
+We therefore flagged these antennas {\bf in} our analysis, resulting in a total of 248 antennas to calibrate.
+In each calibration loop, we form $C_{i,\hat{\boldsymbol{s}}_0}, i=1,2,\ldots N_a$ and update our gain estimates over 10 timestamps (51.2 ms).
 We iterate the loop 30 times for a total of 1.536 seconds.
-The results of this experiment are shown in figures~\ref{fig:data_phase}~--~\ref{fig:data_images}.
+The results of this {\bf calibration} experiment are shown in figures~\ref{fig:data_phase}~--~\ref{fig:data_images}.
 
-Figure~\ref{fig:data_phase} shows the phase of our gain estimates over our 30 calibration iterations, again with each colored line representing a different antenna.
+Figure~\ref{fig:data_phase} shows the phase of our gain estimates over 30 calibration iterations, again with each colored line representing a different antenna.
 Given the quality of uncalibrated image demonstrated in \cite{thy15c}, it is a bit surprising to see the phase variation in our solutions.
@@
 The middle panel shows the image produced after calibration with identical integration time and bandwidth.
-The sidelobes throughout the image are significantly suppressed, and the galactic plane is much more evident, despite only modeling Cyg~A and Cas~A.
+The sidelobes ({\bf or is it better to call them speckles due to uncalibrated phases because sidelobes are typically associated with the psf and not with calibration?}) throughout the image are significantly suppressed, and the galactic plane is much more evident, despite only modeling Cyg~A and Cas~A.
 We also note that the feature just to the right of Cyg~A is dimmer in the calibrated image, better matching the expected flux from the global sky model (GSM, \citealt{deo08}).
-We show the GSM modulated by two factors of the LWA primary beam in the right panel of figure~\ref{fig:data_images} for reference.
+We show the GSM modulated by two factors of the LWA {\bf power pattern} in the right panel of figure~\ref{fig:data_images} for reference.
 
@@
 To compare the image qualities quantitatively we compute the dynamic range, defined as the peak of the image over the noise level.
-We estimate the noise level as the median of the absolute deviation of the image.
+We estimate the noise level as the median of the absolute deviation of the image, {\bf a robust statistic}.
 With this metric we find the calibrated image to have a 55\% dynamic range improvement over the uncalibrated image.

### 2016-02-23 cal_demo/ms.tex
Add a comment in section 3
Add comments in section 5

@@
 Without the damping factor, the oscillation seen in the first couple iterations would have been significantly larger and taken much longer to fade out.
+({\bf Add a sentence or two to explain the oscillation after convergence due to intrinsic source fluctuations.})
 
@@
 As in section~\ref{sec:sim}, we use the MWA core layout (51 antennas), and a 4.4~m square tophat antenna voltage pattern.
-We simulate 64 40~kHz frequency channels.
+We simulate 64 {\bf frequency channels each of width 40~kHz}.
 Because our calibration loop treats frequency channels independently, each channel can be treated as a separate trial of the simulation, and is used to better estimate the statistics.
@@
 Ideally we would use a damping factor of zero to allow for the noise to fully corrupt the gains.
-However, we found that a zero damping factor resulted in non-convergent results, so we adopted a small factor, $\damp=0.05$.
+However, we found that a zero damping factor resulted in non-convergent results ({\bf this statement makes me nervous and makes it sound these results may not be robust.
+It sounds like a higher $\damp$ is just delaying the divergence.}), so we adopted a small factor, $\damp=0.05$.
 For a fair comparison, we use the same factor for the visibility-based solutions.
-We then observe the error in the gain estimates introduced by the noise.
+We then observe the error in the gain estimates introduced by the noise,
 \begin{equation}\label{eq:gain_error}
-\sigma_g = \left[\frac{1}{N_f \Nant} \sum_f \sum_a \frac{\left|g'_a(f)-g^T_a(f)\right|^2}{\left|g^T_a(f)\right|^2}\right]^{1/2}
+\sigma_g = \left[\frac{1}{N_f \Nant} \sum_f \sum_a \frac{\left|g'_a(f)-g^T_a(f)\right|^2}{\left|g^T_a(f)\right|^2}\right]^{1/2}.
 \end{equation}
+({\bf Here is my major concern.
+I feel these statistics should be estimated after making sure gains have converged after few loops, otherwise these statistics may not be accurate.
+Since you do not form images, I don't think it should take very long.})
 
@@
 The error of the visibility-based gains is estimated in the same way as in equation~\ref{eq:gain_error}.
+({\bf How is $\widetilde{V}_{ab}$ defined?
+As the average of all visibilities between that antenna pair in the calibration loop timescale?
+Or are the gains determined on every timestamp using linear algebraic minimization?})
 
@@
 The gain errors from our EPICal method is shown with solid lines, each color representing different levels of receiver noise.
-Each data point represents a separate simulation run (independent realization of the noise).
+Each data point ({\bf I see no discrete data points.
+Can you add those to the plot?}) represents a separate simulation run (independent realization of the noise).
 The general trends are consistent with expectations -- decreasing as the square root of integration time, and increasing with receiver noise.
-The lowest receiver noise level (1.0 Jy) is significantly below the sky noise contribution (about 8 Jy), so the black line is not as low, compared to the red, as one would expect from receiver noise consideration alone.
+The lowest receiver noise level (1.0 Jy) is significantly below the sky noise contribution (about 8 Jy), so the black line is not as low, compared to the red, as one would expect from receiver noise consideration alone ({\bf Need to elaborate.
+It is too terse, I couldn't follow.
+}).
 
@@
 Generally we see the visibility-based calibration achieves slightly lower noise solutions.
-On average the EPICal errors are 1.5\% higher, and at the worst is 3.5\% higher.
+On average the EPICal errors are 1.5\% higher, and at the worst is 3.5\% higher ({\bf but need to also explain this trend reverses and dependent on receiver noise and integration time}).
 
@@
 In the more complicated sky case (about 23 Jy total sky power), we saw the EPICal gains exhibit errors 20\% higher than the visibility derived gains.
-The trend of higher EPICal errors with more complicated skies is due to the single correlated pixel not containing the full signal to noise of the data.
-This reaffirms our intuition that using a pixel with a bright source is ideal, but will still not match visibilities for complicated skies.
+The trend of higher EPICal errors with more complicated skies is due to the single correlated pixel not containing the full signal to noise of the data ({\bf Here is my other major concern.
+There is no sort of numerical substantiation offered to confirm this other than saying we have not done this yet and thus missing information.}).
+This reaffirms our intuition that using a pixel with a bright source is ideal, but will still not match visibilities for complicated skies ({\bf to reaffirm we need some sort of numerical evidence that this is at play and we will work on this further}).
 We suggest additional strategies for improving this signal to noise in the following section.
+({\bf At the moment it sounds like we are conveniently sweeping it under the rug.
+Can we at least qualitatively make arguments that this should improve and what might that information from multiple pixels look like mathematically?})
 

### 2016-02-23 cal_demo/ms.tex
Add comments in section 6

@@
 
-\textbf{Fitting gain models.} With some knowledge of the instrumental bandpass, the signal-to-noise of the gain solutions can be greatly improved by fitting a model to the per-frequency solutions derived here.
+\textbf{Fitting gain models.} With some knowledge of the instrumental bandpass, the signal-to-noise of the gain solutions ({\bf you mean of the image?}) can be greatly improved by fitting a model to the per-frequency solutions derived here.
 We used this method at a rudimentary level in our demonstration to LWA data by assuming the gains were constant over a narrow bandwidth.
@@
 
-\textbf{Dynamic parameters.} In our controlled experiments we fine tuned a number of parameters based on our testing (e.g.
+\textbf{Dynamic parameters.} In our controlled experiments we fine-tuned a number of parameters based on our testing (e.g.
 damping factor, integration time, frequency averaging).

### 2016-02-23 cal_demo/ms.tex
Add comments in Abstract

@@
 This method scales as the number of antennas, and produces results comparable to those from visibilities.
-Through simulations and application to Long Wavelength Array data we show this algorithm is a feasible solution for next generation instruments.
+Through simulations and application to Long Wavelength Array data we show this algorithm is a feasible solution for next generation instruments ({\bf I am not sure if we have explicitly shown this last part to be true.
+For instance, people would raise questions about complex sky, direction-dependent gains, etc.
+when talking about feasibility for future instruments.
+We could tone it down to something like it shows promise for future telescopes.}).
 \end{abstract}

### 2016-02-25 cal_demo/ms.tex
Starting to address Nithyas comments on cal paper

--- 2016-02-23 cal_demo/ms.tex
@@
 This method scales as the number of antennas, and produces results comparable to those from visibilities.
-Through simulations and application to Long Wavelength Array data we show this algorithm is a feasible solution for next generation instruments.
+Through simulations and application to Long Wavelength Array data we show this algorithm is a promising solution for next generation instruments.
 \end{abstract}
@@
 A challenge common to all direct imaging algorithms is calibration of the antenna gains.
-Traditionally, pair-wise visibilities are written to disk and used to calibrate offline.
-However, a direct imaging correlator mixes the signals from all antennas before averaging and writing to disk, making calibration a requirement at the front end ({\bf need to make this distinction more clear by pointing where exactly calibration happens in a X-based system and MOFF}).
+With a traditional FX or XF correlator, pair-wise visibilities are written to disk and used to calibrate offline before further processing such as imaging.
+However, a direct imaging correlator mixes the signals from all antennas before averaging and writing to disk, making calibration a requirement at the front end, before imaging and averaging.
 Previous solutions have involved applying calibration solutions generated from a parallel FX correlator \citep{zhe14, fos14}, or integrating a dedicated FX correlator which periodically formed the full visibility matrix to solve for gains \citep{wij09,dev09}.
@@
 The MOFF algorithm next calls for a calibration.
-We will assume we have formed an estimate of the gains after $n$ iterations of a calibration loop, and derive an updated {\bf estimate for the ($n+1$)-{\it th} iteration}.
+We will assume we have formed an estimate of the gains after $n$ iterations of a calibration loop, and derive an updated estimate for the ($n+1$)$^{\mathrm{th}}$ iteration.
 How the calibration gain is applied can depend on the dominant source of the noise in the measurement.
@@
 It is easy to see from here that the net effect of including the receiver noise term, $\widetilde{n}_a(f,t)$, will result in added noise on the true visibilities, including a bias on the auto-correlation terms.
-In principle a bias can result on any visibility from any noise correlations between antennas, but the implementation included in the EPIC software package restricts the noise term to include only the auto-correlation noise bias, and assumes baseline dependent noise {\bf has} zero mean.
+In principle a bias can result on any visibility from any noise correlations between antennas, but the implementation included in the EPIC software package restricts the noise term to include only the auto-correlation noise bias, and assumes baseline dependent noise has zero mean.
 
@@
 While $g'^{(n+1)}_a$ is the best estimate of the gains after the $n^{\textrm{th}}$ iteration, the damped version, $g^{(n+1)}_a$, is actually used in the iterative calibration loop.
-Once the loop converges the difference is a longer effective integration for the damped version (lower thermal noise) ({\bf the last sentence is too condensed that it is not clear what you are trying to say}).
+Once the loop converges the damped version is essentially a weighted average over the past several iterations, giving it a longer effective integration time.
 
@@
 Computationally expensive steps that must be performed ``on-chip" are shown inside the gray box.
-The uncalibrated antenna signals are tapped out after the F-engine and correlated against the output image pixel of interest ({\bf i.e., equation~\ref{eq:Cna} is ``on-chip''}).
+The uncalibrated antenna signals are tapped out after the F-engine and correlated against the output image pixel of interest (i.e., equation~\ref{eq:Cna} is ``on-chip'').
 The correlated values are then passed off-chip to estimate the gains using equation~\ref{eq:cal_solution}, and additional fitting if desired.
@@
 \begin{equation}
-\Cna \rightarrow \Cna - \frac{1}{\Nant} h^{*(n)}_a\beamtheta^*_a e^{-2\pi i \spix \cdot \ra} \left<|\Er{a}|^2\right>_t
+\Cna \rightarrow \Cna - \frac{1}{\Nant} h^{*(n)}_a\beamtheta^*_a(\spix) e^{-2\pi i \spix \cdot \ra} \left<|\Er{a}|^2\right>_t
 \end{equation}
@@
 
-For our unknown gains, we create a set of random numbers with amplitude $1\pm0.25$ ({\bf replace with amplitude in range [0.75, 1.25]}), and completely random phase.
+For our unknown gains, we create a set of random complex numbers where the amplitude is drawn from approximately gaussian distribution centered around 1 with width 0.25, and completely random phase.
 These are our ``true gains", and we apply them to the frequency-domain simulated antenna electric fields as in equation~\ref{eq:apply_gain}.
@@
 Without the damping factor, the oscillation seen in the first couple iterations would have been significantly larger and taken much longer to fade out.
-({\bf Add a sentence or two to explain the oscillation after convergence due to intrinsic source fluctuations.})
 
@@
 
+After the gains have converged, we see both the phases and amplitudes to continue to fluctuate coherently.
+This is due to the stochastic fluctuations of the simulated sources themselves, and is limited by the sky noise.
+For this simulation we restricted each calibration iteration to only 10~ms integration time.
+In principle a calibration implementation could use short integration times to allow the gains to converge, then increase the integration time to reduce this noise.
+
 Images created at the beginning of calibration and at the end are shown in figure~\ref{fig:sim_images}.
-Each image is {\bf obtained over} 10~ms integration, corresponding to all snapshot images created with a given set of gain estimates.
+Each image is obtained over 10~ms integration, corresponding to all snapshot images created with a given set of gain estimates.
 The top panel shows the image produced with our initialized unity gains.
@@
 The data is from the LWA narrow-band transient buffer (TBN), with time ordered voltage data from 255 antennas within a core radius of 100~m.
-The central frequency is 74.03 MHz, {\bf with} a bandwidth of 100 kHz and writeout timescale of 5.12 ms (frequency channel resolution of 195.3125 Hz).
+The central frequency is 74.03 MHz, with a bandwidth of 100 kHz and writeout timescale of 5.12 ms (frequency channel resolution of 195.3125 Hz).
 For this demonstration we limit ourselves to a single polarization.
@@
 After correcting for geometric cable delays, the instrument is naturally well calibrated, as was seen in the demonstration of the EPIC imager in \citealt{thy15c}.
-However, we will aim to improve {\bf further} on this calibration using our algorithm.
+However, we will aim to further improve on this calibration using our algorithm.
 
@@
 Here we average solutions across 150 channels, or about 29 kHz.
-With a fractional bandwidth $B/f_0 = 3.9 \times 10^{-4} \ll 1$, {\bf we assume a smooth bandpass across the band}.
-% Because there is still a fair amount of noise in the data,
+With a fractional bandwidth $B/f_0 = 3.9 \times 10^{-4} \ll 1$, we assume a smooth bandpass across the band.
 A damping factor of $\gamma = 0.7$ was adopted.
 
-Finally, we found that seven antennas\footnote{LWA antenna IDs 48, 85, 124, 148, 203, 217, and 244 were flagged.} produced unstable gain solutions, and in fact corrupted the ensemble ({\bf specifically what is the ensemble exactly?}).
-We therefore flagged these antennas {\bf in} our analysis, resulting in a total of 248 antennas to calibrate.
+Finally, we found that seven antennas\footnote{LWA antenna IDs 48, 85, 124, 148, 203, 217, and 244 were flagged.} produced unstable gain solutions, and in fact corrupted the entire array.
+We therefore flagged these antennas in our analysis, resulting in a total of 248 antennas to calibrate.
 In each calibration loop, we form $C_{i,\hat{\boldsymbol{s}}_0}, i=1,2,\ldots N_a$ and update our gain estimates over 10 timestamps (51.2 ms).
 We iterate the loop 30 times for a total of 1.536 seconds.
-The results of this {\bf calibration} experiment are shown in figures~\ref{fig:data_phase}~--~\ref{fig:data_images}.
+The results of this calibration experiment are shown in figures~\ref{fig:data_phase}~--~\ref{fig:data_images}.
 
@@
 We also note that the feature just to the right of Cyg~A is dimmer in the calibrated image, better matching the expected flux from the global sky model (GSM, \citealt{deo08}).
-We show the GSM modulated by two factors of the LWA {\bf power pattern} in the right panel of figure~\ref{fig:data_images} for reference.
+We show the GSM modulated by two factors of the LWA power pattern in the right panel of figure~\ref{fig:data_images} for reference.
 
@@
 To compare the image qualities quantitatively we compute the dynamic range, defined as the peak of the image over the noise level.
-We estimate the noise level as the median of the absolute deviation of the image, {\bf a robust statistic}.
+We estimate the noise level as the median of the absolute deviation of the image, a robust statistic.
 With this metric we find the calibrated image to have a 55\% dynamic range improvement over the uncalibrated image.
@@
 As in section~\ref{sec:sim}, we use the MWA core layout (51 antennas), and a 4.4~m square tophat antenna voltage pattern.
-We simulate 64 {\bf frequency channels each of width 40~kHz}.
+We simulate 64 frequency channels each of width 40~kHz.
 Because our calibration loop treats frequency channels independently, each channel can be treated as a separate trial of the simulation, and is used to better estimate the statistics.

### 2016-02-29 cal_demo/ms.tex
starting to incorporate miguels comments on cal paper

@@
 In order to satisfy the survey speeds required for precision cosmology as well as searches for fast radio transients, radio astronomy is undergoing a paradigm shift toward interferometers consisting of hundreds to thousands of small, widefield antennas.
-Many arrays with this design are already built or under construction including the Hydrogen Epoch of Reionization Array\footnote{http://reionization.org} (HERA), the Murchison Widefield Array (MWA; \citealt{tin13,bow13}), the Precision Array for Probing the Epoch of Reionization (PAPER; \citealt{par10}), the LOw Frequency ARray (LOFAR; \citealt{van13}), the Canadian Hydrogen Intensity Mapping Experiment (CHIME,\citealt{ban14}), the Long Wavelength Array (LWA, \citealt{ell13}), and the low frequency Square Kilometer Array (SKA1-Low \citealt{mel13}).
+Many arrays with this design are already built or under construction including the Hydrogen Epoch of Reionization Array\footnote{http://reionization.org} (HERA), the Murchison Widefield Array (MWA; \citealt{tin13,bow13}), the Donald C.
+Backer Precision Array for Probing the Epoch of Reionization (PAPER; \citealt{par10}), the LOw Frequency ARray (LOFAR; \citealt{van13}), the Canadian Hydrogen Intensity Mapping Experiment (CHIME,\citealt{ban14}), the Long Wavelength Array (LWA, \citealt{ell13}), and the low frequency Square Kilometer Array (SKA1-Low \citealt{mel13}).
 
@@
 \end{equation}
-Note that this quantity is neither a true or estimated value.
+Note that this quantity is neither a true or estimated value, but rather the raw electric field signal corrupted by the complex gain before we attempt to estimate the true value.
 The noise term is strictly receiver noise -- noise introduced by the instrument.
@@
 
-The goal of our calibration algorithm will be to estimate the antenna gains.
+The goal of our calibration algorithm will be to estimate the antenna gains, $g_a$.
 We will assume the gains have no time dependence within the timescale of finding our solutions.
@@
 The correlated values are then passed off-chip to estimate the gains using equation~\ref{eq:cal_solution}, and additional fitting if desired.
-The gains are then passed back to the correlator to update the calibration for subsequent data streams.
+The gains are then passed back to the correlator to update the calibration for subsequent integration intervals.
 

### 2016-02-29 cal_demo/ms.tex
Removed ambiguous calibration term "h"

@@
 The MOFF algorithm next calls for a calibration.
-We will assume we have formed an estimate of the gains after $n$ iterations of a calibration loop, and derive an updated estimate for the ($n+1$)$^{\mathrm{th}}$ iteration.
-How the calibration gain is applied can depend on the dominant source of the noise in the measurement.
-If the noise is dominated by sky fluctuations, the noise level is affected by the antenna gains so dividing by the complex gains will flatten the noise across antennas.
-However, if the noise is dominated in the receiver (after the complex gain factor of equation~\ref{eq:apply_gain}), a lower amplitude gain results in lower signal to noise for that antenna.
-Therefore, one should multiply by the conjugate of the gain to down weight {\bf it} appropriately while removing the phase acquired.
-We will abstain from this choice for now, and instead use a multiplicative factor $h^{(n)}_a$ to represent the application of our $n^{\mathrm{th}}$-loop estimate of the gain for antenna $a$:
+We will assume we have formed an estimate of the gains after $n$ iterations of a calibration loop.
+We then correct the incoming electric field data stream using this current estimate.
 \begin{equation}
-\Erest{a} = h^{(n)}_a \Er{a}
+\Erest{a} = \Er{a}/ g^{(n)}_a,
 \end{equation}
-where,
-\begin{equation}
-h^{(n)}_a=\begin{cases}
-1/g^{(n)}_a, & \mbox{sky noise dominated} \\ 
-g^{*(n)}_a, & \mbox{receiver noise dominated}.
-\end{cases}
-\end{equation}
+These corrected fields will be used in the EPIC pipeline to derive an updated estimate for the ($n+1$)$^{\mathrm{th}}$ iteration.
 
@@
 \begin{equation}
-I'(\s_i) = \left<\left| \frac{1}{\Nant}\sum_i e^{2\pi i \mathbf{r}\cdot \s_i} \sum_a \beamr_a(\ri - \ra) h^{(n)}_a g^T_a \Er{a}^T(t) \right|^2 \right>_t.
+I'(\s_i) = \left<\left| \frac{1}{\Nant}\sum_i e^{2\pi i \mathbf{r}\cdot \s_i} \sum_a \beamr_a(\ri - \ra) \frac{g^T_a}{g^{(n)}_a} \Er{a}^T(t) \right|^2 \right>_t.
 \end{equation}
@@
 \begin{equation}
-\Ethetaest (\s_i,t) = \frac{1}{\Nant} \sum_i e^{2\pi i \ri \cdot \s_i} \sum_a \beamr_a(\ri - \ra) h^{(n)}_a g^T_a \Er{a}^T(t)
+\Ethetaest (\s_i,t) = \frac{1}{\Nant} \sum_i e^{2\pi i \ri \cdot \s_i} \sum_a \beamr_a(\ri - \ra) \frac{g^T_a}{g^{(n)}_a} \Er{a}^T(t)
 \end{equation}
@@
 \begin{align}\label{eq:epix}
-\Ethetaest(\s_i,t) & = \frac{1}{\Nant} \sum_a h^{(n)}_a g^T_a\Er{a}^T(t) e^{2\pi i \s_i \cdot \ra} \nonumber \\ 
-  & \qquad\qquad \sum_i \beamr_b(\ri-\ra)e^{2\pi i \s_i \cdot (\ri-\ra)} \nonumber\\
-& = \frac{1}{\Nant} \sum_a h^{(n)}_a g^T_a\Er{a}^T(t) e^{2\pi i \s_i \cdot \ra}\beamtheta_a(\s_i)
+\Ethetaest(\s_i,t) & = \frac{1}{\Nant} \sum_a \frac{g^T_a}{g^{(n)}_a}\Er{a}^T(t) e^{2\pi i \s_i \cdot \ra} \nonumber \\
+  & \qquad\times \sum_i \beamr_b(\ri-\ra)e^{2\pi i \s_i \cdot (\ri-\ra)} \nonumber\\
+& = \frac{1}{\Nant} \sum_a \frac{ g^T_a}{g^{(n)}_a}\Er{a}^T(t) e^{2\pi i \s_i \cdot \ra}\beamtheta_a(\s_i)
 \end{align}
@@
 \begin{align}\label{eq:cna}
-\Cna & = \left<g^T_a \Er{a}^T(t) \frac{1}{\Nant} \sum_b h^{*(n)}_b g^{T*}_b\Er{b}^{T*}(t) e^{-2\pi i \spix \cdot \rb} \beamtheta^*_b(\spix)\right>_t \nonumber \\
-& = \frac{g^T_a}{\Nant} \sum_b h^{*(n)}_b g^{T*}_b \beamtheta^*_b(\spix) e^{-2\pi i \spix \cdot \rb} \left<\Er{a}^T \Er{b}^{*T} \right>_t \nonumber \\
-& = \frac{g^T_a}{\Nant} \sum_b h^{*(n)}_b g^{T*}_b \beamtheta^*_b(\spix) e^{-2\pi i \spix \cdot \rb} \V^T_{ab}
+\Cna & = \left<g^T_a \Er{a}^T(t) \frac{1}{\Nant} \sum_b \frac{g^{T*}_b}{g^{*(n)}_b}\Er{b}^{T*}(t) e^{-2\pi i \spix \cdot \rb} \beamtheta^*_b(\spix)\right>_t \nonumber \\
+& = \frac{g^T_a}{\Nant} \sum_b \frac{g^{T*}_b}{g^{*(n)}_b} \beamtheta^*_b(\spix) e^{-2\pi i \spix \cdot \rb} \left<\Er{a}^T \Er{b}^{*T} \right>_t \nonumber \\
+& = \frac{g^T_a}{\Nant} \sum_b \frac{g^{T*}_b}{g^{*(n)}_b} \beamtheta^*_b(\spix) e^{-2\pi i \spix \cdot \rb} \V^T_{ab}
 \end{align}
@@
 \begin{equation}\label{eq:cal_solution}
-g'^{(n+1)}_a = \Cna \Nant \left[ \sum_b h^{*(n)}_b g^{*(n)}_b \beamtheta^*_b(\spix) e^{-2\pi i \spix \cdot \rb} \V^T_{ab} \right]^{-1}
+g'^{(n+1)}_a = \Cna \Nant \left[ \sum_b \beamtheta^*_b(\spix) e^{-2\pi i \spix \cdot \rb} \V^T_{ab} \right]^{-1}
 \end{equation}
@@
 However, the sky model can be precomputed offline, and the online computation complexity scales only as $\mathcal{O}(\Nant)$ as we form a $\Cna$ for each antenna.
-In the case where $h_b = 1/g_b$ (typical calibration procedure for sky-noise dominated systems), this simplifies slightly.
-\begin{equation}\label{eq:cal_solution_simple}
-g'^{(n+1)}_a = \Cna \Nant \left[ \sum_b \beamtheta^*_b(\spix) e^{-2\pi i \spix \cdot \rb} \V^T_{ab} \right]^{-1}
-\end{equation}
 
@@
 \begin{equation}
-\Cna \rightarrow \Cna - \frac{1}{\Nant} h^{*(n)}_a\beamtheta^*_a(\spix) e^{-2\pi i \spix \cdot \ra} \left<|\Er{a}|^2\right>_t
+\Cna \rightarrow \Cna - \frac{1}{\Nant\, g^{*(n)}_a} \beamtheta^*_a(\spix) e^{-2\pi i \spix \cdot \ra} \left<|\Er{a}|^2\right>_t
 \end{equation}
@@
 \begin{equation}
-\mathcal{C}^{(n)}_{a,0} \rightarrow \frac{1}{\Nant}\sum_b h^{*(n)}_b \V_{ab}
+\mathcal{C}^{(n)}_{a,0} \rightarrow \frac{1}{\Nant}\sum_b \frac{1}{g^{*(n)}_b} \V_{ab}
 \end{equation}
@@
 \begin{equation}
-g'^{(n+1)}_a \rightarrow \left[\sum_b h^{*(n)}_b \V_{ab}\right] \times \left[S_{\mathrm{src}}\sum_b h^{*(n)}_b g^{*(n)}_b\right]^{-1}
+g'^{(n+1)}_a \rightarrow \frac{\sum_b  \V_{ab}/g^{*(n)}_b}{\Nant S_{\mathrm{src}}}
 \end{equation}

### 2016-02-29 cal_demo/ms.tex
Added reference to grav wave followup. Using temporary cite to EM followup paper

@@
 As the number of elements in future arrays grows, the computational cost will become prohibitively expensive, and exploring efficient correlator schemes is essential to enable next generation instruments \citep{lon00}.
-Meanwhile, radio transient monitoring requires access to high time and frequency resolution data.
-For example, fast radio bursts (FRBs) are highly unexplored at low frequencies (< 1 GHz), but are expected to occur on timescales $\Delta t \sim$ 1--10~ms \citep{tho13}.
+Meanwhile, radio transient monitoring requires access to high time and frequency resolution data to identify and characterize events such as fast radio bursts (FRBs, \citealt{lor07}), or to follow up gravitational wave candidates with radio observations \citep{abb16a,abb16b}.
+FRBs are highly unexplored at low frequencies (< 1 GHz), but are expected to occur on timescales $\Delta t \sim$ 1--10~ms \citep{tho13}.
 Recording the full visibility matrix for $\Nant \gtrsim 10^3$ arrays at this timescale leads to extremely high data write rates.

### 2016-02-29 cal_demo/ms.tex
rewrote math section of cal paper to include a review of moff

@@
 
-We establish the mathematical framework and derive the calibration algorithm in \S \ref{sec:math}.
+We review the MOFF algorithm and derive the calibration algorithm in \S \ref{sec:math}.
 We then demonstrate the algorithm in simulations in \S \ref{sec:sim}, and apply to a sample LWA data set in \S \ref{sec:data}.
@@
 \section{Mathematical Framework}\label{sec:math}
-We begin by establishing the mathematical framework for the calibration problem.
-We derive the calibration solutions for the MOFF algorithm (adopting the notation of \citealt{thy15c}), but note the result is easily extended to FFT correlator algorithms by removing the gridding step.
+\subsection{Review of the MOFF algorithm}
+We begin by reviewing the data flow of the MOFF algorithm, highlighting aspects relevant to the calibration.
+The interested reader is encouraged to refer to \citealt{mor11} for a more thorough discussion.
+Figure~\ref{fig:moff_flow} is reproduced from \citealt{thy15c} to illustrate the various steps of the algorithm.
 
-The electric field incident on the ground, $\Er{}(\mathbf{r},f,t)$, is related to the sky electric field, $E(\hat{\mathbf{s}},f,t)$, through a Fourier transform.
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figures/MOFF_flowchart.pdf}
+\caption{Data flow for the MOFF algorithm, reproduced from \citealt{thy15c}.
+The time ordered electric fields are collected by the antenna elements then frequency transformed in the F-engine.
+The calibration is applied, then the electric field measurements are gridded using the antenna voltage patterns.
+The imager performs a spatial 2D FFT on the gridded data before squaring and averaging.
+The accumulated images are then written to disk.}
+\label{fig:moff_flow}
+\end{center}
+\end{figure}
+
+The electric field incident on the ground, $\Er{}(\mathbf{r},t)$, is related to the sky electric field, $E(\hat{\mathbf{s}},t)$, through a Fourier transform.
 \begin{equation}
-\Er{}(\mathbf{r},f,t) = \int E(\hat{\mathbf{s}},f,t) e^{-2\pi i \mathbf{r}\cdot \hat{\mathbf{s}}}\, \dif^2 \hat{\mathbf{s}}
+\Er{}(\mathbf{r},t) = \int E(\hat{\mathbf{s}},t) e^{-2\pi i \mathbf{r}\cdot \hat{\mathbf{s}}}\, \dif^2 \hat{\mathbf{s}}
 \end{equation}
-Here $\hat{\mathbf{s}}$ denotes the sine-projected unit vector for the sky angle, $\mathbf{r}$ is the observer's location (measured in wavelengths relative to an arbitrary origin), and $f$ and $t$ denote the frequency and time dependence, respectively.
+Here $\hat{\mathbf{s}}$ denotes the sine-projected unit vector for the sky angle, $\mathbf{r}$ is the observer's location (measured in wavelengths relative to an arbitrary origin).
 We will encounter several quantities which we attempt to estimate.
-We distinguish the ``true" values with a superscript $T$, while the estimates are denoted with a prime.
-We define the true antenna signal as a convolution of the antenna voltage pattern, $\beamr$, with the electric field on the ground.
+We distinguish the ``true'' values with a superscript $T$, while the estimated quantities are denoted with a prime.
+We define the true antenna signal as a convolution of the antenna voltage pattern, \beamr, with the electric field on the ground.
 \begin{equation}
-\Er{a}^T(f,t) \equiv \int \beamr_a(\mathbf{r}-\ra) \Er{}(\mathbf{r},f,t) \, \dif^2 \mathbf{r}
+\Er{a}^T(t) \equiv \int \beamr_a(\mathbf{r}-\ra) \Er{}(\mathbf{r},t) \, \dif^2 \mathbf{r}
 \end{equation}
@@
 
-We next model the measured, uncalibrated electric field as a multiplicative complex gain and an additive noise term applied to the true antenna electric field. 
+The first step of the MOFF algorithm is identical to the F-engine of a traditional FX correlator where the electric fields are Fourier transformed to frequency channels, $\Er{a}^T(f,t)$.
+Even after transforming along the time axis, we retain the time dependence of the signal to represent a stream of spectra.
+
+Thus far we have assumed the antennas deliver the true electric field signals. Here we introduce a multiplicative complex gain as well as an additive noise term which corrupt the measured signals.
 \begin{equation}\label{eq:apply_gain}
@@
 Any sky noise is implicitly included in the time dependence of the sky electric field.
+We will neglect $\widetilde{n}_a$ for now, but will inspect its effects at the end of this section.
+Each subsequent step treats each frequency channel independently, so we will drop the $f$ to simplify notation.
+
+The MOFF algorithm next calls for a calibration.
+The goal of our new calibration method will be to form an estimate of the gains, $g'_a$.
+For now we will assume we have formed an estimate to proceed with the MOFF pipeline.
+We correct the incoming electric field data stream using our current estimate of the gains.
+\begin{equation}
+\Erest{a}(t) = \Er{a}(t)/ g'_a
+\end{equation}
+
+The next step is to grid the antenna signals.
+MOFF uses the antenna patterns as the gridding kernel according to the software holography/A-transpose technique \citep{mor09,bha08}.
+This achieves both an optimal map \citep{teg97b} and places the data on a regular grid.
+Unlike traditional correlators which grid visibilities, the MOFF grids the electric fields directly.
+This operation must be performed for every time step, before any averaging of the data.
+
+After gridding, the imaging portion of the MOFF performs a 2D spatial FFT to form instantaneous electric field images of the sky.
+The pipeline up to this point can be summarized through the following equation.
+\begin{equation}
+\Ethetaest (\s_i,t) = \underbrace{\frac{1}{\Nant} \sum_i e^{2\pi i \ri \cdot \s_i}}_{\mathrm{2D\;FFT}} 
+\underbrace{\sum_a \beamr_a(\ri - \ra)}_{\mathrm{Gridding}}
+\underbrace{ \frac{g^T_a}{g'_a} \Er{a}^T(t)}_{\mathrm{Calibration}}
+\end{equation}
+The sum over antennas uses the voltage pattern to grid the calibrated electric fields onto regular gridpoint, $i$.
+The sum over $i$ denotes the 2D FFT to sky coordinates, resulting in an estimate for the instantaneous electric field image.
+We can simplify this expression by exchanging the sums to transform the beam term into sky coordinates.
+\begin{align}\label{eq:epix}
+\Ethetaest(\s_i,t) & = \frac{1}{\Nant} \sum_a \frac{g^T_a}{g'_a}\Er{a}^T(t) e^{2\pi i \s_i \cdot \ra} \nonumber \\
+  & \qquad\times \sum_i \beamr_b(\ri-\ra)e^{2\pi i \s_i \cdot (\ri-\ra)} \nonumber\\
+& = \frac{1}{\Nant} \sum_a \frac{ g^T_a}{g'_a}\Er{a}^T(t) e^{2\pi i \s_i \cdot \ra}\beamtheta_a(\s_i)
+\end{align}
+In this form we see that the effect of gridding with the antenna voltage pattern is to attenuate the image by a factor of the antenna sky response, $\beamtheta_a(\s_i)$.
+
+Finally, the images are squared and averaged in time to reduce output bandwidth, then written to disk.
+While operations have been reordered, the resulting image is equivalent to one produced by gridding and imaging visibilities.
+By shifting from cross-correlation of all pairs of antennas to a spatial FFT, the computational cost is significantly lowered for certain classes of compact arrays.
+The output bandwidth is also greatly reduced as visibilities are exchanged for gridded images.
+A full exploration of these scalings was presented in \citealt{thy15c}.
+
+\subsection{Derivation of calibration}
+We saw above that the MOFF algorithm (in general any direct imaging architecture) mixes the signal from all antennas coherently, and thus requires the calibration to be applied in real time at the front end of the correlator.
+Furthermore, visibilities are never formed, which are traditionally the basic measurement used to form calibration solutions.
+Here we derive an alternative method to estimate the antenna gains in real time, using the data products of the MOFF.
+
+Because the gains must be estimated in real time, we will need to use a finite stream of data to form and estimate which will be applied to the subsequent stream of data.
+Therefore the process of finding a solution is an iterative one.
+We will use parenthetical superscripts to denote the calibration loop number.
+For example, we will assume we have already formed an estimate from $n$ loops, $g^{(n)}_a$.
+We will use these estimates to calibrate the next stream of data in order to form an updated estimate, $g^{(n+1)}_a$.
+
+%%%%%%
+%%%%%%
+
+%We begin by establishing the mathematical framework for the calibration problem.
+We derive the calibration solutions for the MOFF algorithm (adopting the notation of \citealt{thy15c}), but note the result is easily extended to FFT correlator algorithms by removing the gridding step.
+%
+%The electric field incident on the ground, $\Er{}(\mathbf{r},f,t)$, is related to the sky electric field, $E(\hat{\mathbf{s}},f,t)$, through a Fourier transform.
+%\begin{equation}
+%\Er{}(\mathbf{r},f,t) = \int E(\hat{\mathbf{s}},f,t) e^{-2\pi i \mathbf{r}\cdot \hat{\mathbf{s}}}\, \dif^2 \hat{\mathbf{s}}
+%\end{equation}
+%Here $\hat{\mathbf{s}}$ denotes the sine-projected unit vector for the sky angle, $\mathbf{r}$ is the observer's location (measured in wavelengths relative to an arbitrary origin), and $f$ and $t$ denote the frequency and time dependence, respectively.
+We will encounter several quantities which we attempt to estimate.
+We distinguish the ``true" values with a superscript $T$, while the estimates are denoted with a prime.
+We define the true antenna signal as a convolution of the antenna voltage pattern, $\beamr$, with the electric field on the ground.
+%\begin{equation}
+%\Er{a}^T(f,t) \equiv \int \beamr_a(\mathbf{r}-\ra) \Er{}(\mathbf{r},f,t) \, \dif^2 \mathbf{r}
+%\end{equation}
+%The subscript $a$ labels the antenna, and $\mathbf{r}_a$ is the location of antenna $a$.
+%
+%We next model the measured, uncalibrated electric field as a multiplicative complex gain and an additive noise term applied to the true antenna electric field. 
+%\begin{equation}\label{eq:apply_gain}
+%\Er{a}(f,t) = g^T_a(f,t) \Er{a}^T(f,t) + \widetilde{n}_a(f,t)
+%\end{equation}
+%Note that this quantity is neither a true or estimated value, but rather the raw electric field signal corrupted by the complex gain before we attempt to estimate the true value.
+The noise term is strictly receiver noise -- noise introduced by the instrument.
+Any sky noise is implicitly included in the time dependence of the sky electric field.
 As the noise of modern low frequency arrays is heavily dominated by sky noise, we will neglect $\widetilde{n}_a$ for now, but will inspect its effects at the end of this section.
-
-The goal of our calibration algorithm will be to estimate the antenna gains, $g_a$.
+%
+%The goal of our calibration algorithm will be to estimate the antenna gains, $g_a$.
 We will assume the gains have no time dependence within the timescale of finding our solutions.
 Furthermore, we will treat each frequency channel independently, and drop the $f$ to simplify notation.
-
-The MOFF algorithm next calls for a calibration.
+%
+%The MOFF algorithm next calls for a calibration.
 We will assume we have formed an estimate of the gains after $n$ iterations of a calibration loop.
 We then correct the incoming electric field data stream using this current estimate.
-\begin{equation}
-\Erest{a} = \Er{a}/ g^{(n)}_a,
-\end{equation}
-These corrected fields will be used in the EPIC pipeline to derive an updated estimate for the ($n+1$)$^{\mathrm{th}}$ iteration.
+%\begin{equation}
+%\Erest{a} = \Er{a}/ g^{(n)}_a,
+%\end{equation}
+%These corrected fields will be used in the EPIC pipeline to derive an updated estimate for the ($n+1$)$^{\mathrm{th}}$ iteration. 
+%
+%A dirty image is next formed by gridding the calibrated fields with the antenna voltage pattern, Fourier transforming, squaring, and averaging in time.
+The estimated value for a pixel, $\s_i$, can be expressed as
+%\begin{equation}
+%I'(\s_i) = \left<\left| \frac{1}{\Nant}\sum_i e^{2\pi i \mathbf{r}\cdot \s_i} \sum_a \beamr_a(\ri - \ra) \frac{g^T_a}{g^{(n)}_a} \Er{a}^T(t) \right|^2 \right>_t.
+%\end{equation}
+%This is the final output of the MOFF correlator. However, for calibration purposes we will be interested in the electric field image just prior to squaring and averaging.
+%\begin{equation}
+%\Ethetaest (\s_i,t) = \frac{1}{\Nant} \sum_i e^{2\pi i \ri \cdot \s_i} \sum_a \beamr_a(\ri - \ra) \frac{g^T_a}{g^{(n)}_a} \Er{a}^T(t)
+%\end{equation}
+%We can simplify this expression by exchanging the sums to transform the beam term into sky coordinates.
+%\begin{align}\label{eq:epix}
+%\Ethetaest(\s_i,t) & = \frac{1}{\Nant} \sum_a \frac{g^T_a}{g^{(n)}_a}\Er{a}^T(t) e^{2\pi i \s_i \cdot \ra} \nonumber \\
+%  & \qquad\times \sum_i \beamr_b(\ri-\ra)e^{2\pi i \s_i \cdot (\ri-\ra)} \nonumber\\
+%& = \frac{1}{\Nant} \sum_a \frac{ g^T_a}{g^{(n)}_a}\Er{a}^T(t) e^{2\pi i \s_i \cdot \ra}\beamtheta_a(\s_i)
+%\end{align}
 
-A dirty image is next formed by gridding the calibrated fields with the antenna voltage pattern, Fourier transforming, squaring, and averaging in time.
-The estimated value for a pixel, $\s_i$, can be expressed as
-\begin{equation}
-I'(\s_i) = \left<\left| \frac{1}{\Nant}\sum_i e^{2\pi i \mathbf{r}\cdot \s_i} \sum_a \beamr_a(\ri - \ra) \frac{g^T_a}{g^{(n)}_a} \Er{a}^T(t) \right|^2 \right>_t.
-\end{equation}
-This is the final output of the MOFF correlator. However, for calibration purposes we will be interested in the electric field image just prior to squaring and averaging.
-\begin{equation}
-\Ethetaest (\s_i,t) = \frac{1}{\Nant} \sum_i e^{2\pi i \ri \cdot \s_i} \sum_a \beamr_a(\ri - \ra) \frac{g^T_a}{g^{(n)}_a} \Er{a}^T(t)
-\end{equation}
-We can simplify this expression by exchanging the sums to transform the beam term into sky coordinates.
-\begin{align}\label{eq:epix}
-\Ethetaest(\s_i,t) & = \frac{1}{\Nant} \sum_a \frac{g^T_a}{g^{(n)}_a}\Er{a}^T(t) e^{2\pi i \s_i \cdot \ra} \nonumber \\
-  & \qquad\times \sum_i \beamr_b(\ri-\ra)e^{2\pi i \s_i \cdot (\ri-\ra)} \nonumber\\
-& = \frac{1}{\Nant} \sum_a \frac{ g^T_a}{g^{(n)}_a}\Er{a}^T(t) e^{2\pi i \s_i \cdot \ra}\beamtheta_a(\s_i)
-\end{align}
-
-Next we move toward the feedback calibration outlined in \citealt{mor11}.
+%%%%
+%%%%
+As a starting point, we consider the feedback calibration outlined in \citealt{mor11}.
 There it was suggested to form a correlation of the uncalibrated antenna measurements with an image pixel from the output of the correlator.
@@
 
-Finally, we find an update to our gain solution by assuming our current estimate of the gains, $g^{(n)}_b$, is approximately correct and substitute into the sum for the true gains.
-We also require model visibilities formed from sky and primary beam models in place of true visibilities.
+The next step is to use this antenna-pixel correlation to update our gain solution.
+We will model the right hand side of equation~\ref{eq:cna} by assuming our current gain estimates are approximately correct, $g^{(n)}_b\approx g^T_b$, and replacing the true visibilities with a set of model visibilities, $\V'_{ab}$.
+These model visibilities rely an knowledge of the sky, and should be precomputed offline.
+We can then solve equation~\ref{eq:cna} for $g^T_a$ to achieve an updated estimate.
 \begin{equation}\label{eq:cal_solution}
-g'^{(n+1)}_a = \Cna \Nant \left[ \sum_b \beamtheta^*_b(\spix) e^{-2\pi i \spix \cdot \rb} \V^T_{ab} \right]^{-1}
+g'^{(n+1)}_a = \Cna \Nant \left[ \sum_b \beamtheta^*_b(\spix) e^{-2\pi i \spix \cdot \rb} \V'_{ab} \right]^{-1}
 \end{equation}
-This equation is our prescription for estimating the antenna gains of a direct imaging array.
-The approach is iterative in nature, and requires a sky model.
-However, the sky model can be precomputed offline, and the online computation complexity scales only as $\mathcal{O}(\Nant)$ as we form a $\Cna$ for each antenna.
-
-While testing we found equation~\ref{eq:cal_solution} resulted in oscillatory gain solutions as it was iterated, as is often the case in iterative minimization methods.
-To mitigate this we introduce a damping factor, $0 \leq \damp <1$, which is used to attenuate the gain update, effectively giving the solutions memory of previous iterations.
-\begin{equation}
-g^{(n+1)}_a = (1-\damp) g'^{(n+1)}_a + \damp g^{(n)}_a
-\end{equation}
-We found that while equation~\ref{eq:cal_solution} does indeed converge on good solutions, the process is made faster by tuning the damping factor.
-While $g'^{(n+1)}_a$ is the best estimate of the gains after the $n^{\textrm{th}}$ iteration, the damped version, $g^{(n+1)}_a$, is actually used in the iterative calibration loop.
-Once the loop converges the damped version is essentially a weighted average over the past several iterations, giving it a longer effective integration time.
+This equation is our prescription for estimating the antenna gains of a direct imaging array. The online computation complexity scales as $\mathcal{O}(\Nant)$ as we form a $\Cna$ for each antenna.
 
@@
 
-An important feature to note is that, like the MOFF-generated images themselves, equations~\ref{eq:cal_solution} and~\ref{eq:cal_solution_simple} include the antenna auto-correlations (the sum is over \emph{all} $b$, not excluding $a$).
+While testing we found equation~\ref{eq:cal_solution} resulted in oscillatory gain solutions as it was iterated, as is often the case in iterative minimization methods.
+To mitigate this we introduce a damping factor, $0 \leq \damp <1$, which is used to attenuate the gain update, effectively giving the solutions memory of previous iterations.
+\begin{equation}
+g^{(n+1)}_a = (1-\damp) g'^{(n+1)}_a + \damp g^{(n)}_a
+\end{equation}
+We found that while equation~\ref{eq:cal_solution} does indeed converge on good solutions, the process is made faster by tuning the damping factor.
+While $g'^{(n+1)}_a$ is the best estimate of the gains after the $n^{\textrm{th}}$ iteration, the damped version, $g^{(n+1)}_a$, is actually used in the iterative calibration loop.
+Once the loop converges the damped version is essentially a weighted average over the past several iterations, giving it a longer effective integration time.
+
+An important feature to note is that, like the MOFF-generated images themselves, equation~\ref{eq:cal_solution} includes the antenna auto-correlations (the sum is over \emph{all} $b$, not excluding $a$).
 It can be difficult to perfectly model the noise bias from auto-correlations, which can often times be far brighter than the visibilities themselves.
@@
 We have confirmed the limiting case equivalence here, and will explore the more general case in more detail in \S~\ref{sec:noise}.
+
+%%%%%
+%%%%%
 

### 2016-03-01 cal_demo/ms.tex
addressing more of nithyas comments

@@
 \end{equation}
+We have absorbed $g_a^T$ into the electric field so that the correlation is now simply the autocorrelation of the uncalibrated antenna measurement.
 This requires generating these correlations, which again only scale as $\mathcal{O}(\Nant)$, and are generally useful for array diagnostics.
-({\bf What happened to the $g_a^T$ term outside the sum?
-And why did you drop $\hat{\mathbf{s}}_0$ in $W_a^*$}?)
 
@@
 We can further drop the exponential phase terms because $\spix=0$.
-We then absorb the true gains and gain corrections into the true visibilities in equation~\ref{eq:cna} to express as a sum of measured visibilities.
+We then absorb the true gains into the true visibilities in equation~\ref{eq:cna} to express as a sum of measured, uncalibrated, visibilities.
 \begin{equation}
@@
 \end{equation}
-({\bf It is not clear what happened to the $g$ terms.})
 

### 2016-03-02 cal_demo/ms.tex
addressing more comments, tweaking schematic figure

@@
 \caption{The general data flow of the MOFF correlator, with a feedback calibration loop.
-A pixel from the (unsquared) image is tapped out and correlated against the input antenna electric field signals to form $\Cna$ coefficients, which are used to estimate the antenna gains.
-These gain estimates are then fed back into the correlator to be applied to subsequent data, and the process is repeated.
+A pixel from the (unsquared) image is tapped out and correlated against the input antenna electric field signals to form $\Cna$ coefficients (equation~\ref{eq:Cna}).
+These coefficients are then shipped off-chip to apply equations~\ref{eq:cal_solution}~and~\ref{eq:damping}, and any other processing to arrive at an updated estimate of the gains, $g^{(n)}_a$.
 The gray box shows operations which must be done at high speed (before averaging in time), which white boxes show operations which can be performed ``off-chip".
@@
 To mitigate this we introduce a damping factor, $0 \leq \damp <1$, which is used to attenuate the gain update, effectively giving the solutions memory of previous iterations.
-\begin{equation}
+\begin{equation}\label{eq:damping}
 g^{(n+1)}_a = (1-\damp) g'^{(n+1)}_a + \damp g^{(n)}_a
@@
 The computation necessary only scales with the number of antenna elements, making it a sub-dominant cost factor when designing the correlator.
-This strategy will enable fast read-out and arrays with many thousands of antennas, which will be necessary for future radio transient and cosmology experiments respectively.
+This strategy will enable fast read-out for arrays with many thousands of antennas, which will be necessary for future radio transient and cosmology experiments.
 
@@
 
-\textbf{Fitting gain models.} With some knowledge of the instrumental bandpass, the signal-to-noise of the gain solutions ({\bf you mean of the image?}) can be greatly improved by fitting a model to the per-frequency solutions derived here.
+\textbf{Multiple pixel correlations.} As was seen in section~\ref{sec:noise}, the single pixel correlation demonstrated in this paper can underperform in the presence of a complex sky, when compared to visibility-based gains with perfect knowledge of the sky.
+These errors can be mitigated by using correlations of multiple image pixels to incorporate a higher fraction of the total sky power into the calibration loop.
+Of course this increases the computational cost to a scaling of $\mathcal{O}(\Nant N_{\mathrm{pix}})$, which will typically be much lower than the $\mathcal{O}(\Ng \log_2 \Ng)$ of the correlator itself.
+These additional correlations would further enable direction dependent gain solutions.
+Each correlation could be used to independently solve for gains in each pixel direction, then fit to a beam model on the sky.
+This updated beam pattern would then feed into the gridding step of the correlator, allowing the imager to convolve the signals with the effective beam pattern in the ground plane.
+
+\textbf{Fitting gain models.} With some knowledge of the instrumental bandpass, the noise of the gain solutions can be greatly reduced by fitting a model to the per-frequency solutions derived here.
 We used this method at a rudimentary level in our demonstration to LWA data by assuming the gains were constant over a narrow bandwidth.
@@
 
-\textbf{Multiple pixel correlations.} As pointed out in section~\ref{sec:noise}, for a sky-noise dominated system, the signal to noise in the calibrator pixel dominates the error on the EPICal generated gain estimates.
-Besides the methods described above to increase signal to noise, one could also use correlations of multiple image pixels to incorporate a higher fraction of the total sky noise.
-Of course this increases the computational cost to a scaling of $\mathcal{O}(\Nant N_{\mathrm{pix}})$, which will typically be much lower than the $\mathcal{O}(\Ng \log_2 \Ng)$ of the correlator itself.
-
-\textbf{Direction dependent gains.} Even below the timescale of a dynamic ionosphere, a widefield instrument may require direction dependent gains to account for a varying thickness ionospheric sheet.
-The holographic mapping framework used by the MOFF algorithm, and in turn the EPIC and EPICal software can naturally account for direction dependence through an updated model of the antenna voltage pattern.
-This can be achieved by using multiple pixel correlations to independently solve for gains in each pixel direction.
-These solutions can then be used to fit a beam model on the sky (analogous to the gain model fitting mentioned above), and updating the antenna voltage pattern.
-This updated pattern would then feed into the gridding step of the correlator, allowing the imager to convolve the signals with the effective beam pattern in the ground plane.
-
 The work here will serve as a foundation for further development.
@@
 With next generation instruments in the planning and development stages, EPICal is poised to greatly contribute to their success.
+The software is integrated into the EPIC package and freely available to use in simulations or offline processing of data.
+Work is underway to port the code to real-time GPU systems for deployment.
 

### 2016-03-02 cal_demo/ms.tex
rewrote noise/error section

@@
 \usepackage{color}
+\usepackage{enumerate}
 
@@
 \newcommand{\itr}{20}
+\newcommand{\tcal}{\ensuremath{t_{\mathrm{cal}}}}
+\newcommand{\teff}{\ensuremath{t_{\mathrm{eff}}}}
 
@@
 
-\section{Noise trends}\label{sec:noise}
-Here we study the properties of our calibration solutions with respect to noise in the system, and compare to solutions found from visibilities.
-For this comparison we run a suite of simulations varying the receiver noise and integration times, while forming antenna-pixel correlations, $\Cna$, for EPICal gain solutions, and simultaneously forming visibilities from the same E-field streams to find visibility-based gain solutions.
+\section{Noise trends and error analysis}\label{sec:noise}
+Now we return to simulations to study the effect of noise in our system, and the consequences of an incomplete sky model.
+We run a suite of simulations varying the receiver noise and integration times, while forming antenna-pixel correlations, $\Cna$, for EPICal gain solutions, and simultaneously forming visibilities from the same E-field streams to find visibility-based gain solutions for comparison.
 
-We use a simulated sky consisting of a 5 Jy calibrator source, and 10 other random sources with apparent flux densities 0.2 to 0.5 Jy (total sky power $\approx$~8 Jy).
+We use a simulated sky consisting of a 5~Jy calibrator source, and 49 other random sources with apparent flux densities 0.2 to 0.5 Jy (total sky power $\approx$ 23 Jy).
 We generated these sources randomly, but kept them fixed for each run.
@@
 The level of the receiver noise, $\sigma_r = \left<\left|\widetilde{n}_a(f,t)\right|^2\right>$, is varied in different simulation runs.
+We include two limiting cases, where the receiver noise power is subdominant to the sky power ($\sigma_r = 10.0$~Jy), and where the receiver dominates the noise ($\sigma_r=100.0$~Jy).
 
-We initialize our gain estimates with the true values (${g^{(0)}_a=g^T_a=1}$), and allow the estimate to be corrupted by the noise through one iteration of the calibration loop.
-Ideally we would use a damping factor of zero to allow for the noise to fully corrupt the gains.
-However, we found that a zero damping factor resulted in non-convergent results ({\bf this statement makes me nervous and makes it sound these results may not be robust.
-It sounds like a higher $\damp$ is just delaying the divergence.}), so we adopted a small factor, $\damp=0.05$.
-For a fair comparison, we use the same factor for the visibility-based solutions.
-We then observe the error in the gain estimates introduced by the noise,
+For each simulation run we find gain solutions using four methods:
+\begin{enumerate}[i.]
+\item EPICal, using a full sky model
+\item EPICal, using a single point source model
+\item Visibility-based, using a full sky model
+\item Visibility-based, using a single point source model.
+\end{enumerate}
+In the cases where we use the full sky model, the model visibilities, $\V'_{ab}$ used in the calibration loop are created using all point sources in the simulated sky.
+For the single point source model, we only allow the calibration loop to be privy to the bright calibrator source.
+While we vary the model used for calibration, the simulated ``true sky'' remains fixed with all 50 point sources.
+
+For the EPICal calibration loops, we initialize our gain estimates with the true values (${g^{(0)}_a=g^T_a=1}$), and allow the estimate to be corrupted by the noise through ten iterations of the calibration loop.
+We adopt a damping factor $\damp=0.35$.
+Because EPICal updates the gain estimate at each calibration loop, \tcal, but retains a memory of previous iterations through the damping factor, the total integration time is not straightforward.
+We define an effective integration time by considering the relative weights of each previous $\Cna$ contributing to the current $g^{(n)}_a$ estimate, each with integration time \tcal.
+In the limit $n\rightarrow \infty$, the geometric series converges to an effective integration time of
+\begin{equation}
+\teff = \tcal \times \frac{1+\damp}{1-\damp}.
+\end{equation}
+With the damping factor we adopted here, $\teff \approx 2.1 \times \tcal$.
+
+We simultaneously form simulated visibilities by correlating all pairs of antenna E-field measurements (including the receiver noise).
+The E-fields are correlated for a duration equal to the effective integration time of the EPICal loop for comparison.
+We then find the visibility-based gain estimates by minimizing
+\begin{equation}\label{eq:vis_cal}
+\chi^2 = \sum_a\sum_{b\ne a} \left|\V_{ab}-g_a g_b^* \V'_{ab}\right|^2,
+\end{equation}
+using both versions of the model visibilities described above.
+
+ For each version of calibration, we observe the error in the gain estimates averaging over both antennas and frequency channels,
 \begin{equation}\label{eq:gain_error}
@@
 \end{equation}
-({\bf Here is my major concern.
-I feel these statistics should be estimated after making sure gains have converged after few loops, otherwise these statistics may not be accurate.
-Since you do not form images, I don't think it should take very long.})
 
-We simultaneously form simulated visibilities by correlating all pairs of antenna E-field measurements (including the receiver noise).
-At the end of the calibration loop we find the visibility-based gain estimates by minimizing
-\begin{equation}\label{eq:vis_cal}
-\chi^2 = \sum_a\sum_{b\ne a} \left|\V_{ab}-g_a g_b^* \V^T_{ab}\right|^2.
-\end{equation}
-The error of the visibility-based gains is estimated in the same way as in equation~\ref{eq:gain_error}.
-({\bf How is $\widetilde{V}_{ab}$ defined?
-As the average of all visibilities between that antenna pair in the calibration loop timescale?
-Or are the gains determined on every timestamp using linear algebraic minimization?})
-
-The results of our simulations are shown in figure~\ref{fig:noise}.
-The gain errors from our EPICal method is shown with solid lines, each color representing different levels of receiver noise.
-Each data point ({\bf I see no discrete data points.
-Can you add those to the plot?}) represents a separate simulation run (independent realization of the noise).
-The general trends are consistent with expectations -- decreasing as the square root of integration time, and increasing with receiver noise.
-The lowest receiver noise level (1.0 Jy) is significantly below the sky noise contribution (about 8 Jy), so the black line is not as low, compared to the red, as one would expect from receiver noise consideration alone ({\bf Need to elaborate.
-It is too terse, I couldn't follow.
-}).
+The results of our simulations are shown in figure~\ref{fig:errors}.
+The gain errors from our EPICal method is shown with solid lines, while the dashed lines represent the visibility-based calibration.
+The thick lines represent the results of using a full sky model, and the thin lines are from the single point source model.
+In the case of the full sky model, we see the errors trend downward with longer integration times, as expected.
+The EPICal errors are slightly higher than the visibility-based errors.
+This is not surprising as EPICal only uses the information in a single pixel, while the visibilities use all information from the full sky model.
+However, we see that with this perfect model the errors do behave like noise, and the same level of noise can be achieved with a longer integration time (or larger damping factor).
 
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{figures/cal_paper_noise_med_sky.pdf}
-\caption{Gain estimate errors as a function of integration time and receiver noise (line color).
+\includegraphics[width=\columnwidth]{figures/cal_paper_errors_complex_sky.pdf}
+\caption{Gain estimate errors as a function of integration time, receiver noise (line color), and calibration method.
 The solid lines represent the error on the EPICal derived gain estimates, while the dashed lines represent the error on the visibility based estimates.
-On average the EPICal error is 1.5\% higher than visibilities, and at worst it was 3.5\% higher.
+The thicker lines were derived using a full sky model in the calibration loops, and the thinner lines used only a single point source model.
+In the full sky model case we see all methods trend down with longer integration time as expected.
+In the case of a single source model, the errors bottom out due to the flux in the measurements that is not modeled.
+In this regime the EPICal method performs slightly better than the visibility-based method due to effectively beamforming to the sky pixel where the incomplete model is most accurate.
 }
-\label{fig:noise}
+\label{fig:errors}
 \end{center}
@@
 
-The dashed lines in figure~\ref{fig:noise} represent the errors from the visibility-based calibration.
-Each data point corresponds to the same simulation run as the corresponding EPICal point (same realization of noise).
-Generally we see the visibility-based calibration achieves slightly lower noise solutions.
-On average the EPICal errors are 1.5\% higher, and at the worst is 3.5\% higher ({\bf but need to also explain this trend reverses and dependent on receiver noise and integration time}).
-
-We repeated this process for two extremes -- a simple sky model of a single point source, and a more complicated sky with 50 sources.
-In the simple sky case, we found the EPICal and visibility gain estimates to be identical to floating point precision, as expected from our discussion in section~\ref{sec:math}.
-In the more complicated sky case (about 23 Jy total sky power), we saw the EPICal gains exhibit errors 20\% higher than the visibility derived gains.
-The trend of higher EPICal errors with more complicated skies is due to the single correlated pixel not containing the full signal to noise of the data ({\bf Here is my other major concern.
-There is no sort of numerical substantiation offered to confirm this other than saying we have not done this yet and thus missing information.}).
-This reaffirms our intuition that using a pixel with a bright source is ideal, but will still not match visibilities for complicated skies ({\bf to reaffirm we need some sort of numerical evidence that this is at play and we will work on this further}).
-We suggest additional strategies for improving this signal to noise in the following section.
-({\bf At the moment it sounds like we are conveniently sweeping it under the rug.
-Can we at least qualitatively make arguments that this should improve and what might that information from multiple pixels look like mathematically?})
-
-%\section{Noise analysis}\label{sec:noise_old}
-%\textcolor{red}{This section will most likely disappear in favor of the Noise Trends section.}
-%
-%Here we study the properties of our calibration solutions with respect to noise in the system.
-In order to compare to solutions from visibility-based calibration, we first derive the theoretical lower bound of the error in these solutions, given a set of measured visibilities, $\mathbf{V}$.
-Motivated by the performance analysis seen in \cite{kaz13}, we will do this through the Fischer information formalism and employ the Cram\'er-Rao lower bound.
-%
-%Assume we did form visibilities with some integration time.
-Assuming the noise on each visibility, $\sigma_{ab}$, is independent, we can write the likelihood function of measuring $V_{ab}$ given the true value, the gains, and the noise.
-%\begin{equation}
-%\mathcal{L}(V_{ab};\mathbf{g}) = \frac{1}{2\pi \sigma_{ab}^2}\exp\left[-\frac{\left|V_{ab} - g_a g_b^* V_{ab}^T\right|^2}{2\sigma_{ab}^2}\right]
-%\end{equation}
-%Then the likelihood of the set of all visibilities is the product of all individual likelihoods.
-%\begin{equation}
-%\mathcal{L}(\mathbf{V};\mathbf{g}) = \prod_a \prod_{b > a} \mathcal{L}(V_{ab};\mathbf{g})
-%\end{equation}
-%The Fischer information matrix for the set of gain parameters is
-%\begin{equation}
-%\mathbf{F}^g_{ij} = \left<\left. \frac{\partial \ln \mathcal{L}(\mathbf{V};\mathbf{g})}{\partial g_i} \frac{\partial \ln \mathcal{L}(\mathbf{V};\mathbf{g})}{\partial g_j}\right|_\mathbf{g}\right>,
-%\end{equation}
-%where the expectation is taken over all possible visibility measurements.
-%We next evaluate the derivate of the log-likelihood.
-%\begin{equation}
-%\frac{\partial \ln \mathcal{L}(\mathbf{V};\mathbf{g})}{\partial g_i} = 
-%\sum_{a \ne i} \frac{g_a^* V_{ai}^{T*} (V_{ai} - g_a g_i^* V_{ai}^T)}{2 \sigma_{ai}^2}
-%\end{equation}
-%
-%In order to find the Cram\'er-Rao lower bound on the variance of the complex parameter, $g_i$, we consider the term of the Fischer matrix where the first derivative is taken with respect to $g_i$, and the second with $g_i^*$.
-The result is
-%\begin{align}
-%\mathbf{F}^g_{ii} = & \left<\left[ \sum_{a \ne i} \frac{g_a^* V_{ai}^{T*} (V_{ai} - g_a g_i^* V_{ai}^T)}{2 \sigma_{ai}^2} \right] \right. \times \nonumber \\
-%& \left. \left[ \sum_{b \ne i} \frac{g_b V_{bi}^{T} (V_{bi}^* - g_b^* g_i V_{bi}^{T*})}{2 \sigma_{bi}^2} \right] \right> \nonumber \\
-%= &\sum_{a\ne i} \sum_{b \ne i} \frac{g_a^* g_b V_{ai}^{T*} V_{bi}^T}{4 \sigma_{ai}^2 \sigma_{bi}^2} \left< V_{ai}V_{bi}^* - g_i g_b^* V_{ai}V_{bi}^{T*}\nonumber \right.\\
-%&\left. - g_a g_i^* V_{ai}^T V_{bi}^* + g_a g_b^* |g_i|^2 V_{ai}^T V_{bi}^{T*}\right>
-%\label{eq:Fischer_before_expect}
-%\end{align}
-%
-%The expected values are easy to evaluate.
-Each visibility will average to the ``true" value times the respective gains.
-The term with two visibilities will include a noise bias on autocorrelation terms.
-%\begin{equation}
-%\left<V_{ai}V_{bi}^*\right> = |g_i|^2 g_a g_b^* V_{ai}^T V_{bi}^{T*} + \sigma_{ai}^2 \delta_{ab}
-%\end{equation}
-%Here $\delta_{ab}$ is the Kronecker delta selecting the term where $a=b$ and the noise correlates. Plugging in the expectation values, equation \ref{eq:Fischer_before_expect} simplifies greatly to
-%\begin{equation}
-%\mathbf{F}^g_{ii} = \sum_{a \ne i} \frac{\left|g_a V_{ia}^T \right|^2}{4\sigma_{ai}^2}
-%\end{equation}
-%
-%Finally we relate our result to the theoretical best uncertainty we can place on our unknown gain parameter using the Cram\'er-Rao lower bound \cite{kay93}.
-%\begin{equation}\label{eq:cramer_rao}
-%\sigma_{g_i}^2 \ge \left[ \mathbf{F}^g_{ii} \right]^{-1} = \left[ \sum_{a \ne i} \frac{\left|g_a V_{ia}^T \right|^2}{4\sigma_{ai}^2} \right]^{-1}
-%\end{equation}
-%
-%We next run a set of EPICal simulations to compare the noisy results to the theoretical lower bound if we had formed an analogous set of visibilities.
-Each simulation is run with a simple sky of a single point source with flux 1 Jy.
-We add noise to the system by adding a Gaussian-random complex value to each channelized electric field antenna measurement as in equation~\ref{eq:apply_gain}.
-The ``true" gains in our simulation are unity, as well as our initial estimate.
-This emulates a system where the initially volatile gain correction has been suppressed and only noise remains to corrupt our estimation, allowing us to run many simulations on a shorter timescale.
-%
-%We vary two parameters in our simulations: the variance of the complex random noise added to our electric field measurements ($\sigma_0$), and the length of integration in each calibration loop, $t_{\mathrm{cal}}$.
-The equivalent visibility variance in equation~\ref{eq:cramer_rao} is given by
-%\begin{equation}
-%\sigma_{ai}^2 = \frac{\sigma_0^2}{B t_{\mathrm{eff}}},
-%\end{equation}
-%where $B$ is the bandwidth of the simulation, and $t_{\mathrm{eff}}$ is the effective integration time based on the length of our calibration loop and the damping factor.
-%\begin{equation}
-%t_{\mathrm{eff}} = t_{\mathrm{cal}} \times \frac{1+\damp}{1-\damp}
-%\end{equation}
-%
-%We demonstrate the trend of the gain errors with respect to these parameters in figure~\ref{fig:noise}.
-For each $\sigma_0$ and $t_{\mathrm{cal}}$ we run the simulation for seven iterations of the calibration loop to allow the noise to fully corrupt our estimates.
-We then compute the variance of the resulting gains across the antennas.
-Each color line represents a different amount of input noise, with the dashed line showing the measured variance and the solid line showing the theoretical lower limit from equation~\ref{eq:cramer_rao}.
-The agreement between the dashed and solid lines demonstrates that not only do we reach an uncertainty consistent with visibility-based solutions, but that we actually reach the theoretical lower limit based on the amount of information contained in an equivalent set of visibility measurements.
-%
-%
+When using the single point source model, all calibration methods trend downward until they reach an error floor due to the confusion sources that were not modeled.
+The EPICal and visibility-based solutions reach a similar floor, but EPICal achieves a marginally lower level.
+This can be attributed to the same reason EPICal underperformed with the full sky model.
+EPICal only used a single pixel in the sky to form its solutions.
+When using the full sky model, it was down weighting the information in the rest of the sky.
+But when the sky model only contains the bright point source, the single pixel used in EPICal is the location where this model is most accurate, effectively down weighting the pixels with incomplete sky model.
 
@@
 Of course this increases the computational cost to a scaling of $\mathcal{O}(\Nant N_{\mathrm{pix}})$, which will typically be much lower than the $\mathcal{O}(\Ng \log_2 \Ng)$ of the correlator itself.
-These additional correlations would further enable direction dependent gain solutions.
+These additional correlations would also enable direction dependent gain solutions.
 Each correlation could be used to independently solve for gains in each pixel direction, then fit to a beam model on the sky.

### 2016-03-02 cal_demo/ms.tex
added reference to PyGSM

@@
 We also note that the feature just to the right of Cyg~A is dimmer in the calibrated image, better matching the expected flux from the global sky model (GSM, \citealt{deo08}).
-We show the GSM modulated by two factors of the LWA power pattern in the right panel of figure~\ref{fig:data_images} for reference.
+We show the GSM\footnote{using the PyGSM module by D.~C.~Price to import the GSM, \textcolor{red}{insert reference}} modulated by two factors of the LWA power pattern in the right panel of figure~\ref{fig:data_images} for reference.
 

### 2016-03-02 cal_demo/ms.tex
more tweaks to cal paper, improved figure, updated reference

@@
 We then demonstrate the algorithm in simulations in \S \ref{sec:sim}, and apply to a sample LWA data set in \S \ref{sec:data}.
-Then we discuss the noise properties of the resulting gain solutions in \S \ref{sec:noise}.
+Then we discuss errors and noise properties of the resulting gain solutions in \S \ref{sec:noise}.
 Finally we conclude and discuss potential extensions to the algorithm in \S \ref{sec:discussion}.
@@
 The middle panel shows the image produced after calibration with identical integration time and bandwidth.
-The sidelobes ({\bf or is it better to call them speckles due to uncalibrated phases because sidelobes are typically associated with the psf and not with calibration?}) throughout the image are significantly suppressed, and the galactic plane is much more evident, despite only modeling Cyg~A and Cas~A.
+The sidelobes throughout the image are significantly suppressed, and the galactic plane is much more evident, despite only modeling Cyg~A and Cas~A.
 We also note that the feature just to the right of Cyg~A is dimmer in the calibrated image, better matching the expected flux from the global sky model (GSM, \citealt{deo08}).
@@
 
- For each version of calibration, we observe the error in the gain estimates averaging over both antennas and frequency channels,
+ For each version of calibration, we observe the error in the gain estimates averaging over both antennas and frequency channels.
 \begin{equation}\label{eq:gain_error}
-\sigma_g = \left[\frac{1}{N_f \Nant} \sum_f \sum_a \frac{\left|g'_a(f)-g^T_a(f)\right|^2}{\left|g^T_a(f)\right|^2}\right]^{1/2}.
+\sigma_g = \left[\frac{1}{N_f \Nant} \sum_f \sum_a \frac{\left|g'_a(f)-g^T_a(f)\right|^2}{\left|g^T_a(f)\right|^2}\right]^{1/2}
 \end{equation}
@@
 The EPICal and visibility-based solutions reach a similar floor, but EPICal achieves a marginally lower level.
-This can be attributed to the same reason EPICal underperformed with the full sky model.
-EPICal only used a single pixel in the sky to form its solutions.
+This can be attributed to the same reason EPICal underperformed with the full sky model: EPICal only used a single pixel in the sky to form its solutions.
 When using the full sky model, it was down weighting the information in the rest of the sky.
 But when the sky model only contains the bright point source, the single pixel used in EPICal is the location where this model is most accurate, effectively down weighting the pixels with incomplete sky model.
+
+Any realistic sky model will lie somewhere between the two extremes explored here.
+In the case where the sky is modeled by a single source, as is often the case for an initial calibration, EPICal actually achieves smaller gain errors compared to the traditional visibility-based calibration.
+As the sky model improves, both visibility-based and EPICal gains improve, though the former quicker than the latter.
+In the limit of a perfect model, both calibration methods produce noise-like errors in their gains which scale down with more integration time.
 

### 2016-03-04 cal_demo/ms.tex
touching up cal paper

@@
 These coefficients are then shipped off-chip to apply equations~\ref{eq:cal_solution}~and~\ref{eq:damping}, and any other processing to arrive at an updated estimate of the gains, $g^{(n)}_a$.
-The gray box shows operations which must be done at high speed (before averaging in time), which white boxes show operations which can be performed ``off-chip".
-({\bf I have some suggestions to tweak this figure.})}
+The gray box shows operations which must be done at high speed (before averaging in time), which white boxes show operations which can be performed ``off-chip".}
 \label{fig:schematic}
@@
 We also note that the feature just to the right of Cyg~A is dimmer in the calibrated image, better matching the expected flux from the global sky model (GSM, \citealt{deo08}).
-We show the GSM\footnote{using the PyGSM module by D.~C.~Price to import the GSM, \textcolor{red}{insert reference}} modulated by two factors of the LWA power pattern in the right panel of figure~\ref{fig:data_images} for reference.
+We show the GSM\footnote{We use the PyGSM module by D.~C.~Price to import the GSM, https://github.com/telegraphic/PyGSM} modulated by two factors of the LWA power pattern in the right panel of figure~\ref{fig:data_images} for reference.
 
@@
 In the case of the full sky model, we see the errors trend downward with longer integration times, as expected.
-The EPICal errors are slightly higher than the visibility-based errors.
+The EPICal errors are slightly higher than the visibility-based errors, on average about 23\% difference.
 This is not surprising as EPICal only uses the information in a single pixel, while the visibilities use all information from the full sky model.
@@
 In the full sky model case we see all methods trend down with longer integration time as expected.
+For a given integration time, EPICal derived errors are on average 23\% higher than those of visibility based calibration.
 In the case of a single source model, the errors bottom out due to the flux in the measurements that is not modeled.
 In this regime the EPICal method performs slightly better than the visibility-based method due to effectively beamforming to the sky pixel where the incomplete model is most accurate.
+On average EPICal errors are 5\% lower in this regime.
 }
@@
 When using the single point source model, all calibration methods trend downward until they reach an error floor due to the confusion sources that were not modeled.
-The EPICal and visibility-based solutions reach a similar floor, but EPICal achieves a marginally lower level.
+The EPICal and visibility-based solutions reach a similar floor, but EPICal achieves a marginally lower level (on average 5\% lower).
 This can be attributed to the same reason EPICal underperformed with the full sky model: EPICal only used a single pixel in the sky to form its solutions.
@@
 
-There remain many avenues of future work before EPICal can be deployed on an instrument.
+EPICal can be further improved through several extensions.
 Here we name a few potential considerations for further study.

### 2016-03-07 cal_demo/ms.tex
Adding comments on other calibration regimes, adding hera memo 10 to bib

@@
 
-When using the single point source model, all calibration methods trend downward until they reach an error floor due to the confusion sources that were not modeled.
+The exact ratio of EPICal gain noise to visibility-based gain noise depends on the fraction of the total sky power contained in our calibrator source.
+In the example here, our calibrator accounted for 22\% of the total sky power.
+We repeated the experiment in the regime where the calibrator dominated the sky and found that the difference in EPICal and visibility-based error goes to zero, as expected.
+For a typical HERA observation, the sky temperature is expected to by about 180~K \citep{jac15}, and a bright calibrator source could be about 10\% of this power.
+In this regime we found that EPICal gain noise was about 60\% higher than visibility-based gain noise for the same effective integration time.
+
+When using the single point source model, all calibration methods in figure~\ref{fig:errors} trend downward until they reach an error floor due to the confusion sources that were not modeled.
 The EPICal and visibility-based solutions reach a similar floor, but EPICal achieves a marginally lower level (on average 5\% lower).
@@
 But when the sky model only contains the bright point source, the single pixel used in EPICal is the location where this model is most accurate, effectively down weighting the pixels with incomplete sky model.
+In the regime typical of HERA discussed above, EPICal's error floor was 10\% lower than that of visibility-based solutions.
 

### 2016-03-07 cal_demo/ms.tex
cleaning up unused text/code
Changes to make arxiv submission work

@@
 % Basic setup. Most papers should leave these options alone.
-\documentclass[a4paper,fleqn,usenatbib]{../mnras}
+\documentclass[a4paper,fleqn,usenatbib]{mnras}
 
-\usepackage{newtxtext,newtxmath}
+%\usepackage{newtxtext,newtxmath}
 
@@
 % Keep the title short and informative.
-\title[EPICal]{An efficient feedback calibration algorithm for direct imaging radio telescopes
+\title[EPICal]{An Efficient Feedback Calibration Algorithm for Direct Imaging Radio Telescopes
 %The E-field Parallel Imaging Calibration Algorithm for Next-Generation Radio Telescopes
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{figures/MOFF_flowchart.pdf}
+\includegraphics[width=\columnwidth]{MOFF_flowchart.pdf}
 \caption{Data flow for the MOFF algorithm, reproduced from \citealt{thy15c}.
@@
 
-%%%%%%
-%%%%%%
-
-%We begin by establishing the mathematical framework for the calibration problem.
-We derive the calibration solutions for the MOFF algorithm (adopting the notation of \citealt{thy15c}), but note the result is easily extended to FFT correlator algorithms by removing the gridding step.
-%
-%The electric field incident on the ground, $\Er{}(\mathbf{r},f,t)$, is related to the sky electric field, $E(\hat{\mathbf{s}},f,t)$, through a Fourier transform.
-%\begin{equation}
-%\Er{}(\mathbf{r},f,t) = \int E(\hat{\mathbf{s}},f,t) e^{-2\pi i \mathbf{r}\cdot \hat{\mathbf{s}}}\, \dif^2 \hat{\mathbf{s}}
-%\end{equation}
-%Here $\hat{\mathbf{s}}$ denotes the sine-projected unit vector for the sky angle, $\mathbf{r}$ is the observer's location (measured in wavelengths relative to an arbitrary origin), and $f$ and $t$ denote the frequency and time dependence, respectively.
-We will encounter several quantities which we attempt to estimate.
-We distinguish the ``true" values with a superscript $T$, while the estimates are denoted with a prime.
-We define the true antenna signal as a convolution of the antenna voltage pattern, $\beamr$, with the electric field on the ground.
-%\begin{equation}
-%\Er{a}^T(f,t) \equiv \int \beamr_a(\mathbf{r}-\ra) \Er{}(\mathbf{r},f,t) \, \dif^2 \mathbf{r}
-%\end{equation}
-%The subscript $a$ labels the antenna, and $\mathbf{r}_a$ is the location of antenna $a$.
-%
-%We next model the measured, uncalibrated electric field as a multiplicative complex gain and an additive noise term applied to the true antenna electric field. 
-%\begin{equation}\label{eq:apply_gain}
-%\Er{a}(f,t) = g^T_a(f,t) \Er{a}^T(f,t) + \widetilde{n}_a(f,t)
-%\end{equation}
-%Note that this quantity is neither a true or estimated value, but rather the raw electric field signal corrupted by the complex gain before we attempt to estimate the true value.
-The noise term is strictly receiver noise -- noise introduced by the instrument.
-Any sky noise is implicitly included in the time dependence of the sky electric field.
-As the noise of modern low frequency arrays is heavily dominated by sky noise, we will neglect $\widetilde{n}_a$ for now, but will inspect its effects at the end of this section.
-%
-%The goal of our calibration algorithm will be to estimate the antenna gains, $g_a$.
-We will assume the gains have no time dependence within the timescale of finding our solutions.
-Furthermore, we will treat each frequency channel independently, and drop the $f$ to simplify notation.
-%
-%The MOFF algorithm next calls for a calibration.
-We will assume we have formed an estimate of the gains after $n$ iterations of a calibration loop.
-We then correct the incoming electric field data stream using this current estimate.
-%\begin{equation}
-%\Erest{a} = \Er{a}/ g^{(n)}_a,
-%\end{equation}
-%These corrected fields will be used in the EPIC pipeline to derive an updated estimate for the ($n+1$)$^{\mathrm{th}}$ iteration. 
-%
-%A dirty image is next formed by gridding the calibrated fields with the antenna voltage pattern, Fourier transforming, squaring, and averaging in time.
-The estimated value for a pixel, $\s_i$, can be expressed as
-%\begin{equation}
-%I'(\s_i) = \left<\left| \frac{1}{\Nant}\sum_i e^{2\pi i \mathbf{r}\cdot \s_i} \sum_a \beamr_a(\ri - \ra) \frac{g^T_a}{g^{(n)}_a} \Er{a}^T(t) \right|^2 \right>_t.
-%\end{equation}
-%This is the final output of the MOFF correlator. However, for calibration purposes we will be interested in the electric field image just prior to squaring and averaging.
-%\begin{equation}
-%\Ethetaest (\s_i,t) = \frac{1}{\Nant} \sum_i e^{2\pi i \ri \cdot \s_i} \sum_a \beamr_a(\ri - \ra) \frac{g^T_a}{g^{(n)}_a} \Er{a}^T(t)
-%\end{equation}
-%We can simplify this expression by exchanging the sums to transform the beam term into sky coordinates.
-%\begin{align}\label{eq:epix}
-%\Ethetaest(\s_i,t) & = \frac{1}{\Nant} \sum_a \frac{g^T_a}{g^{(n)}_a}\Er{a}^T(t) e^{2\pi i \s_i \cdot \ra} \nonumber \\
-%  & \qquad\times \sum_i \beamr_b(\ri-\ra)e^{2\pi i \s_i \cdot (\ri-\ra)} \nonumber\\
-%& = \frac{1}{\Nant} \sum_a \frac{ g^T_a}{g^{(n)}_a}\Er{a}^T(t) e^{2\pi i \s_i \cdot \ra}\beamtheta_a(\s_i)
-%\end{align}
-
-%%%%
-%%%%
 As a starting point, we consider the feedback calibration outlined in \citealt{mor11}.
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{figures/schematic.pdf}
+\includegraphics[width=\columnwidth]{schematic.pdf}
 \caption{The general data flow of the MOFF correlator, with a feedback calibration loop.
@@
 
-%%%%%
-%%%%%
-
 \section{Simulation}\label{sec:sim}
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{figures/cal_paper_sim_phase.pdf}
+\includegraphics[width=\columnwidth]{cal_paper_sim_phase.pdf}
 \caption{Phase error of gain estimates as a function of iteration for simulated calibration.
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{figures/cal_paper_sim_amps.pdf}
+\includegraphics[width=\columnwidth]{cal_paper_sim_amps.pdf}
 \caption{Gain estimate amplitudes as a function of iteration for simulated calibration.
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{figures/cal_paper_sim_image_before.pdf}
-\includegraphics[width=\columnwidth]{figures/cal_paper_sim_image_after.pdf}
+\includegraphics[width=\columnwidth]{cal_paper_sim_image_before.pdf}
+\includegraphics[width=\columnwidth]{cal_paper_sim_image_after.pdf}
 \caption{Images formed during simulated calibration.
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{figures/cal_paper_data_phases.pdf}
+\includegraphics[width=\columnwidth]{cal_paper_data_phases.pdf}
 \caption{Gain phase solutions as a function of calibration iteration for an LWA TBN observation.
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{figures/cal_paper_data_amps.pdf}
+\includegraphics[width=\columnwidth]{cal_paper_data_amps.pdf}
 \caption{Gain amplitudes as a function of calibration iteration for an LWA TBN observation.
@@
 \begin{center}
-\includegraphics[width=0.33\linewidth]{figures/cal_paper_data_image_before.pdf}
-\includegraphics[width=0.33\linewidth]{figures/cal_paper_data_image_after.pdf}
-\includegraphics[width=0.33\linewidth]{figures/cal_paper_gsm_beam_weighted.pdf}
+\includegraphics[width=0.3\linewidth]{cal_paper_data_image_before.pdf}
+\includegraphics[width=0.3\linewidth]{cal_paper_data_image_after.pdf}
+\includegraphics[width=0.3\linewidth]{cal_paper_gsm_beam_weighted.pdf}
 \caption{Images produced before (left) and after (middle) calibrating LWA data.
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{figures/cal_paper_errors_complex_sky.pdf}
+\includegraphics[width=\columnwidth]{cal_paper_errors_complex_sky.pdf}
 \caption{Gain estimate errors as a function of integration time, receiver noise (line color), and calibration method.
@@
 
-\bibliographystyle{../mnras}
-\bibliography{../epic} % if your bibtex file is called example.bib
-
-%%%%%%%%%%%%%%%%% APPENDICES %%%%%%%%%%%%%%%%%%%%%
-
-%\appendix
-
-%\section{Some extra material}
-
-%If you want to present additional material which would interrupt the flow of the main paper,
-%it can be placed in an Appendix which appears after the list of references.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\bibliographystyle{mnras}
+%\bibliography{epic} % if your bibtex file is called example.bib
+\input{ms.bbl}
 

### 2016-03-07 cal_demo/ms.tex
trying to fix branching issues
moving figures for mnras

@@
 \begin{center}
-\includegraphics[width=\columnwidth]{MOFF_flowchart.pdf}
+\includegraphics[width=\columnwidth]{fig1.pdf}
 \caption{Data flow for the MOFF algorithm, reproduced from \citealt{thy15c}.
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{schematic.pdf}
+\includegraphics[width=\columnwidth]{fig2.pdf}
 \caption{The general data flow of the MOFF correlator, with a feedback calibration loop.
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{cal_paper_sim_phase.pdf}
+\includegraphics[width=\columnwidth]{fig3.pdf}
 \caption{Phase error of gain estimates as a function of iteration for simulated calibration.
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{cal_paper_sim_amps.pdf}
+\includegraphics[width=\columnwidth]{fig4.pdf}
 \caption{Gain estimate amplitudes as a function of iteration for simulated calibration.
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{cal_paper_sim_image_before.pdf}
-\includegraphics[width=\columnwidth]{cal_paper_sim_image_after.pdf}
+\includegraphics[width=\columnwidth]{fig5a.pdf}
+\includegraphics[width=\columnwidth]{fig5b.pdf}
 \caption{Images formed during simulated calibration.
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{cal_paper_data_phases.pdf}
+\includegraphics[width=\columnwidth]{fig6.pdf}
 \caption{Gain phase solutions as a function of calibration iteration for an LWA TBN observation.
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{cal_paper_data_amps.pdf}
+\includegraphics[width=\columnwidth]{fig7.pdf}
 \caption{Gain amplitudes as a function of calibration iteration for an LWA TBN observation.
@@
 \begin{center}
-\includegraphics[width=0.3\linewidth]{cal_paper_data_image_before.pdf}
-\includegraphics[width=0.3\linewidth]{cal_paper_data_image_after.pdf}
-\includegraphics[width=0.3\linewidth]{cal_paper_gsm_beam_weighted.pdf}
+\includegraphics[width=0.3\linewidth]{fig8a.pdf}
+\includegraphics[width=0.3\linewidth]{fig8b.pdf}
+\includegraphics[width=0.3\linewidth]{fig8c.pdf}
 \caption{Images produced before (left) and after (middle) calibrating LWA data.
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{cal_paper_errors_complex_sky.pdf}
+\includegraphics[width=\columnwidth]{fig9.pdf}
 \caption{Gain estimate errors as a function of integration time, receiver noise (line color), and calibration method.
@@
 \bibliographystyle{mnras}
-%\bibliography{epic} % if your bibtex file is called example.bib
+%\bibliography{../epic} % if your bibtex file is called example.bib
 \input{ms.bbl}

### 2016-03-30 cal_demo/ms.tex
adding danny prices pyGSM to citations

--- 2016-03-07 cal_demo/ms.tex
@@
 We also note that the feature just to the right of Cyg~A is dimmer in the calibrated image, better matching the expected flux from the global sky model (GSM, \citealt{deo08}).
-We show the GSM\footnote{We use the PyGSM module by D.~C.~Price to import the GSM, https://github.com/telegraphic/PyGSM} modulated by two factors of the LWA power pattern in the right panel of figure~\ref{fig:data_images} for reference.
+We show the GSM \citep{pri16} modulated by two factors of the LWA power pattern in the right panel of figure~\ref{fig:data_images} for reference.
 
@@
 
-\bibliographystyle{mnras}
-%\bibliography{epic} % if your bibtex file is called example.bib
-\input{ms.bbl}
+\bibliographystyle{../mnras}
+\bibliography{../epic} % if your bibtex file is called example.bib
+%\input{ms.bbl}
 

### 2016-04-14 cal_demo/ms.tex

--- 2016-03-07 cal_demo/ms.tex
@@
 
-\bibliographystyle{mnras}
-%\bibliography{../epic} % if your bibtex file is called example.bib
-\input{ms.bbl}
+\bibliographystyle{../mnras}
+\bibliography{../epic} % if your bibtex file is called example.bib
+%\input{ms.bbl}
 

### 2016-04-14 cal_demo/ms.tex
Breaking up lines in cal_demo paper for easier tracking of changes.

@@
 \begin{abstract}
-We present the E-field Parallel Imaging Calibration (EPICal) algorithm, which addresses the need for a real-time calibration method for direct imaging radio astronomy correlators.
-Direct imaging involves a spatial fast Fourier transform of antenna voltages, alleviating the harsh $\mathcal{O}(\Nant^2)$ computational scaling to a more gentle $\mathcal{O}(\Nant \log_2 \Nant)$, which can save orders of magnitude in computation cost for next generation arrays consisting of hundreds to thousands of antennas.
-However, because signals are mixed in the correlator, gain correction must be applied on the front end.
-We develop the EPICal algorithm to form gain solutions in real time without ever forming visibilities.
-This method scales as the number of antennas, and produces results comparable to those from visibilities.
-Through simulations and application to Long Wavelength Array data we show this algorithm is a promising solution for next generation instruments.
+We present the E-field Parallel Imaging Calibration (EPICal) algorithm, which addresses the 
+need for a real-time calibration method for direct imaging radio astronomy correlators. Direct 
+imaging involves a spatial fast Fourier transform of antenna voltages, alleviating the harsh 
+$\mathcal{O}(\Nant^2)$ computational scaling to a more gentle $\mathcal{O}(\Nant \log_2 
+\Nant)$, which can save orders of magnitude in computation cost for next generation arrays 
+consisting of hundreds to thousands of antennas. However, because signals are mixed in the 
+correlator, gain correction must be applied on the front end. We develop the EPICal algorithm 
+to form gain solutions in real time without ever forming visibilities. This method scales as the 
+number of antennas, and produces results comparable to those from visibilities. Through 
+simulations and application to Long Wavelength Array data we show this algorithm is a 
+promising solution for next generation instruments.
 \end{abstract}
@@
 \section{Introduction}
-In order to satisfy the survey speeds required for precision cosmology as well as searches for fast radio transients, radio astronomy is undergoing a paradigm shift toward interferometers consisting of hundreds to thousands of small, widefield antennas.
-Many arrays with this design are already built or under construction including the Hydrogen Epoch of Reionization Array\footnote{http://reionization.org} (HERA), the Murchison Widefield Array (MWA; \citealt{tin13,bow13}), the Donald C.
-Backer Precision Array for Probing the Epoch of Reionization (PAPER; \citealt{par10}), the LOw Frequency ARray (LOFAR; \citealt{van13}), the Canadian Hydrogen Intensity Mapping Experiment (CHIME,\citealt{ban14}), the Long Wavelength Array (LWA, \citealt{ell13}), and the low frequency Square Kilometer Array (SKA1-Low \citealt{mel13}).
+In order to satisfy the survey speeds required for precision cosmology as well as searches for 
+fast radio transients, radio astronomy is undergoing a paradigm shift toward interferometers 
+consisting of hundreds to thousands of small, widefield antennas. Many arrays with this design 
+are already built or under construction including the Hydrogen Epoch of Reionization Array
+\footnote{http://reionization.org} (HERA), the Murchison Widefield Array (MWA; 
+\citealt{tin13,bow13}), the Donald C. Backer Precision Array for Probing the Epoch of 
+Reionization (PAPER; \citealt{par10}), the LOw Frequency ARray (LOFAR; \citealt{van13}), the 
+Canadian Hydrogen Intensity Mapping Experiment (CHIME,\citealt{ban14}), the Long 
+Wavelength Array (LWA, \citealt{ell13}), and the low frequency Square Kilometer Array (SKA1-
+Low \citealt{mel13}).
 
-Traditional radio correlators cross-multiply the voltage signals from all pairs of antennas, and the computation scales as the number of antennas squared, $\mathcal{O}(\Nant^2)$ \citep{bun04}.
-As the number of elements in future arrays grows, the computational cost will become prohibitively expensive, and exploring efficient correlator schemes is essential to enable next generation instruments \citep{lon00}.
-Meanwhile, radio transient monitoring requires access to high time and frequency resolution data to identify and characterize events such as fast radio bursts (FRBs, \citealt{lor07}), or to follow up gravitational wave candidates with radio observations \citep{abb16a,abb16b}.
-FRBs are highly unexplored at low frequencies (< 1 GHz), but are expected to occur on timescales $\Delta t \sim$ 1--10~ms \citep{tho13}.
-Recording the full visibility matrix for $\Nant \gtrsim 10^3$ arrays at this timescale leads to extremely high data write rates.
+Traditional radio correlators cross-multiply the voltage signals from all pairs of antennas, and 
+the computation scales as the number of antennas squared, $\mathcal{O}(\Nant^2)$ 
+\citep{bun04}. As the number of elements in future arrays grows, the computational cost will 
+become prohibitively expensive, and exploring efficient correlator schemes is essential to 
+enable next generation instruments \citep{lon00}. Meanwhile, radio transient monitoring 
+requires access to high time and frequency resolution data to identify and characterize events 
+such as fast radio bursts (FRBs, \citealt{lor07}), or to follow up gravitational wave candidates 
+with radio observations \citep{abb16a,abb16b}. FRBs are highly unexplored at low frequencies 
+(< 1 GHz), but are expected to occur on timescales $\Delta t \sim$ 1--10~ms \citep{tho13}. 
+Recording the full visibility matrix for $\Nant \gtrsim 10^3$ arrays at this timescale leads to 
+extremely high data write rates.
 
-Direct imaging correlators are a new variety of radio correlator which aim to alleviate both the computational strain of forming $\Nant^2$ correlations and the high data throughput associated with short timescale science.
-This is done by performing a spatial fast Fourier transform (FFT) to image the antenna voltages, then squaring and averaging in time.
-This process scales as $\mathcal{O}(\Ng \log_2 \Ng)$, where \Ng~is the number of grid points in the FFT \citep{mor11, teg09, teg10}.
-For certain classes of telescopes, significantly those envisioned for next generation cosmology experiments, this scaling is a large improvement over the $\Nant^2$ scaling of traditional methods.
-Furthermore, because images are generated online, the native output bandwidth will be lowered (assuming $\Ng < \Nant^2$), and has the potential to be lowered even further with online transient processing.
+Direct imaging correlators are a new variety of radio correlator which aim to alleviate both the 
+computational strain of forming $\Nant^2$ correlations and the high data throughput associated 
+with short timescale science. This is done by performing a spatial fast Fourier transform (FFT) 
+to image the antenna voltages, then squaring and averaging in time. This process scales as 
+$\mathcal{O}(\Ng \log_2 \Ng)$, where \Ng~is the number of grid points in the FFT \citep{mor11,
+ teg09, teg10}. For certain classes of telescopes, significantly those envisioned for next 
+ generation cosmology experiments, this scaling is a large improvement over the $\Nant^2$ 
+ scaling of traditional methods. Furthermore, because images are generated online, the native 
+ output bandwidth will be lowered (assuming $\Ng < \Nant^2$), and has the potential to be 
+ lowered even further with online transient processing.
 
-A handful of prototype direct imaging correlators have been tested on arrays including the Basic Element for SKA Training II (BEST-2) array \citep{fos14}, the Omniscope \citep{zhe14}, and an earlier pulsar timing experiment at GHz frequencies \citep{oto94, dai00}.
-Each of these are examples of so-called FFT correlators -- a subclass of direct imaging correlators which rely on identical antennas with restricted placement, which allows the FFT to be performed without gridding.
-We recently released the E-field Parallel Imaging Correlator \citep[EPIC;][]{thy15c}, which is a software implementation of the Modular Optimal Frequency Fourier \citep[MOFF;][]{mor11} imaging algorithm.
-This architecture leverages the software holography/A-transpose framework to grid electric field data streams before performing the spatial FFT, allowing for an optimal map without placing constraints on array layout or requiring identical antennas \citep{mor09,bha08,teg97a}.
+A handful of prototype direct imaging correlators have been tested on arrays including the 
+Basic Element for SKA Training II (BEST-2) array \citep{fos14}, the Omniscope \citep{zhe14}, 
+and an earlier pulsar timing experiment at GHz frequencies \citep{oto94, dai00}. Each of these 
+are examples of so-called FFT correlators -- a subclass of direct imaging correlators which rely 
+on identical antennas with restricted placement, which allows the FFT to be performed without 
+gridding. We recently released the E-field Parallel Imaging Correlator \citep[EPIC;][]{thy15c}, 
+which is a software implementation of the Modular Optimal Frequency Fourier \citep[MOFF;][]
+{mor11} imaging algorithm. This architecture leverages the software holography/A-transpose 
+framework to grid electric field data streams before performing the spatial FFT, allowing for an 
+optimal map without placing constraints on array layout or requiring identical antennas 
+\citep{mor09,bha08,teg97a}.
 
-A challenge common to all direct imaging algorithms is calibration of the antenna gains.
-With a traditional FX or XF correlator, pair-wise visibilities are written to disk and used to calibrate offline before further processing such as imaging.
-However, a direct imaging correlator mixes the signals from all antennas before averaging and writing to disk, making calibration a requirement at the front end, before imaging and averaging.
-Previous solutions have involved applying calibration solutions generated from a parallel FX correlator \citep{zhe14, fos14}, or integrating a dedicated FX correlator which periodically formed the full visibility matrix to solve for gains \citep{wij09,dev09}.
-While these solutions were sufficient to enable the exploration of FFT correlators and beamformers, they will not scale to future arrays with $\Nant \gtrsim 10^3$.
+A challenge common to all direct imaging algorithms is calibration of the antenna gains. With a 
+traditional FX or XF correlator, pair-wise visibilities are written to disk and used to calibrate 
+offline before further processing such as imaging. However, a direct imaging correlator mixes 
+the signals from all antennas before averaging and writing to disk, making calibration a 
+requirement at the front end, before imaging and averaging. Previous solutions have involved 
+applying calibration solutions generated from a parallel FX correlator \citep{zhe14, fos14}, or 
+integrating a dedicated FX correlator which periodically formed the full visibility matrix to solve 
+for gains \citep{wij09,dev09}. While these solutions were sufficient to enable the exploration of 
+FFT correlators and beamformers, they will not scale to future arrays with $\Nant \gtrsim 10^3$.
 
-Here we present the E-field Parallel Imaging Calibration (EPICal) algorithm -- a novel solution to the calibration problem, which can be integrated into direct imaging correlators and scales only as the number of antennas, $\mathcal{O}(\Nant)$.
-This method uses a correlation of the uncalibrated antenna signal stream with an output image pixel from the backend of the correlator to solve for the complex gains of the antennas.
-Because the calibration must be applied before gridding and imaging, our solution requires an iterative approach where the data from one time series is used to update the gains which are applied to the following time series.
-An example implementation of the algorithm is available with the EPIC software package\footnote{http://github.com/nithyanandan/EPIC}.
+Here we present the E-field Parallel Imaging Calibration (EPICal) algorithm -- a novel solution 
+to the calibration problem, which can be integrated into direct imaging correlators and scales 
+only as the number of antennas, $\mathcal{O}(\Nant)$. This method uses a correlation of the 
+uncalibrated antenna signal stream with an output image pixel from the backend of the 
+correlator to solve for the complex gains of the antennas. Because the calibration must be 
+applied before gridding and imaging, our solution requires an iterative approach where the data 
+from one time series is used to update the gains which are applied to the following time series. 
+An example implementation of the algorithm is available with the EPIC software package
+\footnote{http://github.com/nithyanandan/EPIC}.
 
-We review the MOFF algorithm and derive the calibration algorithm in \S \ref{sec:math}.
-We then demonstrate the algorithm in simulations in \S \ref{sec:sim}, and apply to a sample LWA data set in \S \ref{sec:data}.
-Then we discuss errors and noise properties of the resulting gain solutions in \S \ref{sec:noise}.
-Finally we conclude and discuss potential extensions to the algorithm in \S \ref{sec:discussion}.
+We review the MOFF algorithm and derive the calibration algorithm in \S \ref{sec:math}. We 
+then demonstrate the algorithm in simulations in \S \ref{sec:sim}, and apply to a sample LWA 
+data set in \S \ref{sec:data}. Then we discuss errors and noise properties of the resulting gain 
+solutions in \S \ref{sec:noise}. Finally we conclude and discuss potential extensions to the 
+algorithm in \S \ref{sec:discussion}.
 
@@
 \subsection{Review of the MOFF algorithm}
-We begin by reviewing the data flow of the MOFF algorithm, highlighting aspects relevant to the calibration.
-The interested reader is encouraged to refer to \citealt{mor11} for a more thorough discussion.
-Figure~\ref{fig:moff_flow} is reproduced from \citealt{thy15c} to illustrate the various steps of the algorithm.
+We begin by reviewing the data flow of the MOFF algorithm, highlighting aspects relevant to 
+the calibration. The interested reader is encouraged to refer to \citealt{mor11} for a more 
+thorough discussion. Figure~\ref{fig:moff_flow} is reproduced from \citealt{thy15c} to illustrate 
+the various steps of the algorithm.
 
@@
 \includegraphics[width=\columnwidth]{fig1.pdf}
-\caption{Data flow for the MOFF algorithm, reproduced from \citealt{thy15c}.
-The time ordered electric fields are collected by the antenna elements then frequency transformed in the F-engine.
-The calibration is applied, then the electric field measurements are gridded using the antenna voltage patterns.
-The imager performs a spatial 2D FFT on the gridded data before squaring and averaging.
-The accumulated images are then written to disk.}
+\caption{Data flow for the MOFF algorithm, reproduced from \citealt{thy15c}. The time ordered 
+electric fields are collected by the antenna elements then frequency transformed in the F-
+engine. The calibration is applied, then the electric field measurements are gridded using the 
+antenna voltage patterns. The imager performs a spatial 2D FFT on the gridded data before 
+squaring and averaging. The accumulated images are then written to disk.}
 \label{fig:moff_flow}
@@
 
-The electric field incident on the ground, $\Er{}(\mathbf{r},t)$, is related to the sky electric field, $E(\hat{\mathbf{s}},t)$, through a Fourier transform.
+The electric field incident on the ground, $\Er{}(\mathbf{r},t)$, is related to the sky electric field, 
+$E(\hat{\mathbf{s}},t)$, through a Fourier transform.
 \begin{equation}
@@
 \end{equation}
-Here $\hat{\mathbf{s}}$ denotes the sine-projected unit vector for the sky angle, $\mathbf{r}$ is the observer's location (measured in wavelengths relative to an arbitrary origin).
-We will encounter several quantities which we attempt to estimate.
-We distinguish the ``true'' values with a superscript $T$, while the estimated quantities are denoted with a prime.
-We define the true antenna signal as a convolution of the antenna voltage pattern, \beamr, with the electric field on the ground.
+Here $\hat{\mathbf{s}}$ denotes the sine-projected unit vector for the sky angle, $\mathbf{r}$ is 
+the observer's location (measured in wavelengths relative to an arbitrary origin). We will 
+encounter several quantities which we attempt to estimate. We distinguish the ``true'' values 
+with a superscript $T$, while the estimated quantities are denoted with a prime. We define the 
+true antenna signal as a convolution of the antenna voltage pattern, \beamr, with the electric 
+field on the ground.
 \begin{equation}
@@
 
-The first step of the MOFF algorithm is identical to the F-engine of a traditional FX correlator where the electric fields are Fourier transformed to frequency channels, $\Er{a}^T(f,t)$.
-Even after transforming along the time axis, we retain the time dependence of the signal to represent a stream of spectra.
+The first step of the MOFF algorithm is identical to the F-engine of a traditional FX correlator 
+where the electric fields are Fourier transformed to frequency channels, $\Er{a}^T(f,t)$. Even 
+after transforming along the time axis, we retain the time dependence of the signal to represent 
+a stream of spectra.
 
-Thus far we have assumed the antennas deliver the true electric field signals. Here we introduce a multiplicative complex gain as well as an additive noise term which corrupt the measured signals.
+Thus far we have assumed the antennas deliver the true electric field signals. Here we 
+introduce a multiplicative complex gain as well as an additive noise term which corrupt the 
+measured signals.
 \begin{equation}\label{eq:apply_gain}
@@
 \end{equation}
-Note that this quantity is neither a true or estimated value, but rather the raw electric field signal corrupted by the complex gain before we attempt to estimate the true value.
-The noise term is strictly receiver noise -- noise introduced by the instrument.
-Any sky noise is implicitly included in the time dependence of the sky electric field.
-We will neglect $\widetilde{n}_a$ for now, but will inspect its effects at the end of this section.
-Each subsequent step treats each frequency channel independently, so we will drop the $f$ to simplify notation.
+Note that this quantity is neither a true or estimated value, but rather the raw electric field signal 
+corrupted by the complex gain before we attempt to estimate the true value. The noise term is 
+strictly receiver noise -- noise introduced by the instrument. Any sky noise is implicitly included 
+in the time dependence of the sky electric field. We will neglect $\widetilde{n}_a$ for now, but 
+will inspect its effects at the end of this section. Each subsequent step treats each frequency 
+channel independently, so we will drop the $f$ to simplify notation.
 
-The MOFF algorithm next calls for a calibration.
-The goal of our new calibration method will be to form an estimate of the gains, $g'_a$.
-For now we will assume we have formed an estimate to proceed with the MOFF pipeline.
-We correct the incoming electric field data stream using our current estimate of the gains.
+The MOFF algorithm next calls for a calibration. The goal of our new calibration method will be 
+to form an estimate of the gains, $g'_a$. For now we will assume we have formed an estimate 
+to proceed with the MOFF pipeline. We correct the incoming electric field data stream using our 
+current estimate of the gains.
 \begin{equation}
@@
 
-The next step is to grid the antenna signals.
-MOFF uses the antenna patterns as the gridding kernel according to the software holography/A-transpose technique \citep{mor09,bha08}.
-This achieves both an optimal map \citep{teg97b} and places the data on a regular grid.
-Unlike traditional correlators which grid visibilities, the MOFF grids the electric fields directly.
-This operation must be performed for every time step, before any averaging of the data.
+The next step is to grid the antenna signals. MOFF uses the antenna patterns as the gridding 
+kernel according to the software holography/A-transpose technique \citep{mor09,bha08}. This 
+achieves both an optimal map \citep{teg97b} and places the data on a regular grid. Unlike 
+traditional correlators which grid visibilities, the MOFF grids the electric fields directly. This 
+operation must be performed for every time step, before any averaging of the data.
 
-After gridding, the imaging portion of the MOFF performs a 2D spatial FFT to form instantaneous electric field images of the sky.
-The pipeline up to this point can be summarized through the following equation.
+After gridding, the imaging portion of the MOFF performs a 2D spatial FFT to form 
+instantaneous electric field images of the sky. The pipeline up to this point can be summarized 
+through the following equation.
 \begin{equation}
@@
 \end{equation}
-The sum over antennas uses the voltage pattern to grid the calibrated electric fields onto regular gridpoint, $i$.
-The sum over $i$ denotes the 2D FFT to sky coordinates, resulting in an estimate for the instantaneous electric field image.
-We can simplify this expression by exchanging the sums to transform the beam term into sky coordinates.
+The sum over antennas uses the voltage pattern to grid the calibrated electric fields onto 
+regular gridpoint, $i$. The sum over $i$ denotes the 2D FFT to sky coordinates, resulting in an 
+estimate for the instantaneous electric field image. We can simplify this expression by 
+exchanging the sums to transform the beam term into sky coordinates.
 \begin{align}\label{eq:epix}
@@
 \end{align}
-In this form we see that the effect of gridding with the antenna voltage pattern is to attenuate the image by a factor of the antenna sky response, $\beamtheta_a(\s_i)$.
+In this form we see that the effect of gridding with the antenna voltage pattern is to attenuate 
+the image by a factor of the antenna sky response, $\beamtheta_a(\s_i)$.
 
-Finally, the images are squared and averaged in time to reduce output bandwidth, then written to disk.
-While operations have been reordered, the resulting image is equivalent to one produced by gridding and imaging visibilities.
-By shifting from cross-correlation of all pairs of antennas to a spatial FFT, the computational cost is significantly lowered for certain classes of compact arrays.
-The output bandwidth is also greatly reduced as visibilities are exchanged for gridded images.
-A full exploration of these scalings was presented in \citealt{thy15c}.
+Finally, the images are squared and averaged in time to reduce output bandwidth, then written 
+to disk. While operations have been reordered, the resulting image is equivalent to one 
+produced by gridding and imaging visibilities. By shifting from cross-correlation of all pairs of 
+antennas to a spatial FFT, the computational cost is significantly lowered for certain classes of 
+compact arrays. The output bandwidth is also greatly reduced as visibilities are exchanged for 
+gridded images. A full exploration of these scalings was presented in \citealt{thy15c}.
 
 \subsection{Derivation of calibration}
-We saw above that the MOFF algorithm (in general any direct imaging architecture) mixes the signal from all antennas coherently, and thus requires the calibration to be applied in real time at the front end of the correlator.
-Furthermore, visibilities are never formed, which are traditionally the basic measurement used to form calibration solutions.
-Here we derive an alternative method to estimate the antenna gains in real time, using the data products of the MOFF.
+We saw above that the MOFF algorithm (in general any direct imaging architecture) mixes the 
+signal from all antennas coherently, and thus requires the calibration to be applied in real time 
+at the front end of the correlator. Furthermore, visibilities are never formed, which are 
+traditionally the basic measurement used to form calibration solutions. Here we derive an 
+alternative method to estimate the antenna gains in real time, using the data products of the 
+MOFF.
 
-Because the gains must be estimated in real time, we will need to use a finite stream of data to form and estimate which will be applied to the subsequent stream of data.
-Therefore the process of finding a solution is an iterative one.
-We will use parenthetical superscripts to denote the calibration loop number.
-For example, we will assume we have already formed an estimate from $n$ loops, $g^{(n)}_a$.
-We will use these estimates to calibrate the next stream of data in order to form an updated estimate, $g^{(n+1)}_a$.
+Because the gains must be estimated in real time, we will need to use a finite stream of data to 
+form and estimate which will be applied to the subsequent stream of data. Therefore the 
+process of finding a solution is an iterative one. We will use parenthetical superscripts to 
+denote the calibration loop number. For example, we will assume we have already formed an 
+estimate from $n$ loops, $g^{(n)}_a$. We will use these estimates to calibrate the next stream 
+of data in order to form an updated estimate, $g^{(n+1)}_a$.
 
-As a starting point, we consider the feedback calibration outlined in \citealt{mor11}.
-There it was suggested to form a correlation of the uncalibrated antenna measurements with an image pixel from the output of the correlator.
-This is a statistically stationary quantity, which can be related to the sum of visibilities involving the antenna used in the correlation -- exactly the sum needed to calibrate a simple sky of a single point source.
-However, we aim for a more generalized solution for arbitrarily complex sky models.
-We therefore study the full expression for the antenna-pixel correlation,
+As a starting point, we consider the feedback calibration outlined in \citealt{mor11}. There it 
+was suggested to form a correlation of the uncalibrated antenna measurements with an image 
+pixel from the output of the correlator. This is a statistically stationary quantity, which can be 
+related to the sum of visibilities involving the antenna used in the correlation -- exactly the sum 
+needed to calibrate a simple sky of a single point source. However, we aim for a more 
+generalized solution for arbitrarily complex sky models. We therefore study the full expression 
+for the antenna-pixel correlation,
 \begin{equation}\label{eq:Cna}
@@
 \end{equation}
-where the superscript $n$ again represents the quantity formed in the $n^\mathrm{th}$ calibration loop, and \spix\, is the pixel center nearest a bright calibrator of interest.
-The following will hold for any chosen pixel, \spix, though it is advantageous to choose a pixel which contains a bright source to achieve a high signal to noise.
+where the superscript $n$ again represents the quantity formed in the $n^\mathrm{th}$ 
+calibration loop, and \spix\, is the pixel center nearest a bright calibrator of interest. The 
+following will hold for any chosen pixel, \spix, though it is advantageous to choose a pixel which 
+contains a bright source to achieve a high signal to noise.
 
@@
 \end{align}
-where in the second step we group time-dependent terms, and in the third we define the true visibilities as the correlation between true antenna electric field measurements.
-It is easy to see from here that the net effect of including the receiver noise term, $\widetilde{n}_a(f,t)$, will result in added noise on the true visibilities, including a bias on the auto-correlation terms.
-In principle a bias can result on any visibility from any noise correlations between antennas, but the implementation included in the EPIC software package restricts the noise term to include only the auto-correlation noise bias, and assumes baseline dependent noise has zero mean.
+where in the second step we group time-dependent terms, and in the third we define the true 
+visibilities as the correlation between true antenna electric field measurements. It is easy to see 
+from here that the net effect of including the receiver noise term, $\widetilde{n}_a(f,t)$, will 
+result in added noise on the true visibilities, including a bias on the auto-correlation terms. In 
+principle a bias can result on any visibility from any noise correlations between antennas, but 
+the implementation included in the EPIC software package restricts the noise term to include 
+only the auto-correlation noise bias, and assumes baseline dependent noise has zero mean.
 
-The next step is to use this antenna-pixel correlation to update our gain solution.
-We will model the right hand side of equation~\ref{eq:cna} by assuming our current gain estimates are approximately correct, $g^{(n)}_b\approx g^T_b$, and replacing the true visibilities with a set of model visibilities, $\V'_{ab}$.
-These model visibilities rely an knowledge of the sky, and should be precomputed offline.
-We can then solve equation~\ref{eq:cna} for $g^T_a$ to achieve an updated estimate.
+The next step is to use this antenna-pixel correlation to update our gain solution. We will model 
+the right hand side of equation~\ref{eq:cna} by assuming our current gain estimates are 
+approximately correct, $g^{(n)}_b\approx g^T_b$, and replacing the true visibilities with a set of 
+model visibilities, $\V'_{ab}$. These model visibilities rely an knowledge of the sky, and should 
+be precomputed offline. We can then solve equation~\ref{eq:cna} for $g^T_a$ to achieve an 
+updated estimate.
 \begin{equation}\label{eq:cal_solution}
@@
 \end{equation}
-This equation is our prescription for estimating the antenna gains of a direct imaging array. The online computation complexity scales as $\mathcal{O}(\Nant)$ as we form a $\Cna$ for each antenna.
+This equation is our prescription for estimating the antenna gains of a direct imaging array. The 
+online computation complexity scales as $\mathcal{O}(\Nant)$ as we form a $\Cna$ for each 
+antenna.
 
-We show schematically the process of calibrating a direct imaging correlator in figure~\ref{fig:schematic}.
-Computationally expensive steps that must be performed ``on-chip" are shown inside the gray box.
-The uncalibrated antenna signals are tapped out after the F-engine and correlated against the output image pixel of interest (i.e., equation~\ref{eq:Cna} is ``on-chip'').
-The correlated values are then passed off-chip to estimate the gains using equation~\ref{eq:cal_solution}, and additional fitting if desired.
-The gains are then passed back to the correlator to update the calibration for subsequent integration intervals.
+We show schematically the process of calibrating a direct imaging correlator in 
+figure~\ref{fig:schematic}. Computationally expensive steps that must be performed ``on-chip" 
+are shown inside the gray box. The uncalibrated antenna signals are tapped out after the F-
+engine and correlated against the output image pixel of interest (i.e., equation~\ref{eq:Cna} is 
+``on-chip''). The correlated values are then passed off-chip to estimate the gains using 
+equation~\ref{eq:cal_solution}, and additional fitting if desired. The gains are then passed back 
+to the correlator to update the calibration for subsequent integration intervals.
 
@@
 \includegraphics[width=\columnwidth]{fig2.pdf}
-\caption{The general data flow of the MOFF correlator, with a feedback calibration loop.
-A pixel from the (unsquared) image is tapped out and correlated against the input antenna electric field signals to form $\Cna$ coefficients (equation~\ref{eq:Cna}).
-These coefficients are then shipped off-chip to apply equations~\ref{eq:cal_solution}~and~\ref{eq:damping}, and any other processing to arrive at an updated estimate of the gains, $g^{(n)}_a$.
-The gray box shows operations which must be done at high speed (before averaging in time), which white boxes show operations which can be performed ``off-chip".}
+\caption{The general data flow of the MOFF correlator, with a feedback calibration loop. A pixel 
+from the (unsquared) image is tapped out and correlated against the input antenna electric field 
+signals to form $\Cna$ coefficients (equation~\ref{eq:Cna}). These coefficients are then 
+shipped off-chip to apply equations~\ref{eq:cal_solution}~and~\ref{eq:damping}, and any other 
+processing to arrive at an updated estimate of the gains, $g^{(n)}_a$. The gray box shows 
+operations which must be done at high speed (before averaging in time), which white boxes 
+show operations which can be performed ``off-chip".}
 \label{fig:schematic}
@@
 
-While testing we found equation~\ref{eq:cal_solution} resulted in oscillatory gain solutions as it was iterated, as is often the case in iterative minimization methods.
-To mitigate this we introduce a damping factor, $0 \leq \damp <1$, which is used to attenuate the gain update, effectively giving the solutions memory of previous iterations.
+While testing we found equation~\ref{eq:cal_solution} resulted in oscillatory gain solutions as it 
+was iterated, as is often the case in iterative minimization methods. To mitigate this we 
+introduce a damping factor, $0 \leq \damp <1$, which is used to attenuate the gain update, 
+effectively giving the solutions memory of previous iterations.
 \begin{equation}\label{eq:damping}
@@
 \end{equation}
-We found that while equation~\ref{eq:cal_solution} does indeed converge on good solutions, the process is made faster by tuning the damping factor.
-While $g'^{(n+1)}_a$ is the best estimate of the gains after the $n^{\textrm{th}}$ iteration, the damped version, $g^{(n+1)}_a$, is actually used in the iterative calibration loop.
-Once the loop converges the damped version is essentially a weighted average over the past several iterations, giving it a longer effective integration time.
+We found that while equation~\ref{eq:cal_solution} does indeed converge on good solutions, 
+the process is made faster by tuning the damping factor. While $g'^{(n+1)}_a$ is the best 
+estimate of the gains after the $n^{\textrm{th}}$ iteration, the damped version, $g^{(n+1)}_a$, is 
+actually used in the iterative calibration loop. Once the loop converges the damped version is 
+essentially a weighted average over the past several iterations, giving it a longer effective 
+integration time.
 
-An important feature to note is that, like the MOFF-generated images themselves, equation~\ref{eq:cal_solution} includes the antenna auto-correlations (the sum is over \emph{all} $b$, not excluding $a$).
-It can be difficult to perfectly model the noise bias from auto-correlations, which can often times be far brighter than the visibilities themselves.
-It can therefore be beneficial to subtract this term directly from \Cna, and exclude the $b=a$ term in the sum.
+An important feature to note is that, like the MOFF-generated images themselves, 
+equation~\ref{eq:cal_solution} includes the antenna auto-correlations (the sum is over 
+\emph{all} $b$, not excluding $a$). It can be difficult to perfectly model the noise bias from 
+auto-correlations, which can often times be far brighter than the visibilities themselves. It can 
+therefore be beneficial to subtract this term directly from \Cna, and exclude the $b=a$ term in 
+the sum.
 \begin{equation}
@@
 \end{equation}
-We have absorbed $g_a^T$ into the electric field so that the correlation is now simply the autocorrelation of the uncalibrated antenna measurement.
-This requires generating these correlations, which again only scale as $\mathcal{O}(\Nant)$, and are generally useful for array diagnostics.
+We have absorbed $g_a^T$ into the electric field so that the correlation is now simply the 
+autocorrelation of the uncalibrated antenna measurement. This requires generating these 
+correlations, which again only scale as $\mathcal{O}(\Nant)$, and are generally useful for array 
+diagnostics.
 
-We conclude this section by connecting our calibration expression to that found in a visibility framework.
-In the limit of a single bright calibrating source at phase center, we can greatly simplify equations~\ref{eq:cna} and~\ref{eq:cal_solution}.
-We will assume the beams are normalized such that $\beamtheta(0)=1$.
-We can further drop the exponential phase terms because $\spix=0$.
-We then absorb the true gains into the true visibilities in equation~\ref{eq:cna} to express as a sum of measured, uncalibrated, visibilities.
+We conclude this section by connecting our calibration expression to that found in a visibility 
+framework. In the limit of a single bright calibrating source at phase center, we can greatly 
+simplify equations~\ref{eq:cna} and~\ref{eq:cal_solution}. We will assume the beams are 
+normalized such that $\beamtheta(0)=1$. We can further drop the exponential phase terms 
+because $\spix=0$. We then absorb the true gains into the true visibilities in 
+equation~\ref{eq:cna} to express as a sum of measured, uncalibrated, visibilities.
 \begin{equation}
@@
 
-We next plug this expression into equation~\ref{eq:cal_solution} to find our simplified calibration solution for a single bright point source.
-Because our sky is a single bright point source, the model visibilities are simply the flux of the source, $S_{\mathrm{src}}$.
+We next plug this expression into equation~\ref{eq:cal_solution} to find our simplified 
+calibration solution for a single bright point source. Because our sky is a single bright point 
+source, the model visibilities are simply the flux of the source, $S_{\mathrm{src}}$.
 \begin{equation}
@@
 \end{equation}
-This is simply a gain-weighted sum of the measured visibilities over the flux of the source, which is indeed the limiting result from a visibility approach, for example seen in \citealt{mit08}.
-The ability to recover the equivalent expression despite not actually forming the visibilities is a result of the fact that only sums over visibilities come into the FX solution, as was described in \citealt{mor11}.
-We have confirmed the limiting case equivalence here, and will explore the more general case in more detail in \S~\ref{sec:noise}.
+This is simply a gain-weighted sum of the measured visibilities over the flux of the source, 
+which is indeed the limiting result from a visibility approach, for example seen in \citealt{mit08}. 
+The ability to recover the equivalent expression despite not actually forming the visibilities is a 
+result of the fact that only sums over visibilities come into the FX solution, as was described in 
+\citealt{mor11}. We have confirmed the limiting case equivalence here, and will explore the 
+more general case in more detail in \S~\ref{sec:noise}.
 
 \section{Simulation}\label{sec:sim}
-We first demonstrate our calibration method through a controlled simulation.
-A complex gain is created for each antenna with random phase and amplitude, which is used to corrupt the simulated data stream, then we attempt to recover the gains using our calibration routine.
-The simulation software used is included in the EPIC package.
+We first demonstrate our calibration method through a controlled simulation. A complex gain is 
+created for each antenna with random phase and amplitude, which is used to corrupt the 
+simulated data stream, then we attempt to recover the gains using our calibration routine. The 
+simulation software used is included in the EPIC package.
 
-Our simulated signal consists of 10 random point sources with flux densities 0.5~Jy~$\lesssim S \lesssim 1$~Jy.
-For an antenna array we use the inner 51 antennas of the MWA layout \citep{bea12}, within a bounding box of 150~m.
-The antenna voltage pattern used is a 4.4~m square tophat on the ground.
-Because our algorithm treats frequency channels independently, we simulate only one channel.
-For context we treat this channel as a single 40 kHz, meaning each subsequent timestep is separated by 25~$\mu$s.
+Our simulated signal consists of 10 random point sources with flux densities 0.5~Jy~$\lesssim 
+S \lesssim 1$~Jy. For an antenna array we use the inner 51 antennas of the MWA layout 
+\citep{bea12}, within a bounding box of 150~m. The antenna voltage pattern used is a 4.4~m 
+square tophat on the ground. Because our algorithm treats frequency channels independently, 
+we simulate only one channel. For context we treat this channel as a single 40 kHz, meaning 
+each subsequent timestep is separated by 25~$\mu$s.
 
-For our unknown gains, we create a set of random complex numbers where the amplitude is drawn from approximately gaussian distribution centered around 1 with width 0.25, and completely random phase.
-These are our ``true gains", and we apply them to the frequency-domain simulated antenna electric fields as in equation~\ref{eq:apply_gain}.
-Our analysis is blind to these values until the end of the process to check accuracy.
-The gain estimates are initialized with unity, $g^{(0)}_a=1$.
+For our unknown gains, we create a set of random complex numbers where the amplitude is 
+drawn from approximately gaussian distribution centered around 1 with width 0.25, and 
+completely random phase. These are our ``true gains", and we apply them to the frequency-
+domain simulated antenna electric fields as in equation~\ref{eq:apply_gain}. Our analysis is 
+blind to these values until the end of the process to check accuracy. The gain estimates are 
+initialized with unity, $g^{(0)}_a=1$.
 
-We next process and image \caliter~time steps (10~ms).
-We also form the correlations, \Cna[0], used in our calibration loop.
-The pixel used for the correlation is the source with the largest apparent flux (intrinsic flux attenuated by the primary beam).
-These correlation values are used to update the gain estimates, which in turn are used to calibrate the following \caliter~time steps.
-Through experimentation we found a damping factor of $\damp=0.35$ resulted in the quickest convergence in this simulation.
+We next process and image \caliter~time steps (10~ms). We also form the correlations, 
+\Cna[0], used in our calibration loop. The pixel used for the correlation is the source with the 
+largest apparent flux (intrinsic flux attenuated by the primary beam). These correlation values 
+are used to update the gain estimates, which in turn are used to calibrate the following 
+\caliter~time steps. Through experimentation we found a damping factor of $\damp=0.35$ 
+resulted in the quickest convergence in this simulation.
 
-The calibration loop continues by updating the gain estimates every \caliter~time steps.
-The phases of our gain estimates are shown in figure~\ref{fig:sim_phase} for 20 such iterations.
-The phase error plotted is the phase relative to the true gain for each antenna (various colored lines).
-One antenna was used as a reference to fix the absolute phase, so has zero phase error.
-The other 50 antennas are shown to have error spanning $2\pi$ initially, and after about 10 iterations lock into a solution, settling down to noise levels around iteration 15 (0.15~s).
-We stop the simulation when the updated gains trace the thermal noise of the simulated sources, which can be seen by the coherence of the 50 antenna gains after iteration 15.
+The calibration loop continues by updating the gain estimates every \caliter~time steps. The 
+phases of our gain estimates are shown in figure~\ref{fig:sim_phase} for 20 such iterations. 
+The phase error plotted is the phase relative to the true gain for each antenna (various colored 
+lines). One antenna was used as a reference to fix the absolute phase, so has zero phase 
+error. The other 50 antennas are shown to have error spanning $2\pi$ initially, and after about 
+10 iterations lock into a solution, settling down to noise levels around iteration 15 (0.15~s). We 
+stop the simulation when the updated gains trace the thermal noise of the simulated sources, 
+which can be seen by the coherence of the 50 antenna gains after iteration 15.
 
@@
 \includegraphics[width=\columnwidth]{fig3.pdf}
-\caption{Phase error of gain estimates as a function of iteration for simulated calibration.
-The gains were initialized with random phases, but the calibration loop was able to recover the correct phases after about 15 iterations.
-Each line represents an antenna in our 51 MWA antenna sample.
+\caption{Phase error of gain estimates as a function of iteration for simulated calibration. The 
+gains were initialized with random phases, but the calibration loop was able to recover the 
+correct phases after about 15 iterations. Each line represents an antenna in our 51 MWA 
+antenna sample.
 }
@@
 
-The estimated gain amplitudes for the simulations are shown in figure~\ref{fig:sim_amp}.
-The quantity plotted is the magnitude of the estimated gains over the true gains, $\left|g^{(n)}_a/g^T_a\right|$, which places all antennas on the same scale.
-We can see the amplitudes converge toward their true values around the same time as the phases (iteration $\sim$15).
-At the beginning of calibration we can see the value of the damping factor.
-At $n=0$, a couple of gains are shown to have abnormally high amplitude estimates, notably one about 3.3 times its true value (red line).
-These unbalanced high estimates caused the entire set of gains to be under estimated at $n=1$, even with a damping factor of 0.35.
-By $n=5$ the unbalanced amplitudes have been damped out and the calibration continues.
-Without the damping factor, the oscillation seen in the first couple iterations would have been significantly larger and taken much longer to fade out.
+The estimated gain amplitudes for the simulations are shown in figure~\ref{fig:sim_amp}. The 
+quantity plotted is the magnitude of the estimated gains over the true gains, $\left|g^{(n)}_a/
+g^T_a\right|$, which places all antennas on the same scale. We can see the amplitudes 
+converge toward their true values around the same time as the phases (iteration $\sim$15). At 
+the beginning of calibration we can see the value of the damping factor. At $n=0$, a couple of 
+gains are shown to have abnormally high amplitude estimates, notably one about 3.3 times its 
+true value (red line). These unbalanced high estimates caused the entire set of gains to be 
+under estimated at $n=1$, even with a damping factor of 0.35. By $n=5$ the unbalanced 
+amplitudes have been damped out and the calibration continues. Without the damping factor, 
+the oscillation seen in the first couple iterations would have been significantly larger and taken 
+much longer to fade out.
 
@@
 \includegraphics[width=\columnwidth]{fig4.pdf}
-\caption{Gain estimate amplitudes as a function of iteration for simulated calibration.
-Again, each line represents an antenna in the 51 MWA antenna sample.
-The gain estimates were initialized randomly, while the true values were unity.
-After about 15 iterations we see the calibration loop has settled around the correct values, with only noise remaining.}
+\caption{Gain estimate amplitudes as a function of iteration for simulated calibration. Again, 
+each line represents an antenna in the 51 MWA antenna sample. The gain estimates were 
+initialized randomly, while the true values were unity. After about 15 iterations we see the 
+calibration loop has settled around the correct values, with only noise remaining.}
 \label{fig:sim_amp}
@@
 
-After the gains have converged, we see both the phases and amplitudes to continue to fluctuate coherently.
-This is due to the stochastic fluctuations of the simulated sources themselves, and is limited by the sky noise.
-For this simulation we restricted each calibration iteration to only 10~ms integration time.
-In principle a calibration implementation could use short integration times to allow the gains to converge, then increase the integration time to reduce this noise.
+After the gains have converged, we see both the phases and amplitudes to continue to 
+fluctuate coherently. This is due to the stochastic fluctuations of the simulated sources 
+themselves, and is limited by the sky noise. For this simulation we restricted each calibration 
+iteration to only 10~ms integration time. In principle a calibration implementation could use 
+short integration times to allow the gains to converge, then increase the integration time to 
+reduce this noise.
 
-Images created at the beginning of calibration and at the end are shown in figure~\ref{fig:sim_images}.
-Each image is obtained over 10~ms integration, corresponding to all snapshot images created with a given set of gain estimates.
-The top panel shows the image produced with our initialized unity gains.
-Because the phases are completely random, the image is essentially noise with the primary beam evident.
-After 20 iterations, the image is far more clear, shown in the bottom panel.
-Each of the ten simulated sources are clearly visible, indicated with red circles.
+Images created at the beginning of calibration and at the end are shown in 
+figure~\ref{fig:sim_images}. Each image is obtained over 10~ms integration, corresponding to 
+all snapshot images created with a given set of gain estimates. The top panel shows the image 
+produced with our initialized unity gains. Because the phases are completely random, the 
+image is essentially noise with the primary beam evident. After 20 iterations, the image is far 
+more clear, shown in the bottom panel. Each of the ten simulated sources are clearly visible, 
+indicated with red circles.
 
@@
 \includegraphics[width=\columnwidth]{fig5b.pdf}
-\caption{Images formed during simulated calibration.
-\emph{Top:} An image generated from a single frequency channel and 10 ms integration after the gain estimates are randomized.
-As expected with random phases, the image is completely noisy, with the shape of the primary beam evident.
-\emph{Bottom:} An image formed after calibration, again with a single frequency channel and 10 ms integration.
-The ten simulated (and modeled) point sources are easily visible, and highlighted with red circles.
+\caption{Images formed during simulated calibration. \emph{Top:} An image generated from a 
+single frequency channel and 10 ms integration after the gain estimates are randomized. As 
+expected with random phases, the image is completely noisy, with the shape of the primary 
+beam evident. \emph{Bottom:} An image formed after calibration, again with a single frequency 
+channel and 10 ms integration. The ten simulated (and modeled) point sources are easily 
+visible, and highlighted with red circles.
 }
@@
 \section{Application to LWA data}\label{sec:data}
-We next demonstrate our calibration algorithm using an observation from the LWA station in New Mexico.
-The data is from the LWA narrow-band transient buffer (TBN), with time ordered voltage data from 255 antennas within a core radius of 100~m.
-The central frequency is 74.03 MHz, with a bandwidth of 100 kHz and writeout timescale of 5.12 ms (frequency channel resolution of 195.3125 Hz).
-For this demonstration we limit ourselves to a single polarization.
+We next demonstrate our calibration algorithm using an observation from the LWA station in 
+New Mexico. The data is from the LWA narrow-band transient buffer (TBN), with time ordered 
+voltage data from 255 antennas within a core radius of 100~m. The central frequency is 74.03 
+MHz, with a bandwidth of 100 kHz and writeout timescale of 5.12 ms (frequency channel 
+resolution of 195.3125 Hz). For this demonstration we limit ourselves to a single polarization.
 
-After correcting for geometric cable delays, the instrument is naturally well calibrated, as was seen in the demonstration of the EPIC imager in \citealt{thy15c}.
-However, we will aim to further improve on this calibration using our algorithm.
+After correcting for geometric cable delays, the instrument is naturally well calibrated, as was 
+seen in the demonstration of the EPIC imager in \citealt{thy15c}. However, we will aim to 
+further improve on this calibration using our algorithm.
 
-We proceed by forming model visibilities.
-We model only two bright objects as point sources: Cyg A with flux 16611.68 Jy \citep{coh07}; and Cas A with flux 17693.9 Jy \citep{kas07}.
-Because the raw data is attenuated by the primary beam of the instrument, we also account for this in our model using beam values consistent with \cite{hic12}.
+We proceed by forming model visibilities. We model only two bright objects as point sources: 
+Cyg A with flux 16611.68 Jy \citep{coh07}; and Cas A with flux 17693.9 Jy \citep{kas07}. 
+Because the raw data is attenuated by the primary beam of the instrument, we also account for 
+this in our model using beam values consistent with \cite{hic12}.
 
-We made several choices while studying the behavior of the LWA data to improve our calibration.
-Through our previous imaging work, we noted that the flux scale of uncalibrated images was consistent with average gain amplitudes of 0.25.
-To allow the calibration to converge quickly we initialized our gain estimates at this level.
-We also found a boost in signal to noise is achieved easily by averaging frequency channels and assuming the gains are constant within a sub-band.
-Here we average solutions across 150 channels, or about 29 kHz.
-With a fractional bandwidth $B/f_0 = 3.9 \times 10^{-4} \ll 1$, we assume a smooth bandpass across the band.
+We made several choices while studying the behavior of the LWA data to improve our 
+calibration. Through our previous imaging work, we noted that the flux scale of uncalibrated 
+images was consistent with average gain amplitudes of 0.25. To allow the calibration to 
+converge quickly we initialized our gain estimates at this level. We also found a boost in signal 
+to noise is achieved easily by averaging frequency channels and assuming the gains are 
+constant within a sub-band. Here we average solutions across 150 channels, or about 29 kHz. 
+With a fractional bandwidth $B/f_0 = 3.9 \times 10^{-4} \ll 1$, we assume a smooth bandpass 
+across the band. 
 A damping factor of $\gamma = 0.7$ was adopted.
 
-Finally, we found that seven antennas\footnote{LWA antenna IDs 48, 85, 124, 148, 203, 217, and 244 were flagged.} produced unstable gain solutions, and in fact corrupted the entire array.
-We therefore flagged these antennas in our analysis, resulting in a total of 248 antennas to calibrate.
-In each calibration loop, we form $C_{i,\hat{\boldsymbol{s}}_0}, i=1,2,\ldots N_a$ and update our gain estimates over 10 timestamps (51.2 ms).
-We iterate the loop 30 times for a total of 1.536 seconds.
-The results of this calibration experiment are shown in figures~\ref{fig:data_phase}~--~\ref{fig:data_images}.
+Finally, we found that seven antennas\footnote{LWA antenna IDs 48, 85, 124, 148, 203, 217, 
+and 244 were flagged.} produced unstable gain solutions, and in fact corrupted the entire array. 
+We therefore flagged these antennas in our analysis, resulting in a total of 248 antennas to 
+calibrate. In each calibration loop, we form $C_{i,\hat{\boldsymbol{s}}_0}, i=1,2,\ldots N_a$ and 
+update our gain estimates over 10 timestamps (51.2 ms). We iterate the loop 30 times for a 
+total of 1.536 seconds. The results of this calibration experiment are shown in figures~
+\ref{fig:data_phase}~--~\ref{fig:data_images}.
 
-Figure~\ref{fig:data_phase} shows the phase of our gain estimates over 30 calibration iterations, again with each colored line representing a different antenna.
-Given the quality of uncalibrated image demonstrated in \cite{thy15c}, it is a bit surprising to see the phase variation in our solutions.
-However, the phases are relatively flat after about 15 iterations (modulo noise), and exhibit a central ``trunk" where the majority of phases are congregated.
-This behavior is suggestive that while the uncalibrated voltages were able to produce a viable image, the minor changes from our solutions will focus the image and improve the quality.
-The actual location of the ``trunk" (slightly negative) is simply determined by the reference antenna chosen to have identically zero phase, but happens to be slightly more positive than the bulk of antennas.
-For plotting clarity, we unwrapped phases resulting in phases that appear to exceed $\pm \pi$.
+Figure~\ref{fig:data_phase} shows the phase of our gain estimates over 30 calibration 
+iterations, again with each colored line representing a different antenna. Given the quality of 
+uncalibrated image demonstrated in \cite{thy15c}, it is a bit surprising to see the phase 
+variation in our solutions. However, the phases are relatively flat after about 15 iterations 
+(modulo noise), and exhibit a central ``trunk" where the majority of phases are congregated. 
+This behavior is suggestive that while the uncalibrated voltages were able to produce a viable 
+image, the minor changes from our solutions will focus the image and improve the quality. The 
+actual location of the ``trunk" (slightly negative) is simply determined by the reference antenna 
+chosen to have identically zero phase, but happens to be slightly more positive than the bulk of 
+antennas. For plotting clarity, we unwrapped phases resulting in phases that appear to exceed 
+$\pm \pi$.
 
@@
 \caption{Gain phase solutions as a function of calibration iteration for an LWA TBN observation. 
-The gain estimates are initialized with zero phase, but quickly span a $2\pi$ range, and settle into relatively flat, albeit noisy, solutions.
-The majority of phases congregate near zero, which is not surprising given the fairly good quality image produced from uncalibrated data.
+The gain estimates are initialized with zero phase, but quickly span a $2\pi$ range, and settle 
+into relatively flat, albeit noisy, solutions. The majority of phases congregate near zero, which is 
+not surprising given the fairly good quality image produced from uncalibrated data.
 }
@@
 
-The gain amplitudes as a function of calibration iteration are shown in figure~\ref{fig:data_amp}.
-There is a fairly wide range in gain amplitudes (from 0.12 to 0.59).
-However, this is not surprising due to the non-uniformity in the cables from the antennas to the receivers.
-Again, the amplitudes are noisy but relatively flat.
+The gain amplitudes as a function of calibration iteration are shown in 
+figure~\ref{fig:data_amp}. There is a fairly wide range in gain amplitudes (from 0.12 to 0.59). 
+However, this is not surprising due to the non-uniformity in the cables from the antennas to the 
+receivers. Again, the amplitudes are noisy but relatively flat.
 
@@
 \includegraphics[width=\columnwidth]{fig7.pdf}
-\caption{Gain amplitudes as a function of calibration iteration for an LWA TBN observation.
-The gains estimates were initialized with amplitude 0.25 after inspection of the electric field values compared to our model sky.
-The solutions are noisy, but flat.
-The range in amplitudes is due to the non-uniformity of cables between the LWA antennas and receivers.
+\caption{Gain amplitudes as a function of calibration iteration for an LWA TBN observation. The 
+gains estimates were initialized with amplitude 0.25 after inspection of the electric field values 
+compared to our model sky. The solutions are noisy, but flat. The range in amplitudes is due to 
+the non-uniformity of cables between the LWA antennas and receivers.
 }
@@
 
-Figure~\ref{fig:data_images} shows the improvement in the images due to our calibration.
-The left panel shows the uncalibrated image integrated over 51.2 ms, 29 kHz.
-Cyg~A is prominent near the center of the image, and Cas~A is also clearly visible in the upper right.
-The middle panel shows the image produced after calibration with identical integration time and bandwidth.
-The sidelobes throughout the image are significantly suppressed, and the galactic plane is much more evident, despite only modeling Cyg~A and Cas~A.
-We also note that the feature just to the right of Cyg~A is dimmer in the calibrated image, better matching the expected flux from the global sky model (GSM, \citealt{deo08}).
-We show the GSM \citep{pri16} modulated by two factors of the LWA power pattern in the right panel of figure~\ref{fig:data_images} for reference.
+Figure~\ref{fig:data_images} shows the improvement in the images due to our calibration. The 
+left panel shows the uncalibrated image integrated over 51.2 ms, 29 kHz. Cyg~A is prominent 
+near the center of the image, and Cas~A is also clearly visible in the upper right. The middle 
+panel shows the image produced after calibration with identical integration time and bandwidth. 
+The sidelobes throughout the image are significantly suppressed, and the galactic plane is 
+much more evident, despite only modeling Cyg~A and Cas~A. We also note that the feature 
+just to the right of Cyg~A is dimmer in the calibrated image, better matching the expected flux 
+from the global sky model (GSM, \citealt{deo08}). We show the GSM \citep{pri16} modulated 
+by two factors of the LWA power pattern in the right panel of figure~\ref{fig:data_images} for 
+reference.
 
@@
 \includegraphics[width=0.3\linewidth]{fig8c.pdf}
-\caption{Images produced before (left) and after (middle) calibrating LWA data.
-These images were produces with 29 kHz bandwidth and 51.2 ms integration.
-The calibrated image shows significant reduction in sidelobe rumble throughout, while retaining the prominent Cyg~A and Cas~A sources.
-The galactic plane is also substantially more evident after calibration.
-For reference, the GSM is shown in the same coordinates and weighted by two factors of the LWA primary beam in the right panel.
-While much of the GSM is visible in the calibrated image, we note that the model used to calibrate was only the two bright sources, Cyg~A and Cas~A.
+\caption{Images produced before (left) and after (middle) calibrating LWA data. These images 
+were produces with 29 kHz bandwidth and 51.2 ms integration. The calibrated image shows 
+significant reduction in sidelobe rumble throughout, while retaining the prominent Cyg~A and 
+Cas~A sources. The galactic plane is also substantially more evident after calibration. For 
+reference, the GSM is shown in the same coordinates and weighted by two factors of the LWA 
+primary beam in the right panel. While much of the GSM is visible in the calibrated image, we 
+note that the model used to calibrate was only the two bright sources, Cyg~A and Cas~A.
 }
@@
 
-To compare the image qualities quantitatively we compute the dynamic range, defined as the peak of the image over the noise level.
-We estimate the noise level as the median of the absolute deviation of the image, a robust statistic.
-With this metric we find the calibrated image to have a 55\% dynamic range improvement over the uncalibrated image.
+To compare the image qualities quantitatively we compute the dynamic range, defined as the 
+peak of the image over the noise level. We estimate the noise level as the median of the 
+absolute deviation of the image, a robust statistic. With this metric we find the calibrated image 
+to have a 55\% dynamic range improvement over the uncalibrated image.
 
 \section{Noise trends and error analysis}\label{sec:noise}
-Now we return to simulations to study the effect of noise in our system, and the consequences of an incomplete sky model.
-We run a suite of simulations varying the receiver noise and integration times, while forming antenna-pixel correlations, $\Cna$, for EPICal gain solutions, and simultaneously forming visibilities from the same E-field streams to find visibility-based gain solutions for comparison.
+Now we return to simulations to study the effect of noise in our system, and the consequences 
+of an incomplete sky model. We run a suite of simulations varying the receiver noise and 
+integration times, while forming antenna-pixel correlations, $\Cna$, for EPICal gain solutions, 
+and simultaneously forming visibilities from the same E-field streams to find visibility-based 
+gain solutions for comparison.
 
-We use a simulated sky consisting of a 5~Jy calibrator source, and 49 other random sources with apparent flux densities 0.2 to 0.5 Jy (total sky power $\approx$ 23 Jy).
-We generated these sources randomly, but kept them fixed for each run.
-As in section~\ref{sec:sim}, we use the MWA core layout (51 antennas), and a 4.4~m square tophat antenna voltage pattern.
-We simulate 64 frequency channels each of width 40~kHz.
-Because our calibration loop treats frequency channels independently, each channel can be treated as a separate trial of the simulation, and is used to better estimate the statistics.
+We use a simulated sky consisting of a 5~Jy calibrator source, and 49 other random sources 
+with apparent flux densities 0.2 to 0.5 Jy (total sky power $\approx$ 23 Jy). We generated 
+these sources randomly, but kept them fixed for each run. As in section~\ref{sec:sim}, we use 
+the MWA core layout (51 antennas), and a 4.4~m square tophat antenna voltage pattern. We 
+simulate 64 frequency channels each of width 40~kHz. Because our calibration loop treats 
+frequency channels independently, each channel can be treated as a separate trial of the 
+simulation, and is used to better estimate the statistics.
 
-To simulate receiver noise, we add a gaussian distributed complex random number to each antenna electric field measurement at each time stamp according to equation~\ref{eq:apply_gain}.
-The level of the receiver noise, $\sigma_r = \left<\left|\widetilde{n}_a(f,t)\right|^2\right>$, is varied in different simulation runs.
-We include two limiting cases, where the receiver noise power is subdominant to the sky power ($\sigma_r = 10.0$~Jy), and where the receiver dominates the noise ($\sigma_r=100.0$~Jy).
+To simulate receiver noise, we add a gaussian distributed complex random number to each 
+antenna electric field measurement at each time stamp according to 
+equation~\ref{eq:apply_gain}. The level of the receiver noise, $\sigma_r = \left<\left|
+\widetilde{n}_a(f,t)\right|^2\right>$, is varied in different simulation runs. We include two limiting 
+cases, where the receiver noise power is subdominant to the sky power ($\sigma_r = 
+10.0$~Jy), and where the receiver dominates the noise ($\sigma_r=100.0$~Jy).
 
@@
 \end{enumerate}
-In the cases where we use the full sky model, the model visibilities, $\V'_{ab}$ used in the calibration loop are created using all point sources in the simulated sky.
-For the single point source model, we only allow the calibration loop to be privy to the bright calibrator source.
-While we vary the model used for calibration, the simulated ``true sky'' remains fixed with all 50 point sources.
+In the cases where we use the full sky model, the model visibilities, $\V'_{ab}$ used in the 
+calibration loop are created using all point sources in the simulated sky. For the single point 
+source model, we only allow the calibration loop to be privy to the bright calibrator source. 
+While we vary the model used for calibration, the simulated ``true sky'' remains fixed with all 50 
+point sources.
 
-For the EPICal calibration loops, we initialize our gain estimates with the true values (${g^{(0)}_a=g^T_a=1}$), and allow the estimate to be corrupted by the noise through ten iterations of the calibration loop.
-We adopt a damping factor $\damp=0.35$.
-Because EPICal updates the gain estimate at each calibration loop, \tcal, but retains a memory of previous iterations through the damping factor, the total integration time is not straightforward.
-We define an effective integration time by considering the relative weights of each previous $\Cna$ contributing to the current $g^{(n)}_a$ estimate, each with integration time \tcal.
-In the limit $n\rightarrow \infty$, the geometric series converges to an effective integration time of
+For the EPICal calibration loops, we initialize our gain estimates with the true values 
+(${g^{(0)}_a=g^T_a=1}$), and allow the estimate to be corrupted by the noise through ten 
+iterations of the calibration loop. We adopt a damping factor $\damp=0.35$. Because EPICal 
+updates the gain estimate at each calibration loop, \tcal, but retains a memory of previous 
+iterations through the damping factor, the total integration time is not straightforward. We define 
+an effective integration time by considering the relative weights of each previous $\Cna$ 
+contributing to the current $g^{(n)}_a$ estimate, each with integration time \tcal. In the limit $n
+\rightarrow \infty$, the geometric series converges to an effective integration time of
 \begin{equation}
@@
 
-We simultaneously form simulated visibilities by correlating all pairs of antenna E-field measurements (including the receiver noise).
-The E-fields are correlated for a duration equal to the effective integration time of the EPICal loop for comparison.
-We then find the visibility-based gain estimates by minimizing
+We simultaneously form simulated visibilities by correlating all pairs of antenna E-field 
+measurements (including the receiver noise). The E-fields are correlated for a duration equal to 
+the effective integration time of the EPICal loop for comparison. We then find the visibility-
+based gain estimates by minimizing
 \begin{equation}\label{eq:vis_cal}
@@
 
- For each version of calibration, we observe the error in the gain estimates averaging over both antennas and frequency channels.
+ For each version of calibration, we observe the error in the gain estimates averaging over both 
+ antennas and frequency channels.
 \begin{equation}\label{eq:gain_error}
@@
 
-The results of our simulations are shown in figure~\ref{fig:errors}.
-The gain errors from our EPICal method is shown with solid lines, while the dashed lines represent the visibility-based calibration.
-The thick lines represent the results of using a full sky model, and the thin lines are from the single point source model.
-In the case of the full sky model, we see the errors trend downward with longer integration times, as expected.
-The EPICal errors are slightly higher than the visibility-based errors, on average about 23\% difference.
-This is not surprising as EPICal only uses the information in a single pixel, while the visibilities use all information from the full sky model.
-However, we see that with this perfect model the errors do behave like noise, and the same level of noise can be achieved with a longer integration time (or larger damping factor).
+The results of our simulations are shown in figure~\ref{fig:errors}. The gain errors from our 
+EPICal method is shown with solid lines, while the dashed lines represent the visibility-based 
+calibration. The thick lines represent the results of using a full sky model, and the thin lines are 
+from the single point source model.  In the case of the full sky model, we see the errors trend 
+downward with longer integration times, as expected. The EPICal errors are slightly higher than 
+the visibility-based errors, on average about 23\% difference. This is not surprising as EPICal 
+only uses the information in a single pixel, while the visibilities use all information from the full 
+sky model. However, we see that with this perfect model the errors do behave like noise, and 
+the same level of noise can be achieved with a longer integration time (or larger damping 
+factor).
 
@@
 \includegraphics[width=\columnwidth]{fig9.pdf}
-\caption{Gain estimate errors as a function of integration time, receiver noise (line color), and calibration method.
-The solid lines represent the error on the EPICal derived gain estimates, while the dashed lines represent the error on the visibility based estimates.
-The thicker lines were derived using a full sky model in the calibration loops, and the thinner lines used only a single point source model.
-In the full sky model case we see all methods trend down with longer integration time as expected.
-For a given integration time, EPICal derived errors are on average 23\% higher than those of visibility based calibration.
-In the case of a single source model, the errors bottom out due to the flux in the measurements that is not modeled.
-In this regime the EPICal method performs slightly better than the visibility-based method due to effectively beamforming to the sky pixel where the incomplete model is most accurate.
-On average EPICal errors are 5\% lower in this regime.
+\caption{Gain estimate errors as a function of integration time, receiver noise (line color), and 
+calibration method. The solid lines represent the error on the EPICal derived gain estimates, 
+while the dashed lines represent the error on the visibility based estimates. The thicker lines 
+were derived using a full sky model in the calibration loops, and the thinner lines used only a 
+single point source model. In the full sky model case we see all methods trend down with 
+longer integration time as expected. For a given integration time, EPICal derived errors are on 
+average 23\% higher than those of visibility based calibration. In the case of a single source 
+model, the errors bottom out due to the flux in the measurements that is not modeled. In this 
+regime the EPICal method performs slightly better than the visibility-based method due to 
+effectively beamforming to the sky pixel where the incomplete model is most accurate. On 
+average EPICal errors are 5\% lower in this regime.
 }
@@
 
-The exact ratio of EPICal gain noise to visibility-based gain noise depends on the fraction of the total sky power contained in our calibrator source.
-In the example here, our calibrator accounted for 22\% of the total sky power.
-We repeated the experiment in the regime where the calibrator dominated the sky and found that the difference in EPICal and visibility-based error goes to zero, as expected.
-For a typical HERA observation, the sky temperature is expected to by about 180~K \citep{jac15}, and a bright calibrator source could be about 10\% of this power.
-In this regime we found that EPICal gain noise was about 60\% higher than visibility-based gain noise for the same effective integration time.
+The exact ratio of EPICal gain noise to visibility-based gain noise depends on the fraction of 
+the total sky power contained in our calibrator source. In the example here, our calibrator 
+accounted for 22\% of the total sky power. We repeated the experiment in the regime where 
+the calibrator dominated the sky and found that the difference in EPICal and visibility-based 
+error goes to zero, as expected. For a typical HERA observation, the sky temperature is 
+expected to by about 180~K \citep{jac15}, and a bright calibrator source could be about 10\% 
+of this power. In this regime we found that EPICal gain noise was about 60\% higher than 
+visibility-based gain noise for the same effective integration time.
 
-When using the single point source model, all calibration methods in figure~\ref{fig:errors} trend downward until they reach an error floor due to the confusion sources that were not modeled.
-The EPICal and visibility-based solutions reach a similar floor, but EPICal achieves a marginally lower level (on average 5\% lower).
-This can be attributed to the same reason EPICal underperformed with the full sky model: EPICal only used a single pixel in the sky to form its solutions.
-When using the full sky model, it was down weighting the information in the rest of the sky.
-But when the sky model only contains the bright point source, the single pixel used in EPICal is the location where this model is most accurate, effectively down weighting the pixels with incomplete sky model.
-In the regime typical of HERA discussed above, EPICal's error floor was 10\% lower than that of visibility-based solutions.
+When using the single point source model, all calibration methods in figure~\ref{fig:errors} trend 
+downward until they reach an error floor due to the confusion sources that were not modeled. 
+The EPICal and visibility-based solutions reach a similar floor, but EPICal achieves a 
+marginally lower level (on average 5\% lower). This can be attributed to the same reason 
+EPICal underperformed with the full sky model: EPICal only used a single pixel in the sky to 
+form its solutions. When using the full sky model, it was down weighting the information in the 
+rest of the sky. But when the sky model only contains the bright point source, the single pixel 
+used in EPICal is the location where this model is most accurate, effectively down weighting 
+the pixels with incomplete sky model. In the regime typical of HERA discussed above, EPICal's 
+error floor was 10\% lower than that of visibility-based solutions.
 
-Any realistic sky model will lie somewhere between the two extremes explored here.
-In the case where the sky is modeled by a single source, as is often the case for an initial calibration, EPICal actually achieves smaller gain errors compared to the traditional visibility-based calibration.
-As the sky model improves, both visibility-based and EPICal gains improve, though the former quicker than the latter.
-In the limit of a perfect model, both calibration methods produce noise-like errors in their gains which scale down with more integration time.
+Any realistic sky model will lie somewhere between the two extremes explored here. In the 
+case where the sky is modeled by a single source, as is often the case for an initial calibration, 
+EPICal actually achieves smaller gain errors compared to the traditional visibility-based 
+calibration. As the sky model improves, both visibility-based and EPICal gains improve, though 
+the former quicker than the latter. In the limit of a perfect model, both calibration methods 
+produce noise-like errors in their gains which scale down with more integration time.
 
 \section{Discussion}\label{sec:discussion}
-Through simulations and application to real data, we have shown that the EPICal algorithm is a viable solution for calibrating direct imaging arrays in real time.
-The computation necessary only scales with the number of antenna elements, making it a sub-dominant cost factor when designing the correlator.
-This strategy will enable fast read-out for arrays with many thousands of antennas, which will be necessary for future radio transient and cosmology experiments.
+Through simulations and application to real data, we have shown that the EPICal algorithm is a 
+viable solution for calibrating direct imaging arrays in real time. The computation necessary 
+only scales with the number of antenna elements, making it a sub-dominant cost factor when 
+designing the correlator. This strategy will enable fast read-out for arrays with many thousands 
+of antennas, which will be necessary for future radio transient and cosmology experiments.
 
-EPICal can be further improved through several extensions.
-Here we name a few potential considerations for further study.
-These topics can be thought of as extensions to the white ``estimate gains" box in figure~\ref{fig:schematic}, which can be performed at a much lower cadence than the actual correlations.
+EPICal can be further improved through several extensions. Here we name a few potential 
+considerations for further study. These topics can be thought of as extensions to the white 
+``estimate gains" box in figure~\ref{fig:schematic}, which can be performed at a much lower 
+cadence than the actual correlations.
 
-\textbf{Multiple pixel correlations.} As was seen in section~\ref{sec:noise}, the single pixel correlation demonstrated in this paper can underperform in the presence of a complex sky, when compared to visibility-based gains with perfect knowledge of the sky.
-These errors can be mitigated by using correlations of multiple image pixels to incorporate a higher fraction of the total sky power into the calibration loop.
-Of course this increases the computational cost to a scaling of $\mathcal{O}(\Nant N_{\mathrm{pix}})$, which will typically be much lower than the $\mathcal{O}(\Ng \log_2 \Ng)$ of the correlator itself.
-These additional correlations would also enable direction dependent gain solutions.
-Each correlation could be used to independently solve for gains in each pixel direction, then fit to a beam model on the sky.
-This updated beam pattern would then feed into the gridding step of the correlator, allowing the imager to convolve the signals with the effective beam pattern in the ground plane.
+\textbf{Multiple pixel correlations.} As was seen in section~\ref{sec:noise}, the single pixel 
+correlation demonstrated in this paper can underperform in the presence of a complex sky, 
+when compared to visibility-based gains with perfect knowledge of the sky. These errors can be 
+mitigated by using correlations of multiple image pixels to incorporate a higher fraction of the 
+total sky power into the calibration loop. Of course this increases the computational cost to a 
+scaling of $\mathcal{O}(\Nant N_{\mathrm{pix}})$, which will typically be much lower than the 
+$\mathcal{O}(\Ng \log_2 \Ng)$ of the correlator itself. These additional correlations would also 
+enable direction dependent gain solutions. Each correlation could be used to independently 
+solve for gains in each pixel direction, then fit to a beam model on the sky. This updated beam 
+pattern would then feed into the gridding step of the correlator, allowing the imager to convolve 
+the signals with the effective beam pattern in the ground plane.
 
-\textbf{Fitting gain models.} With some knowledge of the instrumental bandpass, the noise of the gain solutions can be greatly reduced by fitting a model to the per-frequency solutions derived here.
-We used this method at a rudimentary level in our demonstration to LWA data by assuming the gains were constant over a narrow bandwidth.
-One could easily improve on this by extending the bandwidth and fitting for a low order polynomial in phase and amplitude.
-Additionally, with knowledge of the poly-phase filters applied to the data stream, we could include channels closer to the edge of the band.
+\textbf{Fitting gain models.} With some knowledge of the instrumental bandpass, the noise of 
+the gain solutions can be greatly reduced by fitting a model to the per-frequency solutions 
+derived here. We used this method at a rudimentary level in our demonstration to LWA data by 
+assuming the gains were constant over a narrow bandwidth. One could easily improve on this 
+by extending the bandwidth and fitting for a low order polynomial in phase and amplitude. 
+Additionally, with knowledge of the poly-phase filters applied to the data stream, we could 
+include channels closer to the edge of the band.
 
-\textbf{Improving sky model.} In this work we used a perfect sky model (for simulation), or a very simple sky model (for LWA data).
-In principle the images produced by the EPIC correlator can be used to improve the sky model used in calibration -- similar to a major loop in self-calibration.
-This can be especially useful for compact, widefield arrays which require a model of both compact and diffuse sources over a large patch of sky.
-In addition to requiring the sky model, the calibration also requires this model to be attenuated by the primary beam, which can be difficult to measure at the precision necessary \citep[e.g.][]{neb15,vir14,thy15b}.
-However, the direct imaging correlator provides exactly the model necessary in real time, and can be iterated over to improve the images and gain solutions.
+\textbf{Improving sky model.} In this work we used a perfect sky model (for simulation), or a 
+very simple sky model (for LWA data). In principle the images produced by the EPIC correlator 
+can be used to improve the sky model used in calibration -- similar to a major loop in self-
+calibration. This can be especially useful for compact, widefield arrays which require a model of 
+both compact and diffuse sources over a large patch of sky. In addition to requiring the sky 
+model, the calibration also requires this model to be attenuated by the primary beam, which 
+can be difficult to measure at the precision necessary \citep[e.g.][]{neb15,vir14,thy15b}. 
+However, the direct imaging correlator provides exactly the model necessary in real time, and 
+can be iterated over to improve the images and gain solutions.
 
-\textbf{Dynamic parameters.} In our controlled experiments we fine-tuned a number of parameters based on our testing (e.g.
-damping factor, integration time, frequency averaging).
-A deployed system will require robust determination of these parameters to operate continuously.
-The specifics will be heavily dependent on the stability of the instrument, the frequency of observation, and the sky.
-For example, a stable instrument may be able to use short integrations to determine a rough estimate of the gains before switching to much longer integration (on order seconds to minutes) to highly increase signal to noise.
-At low frequencies or high imaging resolution, the dynamics of the ionosphere are important, and will likely drive the limit of time integration allowed.
+\textbf{Dynamic parameters.} In our controlled experiments we fine-tuned a number of 
+parameters based on our testing (e.g. damping factor, integration time, frequency averaging). A 
+deployed system will require robust determination of these parameters to operate continuously. 
+The specifics will be heavily dependent on the stability of the instrument, the frequency of 
+observation, and the sky. For example, a stable instrument may be able to use short 
+integrations to determine a rough estimate of the gains before switching to much longer 
+integration (on order seconds to minutes) to highly increase signal to noise. At low frequencies 
+or high imaging resolution, the dynamics of the ionosphere are important, and will likely drive 
+the limit of time integration allowed.
 
-The work here will serve as a foundation for further development.
-We have shown that the EPICal algorithm produces reliable calibration solutions, and have identified several aspects to increase the scope.
-With next generation instruments in the planning and development stages, EPICal is poised to greatly contribute to their success.
-The software is integrated into the EPIC package and freely available to use in simulations or offline processing of data.
-Work is underway to port the code to real-time GPU systems for deployment.
+The work here will serve as a foundation for further development. We have shown that the 
+EPICal algorithm produces reliable calibration solutions, and have identified several aspects to 
+increase the scope. With next generation instruments in the planning and development stages, 
+EPICal is poised to greatly contribute to their success. The software is integrated into the EPIC 
+package and freely available to use in simulations or offline processing of data. Work is 
+underway to port the code to real-time GPU systems for deployment.
 
 \section*{Acknowledgements}
-This work has been supported by the National Science Foundation through award AST-1206552. 
-We thank Danny Jacobs for his valuable inputs, and Greg Taylor for providing us with LWA data. 
-Construction of the LWA has been supported by the Office of Naval Research under 
-Contract N00014-07-C-0147. Support for operations and continuing development of 
-the LWA1 is provided by the National Science Foundation under grant AST-1139974 
-of the University Radio Observatory program.
+This work has been supported by the National Science Foundation through award 
+AST-1206552. We thank Danny Jacobs for his valuable inputs, and Greg Taylor for providing 
+us with LWA data. Construction of the LWA has been supported by the Office of Naval 
+Research under Contract N00014-07-C-0147. Support for operations and continuing 
+development of the LWA1 is provided by the National Science Foundation under grant 
+AST-1139974 of the University Radio Observatory program.
 

### 2016-04-18 cal_demo/ms.tex
Addressing some of the easy/straightforward referee comments to cal_demo paper

@@
  scaling of traditional methods. Furthermore, because images are generated online, the native 
- output bandwidth will be lowered (assuming $\Ng < \Nant^2$), and has the potential to be 
+ output bandwidth will be lowered in cases where $\Ng < \Nant^2$, and has the potential to be 
  lowered even further with online transient processing.
@@
 kernel according to the software holography/A-transpose technique \citep{mor09,bha08}. This 
-achieves both an optimal map \citep{teg97b} and places the data on a regular grid. Unlike 
+achieves an optimal map \citep{teg97b} and places the data on a regular grid. Unlike 
 traditional correlators which grid visibilities, the MOFF grids the electric fields directly. This 
@@
 \Ethetaest(\s_i,t) & = \frac{1}{\Nant} \sum_a \frac{g^T_a}{g'_a}\Er{a}^T(t) e^{2\pi i \s_i \cdot \ra} \nonumber \\
-  & \qquad\times \sum_i \beamr_b(\ri-\ra)e^{2\pi i \s_i \cdot (\ri-\ra)} \nonumber\\
+  & \qquad\times \sum_i \beamr_a(\ri-\ra)e^{2\pi i \s_i \cdot (\ri-\ra)} \nonumber\\
 & = \frac{1}{\Nant} \sum_a \frac{ g^T_a}{g'_a}\Er{a}^T(t) e^{2\pi i \s_i \cdot \ra}\beamtheta_a(\s_i)
@@
 converge toward their true values around the same time as the phases (iteration $\sim$15). At 
-the beginning of calibration we can see the value of the damping factor. At $n=0$, a couple of 
+the beginning of calibration we can see the importance of the damping factor. At $n=0$, a couple of 
 gains are shown to have abnormally high amplitude estimates, notably one about 3.3 times its 
@@
 Finally, we found that seven antennas\footnote{LWA antenna IDs 48, 85, 124, 148, 203, 217, 
-and 244 were flagged.} produced unstable gain solutions, and in fact corrupted the entire array. 
-We therefore flagged these antennas in our analysis, resulting in a total of 248 antennas to 
+and 244 were omitted.} produced unstable gain solutions, and in fact corrupted the entire array. 
+We therefore omitted these antennas from our analysis, resulting in a total of 248 antennas to 
 calibrate. In each calibration loop, we form $C_{i,\hat{\boldsymbol{s}}_0}, i=1,2,\ldots N_a$ and 
@@
 The results of our simulations are shown in figure~\ref{fig:errors}. The gain errors from our 
-EPICal method is shown with solid lines, while the dashed lines represent the visibility-based 
+EPICal method are shown with solid lines, while the dashed lines represent the visibility-based 
 calibration. The thick lines represent the results of using a full sky model, and the thin lines are 
@@
 only uses the information in a single pixel, while the visibilities use all information from the full 
-sky model. However, we see that with this perfect model the errors do behave like noise, and 
-the same level of noise can be achieved with a longer integration time (or larger damping 
+sky model. However, we see that with this perfect model the errors do behave like noise
+in the sense that they integrate down with longer integrations with the expected $t^{-1/2}$ 
+trend. 
+The same level of gain error as visibility based solutions can be achieved with a longer integration time (or larger damping 
 factor).

### 2016-05-02 cal_demo/ms.tex
Starting major revisions in cal sec 2.1

@@
 
-The electric field incident on the ground, $\Er{}(\mathbf{r},t)$, is related to the sky electric field, 
-$E(\hat{\mathbf{s}},t)$, through a Fourier transform.
+%%% New text
+We begin by considering the electric field vector incident on the ground as a function of time, 
+$\mathbf{E}(\mathbf{r},t)$. The position vector, $\mathbf{r}$, is measured from some arbitrary 
+origin. Following \citealt{cla99} we make a few assumptions in order to clarify notation and 
+simplify the problem. First we note that the time ordered electric field can be expressed as a 
+Fourier series with coefficients corresponding to frequencies $f$. This allows us to analyze the 
+quasi-monochromatic components of the electric field, $\mathbf{E}(\mathbf{r},f,t)$, with the 
+understanding that the full field is composed of all components and their corresponding Fourier 
+functions. We retain the time dependence because in practice each coefficient will only be 
+measured over some sampling period.
+
+It is convenient to simplify notation by expressing the position vector in units of wavelengths, 
+$\mathbf{r}\rightarrow \mathbf{r}/\lambda$. We will also assume all antennas lie in a plane, 
+reducing the problem to two dimensions. While the MOFF algorithm in principle can account for 
+non-coplanar arrays through the W-projection algorithm \citep{cor08}, the EPIC code does not 
+currently support this so we leave it to future work.
+
+Next we treat the electric field as a scalar quantity, ignoring polarization effects. This allows us 
+to express the electric field as a scalar, $E(\mathbf{r},f,t)$, and treat subsequent field 
+propagation and measurements as scalar rather than tensor multiplications. A full polarization 
+treatment is required to construct polarized sky images, however the calibration will only 
+require a sky model expressed in the instrumental polarization basis. We can therefore treat 
+each polarization independently.
+
+Finally we assume the sources of the electric field are far away, allowing us to describe the 
+electric fields propagating from sky sources as originating from an imaginary celestial sphere 
+located a large distance from the ground. We express this intermediary electric field as an 
+angular distribution, $\mathcal{E}(\hat{\mathbf{s}},f,t)$, where $\hat{\mathbf{s}}$ is the unit 
+vector in the direction of a patch of the celestial sphere. We further assume the space between 
+the celestial sphere and the ground is empty, allowing us to express the electric field incident 
+on the ground as the integral of the propagated fields from the sphere.
 \begin{equation}
-\Er{}(\mathbf{r},t) = \int E(\hat{\mathbf{s}},t) e^{-2\pi i \mathbf{r}\cdot \hat{\mathbf{s}}}\, \dif^2 \hat{\mathbf{s}}
+E(\mathbf{r},f,t) = \int \mathcal{E}(\hat{\mathbf{s}},f,t) e^{-2\pi i \mathbf{r} \cdot \hat{\mathbf{s}}}\, d\Omega
 \end{equation}  
-Here $\hat{\mathbf{s}}$ denotes the sine-projected unit vector for the sky angle, $\mathbf{r}$ is 
-the observer's location (measured in wavelengths relative to an arbitrary origin). We will 
-encounter several quantities which we attempt to estimate. We distinguish the ``true'' values 
-with a superscript $T$, while the estimated quantities are denoted with a prime. We define the 
-true antenna signal as a convolution of the antenna voltage pattern, \beamr, with the electric 
-field on the ground.
+By making the latter assumptions we have left treatment of the ionosphere and terrestrial 
+emission such as local interference or thermal radiation from the ground for future work.
+
+Next we turn to the measurement performed by an antenna with far-field radiation pattern 
+$\mathcal{W}_a(\hat{\mathbf{s}})$, where the subscript $a$ denotes a specific antenna. We 
+adopt the common convention of normalizing the radiation pattern to have response of 1 at the 
+center of the main lobe \citep[e.g.][]{nap99}. Consequently, the translation between the terminal 
+voltages to electric field measurements has been absorbed in our definition of the radiation 
+pattern. The ideal antenna electric field measurement can be expressed as 
 \begin{equation}
-\Er{a}^T(t) \equiv \int \beamr_a(\mathbf{r}-\ra) \Er{}(\mathbf{r},t) \, \dif^2 \mathbf{r}
+E^I_a(f,t) = \int \mathcal{W}_a(\hat{\mathbf{s}})\mathcal{E}(\hat{\mathbf{s}},f,t) e^{-2\pi i \mathbf{r} \cdot \hat{\mathbf{s}}}\, d\Omega. 
 \end{equation}
-The subscript $a$ labels the antenna, and $\mathbf{r}_a$ is the location of antenna $a$.
+Equivalently we can express the effect of the antenna radiation pattern in the plane of the 
+aperture.
+\begin{equation}
+E^I_a(f,t) = \int_{\mathrm{aperture}}W_a(\mathbf{r}-\mathbf{r}_a)E_a(\mathbf{r},t)\,d^2\mathbf{r}
+\end{equation}
+Here we used the Fourier relationship between the antenna's far-field radiation pattern and the 
+aperture distribution, $W_a(\mathbf{r}-\mathbf{r}_a)=\int  \mathcal{W}_a(\hat{\mathbf{s}}) \, d
+\Omega$. Although any antenna extends over a region of space, we include a nominal antenna 
+position, $\mathbf{r}_a$, for multiple reasons. In the case where all antennas of an array are 
+identical, a global aperture distribution can be adopted and the translation of the antennas 
+distinguishes them. However, even if all antennas have unique aperture distributions, it can be 
+convenient to retain the relative positions of antennas in the analysis. This allows for easy 
+comparison with visibility-based analyses where the baseline vectors are used to describe 
+sampling of the $uv$ plane. From a practical point of view, measurements of the antenna 
+radiation pattern may not be performed in-situ, so it is convenient to describe the antenna 
+aperture distribution relative to a point fixed within the aperture, along with a translational 
+position.
+
+%The electric field incident on the ground, $\Er{}(\mathbf{r},t)$, is related to the sky electric field, 
+%$E(\hat{\mathbf{s}},t)$, through a Fourier transform.
+%\begin{equation}
+%\Er{}(\mathbf{r},t) = \int E(\hat{\mathbf{s}},t) e^{-2\pi i \mathbf{r}\cdot \hat{\mathbf{s}}}\, \dif^2 \hat{\mathbf{s}}
+%\end{equation}
+%Here $\hat{\mathbf{s}}$ denotes the sine-projected unit vector for the sky angle, $\mathbf{r}$ is 
+%the observer's location (measured in wavelengths relative to an arbitrary origin). We will 
+%encounter several quantities which we attempt to estimate. We distinguish the ``true'' values 
+%with a superscript $T$, while the estimated quantities are denoted with a prime. We define the 
+%true antenna signal as a convolution of the antenna voltage pattern, \beamr, with the electric 
+%field on the ground.
+%\begin{equation}
+%\Er{a}^T(t) \equiv \int \beamr_a(\mathbf{r}-\ra) \Er{}(\mathbf{r},t) \, \dif^2 \mathbf{r}
+%\end{equation}
+%The subscript $a$ labels the antenna, and $\mathbf{r}_a$ is the location of antenna $a$.
 

### 2016-05-02 cal_demo/ms.tex
describe F-engine and FX in cal paper

@@
 A challenge common to all direct imaging algorithms is calibration of the antenna gains. With a 
-traditional FX or XF correlator, pair-wise visibilities are written to disk and used to calibrate 
+traditional cross correlator, pair-wise visibilities are written to disk and used to calibrate 
 offline before further processing such as imaging. However, a direct imaging correlator mixes 
@@
 \begin{equation}
-E^I_a(f,t) = \int_{\mathrm{aperture}}W_a(\mathbf{r}-\mathbf{r}_a)E_a(\mathbf{r},t)\,d^2\mathbf{r}
+E^I_a(f,t) = \int_{\mathrm{aperture}}W_a(\mathbf{r}-\mathbf{r}_a)E_a(\mathbf{r},f,t)\,d^2\mathbf{r}
 \end{equation}
@@
 
-The first step of the MOFF algorithm is identical to the F-engine of a traditional FX correlator 
-where the electric fields are Fourier transformed to frequency channels, $\Er{a}^T(f,t)$. Even 
-after transforming along the time axis, we retain the time dependence of the signal to represent 
-a stream of spectra.
+The first step of the MOFF algorithm is to digitize the antenna signals and perform an FFT to 
+channelize the time-ordered data. This is effectively estimating the quasi-monochromatic 
+components of the electric field we already discussed. The sampling rate of the digitization 
+determines the bandwidth of the observation, and the sampling period determines the minimum 
+frequency channel resolution. The frequency Fourier transform operation is often referred to as 
+the F-engine of a correlator. The MOFF shares the same F-engine design as an FX cross-
+correlator which performs the frequency transform (F-engine) before cross multiplying (X-
+engine).
+
+%The first step of the MOFF algorithm is identical to the F-engine of a traditional FX correlator 
+%where the electric fields are Fourier transformed to frequency channels, $\Er{a}^T(f,t)$. Even 
+%after transforming along the time axis, we retain the time dependence of the signal to represent 
+%a stream of spectra.
 

### 2016-05-03 cal_demo/ms.tex
Starting to add text to explain FX and clarify noise assumptions

@@
 frequency channel resolution. The frequency Fourier transform operation is often referred to as 
-the F-engine of a correlator. The MOFF shares the same F-engine design as an FX cross-
-correlator which performs the frequency transform (F-engine) before cross multiplying (X-
-engine).
+the F-engine of a correlator. The MOFF shares the same F-engine design as an FX cross
+correlator which performs the frequency transform (F-engine) before cross multiplying 
+(X-engine).
 
@@
 
-Thus far we have assumed the antennas deliver the true electric field signals. Here we 
-introduce a multiplicative complex gain as well as an additive noise term which corrupt the 
-measured signals.
+Thus far we have assumed the antennas perfectly measure the sky electric field distribution according to their far-field radiation patterns.
+Here we introduce a multiplicative complex gain as well as an additive noise term which corrupt the measured signals.
 \begin{equation}\label{eq:apply_gain}
-\Er{a}(f,t) = g^T_a(f,t) \Er{a}^T(f,t) + \widetilde{n}_a(f,t)
+E_a(f,t) = g_a(f,t) E_a^I(f,t) + n_a(f,t)
 \end{equation}
-Note that this quantity is neither a true or estimated value, but rather the raw electric field signal 
-corrupted by the complex gain before we attempt to estimate the true value. The noise term is 
+The noise term is 
 strictly receiver noise -- noise introduced by the instrument. Any sky noise is implicitly included 
-in the time dependence of the sky electric field. We will neglect $\widetilde{n}_a$ for now, but 
-will inspect its effects at the end of this section. Each subsequent step treats each frequency 
+in the time dependence of the sky electric field. 
+We have carried the frequency dependence to this point to make clear the complex gains are
+frequency dependent. However, each subsequent step treats each frequency 
 channel independently, so we will drop the $f$ to simplify notation. 
+Furthermore, while antenna gains may vary with time, we will assume they are constant over
+the period of our calibration, and thus drop their time dependence from the notation.
 
-The MOFF algorithm next calls for a calibration. The goal of our new calibration method will be 
-to form an estimate of the gains, $g'_a$. For now we will assume we have formed an estimate 
+The MOFF algorithm next calls for a calibration of the electric field measurements. The goal of our new calibration method will be 
+to form an estimate of the gains, $g'_a$, where the tick represents the estimate. For now we will assume we have formed an estimate 
 to proceed with the MOFF pipeline. We correct the incoming electric field data stream using our 
@@
 \begin{equation}
-\Erest{a}(t) = \Er{a}(t)/ g'_a
+E'_a(t) = E_a(t)/ g'_a
 \end{equation}
 
-The next step is to grid the antenna signals. MOFF uses the antenna patterns as the gridding 
-kernel according to the software holography/A-transpose technique \citep{mor09,bha08}. This 
+The next step is to grid the calibrated electric field measurements. MOFF uses the antenna patterns as the gridding 
+kernel according to the software holography/A-transpose technique \citep{mor09,bha08}. 
+Mathematically we can express the gridding operation as
+\begin{equation}
+E'_j=\sum_a W_a(\mathbf{r}_i-\mathbf{r}_a) E'_a(t)
+\end{equation}
+This 
 achieves an optimal map \citep{teg97b} and places the data on a regular grid. Unlike 

### 2016-05-03 cal_demo/ms.tex
More smoothing of text, expanding on explanation of gridding step

@@
 \end{equation}
-The noise term is 
-strictly receiver noise -- noise introduced by the instrument. Any sky noise is implicitly included 
-in the time dependence of the sky electric field. 
+%The noise term is 
+%strictly receiver noise -- noise introduced by the instrument. Any sky noise is implicitly included 
+%in the time dependence of the sky electric field. 
 We have carried the frequency dependence to this point to make clear the complex gains are
@@
 The MOFF algorithm next calls for a calibration of the electric field measurements. The goal of our new calibration method will be 
-to form an estimate of the gains, $g'_a$, where the tick represents the estimate. For now we will assume we have formed an estimate 
+to form an estimate of the gains, $g'_a$, where the prime represents an estimate. For now we will assume we have formed an estimate 
 to proceed with the MOFF pipeline. We correct the incoming electric field data stream using our 
@@
 
-The next step is to grid the calibrated electric field measurements. MOFF uses the antenna patterns as the gridding 
+The next step is to grid the calibrated electric field measurements to regularly spaced 
+positions on the 2D plane of the ground, $\mathbf{r}_j$. 
+MOFF uses the antenna patterns as the gridding 
 kernel according to the software holography/A-transpose technique \citep{mor09,bha08}. 
@@
 \begin{equation}
-E'_j=\sum_a W_a(\mathbf{r}_i-\mathbf{r}_a) E'_a(t)
+E'_j=\sum_a W_a(\mathbf{r}_j-\mathbf{r}_a) E'_a(t)
 \end{equation}

### 2016-05-03 cal_demo/ms.tex
minor tweaks of wording

@@
 achieves an optimal map \citep{teg97b} and places the data on a regular grid. Unlike 
-traditional correlators which grid visibilities, the MOFF grids the electric fields directly. This 
-operation must be performed for every time step, before any averaging of the data.
+traditional correlators which grid visibilities, the MOFF grids the electric field measurements directly. This 
+operation must be performed for every sample, before any averaging of the data.
 

### 2016-05-03 cal_demo/ms.tex
More comments on gridding, justifying the 2D FFT

@@
 the celestial sphere and the ground is empty, allowing us to express the electric field incident 
-on the ground as the integral of the propagated fields from the sphere.
-\begin{equation}
+on the ground as a Fourier integral of the propagated fields across the hemisphere above the horizon.
+\begin{equation}\label{eq:sky_propagated}
 E(\mathbf{r},f,t) = \int \mathcal{E}(\hat{\mathbf{s}},f,t) e^{-2\pi i \mathbf{r} \cdot \hat{\mathbf{s}}}\, d\Omega
@@
 
-The next step is to grid the calibrated electric field measurements to regularly spaced 
+The next step is to grid the calibrated electric field measurements to a finites set of regularly spaced 
 positions on the 2D plane of the ground, $\mathbf{r}_j$.
-MOFF uses the antenna patterns as the gridding 
+The extent of the grid is chosen to encompass the footprint of the antenna array.
+MOFF uses the antenna aperture distributions as the gridding 
 kernel according to the software holography/A-transpose technique \citep{mor09,bha08}. 
@@
 After gridding, the imaging portion of the MOFF performs a 2D spatial FFT to form 
-instantaneous electric field images of the sky. The pipeline up to this point can be summarized 
+instantaneous electric field images of the sky at fixed locations $\s_k$.
+The Fourier dual coordinates to $\mathbf{r}$ are $(l,m)$, where $l=\sin\theta\cos\phi$,
+$m=\nobreak\sin\theta\sin\phi$, and $(\theta,\phi)$ are the zenith 
+and azimuthal angles. The full unit vector towards the sky is 
+$\s=(l,m,\sqrt{1-l^2-m^2})$.
+The sample spacing and resolution of $\s_k$ are given 
+by the choices of $\mathbf{r}_j$ grid parameters, according to the usual FFT relations.
+The pipeline up to this point can be summarized 
 through the following equation.
-\begin{equation}
-\Ethetaest (\s_i,t) = \underbrace{\frac{1}{\Nant} \sum_i e^{2\pi i \ri \cdot \s_i}}_{\mathrm{2D\;FFT}} 
-\underbrace{\sum_a \beamr_a(\ri - \ra)}_{\mathrm{Gridding}}
-\underbrace{ \frac{g^T_a}{g'_a} \Er{a}^T(t)}_{\mathrm{Calibration}}
-\end{equation}
+\begin{align}
+\mathcal{E}'(\s_k,t) & = \underbrace{\frac{1}{\Nant} \sum_j e^{2\pi i \mathbf{r}_j \cdot \s_k}}_{\mathrm{2D\;FFT}} 
+\underbrace{\sum_a W_a(\mathbf{r}_k - \ra)}_{\mathrm{Gridding}} \\
+& \qquad \times \underbrace{ \frac{1}{g'_a} \left(g_a E_a^I(t)+n_a(t)\right)}_{\mathrm{Calibration}} \nonumber
+\end{align}
 The sum over antennas uses the voltage pattern to grid the calibrated electric fields onto 

### 2016-05-03 cal_demo/ms.tex
Finished first attempt at rewriting sec 2.1

@@
 \end{align}
-The sum over antennas uses the voltage pattern to grid the calibrated electric fields onto 
-regular gridpoint, $i$. The sum over $i$ denotes the 2D FFT to sky coordinates, resulting in an 
+The sum over antennas uses the antenna aperture distribution to grid the calibrated electric field measurements onto 
+regular gridpoints, $\mathbf{r}_j$. The sum over $j$ denotes the 2D FFT to sky coordinates, resulting in an 
 estimate for the instantaneous electric field image. We can simplify this expression by 
-exchanging the sums to transform the beam term into sky coordinates.
+exchanging the sums and use the DFT to approximate the Fourier relationship between
+the antenna aperture distribution and the far-field radiation pattern.
 \begin{align}\label{eq:epix}
-\Ethetaest(\s_i,t) & = \frac{1}{\Nant} \sum_a \frac{g^T_a}{g'_a}\Er{a}^T(t) e^{2\pi i \s_i \cdot \ra} \nonumber \\
-  & \qquad\times \sum_i \beamr_a(\ri-\ra)e^{2\pi i \s_i \cdot (\ri-\ra)} \nonumber\\
-& = \frac{1}{\Nant} \sum_a \frac{ g^T_a}{g'_a}\Er{a}^T(t) e^{2\pi i \s_i \cdot \ra}\beamtheta_a(\s_i)
+\mathcal{E}(\s_k,t) & = \frac{1}{\Nant} \sum_a \frac{1}{g'_a}\left(g_a E^I_a(t)+n_a(t)\right) e^{2\pi i \s_k \cdot \ra} \nonumber \\
+  & \qquad\times \sum_j W_a(\mathbf{r}_j-\ra)e^{2\pi i \s_k \cdot (\mathbf{r}_j-\ra)} \nonumber\\
+& \approx \frac{1}{\Nant} \sum_a \frac{1}{g'_a}\left(g_aE^I_a(t)+n_a(t)\right) e^{2\pi i \s_k \cdot \ra}\mathcal{W}_a(\s_k)
 \end{align}
-In this form we see that the effect of gridding with the antenna voltage pattern is to attenuate 
-the image by a factor of the antenna sky response, $\beamtheta_a(\s_i)$.
+In this form we see that the effect of gridding with the antenna aperture distribution is to attenuate 
+the image by a factor of the far-field radiation response, $\mathcal{W}_a(\s)$.
 

### 2016-05-04 cal_demo/ms.tex
cleaning up derivation of cal, retaining noise throughout

@@
 Because the gains must be estimated in real time, we will need to use a finite stream of data to 
-form and estimate which will be applied to the subsequent stream of data. Therefore the 
+form an estimate which will be applied to the subsequent stream of data. Therefore the 
 process of finding a solution is an iterative one. We will use parenthetical superscripts to 
@@
 As a starting point, we consider the feedback calibration outlined in \citealt{mor11}. There it 
-was suggested to form a correlation of the uncalibrated antenna measurements with an image 
+was suggested to form a correlation of the uncalibrated antenna measurements with the complex conjugate of an image 
 pixel from the output of the correlator. This is a statistically stationary quantity, which can be 
@@
 \begin{equation}\label{eq:Cna}
-\Cna \equiv \left<\Er{a}(t) E'^*(\spix,t)\right>_t,
+\Cna \equiv \left<E_a^{(n)}(t) \mathcal{E}'^{*(n)}(\spix,t)\right>_t,
 \end{equation}
 where the superscript $n$ again represents the quantity formed in the $n^\mathrm{th}$ 
-calibration loop, and \spix\, is the pixel center nearest a bright calibrator of interest. The 
+calibration loop, 
+the superscript $^*$ indicates complex conjugation,
+and \spix\, is the pixel center nearest a bright calibrator of interest. The 
 following will hold for any chosen pixel, \spix, though it is advantageous to choose a pixel which 
@@
 
-Plugging equation \ref{eq:epix} into equation \ref{eq:Cna}, we find,
+Plugging equation \ref{eq:epix} into equation \ref{eq:Cna}, 
+and substituting the $g^{(n)}_b$ for $g'_b$ to indicate the specific loop number,
+we find,
 \begin{align}\label{eq:cna}
-\Cna & = \left<g^T_a \Er{a}^T(t) \frac{1}{\Nant} \sum_b \frac{g^{T*}_b}{g^{*(n)}_b}\Er{b}^{T*}(t) e^{-2\pi i \spix \cdot \rb} \beamtheta^*_b(\spix)\right>_t \nonumber \\
-& = \frac{g^T_a}{\Nant} \sum_b \frac{g^{T*}_b}{g^{*(n)}_b} \beamtheta^*_b(\spix) e^{-2\pi i \spix \cdot \rb} \left<\Er{a}^T \Er{b}^{*T} \right>_t \nonumber \\
-& = \frac{g^T_a}{\Nant} \sum_b \frac{g^{T*}_b}{g^{*(n)}_b} \beamtheta^*_b(\spix) e^{-2\pi i \spix \cdot \rb} \V^T_{ab}
+\Cna & = \left<\left(g_a E_a^I(t)+n_a(t)\right) \frac{1}{\Nant} \right. \nonumber \\
+& \qquad \left. \times \sum_b \frac{1}{g^{*(n)}_b}\left(g^*_bE_b^{I*}(t)+n^*_b(t)\right) e^{-2\pi i \spix \cdot \rb} \mathcal{W}^*_b(\spix)\right>_t \nonumber \\
+& = \frac{1}{\Nant} \sum_b \frac{1}{g^{*(n)}_b} \mathcal{W}^*_b(\spix) e^{-2\pi i \spix \cdot \rb} \nonumber \\
+& \qquad \times \left<\left(g_aE_a^I(t)+n_a(t)\right)\left(g^*_bE_b^{*I}+n^*_b(t)\right) \right>_t \nonumber \\
+& = \frac{1}{\Nant} \sum_b \frac{1}{g^{*(n)}_b} \mathcal{W}^*_b(\spix) e^{-2\pi i \spix \cdot \rb} \nonumber \\
+& \qquad \times \left(g_a g^*_b \left<E^I_a(t)E^{*I}_b(t)\right>_t + \left<n_a(t) n^*_b(t)\right>_t\right) \nonumber \\
+& = \frac{1}{\Nant} \sum_b \frac{1}{g^{*(n)}_b} \mathcal{W}^*_b(\spix) e^{-2\pi i \spix \cdot \rb} \left(g_ag^*_bV^I_{ab}+n_{ab}\right)
 \end{align}
-where in the second step we group time-dependent terms, and in the third we define the true 
-visibilities as the correlation between true antenna electric field measurements. It is easy to see 
-from here that the net effect of including the receiver noise term, $\widetilde{n}_a(f,t)$, will 
-result in added noise on the true visibilities, including a bias on the auto-correlation terms. In 
-principle a bias can result on any visibility from any noise correlations between antennas, but 
-the implementation included in the EPIC software package restricts the noise term to include 
-only the auto-correlation noise bias, and assumes baseline dependent noise has zero mean.
+where in the second step we group time-dependent terms, 
+in the third step we assume the noise does not correlate with the ideal electric field 
+measurements, and in the final step we introduce the baseline noise $n_{ab}\equiv
+\left<n_a(t)n^*_b(t)\right>_t$, and ideal visibilities as the correlation between ideal
+electric field measurements, $V^I_{ab} \equiv\left<E^I_a(t)E^{*I}_b(t)\right>_t$.
+To be clear, visibilities are never directly formed from the data. Rather it is convenient to
+express $\Cna$ in terms of the ideal visibilities, which we can model when solving for the
+gains below.
+
+\textcolor{red}{need a brief comment on the noise term}
 

### 2016-05-06 cal_demo/ms.tex
more progress towards addressing reviewer comments

@@
 \newcommand{\Cna}[1][n]{\ensuremath{\mathcal{C}^{(#1)}_{a,\spix}}}
+\newcommand{\Kna}[1][n]{\ensuremath{\mathcal{K}^{(#1)}_{a,\spix}}}
 \newcommand{\ri}{\ensuremath{\mathbf{r}_i}}
@@
 \begin{equation}\label{eq:Cna}
-\Cna \equiv \left<E_a^{(n)}(t) \mathcal{E}'^{*(n)}(\spix,t)\right>_t,
+\Kna \equiv \left<E_a^{(n)}(t) \mathcal{E}'^{*(n)}(\spix,t)\right>_t,
 \end{equation}
@@
 we find,
-\begin{align}\label{eq:cna}
-\Cna & = \left<\left(g_a E_a^I(t)+n_a(t)\right) \frac{1}{\Nant} \right. \nonumber \\
+\begin{align}
+\Kna & = \left<\left(g_a E_a^I(t)+n_a(t)\right) \frac{1}{\Nant} \right. \nonumber \\
 & \qquad \left. \times \sum_b \frac{1}{g^{*(n)}_b}\left(g^*_bE_b^{I*}(t)+n^*_b(t)\right) e^{-2\pi i \spix \cdot \rb} \mathcal{W}^*_b(\spix)\right>_t \nonumber \\
 & = \frac{1}{\Nant} \sum_b \frac{1}{g^{*(n)}_b} \mathcal{W}^*_b(\spix) e^{-2\pi i \spix \cdot \rb} \nonumber \\
-& \qquad \times \left<\left(g_aE_a^I(t)+n_a(t)\right)\left(g^*_bE_b^{*I}+n^*_b(t)\right) \right>_t \nonumber \\
+& \qquad \times \left<\left(g_aE_a^I(t)+n_a(t)\right)\left(g^*_bE_b^{I*}+n^*_b(t)\right) \right>_t \nonumber \\
 & = \frac{1}{\Nant} \sum_b \frac{1}{g^{*(n)}_b} \mathcal{W}^*_b(\spix) e^{-2\pi i \spix \cdot \rb} \nonumber \\
-& \qquad \times \left(g_a g^*_b \left<E^I_a(t)E^{*I}_b(t)\right>_t + \left<n_a(t) n^*_b(t)\right>_t\right) \nonumber \\
-& = \frac{1}{\Nant} \sum_b \frac{1}{g^{*(n)}_b} \mathcal{W}^*_b(\spix) e^{-2\pi i \spix \cdot \rb} \left(g_ag^*_bV^I_{ab}+n_{ab}\right)
+& \qquad \times \left(g_a g^*_b \left<E^I_a(t)E^{*I}_b(t)\right>_t + g_a\left<E_a^I(t)n^*_b(t)\right>_t \right. \nonumber \\
+& \qquad \left. + g^*_b\left<n_a(t)E^{I*}_b(t)\right>_t + \left<n_a(t) n^*_b(t)\right>_t\right)  
+%& = \frac{1}{\Nant} \sum_b \frac{1}{g^{*(n)}_b} \mathcal{W}^*_b(\spix) e^{-2\pi i \spix \cdot \rb} \left(g_ag^*_bV^I_{ab}+n_{ab}\right)
 \end{align}
 where in the second step we group time-dependent terms, 
-in the third step we assume the noise does not correlate with the ideal electric field 
-measurements, and in the final step we introduce the baseline noise $n_{ab}\equiv
-\left<n_a(t)n^*_b(t)\right>_t$, and ideal visibilities as the correlation between ideal
-electric field measurements, $V^I_{ab} \equiv\left<E^I_a(t)E^{*I}_b(t)\right>_t$.
-To be clear, visibilities are never directly formed from the data. Rather it is convenient to
-express $\Cna$ in terms of the ideal visibilities, which we can model when solving for the
-gains below.
+and in the third step we expanded the product within the time average.
+Because the noise does not correlate with the ideal antenna measurements we can
+drop the middle terms. We will also assume that the noise of antenna $a$ correlated with
+the noise of antenna $b$ is also zero mean, unless $a=b$. By defining an ideal visibility as
+$V_{ab}^I \equiv \left<E^I_a(t)E^{*I}_b(t)\right>_t$, we can simplify to
+\begin{align}\label{eq:Kna}
+\Kna & = \frac{1}{\Nant} \sum_b \frac{1}{g^{*(n)}_b} \mathcal{W}^*_b(\spix) e^{-2\pi i \spix \cdot \rb}\nonumber \\
+& \qquad \times \left( g_ag_b^* V_{ab}^I + \delta_{ab}\left<|n_a(t)|^2\right>_t\right)
+\end{align}
+where $\delta_{ab}$ is the Kronecker-delta function. To be clear, visibilities are never directly 
+formed from the data. Rather it is convenient to express $\Kna$ in terms of the ideal 
+visibilities, which we can model when solving for the gains below.
 
-\textcolor{red}{need a brief comment on the noise term}
+We will ultimately wish to model the right side of this equation to in order to solve for the
+antenna gains. However, the self-correlated noise bias, $\left<|n_a(t)|^2\right>$, can often be 
+significantly larger than the cross correlation power, and difficult to model precisely. It 
+is therefore beneficial to subtract the self-correlation directly from $\Kna$. By properly
+accounting for the various weighting and phasing terms, we can use self-correlation of the
+uncalibrated antenna electric field measurements to remove the noise bias and $V^I_{aa}$
+term from equation~\ref{eq:Kna}.
+\begin{align}\label{eq:cna}
+\Cna & \equiv \Kna - \frac{1}{\Nant g^{*(n)}_a}\mathcal{W}^*_a(\spix)e^{-2\pi i \spix \cdot \ra} \left<|E_a|^2\right>_t\nonumber \\
+& = \frac{g_a}{\Nant} \sum_{b\ne a} \frac{g^*_b}{g^{*(n)}_b} \mathcal{W}^*_b(\spix) e^{-2\pi i \spix \cdot \rb} V^I_{ab}
+\end{align}
 
@@
 the right hand side of equation~\ref{eq:cna} by assuming our current gain estimates are 
-approximately correct, $g^{(n)}_b\approx g^T_b$, and replacing the true visibilities with a set of 
-model visibilities, $\V'_{ab}$. These model visibilities rely an knowledge of the sky, and should 
-be precomputed offline. We can then solve equation~\ref{eq:cna} for $g^T_a$ to achieve an 
+approximately correct, $g^{(n)}_b\approx g_b$, and replacing the ideal visibilities with a set of 
+model visibilities, $V^M_{ab}$. These model visibilities rely on knowledge of the sky, and should 
+be precomputed before observing. We can then solve equation~\ref{eq:cna} for $g_a$ to achieve an 
 updated estimate.
 \begin{equation}\label{eq:cal_solution}
-g'^{(n+1)}_a = \Cna \Nant \left[ \sum_b \beamtheta^*_b(\spix) e^{-2\pi i \spix \cdot \rb} \V'_{ab} \right]^{-1}
+g^{(n+1)}_a = \Cna \Nant \left[ \sum_{b\ne a} \mathcal{W}^*_b(\spix) e^{-2\pi i \spix \cdot \rb} V^M_{ab} \right]^{-1}
 \end{equation}
 This equation is our prescription for estimating the antenna gains of a direct imaging array. The 
-online computation complexity scales as $\mathcal{O}(\Nant)$ as we form a $\Cna$ for each 
+real-time computation complexity scales as $\mathcal{O}(\Nant)$ as we form 
+an antenna-pixel correlation, $\Cna$, for each 
 antenna.
+
+\textcolor{red}{pick up here}
 

### 2016-05-16 cal_demo/ms.tex
Fixing more terminology, going through sec 2 with fine toothed comb

@@
 \caption{Data flow for the MOFF algorithm, reproduced from \citealt{thy15c}. The time ordered 
-electric fields are collected by the antenna elements then frequency transformed in the F-
-engine. The calibration is applied, then the electric field measurements are gridded using the 
-antenna voltage patterns. The imager performs a spatial 2D FFT on the gridded data before 
+electric field measurements are collected by the antenna elements then frequency transformed in the 
+F-engine. The calibration is applied, then the electric field measurements are gridded using the 
+antenna illumination patterns. The imager performs a spatial 2D FFT on the gridded data before 
 squaring and averaging. The accumulated images are then written to disk.}
@@
 Here we used the Fourier relationship between the antenna's far-field radiation pattern and the 
-aperture distribution, $W_a(\mathbf{r}-\mathbf{r}_a)=\int  \mathcal{W}_a(\hat{\mathbf{s}}) \, d
+aperture illumination pattern (sometimes referred to as the aperture distribution), $W_a(\mathbf{r}-\mathbf{r}_a)=\int  \mathcal{W}_a(\hat{\mathbf{s}}) \, d
 \Omega$. Although any antenna extends over a region of space, we include a nominal antenna 
@@
 
-The next step is to grid the calibrated electric field measurements to a finites set of regularly spaced 
+The next step is to grid the calibrated electric field measurements to a finite set of regularly spaced 
 positions on the 2D plane of the ground, $\mathbf{r}_j$.
@@
 Mathematically we can express the gridding operation as
-\begin{equation}
+\begin{equation}\label{eq:gridding}
 E'_j=\sum_a W_a(\mathbf{r}_j-\mathbf{r}_a) E'_a(t)
 \end{equation}
-This 
+Formally each pixel is calculated by summing over all antennas and their illumination patterns.
+However the illumination patterns are typically compact and non-overlapping, so most pixels
+will only contain contribution from at most one antenna, allowing equation~\ref{eq:gridding}
+to be calculated through a sparse matrix multiplication.
+This gridding 
 achieves an optimal map \citep{teg97b} and places the data on a regular grid. Unlike 
 traditional correlators which grid visibilities, the MOFF grids the electric field measurements directly. This 
-operation must be performed for every sample, before any averaging of the data.
+operation must be performed for every time sample, before any averaging of the data.
 
@@
 by the choices of $\mathbf{r}_j$ grid parameters, according to the usual FFT relations. 
+Fine $\mathbf{r}_j$ grid spacing can result in non-physical angular pixels ($l^2+m^2>1$),
+which should be discarded.
+
 The pipeline up to this point can be summarized 
@@
 \mathcal{E}'(\s_k,t) & = \underbrace{\frac{1}{\Nant} \sum_j e^{2\pi i \mathbf{r}_j \cdot \s_k}}_{\mathrm{2D\;FFT}} 
-\underbrace{\sum_a W_a(\mathbf{r}_k - \ra)}_{\mathrm{Gridding}} \\
+\underbrace{\sum_a W_a(\mathbf{r}_j - \ra)}_{\mathrm{Gridding}} \\
 & \qquad \times \underbrace{ \frac{1}{g'_a} \left(g_a E_a^I(t)+n_a(t)\right)}_{\mathrm{Calibration}} \nonumber
 \end{align}
-The sum over antennas uses the antenna aperture distribution to grid the calibrated electric field measurements onto 
+The sum over antennas uses the antenna illumination pattern to grid the calibrated electric field measurements onto 
 regular gridpoints, $\mathbf{r}_j$. The sum over $j$ denotes the 2D FFT to sky coordinates, resulting in an 
@@
 exchanging the sums and use the DFT to approximate the Fourier relationship between
-the antenna aperture distribution and the far-field radiation pattern.
+the antenna illumination pattern and the far-field radiation pattern.
 \begin{align}\label{eq:epix}
@@
 \end{align}
-In this form we see that the effect of gridding with the antenna aperture distribution is to attenuate 
+In this form we see that the effect of gridding with the antenna illumination pattern is to attenuate 
 the image by a factor of the far-field radiation response, $\mathcal{W}_a(\s)$.
@@
 for the antenna-pixel correlation,
-\begin{equation}\label{eq:Cna}
+\begin{equation}\label{eq:Kna_def}
 \Kna \equiv \left<E_a^{(n)}(t) \mathcal{E}'^{*(n)}(\spix,t)\right>_t,
@@
 
-Plugging equation \ref{eq:epix} into equation \ref{eq:Cna}, 
+Plugging equation \ref{eq:epix} into equation \ref{eq:Kna_def}, 
 and substituting the $g^{(n)}_b$ for $g'_b$ to indicate the specific loop number,
@@
 Because the noise does not correlate with the ideal antenna measurements we can
-drop the middle terms. We will also assume that the noise of antenna $a$ correlated with
-the noise of antenna $b$ is also zero mean, unless $a=b$. By defining an ideal visibility as
+drop the middle terms. We will also assume that the cross-correlation noise ($a\ne b$) 
+is zero mean. Then by defining an ideal visibility as
 $V_{ab}^I \equiv \left<E^I_a(t)E^{*I}_b(t)\right>_t$, we can simplify to
@@
 
-We will ultimately wish to model the right side of this equation to in order to solve for the
-antenna gains. However, the self-correlated noise bias, $\left<|n_a(t)|^2\right>$, can often be 
+We will ultimately wish to model the right side of equation~\ref{eq:Kna} in order to solve for the
+antenna gains. However, the self-correlated noise, $\left<|n_a(t)|^2\right>$, can often be 
 significantly larger than the cross correlation power, and difficult to model precisely. It 
 is therefore beneficial to subtract the self-correlation directly from $\Kna$. By properly
-accounting for the various weighting and phasing terms, we can use self-correlation of the
-uncalibrated antenna electric field measurements to remove the noise bias and $V^I_{aa}$
-term from equation~\ref{eq:Kna}.
-\begin{align}\label{eq:cna}
+accounting for the various weighting and phasing terms, we can use self-correlations of the
+uncalibrated antenna electric field measurements to remove the corresponding
+terms from equation~\ref{eq:Kna}.
+\begin{align}\label{eq:Cna}
 \Cna & \equiv \Kna - \frac{1}{\Nant g^{*(n)}_a}\mathcal{W}^*_a(\spix)e^{-2\pi i \spix \cdot \ra} \left<|E_a|^2\right>_t\nonumber \\
@@
 The next step is to use this antenna-pixel correlation to update our gain solution. We will model 
-the right hand side of equation~\ref{eq:cna} by assuming our current gain estimates are 
+the right hand side of equation~\ref{eq:Cna} by assuming our current gain estimates are 
 approximately correct, $g^{(n)}_b\approx g_b$, and replacing the ideal visibilities with a set of 
 model visibilities, $V^M_{ab}$. These model visibilities rely on knowledge of the sky, and should 
-be precomputed before observing. We can then solve equation~\ref{eq:cna} for $g_a$ to achieve an 
+be precomputed before observing. We can then solve equation~\ref{eq:Cna} for $g_a$ to achieve an 
 updated estimate.
 \begin{equation}\label{eq:cal_solution}
-g^{(n+1)}_a = \Cna \Nant \left[ \sum_{b\ne a} \mathcal{W}^*_b(\spix) e^{-2\pi i \spix \cdot \rb} V^M_{ab} \right]^{-1}
+%g^{(n+1)}_a = \Cna \Nant \left[ \sum_{b\ne a} \mathcal{W}^*_b(\spix) e^{-2\pi i \spix \cdot \rb} V^M_{ab} \right]^{-1}
+g^{(n+1)}_a = \frac{\Cna \Nant }{ \sum_{b\ne a} \mathcal{W}^*_b(\spix) e^{-2\pi i \spix \cdot \rb} V^M_{ab} }
 \end{equation}
@@
 
-\textcolor{red}{pick up here}
-
 We show schematically the process of calibrating a direct imaging correlator in 
-figure~\ref{fig:schematic}. Computationally expensive steps that must be performed ``on-chip" 
-are shown inside the gray box. The uncalibrated antenna signals are tapped out after the F-
-engine and correlated against the output image pixel of interest (i.e., equation~\ref{eq:Cna} is 
-``on-chip''). The correlated values are then passed off-chip to estimate the gains using 
-equation~\ref{eq:cal_solution}, and additional fitting if desired. The gains are then passed back 
+figure~\ref{fig:schematic}. Computationally expensive steps that must be performed in near real-time 
+are shown inside the gray box. The uncalibrated antenna signals are tapped out after the 
+F-engine and correlated against the output image pixel of interest (i.e., equation~\ref{eq:Cna} is 
+computed in near real-time). The correlated values are then used to estimate the gains
+using equation~\ref{eq:cal_solution}, and additional fitting if desired. This step does not have
+the speed requirement of the steps in the gray box because the correlations have been 
+averaged in time.
+The gains are then passed back 
 to the correlator to update the calibration for subsequent integration intervals.
@@
 from the (unsquared) image is tapped out and correlated against the input antenna electric field 
-signals to form $\Cna$ coefficients (equation~\ref{eq:Cna}). These coefficients are then 
-shipped off-chip to apply equations~\ref{eq:cal_solution}~and~\ref{eq:damping}, and any other 
-processing to arrive at an updated estimate of the gains, $g^{(n)}_a$. The gray box shows 
-operations which must be done at high speed (before averaging in time), which white boxes 
-show operations which can be performed ``off-chip".}
+measurements to form $\Cna$ coefficients (equation~\ref{eq:Cna}). These coefficients are then 
+used to update the gain estimates using equations~\ref{eq:cal_solution}~and~\ref{eq:damping},
+and any additional fitting.
+The gray box shows operations which must be done at high speed (before averaging in time),
+while the white boxes show operations which can be performed at lower speeds because
+they involve time averages quantities.
+}
 \label{fig:schematic}
@@
 
-While testing we found equation~\ref{eq:cal_solution} resulted in oscillatory gain solutions as it 
-was iterated, as is often the case in iterative minimization methods. To mitigate this we 
+While testing we found equation~\ref{eq:cal_solution} resulted in oscillatory gain solutions over time, 
+as is often the case in iterative methods. To mitigate this we 
 introduce a damping factor, $0 \leq \damp <1$, which is used to attenuate the gain update, 
@@
 \begin{equation}\label{eq:damping}
-g^{(n+1)}_a = (1-\damp) g'^{(n+1)}_a + \damp g^{(n)}_a
+%g^{(n+1)}_a = (1-\damp) g'^{(n+1)}_a + \damp g^{(n)}_a
+g^{(n+1)}_a = \frac{(1-\damp)\Cna \Nant }{ \sum_{b\ne a} \mathcal{W}^*_b(\spix) e^{-2\pi i \spix \cdot \rb} V^M_{ab} }+ \damp g^{(n)}_a
 \end{equation}
 We found that while equation~\ref{eq:cal_solution} does indeed converge on good solutions, 
-the process is made faster by tuning the damping factor. While $g'^{(n+1)}_a$ is the best 
-estimate of the gains after the $n^{\textrm{th}}$ iteration, the damped version, $g^{(n+1)}_a$, is 
-actually used in the iterative calibration loop. Once the loop converges the damped version is 
+the process is made faster by tuning the damping factor. 
+ Once the loop converges the damped version is 
 essentially a weighted average over the past several iterations, giving it a longer effective 
@@
 
-An important feature to note is that, like the MOFF-generated images themselves, 
-equation~\ref{eq:cal_solution} includes the antenna auto-correlations (the sum is over 
-\emph{all} $b$, not excluding $a$). It can be difficult to perfectly model the noise bias from 
-auto-correlations, which can often times be far brighter than the visibilities themselves. It can 
-therefore be beneficial to subtract this term directly from \Cna, and exclude the $b=a$ term in 
-the sum.
-\begin{equation}
-\Cna \rightarrow \Cna - \frac{1}{\Nant\, g^{*(n)}_a} \beamtheta^*_a(\spix) e^{-2\pi i \spix \cdot \ra} \left<|\Er{a}|^2\right>_t
-\end{equation}
-We have absorbed $g_a^T$ into the electric field so that the correlation is now simply the 
-autocorrelation of the uncalibrated antenna measurement. This requires generating these 
-correlations, which again only scale as $\mathcal{O}(\Nant)$, and are generally useful for array 
-diagnostics.
-
 We conclude this section by connecting our calibration expression to that found in a visibility 
-framework. In the limit of a single bright calibrating source at phase center, we can greatly 
-simplify equations~\ref{eq:cna} and~\ref{eq:cal_solution}. We will assume the beams are 
-normalized such that $\beamtheta(0)=1$. We can further drop the exponential phase terms 
-because $\spix=0$. We then absorb the true gains into the true visibilities in 
-equation~\ref{eq:cna} to express as a sum of measured, uncalibrated, visibilities.
+framework. In the limit of a single bright calibrating source at center, we can greatly 
+simplify equations~\ref{eq:Cna} and~\ref{eq:cal_solution}. We will assume the radiation patterns are 
+normalized such that $\mathcal{W}(0)=1$. We can further drop the exponential phase terms 
+because $\spix$ is perpendicular to our planar antenna array. We then absorb the true gains into the true visibilities in 
+equation~\ref{eq:Cna} to express as a sum of measured, uncalibrated, visibilities.
 \begin{equation}
@@
 \begin{equation}
-g'^{(n+1)}_a \rightarrow \frac{\sum_b  \V_{ab}/g^{*(n)}_b}{\Nant S_{\mathrm{src}}}
+g^{(n+1)}_a \rightarrow \frac{\sum_b  V_{ab}/g^{*(n)}_b}{\Nant S_{\mathrm{src}}}
 \end{equation}
@@
 The ability to recover the equivalent expression despite not actually forming the visibilities is a 
-result of the fact that only sums over visibilities come into the FX solution, as was described in 
+result of the fact that only sums over visibilities come into the visibility-based solution, as was described in 
 \citealt{mor11}. We have confirmed the limiting case equivalence here, and will explore the 

### 2016-05-20 cal_demo/ms.tex
Starting edits on section 3

@@
 
-Our simulated signal consists of 10 random point sources with flux densities 0.5~Jy~$\lesssim 
-S \lesssim 1$~Jy. For an antenna array we use the inner 51 antennas of the MWA layout 
-\citep{bea12}, within a bounding box of 150~m. The antenna voltage pattern used is a 4.4~m 
-square tophat on the ground. Because our algorithm treats frequency channels independently, 
-we simulate only one channel. For context we treat this channel as a single 40 kHz, meaning 
-each subsequent timestep is separated by 25~$\mu$s.
+The physical parameters of our simulation can be arbitrarily scaled to other cases. However, 
+we provide typical units for an MWA observation as an example. Our simulated antenna 
+array consists of the inner 51 antennas of the MWA layout \citep{bea12}, within a bounding 
+box of 150~m. The antenna illumination pattern used is a 4.4~m square tophat; a very rough 
+approximation to the square shaped tiles of the MWA. Because our algorithm treats frequency 
+channels independently, we simulate only one channel. For context we treat this as a channel
+at 150~MHz with 40 kHz bandwidth, and adopt a sampling period of 25~$\mu$s. The simulated signal 
+consists of 10 random point sources with flux densitities 0.5~Jy~$\lesssim S \lesssim$~1~Jy 
+within the main lobe of the primary beam.
 
@@
 drawn from approximately gaussian distribution centered around 1 with width 0.25, and 
-completely random phase. These are our ``true gains", and we apply them to the frequency-
-domain simulated antenna electric fields as in equation~\ref{eq:apply_gain}. Our analysis is 
+completely random phase. These are our ``true gains", and we apply them to the frequency
+domain simulated antenna electric field measurements as in equation~\ref{eq:apply_gain}. Our analysis is 
 blind to these values until the end of the process to check accuracy. The gain estimates are 
@@
 
-We next process and image \caliter~time steps (10~ms). We also form the correlations, 
+We next process and image \caliter~time samples (10~ms). We also form the correlations, 
 \Cna[0], used in our calibration loop. The pixel used for the correlation is the source with the 
 largest apparent flux (intrinsic flux attenuated by the primary beam). These correlation values 
-are used to update the gain estimates, which in turn are used to calibrate the following 
-\caliter~time steps. Through experimentation we found a damping factor of $\damp=0.35$ 
+are used to update the gain estimates. For this simple demonstration we create perfect model 
+visibilities of our 10 simulated sources to be used in equation~\ref{eq:damping}. 
+The updated gain estimates are used to calibrate the following 
+\caliter~time samples. Through experimentation we found a damping factor of $\damp=0.35$ 
 resulted in the quickest convergence in this simulation.
 
-The calibration loop continues by updating the gain estimates every \caliter~time steps. The 
+The calibration loop continues by updating the gain estimates every \caliter~time samples. The 
 phases of our gain estimates are shown in figure~\ref{fig:sim_phase} for 20 such iterations. 

### 2016-06-07 cal_demo/ms.tex
Addressing more small reviewer comments

@@
 image is essentially noise with the primary beam evident. After 20 iterations, the image is far 
-more clear, shown in the bottom panel. Each of the ten simulated sources are clearly visible, 
+more clear, shown in the bottom panel. Each of the ten simulated sources is clearly visible, 
 indicated with red circles.
@@
 We next demonstrate our calibration algorithm using an observation from the LWA station in 
-New Mexico. The data is from the LWA narrow-band transient buffer (TBN), with time ordered 
-voltage data from 255 antennas within a core radius of 100~m. The central frequency is 74.03 
-MHz, with a bandwidth of 100 kHz and writeout timescale of 5.12 ms (frequency channel 
+New Mexico \citep{ell13}. The data is from the LWA narrow-band transient buffer (TBN), with time ordered 
+voltage data from the 255 core antennas within a 110~m x 100~m ellipse. 
+The central frequency is 74.03 
+MHz, with a bandwidth of 100 kHz and sampling period of 5.12 ms (frequency channel 
 resolution of 195.3125 Hz). For this demonstration we limit ourselves to a single polarization.
@@
 images was consistent with average gain amplitudes of 0.25. To allow the calibration to 
-converge quickly we initialized our gain estimates at this level. We also found a boost in signal 
-to noise is achieved easily by averaging frequency channels and assuming the gains are 
-constant within a sub-band. Here we average solutions across 150 channels, or about 29 kHz. 
+converge quickly we initialized our gain estimates at this level. 
+We also found a boost in signal to noise is achieved easily by assuming the gains are
+constant across a range of frequencies, and averaging solutions across channels.
+Here we average solutions across 150 channels, or about 29 kHz. 
 With a fractional bandwidth $B/f_0 = 3.9 \times 10^{-4} \ll 1$, we assume a smooth bandpass 
@@
 A damping factor of $\gamma = 0.7$ was adopted.
+The large damping factor, relative to what was used in simulations, was necessary due to
+the larger noise levels in the LWA data compared to the ideal simulations.
 
@@
 calibrate. In each calibration loop, we form $C_{i,\hat{\boldsymbol{s}}_0}, i=1,2,\ldots N_a$ and 
-update our gain estimates over 10 timestamps (51.2 ms). We iterate the loop 30 times for a 
+update our gain estimates over 10 samples (51.2 ms). We iterate the loop 30 times for a 
 total of 1.536 seconds. The results of this calibration experiment are shown in figures~
@@
 \caption{Images produced before (left) and after (middle) calibrating LWA data. These images 
-were produces with 29 kHz bandwidth and 51.2 ms integration. The calibrated image shows 
+were produced with 29 kHz bandwidth and 51.2 ms integration. The calibrated image shows 
 significant reduction in sidelobe rumble throughout, while retaining the prominent Cyg~A and 
@@
 peak of the image over the noise level. We estimate the noise level as the median of the 
-absolute deviation of the image, a robust statistic. With this metric we find the calibrated image 
-to have a 55\% dynamic range improvement over the uncalibrated image.
+absolute deviation of the image, a robust statistic. With this metric we find the uncalibrated 
+and calibrated images to have dynamic ranges of 88.5 and 137, respectively, a 55\% 
+improvement.
 
-\section{Noise trends and error analysis}\label{sec:noise}
+\section{Noise trends}\label{sec:noise}
 Now we return to simulations to study the effect of noise in our system, and the consequences 

### 2016-06-08 cal_demo/ms.tex
more edits... updated sim images figure, changed model for data

@@
 figure~\ref{fig:sim_images}. Each image is obtained over 10~ms integration, corresponding to 
-all snapshot images created with a given set of gain estimates. The top panel shows the image 
+all snapshot images created with a given set of gain estimates. The left panel shows the image 
 produced with our initialized unity gains. Because the phases are completely random, the 
 image is essentially noise with the primary beam evident. After 20 iterations, the image is far 
-more clear, shown in the bottom panel. Each of the ten simulated sources is clearly visible, 
-indicated with red circles.
+more clear, shown in the middle panel. 
+The ten sources are visible, along with rumble throughout the image. By comparing with
+a perfectly calibrated image (right panel), we see that the rumble and the shapes of the
+point sources are matched between the two images. We therefore conclude that the 
+artifacts visible are dominated by the point spread function, rather than calibration errors.
 
-\begin{figure}
+\begin{figure*}
 \begin{center}
-\includegraphics[width=\columnwidth]{fig5a.pdf}
-\includegraphics[width=\columnwidth]{fig5b.pdf}
-\caption{Images formed during simulated calibration. \emph{Top:} An image generated from a 
+\includegraphics[width=.3\linewidth]{fig5a.pdf}
+\includegraphics[width=.3\linewidth]{fig5b.pdf}
+\includegraphics[width=.3\linewidth]{fig5c.pdf}
+\caption{Images formed during simulated calibration. \emph{Left:} An image generated from a 
 single frequency channel and 10 ms integration after the gain estimates are randomized. As 
 expected with random phases, the image is completely noisy, with the shape of the primary 
-beam evident. \emph{Bottom:} An image formed after calibration, again with a single frequency 
-channel and 10 ms integration. The ten simulated (and modeled) point sources are easily 
-visible, and highlighted with red circles.
+beam evident. \emph{Middle:} An image formed after calibration, again with a single frequency 
+channel and 10 ms integration. Now the simulated point sources are visible in the image. 
+\emph{Right:} An image generated from the same simulated data as the middle image, but 
+with perfect calibration correction. Many of the features seen in the middle image are also
+seen here, confirming that they are point spread function artifacts. The ten simulated (and modeled) point sources are highlighted with red circles.
+\label{fig:sim_images}
 }
-\label{fig:sim_images}
 \end{center}
-\end{figure}
+\end{figure*}
 
@@
 We proceed by forming model visibilities. We model only two bright objects as point sources: 
-Cyg A with flux 16611.68 Jy \citep{coh07}; and Cas A with flux 17693.9 Jy \citep{kas07}. 
+Cyg A with flux 20,539 Jy, and Cas A with flux 19,328 Jy \citep{lan12}. 
 Because the raw data is attenuated by the primary beam of the instrument, we also account for 
@@
 just to the right of Cyg~A is dimmer in the calibrated image, better matching the expected flux 
-from the global sky model (GSM, \citealt{deo08}). We show the GSM \citep{pri16} modulated 
-by two factors of the LWA power pattern in the right panel of figure~\ref{fig:data_images} for 
-reference.
+from the global sky model (GSM, \citealt{deo08}). 
+For reference we should the GSM \citep{pri16} in the right panel of 
+Figure~\ref{fig:data_images}, convolved by the LWA point spread function and weighted
+by  two factors of the primary beam to match the data images.
 
@@
 were produced with 29 kHz bandwidth and 51.2 ms integration. The calibrated image shows 
-significant reduction in sidelobe rumble throughout, while retaining the prominent Cyg~A and 
+significant reduction in rumble throughout, while retaining the prominent Cyg~A and 
 Cas~A sources. The galactic plane is also substantially more evident after calibration. For 
-reference, the GSM is shown in the same coordinates and weighted by two factors of the LWA 
-primary beam in the right panel. While much of the GSM is visible in the calibrated image, we 
+reference, the GSM is shown in the 
+right panel convolved with the LWA point spread function and weighted by two factors
+of the primary beam to place it in the same holographic frame as the data images. While much of the GSM is visible in the calibrated image, we 
 note that the model used to calibrate was only the two bright sources, Cyg~A and Cas~A.

### 2016-06-08 cal_demo/ms.tex
Updated notation in section 5

@@
 \end{enumerate}
-In the cases where we use the full sky model, the model visibilities, $\V'_{ab}$ used in the 
+In the cases where we use the full sky model, the model visibilities, $V^M_{ab}$ used in the 
 calibration loop are created using all point sources in the simulated sky. For the single point 
@@
 For the EPICal calibration loops, we initialize our gain estimates with the true values 
-(${g^{(0)}_a=g^T_a=1}$), and allow the estimate to be corrupted by the noise through ten 
+(${g^{(0)}_a=g_a=1}$), and allow the estimate to be corrupted by the noise through ten 
 iterations of the calibration loop. We adopt a damping factor $\damp=0.35$. Because EPICal 
@@
 \begin{equation}\label{eq:vis_cal}
-\chi^2 = \sum_a\sum_{b\ne a} \left|\V_{ab}-g_a g_b^* \V'_{ab}\right|^2,
+\chi^2 = \sum_a\sum_{b\ne a} \left|V_{ab}-g_a g_b^* V^M_{ab}\right|^2,
 \end{equation}
@@
 \begin{equation}\label{eq:gain_error}
-\sigma_g = \left[\frac{1}{N_f \Nant} \sum_f \sum_a \frac{\left|g'_a(f)-g^T_a(f)\right|^2}{\left|g^T_a(f)\right|^2}\right]^{1/2}
+\sigma_g = \left[\frac{1}{N_f \Nant} \sum_f \sum_a \frac{\left|g^{(n)}_a(f)-g_a(f)\right|^2}{\left|g_a(f)\right|^2}\right]^{1/2}
 \end{equation}

### 2016-06-09 cal_demo/ms.tex
Switched data and noise sections. Needs to be ironed out for new flow
touching up noise section after switching order with data section

@@
 
+\section{Noise trends}\label{sec:noise}
+Now next study the effect of noise in our system, and the consequences 
+of an incomplete sky model. We run a suite of simulations varying the receiver noise and 
+integration times, while forming antenna-pixel correlations, $\Cna$, for EPICal gain solutions, 
+and simultaneously forming visibilities from the same electric field measurement streams to find visibility-based 
+gain solutions for comparison.
+
+We use a simulated sky consisting of a 5~Jy calibrator source, and 49 other random sources 
+with apparent flux densities 0.2 to 0.5 Jy (total sky power $\approx$ 23 Jy). We generated 
+these sources randomly, but kept them fixed for each run. 
+We simulate the same antenna layout as in Section~\ref{sec:sim}. 
+However, we increase the number of frequency channels simulated to 64.
+Because our calibration loop treats 
+frequency channels independently, each channel can be treated as a separate trial of the 
+simulation, and is used to better estimate the statistics.
+
+To simulate receiver noise, we add a gaussian distributed complex random number to each 
+antenna electric field measurement at each time sample according to 
+equation~\ref{eq:apply_gain}. The level of the receiver noise, $\sigma_r = \left<\left|
+n_a(f,t)\right|^2\right>$, is varied in different simulation runs. We include two limiting 
+cases, where the receiver noise power is subdominant to the sky power ($\sigma_r = 
+10.0$~Jy), and where the receiver dominates the noise ($\sigma_r=100.0$~Jy).
+
+For each simulation run we find gain solutions using four methods:
+\begin{enumerate}[i.]
+\item EPICal, using a full sky model
+\item EPICal, using a single point source model
+\item Visibility-based, using a full sky model
+\item Visibility-based, using a single point source model.
+\end{enumerate}
+In the cases where we use the full sky model, the model visibilities, $V^M_{ab}$, used in the 
+calibration loop are created using all point sources in the simulated sky. For the single point 
+source model, we only allow the calibration loop to be privy to the bright calibrator source. 
+While we vary the model used for calibration, the simulated ``true sky'' remains fixed with all 50 
+point sources.
+
+For the EPICal calibration loops, we initialize our gain estimates with the true values 
+(${g^{(0)}_a=g_a=1}$), and allow the estimate to be corrupted by the noise through ten 
+iterations of the calibration loop. We again adopt a damping factor $\damp=0.35$. Because EPICal 
+updates the gain estimate at each calibration loop, \tcal, but retains a memory of previous 
+iterations through the damping factor, the total integration time is not straightforward. We define 
+an effective integration time by considering the relative weights of each previous $\Cna$ 
+contributing to the current $g^{(n)}_a$ estimate, each with integration time \tcal. In the limit $n
+\rightarrow \infty$, the geometric series converges to an effective integration time of
+\begin{equation}
+\teff = \tcal \times \frac{1+\damp}{1-\damp}.
+\end{equation}
+With the damping factor we adopted here, $\teff \approx 2.1 \times \tcal$.
+
+We simultaneously form simulated visibilities, $V_{ab}$, by correlating all pairs of antenna electric field 
+measurements (including the receiver noise). The electric field measurements are correlated for a duration equal to 
+the effective integration time of the EPICal loop for comparison. We then find the visibility-
+based gain estimates by minimizing
+\begin{equation}\label{eq:vis_cal}
+\chi^2 = \sum_a\sum_{b\ne a} \left|V_{ab}-g_a g_b^* V^M_{ab}\right|^2,
+\end{equation}
+using both versions of the model visibilities described above.
+
+ For each version of calibration, we observe the error in the gain estimates by averaging over both 
+ antennas and frequency channels.
+\begin{equation}\label{eq:gain_error}
+\sigma_g = \left[\frac{1}{N_f \Nant} \sum_f \sum_a \frac{\left|g^{(n)}_a(f)-g_a(f)\right|^2}{\left|g_a(f)\right|^2}\right]^{1/2}
+\end{equation}
+
+The results of our simulations are shown in Figure~\ref{fig:errors}. The gain errors from our 
+EPICal method are shown with solid lines, while the dashed lines represent the visibility-based 
+calibration. The thick lines represent the results of using a full sky model, and the thin lines are 
+from the single point source model.  In the case of the full sky model, we see the errors trend 
+downward with longer integration times, as expected. The EPICal errors are slightly higher than 
+the visibility-based errors, on average about 23\% difference. This is not surprising as EPICal 
+only uses the information in a single pixel, while the visibilities use all information from the full 
+sky model. However, we see that with this perfect model the errors do behave like noise
+in the sense that they integrate down with longer integrations with the expected $t^{-1/2}$ 
+trend. 
+The same level of gain error as visibility based solutions can be achieved with a longer integration time (or larger damping 
+factor).
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{fig9.pdf}
+\caption{Gain estimate errors as a function of integration time, receiver noise (line color), and 
+calibration method. The solid lines represent the error on the EPICal derived gain estimates, 
+while the dashed lines represent the error on the visibility based estimates. The thicker lines 
+were derived using a full sky model in the calibration loops, and the thinner lines used only a 
+single point source model. In the full sky model case we see all methods trend down with 
+longer integration time as expected. For a given integration time, EPICal derived errors are on 
+average 23\% higher than those of visibility based calibration. In the case of a single source 
+model, the errors bottom out due to the flux in the measurements that is not modeled. In this 
+regime the EPICal method performs slightly better than the visibility-based method due to 
+effectively beamforming to the sky pixel where the incomplete model is most accurate. On 
+average EPICal errors are 5\% lower in this regime.
+}
+\label{fig:errors}
+\end{center}
+\end{figure}
+
+The exact ratio of EPICal gain noise to visibility-based gain noise depends on the fraction of 
+the total sky power contained in our calibrator source. In the example here, our calibrator 
+accounted for 22\% of the total sky power. We repeated the experiment in the regime where 
+the calibrator dominated the sky and found that the difference in EPICal and visibility-based 
+error goes to zero, as expected. For a typical HERA observation, the sky temperature is 
+expected to by about 180~K \citep{jac15}, and a bright calibrator source could be about 10\% 
+of this power. In this regime we found that EPICal gain noise was about 60\% higher than 
+visibility-based gain noise for the same effective integration time.
+
+When using the single point source model, all calibration methods in figure~\ref{fig:errors} trend 
+downward until they reach an error floor due to the confusion sources that were not modeled. 
+The EPICal and visibility-based solutions reach a similar floor, but EPICal achieves a 
+marginally lower level (on average 5\% lower). This can be attributed to the same reason 
+EPICal underperformed with the full sky model: EPICal only used a single pixel in the sky to 
+form its solutions. When using the full sky model, it was down weighting the information in the 
+rest of the sky. But when the sky model only contains the bright point source, the single pixel 
+used in EPICal is the location where this model is most accurate, effectively down weighting 
+the pixels with incomplete sky model. In the regime typical of HERA discussed above, EPICal's 
+error floor was 10\% lower than that of visibility-based solutions.
+
+Any realistic sky model will lie somewhere between the two extremes explored here. In the 
+case where the sky is modeled by a single source, as is often the case for an initial calibration, 
+EPICal actually achieves smaller gain errors compared to the traditional visibility-based 
+calibration. As the sky model improves, both visibility-based and EPICal gains improve, though 
+the former quicker than the latter. In the limit of a perfect model, both calibration methods 
+produce noise-like errors in their gains which scale down with more integration time.
+
 \section{Application to LWA data}\label{sec:data}
@@
 
-\section{Noise trends}\label{sec:noise}
-Now we return to simulations to study the effect of noise in our system, and the consequences 
-of an incomplete sky model. We run a suite of simulations varying the receiver noise and 
-integration times, while forming antenna-pixel correlations, $\Cna$, for EPICal gain solutions, 
-and simultaneously forming visibilities from the same E-field streams to find visibility-based 
-gain solutions for comparison.
-
-We use a simulated sky consisting of a 5~Jy calibrator source, and 49 other random sources 
-with apparent flux densities 0.2 to 0.5 Jy (total sky power $\approx$ 23 Jy). We generated 
-these sources randomly, but kept them fixed for each run. As in section~\ref{sec:sim}, we use 
-the MWA core layout (51 antennas), and a 4.4~m square tophat antenna voltage pattern. We 
-simulate 64 frequency channels each of width 40~kHz. Because our calibration loop treats 
-frequency channels independently, each channel can be treated as a separate trial of the 
-simulation, and is used to better estimate the statistics.
-
-To simulate receiver noise, we add a gaussian distributed complex random number to each 
-antenna electric field measurement at each time stamp according to 
-equation~\ref{eq:apply_gain}. The level of the receiver noise, $\sigma_r = \left<\left|
-\widetilde{n}_a(f,t)\right|^2\right>$, is varied in different simulation runs. We include two limiting 
-cases, where the receiver noise power is subdominant to the sky power ($\sigma_r = 
-10.0$~Jy), and where the receiver dominates the noise ($\sigma_r=100.0$~Jy).
-
-For each simulation run we find gain solutions using four methods:
-\begin{enumerate}[i.]
-\item EPICal, using a full sky model
-\item EPICal, using a single point source model
-\item Visibility-based, using a full sky model
-\item Visibility-based, using a single point source model.
-\end{enumerate}
-In the cases where we use the full sky model, the model visibilities, $V^M_{ab}$ used in the 
-calibration loop are created using all point sources in the simulated sky. For the single point 
-source model, we only allow the calibration loop to be privy to the bright calibrator source. 
-While we vary the model used for calibration, the simulated ``true sky'' remains fixed with all 50 
-point sources.
-
-For the EPICal calibration loops, we initialize our gain estimates with the true values 
-(${g^{(0)}_a=g_a=1}$), and allow the estimate to be corrupted by the noise through ten 
-iterations of the calibration loop. We adopt a damping factor $\damp=0.35$. Because EPICal 
-updates the gain estimate at each calibration loop, \tcal, but retains a memory of previous 
-iterations through the damping factor, the total integration time is not straightforward. We define 
-an effective integration time by considering the relative weights of each previous $\Cna$ 
-contributing to the current $g^{(n)}_a$ estimate, each with integration time \tcal. In the limit $n
-\rightarrow \infty$, the geometric series converges to an effective integration time of
-\begin{equation}
-\teff = \tcal \times \frac{1+\damp}{1-\damp}.
-\end{equation}
-With the damping factor we adopted here, $\teff \approx 2.1 \times \tcal$.
-
-We simultaneously form simulated visibilities by correlating all pairs of antenna E-field 
-measurements (including the receiver noise). The E-fields are correlated for a duration equal to 
-the effective integration time of the EPICal loop for comparison. We then find the visibility-
-based gain estimates by minimizing
-\begin{equation}\label{eq:vis_cal}
-\chi^2 = \sum_a\sum_{b\ne a} \left|V_{ab}-g_a g_b^* V^M_{ab}\right|^2,
-\end{equation}
-using both versions of the model visibilities described above.
-
- For each version of calibration, we observe the error in the gain estimates averaging over both 
- antennas and frequency channels.
-\begin{equation}\label{eq:gain_error}
-\sigma_g = \left[\frac{1}{N_f \Nant} \sum_f \sum_a \frac{\left|g^{(n)}_a(f)-g_a(f)\right|^2}{\left|g_a(f)\right|^2}\right]^{1/2}
-\end{equation}
-
-The results of our simulations are shown in figure~\ref{fig:errors}. The gain errors from our 
-EPICal method are shown with solid lines, while the dashed lines represent the visibility-based 
-calibration. The thick lines represent the results of using a full sky model, and the thin lines are 
-from the single point source model.  In the case of the full sky model, we see the errors trend 
-downward with longer integration times, as expected. The EPICal errors are slightly higher than 
-the visibility-based errors, on average about 23\% difference. This is not surprising as EPICal 
-only uses the information in a single pixel, while the visibilities use all information from the full 
-sky model. However, we see that with this perfect model the errors do behave like noise
-in the sense that they integrate down with longer integrations with the expected $t^{-1/2}$ 
-trend. 
-The same level of gain error as visibility based solutions can be achieved with a longer integration time (or larger damping 
-factor).
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{fig9.pdf}
-\caption{Gain estimate errors as a function of integration time, receiver noise (line color), and 
-calibration method. The solid lines represent the error on the EPICal derived gain estimates, 
-while the dashed lines represent the error on the visibility based estimates. The thicker lines 
-were derived using a full sky model in the calibration loops, and the thinner lines used only a 
-single point source model. In the full sky model case we see all methods trend down with 
-longer integration time as expected. For a given integration time, EPICal derived errors are on 
-average 23\% higher than those of visibility based calibration. In the case of a single source 
-model, the errors bottom out due to the flux in the measurements that is not modeled. In this 
-regime the EPICal method performs slightly better than the visibility-based method due to 
-effectively beamforming to the sky pixel where the incomplete model is most accurate. On 
-average EPICal errors are 5\% lower in this regime.
-}
-\label{fig:errors}
-\end{center}
-\end{figure}
-
-The exact ratio of EPICal gain noise to visibility-based gain noise depends on the fraction of 
-the total sky power contained in our calibrator source. In the example here, our calibrator 
-accounted for 22\% of the total sky power. We repeated the experiment in the regime where 
-the calibrator dominated the sky and found that the difference in EPICal and visibility-based 
-error goes to zero, as expected. For a typical HERA observation, the sky temperature is 
-expected to by about 180~K \citep{jac15}, and a bright calibrator source could be about 10\% 
-of this power. In this regime we found that EPICal gain noise was about 60\% higher than 
-visibility-based gain noise for the same effective integration time.
-
-When using the single point source model, all calibration methods in figure~\ref{fig:errors} trend 
-downward until they reach an error floor due to the confusion sources that were not modeled. 
-The EPICal and visibility-based solutions reach a similar floor, but EPICal achieves a 
-marginally lower level (on average 5\% lower). This can be attributed to the same reason 
-EPICal underperformed with the full sky model: EPICal only used a single pixel in the sky to 
-form its solutions. When using the full sky model, it was down weighting the information in the 
-rest of the sky. But when the sky model only contains the bright point source, the single pixel 
-used in EPICal is the location where this model is most accurate, effectively down weighting 
-the pixels with incomplete sky model. In the regime typical of HERA discussed above, EPICal's 
-error floor was 10\% lower than that of visibility-based solutions.
-
-Any realistic sky model will lie somewhere between the two extremes explored here. In the 
-case where the sky is modeled by a single source, as is often the case for an initial calibration, 
-EPICal actually achieves smaller gain errors compared to the traditional visibility-based 
-calibration. As the sky model improves, both visibility-based and EPICal gains improve, though 
-the former quicker than the latter. In the limit of a perfect model, both calibration methods 
-produce noise-like errors in their gains which scale down with more integration time.
-
 \section{Discussion}\label{sec:discussion}

### 2016-06-09 cal_demo/ms.tex
Switched around figure filenames after switching sections
Added text re:noise of data solutions, updated freq averaging in data

@@
 \begin{center}
-\includegraphics[width=\columnwidth]{fig9.pdf}
+\includegraphics[width=\columnwidth]{fig6.pdf}
 \caption{Gain estimate errors as a function of integration time, receiver noise (line color), and 
@@
 constant across a range of frequencies, and averaging solutions across channels.
-Here we average solutions across 150 channels, or about 29 kHz. 
-With a fractional bandwidth $B/f_0 = 3.9 \times 10^{-4} \ll 1$, we assume a smooth bandpass 
+Here we average solutions across 300 channels, or about 58.6 kHz. 
+With a fractional bandwidth $B/f_0 = 7.9 \times 10^{-4} \ll 1$, we assume a smooth bandpass 
 across the band. 
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{fig6.pdf}
+\includegraphics[width=\columnwidth]{fig7.pdf}
 \caption{Gain phase solutions as a function of calibration iteration for an LWA TBN observation. 
@@
 
+To ensure the level of noise in our solutions is reasonable we compared to the fractional errors
+from our simulations shown in Figure~\ref{fig:errors}. We first use the Global Sky Model
+\citep{deo08} to generate a mock observed sky at 74.03 MHz using the PyGSM python 
+module\footnote{https://github.com/telegraphic/PyGSM} \citep{pri16}. We integrate this model
+to estimate a total sky flux power of 171,000~Jy, meaning our calibration pixel containing
+Cyg A accounts for  approximately $20,539/171,000 \approx 12\%$ of the total sky power. 
+This places us in a regime very similar to the incomplete sky model simulations of the 
+previous section, where the calibration pixel accounted for 21.7\% of the sky power. We 
+should therefore expect our fractional gain noise levels to be at best around 0.1, the level 
+to which the simulations asymptotically approached.
+
+To estimate our fractional gain noise we calculate the standard deviation of the gain
+estimates after iteration 15 for each antenna separately. We then divide these standard
+deviations by the respective mean gain amplitudes. This resulted in a distribution of fractional
+gain noise centered at 0.15 and a width of 0.04 ($\approx 0.15/\sqrt{15}$). This noise level
+is consistent with what we expected from the simulations.
+There were four outliers with noise levels significantly outside the distribution 
+($\sigma_g \approx 0.29$). Upon closer inspection the gain amplitudes of these outliers had
+converged later in the calibration loop, after iteration 15.
+
 \begin{figure}
 \begin{center}
-\includegraphics[width=\columnwidth]{fig7.pdf}
+\includegraphics[width=\columnwidth]{fig8.pdf}
 \caption{Gain amplitudes as a function of calibration iteration for an LWA TBN observation. The 
@@
 just to the right of Cyg~A is dimmer in the calibrated image, better matching the expected flux 
-from the global sky model (GSM, \citealt{deo08}). 
-For reference we should the GSM \citep{pri16} in the right panel of 
+from the GSM. 
+For reference we show the GSM in the right panel of 
 Figure~\ref{fig:data_images}, convolved by the LWA point spread function and weighted
-by  two factors of the primary beam to match the data images.
+by  two factors of the primary beam to match the holographic frame of the data images.
 
@@
 \begin{center}
-\includegraphics[width=0.3\linewidth]{fig8a.pdf}
-\includegraphics[width=0.3\linewidth]{fig8b.pdf}
-\includegraphics[width=0.3\linewidth]{fig8c.pdf}
+\includegraphics[width=0.3\linewidth]{fig9a.pdf}
+\includegraphics[width=0.3\linewidth]{fig9b.pdf}
+\includegraphics[width=0.3\linewidth]{fig9c.pdf}
 \caption{Images produced before (left) and after (middle) calibrating LWA data. These images 
@@
 by extending the bandwidth and fitting for a low order polynomial in phase and amplitude. 
-Additionally, with knowledge of the poly-phase filters applied to the data stream, we could 
-include channels closer to the edge of the band.
+Additionally, with knowledge of 
+any filters applied to the data stream in the receiver chain,
+we could include channels closer to the edge of the band.
 

### 2016-06-20 cal_demo/ms.tex
Fine tuning notation, language, and consistency. Ready to write up response to referee

@@
 need for a real-time calibration method for direct imaging radio astronomy correlators. Direct 
-imaging involves a spatial fast Fourier transform of antenna voltages, alleviating the harsh 
+imaging involves a spatial fast Fourier transform of antenna signals, alleviating the harsh 
 $\mathcal{O}(\Nant^2)$ computational scaling to a more gentle $\mathcal{O}(\Nant \log_2 
@@
 consisting of hundreds to thousands of small, widefield antennas. Many arrays with this design 
-are already built or under construction including the Hydrogen Epoch of Reionization Array
-\footnote{http://reionization.org} (HERA), the Murchison Widefield Array (MWA; 
+are already built or under construction including the Hydrogen Epoch of Reionization 
+Array\footnote{http://reionization.org} (HERA), the Murchison Widefield Array (MWA; 
 \citealt{tin13,bow13}), the Donald C. Backer Precision Array for Probing the Epoch of 
@@
 
-Traditional radio correlators cross-multiply the voltage signals from all pairs of antennas, and 
+Traditional radio correlators cross-multiply the signals from all pairs of antennas, and 
 the computation scales as the number of antennas squared, $\mathcal{O}(\Nant^2)$ 
@@
 with short timescale science. This is done by performing a spatial fast Fourier transform (FFT) 
-to image the antenna voltages, then squaring and averaging in time. This process scales as 
+to image the antenna signals, then squaring and averaging in time. This process scales as 
 $\mathcal{O}(\Ng \log_2 \Ng)$, where \Ng~is the number of grid points in the FFT \citep{mor11,
@@
  generation cosmology experiments, this scaling is a large improvement over the $\Nant^2$ 
- scaling of traditional methods. Furthermore, because images are generated online, the native 
+ scaling of traditional methods. Furthermore, because images are generated in real-time, the native 
  output bandwidth will be lowered in cases where $\Ng < \Nant^2$, and has the potential to be 
- lowered even further with online transient processing.
+ lowered even further with real-time transient processing.
 
@@
 from one time series is used to update the gains which are applied to the following time series. 
-An example implementation of the algorithm is available with the EPIC software package
-\footnote{http://github.com/nithyanandan/EPIC}.
+An example implementation of the algorithm is available with the EPIC software 
+package\footnote{http://github.com/nithyanandan/EPIC}.
 
 We review the MOFF algorithm and derive the calibration algorithm in \S \ref{sec:math}. We 
-then demonstrate the algorithm in simulations in \S \ref{sec:sim}, and apply to a sample LWA 
-data set in \S \ref{sec:data}. Then we discuss errors and noise properties of the resulting gain 
-solutions in \S \ref{sec:noise}. Finally we conclude and discuss potential extensions to the 
+then demonstrate the algorithm in simulations in \S \ref{sec:sim}, 
+and discuss noise trends in \S \ref{sec:noise}. We then apply the algorithm to a sample LWA 
+data set in \S \ref{sec:data}.
+Finally we conclude and discuss potential extensions to the 
 algorithm in \S \ref{sec:discussion}.
@@
 F-engine. The calibration is applied, then the electric field measurements are gridded using the 
-antenna illumination patterns. The imager performs a spatial 2D FFT on the gridded data before 
+antenna aperture illumination patterns. The imager performs a spatial 2D FFT on the gridded data before 
 squaring and averaging. The accumulated images are then written to disk.}
@@
 aperture illumination pattern (sometimes referred to as the aperture distribution), $W_a(\mathbf{r}-\mathbf{r}_a)=\int  \mathcal{W}_a(\hat{\mathbf{s}}) \, d
-\Omega$. Although any antenna extends over a region of space, we include a nominal antenna 
-position, $\mathbf{r}_a$, for multiple reasons. In the case where all antennas of an array are 
-identical, a global aperture distribution can be adopted and the translation of the antennas 
-distinguishes them. However, even if all antennas have unique aperture distributions, it can be 
-convenient to retain the relative positions of antennas in the analysis. This allows for easy 
-comparison with visibility-based analyses where the baseline vectors are used to describe 
-sampling of the $uv$ plane. From a practical point of view, measurements of the antenna 
-radiation pattern may not be performed in-situ, so it is convenient to describe the antenna 
-aperture distribution relative to a point fixed within the aperture, along with a translational 
-position.
+\Omega$. 
+While it is possible to define each antenna aperture illumination pattern in a way to absorb the 
+relative position of the antenna, we keep the $\mathbf{r}_a$ coordinate to retain the separable
+concepts of the primary beam (the product of two antenna radiation patterns) and the array
+point spread function (square of the Fourier transform of the array layout).
 
@@
 The extent of the grid is chosen to encompass the footprint of the antenna array.
-MOFF uses the antenna aperture distributions as the gridding 
+MOFF uses the antenna illumination patterns as the gridding 
 kernel according to the software holography/A-transpose technique \citep{mor09,bha08}. 
@@
 \begin{equation}
-\mathcal{C}^{(n)}_{a,0} \rightarrow \frac{1}{\Nant}\sum_b \frac{1}{g^{*(n)}_b} \V_{ab}
+\mathcal{C}^{(n)}_{a,0} \rightarrow \frac{1}{\Nant}\sum_b \frac{1}{g^{*(n)}_b} V_{ab}
 \end{equation}
@@
 \citealt{mor11}. We have confirmed the limiting case equivalence here, and will explore the 
-more general case in more detail in \S~\ref{sec:noise}.
+more general case in more detail in the following sections.
 
@@
 error. The other 50 antennas are shown to have error spanning $2\pi$ initially, and after about 
-10 iterations lock into a solution, settling down to noise levels around iteration 15 (0.15~s). We 
+10 iterations lock into a solution, settling down to noise levels around iteration 12 (0.12~s). We 
 stop the simulation when the updated gains trace the thermal noise of the simulated sources, 
-which can be seen by the coherence of the 50 antenna gains after iteration 15.
+which can be seen by the coherence of the 50 antenna gains after iteration 12.
 
@@
 gains were initialized with random phases, but the calibration loop was able to recover the 
-correct phases after about 15 iterations. Each line represents an antenna in our 51 MWA 
+correct phases after about 12 iterations. Each line represents an antenna in our 51 MWA 
 antenna sample.
@@
 quantity plotted is the magnitude of the estimated gains over the true gains, $\left|g^{(n)}_a/
-g^T_a\right|$, which places all antennas on the same scale. We can see the amplitudes 
-converge toward their true values around the same time as the phases (iteration $\sim$15). At 
+g_a\right|$, which places all antennas on the same scale. We can see the amplitudes 
+converge toward their true values around the same time as the phases (iteration $\sim$12). At 
 the beginning of calibration we can see the importance of the damping factor. At $n=0$, a couple of 
@@
 each line represents an antenna in the 51 MWA antenna sample. The gain estimates were 
-initialized randomly, while the true values were unity. After about 15 iterations we see the 
+initialized randomly, while the true values were unity. After about 12 iterations we see the 
 calibration loop has settled around the correct values, with only noise remaining.}
@@
 After the gains have converged, we see both the phases and amplitudes to continue to 
-fluctuate coherently. This is due to the stochastic fluctuations of the simulated sources 
-themselves, and is limited by the sky noise. For this simulation we restricted each calibration 
+fluctuate coherently. This is due to the stochastic fluctuations of the sources 
+themselves, which are simulated as zero mean random processes with variance proportional
+to their flux density.
+For this simulation we restricted each calibration 
 iteration to only 10~ms integration time. In principle a calibration implementation could use 
@@
 Images created at the beginning of calibration and at the end are shown in 
-figure~\ref{fig:sim_images}. Each image is obtained over 10~ms integration, corresponding to 
+figure~\ref{fig:sim_images}. Each image is obtained over 400 time sample integration, corresponding to 
 all snapshot images created with a given set of gain estimates. The left panel shows the image 
 produced with our initialized unity gains. Because the phases are completely random, the 
-image is essentially noise with the primary beam evident. After 20 iterations, the image is far 
+image is essentially noise with the primary beam evident. After 20 calibration iterations, the image is far 
 more clear, shown in the middle panel. 
@@
 \caption{Images formed during simulated calibration. \emph{Left:} An image generated from a 
-single frequency channel and 10 ms integration after the gain estimates are randomized. As 
+single frequency channel and 400 time sample integration after the gain estimates are randomized. As 
 expected with random phases, the image is completely noisy, with the shape of the primary 
 beam evident. \emph{Middle:} An image formed after calibration, again with a single frequency 
-channel and 10 ms integration. Now the simulated point sources are visible in the image. 
+channel and 400 time sample integration. Now the simulated point sources are visible in the image. 
 \emph{Right:} An image generated from the same simulated data as the middle image, but 
@@
 \section{Noise trends}\label{sec:noise}
-Now next study the effect of noise in our system, and the consequences 
+Next we study the effect of noise in our system, and the consequences 
 of an incomplete sky model. We run a suite of simulations varying the receiver noise and 
@@
 cases, where the receiver noise power is subdominant to the sky power ($\sigma_r = 
-10.0$~Jy), and where the receiver dominates the noise ($\sigma_r=100.0$~Jy).
+10.0$~Jy), and where the receiver dominates the total power ($\sigma_r=100.0$~Jy).
 
@@
 measurements (including the receiver noise). The electric field measurements are correlated for a duration equal to 
-the effective integration time of the EPICal loop for comparison. We then find the visibility-
-based gain estimates by minimizing
+the effective integration time of the EPICal loop for comparison. We then find the 
+visibility-based gain estimates by minimizing
 \begin{equation}\label{eq:vis_cal}
@@
 calibration. Through our previous imaging work, we noted that the flux scale of uncalibrated 
-images was consistent with average gain amplitudes of 0.25. To allow the calibration to 
+images was consistent with average gain amplitudes of roughly 0.25. To allow the calibration to 
 converge quickly we initialized our gain estimates at this level. 
@@
 (modulo noise), and exhibit a central ``trunk" where the majority of phases are congregated. 
-This behavior is suggestive that while the uncalibrated voltages were able to produce a viable 
+This behavior is suggestive that while the uncalibrated data were able to produce a viable 
 image, the minor changes from our solutions will focus the image and improve the quality. The 
@@
 
-To estimate our fractional gain noise we calculate the standard deviation of the gain
+To estimate the fractional gain noise of our solutions we calculate the standard deviation of the gain
 estimates after iteration 15 for each antenna separately. We then divide these standard
@@
 very simple sky model (for LWA data). In principle the images produced by the EPIC correlator 
-can be used to improve the sky model used in calibration -- similar to a major loop in self-
-calibration. This can be especially useful for compact, widefield arrays which require a model of 
+can be used to improve the sky model used in calibration -- similar to a major loop in 
+self-calibration. This can be especially useful for compact, widefield arrays which require a model of 
 both compact and diffuse sources over a large patch of sky. In addition to requiring the sky 

### 2016-06-21 imaging_demo/ms.tex
Make few changes addressing referee comments
From 2add9d8 (mnras):
====================

In ms.tex,

(*) Make few changes addressing referee comments

--- 2015-11-08 imaging_demo/ms.tex
@@
 digitized by the A/D converter is Fourier transformed to generate electric 
-field spectra. This step can be parallelized across antennas as shown in 
-Fig.~\ref{fig:f-engine}. The output is then fed to either MOFF or 
+field spectra. This step can be parallelized across antennas%  as shown in 
+% Fig.~\ref{fig:f-engine}
+. The output is then fed to either MOFF or 
 visibility-based imaging pipelines.
 
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figure4}
-  % \includegraphics[width=\columnwidth]{parallel_F_engine}
-  % \includegraphics[width=\columnwidth]{parallel_F_engine.eps}
-  \caption{Block diagram of a F-engine. The electric field data streams from
-    antennas are Fourier transformed in parallel to generate electric field
-    spectra.}
-  \label{fig:f-engine}
-\end{figure}
+% \begin{figure}
+%   \includegraphics[width=\columnwidth]{figure4}
+%   % \includegraphics[width=\columnwidth]{parallel_F_engine}
+%   % \includegraphics[width=\columnwidth]{parallel_F_engine.eps}
+%   \caption{Block diagram of a F-engine. The electric field data streams from
+%     antennas are Fourier transformed in parallel to generate electric field
+%     spectra.}
+%   \label{fig:f-engine}
+% \end{figure}
 
@@
 
-In a companion paper to appear soon, we demonstrate a novel calibration 
+In a companion paper \citep{bea16}, we demonstrate a novel calibration 
 technique (EPICal) which leverages the data products formed by direct imaging 

### 2016-06-21 cal_demo/ms.tex
More fine tuning as I re-read referee reports. Also updating .pdf

@@
 
-Traditional radio correlators cross-multiply the signals from all pairs of antennas, and 
+The most common radio correlator designs of today, the FX and lag correlators \citep{rom99},
+cross-multiply the signals from all pairs of antennas, and 
 the computation scales as the number of antennas squared, $\mathcal{O}(\Nant^2)$ 
@@
 traditional cross correlator, pair-wise visibilities are written to disk and used to calibrate 
-offline before further processing such as imaging. However, a direct imaging correlator mixes 
+after observing before further processing such as imaging. However, a direct imaging correlator mixes 
 the signals from all antennas before averaging and writing to disk, making calibration a 
@@
 
-Thus far we have assumed the antennas perfectly measure the sky electric field distribution according to their far-field radiation patterns.
-Here we introduce a multiplicative complex gain as well as an additive noise term which corrupt the measured signals.
+Thus far we have considered the perfectly measured electric field distribution according to 
+the antenna far-field radiation patterns. Here we introduce a multiplicative complex gain as 
+well as an additive noise term which corrupt the measured signals.
 \begin{equation}\label{eq:apply_gain}
@@
 \begin{align}\label{eq:epix}
-\mathcal{E}(\s_k,t) & = \frac{1}{\Nant} \sum_a \frac{1}{g'_a}\left(g_a E^I_a(t)+n_a(t)\right) e^{2\pi i \s_k \cdot \ra} \nonumber \\
+\mathcal{E}'(\s_k,t) & = \frac{1}{\Nant} \sum_a \frac{1}{g'_a}\left(g_a E^I_a(t)+n_a(t)\right) e^{2\pi i \s_k \cdot \ra} \nonumber \\
   & \qquad\times \sum_j W_a(\mathbf{r}_j-\ra)e^{2\pi i \s_k \cdot (\mathbf{r}_j-\ra)} \nonumber\\
@@
 
-Finally, the images are squared and averaged in time to reduce output bandwidth, then written 
+Finally, the images are squared and averaged in time to form dirty images, 
+$I'(\hat{\mathbf{s}}_k)$, then written 
 to disk. While operations have been reordered, the resulting image is equivalent to one 

### 2016-06-22 cal_demo/ms.tex
point to mnras.cls in parent directory
Added "aperture" before all cases of "illumination pattern"

@@
 The extent of the grid is chosen to encompass the footprint of the antenna array.
-MOFF uses the antenna illumination patterns as the gridding 
+MOFF uses the antenna aperture illumination patterns as the gridding 
 kernel according to the software holography/A-transpose technique \citep{mor09,bha08}. 
@@
 \end{equation}
-Formally each pixel is calculated by summing over all antennas and their illumination patterns.
-However the illumination patterns are typically compact and non-overlapping, so most pixels
+Formally each pixel is calculated by summing over all antennas and their aperture illumination patterns.
+However the aperture illumination patterns are typically compact and non-overlapping, so most pixels
 will only contain contribution from at most one antenna, allowing equation~\ref{eq:gridding}
@@
 \end{align}
-The sum over antennas uses the antenna illumination pattern to grid the calibrated electric field measurements onto 
+The sum over antennas uses the antenna aperture illumination pattern to grid the calibrated electric field measurements onto 
 regular gridpoints, $\mathbf{r}_j$. The sum over $j$ denotes the 2D FFT to sky coordinates, resulting in an 
@@
 exchanging the sums and use the DFT to approximate the Fourier relationship between
-the antenna illumination pattern and the far-field radiation pattern.
+the antenna aperture illumination pattern and the far-field radiation pattern.
 \begin{align}\label{eq:epix}
@@
 \end{align}
-In this form we see that the effect of gridding with the antenna illumination pattern is to attenuate 
+In this form we see that the effect of gridding with the antenna aperture illumination pattern is to attenuate 
 the image by a factor of the far-field radiation response, $\mathcal{W}_a(\s)$.
@@
 array consists of the inner 51 antennas of the MWA layout \citep{bea12}, within a bounding 
-box of 150~m. The antenna illumination pattern used is a 4.4~m square tophat; a very rough 
+box of 150~m. The antenna aperture illumination pattern used is a 4.4~m square tophat; a very rough 
 approximation to the square shaped tiles of the MWA. Because our algorithm treats frequency 

### 2016-06-22 cal_demo/ms.tex
Made references to figures consistent

@@
 We show schematically the process of calibrating a direct imaging correlator in 
-figure~\ref{fig:schematic}. Computationally expensive steps that must be performed in near real-time 
+Fig.~\ref{fig:schematic}. Computationally expensive steps that must be performed in near real-time 
 are shown inside the gray box. The uncalibrated antenna signals are tapped out after the 
@@
 The calibration loop continues by updating the gain estimates every \caliter~time samples. The 
-phases of our gain estimates are shown in figure~\ref{fig:sim_phase} for 20 such iterations. 
+phases of our gain estimates are shown in Fig.~\ref{fig:sim_phase} for 20 such iterations. 
 The phase error plotted is the phase relative to the true gain for each antenna (various colored 
@@
 
-The estimated gain amplitudes for the simulations are shown in figure~\ref{fig:sim_amp}. The 
+The estimated gain amplitudes for the simulations are shown in Fig.~\ref{fig:sim_amp}. The 
 quantity plotted is the magnitude of the estimated gains over the true gains, $\left|g^{(n)}_a/
@@
 Images created at the beginning of calibration and at the end are shown in 
-figure~\ref{fig:sim_images}. Each image is obtained over 400 time sample integration, corresponding to 
+Fig.~\ref{fig:sim_images}. Each image is obtained over 400 time sample integration, corresponding to 
 all snapshot images created with a given set of gain estimates. The left panel shows the image 
@@
 
-The results of our simulations are shown in Figure~\ref{fig:errors}. The gain errors from our 
+The results of our simulations are shown in Fig.~\ref{fig:errors}. The gain errors from our 
 EPICal method are shown with solid lines, while the dashed lines represent the visibility-based 
@@
 
-When using the single point source model, all calibration methods in figure~\ref{fig:errors} trend 
+When using the single point source model, all calibration methods in Fig.~\ref{fig:errors} trend 
 downward until they reach an error floor due to the confusion sources that were not modeled. 
@@
 update our gain estimates over 10 samples (51.2 ms). We iterate the loop 30 times for a 
-total of 1.536 seconds. The results of this calibration experiment are shown in figures~
+total of 1.536 seconds. The results of this calibration experiment are shown in Figs.~
 \ref{fig:data_phase}~--~\ref{fig:data_images}.
@@
 The gain amplitudes as a function of calibration iteration are shown in 
-figure~\ref{fig:data_amp}. There is a fairly wide range in gain amplitudes (from 0.12 to 0.59). 
+Fig.~\ref{fig:data_amp}. There is a fairly wide range in gain amplitudes (from 0.12 to 0.59). 
 However, this is not surprising due to the non-uniformity in the cables from the antennas to the 
@@
 To ensure the level of noise in our solutions is reasonable we compared to the fractional errors
-from our simulations shown in Figure~\ref{fig:errors}. We first use the Global Sky Model
+from our simulations shown in Fig.~\ref{fig:errors}. We first use the Global Sky Model
 \citep{deo08} to generate a mock observed sky at 74.03 MHz using the PyGSM python 
@@
 For reference we show the GSM in the right panel of 
-Figure~\ref{fig:data_images}, convolved by the LWA point spread function and weighted
+Fig.~\ref{fig:data_images}, convolved by the LWA point spread function and weighted
 by  two factors of the primary beam to match the holographic frame of the data images.
@@
 considerations for further study. These topics can be thought of as extensions to the white 
-``estimate gains" box in figure~\ref{fig:schematic}, which can be performed at a much lower 
+``estimate gains" box in Fig.~\ref{fig:schematic}, which can be performed at a much lower 
 cadence than the actual correlations.

### 2016-06-22 cal_demo/ms.tex
Minor fixes when rereading text

@@
 \begin{equation}
-\mathcal{C}^{(n)}_{a,0} \rightarrow \frac{1}{\Nant}\sum_b \frac{1}{g^{*(n)}_b} V_{ab}
+\mathcal{C}^{(n)}_{a,0} \rightarrow \frac{1}{\Nant}\sum_{b\ne a} \frac{1}{g^{*(n)}_b} V_{ab}
 \end{equation}
@@
 \begin{equation}
-g^{(n+1)}_a \rightarrow \frac{\sum_b  V_{ab}/g^{*(n)}_b}{\Nant S_{\mathrm{src}}}
+g^{(n+1)}_a \rightarrow \frac{\sum_{b\ne a}  V_{ab}/g^{*(n)}_b}{\Nant S_{\mathrm{src}}}
 \end{equation}
@@
 For our unknown gains, we create a set of random complex numbers where the amplitude is 
-drawn from approximately gaussian distribution centered around 1 with width 0.25, and 
+drawn from an approximately gaussian distribution centered around 1 with width 0.25, and 
 completely random phase. These are our ``true gains", and we apply them to the frequency
@@
 Images created at the beginning of calibration and at the end are shown in 
-Fig.~\ref{fig:sim_images}. Each image is obtained over 400 time sample integration, corresponding to 
+Fig.~\ref{fig:sim_images}. Each image is obtained over 400 time sample (10 ms) integration, corresponding to 
 all snapshot images created with a given set of gain estimates. The left panel shows the image 
@@
 \caption{Images formed during simulated calibration. \emph{Left:} An image generated from a 
-single frequency channel and 400 time sample integration after the gain estimates are randomized. As 
+single frequency channel and 400 time sample (10 ms) integration after the gain estimates are randomized. As 
 expected with random phases, the image is completely noisy, with the shape of the primary 
@@
 Figure~\ref{fig:data_images} shows the improvement in the images due to our calibration. The 
-left panel shows the uncalibrated image integrated over 51.2 ms, 29 kHz. Cyg~A is prominent 
+left panel shows the uncalibrated image integrated over 51.2 ms, 83 kHz. Cyg~A is prominent 
 near the center of the image, and Cas~A is also clearly visible in the upper right. The middle 
@@
 \caption{Images produced before (left) and after (middle) calibrating LWA data. These images 
-were produced with 29 kHz bandwidth and 51.2 ms integration. The calibrated image shows 
+were produced with 83 kHz bandwidth and 51.2 ms integration. The calibrated image shows 
 significant reduction in rumble throughout, while retaining the prominent Cyg~A and 
@@
 EPICal is poised to greatly contribute to their success. The software is integrated into the EPIC 
-package and freely available to use in simulations or offline processing of data. Work is 
+package and freely available to use in simulations or post-processing of data. Work is 
 underway to port the code to real-time GPU systems for deployment.

### 2016-06-22 cal_demo/ms.tex
Bolded changes in cal demo paper in prep for resubmission

@@
 need for a real-time calibration method for direct imaging radio astronomy correlators. Direct 
-imaging involves a spatial fast Fourier transform of antenna signals, alleviating the harsh 
+imaging involves a spatial fast Fourier transform of antenna \textbf{signals}, alleviating the harsh 
 $\mathcal{O}(\Nant^2)$ computational scaling to a more gentle $\mathcal{O}(\Nant \log_2 
@@
 
-The most common radio correlator designs of today, the FX and lag correlators \citep{rom99},
-cross-multiply the signals from all pairs of antennas, and 
+\textbf{The most common radio correlator designs of today, the FX and lag correlators \citep{rom99}},
+cross-multiply the \textbf{signals} from all pairs of antennas, and 
 the computation scales as the number of antennas squared, $\mathcal{O}(\Nant^2)$ 
@@
  generation cosmology experiments, this scaling is a large improvement over the $\Nant^2$ 
- scaling of traditional methods. Furthermore, because images are generated in real-time, the native 
- output bandwidth will be lowered in cases where $\Ng < \Nant^2$, and has the potential to be 
- lowered even further with real-time transient processing.
+ scaling of traditional methods. Furthermore, because images are generated \textbf{in real-time}, the native 
+ output bandwidth will be lowered \textbf{in cases where $\Ng < \Nant^2$}, and has the potential to be 
+ lowered even further with \textbf{real-time} transient processing.
 
@@
 A challenge common to all direct imaging algorithms is calibration of the antenna gains. With a 
-traditional cross correlator, pair-wise visibilities are written to disk and used to calibrate 
-after observing before further processing such as imaging. However, a direct imaging correlator mixes 
+traditional \textbf{cross correlator}, pair-wise visibilities are written to disk and used to calibrate 
+\textbf{after observing before further processing} such as imaging. However, a direct imaging correlator mixes 
 the signals from all antennas before averaging and writing to disk, making calibration a 
@@
 then demonstrate the algorithm in simulations in \S \ref{sec:sim}, 
-and discuss noise trends in \S \ref{sec:noise}. We then apply the algorithm to a sample LWA 
-data set in \S \ref{sec:data}.
+\textbf{and discuss noise trends in \S \ref{sec:noise}. We then apply the algorithm to a sample LWA 
+data set in \S \ref{sec:data}}.
 Finally we conclude and discuss potential extensions to the 
@@
 \caption{Data flow for the MOFF algorithm, reproduced from \citealt{thy15c}. The time ordered 
-electric field measurements are collected by the antenna elements then frequency transformed in the 
+electric field \textbf{measurements} are collected by the antenna elements then frequency transformed in the 
 F-engine. The calibration is applied, then the electric field measurements are gridded using the 
-antenna aperture illumination patterns. The imager performs a spatial 2D FFT on the gridded data before 
+\textbf{antenna aperture illumination patterns}. The imager performs a spatial 2D FFT on the gridded data before 
 squaring and averaging. The accumulated images are then written to disk.}
@@
 %%% New text
-We begin by considering the electric field vector incident on the ground as a function of time, 
+\textbf{We begin by considering the electric field vector incident on the ground as a function of time, 
 $\mathbf{E}(\mathbf{r},t)$. The position vector, $\mathbf{r}$, is measured from some arbitrary 
@@
 functions. We retain the time dependence because in practice each coefficient will only be 
-measured over some sampling period.
+measured over some sampling period. }
 
-It is convenient to simplify notation by expressing the position vector in units of wavelengths, 
+\textbf{It is convenient to simplify notation by expressing the position vector in units of wavelengths, 
 $\mathbf{r}\rightarrow \mathbf{r}/\lambda$. We will also assume all antennas lie in a plane, 
@@
 non-coplanar arrays through the W-projection algorithm \citep{cor08}, the EPIC code does not 
-currently support this so we leave it to future work.
+currently support this so we leave it to future work.}
 
-Next we treat the electric field as a scalar quantity, ignoring polarization effects. This allows us 
+\textbf{Next we treat the electric field as a scalar quantity, ignoring polarization effects. This allows us 
 to express the electric field as a scalar, $E(\mathbf{r},f,t)$, and treat subsequent field 
@@
 require a sky model expressed in the instrumental polarization basis. We can therefore treat 
-each polarization independently.
+each polarization independently.}
 
-Finally we assume the sources of the electric field are far away, allowing us to describe the 
+\textbf{Finally we assume the sources of the electric field are far away, allowing us to describe the 
 electric fields propagating from sky sources as originating from an imaginary celestial sphere 
@@
 By making the latter assumptions we have left treatment of the ionosphere and terrestrial 
-emission such as local interference or thermal radiation from the ground for future work.
+emission such as local interference or thermal radiation from the ground for future work. }
 
-Next we turn to the measurement performed by an antenna with far-field radiation pattern 
+\textbf{Next we turn to the measurement performed by an antenna with far-field radiation pattern 
 $\mathcal{W}_a(\hat{\mathbf{s}})$, where the subscript $a$ denotes a specific antenna. We 
@@
 concepts of the primary beam (the product of two antenna radiation patterns) and the array
-point spread function (square of the Fourier transform of the array layout).
+point spread function (square of the Fourier transform of the array layout).}
 
-%The electric field incident on the ground, $\Er{}(\mathbf{r},t)$, is related to the sky electric field, 
-%$E(\hat{\mathbf{s}},t)$, through a Fourier transform.
-%\begin{equation}
-%\Er{}(\mathbf{r},t) = \int E(\hat{\mathbf{s}},t) e^{-2\pi i \mathbf{r}\cdot \hat{\mathbf{s}}}\, \dif^2 \hat{\mathbf{s}}
-%\end{equation}
-%Here $\hat{\mathbf{s}}$ denotes the sine-projected unit vector for the sky angle, $\mathbf{r}$ is 
-%the observer's location (measured in wavelengths relative to an arbitrary origin). We will 
-%encounter several quantities which we attempt to estimate. We distinguish the ``true'' values 
-%with a superscript $T$, while the estimated quantities are denoted with a prime. We define the 
-%true antenna signal as a convolution of the antenna voltage pattern, \beamr, with the electric 
-%field on the ground.
-%\begin{equation}
-%\Er{a}^T(t) \equiv \int \beamr_a(\mathbf{r}-\ra) \Er{}(\mathbf{r},t) \, \dif^2 \mathbf{r}
-%\end{equation}
-%The subscript $a$ labels the antenna, and $\mathbf{r}_a$ is the location of antenna $a$.
-
-The first step of the MOFF algorithm is to digitize the antenna signals and perform an FFT to 
+\textbf{The first step of the MOFF algorithm is to digitize the antenna signals and perform an FFT to 
 channelize the time-ordered data. This is effectively estimating the quasi-monochromatic 
@@
 correlator which performs the frequency transform (F-engine) before cross multiplying 
-(X-engine).
+(X-engine). }
 
-%The first step of the MOFF algorithm is identical to the F-engine of a traditional FX correlator 
-%where the electric fields are Fourier transformed to frequency channels, $\Er{a}^T(f,t)$. Even 
-%after transforming along the time axis, we retain the time dependence of the signal to represent 
-%a stream of spectra.
-
-Thus far we have considered the perfectly measured electric field distribution according to 
-the antenna far-field radiation patterns. Here we introduce a multiplicative complex gain as 
+Thus far we have \textbf{considered the perfectly measured electric field distribution according to 
+the antenna far-field radiation patterns}. Here we introduce a multiplicative complex gain as 
 well as an additive noise term which corrupt the measured signals.
@@
 \end{equation}
-%The noise term is 
-%strictly receiver noise -- noise introduced by the instrument. Any sky noise is implicitly included 
-%in the time dependence of the sky electric field. 
-We have carried the frequency dependence to this point to make clear the complex gains are
-frequency dependent. However, each subsequent step treats each frequency 
+\textbf{We have carried the frequency dependence to this point to make clear the complex gains are
+frequency dependent. However,} each subsequent step treats each frequency 
 channel independently, so we will drop the $f$ to simplify notation. 
-Furthermore, while antenna gains may vary with time, we will assume they are constant over
-the period of our calibration, and thus drop their time dependence from the notation.
+\textbf{Furthermore, while antenna gains may vary with time, we will assume they are constant over
+the period of our calibration, and thus drop their time dependence from the notation.}
 
-The MOFF algorithm next calls for a calibration of the electric field measurements. The goal of our new calibration method will be 
-to form an estimate of the gains, $g'_a$, where the prime represents an estimate. For now we will assume we have formed an estimate 
+The MOFF algorithm next calls for a calibration \textbf{of the electric field measurements}. The goal of our new calibration method will be 
+to form an estimate of the gains, $g'_a$, \textbf{where the prime represents an estimate}. For now we will assume we have formed an estimate 
 to proceed with the MOFF pipeline. We correct the incoming electric field data stream using our 
@@
 
-The next step is to grid the calibrated electric field measurements to a finite set of regularly spaced 
+The next step is to grid the \textbf{calibrated electric field measurements to a finite set of regularly spaced 
 positions on the 2D plane of the ground, $\mathbf{r}_j$.
 The extent of the grid is chosen to encompass the footprint of the antenna array.
-MOFF uses the antenna aperture illumination patterns as the gridding 
+MOFF uses the antenna aperture illumination patterns as the} gridding 
 kernel according to the software holography/A-transpose technique \citep{mor09,bha08}. 
-Mathematically we can express the gridding operation as
+\textbf{Mathematically we can express the gridding operation as}
 \begin{equation}\label{eq:gridding}
@@
 \end{equation}
-Formally each pixel is calculated by summing over all antennas and their aperture illumination patterns.
+\textbf{Formally each pixel is calculated by summing over all antennas and their aperture illumination patterns.
 However the aperture illumination patterns are typically compact and non-overlapping, so most pixels
 will only contain contribution from at most one antenna, allowing equation~\ref{eq:gridding}
-to be calculated through a sparse matrix multiplication.
+to be calculated through a sparse matrix multiplication.}
 This gridding 
@@
 After gridding, the imaging portion of the MOFF performs a 2D spatial FFT to form 
-instantaneous electric field images of the sky at fixed locations $\s_k$.
+instantaneous electric field images of the sky \textbf{at fixed locations $\s_k$.
 The Fourier dual coordinates to $\mathbf{r}$ are $(l,m)$, where $l=\sin\theta\cos\phi$,
@@
 Fine $\mathbf{r}_j$ grid spacing can result in non-physical angular pixels ($l^2+m^2>1$),
-which should be discarded.
+which should be discarded.}
 
@@
 \end{align}
-The sum over antennas uses the antenna aperture illumination pattern to grid the calibrated electric field measurements onto 
-regular gridpoints, $\mathbf{r}_j$. The sum over $j$ denotes the 2D FFT to sky coordinates, resulting in an 
+The sum over antennas uses the \textbf{antenna aperture illumination pattern} to grid the calibrated electric field \textbf{measurements} onto 
+regular gridpoints, $\mathbf{r}_j$. The \textbf{sum over $j$} denotes the 2D FFT to sky coordinates, resulting in an 
 estimate for the instantaneous electric field image. We can simplify this expression by 
-exchanging the sums and use the DFT to approximate the Fourier relationship between
-the antenna aperture illumination pattern and the far-field radiation pattern.
+exchanging the sums \textbf{and use the DFT to approximate the Fourier relationship between
+the antenna aperture illumination pattern and the far-field radiation pattern.}
 \begin{align}\label{eq:epix}
@@
 \end{align}
-In this form we see that the effect of gridding with the antenna aperture illumination pattern is to attenuate 
-the image by a factor of the far-field radiation response, $\mathcal{W}_a(\s)$.
+In this form we see that the effect of gridding with the antenna \textbf{aperture illumination pattern} is to attenuate 
+the image by a factor of the \textbf{far-field radiation response, $\mathcal{W}_a(\s)$.}
 
-Finally, the images are squared and averaged in time to form dirty images, 
-$I'(\hat{\mathbf{s}}_k)$, then written 
+Finally, the images are squared and averaged in time to form \textbf{dirty images, 
+$I'(\hat{\mathbf{s}}_k)$, then} written 
 to disk. While operations have been reordered, the resulting image is equivalent to one 
@@
 As a starting point, we consider the feedback calibration outlined in \citealt{mor11}. There it 
-was suggested to form a correlation of the uncalibrated antenna measurements with the complex conjugate of an image 
+was suggested to form a correlation of the uncalibrated antenna measurements with the \textbf{complex conjugate of} an image 
 pixel from the output of the correlator. This is a statistically stationary quantity, which can be 
@@
 calibration loop, 
-the superscript $^*$ indicates complex conjugation,
+\textbf{the superscript $^*$ indicates complex conjugation,}
 and \spix\, is the pixel center nearest a bright calibrator of interest. The 
@@
 Plugging equation \ref{eq:epix} into equation \ref{eq:Kna_def}, 
-and substituting the $g^{(n)}_b$ for $g'_b$ to indicate the specific loop number,
+\textbf{and substituting the $g^{(n)}_b$ for $g'_b$ to indicate the specific loop number,}
 we find,
@@
 where in the second step we group time-dependent terms, 
-and in the third step we expanded the product within the time average.
+and in the third \textbf{step we expanded the product within the time average.
 Because the noise does not correlate with the ideal antenna measurements we can
@@
 is zero mean. Then by defining an ideal visibility as
-$V_{ab}^I \equiv \left<E^I_a(t)E^{*I}_b(t)\right>_t$, we can simplify to
+$V_{ab}^I \equiv \left<E^I_a(t)E^{*I}_b(t)\right>_t$, we can simplify to}
 \begin{align}\label{eq:Kna}
@@
 \end{align}
-where $\delta_{ab}$ is the Kronecker-delta function. To be clear, visibilities are never directly 
+\textbf{where $\delta_{ab}$ is the Kronecker-delta function. To be clear, visibilities are never directly 
 formed from the data. Rather it is convenient to express $\Kna$ in terms of the ideal 
-visibilities, which we can model when solving for the gains below.
+visibilities, which we can model when solving for the gains below.}
 
-We will ultimately wish to model the right side of equation~\ref{eq:Kna} in order to solve for the
+\textbf{We will ultimately wish to model the right side of equation~\ref{eq:Kna} in order to solve for the
 antenna gains. However, the self-correlated noise, $\left<|n_a(t)|^2\right>$, can often be 
@@
 uncalibrated antenna electric field measurements to remove the corresponding
-terms from equation~\ref{eq:Kna}.
+terms from equation~\ref{eq:Kna}.}
 \begin{align}\label{eq:Cna}
@@
 the right hand side of equation~\ref{eq:Cna} by assuming our current gain estimates are 
-approximately correct, $g^{(n)}_b\approx g_b$, and replacing the ideal visibilities with a set of 
+approximately correct, $g^{(n)}_b\approx g_b$, and replacing the \textbf{ideal} visibilities with a set of 
 model visibilities, $V^M_{ab}$. These model visibilities rely on knowledge of the sky, and should 
-be precomputed before observing. We can then solve equation~\ref{eq:Cna} for $g_a$ to achieve an 
+be precomputed \textbf{before observing}. We can then solve equation~\ref{eq:Cna} for $g_a$ to achieve an 
 updated estimate.
 \begin{equation}\label{eq:cal_solution}
-%g^{(n+1)}_a = \Cna \Nant \left[ \sum_{b\ne a} \mathcal{W}^*_b(\spix) e^{-2\pi i \spix \cdot \rb} V^M_{ab} \right]^{-1}
 g^{(n+1)}_a = \frac{\Cna \Nant }{ \sum_{b\ne a} \mathcal{W}^*_b(\spix) e^{-2\pi i \spix \cdot \rb} V^M_{ab} }
@@
 This equation is our prescription for estimating the antenna gains of a direct imaging array. The 
-real-time computation complexity scales as $\mathcal{O}(\Nant)$ as we form 
-an antenna-pixel correlation, $\Cna$, for each 
+\textbf{real-time} computation complexity scales as $\mathcal{O}(\Nant)$ as we form 
+\textbf{an antenna-pixel correlation,} $\Cna$, for each 
 antenna.
@@
 We show schematically the process of calibrating a direct imaging correlator in 
-Fig.~\ref{fig:schematic}. Computationally expensive steps that must be performed in near real-time 
+Fig.~\ref{fig:schematic}. Computationally expensive steps that must be performed \textbf{in near real-time} 
 are shown inside the gray box. The uncalibrated antenna signals are tapped out after the 
 F-engine and correlated against the output image pixel of interest (i.e., equation~\ref{eq:Cna} is 
-computed in near real-time). The correlated values are then used to estimate the gains
-using equation~\ref{eq:cal_solution}, and additional fitting if desired. This step does not have
+\textbf{computed in near real-time}). The correlated values are then \textbf{used to estimate the gains}
+using equation~\ref{eq:cal_solution}, and additional fitting if desired. \textbf{This step does not have
 the speed requirement of the steps in the gray box because the correlations have been 
-averaged in time.
+averaged in time.}
 The gains are then passed back 
@@
 from the (unsquared) image is tapped out and correlated against the input antenna electric field 
-measurements to form $\Cna$ coefficients (equation~\ref{eq:Cna}). These coefficients are then 
-used to update the gain estimates using equations~\ref{eq:cal_solution}~and~\ref{eq:damping},
-and any additional fitting.
+\textbf{measurements} to form $\Cna$ coefficients (equation~\ref{eq:Cna}). These coefficients are then 
+\textbf{used to update the gain estimates using} equations~\ref{eq:cal_solution}~and~\ref{eq:damping},
+and \textbf{any additional fitting.}
 The gray box shows operations which must be done at high speed (before averaging in time),
-while the white boxes show operations which can be performed at lower speeds because
-they involve time averages quantities.
+\textbf{while the} white boxes show operations which can be performed \textbf{at lower speeds because
+they involve time averages quantities.}
 }
@@
 
-While testing we found equation~\ref{eq:cal_solution} resulted in oscillatory gain solutions over time, 
+While testing we found equation~\ref{eq:cal_solution} resulted in oscillatory gain solutions \textbf{over time,} 
 as is often the case in iterative methods. To mitigate this we 
@@
 We conclude this section by connecting our calibration expression to that found in a visibility 
-framework. In the limit of a single bright calibrating source at center, we can greatly 
-simplify equations~\ref{eq:Cna} and~\ref{eq:cal_solution}. We will assume the radiation patterns are 
+framework. In the limit of a single bright calibrating source \textbf{at center,} we can greatly 
+simplify equations~\ref{eq:Cna} and~\ref{eq:cal_solution}. We will assume the \textbf{radiation patterns} are 
 normalized such that $\mathcal{W}(0)=1$. We can further drop the exponential phase terms 
-because $\spix$ is perpendicular to our planar antenna array. We then absorb the true gains into the true visibilities in 
+because $\spix$ \textbf{is perpendicular to our planar antenna array.} We then absorb the true gains into the true visibilities in 
 equation~\ref{eq:Cna} to express as a sum of measured, uncalibrated, visibilities.
@@
 The ability to recover the equivalent expression despite not actually forming the visibilities is a 
-result of the fact that only sums over visibilities come into the visibility-based solution, as was described in 
+result of the fact that only sums over visibilities come into the \textbf{visibility-based} solution, as was described in 
 \citealt{mor11}. We have confirmed the limiting case equivalence here, and will explore the 
-more general case in more detail in the following sections.
+more general case in more detail in \textbf{the following sections.}
 
@@
 
-The physical parameters of our simulation can be arbitrarily scaled to other cases. However, 
+\textbf{The physical parameters of our simulation can be arbitrarily scaled to other cases. However, 
 we provide typical units for an MWA observation as an example. Our simulated antenna 
-array consists of the inner 51 antennas of the MWA layout \citep{bea12}, within a bounding 
-box of 150~m. The antenna aperture illumination pattern used is a 4.4~m square tophat; a very rough 
-approximation to the square shaped tiles of the MWA. Because our algorithm treats frequency 
-channels independently, we simulate only one channel. For context we treat this as a channel
+array consists of} the inner 51 antennas of the MWA layout \citep{bea12}, within a bounding 
+box of 150~m. The \textbf{antenna aperture illumination pattern} used is a 4.4~m square tophat; \textbf{a very rough 
+approximation to the square shaped tiles of the MWA.} Because our algorithm treats frequency 
+channels independently, we simulate only one channel. For context we treat this as \textbf{a channel
 at 150~MHz with 40 kHz bandwidth, and adopt a sampling period of 25~$\mu$s. The simulated signal 
 consists of 10 random point sources with flux densitities 0.5~Jy~$\lesssim S \lesssim$~1~Jy 
-within the main lobe of the primary beam.
+within the main lobe of the primary beam.}
 
@@
 completely random phase. These are our ``true gains", and we apply them to the frequency
-domain simulated antenna electric field measurements as in equation~\ref{eq:apply_gain}. Our analysis is 
+domain simulated antenna electric field \textbf{measurements} as in equation~\ref{eq:apply_gain}. Our analysis is 
 blind to these values until the end of the process to check accuracy. The gain estimates are 
@@
 
-We next process and image \caliter~time samples (10~ms). We also form the correlations, 
+We next process and image \caliter~time \textbf{samples} (10~ms). We also form the correlations, 
 \Cna[0], used in our calibration loop. The pixel used for the correlation is the source with the 
 largest apparent flux (intrinsic flux attenuated by the primary beam). These correlation values 
-are used to update the gain estimates. For this simple demonstration we create perfect model 
+are used to update the gain estimates. \textbf{For this simple demonstration we create perfect model 
 visibilities of our 10 simulated sources to be used in equation~\ref{eq:damping}. 
-The updated gain estimates are used to calibrate the following 
+The updated gain estimates are} used to calibrate the following 
 \caliter~time samples. Through experimentation we found a damping factor of $\damp=0.35$ 
@@
 
-The calibration loop continues by updating the gain estimates every \caliter~time samples. The 
+The calibration loop continues by updating the gain estimates every \caliter~time \textbf{samples.} The 
 phases of our gain estimates are shown in Fig.~\ref{fig:sim_phase} for 20 such iterations. 
@@
 error. The other 50 antennas are shown to have error spanning $2\pi$ initially, and after about 
-10 iterations lock into a solution, settling down to noise levels around iteration 12 (0.12~s). We 
+10 iterations lock into a solution, settling down to noise levels around iteration \textbf{12 (0.12~s)}. We 
 stop the simulation when the updated gains trace the thermal noise of the simulated sources, 
@@
 gains were initialized with random phases, but the calibration loop was able to recover the 
-correct phases after about 12 iterations. Each line represents an antenna in our 51 MWA 
+correct phases after about \textbf{12} iterations. Each line represents an antenna in our 51 MWA 
 antenna sample.
@@
 g_a\right|$, which places all antennas on the same scale. We can see the amplitudes 
-converge toward their true values around the same time as the phases (iteration $\sim$12). At 
-the beginning of calibration we can see the importance of the damping factor. At $n=0$, a couple of 
+converge toward their true values around the same time as the phases (iteration \textbf{$\sim$12}). At 
+the beginning of calibration we can see the \textbf{importance} of the damping factor. At $n=0$, a couple of 
 gains are shown to have abnormally high amplitude estimates, notably one about 3.3 times its 
@@
 each line represents an antenna in the 51 MWA antenna sample. The gain estimates were 
-initialized randomly, while the true values were unity. After about 12 iterations we see the 
+initialized randomly, while the true values were unity. After about \textbf{12} iterations we see the 
 calibration loop has settled around the correct values, with only noise remaining.}
@@
 fluctuate coherently. This is due to the stochastic fluctuations of the sources 
-themselves, which are simulated as zero mean random processes with variance proportional
-to their flux density.
+themselves, \textbf{which are simulated as zero mean random processes with variance proportional
+to their flux density.}
 For this simulation we restricted each calibration 
@@
 Images created at the beginning of calibration and at the end are shown in 
-Fig.~\ref{fig:sim_images}. Each image is obtained over 400 time sample (10 ms) integration, corresponding to 
-all snapshot images created with a given set of gain estimates. The left panel shows the image 
+Fig.~\ref{fig:sim_images}. Each image is obtained over a 400 time sample \textbf{(10 ms)} integration, corresponding to 
+all snapshot images created with a given set of gain estimates. The \textbf{left} panel shows the image 
 produced with our initialized unity gains. Because the phases are completely random, the 
 image is essentially noise with the primary beam evident. After 20 calibration iterations, the image is far 
-more clear, shown in the middle panel. 
-The ten sources are visible, along with rumble throughout the image. By comparing with
+more clear, shown in the \textbf{middle} panel. 
+\textbf{The ten sources are visible, along with rumble throughout the image. By comparing with
 a perfectly calibrated image (right panel), we see that the rumble and the shapes of the
 point sources are matched between the two images. We therefore conclude that the 
-artifacts visible are dominated by the point spread function, rather than calibration errors.
+artifacts visible are dominated by the point spread function, rather than calibration errors.}
 
@@
 \caption{Images formed during simulated calibration. \emph{Left:} An image generated from a 
-single frequency channel and 400 time sample (10 ms) integration after the gain estimates are randomized. As 
+single frequency channel and \textbf{400 time sample (10 ms)} integration after the gain estimates are randomized. As 
 expected with random phases, the image is completely noisy, with the shape of the primary 
 beam evident. \emph{Middle:} An image formed after calibration, again with a single frequency 
-channel and 400 time sample integration. Now the simulated point sources are visible in the image. 
+channel and \textbf{400 time sample} integration. \textbf{Now the simulated point sources are visible in the image. 
 \emph{Right:} An image generated from the same simulated data as the middle image, but 
 with perfect calibration correction. Many of the features seen in the middle image are also
-seen here, confirming that they are point spread function artifacts. The ten simulated (and modeled) point sources are highlighted with red circles.
+seen here, confirming that they are point spread function artifacts. The ten simulated (and modeled) point sources are highlighted with red circles.}
 \label{fig:sim_images}
@@
 \section{Noise trends}\label{sec:noise}
-Next we study the effect of noise in our system, and the consequences 
+\textbf{Next we} study the effect of noise in our system, and the consequences 
 of an incomplete sky model. We run a suite of simulations varying the receiver noise and 
 integration times, while forming antenna-pixel correlations, $\Cna$, for EPICal gain solutions, 
-and simultaneously forming visibilities from the same electric field measurement streams to find visibility-based 
+and simultaneously forming visibilities from the same \textbf{electric field measurement} streams to find visibility-based 
 gain solutions for comparison.
@@
 these sources randomly, but kept them fixed for each run. 
-We simulate the same antenna layout as in Section~\ref{sec:sim}. 
-However, we increase the number of frequency channels simulated to 64.
+\textbf{We simulate the same antenna layout as in Section~\ref{sec:sim}. 
+However, we increase the number of frequency channels simulated to 64.}
 Because our calibration loop treats 
@@
 To simulate receiver noise, we add a gaussian distributed complex random number to each 
-antenna electric field measurement at each time sample according to 
+antenna electric field measurement at each time \textbf{sample} according to 
 equation~\ref{eq:apply_gain}. The level of the receiver noise, $\sigma_r = \left<\left|
@@
 (${g^{(0)}_a=g_a=1}$), and allow the estimate to be corrupted by the noise through ten 
-iterations of the calibration loop. We again adopt a damping factor $\damp=0.35$. Because EPICal 
+iterations of the calibration loop. We \textbf{again} adopt a damping factor $\damp=0.35$. Because EPICal 
 updates the gain estimate at each calibration loop, \tcal, but retains a memory of previous 
@@
 
-We simultaneously form simulated visibilities, $V_{ab}$, by correlating all pairs of antenna electric field 
-measurements (including the receiver noise). The electric field measurements are correlated for a duration equal to 
+We simultaneously form simulated visibilities, $V_{ab}$, by correlating all pairs of antenna \textbf{electric field 
+measurements} (including the receiver noise). The \textbf{electric field measurements are correlated} for a duration equal to 
 the effective integration time of the EPICal loop for comparison. We then find the 
@@
 sky model. However, we see that with this perfect model the errors do behave like noise
-in the sense that they integrate down with longer integrations with the expected $t^{-1/2}$ 
-trend. 
-The same level of gain error as visibility based solutions can be achieved with a longer integration time (or larger damping 
+\textbf{in the sense that they integrate down with longer integrations with the expected $t^{-1/2}$ 
+trend. }
+The same level of \textbf{gain error as visibility based solutions} can be achieved with a longer integration time (or larger damping 
 factor).
@@
 We next demonstrate our calibration algorithm using an observation from the LWA station in 
-New Mexico \citep{ell13}. The data is from the LWA narrow-band transient buffer (TBN), with time ordered 
-voltage data from the 255 core antennas within a 110~m x 100~m ellipse. 
+New Mexico \textbf{\citep{ell13}.} The data is from the LWA narrow-band transient buffer (TBN), with time ordered 
+voltage data from the 255 core antennas within a \textbf{110~m x 100~m ellipse. }
 The central frequency is 74.03 
-MHz, with a bandwidth of 100 kHz and sampling period of 5.12 ms (frequency channel 
+MHz, with a bandwidth of 100 kHz and \textbf{sampling period} of 5.12 ms (frequency channel 
 resolution of 195.3125 Hz). For this demonstration we limit ourselves to a single polarization.
@@
 We proceed by forming model visibilities. We model only two bright objects as point sources: 
-Cyg A with flux 20,539 Jy, and Cas A with flux 19,328 Jy \citep{lan12}. 
+Cyg A with flux \textbf{20,539 Jy, and Cas A with flux 19,328 Jy \citep{lan12}. }
 Because the raw data is attenuated by the primary beam of the instrument, we also account for 
@@
 converge quickly we initialized our gain estimates at this level. 
-We also found a boost in signal to noise is achieved easily by assuming the gains are
-constant across a range of frequencies, and averaging solutions across channels.
-Here we average solutions across 300 channels, or about 58.6 kHz. 
+We also found a boost in signal to noise is achieved easily by \textbf{assuming the gains are
+constant across a range of frequencies, and averaging solutions across channels.}
+Here we average solutions across \textbf{300} channels, or about \textbf{58.6 kHz.} 
 With a fractional bandwidth $B/f_0 = 7.9 \times 10^{-4} \ll 1$, we assume a smooth bandpass 
@@
 A damping factor of $\gamma = 0.7$ was adopted.
-The large damping factor, relative to what was used in simulations, was necessary due to
-the larger noise levels in the LWA data compared to the ideal simulations.
+\textbf{The large damping factor, relative to what was used in simulations, was necessary due to
+the larger noise levels in the LWA data compared to the ideal simulations.}
 
@@
 and 244 were omitted.} produced unstable gain solutions, and in fact corrupted the entire array. 
-We therefore omitted these antennas from our analysis, resulting in a total of 248 antennas to 
+We therefore \textbf{omitted these antennas from our analysis,} resulting in a total of 248 antennas to 
 calibrate. In each calibration loop, we form $C_{i,\hat{\boldsymbol{s}}_0}, i=1,2,\ldots N_a$ and 
-update our gain estimates over 10 samples (51.2 ms). We iterate the loop 30 times for a 
+update our gain estimates over 10 \textbf{samples} (51.2 ms). We iterate the loop 30 times for a 
 total of 1.536 seconds. The results of this calibration experiment are shown in Figs.~
@@
 
-To ensure the level of noise in our solutions is reasonable we compared to the fractional errors
+\textbf{To ensure the level of noise in our solutions is reasonable we compared to the fractional errors
 from our simulations shown in Fig.~\ref{fig:errors}. We first use the Global Sky Model
@@
 should therefore expect our fractional gain noise levels to be at best around 0.1, the level 
-to which the simulations asymptotically approached.
+to which the simulations asymptotically approached. }
 
-To estimate the fractional gain noise of our solutions we calculate the standard deviation of the gain
+\textbf{To estimate the fractional gain noise of our solutions we calculate the standard deviation of the gain
 estimates after iteration 15 for each antenna separately. We then divide these standard
@@
 ($\sigma_g \approx 0.29$). Upon closer inspection the gain amplitudes of these outliers had
-converged later in the calibration loop, after iteration 15.
+converged later in the calibration loop, after iteration 15.}
 
@@
 Figure~\ref{fig:data_images} shows the improvement in the images due to our calibration. The 
-left panel shows the uncalibrated image integrated over 51.2 ms, 83 kHz. Cyg~A is prominent 
+left panel shows the uncalibrated image integrated over 51.2 ms, \textbf{83 kHz.} Cyg~A is prominent 
 near the center of the image, and Cas~A is also clearly visible in the upper right. The middle 
@@
 from the GSM. 
-For reference we show the GSM in the right panel of 
+\textbf{For reference we show the GSM in the right panel of 
 Fig.~\ref{fig:data_images}, convolved by the LWA point spread function and weighted
-by  two factors of the primary beam to match the holographic frame of the data images.
+by  two factors of the primary beam to match the holographic frame of the data images.}
 
@@
 \caption{Images produced before (left) and after (middle) calibrating LWA data. These images 
-were produced with 83 kHz bandwidth and 51.2 ms integration. The calibrated image shows 
-significant reduction in rumble throughout, while retaining the prominent Cyg~A and 
+were produced with \textbf{83 kHz} bandwidth and 51.2 ms integration. The calibrated image shows 
+significant reduction in \textbf{rumble} throughout, while retaining the prominent Cyg~A and 
 Cas~A sources. The galactic plane is also substantially more evident after calibration. For 
-reference, the GSM is shown in the 
+reference, the GSM is shown \textbf{in the 
 right panel convolved with the LWA point spread function and weighted by two factors
-of the primary beam to place it in the same holographic frame as the data images. While much of the GSM is visible in the calibrated image, we 
+of the primary beam to place it in the same holographic frame as the data images.} While much of the GSM is visible in the calibrated image, we 
 note that the model used to calibrate was only the two bright sources, Cyg~A and Cas~A.
@@
 peak of the image over the noise level. We estimate the noise level as the median of the 
-absolute deviation of the image, a robust statistic. With this metric we find the uncalibrated 
-and calibrated images to have dynamic ranges of 88.5 and 137, respectively, a 55\% 
-improvement.
+absolute deviation of the image, a robust statistic. With this metric we find the \textbf{uncalibrated 
+and} calibrated images to have \textbf{dynamic ranges of 88.5 and 137, respectively, a 55\% 
+improvement.}
 
@@
 Additionally, with knowledge of 
-any filters applied to the data stream in the receiver chain,
+\textbf{any filters applied to the data stream in the receiver chain,}
 we could include channels closer to the edge of the band.
@@
 EPICal is poised to greatly contribute to their success. The software is integrated into the EPIC 
-package and freely available to use in simulations or post-processing of data. Work is 
+package and freely available to use in simulations or \textbf{post-processing} of data. Work is 
 underway to port the code to real-time GPU systems for deployment.

### 2016-07-07 cal_demo/ms.tex
First round addressing Judds comments

@@
 need for a real-time calibration method for direct imaging radio astronomy correlators. Direct 
-imaging involves a spatial fast Fourier transform of antenna \textbf{signals}, alleviating the harsh 
-$\mathcal{O}(\Nant^2)$ computational scaling to a more gentle $\mathcal{O}(\Nant \log_2 
-\Nant)$, which can save orders of magnitude in computation cost for next generation arrays 
+imaging involves a spatial fast Fourier transform of antenna \textbf{signals}, alleviating 
+\textbf{an 
+$\mathcal{O}(\Nant^2)$ computational bottleneck typical in radio correlators, and yielding a more gentle $\mathcal{O}(\Nant \log_2 
+\Nant)$ scaling.} \textbf{This} can save orders of magnitude in computation cost for next generation arrays 
 consisting of hundreds to thousands of antennas. However, because signals are mixed in the 
-correlator, gain correction must be applied on the front end. We develop the EPICal algorithm 
+correlator, gain correction must be applied \textbf{prior to correlation, rather than on visibilities post-correlation.} We develop the EPICal algorithm 
 to form gain solutions in real time without ever forming visibilities. This method scales as the 
@@
 \textbf{The most common radio correlator designs of today, the FX and lag correlators \citep{rom99}},
-cross-multiply the \textbf{signals} from all pairs of antennas, and 
-the computation scales as the number of antennas squared, $\mathcal{O}(\Nant^2)$ 
+cross-multiply the \textbf{signals} from all pairs of \textbf{antennas.
+This} computation scales as the number of antennas squared, $\mathcal{O}(\Nant^2)$ 
 \citep{bun04}. As the number of elements in future arrays grows, the computational cost will 
@@
 traditional \textbf{cross correlator}, pair-wise visibilities are written to disk and used to calibrate 
-\textbf{after observing before further processing} such as imaging. However, a direct imaging correlator mixes 
+\textbf{after observing and before further processing} such as imaging. However, a direct imaging correlator mixes 
 the signals from all antennas before averaging and writing to disk, making calibration a 
@@
 understanding that the full field is composed of all components and their corresponding Fourier 
-functions. We retain the time dependence because in practice each coefficient will only be 
-measured over some sampling period. }
+functions. We retain the time dependence because in practice each Fourier component will only be 
+measured over discrete sampling intervals. }
 
-\textbf{It is convenient to simplify notation by expressing the position vector in units of wavelengths, 
+\textbf{It is convenient to simplify notation by expressing the position vector in units of wavelengths relative to the monochromatic component of interest, 
 $\mathbf{r}\rightarrow \mathbf{r}/\lambda$. We will also assume all antennas lie in a plane, 
@@
 require a sky model expressed in the instrumental polarization basis. We can therefore treat 
-each polarization independently.}
+each instrument polarization independently.}
 
@@
 $\mathcal{W}_a(\hat{\mathbf{s}})$, where the subscript $a$ denotes a specific antenna. We 
-adopt the common convention of normalizing the radiation pattern to have response of 1 at the 
+adopt the common convention of normalizing the radiation pattern to have response of unity at the 
 center of the main lobe \citep[e.g.][]{nap99}. Consequently, the translation between the terminal 
@@
 Here we used the Fourier relationship between the antenna's far-field radiation pattern and the 
-aperture illumination pattern (sometimes referred to as the aperture distribution), $W_a(\mathbf{r}-\mathbf{r}_a)=\int  \mathcal{W}_a(\hat{\mathbf{s}}) \, d
-\Omega$. 
+aperture illumination pattern (sometimes referred to as the aperture distribution), 
+\begin{equation}
+W_a(\mathbf{r}-\mathbf{r}_a)=\int  \mathcal{W}_a(\hat{\mathbf{s}}) e^{-2\pi i \mathbf{r} \cdot \hat{\mathbf{s}}}\, d
+\Omega.
+\end{equation} 
 While it is possible to define each antenna aperture illumination pattern in a way to absorb the 
@@
 channelize the time-ordered data. This is effectively estimating the quasi-monochromatic 
-components of the electric field we already discussed. The sampling rate of the digitization 
+components of the electric field. The sampling rate of the digitization 
 determines the bandwidth of the observation, and the sampling period determines the minimum 
@@
 The next step is to grid the \textbf{calibrated electric field measurements to a finite set of regularly spaced 
-positions on the 2D plane of the ground, $\mathbf{r}_j$.
+positions on the plane of the ground, $\mathbf{r}_j$.
 The extent of the grid is chosen to encompass the footprint of the antenna array.
@@
 However the aperture illumination patterns are typically compact and non-overlapping, so most pixels
-will only contain contribution from at most one antenna, allowing equation~\ref{eq:gridding}
+will only contain contribution from one antenna, allowing equation~\ref{eq:gridding}
 to be calculated through a sparse matrix multiplication.}
@@
 antennas to a spatial FFT, the computational cost is significantly lowered for certain classes of 
-compact arrays. The output bandwidth is also greatly reduced as visibilities are exchanged for 
+compact arrays. The output bandwidth is also greatly reduced \textbf{in many scenarios} as visibilities are exchanged for 
 gridded images. A full exploration of these scalings was presented in \citealt{thy15c}.
@@
 related to the sum of visibilities involving the antenna used in the correlation -- exactly the sum 
-needed to calibrate a simple sky of a single point source. However, we aim for a more 
+needed to calibrate \textbf{in the case of a single point source sky}. However, we aim for a more 
 generalized solution for arbitrarily complex sky models. We therefore study the full expression 
@@
 following will hold for any chosen pixel, \spix, though it is advantageous to choose a pixel which 
-contains a bright source to achieve a high signal to noise.
+contains a bright source to achieve a high signal to noise \textbf{ratio}.
 
@@
 uncalibrated antenna electric field measurements to remove the corresponding
-terms from equation~\ref{eq:Kna}.}
+terms from equation~\ref{eq:Kna}:}
 \begin{align}\label{eq:Cna}
 \Cna & \equiv \Kna - \frac{1}{\Nant g^{*(n)}_a}\mathcal{W}^*_a(\spix)e^{-2\pi i \spix \cdot \ra} \left<|E_a|^2\right>_t\nonumber \\
-& = \frac{g_a}{\Nant} \sum_{b\ne a} \frac{g^*_b}{g^{*(n)}_b} \mathcal{W}^*_b(\spix) e^{-2\pi i \spix \cdot \rb} V^I_{ab}
+& = \frac{g_a}{\Nant} \sum_{b\ne a} \frac{g^*_b}{g^{*(n)}_b} \mathcal{W}^*_b(\spix) e^{-2\pi i \spix \cdot \rb} V^I_{ab}.
 \end{align}
@@
 using equation~\ref{eq:cal_solution}, and additional fitting if desired. \textbf{This step does not have
-the speed requirement of the steps in the gray box because the correlations have been 
+the stringent cadence requirement of the steps in the gray box because the correlations have been 
 averaged in time.}
@@
 We conclude this section by connecting our calibration expression to that found in a visibility 
-framework. In the limit of a single bright calibrating source \textbf{at center,} we can greatly 
+framework. In the limit of a single bright calibrating source \textbf{at zenith,} we can greatly 
 simplify equations~\ref{eq:Cna} and~\ref{eq:cal_solution}. We will assume the \textbf{radiation patterns} are 
@@
 For our unknown gains, we create a set of random complex numbers where the amplitude is 
-drawn from an approximately gaussian distribution centered around 1 with width 0.25, and 
+drawn from an approximately gaussian distribution centered around \textbf{unity} with width 0.25, and 
 completely random phase. These are our ``true gains", and we apply them to the frequency
@@
 
-After the gains have converged, we see both the phases and amplitudes to continue to 
+After the gains have converged, we see both the phases and amplitudes continue to 
 fluctuate coherently. This is due to the stochastic fluctuations of the sources 
@@
 \textbf{We simulate the same antenna layout as in Section~\ref{sec:sim}. 
-However, we increase the number of frequency channels simulated to 64.}
+However, we increase the number of frequency channels simulated to 64, while keeping the flux of all sources constant across frequency.}
 Because our calibration loop treats 
@@
 calibration loop are created using all point sources in the simulated sky. For the single point 
-source model, we only allow the calibration loop to be privy to the bright calibrator source. 
+source model, we only \textbf{include the bright calibrator source in the model visibilities.}
 While we vary the model used for calibration, the simulated ``true sky'' remains fixed with all 50 
@@
 EPICal underperformed with the full sky model: EPICal only used a single pixel in the sky to 
-form its solutions. When using the full sky model, it was down weighting the information in the 
+form its solutions. When using the full sky model, it was \textbf{under utilizing} the information in the 
 rest of the sky. But when the sky model only contains the bright point source, the single pixel 
@@
 update our gain estimates over 10 \textbf{samples} (51.2 ms). We iterate the loop 30 times for a 
-total of 1.536 seconds. The results of this calibration experiment are shown in Figs.~
+total of 1.536 seconds of data processed. The results of this calibration experiment are shown in Figs.~
 \ref{fig:data_phase}~--~\ref{fig:data_images}.
@@
 iterations, again with each colored line representing a different antenna. Given the quality of 
-uncalibrated image demonstrated in \cite{thy15c}, it is a bit surprising to see the phase 
-variation in our solutions. However, the phases are relatively flat after about 15 iterations 
+uncalibrated image demonstrated in \cite{thy15c}, \textbf{we had anticipated that the phase
+solutions would be small, hence the relatively large amplitudes of the phase solutions we
+found was somewhat unexpected.}
+However, the phases are relatively flat after about 15 iterations 
 (modulo noise), and exhibit a central ``trunk" where the majority of phases are congregated. 
@@
 Cyg A accounts for  approximately $20,539/171,000 \approx 12\%$ of the total sky power. 
-This places us in a regime very similar to the incomplete sky model simulations of the 
+This places us in a regime similar to the incomplete sky model simulations of the 
 previous section, where the calibration pixel accounted for 21.7\% of the sky power. We 
@@
 peak of the image over the noise level. We estimate the noise level as the median of the 
-absolute deviation of the image, a robust statistic. With this metric we find the \textbf{uncalibrated 
+absolute deviation of the image. With this metric we find the \textbf{uncalibrated 
 and} calibrated images to have \textbf{dynamic ranges of 88.5 and 137, respectively, a 55\% 
@@
 can be difficult to measure at the precision necessary \citep[e.g.][]{neb15,vir14,thy15b}. 
-However, the direct imaging correlator provides exactly the model necessary in real time, and 
-can be iterated over to improve the images and gain solutions.
+However, the direct imaging correlator provides exactly the \textbf{product of the sky and beam} necessary in real time, and 
+can be iterated over to improve the images and gain solutions, \textbf{leaving final beam
+correction to post-processing as in traditional processing.}
 
@@
 increase the scope. With next generation instruments in the planning and development stages, 
-EPICal is poised to greatly contribute to their success. The software is integrated into the EPIC 
+EPICal is poised to \textbf{enable new design spaces}. The software is integrated into the EPIC 
 package and freely available to use in simulations or \textbf{post-processing} of data. Work is 
@@
 
+\interlinepenalty=10000
 \bibliographystyle{../mnras}

### 2016-07-11 cal_demo/ms.tex
Finished JDBs comments

@@
 to form gain solutions in real time without ever forming visibilities. This method scales as the 
-number of antennas, and produces results comparable to those from visibilities. Through 
-simulations and application to Long Wavelength Array data we show this algorithm is a 
-promising solution for next generation instruments.
+number of antennas, and produces results comparable to those from visibilities. 
+\textbf{We use simulations to demonstrate the EPICal technique and study the noise properties of
+our gain solutions, showing they are similar to visibility based solutions in realistic 
+situations.
+By applying EPICal to Long Wavelength Array data we achieve a 65\% dynamic range
+improvement compared to uncalibrated images, showing this algorithm is a promising
+solution for next generation instruments.}
 \end{abstract}
@@
 A challenge common to all direct imaging algorithms is calibration of the antenna gains. With a 
-traditional \textbf{cross correlator}, pair-wise visibilities are written to disk and used to calibrate 
+traditional \textbf{cross correlator}, visibilities are written to disk and used to calibrate 
 \textbf{after observing and before further processing} such as imaging. However, a direct imaging correlator mixes 
@@
 
+\textbf{Our simulated antenna 
+array consists of} the inner 51 antennas of the MWA layout \citep{bea12}, within a bounding 
+box of 150~m. 
 \textbf{The physical parameters of our simulation can be arbitrarily scaled to other cases. However, 
-we provide typical units for an MWA observation as an example. Our simulated antenna 
-array consists of} the inner 51 antennas of the MWA layout \citep{bea12}, within a bounding 
-box of 150~m. The \textbf{antenna aperture illumination pattern} used is a 4.4~m square tophat; \textbf{a very rough 
+we provide typical units for an MWA observation as an example. 
+The antenna aperture illumination pattern} used is a 4.4~m square tophat; \textbf{a very rough 
 approximation to the square shaped tiles of the MWA.} Because our algorithm treats frequency 
@@
 For our unknown gains, we create a set of random complex numbers where the amplitude is 
-drawn from an approximately gaussian distribution centered around \textbf{unity} with width 0.25, and 
-completely random phase. These are our ``true gains", and we apply them to the frequency
+drawn from \textbf{a gaussian distribution centered around unity with width 0.25, and truncated at zero. The gain phases are drawn from a uniform distribution between $\pm \pi$. }
+These are our ``true gains", and we apply them to the frequency
 domain simulated antenna electric field \textbf{measurements} as in equation~\ref{eq:apply_gain}. Our analysis is 
@@
 largest apparent flux (intrinsic flux attenuated by the primary beam). These correlation values 
-are used to update the gain estimates. \textbf{For this simple demonstration we create perfect model 
+are used to update the gain estimates. \textbf{We create perfect model 
 visibilities of our 10 simulated sources to be used in equation~\ref{eq:damping}. 
@@
 \includegraphics[width=\columnwidth]{fig4.pdf}
-\caption{Gain estimate amplitudes as a function of iteration for simulated calibration. Again, 
-each line represents an antenna in the 51 MWA antenna sample. The gain estimates were 
-initialized randomly, while the true values were unity. After about \textbf{12} iterations we see the 
+\caption{\textbf{Same as Fig.~\ref{fig:sim_phase}, but for gain amplitudes.
+Each} line represents an antenna in the 51 MWA antenna sample. 
+\textbf{The relative amplitude plotted is the magnitude of the estimated gains divided by the true gains.}
+After about \textbf{12} iterations we see the 
 calibration loop has settled around the correct values, with only noise remaining.}
@@
 
-The results of our simulations are shown in Fig.~\ref{fig:errors}. The gain errors from our 
-EPICal method are shown with solid lines, while the dashed lines represent the visibility-based 
-calibration. The thick lines represent the results of using a full sky model, and the thin lines are 
-from the single point source model.  In the case of the full sky model, we see the errors trend 
+The results of our simulations \textbf{for all four calibration methods are shown in Fig.~\ref{fig:errors}. }
+In the case of the full sky model, we see the errors trend 
 downward with longer integration times, as expected. The EPICal errors are slightly higher than 
@@
 trend. }
-The same level of \textbf{gain error as visibility based solutions} can be achieved with a longer integration time (or larger damping 
-factor).
+\textbf{With longer integration time (or larger damping factor), EPICal can achieve the same 
+level of gain error as visibility based solutions.}
 
@@
 
+\textbf{In a more realistic situation, the observer will not have a perfectly complete model of the sky.
+A common method is to instead model the brightest sources in the field,
+or even a single dominant source.}
 When using the single point source model, all calibration methods in Fig.~\ref{fig:errors} trend 
@@
 used in EPICal is the location where this model is most accurate, effectively down weighting 
-the pixels with incomplete sky model. In the regime typical of HERA discussed above, EPICal's 
+the pixels with incomplete sky model. 
+\textbf{In contrast, the visibility-based calibration method weights the apparent sky equally,
+and experiences a more harsh penalty for missing sources in the model.}
+In the regime typical of HERA discussed above, EPICal's 
 error floor was 10\% lower than that of visibility-based solutions.
@@
 \includegraphics[width=\columnwidth]{fig8.pdf}
-\caption{Gain amplitudes as a function of calibration iteration for an LWA TBN observation. The 
+\caption{\textbf{Same as Fig.~\ref{fig:data_phase}, but for gain amplitudes.}
+The 
 gains estimates were initialized with amplitude 0.25 after inspection of the electric field values 
@@
 Figure~\ref{fig:data_images} shows the improvement in the images due to our calibration. The 
-left panel shows the uncalibrated image integrated over 51.2 ms, \textbf{83 kHz.} Cyg~A is prominent 
+left panel shows the uncalibrated image integrated over 51.2 ms, \textbf{58.6 kHz.} Cyg~A is prominent 
 near the center of the image, and Cas~A is also clearly visible in the upper right. The middle 
@@
 \caption{Images produced before (left) and after (middle) calibrating LWA data. These images 
-were produced with \textbf{83 kHz} bandwidth and 51.2 ms integration. The calibrated image shows 
+were produced with \textbf{58.6 kHz} bandwidth and 51.2 ms integration. The calibrated image shows 
 significant reduction in \textbf{rumble} throughout, while retaining the prominent Cyg~A and 
@@
 absolute deviation of the image. With this metric we find the \textbf{uncalibrated 
-and} calibrated images to have \textbf{dynamic ranges of 88.5 and 137, respectively, a 55\% 
+and} calibrated images to have \textbf{dynamic ranges of 80.4 and 133.1, respectively, a 65\% 
 improvement.}
@@
 self-calibration. This can be especially useful for compact, widefield arrays which require a model of 
-both compact and diffuse sources over a large patch of sky. In addition to requiring the sky 
-model, the calibration also requires this model to be attenuated by the primary beam, which 
+both compact and diffuse sources over a large patch of sky. 
+\textbf{As with any sky-based calibration scheme, EPICal requires that the sky model be attenuated
+by the primary beam,} which 
 can be difficult to measure at the precision necessary \citep[e.g.][]{neb15,vir14,thy15b}. 

### 2016-07-26 cal_demo/ms.tex
Lots of small changes re: Nithya's comments

@@
 \Nant)$ scaling.} \textbf{This} can save orders of magnitude in computation cost for next generation arrays 
-consisting of hundreds to thousands of antennas. However, because signals are mixed in the 
-correlator, gain correction must be applied \textbf{prior to correlation, rather than on visibilities post-correlation.} We develop the EPICal algorithm 
+consisting of hundreds to thousands of antennas. However, because \textbf{antenna} signals are mixed in the 
+correlator, gain correction must be applied \textbf{prior to imaging, rather than on visibilities post-correlation.} We develop the EPICal algorithm 
 to form gain solutions in real time without ever forming visibilities. This method scales as the 
@@
 situations.
-By applying EPICal to Long Wavelength Array data we achieve a 65\% dynamic range
+By applying EPICal to two seconds of Long Wavelength Array data we achieve a 65\% dynamic range
 improvement compared to uncalibrated images, showing this algorithm is a promising
@@
 Recording the full visibility matrix for $\Nant \gtrsim 10^3$ arrays at this timescale leads to 
-extremely high data write rates.
+extremely high data rates.
 
@@
 
-\textbf{It is convenient to simplify notation by expressing the position vector in units of wavelengths relative to the monochromatic component of interest, 
-$\mathbf{r}\rightarrow \mathbf{r}/\lambda$. We will also assume all antennas lie in a plane, 
+\textbf{We will assume all antennas lie in a plane, 
 reducing the problem to two dimensions. While the MOFF algorithm in principle can account for 
@@
 \begin{equation}\label{eq:sky_propagated}
-E(\mathbf{r},f,t) = \int \mathcal{E}(\hat{\mathbf{s}},f,t) e^{-2\pi i \mathbf{r} \cdot \hat{\mathbf{s}}}\, d\Omega
+E(\mathbf{r},f,t) = \int \mathcal{E}(\hat{\mathbf{s}},f,t) e^{-2\pi i f\mathbf{r} \cdot \hat{\mathbf{s}}/c}\, d\Omega
 \end{equation}  
@@
 
-\textbf{Next we turn to the measurement performed by an antenna with far-field radiation pattern 
-$\mathcal{W}_a(\hat{\mathbf{s}})$, where the subscript $a$ denotes a specific antenna. We 
+\textbf{Next we turn to the measurement performed by an antenna with frequency dependent far-field radiation pattern 
+$\mathcal{W}_a(\hat{\mathbf{s}},f)$, where the subscript $a$ denotes a specific antenna. We 
 adopt the common convention of normalizing the radiation pattern to have response of unity at the 
@@
 \begin{equation}
-E^I_a(f,t) = \int \mathcal{W}_a(\hat{\mathbf{s}})\mathcal{E}(\hat{\mathbf{s}},f,t) e^{-2\pi i \mathbf{r} \cdot \hat{\mathbf{s}}}\, d\Omega. 
+E^I_a(f,t) = \int \mathcal{W}_a(\hat{\mathbf{s}},f)\mathcal{E}(\hat{\mathbf{s}},f,t) e^{-2\pi i f\mathbf{r} \cdot \hat{\mathbf{s}}/c}\, d\Omega. 
 \end{equation}
@@
 \begin{equation}
-E^I_a(f,t) = \int_{\mathrm{aperture}}W_a(\mathbf{r}-\mathbf{r}_a)E_a(\mathbf{r},f,t)\,d^2\mathbf{r}
+E^I_a(f,t) = \int_{\mathrm{aperture}}W_a(\mathbf{r}-\mathbf{r}_a,f)E(\mathbf{r},f,t)\,d^2\mathbf{r}
 \end{equation}
@@
 \begin{equation}
-W_a(\mathbf{r}-\mathbf{r}_a)=\int  \mathcal{W}_a(\hat{\mathbf{s}}) e^{-2\pi i \mathbf{r} \cdot \hat{\mathbf{s}}}\, d
+W_a(\mathbf{r}-\mathbf{r}_a,f)=\int  \mathcal{W}_a(\hat{\mathbf{s}},f) e^{-2\pi i f\mathbf{r} \cdot \hat{\mathbf{s}}/c}\, d
 \Omega.
@@
 channelize the time-ordered data. This is effectively estimating the quasi-monochromatic 
-components of the electric field. The sampling rate of the digitization 
-determines the bandwidth of the observation, and the sampling period determines the minimum 
-frequency channel resolution. The frequency Fourier transform operation is often referred to as 
+components of the electric field. 
+The bandwidth of the observation and the time sampling rate are related by $B=1/\Delta t$,
+and the channelized frequency resolution and sampling period have the relation
+$\Delta f = 1/T$, according to the Nyquist theorem.
+The frequency Fourier transform operation is often referred to as 
 the F-engine of a correlator. The MOFF shares the same F-engine design as an FX cross
@@
 the antenna far-field radiation patterns}. Here we introduce a multiplicative complex gain as 
-well as an additive noise term which corrupt the measured signals.
+well as an additive noise term which corrupt the measured signals \textbf{at each of the antennas}.
 \begin{equation}\label{eq:apply_gain}
@@
 \end{equation}
-\textbf{Formally each pixel is calculated by summing over all antennas and their aperture illumination patterns.
+\textbf{Formally the value at each pixel is calculated by summing the antenna
+measurements over all antennas, each weighted by its respective aperture illumination
+pattern.
 However the aperture illumination patterns are typically compact and non-overlapping, so most pixels
@@
 instantaneous electric field images of the sky \textbf{at fixed locations $\s_k$.
-The Fourier dual coordinates to $\mathbf{r}$ are $(l,m)$, where $l=\sin\theta\cos\phi$,
+The FFT is possible because we have gridded to regularly spaced samples in a 2D plane.
+The Fourier dual coordinates to $\mathbf{r}f/c$ are $(l,m)$, where $l=\sin\theta\cos\phi$,
 $m=\nobreak\sin\theta\sin\phi$, and $(\theta,\phi)$ are the zenith 
@@
 $\s=(l,m,\sqrt{1-l^2-m^2})$.
-The sample spacing and resolution of $\s_k$ are given 
-by the choices of $\mathbf{r}_j$ grid parameters, according to the usual FFT relations. 
-Fine $\mathbf{r}_j$ grid spacing can result in non-physical angular pixels ($l^2+m^2>1$),
-which should be discarded.}
+Being direction cosines, grid locations with ${l^2+m^2>1}$ are non-physical and should
+be ignored.}
 
-The pipeline up to this point can be summarized 
+The algorithm up to this point can be summarized 
 through the following equation.
 \begin{align}
-\mathcal{E}'(\s_k,t) & = \underbrace{\frac{1}{\Nant} \sum_j e^{2\pi i \mathbf{r}_j \cdot \s_k}}_{\mathrm{2D\;FFT}} 
+\mathcal{E}'(\s_k,t) & = \underbrace{\frac{1}{\Nant} \sum_j e^{2\pi i f\mathbf{r}_j \cdot \s_k/c}}_{\mathrm{2D\;FFT}} 
 \underbrace{\sum_a W_a(\mathbf{r}_j - \ra)}_{\mathrm{Gridding}} \\
@@
 regular gridpoints, $\mathbf{r}_j$. The \textbf{sum over $j$} denotes the 2D FFT to sky coordinates, resulting in an 
-estimate for the instantaneous electric field image. We can simplify this expression by 
+estimate for the instantaneous electric field image. 
+\textbf{The sums over $a$ and $j$ mix the antenna signals together, which is the primary 
+reason calibration must be performed beforehand where individual antennas retain their 
+identities.}
+
+We can simplify \textbf{the above} expression by 
 exchanging the sums \textbf{and use the DFT to approximate the Fourier relationship between
@@
 \begin{align}\label{eq:epix}
-\mathcal{E}'(\s_k,t) & = \frac{1}{\Nant} \sum_a \frac{1}{g'_a}\left(g_a E^I_a(t)+n_a(t)\right) e^{2\pi i \s_k \cdot \ra} \nonumber \\
-  & \qquad\times \sum_j W_a(\mathbf{r}_j-\ra)e^{2\pi i \s_k \cdot (\mathbf{r}_j-\ra)} \nonumber\\
-& \approx \frac{1}{\Nant} \sum_a \frac{1}{g'_a}\left(g_aE^I_a(t)+n_a(t)\right) e^{2\pi i \s_k \cdot \ra}\mathcal{W}_a(\s_k)
+\mathcal{E}'(\s_k,t) & = \frac{1}{\Nant} \sum_a \frac{1}{g'_a}\left(g_a E^I_a(t)+n_a(t)\right) e^{2\pi i f\s_k \cdot \ra/c} \nonumber \\
+  & \qquad\times \sum_j W_a(\mathbf{r}_j-\ra)e^{2\pi i f\s_k \cdot (\mathbf{r}_j-\ra)/c} \nonumber\\
+& \approx \frac{1}{\Nant} \sum_a \frac{1}{g'_a}\left(g_aE^I_a(t)+n_a(t)\right) e^{2\pi i f\s_k \cdot \ra/c}\mathcal{W}_a(\s_k)
 \end{align}
@@
 the image by a factor of the \textbf{far-field radiation response, $\mathcal{W}_a(\s)$.}
+\textbf{This added factor of the instrument response is sometimes referred to as the
+\emph{holographic frame} in the literature \citep[e.g.][]{mor09, sul12}.}
 
@@
 antennas to a spatial FFT, the computational cost is significantly lowered for certain classes of 
-compact arrays. The output bandwidth is also greatly reduced \textbf{in many scenarios} as visibilities are exchanged for 
-gridded images. A full exploration of these scalings was presented in \citealt{thy15c}.
+compact arrays. The output bandwidth is also greatly reduced \textbf{in many scenarios} as visibilities are \textbf{replaced with} 
+gridded images. \textbf{These scalings were explored in \citealt{thy15c}.}
 
@@
 signal from all antennas coherently, and thus requires the calibration to be applied in real time 
-at the front end of the correlator. Furthermore, visibilities are never formed, which are 
+\textbf{before imaging}. Furthermore, visibilities are never formed, which are 
 traditionally the basic measurement used to form calibration solutions. Here we derive an 
@@
 Because the gains must be estimated in real time, we will need to use a finite stream of data to 
-form an estimate which will be applied to the subsequent stream of data. Therefore the 
+form an estimate which will be applied to the subsequent stream of data. 
+\textbf{We assume that the changes in gains are negligible over adjacent intervals, which 
+in some cases may limit the length of data streams.
+Because data from one interval is used to update the gain solutions for the next,} the 
 process of finding a solution is an iterative one. We will use parenthetical superscripts to 
@@
 \begin{equation}\label{eq:Kna_def}
-\Kna \equiv \left<E_a^{(n)}(t) \mathcal{E}'^{*(n)}(\spix,t)\right>_t,
+\Kna \equiv \left<E_a(t) \mathcal{E}'(\spix,t)\right>_{t^{(n)}},
 \end{equation}
@@
 calibration loop,
-\textbf{the superscript $^*$ indicates complex conjugation,}
+\textbf{and we use $t^{(n)}$ outside the brackets on the right-hand side to indicate all quantities
+within the brackets are drawn from time samples in the $n^\mathrm{th}$ calibration loop.}
+\textbf{The superscript $^*$ indicates complex conjugation,}
 and \spix\, is the pixel center nearest a bright calibrator of interest. The 
@@
 \Kna & = \left<\left(g_a E_a^I(t)+n_a(t)\right) \frac{1}{\Nant} \right. \nonumber \\
-& \qquad \left. \times \sum_b \frac{1}{g^{*(n)}_b}\left(g^*_bE_b^{I*}(t)+n^*_b(t)\right) e^{-2\pi i \spix \cdot \rb} \mathcal{W}^*_b(\spix)\right>_t \nonumber \\
-& = \frac{1}{\Nant} \sum_b \frac{1}{g^{*(n)}_b} \mathcal{W}^*_b(\spix) e^{-2\pi i \spix \cdot \rb} \nonumber \\
-& \qquad \times \left<\left(g_aE_a^I(t)+n_a(t)\right)\left(g^*_bE_b^{I*}+n^*_b(t)\right) \right>_t \nonumber \\
-& = \frac{1}{\Nant} \sum_b \frac{1}{g^{*(n)}_b} \mathcal{W}^*_b(\spix) e^{-2\pi i \spix \cdot \rb} \nonumber \\
-& \qquad \times \left(g_a g^*_b \left<E^I_a(t)E^{*I}_b(t)\right>_t + g_a\left<E_a^I(t)n^*_b(t)\right>_t \right. \nonumber \\
-& \qquad \left. + g^*_b\left<n_a(t)E^{I*}_b(t)\right>_t + \left<n_a(t) n^*_b(t)\right>_t\right)  
+& \qquad \left. \times \sum_b \frac{1}{g^{*(n)}_b}\left(g^*_bE_b^{I*}(t)+n^*_b(t)\right) \right.\nonumber \\
+& \qquad \left. \times e^{-2\pi i f\spix \cdot \rb/c} \mathcal{W}^*_b(\spix)\right>_{t^{(n)}} \nonumber \\
+& = \frac{1}{\Nant} \sum_b \frac{1}{g^{*(n)}_b} \mathcal{W}^*_b(\spix) e^{-2\pi i f\spix \cdot \rb/c} \nonumber \\
+& \qquad \times \left<\left(g_aE_a^I(t)+n_a(t)\right)\left(g^*_bE_b^{I*}+n^*_b(t)\right) \right>_{t^{(n)}} \nonumber \\
+& = \frac{1}{\Nant} \sum_b \frac{1}{g^{*(n)}_b} \mathcal{W}^*_b(\spix) e^{-2\pi i f\spix \cdot \rb/c} \nonumber \\
+& \qquad \times \left(g_a g^*_b \left<E^I_a(t)E^{*I}_b(t)\right>_{t^{(n)}} + g_a\left<E_a^I(t)n^*_b(t)\right>_{t^{(n)}} \right. \nonumber \\
+& \qquad \left. + g^*_b\left<n_a(t)E^{I*}_b(t)\right>_{t^{(n)}} + \left<n_a(t) n^*_b(t)\right>_{t^{(n)}}\right)  
 %& = \frac{1}{\Nant} \sum_b \frac{1}{g^{*(n)}_b} \mathcal{W}^*_b(\spix) e^{-2\pi i \spix \cdot \rb} \left(g_ag^*_bV^I_{ab}+n_{ab}\right)
@@
 \begin{align}\label{eq:Kna}
-\Kna & = \frac{1}{\Nant} \sum_b \frac{1}{g^{*(n)}_b} \mathcal{W}^*_b(\spix) e^{-2\pi i \spix \cdot \rb}\nonumber \\
-& \qquad \times \left( g_ag_b^* V_{ab}^I + \delta_{ab}\left<|n_a(t)|^2\right>_t\right)
+\Kna & = \frac{1}{\Nant} \sum_b \frac{1}{g^{*(n)}_b} \mathcal{W}^*_b(\spix) e^{-2\pi i f\spix \cdot \rb/c}\nonumber \\
+& \qquad \times \left( g_ag_b^* V_{ab}^I + \delta_{ab}\left<|n_a(t)|^2\right>_{t^{(n)}}\right)
 \end{align}
@@
 \begin{align}\label{eq:Cna}
-\Cna & \equiv \Kna - \frac{1}{\Nant g^{*(n)}_a}\mathcal{W}^*_a(\spix)e^{-2\pi i \spix \cdot \ra} \left<|E_a|^2\right>_t\nonumber \\
-& = \frac{g_a}{\Nant} \sum_{b\ne a} \frac{g^*_b}{g^{*(n)}_b} \mathcal{W}^*_b(\spix) e^{-2\pi i \spix \cdot \rb} V^I_{ab}.
+\Cna & \equiv \Kna - \frac{1}{\Nant g^{*(n)}_a}\mathcal{W}^*_a(\spix)e^{-2\pi i f\spix \cdot \ra/c} \left<|E_a|^2\right>_{t^{(n)}}\nonumber \\
+& = \frac{g_a}{\Nant} \sum_{b\ne a} \frac{g^*_b}{g^{*(n)}_b} \mathcal{W}^*_b(\spix) e^{-2\pi i f\spix \cdot \rb/c} V^I_{ab}.
 \end{align}
+\textbf{Due to the above subtraction, the sum is now over $b \ne a$, explicitly removing the
+dependence on the self-correlated antenna signals.}
 
@@
 \begin{equation}\label{eq:cal_solution}
-g^{(n+1)}_a = \frac{\Cna \Nant }{ \sum_{b\ne a} \mathcal{W}^*_b(\spix) e^{-2\pi i \spix \cdot \rb} V^M_{ab} }
+g^{(n+1)}_a = \frac{\Cna \Nant }{ \sum_{b\ne a} \mathcal{W}^*_b(\spix) e^{-2\pi i f\spix \cdot \rb/c} V^M_{ab} }
 \end{equation}
@@
 \textbf{computed in near real-time}). The correlated values are then \textbf{used to estimate the gains}
-using equation~\ref{eq:cal_solution}, and additional fitting if desired. \textbf{This step does not have
-the stringent cadence requirement of the steps in the gray box because the correlations have been 
+using equation~\ref{eq:cal_solution}, and additional fitting if desired 
+\textbf{(see Section \ref{sec:discussion}).}
+\textbf{This step does not have
+the stringent cadence requirement of the steps in the gray box because the 
+antenna-pixel correlations have been 
 averaged in time.}
@@
 \textbf{used to update the gain estimates using} equations~\ref{eq:cal_solution}~and~\ref{eq:damping},
-and \textbf{any additional fitting.}
+and \textbf{any additional fitting (Section \ref{sec:discussion}).}
 The gray box shows operations which must be done at high speed (before averaging in time),
 \textbf{while the} white boxes show operations which can be performed \textbf{at lower speeds because
-they involve time averages quantities.}
+they involve time averaged quantities, and need to be updated on slower timescales.}
 }
@@
 %g^{(n+1)}_a = (1-\damp) g'^{(n+1)}_a + \damp g^{(n)}_a
-g^{(n+1)}_a = \frac{(1-\damp)\Cna \Nant }{ \sum_{b\ne a} \mathcal{W}^*_b(\spix) e^{-2\pi i \spix \cdot \rb} V^M_{ab} }+ \damp g^{(n)}_a
+g^{(n+1)}_a = \frac{(1-\damp)\Cna \Nant }{ \sum_{b\ne a} \mathcal{W}^*_b(\spix) e^{-2\pi i f\spix \cdot \rb/c} V^M_{ab} }+ \damp g^{(n)}_a
 \end{equation}
@@
  Once the loop converges the damped version is 
-essentially a weighted average over the past several iterations, giving it a longer effective 
+essentially a weighted average over the past several iterations, giving \textbf{the solutions} a longer effective 
 integration time.
@@
 created for each antenna with random phase and amplitude, which is used to corrupt the 
-simulated data stream, then we attempt to recover the gains using our calibration routine. The 
+simulated data stream, then we attempt to \textbf{estimate} the gains using our calibration routine. The 
 simulation software used is included in the EPIC package.
@@
 For our unknown gains, we create a set of random complex numbers where the amplitude is 
-drawn from \textbf{a gaussian distribution centered around unity with width 0.25, and truncated at zero. The gain phases are drawn from a uniform distribution between $\pm \pi$. }
+drawn from \textbf{a gaussian distribution centered around unity with width 0.25, and truncated at zero. The gain phases are drawn from a uniform distribution in the range $[-\pi, \pi)$. }
 These are our ``true gains", and we apply them to the frequency
@@
 
-We next process and image \caliter~time \textbf{samples} (10~ms). We also form the correlations, 
+We next process and image \caliter~time \textbf{samples} (10~ms). We also form the \textbf{antenna-pixel} correlations, 
 \Cna[0], used in our calibration loop. The pixel used for the correlation is the source with the 
@@
 The updated gain estimates are} used to calibrate the following 
-\caliter~time samples. Through experimentation we found a damping factor of $\damp=0.35$ 
+\caliter~time samples. Through \textbf{trial and error} we found a damping factor of $\damp=0.35$ 
 resulted in the quickest convergence in this simulation.
@@
 The calibration loop continues by updating the gain estimates every \caliter~time \textbf{samples.} The 
-phases of our gain estimates are shown in Fig.~\ref{fig:sim_phase} for 20 such iterations. 
+\textbf{phase errors} of our gain estimates are shown in Fig.~\ref{fig:sim_phase} for 20 such iterations. 
 The phase error plotted is the phase relative to the true gain for each antenna (various colored 
 lines). One antenna was used as a reference to fix the absolute phase, so has zero phase 
-error. The other 50 antennas are shown to have error spanning $2\pi$ initially, and after about 
+error. The other 50 antennas are shown to have error spanning the range $[-\pi,\pi)$ initially, and after about 
 10 iterations lock into a solution, settling down to noise levels around iteration \textbf{12 (0.12~s)}. We 
@@
 integration times, while forming antenna-pixel correlations, $\Cna$, for EPICal gain solutions, 
-and simultaneously forming visibilities from the same \textbf{electric field measurement} streams to find visibility-based 
+and simultaneously forming visibilities from the same streams \textbf{of measured electric fields} to find visibility-based 
 gain solutions for comparison.
@@
 n_a(f,t)\right|^2\right>$, is varied in different simulation runs. We include two limiting 
-cases, where the receiver noise power is subdominant to the sky power ($\sigma_r = 
-10.0$~Jy), and where the receiver dominates the total power ($\sigma_r=100.0$~Jy).
+cases, \textbf{one} where the receiver noise power is subdominant to the sky power ($\sigma_r = 
+10.0$~Jy), and \textbf{the other} where the receiver \textbf{noise} dominates the total power ($\sigma_r=100.0$~Jy).
 
@@
 
-For the EPICal calibration loops, we initialize our gain estimates with the true values 
+For EPICal, we initialize our gain estimates with the true values 
 (${g^{(0)}_a=g_a=1}$), and allow the estimate to be corrupted by the noise through ten 
-iterations of the calibration loop. We \textbf{again} adopt a damping factor $\damp=0.35$. Because EPICal 
+iterations of the calibration loop. 
+\textbf{This allows us to bypass the initial convergence time and quickly study the effect of the
+noise over many simuations.}
+We \textbf{again} adopt a damping factor $\damp=0.35$. Because EPICal 
 updates the gain estimate at each calibration loop, \tcal, but retains a memory of previous 
@@
 \rightarrow \infty$, the geometric series converges to an effective integration time of
-\begin{equation}
+\begin{equation}\label{eq:teff}
 \teff = \tcal \times \frac{1+\damp}{1-\damp}.
@@
 average EPICal errors are 5\% lower in this regime.
+\textbf{The results from the high and low receiver noise regimes are similar, with both approaching the same floor in the single source model case.}
 }
@@
 
-The exact ratio of EPICal gain noise to visibility-based gain noise depends on the fraction of 
+The exact ratio of \textbf{noise in gains from EPICal and the visibility-based estimator}
+depends on the fraction of 
 the total sky power contained in our calibrator source. In the example here, our calibrator 
@@
 the pixels with incomplete sky model. 
-\textbf{In contrast, the visibility-based calibration method weights the apparent sky equally,
-and experiences a more harsh penalty for missing sources in the model.}
+\textbf{In contrast, the visibility-based calibration method 
+does not have any preference to a certain direction, and
+weights the sky equally (beyond the inherent response
+of the instrument).
+This results in a more harsh penalty for missing sources in the model.}
 In the regime typical of HERA discussed above, EPICal's 
@@
 produce noise-like errors in their gains which scale down with more integration time.
+
+\textbf{Comparing the two regimes of receiver noise levels (red versus black lines in 
+Fig.~\ref{fig:errors}), we see that the gains errors behave as one might expect given our
+above interpretations. In the case of a perfect sky model, the higher receiver noise results
+in larger gain errors for a given integration time, but follows the same $t^{-1/2}$ noise 
+trend. The imperfect model lines begin exhibit the same offset at low integration time
+because the errors are dominated by the thermal noise. But as integration time increases,
+they approach the same floor due to the same level of power from unmodeled sources.}
 
@@
 \textbf{The large damping factor, relative to what was used in simulations, was necessary due to
-the larger noise levels in the LWA data compared to the ideal simulations.}
+the larger noise levels in the LWA data compared to the ideal simulations. This yields
+a larger effective integration time, and thus lower noise on each calibration solution
+(Eq.~\ref{eq:teff})}
 
@@
 uncalibrated image demonstrated in \cite{thy15c}, \textbf{we had anticipated that the phase
-solutions would be small, hence the relatively large amplitudes of the phase solutions we
-found was somewhat unexpected.}
+solutions would not exhibit much variation between antennas. Hence the relatively large variation was unexpected.}
 However, the phases are relatively flat after about 15 iterations 
@@
 chosen to have identically zero phase, but happens to be slightly more positive than the bulk of 
-antennas. For plotting clarity, we unwrapped phases resulting in phases that appear to exceed 
-$\pm \pi$.
+antennas.
 
@@
 not surprising given the fairly good quality image produced from uncalibrated data.
+\textbf{For plotting clarity, we unwrapped phases resulting in phases that appear to exceed 
+$\pm \pi$.}
 }
@@
 panel shows the image produced after calibration with identical integration time and bandwidth. 
-The sidelobes throughout the image are significantly suppressed, and the galactic plane is 
+\textbf{We show each of these images with a color scale which saturates at half the
+maximum value in the respective image data. We find that this choice allows us to
+qualitatively compare the quality of the images, independent of the absolute scale which
+changed during the calibration.}
+The sidelobes throughout the \textbf{calibrated} image are significantly suppressed, and the galactic plane is 
 much more evident, despite only modeling Cyg~A and Cas~A. We also note that the feature 
@@
 \caption{Images produced before (left) and after (middle) calibrating LWA data. These images 
-were produced with \textbf{58.6 kHz} bandwidth and 51.2 ms integration. The calibrated image shows 
+were produced with \textbf{58.6 kHz} bandwidth and 51.2 ms integration. 
+\textbf{The color scales are saturated at half the peak of the respective images.}
+The calibrated image shows 
 significant reduction in \textbf{rumble} throughout, while retaining the prominent Cyg~A and 
@@
 considerations for further study. These topics can be thought of as extensions to the white 
-``estimate gains" box in Fig.~\ref{fig:schematic}, which can be performed at a much lower 
-cadence than the actual correlations.
+``estimate gains" box in Fig.~\ref{fig:schematic}, which can be performed at a much 
+\textbf{slower rate} than the actual antenna-pixel correlations.
 
@@
 can be difficult to measure at the precision necessary \citep[e.g.][]{neb15,vir14,thy15b}. 
-However, the direct imaging correlator provides exactly the \textbf{product of the sky and beam} necessary in real time, and 
+However, the direct imaging correlator provides exactly the \textbf{product of the sky and beam} necessary, and 
 can be iterated over to improve the images and gain solutions, \textbf{leaving final beam

### 2016-08-05 imaging_demo/ms.tex
Add, remove and update figures as per review
From df79977 (mnras):
====================

Update figures as per referee report

In ms.tex,

(*) Add, remove and update figures as per referee report

@@
 % out the following line
-\usepackage{newtxtext,newtxmath}
+% \usepackage{newtxtext,newtxmath}
 % Depending on your LaTeX fonts installation, you might get better results with one of these:
@@
 \usepackage[flushleft]{threeparttable} % For table notes
+\usepackage[caption=false]{subfig} % For subfloats
 
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figure5}
+  \includegraphics[width=\columnwidth]{figure4}
   % \includegraphics[width=\columnwidth]{autocorr_uvwts_pbeam}
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figure6}
+  \includegraphics[width=\columnwidth]{figure5}
   % \includegraphics[width=\columnwidth]{MOFF_FX_comparison_10_random_source_positions_4_iterations_test_aperture_zoomed}
@@
   \label{fig:MOFF-FX-image}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figure6}
+  \caption{.}
+  \label{fig:f-engine}
 \end{figure}
@@
   % \includegraphics[width=\columnwidth]{MOFF_FX_comparison_uvwts_test_aperture_zero_spacing_removed.eps}
-  \caption{Comparison of $uv$ plane antenna power response with MOFF (left) and 
-    visibility-based imaging (right). The zero-spacing footprint from the $uv$ 
-    weights obtained with MOFF has been subtracted. Differences in the void 
-    centreed around zero-spacing and other outer regions is due to loss of 
-    sampled antenna footprint precision due to coarseness of grid spacing 
-    further made severe by the correlation process.}
+  \caption{.}
   \label{fig:MOFF-FX-uvwts}
@@
 \end{figure}
+
+\section{Imaging with Heterogeneous Antenna Array}\label{sec:versatility}
+
+\begin{figure*}
+\subfloat[][Simulated Image]{\label{fig:nonphysical-image}\includegraphics[width=0.5\linewidth]{figure10a.eps}}
+\subfloat[][Normalized Radial Flux Density Profile]{\label{fig:norm-flux-density}\includegraphics[width=0.42\linewidth]{figure10b.eps}}
+\caption{.}
+\label{fig:versatility}
+\end{figure*}
 
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figure10}
+  \includegraphics[width=\columnwidth]{figure11}
   % \includegraphics[width=\columnwidth]{MOFF_FX_computations_fov_gridding_annotated}
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figure11}
+  \includegraphics[width=\columnwidth]{figure12}
   % \includegraphics[width=\columnwidth]{MOFF_FX_crossover_baseline_n-antennas_rho_fov_gridding_legended}

### 2016-08-06 imaging_demo/ms.tex
Make changes in notation in mathematical framework
From 8448190 (mnras):
====================

(*) Make changes in notation in section 2 (mathematical framework)

@@
 
-Electric fields from astrophysical sources, $E(\hat{\mathbf{s}})$, in the sky
-coordinate system denoted by sine-projected unit vector $\hat{\mathbf{s}}$, 
-propagate towards the observer as:
+Electric fields from astrophysical sources, $\mathcal{E}(\hat{\mathbf{s}}, f, t)$, in the sky coordinate system denoted by the unit vector $\hat{\mathbf{s}}$, propagate towards the observer as:
 \begin{align}
-  \widetilde{E}(\mathbf{r}) &= \int E(\hat{\mathbf{s}})\,e^{-i2\pi\mathbf{r}\cdot\hat{\mathbf{s}}}\,\dif^2\hat{\mathbf{s}},
-\end{align}
-where, $\mathbf{r}$ denotes the observer's location (measured in wavelengths 
-relative to some arbitrary origin) and $\widetilde{E}
-(\mathbf{r})$ is the propagated electric field. Thus the propagated electric
-field is a linear superposition of the electric fields emanating from
-astronomical sources with appropriate complex phases. It can also be described
-as a Fourier transform of the electric fields in the sky coordinates.
-
-An antenna, $a$, measures a phased sum of these propagated electric fields over
-its effective collecting area with an additive receiver noise:
+  E(\mathbf{r}, f, t) &= \int E(\hat{\mathbf{s}},f,t)\,e^{-2\pi i f\mathbf{r}\cdot\hat{\mathbf{s}}/c}\,\dif\Omega,
+\end{align}
+where, $\mathbf{r}$ denotes the observer's location, {\bf $f$ is the frequency of radiation, $c$ is the speed of light, $t$ denotes time} and $E(\mathbf{r}, f, t)$ is the propagated electric field.
+Thus the propagated electric field is a linear superposition of the electric fields emanating from astronomical sources with appropriate complex phases.
+{\bf Ignoring wide-field effects}, it can be {\bf simplified as} a Fourier transform of the electric fields in the sky coordinates.
+
+An antenna, {\bf located at $\mathbf{r}_a$ (indexed by $a$)}, measures a phased sum of these propagated electric fields over its effective collecting area with an additive receiver noise:
 \begin{align}\label{eqn:measured-E-field}
-  \widetilde{E}_a &= \int \widetilde{W}_a(\mathbf{r}-\mathbf{r}_a)\,\widetilde{E}(\mathbf{r})\,\dif^2\mathbf{r} + \widetilde{n}_a \\
-                  &= \int \widetilde{W}_a(\mathbf{r}-\mathbf{r}_a) \left[ \int E(\hat{\mathbf{s}})\,e^{-i2\pi\mathbf{r}\cdot\hat{\mathbf{s}}}\,\dif^2\hat{\mathbf{s}} \right] \dif^2\mathbf{r} + \widetilde{n}_a \\
-                  &= \int {W}_a(\hat{\mathbf{s}})\,E(\hat{\mathbf{s}})\,e^{-i2\pi\mathbf{r}_a\!\cdot\,\hat{\mathbf{s}}}\,\dif^2\hat{\mathbf{s}} + \widetilde{n}_a
-\end{align}
-where, $\widetilde{W}_a(\mathbf{r})$ is the aperture electric field illumination
-pattern of the antenna and its Fourier transform, $W_a(\hat{\mathbf{s}})$, is the
-directional antenna voltage response.
-
-Interferometers measure {\it visibilities} -- the degree of coherence between
-electric fields measured by a pair of antennas \citep{van34,zer38,tho01}.
-A visibility, $\widetilde{V}_p$, can be written as:
+  E_a(f,t) &= \int W_a(\mathbf{r}-\mathbf{r}_a,f,t)\,E(\mathbf{r},f,t)\,\dif^2\mathbf{r} + n_a(f,t) \\
+           &= \int W_a(\mathbf{r}-\mathbf{r}_a,f,t) \nonumber \\
+           &\quad \times \left[ \int \mathcal{E}(\hat{\mathbf{s}},f,t)\,e^{-2\pi i f\mathbf{r}\cdot\hat{\mathbf{s}}/c}\,\dif\Omega \right] \dif^2\mathbf{r} + n_a(f,t) \\
+           &= \int \mathcal{W}_a(\hat{\mathbf{s}},f,t)\,\mathcal{E}(\hat{\mathbf{s}},f,t)\,e^{-2\pi i f\mathbf{r}_a\!\cdot\,\hat{\mathbf{s}}/c}\,\dif\Omega + n_a(f,t)
+\end{align}
+where, {\bf $W_a(\mathbf{r},f,t)$ is the aperture electric field illumination pattern of the antenna and its Fourier transform, $\mathcal{W}_a(\hat{\mathbf{s}},f,t)$, is the directional antenna voltage response at a given frequency and time}.
+
+Interferometers measure {\it visibilities} -- the degree of coherence between electric fields measured by a pair of antennas \citep{van34,zer38,tho01}. A visibility, $V_{ab}$, can be written as:
 \begin{align}
-  \widetilde{V}_p &= \left\langle \widetilde{E}_a\widetilde{E}_b^\star \right\rangle_t \label{eqn:cc-vis}\hfill\\
-  		&= \left\langle \left[ \int {W}_a(\hat{\mathbf{s}})\,E(\hat{\mathbf{s}})\,e^{-i2\pi\mathbf{r}_a\!\cdot\,\hat{\mathbf{s}}}\,\dif^2\hat{\mathbf{s}} + \widetilde{n}_a \right] \right. \nonumber\\
-		&\qquad \times \left.\left[ \int {W}^\star_b(\hat{\mathbf{s'}})\,E^\star (\hat{\mathbf{s'}})\,e^{i2\pi\mathbf{r}_b\!\cdot\,\hat{\mathbf{s'}}}\,\dif^2\hat{\mathbf{s'}} + \widetilde{n}^\star_b \right] \right\rangle_t \\
-                  &= \iint {W}_a(\hat{\mathbf{s}}) {W}^\star_b(\hat{\mathbf{s'}}) \left\langle E(\hat{\mathbf{s}}) E^\star(\hat{\mathbf{s'}}) \right\rangle_t e^{-i2\pi(\mathbf{r}_a\!\cdot\,\hat{\mathbf{s}}-\mathbf{r}_b\!\cdot\,\hat{\mathbf{s'}})}\,\dif^2\hat{\mathbf{s}}\,\dif^2\hat{\mathbf{s'}},
-\end{align}
-where we have brought the time average into the integral under the assumption that 
-the aperture illumination pattern does not change over the time-scale of the 
-averaging and $\star$ denotes complex conjugation. This expression can be further 
-simplified with the sky brightness, 
-$I(\hat{\mathbf{s}})= \left\langle E(\hat{\mathbf{s}}) E^\star(\hat{\mathbf{s'}}) 
-\right\rangle_t  \delta(\hat{\mathbf{s}}-\hat{\mathbf{s'}})$, and defining the 
-antenna pair sky power response function (or the directional antenna power pattern), 
-$B_p(\hat{\mathbf{s}})~\equiv~{W}_a(\hat{\mathbf{s}})~{W}^\star_b(\hat{\mathbf{s}})$. 
-The result is the visibility expressed in terms of the sky brightness, the power 
-response, and uncorrelated noise terms which we group into $\widetilde{n}_p$.
-\begin{equation}
-\widetilde{V}_p = \int e^{-i2\pi\mathbf{u}_p\!\cdot\,\hat{\mathbf{s}}}\,B_p(\hat{\mathbf{s}})\,I(\hat{\mathbf{s}})\,\dif^2\hat{\mathbf{s}} + \widetilde{n}_p,
-\end{equation}
-where, the baseline coordinate $\mathbf{u}_p=\mathbf{r}_a-\mathbf{r}_b$ is the 
-vector separation between the two antennas. This signifies that the visibility 
-($\widetilde{V}_p$) measured between a
-pair of antennas ($p$) is obtained by the multiplying the sky brightness
-$I(\hat{\mathbf{s}})$ by the antenna power response $B_p(\hat{\mathbf{s}})$ and
-Fourier transforming from the directional coordinates ($\hat{\mathbf{s}}$) to $uv$
-coordinates, which are then sampled at the locations of the antenna spacings (or
-baselines), namely, $\mathbf{u}_p$, and added to the receiver noise $n_p$.
+  V_{ab}(f,t) &= \left\langle E_a(f,t)E_b^\star(f,t) \right\rangle_t \label{eqn:cc-vis}\hfill\\
+              &= \left\langle \left[ \int \mathcal{W}_a(\hat{\mathbf{s}},f,t)\,\mathcal{E}(\hat{\mathbf{s}},f,t)\,e^{-2\pi i f\mathbf{r}_a\!\cdot\,\hat{\mathbf{s}}/c}\,\dif\Omega\right.\right.
+\nonumber\\
+              &\qquad\qquad\qquad\qquad\qquad\qquad\qquad\quad + \left. n_a(f,t) \vphantom{\int}\right] \nonumber\\
+              &\quad \times \left[ \int \mathcal{W}^\star_b(\hat{\mathbf{s'}},f,t)\,\mathcal{E}^\star (\hat{\mathbf{s'}},f,t)\,e^{2\pi i f\mathbf{r}_b\!\cdot\,\hat{\mathbf{s'}}/c}\,\dif\Omega^\prime \right.
+\nonumber\\
+              &\qquad\qquad\qquad\qquad\qquad\qquad\qquad\quad + \left.\left. n^\star_b(f,t) \vphantom{\int}\right] \right\rangle_t \\
+              &= \iint \mathcal{W}_a(\hat{\mathbf{s}},f,t) \mathcal{W}^\star_b(\hat{\mathbf{s'}},f,t) \left\langle \mathcal{E}(\hat{\mathbf{s}},f,t) \mathcal{E}^\star(\hat{\mathbf{s'}},f,t) \right\rangle_t \nonumber\\
+              &\qquad\qquad\qquad\qquad \times e^{-2\pi i f(\mathbf{r}_a\!\cdot\,\hat{\mathbf{s}}-\mathbf{r}_b\!\cdot\,\hat{\mathbf{s'}})/c}\,\dif\Omega\,\dif\Omega^\prime \nonumber\\
+              &\qquad +\textrm{noise-like cross-terms.}
+\end{align}
+where we have brought the time average into the integral under the assumption that the aperture illumination pattern does not change over the time-scale of the averaging and $\star$ denotes complex conjugation.
+{\bf The noise-like cross-terms ideally have zero mean}.
+This expression can be further simplified with the sky brightness, $\mathcal{I}(\hat{\mathbf{s}},f,t)= \left\langle \mathcal{E}(\hat{\mathbf{s}},f,t)\,\mathcal{E}^\star(\hat{\mathbf{s'}},f,t) \right\rangle_t \delta(\hat{\mathbf{s}}-\hat{\mathbf{s'}})$, and defining the antenna-pair sky power response function (or the directional antenna-pair power pattern), $\mathcal{W}_{ab}(\hat{\mathbf{s}})~\equiv~\mathcal{W}_a(\hat{\mathbf{s}},f,t)~\mathcal{W}^\star_b(\hat{\mathbf{s}},f,t)$.
+The result is the visibility expressed in terms of the sky brightness, the power response, and uncorrelated noise terms which we group into $n_{ab}(f,t)$.
+\begin{align}
+V_{ab}(f,t) &= \int \mathcal{W}_{ab}(\hat{\mathbf{s}},f,t)\,\mathcal{I}(\hat{\mathbf{s}},f,t)\,e^{-2\pi i f\mathbf{r}_{ab}\!\cdot\,\hat{\mathbf{s}}/c}\,\dif\Omega \nonumber\\
+  &\qquad\qquad + n_{ab}(f,t),
+\end{align}
+where, the baseline coordinate $\mathbf{r}_{ab}\equiv\mathbf{r}_a-\mathbf{r}_b$ is the vector separation between the two antennas.
+This signifies that the visibility ($V_{ab}$) measured between a pair of antennas is obtained by the multiplying the sky brightness $\mathcal{I}(\hat{\mathbf{s}},f,t)$ by the antenna power response $\mathcal{W}_{ab}(\hat{\mathbf{s}},f,t)$ and Fourier transforming from the directional coordinates ($\hat{\mathbf{s}}$) to {\bf the measurement plane}, which are then sampled at the locations of the antenna spacings (or baselines), namely, $\mathbf{r}_{ab}$, and added to the noise $n_{ab}$.
 
@@
 \begin{align}\label{eqn:software-holography}
-  \widetilde{V}_p &= \int \widetilde{B}(\mathbf{u}^\prime-\mathbf{u}) \left[\int e^{-i2\pi\mathbf{u}.\hat{\mathbf{s}}}\,I(\hat{\mathbf{s}})\,\dif^2\hat{\mathbf{s}}\right]\dif^2\mathbf{u} + n_p,
-\end{align}
-where, $\widetilde{B}(\mathbf{u})$ denotes the $uv$-space antenna power response
-obtained by a Fourier transform of $B(\hat{\mathbf{s}})$. Effectively, the
-multiplication in image space by $B(\hat{\mathbf{s}})$ has been replaced by a
-convolution with $\widetilde{B}(\mathbf{u})$ in $uv$-space. This is the software
-holographic equivalent of a traditional FX correlator output.
-
-Hereafter, we adopt the matrix notation of \citet{mor11}, where vectors are 
-represented with single coordinates, and matrices are represented by two 
-coordinates denoting the spaces the operator transforms between. 
+  V_{ab}(f,t) &= \int W_{ab}(\mathbf{r}_{ab}^\prime-\mathbf{r},f,t) \nonumber\\ 
+              &\qquad \times \left[\int I(\hat{\mathbf{s}},f,t)\,e^{-2\pi i f\mathbf{r}.\hat{\mathbf{s}}/c}\,\dif\Omega\right]\dif^2\mathbf{r}\,+\, n_{ab}(f,t),
+\end{align}
+where, $W_{ab}(\mathbf{r},f,t)$ denotes the {\bf power response of the antenna pair obtained in the measurement plane} by a spatial Fourier transform of $\mathcal{W}_{ab}(\hat{\mathbf{s}},f,t)$.
+Effectively, the multiplication in image space by $\mathcal{W}_{ab}(\hat{\mathbf{s}},f,t)$ has been replaced by a convolution with $W_{ab}(\mathbf{r},f,t)$ in {\bf measurement plane}.
+This is the software holographic equivalent of a traditional FX correlator output.
+
+Hereafter, we adopt the matrix notation of \citet{mor11}, where vectors are represented with single coordinates, and matrices are represented by two coordinates denoting the spaces the operator transforms between.
+Since each frequency is processed independently, we drop the dependence on $f$.
 In this notation, the above measurement equation can be expressed as:
 \begin{align}
-  \mathbf{m}(\mathbf{v}) &= \widetilde{\mathbf{B}}(\mathbf{v},\mathbf{u})\,\mathbf{F}(\mathbf{u},\hat{\mathbf{s}})\,\mathbf{I}(\hat{\mathbf{s}}) + \mathbf{n}(\mathbf{v}),
-\end{align}
-where the sky brightness $\mathbf{I}(\hat{\mathbf{s}})$ is Fourier transformed using
-$\mathbf{F}(\mathbf{u},\hat{\mathbf{s}})$ and the resultant spatial coherence
-function is weighted and summed using the antenna power response,
-$\widetilde{\mathbf{B}}(\mathbf{v},\mathbf{u})$ in $uv$-space sampled at the baseline
-location to obtain the measured visibilities:
+  \mathbf{m}(\mathbf{v}) &= \mathbf{B}(\mathbf{v},\mathbf{u})\,\mathbf{F}(\mathbf{u},\hat{\mathbf{s}})\,\boldsymbol{\mathcal{I}}(\hat{\mathbf{s}}) + \mathbf{n}(\mathbf{v}),
+\end{align}
+where the sky brightness $\boldsymbol{\mathcal{I}}(\hat{\mathbf{s}})$ is Fourier transformed using $\mathbf{F}(\mathbf{u},\hat{\mathbf{s}})$ and the resultant spatial coherence function is weighted and summed using the antenna power response, $\mathbf{B}(\mathbf{v},\mathbf{u})$ in $uv$-space sampled at the baseline location to obtain the measured visibilities:
 \begin{align}
-  \mathbf{m}(\mathbf{v}) &= \left\langle \widetilde{\mathbf{E}}(\mathbf{a})\,\widetilde{\mathbf{E}}^\star(\mathbf{a}^\prime)\right\rangle_t, \label{eqn:matrix-cc-vis}
-\end{align}
-where $\mathbf{m}(\mathbf{v})$ denotes visibilities measured by cross-correlating
-measured antenna electric fields over all possible pairs of $\mathbf{a}$ and
-$\mathbf{a}^\prime$. It is the same as equation~\ref{eqn:cc-vis} written in
-matrix notation.
-
-Using the optimal map-making formalism \citep{teg97a,teg97b}, a software
-holography image is formed using \citep{mor09}:
+  \mathbf{m}(\mathbf{v}) &= \left\langle \mathbf{E}(\mathbf{a})\,\mathbf{E}^\star(\mathbf{a}^\prime)\right\rangle_t, \label{eqn:matrix-cc-vis}
+\end{align}
+where $\mathbf{m}(\mathbf{v})$ denotes visibilities measured by cross-correlating measured antenna electric fields over all possible pairs of $\mathbf{a}$ and $\mathbf{a}^\prime$.
+It is the same as equation~\ref{eqn:cc-vis} written in matrix notation.
+
+Using the optimal map-making formalism \citep{teg97a,teg97b}, a software holography image is formed using \citep{mor09}:
 \begin{align}
-  \mathbf{I}^\prime(\hat{\mathbf{s}}) &= \mathbf{F}^\textrm{T}(\hat{\mathbf{s}},\mathbf{u})\,\widetilde{\mathbf{B}}^{\,\textrm{T}}(\mathbf{u},\mathbf{v})\,\mathbf{N}^{-1}(\mathbf{v},\mathbf{v})\,\mathbf{m}(\mathbf{v}) \label{eqn:dirty-image-FX}
-\end{align}
-where the measured visibilities are weighted by the inverse of the system noise,
-followed by a gridding process using the holographic antenna power response as 
-the gridding kernel, followed by a Fourier transform to create an image
-$\mathbf{I}^\prime(\hat{\mathbf{s}})$. This is the optimal estimate of the true 
-image $\mathbf{I}(\hat{\mathbf{s}})$ given the visibility measurements.
-
-The intermediate step of gridding with the antenna power response can be 
-expressed as a convolution of a data vector generated by gridding the electric 
-fields directly with the antenna illumination pattern.
+  \boldsymbol{\mathcal{I}}^\prime(\hat{\mathbf{s}}) &= \mathbf{F}^\textrm{T}(\hat{\mathbf{s}},\mathbf{u})\,\mathbf{B}^{\,\textrm{T}}(\mathbf{u},\mathbf{v})\,\mathbf{N}^{-1}(\mathbf{v},\mathbf{v})\,\mathbf{m}(\mathbf{v}) \label{eqn:dirty-image-FX}
+\end{align}
+where the measured visibilities are weighted by the inverse of the system noise, followed by a gridding process using the holographic antenna power response as the gridding kernel, followed by a Fourier transform to create an image $\boldsymbol{\mathcal{I}}^\prime(\hat{\mathbf{s}})$.
+This is the optimal estimate of the true image $\boldsymbol{\mathcal{I}}(\hat{\mathbf{s}})$ given the visibility measurements.
+
+The intermediate step of gridding with the antenna power response can be expressed as a convolution of a data vector generated by gridding the electric fields directly with the antenna illumination pattern.
 \begin{align}
-\widetilde{\mathbf{B}}^{\,\textrm{T}}(\mathbf{u},\mathbf{v})\,\mathbf{N}^{-1}(\mathbf{v} ,\mathbf{v})\,\mathbf{m}(\mathbf{v}) &= \left\langle \left[\widetilde{\mathbf{W}}^\textrm{T}_a(\mathbf{r},\mathbf{a})\, \widetilde{\mathbf{N}}^{-1}\!(\mathbf{a},\mathbf{a})\, \widetilde{\mathbf{E}}(\mathbf{a})\right]\right.
+\mathbf{B}^{\,\textrm{T}}(\mathbf{u},\mathbf{v})\,\mathbf{N}^{-1}(\mathbf{v} ,\mathbf{v})\,\mathbf{m}(\mathbf{v}) &= \left\langle \left[\mathbf{W}_a(\mathbf{r},\mathbf{a})\,\mathbf{N}^{-\mathrm{T}}\!(\mathbf{a},\mathbf{a})\, \mathbf{E}^\star(\mathbf{a})\right]\right.
 \nonumber\\
-&\quad\ast\left.\left[\widetilde{\mathbf{W}}_a(\mathbf{r},\mathbf{a})\, \mathbf{N}^{-1}\!(\mathbf{a},\mathbf{a})\, \widetilde{\mathbf{E}}^\star(\mathbf{a})\right]\right\rangle_t\label{eqn:e-field-conv}
-\end{align}
-
-We can then use the multiplication-convolution theorem to move the convolution in 
-Equation \ref{eqn:e-field-conv} to a square after the Fourier transform in 
-Equation 
-\ref{eqn:dirty-image-FX}.
+&\quad\ast\left.\left[\mathbf{W}^\textrm{T}_a(\mathbf{r},\mathbf{a})\,\mathbf{N}^{-1}\!(\mathbf{a},\mathbf{a})\,\mathbf{E}(\mathbf{a})\right]\right\rangle_t\label{eqn:e-field-conv}
+\end{align}
+
+We can then use the multiplication-convolution theorem to move the convolution in Equation \ref{eqn:e-field-conv} to a square after the Fourier transform in Equation~\ref{eqn:dirty-image-FX}.
 \begin{align}
-  \mathbf{I}^\prime(\hat{\mathbf{s}}) &= \left\langle \left|\,\mathbf{F}^\textrm{T}(\hat{\mathbf{s}},\mathbf{r})\,\widetilde{\mathbf{W}}^\textrm{T}(\mathbf{r},\mathbf{a})\,\widetilde{\mathbf{N}}^{-1}(\mathbf{a},\mathbf{a})\,\widetilde{\mathbf{E}}(\mathbf{a})\,\right|^2\right\rangle_t.
+  \boldsymbol{\mathcal{I}}^\prime(\hat{\mathbf{s}}) &= \left\langle \left|\,\mathbf{F}^\textrm{T}(\hat{\mathbf{s}},\mathbf{r})\,\mathbf{W}^\textrm{T}(\mathbf{r},\mathbf{a})\,\mathbf{N}^{-1}(\mathbf{a},\mathbf{a})\,\mathbf{E}(\mathbf{a})\,\right|^2\right\rangle_t.
 \label{eqn:dirty-image-MOFF}
 \end{align}
-The term inside the angular brackets before squaring has a very similar form as
-that in equation~\ref{eqn:dirty-image-FX}. It signifies that the measured antenna
-electric fields are weighted by the antenna noise, weighted and gridded by the
-antenna aperture kernel, Fourier transformed and finally squared to obtain the
-same image estimate that would have been obtained using equation~
-\ref{eqn:dirty-image-FX}.
+The term inside the angular brackets before squaring has a very similar form as that in equation~\ref{eqn:dirty-image-FX}.
+It signifies that the measured antenna electric fields are weighted by the antenna noise, weighted and gridded by the antenna aperture kernel, Fourier transformed and finally squared to obtain the same image estimate that would have been obtained using equation~\ref{eqn:dirty-image-FX}.
 
@@
 
-% \subsection{Data Flow}
-
 EPIC can accept dual-polarization inputs and produce images of all four 
@@
 
-Fig.~\ref{fig:MOFF-flowchart} shows the flowchart for MOFF imaging. The
-propagated electric fields are shown on the left at different time stamps,
-$t_1\ldots t_\textrm{M}$. At each time stamp, the electric fields measured by
-antennas are denoted by $E_1(t)\ldots E_\textrm{N}(t)$. The F-engine performs a
-temporal Fourier transform on the electric field time-series to obtain electric
-field spectra $\widetilde{E}_1(f)\ldots \widetilde{E}_\textrm{N}(f)$ 
-($\widetilde{\mathbf{E}}(\mathbf{a})$ in matrix notation) for each of the 
-antennas. Each of the complex antenna gains are calibrated to correct the 
-corresponding electric field spectra. These calibrated electric fields are 
-gridded using an antenna-based gridding convolution function after which it is 
-spatially Fourier transformed and squared to obtain an image cube for every time 
-stamp. These images are then time-averaged to obtain the accumulated image 
-$I(f)$ ($\mathbf{I}(\hat{\mathbf{s}})$ in matrix notation).
+Fig.~\ref{fig:MOFF-flowchart} shows the flowchart for MOFF imaging.
+The propagated electric fields are shown on the left at different time stamps, $t_1\ldots t_\textrm{M}$.
+At each time stamp, the electric fields measured by antennas are denoted by $\widetilde{E}_1(t)\ldots \widetilde{E}_\textrm{N}(t)$.
+The F-engine performs a temporal Fourier transform on the electric field time-series to obtain electric field spectra $E_1(f)\ldots E_\textrm{N}(f)$ ($\mathbf{E}(\mathbf{a})$ in matrix notation) for each of the antennas.
+Each of the complex antenna gains are calibrated to correct the corresponding electric field spectra.
+These calibrated electric fields are gridded using an antenna-based gridding convolution function after which it is spatially Fourier transformed and squared to obtain an image cube for every time stamp.
+These images are then time-averaged to obtain the accumulated image $\mathcal{I}^\prime(f)$ ($\boldsymbol{\mathcal{I}}(\hat{\mathbf{s}})$ in matrix notation).
 
@@
   % \includegraphics[width=\columnwidth]{MOFF_flowchart.eps}
-  \caption{A flowchart of MOFF imaging in EPIC. The propagated electric fields
-    shown on the left are measured as time-series $E_1(t)\ldots E_\textrm{N}(t)$
-    by the antennas $\textrm{A}_1\ldots \textrm{A}_\textrm{N}$ which are then 
-    Fourier transformed by the F-engine to produce electric field spectra 
-    $\widetilde{E}_1(f)\ldots \widetilde{E}_\textrm{N}(f)$. They are calibrated 
-    and gridded. The gridded electric fields $\widetilde{E}_\textrm{g}(f)$ from 
-    each time series are imaged to produce images 
-    $I_1(f)\ldots I_\textrm{M}(f)$. These images are time-averaged to obtain the 
-    final image $I(f)$.}
+  \caption{A flowchart of MOFF imaging in EPIC.
+The propagated electric fields shown on the left are measured as time-series $\widetilde{E}_1(t)\ldots \widetilde{E}_\textrm{N}(t)$ by the antennas $\textrm{A}_1\ldots \textrm{A}_\textrm{N}$ which are then Fourier transformed by the F-engine to produce electric field spectra $E_1(f)\ldots E_\textrm{N}(f)$.
+They are calibrated and gridded.
+The gridded electric fields $E_\textrm{g}(f)$ from each time series are imaged to produce images $\mathcal{I}^\prime_1(f)\ldots \mathcal{I}^\prime_\textrm{M}(f)$.
+These images are time-averaged to obtain the final image $\mathcal{I}^\prime(f)$.}
   \label{fig:MOFF-flowchart}
@@
 
-Fig.~\ref{fig:FX-flowchart} shows the flowchart for a visibility-based software 
-holographic imaging from a FX correlator. The antenna-based F-engine is identical 
-to that in the MOFF processing. The electric field spectra from each antenna are 
-then cross-multiplied in the X-engine with those from all other antennas to 
-obtain the visibilities $\widetilde{V}_\textrm{p}(f)$ ($\mathbf{m}(\mathbf{v})$ 
-in matrix notation). They are calibrated and time-averaged to obtain 
-$\langle \widetilde{V}_\textrm{p}(f)\rangle$ which are then gridded and imaged to 
-obtain the image $I(f)$. The $I(f)$ obtained from both techniques are 
-theoretically identical as explained in \S\ref{sec:math}.
+Fig.~\ref{fig:FX-flowchart} shows the flowchart for a visibility-based software holographic imaging from a FX correlator.
+The antenna-based F-engine is identical to that in the MOFF processing.
+The electric field spectra from each antenna are then cross-multiplied in the X-engine with those from all other antennas to obtain the visibilities $V_\textrm{ab}(f,t)$ ($\mathbf{m}(\mathbf{v})$ in matrix notation).
+They are calibrated and time-averaged to obtain $\langle V_\textrm{ab}(f)\rangle$ which are then gridded and imaged to obtain the image $\mathcal{I}^\prime(f)$.
+The $\mathcal{I}^\prime(f)$ obtained from both techniques are theoretically identical as explained in \S\ref{sec:math}.
 
@@
   \caption{A flowchart of visibility-based software holographic imaging in EPIC.
-    The FX process flow shares the F-engine with the MOFF process. Following the 
-    F-engine, the electric fields pass through the X-engine to produce 
-    visibilities $\widetilde{V}_\textrm{p}(f)$ which are calibrated and 
-    time-averaged. Then they are gridded to obtain the gridded visibilities 
-    $\widetilde{V}_\textrm{g}(f)$ which are then Fourier transformed to obtain 
-    the image, $I(f)$.}
+The FX process flow shares the F-engine with the MOFF process.
+Following the F-engine, the electric fields pass through the X-engine to produce visibilities $V_\textrm{ab}(f,t)$ which are calibrated and time-averaged.
+Then they are gridded to obtain the gridded visibilities $V_\textrm{g}(f)$ which are then Fourier transformed to obtain the image, $\mathcal{I}^\prime(f)$.}
   \label{fig:FX-flowchart}
@@
 The gridding kernel in the simplest case is given by the antenna aperture
-illumination function, $\widetilde{B}(\mathbf{r}-\mathbf{r}_a)$, which can
+illumination function, $W(\mathbf{r}-\mathbf{r}_a)$, which can
 be specified either by a functional form or as a table of values against 
@@
 \par\medskip
-\noindent The antenna array aperture illumination over the entire grid,
-$\widetilde{\mathbf{W}}(\mathbf{r})$, is obtained by a projection of the 
-individual antenna aperture illuminations:
+\noindent The antenna array aperture illumination over the entire grid, $W(\mathbf{r})$, is obtained by a projection of the individual antenna aperture illuminations:
 \begin{align}\label{eqn:gridding-convolution}
-  \widetilde{\mathbf{W}}(\mathbf{r}) &= \sum_a \widetilde{\mathbf{W}}_a(\mathbf{r}-\mathbf{r}_a) \\
-                            &= \mathbf{M}(\mathbf{r},\mathbf{a})\,\mathcal{I}(\mathbf{a}),
-\end{align}
-where, $\mathcal{I}(\mathbf{a})$ is a row of ones. This is achieved by
-efficient multiplication with the sparse matrix created in the antenna-to-grid
-mapping process using the sparse matrices module in Python SciPy package. 
-Unless $\widetilde{\mathbf{W}}(\mathbf{r})$ includes time-dependent
-effects of the ionosphere or the instrument, it needs to be computed just once
-for the entire observation. However, the gridding of electric fields must be
-computed at every readout of the electric field spectra,
+  W(\mathbf{r}) &= \sum_a W_a(\mathbf{r}-\mathbf{r}_a),
+\end{align}
+or in matrix notation,
 \begin{align}
-  \widetilde{\mathbf{E}}(\mathbf{r}) &= \mathbf{M}(\mathbf{r},\mathbf{a})\,\widetilde{\mathbf{E}}(\mathbf{a}).
+  \mathbf{W}(\mathbf{r}) &= \mathbf{M}(\mathbf{r},\mathbf{a})\,\mathbf{I}(\mathbf{a}),
+\end{align}
+where, $\mathbf{I}(\mathbf{a})$ is a row of ones denoting coverage of the grid by kernel footprints of antennas.
+This is achieved by efficient multiplication with the sparse matrix created in the antenna-to-grid mapping process using the sparse matrices module in Python SciPy package.
+Unless $\mathbf{W}(\mathbf{r})$ includes time-dependent effects of the ionosphere or the instrument, it needs to be computed just once for the entire observation.
+However, the gridding of electric fields must be computed at every readout of the electric field spectra,
+\begin{align}
+  \mathbf{E}(\mathbf{r}) &= \mathbf{M}(\mathbf{r},\mathbf{a})\,\mathbf{E}(\mathbf{a}).
 \end{align}
@@
 plane and subtract the estimated auto-correlation kernel scaled to the peak 
-value centreed at the zero spacing pixel. The final averaged image is obtained 
+value centred at the zero spacing pixel. The final averaged image is obtained 
 by Fourier transforming the $uv$ plane data and weights with the 
@@
 plane antenna power response (which is theoretically identical to the correlation 
-of individual antenna footprints) centreed on a baseline has a loss in precision 
+of individual antenna footprints) centred on a baseline has a loss in precision 
 at most to first order. Thus, although in the limit of infinitesimally small 
@@
 mitigated by making the grid spacing finer at the expense of increased 
-computational cost. Residuals centreed around zero spacing can also be lowered 
+computational cost. Residuals centred around zero spacing can also be lowered 
 by subtracting each auto-correlation of antenna weights separately by using the 
@@
 in New Mexico. This data is in LWA narrow-band transient buffer (TBN) format 
-from 255 antennas within roughly a diameter of 100~m. The data is centreed at a 
+from 255 antennas within roughly a diameter of 100~m. The data is centred at a 
 frequency of 74.03~MHz, with a sample rate (equal to the bandwidth) of 100~kHz 
@@
 stack, accumulate and average images, and optionally remove the antenna
-auto-correlations centreed around the zero-spacing pixel in the $uv$ plane. 
+auto-correlations centred around the zero-spacing pixel in the $uv$ plane. 
 It also handles all four cross-polarization products. Currently, it supports 

### 2016-08-13 imaging_demo/ms.tex
Add section on imaging with heterogeneous arrays
From e6e6612 (mnras):
====================

(*) Add section on imaging with heterogeneous arrays
(*) Make associated changes in Abstract, Introduction and Conclusions

@@
 %\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
-\newcommand{\Nant}{N_\textrm{a}}
+\newcommand{\Nant}{N_\textrm{A}}
 \newcommand{\Ngrid}{N_\textrm{g}} 
@@
 \begin{abstract}
-Modern radio telescopes are favouring densely packed array layouts with 
-large numbers of antennas ($\Nant\gtrsim 1000$). Since the complexity of 
-traditional correlators scales as $\mathcal{O}(\Nant^2)$, there will be a steep 
-cost for realizing the full imaging potential of these powerful instruments. 
-Through our generic and efficient E-field Parallel Imaging Correlator (EPIC), 
-we present the first software demonstration of a generalized direct imaging 
-algorithm, namely, the Modular Optimal Frequency Fourier (MOFF) imager. It takes 
-advantage of the multiplication-convolution theorem of Fourier transforms. Not 
-only does it bring down the cost for dense layouts to 
-$\mathcal{O}(\Nant\log_2\Nant)$ but can also image from irregularly arranged 
-heterogeneous antenna. EPIC is highly modular and parallelizable, implemented in 
-object oriented Python, and publicly available. We have verified the images 
-produced to be equivalent to those produced using traditional techniques to 
-within a precision determined by coarseness of gridding. We have also validated 
-our implementation on data observed with the Long Wavelength Array (LWA). Antenna 
-layouts with a dense filling factor consisting of a large number of antennas such 
-as LWA, the Square Kilometre Array, Hydrogen Epoch of Reionization Array, and 
-Canadian Hydrogen Intensity Mapping Experiment will gain significant 
-computational advantage by deploying EPIC. Inherent availability of calibrated 
-time-domain images on digitizer writeout time-scales and vastly lower I/O 
-bandwidth relative to visibility-based systems will make it a prime candidate 
-for transient searches of Fast Radio Bursts (FRB) as well as planetary and 
-exoplanetary phenomena. 
+Modern radio telescopes are favouring densely packed array layouts with large numbers of antennas ($\Nant\gtrsim 1000$).
+Since the complexity of traditional correlators scales as $\mathcal{O}(\Nant^2)$, there will be a steep cost for realizing the full imaging potential of these powerful instruments.
+Through our generic and efficient E-field Parallel Imaging Correlator (EPIC), we present the first software demonstration of a generalized direct imaging algorithm, namely, the Modular Optimal Frequency Fourier (MOFF) imager.
+It takes advantage of the multiplication-convolution theorem of Fourier transforms.
+Not only does it bring down the cost for dense layouts to $\mathcal{O}(\Nant\log_2\Nant)$ but can also image from irregularly arranged heterogeneous antenna.
+We have demonstrated its versatility to image with heterogeneous arrays by showing that it robustly estimates the input sky model while wrong assumptions about array homogeneity lead to severe mis-estimates.
+EPIC is highly modular and parallelizable, implemented in object oriented Python, and publicly available.
+We have verified the images produced to be equivalent to those produced using traditional techniques to within a precision set by gridding coarseness.
+We have also validated our implementation on data observed with the Long Wavelength Array (LWA).
+Antenna layouts with dense filling factors consisting of a large number of antennas such as LWA, the Square Kilometre Array, Hydrogen Epoch of Reionization Array, and Canadian Hydrogen Intensity Mapping Experiment will gain significant computational advantage by deploying EPIC.
+Inherent availability of calibrated time-domain images on digitizer writeout time-scales and vastly lower I/O bandwidth relative to visibility-based systems will make it a prime candidate for transient searches of Fast Radio Bursts (FRB) as well as planetary and exoplanetary phenomena.
 \end{abstract}
@@
 
-We begin with a technical description of the algorithm in \S\ref{sec:math}, 
-then discuss our particular implementation in \S\ref{sec:software}. We then 
-verify the output data quality from our code in \S\ref{sec:verify} by presenting 
-simulated images from both the EPIC correlator and comparing to a simulated FX 
-correlator. We also demonstrate the performance with real-world data from the 
-LWA. In \S \ref{sec:analysis}, we explore the scalability of the algorithm in the 
-context of several array design choices. We identify specific array design 
-classes where the EPIC correlator -- is computationally more efficient; and in 
-the field of transients, demands significantly lesser I/O bandwidth relative to 
-visibility-based approaches. We conclude and discuss future research prospects 
-in \S\ref{sec:conclusions}.
+We begin with a technical description of the algorithm in \S\ref{sec:math}, then discuss our particular implementation in \S\ref{sec:software}.
+We then verify the output data quality from our code in \S\ref{sec:verify} by presenting simulated images from both the EPIC correlator and comparing to a simulated FX correlator.
+We also demonstrate the performance with real-world data from the LWA.
+\S\ref{sec:versatility} demonstrates its ability to robustly image data from heterogeneous arrays.
+In \S\ref{sec:analysis}, we explore the scalability of the algorithm in the context of several array design choices.
+We identify specific array design classes where the EPIC correlator -- is computationally more efficient; and in the field of transients, demands significantly lesser I/O bandwidth relative to visibility-based approaches.
+We conclude and discuss future research prospects in \S\ref{sec:conclusions}.
 
@@
 The result is the visibility expressed in terms of the sky brightness, the power response, and uncorrelated noise terms which we group into $n_{ab}(f,t)$.
-\begin{align}
+\begin{align}\label{eqn:measurement-eqn-1}
   V_{ab}(f,t) &= \int \mathcal{W}_{ab}(\hat{\mathbf{s}},f,t)\,\mathcal{I}(\hat{\mathbf{s}},f,t)\,e^{-2\pi i f\mathbf{r}_{ab}\!\cdot\,\hat{\mathbf{s}}/c}\,\dif\Omega \nonumber\\
@@
 \begin{align}
-\mathbf{B}^{\,\textrm{T}}(\mathbf{u},\mathbf{v})\,\mathbf{N}^{-1}(\mathbf{v} ,\mathbf{v})\,\mathbf{m}(\mathbf{v}) &= \left\langle \left[\mathbf{W}_a(\mathbf{r},\mathbf{a})\,\mathbf{N}^{-\mathrm{T}}\!(\mathbf{a},\mathbf{a})\, \mathbf{E}^\star(\mathbf{a})\right]\right.
+\mathbf{B}^{\,\textrm{T}}(\mathbf{u},\mathbf{v})\,\mathbf{N}^{-1}(\mathbf{v} ,\mathbf{v})\,\mathbf{m}(\mathbf{v}) &= \left\langle \left[\mathbf{W}_a(\mathbf{r},\mathbf{a})\,\widetilde{\mathbf{N}}^\textrm{T}\!(\mathbf{a},\mathbf{a})\, \mathbf{E}^\star(\mathbf{a})\right]\right.
 \nonumber\\
-&\quad\ast\left.\left[\mathbf{W}^\textrm{T}_a(\mathbf{r},\mathbf{a})\,\mathbf{N}^{-1}\!(\mathbf{a},\mathbf{a})\,\mathbf{E}(\mathbf{a})\right]\right\rangle_t\label{eqn:e-field-conv}
-\end{align}
-
-We can then use the multiplication-convolution theorem to move the convolution in Equation \ref{eqn:e-field-conv} to a square after the Fourier transform in Equation~\ref{eqn:dirty-image-FX}.
+&\,\quad\ast\left.\left[\mathbf{W}^\textrm{T}_a(\mathbf{r},\mathbf{a})\,\widetilde{\mathbf{N}}\!(\mathbf{a},\mathbf{a})\,\mathbf{E}(\mathbf{a})\right]\right\rangle_t, \label{eqn:e-field-conv}
+\end{align}
+where, $\mathbf{N}^{-1}$ has been expressed as $\widetilde{\mathbf{N}}^\textrm{T}\widetilde{\mathbf{N}}$.
+
+We can then use the multiplication-convolution theorem to move the convolution in Equation~\ref{eqn:e-field-conv} to a square after the Fourier transform in equation~\ref{eqn:dirty-image-FX}.
 \begin{align}
@@
 
-Equation \ref{eqn:dirty-image-MOFF} is the optimal imaging equation used by the 
-MOFF algorithm. While mathematically equivalent to Equation~
-\ref{eqn:dirty-image-FX}, squaring in image space rather than convolving in $uv$ 
-space potentially saves orders of magnitude in computation.
+Equation~\ref{eqn:dirty-image-MOFF} is the optimal imaging equation used by the MOFF algorithm.
+While mathematically equivalent to equation~\ref{eqn:dirty-image-FX}, squaring in image space rather than convolving in $uv$ space potentially saves orders of magnitude in computation.
 
@@
 in these regions is the same. For instance, note the difference in weights at 
-$(u,v)\approx (2,-10)$. The reason for these differences is described below.
+$(u,v)\approx (2.5,4)$. The reason for these differences is described below.
 
@@
 
-\section{Imaging with Heterogeneous Antenna Array}\label{sec:versatility}
+\section{Imaging with Heterogeneous Antenna Arrays}\label{sec:versatility}
+
+One of the advanced features of the MOFF algorithm and its implementation via EPIC is the ability to naturally account for heterogeneity of antennas while still producing images with maximal information \citep{mor11}.
+Unlike the assumption frequently made in standard radio interferometry that all antennas are identical, EPIC can implement the most generic case where no such assumption can be made.
+In other words, EPIC does not assume $W_a(\mathbf{r}-\mathbf{r}_a)$ or $W_{ab}(\mathbf{r}-\mathbf{r}_{ab})$ for different antennas are identical.
+However, any {\it a priori} knowledge available that certain sets of antennas have identical illumination patterns can be easily passed on to EPIC to avoid redundant computations and thus increase efficiency.
+
+We present a methodology to understand the effective angular weighting in the image obtained with EPIC from a heterogeneous array and compare it with that from standard interferometry that assumes a homogeneous array (all antennas are identical).
+We also demonstrate how the assumption in the latter results in a mis-estimation of recovered flux densities of sources.
+
+Dropping the noise term and time dependence, the visibility measured by an antenna pair can be written from equation~\ref{eqn:measurement-eqn-1} as:
+\begin{align}\label{eqn:measurement-eqn-2}
+  V_{ab}(f) &= \int \mathcal{W}^\textrm{I}_a(\hat{\mathbf{s}},f)\mathcal{W}^{\textrm{I}\star}_b(\hat{\mathbf{s}},f)\,\mathcal{I}(\hat{\mathbf{s}},f)\,e^{-2\pi i f\mathbf{r}_{ab}\!\cdot\,\hat{\mathbf{s}}/c}\,\dif\Omega,
+\end{align}
+where, superscript `$\textrm{I}$' in the $\mathcal{W}$ term indicates the weighting is inherently introduced by the instrument during the measurement process. The dirty beam-weighted image is:
+\begin{align}
+  \mathcal{I}_\textrm{D}(\hat{\mathbf{s}},f) &= \frac{1}{\Nant(\Nant-1)/2}\,\sum_{ab}\, V_{ab}(f)\,e^{2\pi i f\mathbf{r}_{ab}\!\cdot\,\hat{\mathbf{s}}/c}.
+\end{align}
+
+When the array is homogeneous, $\mathcal{W}^\textrm{I} \equiv \mathcal{W}^\textrm{I}_a \equiv \mathcal{W}^\textrm{I}_b$ and the above equation reduces to:
+\begin{align}\label{eqn:dirty-image-homogeneous}
+  \mathcal{I}_\textrm{D}(\hat{\mathbf{s}},f) &= |\mathcal{W}^\textrm{I}(\hat{\mathbf{s}},f)|^2\,\mathcal{I}_\textrm{D}^\textrm{iso}(\hat{\mathbf{s}},f),
+\end{align}
+where, $\mathcal{I}_\textrm{D}^\textrm{iso}(\hat{\mathbf{s}},f)$ is the dirty image with no beam weighting (isotropic uniform weighting) determined only by the array layout and $|\mathcal{W}^\textrm{I}|^2$ is the directional power pattern of the antenna pair familiar in standard interferometry.
+
+While imaging, EPIC introduces a weighting to each antenna during gridding as given by equations~\ref{eqn:dirty-image-FX} and \ref{eqn:e-field-conv}, and equivalently, the weighted visibility is:
+\begin{align}
+  V^\prime_{ab}(\mathbf{r},f) &= B^{\textrm{G}\star}_{ab}(\mathbf{r}-\mathbf{r}_{ab},f)\,V_{ab},
+\end{align}
+where, $B^\textrm{G}_{ab}(\mathbf{r}-\mathbf{r}_{ab},f)$ is obtained by the spatial cross-correlation of weighting kernels associated with the individual antennas using the definition:
+\begin{align}
+  B^\textrm{G}_{ab}(\mathbf{r}-\mathbf{r}_{ab},f) &= \int W^\textrm{G}_a(\mathbf{r}^\prime-\mathbf{r}_a)\,W^{\textrm{G}\star}_b(\mathbf{r}^\prime+\mathbf{r}-\mathbf{r}_a)\,\dif^2\mathbf{r}^\prime,
+\end{align}
+and the superscript `$\textrm{G}$' denotes the weighting introduced in analysis during the gridding process.
+Optimal imaging requires $W^\textrm{G}_a = W^\textrm{I}_a$ \citep{mor09,mor11}.
+However, we keep the two superscripts to describe output images under the wrong assumption that the array is homogeneous when instead it is actually heterogeneous.
+
+Thus the sky response of the weighted visibility is:
+\begin{align}
+  \mathcal{I}^\prime_{ab}(\hat{\mathbf{s}},f) &= \int V^\prime_{ab}(\mathbf{r},f)\,e^{2\pi i f\mathbf{r}\cdot\,\hat{\mathbf{s}}/c}\,\dif^2\mathbf{r} \nonumber\\
+                                              &= \int B^{\textrm{G}\star}_{ab}(\mathbf{r}-\mathbf{r}_{ab},f)\,V_{ab}\,e^{2\pi i f\mathbf{r}\cdot\,\hat{\mathbf{s}}/c}\,\dif^2\mathbf{r} \nonumber\\
+                                              &= V_{ab}\,\int \left[\int W^{\textrm{G}\star}_a(\mathbf{r}^\prime-\mathbf{r}_a)\,W^\textrm{G}_b(\mathbf{r}^\prime+\mathbf{r}-\mathbf{r}_a)\,\dif^2\mathbf{r}^\prime\right] \nonumber\\
+                                              &\qquad\qquad \times e^{2\pi i f\mathbf{r}\cdot\,\hat{\mathbf{s}}/c}\,\dif^2\mathbf{r} \nonumber\\
+                                              &= \mathcal{W}^{\textrm{G}\star}_a(\hat{\mathbf{s}},f)\mathcal{W}^\textrm{G}_b(\hat{\mathbf{s}},f)\,V_{ab}\,e^{2\pi i f\mathbf{r}_{ab}\!\cdot\,\hat{\mathbf{s}}/c}
+\end{align}
+and denotes the fringe from the weighted visibility. The image output of EPIC is equivalent to summing these weighted fringes from all baselines:
+\begin{align}\label{eqn:wt-dirty-image-EPIC}
+  \mathcal{I}^\prime(\hat{\mathbf{s}},f) &= \frac{1}{\Nant(\Nant-1)/2}\,\sum_{\substack{ab\\a\ne b}}\,\mathcal{I}^\prime_{ab}(\hat{\mathbf{s}},f) \nonumber\\
+                                         &= \frac{1}{\Nant(\Nant-1)/2}\,\sum_{\substack{ab\\a\ne b}}\,\left[\mathcal{W}^{\textrm{G}\star}_a(\hat{\mathbf{s}},f)\mathcal{W}^\textrm{G}_b(\hat{\mathbf{s}},f)\,V_{ab}\right.
+\nonumber\\
+  &\qquad\qquad\qquad\qquad\qquad\qquad \times \left. e^{2\pi i f\mathbf{r}_{ab}\!\cdot\,\hat{\mathbf{s}}/c}\right].
+\end{align}
+
+If all antennas are identically weighted ($W^\textrm{G} \equiv W^\textrm{G}_a \equiv W^\textrm{G}_b$), using equation~\ref{eqn:dirty-image-homogeneous}, equation~\ref{eqn:wt-dirty-image-EPIC} reduces to:
+\begin{align}\label{eqn:wt-dirty-image-homogeneous}
+  \mathcal{I}^\prime(\hat{\mathbf{s}},f) &= \frac{\left|\mathcal{W}^\textrm{G}(\hat{\mathbf{s}},f)\right|^2}{\Nant(\Nant-1)/2}\,\sum_{\substack{ab\\a\ne b}}\,V_{ab}\,e^{2\pi i f\mathbf{r}_{ab}\!\cdot\,\hat{\mathbf{s}}/c} \nonumber\\
+  &= \left|\mathcal{W}^\textrm{G}(\hat{\mathbf{s}},f)\right|^2\,\mathcal{I}_\textrm{D}(\hat{\mathbf{s}},f) \nonumber\\
+  &= \left|\mathcal{W}^\textrm{G}(\hat{\mathbf{s}},f)\right|^2\,\left|\mathcal{W}^\textrm{I}(\hat{\mathbf{s}},f)\right|^2\,\mathcal{I}_\textrm{D}^\textrm{iso}(\hat{\mathbf{s}},f).
+\end{align}
+Thus, the dirty image, which is already attenuated by the instrumental power pattern, gets further attenuated by the power pattern introduced in the gridding step in EPIC imager.
+This is consistent with \citet{mor09}.
+
+We consider a heterogeneous array consisting of a total of $\Nant$ antennas under $N_\textrm{T}$ different types.
+And under each antenna type $i$, there are $n_i$ antennas such that $\sum_i n_i = \Nant$.
+The total number of unique non-zero spacing antenna pairs is $\Nant(\Nant-1)/2$ and there are potentially $N_\textrm{T}(N_\textrm{T}+1)/2$ unique types of antenna pairs obtained by pairwise combination of $N_\textrm{T}$ antenna types.
+Thus,
+\begin{align}\label{eqn:total-baselines}
+  \frac{1}{2}\Nant(\Nant-1) &= \frac{1}{2}\,\mathop{\sum_{i=1}^{N_\textrm{T}}\sum_{j=1}^{N_\textrm{T}}}_{i\ne j} n_i\,n_j + \frac{1}{2}\,\sum_{i=1}^{N_\textrm{T}} n_i(n_i-1)
+\end{align}
+is obtained by counting antenna pairs from different antenna types (first term) and those from the same antenna type (second term).
+
+Since equation~\ref{eqn:wt-dirty-image-EPIC} is obtained by summing weighted fringes over all unique antenna pairs, it can be expressed as: 
+\begin{align}\label{eqn:wt-dirty-image-EPIC-decomp}
+  \mathcal{I}^\prime(\hat{\mathbf{s}},f) &= \frac{1}{\Nant(\Nant-1)/2}\,\left[\frac{1}{2}\mathop{\sum_{i=1}^{N_\textrm{T}}\sum_{j=1}^{N_\textrm{T}}}_{i\ne j}\sum_{\substack{ab\\a\ne b\\a\in i, b\in j}}\,\mathcal{I}^\prime_{ab}(\hat{\mathbf{s}},f)\right.
+\nonumber\\
+  &\qquad\qquad\qquad\qquad\qquad\quad + \left.\frac{1}{2}\sum_{i=1}^{N_\textrm{T}}\sum_{\substack{ab\\a\ne b\\a,b\in i}}\,\mathcal{I}^\prime_{ab}(\hat{\mathbf{s}},f)\right]. 
+\end{align}
+Here, equation~\ref{eqn:wt-dirty-image-EPIC} has been decomposed as the sum of fringes over all unique antenna pairs in an antenna pair type (innermost summation) and subsequently summed over all antenna pair types.
+
+In order to simplify the analysis, we assume that the spatial distribution of antenna pairs in each antenna pair type are similar and thus result in a similar dirty image, $\mathcal{I}_\textrm{D}^\textrm{iso}(\hat{\mathbf{s}},f)$.
+This usually holds when $\Nant$ is large and antennas under each antenna type are chosen randomly such that differences in the point spread functions are insignificant.
+Under this assumption, equation~\ref{eqn:wt-dirty-image-EPIC-decomp} can be written as:
+\begin{align}\label{eqn:wt-dirty-image-EPIC-decomp-approx1}
+  \mathcal{I}^\prime(\hat{\mathbf{s}},f) &\approx \frac{\mathcal{I}_\textrm{D}^\textrm{iso}(\hat{\mathbf{s}},f)}{\Nant(\Nant-1)/2} \nonumber\\
+  &\,\times \left[\frac{1}{2}\mathop{\sum_{i=1}^{N_\textrm{T}}\sum_{j=1}^{N_\textrm{T}}}_{i\ne j}\sum_{\substack{ab\\a\ne b\\a\in i, b\in j}}\left\{\mathcal{W}^{\textrm{G}\star}_a(\hat{\mathbf{s}},f)\mathcal{W}^\textrm{G}_b(\hat{\mathbf{s}},f)\right.\right.\nonumber\\
+  &\qquad\qquad\qquad\qquad\qquad\qquad \left. \times \mathcal{W}^\textrm{I}_a(\hat{\mathbf{s}},f)\mathcal{W}^{\textrm{I}\star}_b(\hat{\mathbf{s}},f)\right\} \nonumber\\ 
+  &\qquad + \frac{1}{2}\sum_{i=1}^{N_\textrm{T}}\sum_{\substack{ab\\a\ne b\\a,b\in i}}\left\{\mathcal{W}^{\textrm{G}\star}_a(\hat{\mathbf{s}},f)\mathcal{W}^\textrm{G}_b(\hat{\mathbf{s}},f)\right.
+\nonumber\\
+  &\qquad\qquad\qquad\qquad\qquad\qquad \left.\left.\times\mathcal{W}^\textrm{I}_a(\hat{\mathbf{s}},f)\mathcal{W}^{\textrm{I}\star}_b(\hat{\mathbf{s}},f)\right\}\vphantom{\sum_{\substack{ab\\a\ne b\\a\in i, b\in j}}}\right],
+\end{align}
+where, $\mathcal{I}_\textrm{D}^\textrm{iso}(\hat{\mathbf{s}},f)$ is approximately the same for all terms and has been pulled outside the summations.
+For all antennas indexed by $a$ that belong to a particular type $i$, we replace $\mathcal{W}_a$ with $\mathcal{W}_{(i)}$ that arise from both instrumental (superscript `$I$') and gridding (superscript `$G$') origins.
+Using arguments similar to those in equation~\ref{eqn:total-baselines}, equation~\ref{eqn:wt-dirty-image-EPIC-decomp-approx1} can be further simplified to:
+\begin{align}\label{eqn:wt-dirty-image-EPIC-decomp-approx2}
+  \mathcal{I}^\prime(\hat{\mathbf{s}},f) &\approx \frac{\mathcal{I}_\textrm{D}^\textrm{iso}(\hat{\mathbf{s}},f)}{\Nant(\Nant-1)/2} \nonumber\\
+  &\,\times \left[\frac{1}{2}\mathop{\sum_{i=1}^{N_\textrm{T}}\sum_{j=1}^{N_\textrm{T}}}_{i\ne j}n_in_j\left\{\mathcal{W}^{\textrm{G}\star}_{(i)}(\hat{\mathbf{s}},f)\mathcal{W}^\textrm{G}_{(j)}(\hat{\mathbf{s}},f)\right.\right.\nonumber\\
+  &\qquad\qquad\qquad\qquad\qquad\qquad \left. \times \mathcal{W}^\textrm{I}_{(i)}(\hat{\mathbf{s}},f)\mathcal{W}^{\textrm{I}\star}_{(j)}(\hat{\mathbf{s}},f)\right\} \nonumber\\ 
+  &\qquad + \frac{1}{2}\sum_{i=1}^{N_\textrm{T}}n_i(n_i-1)\,\left|\mathcal{W}^\textrm{G}_{(i)}(\hat{\mathbf{s}},f)\right|^2\,\left|\mathcal{W}^\textrm{I}_{(i)}(\hat{\mathbf{s}},f)\right|^2\left.\vphantom{\sum_{\substack{ab\\a\ne b\\a\in i, b\in j}}}\right].
+\end{align}
+Then the effective attenuation of the dirty image due to instrumental and gridding weights is:
+\begin{align}\label{eqn:effective-weighting}
+  \mathcal{W}_\textrm{eff} &= \mathcal{I}^\prime(\hat{\mathbf{s}},f)\, / \,\mathcal{I}_\textrm{D}^\textrm{iso}(\hat{\mathbf{s}},f).
+\end{align}
+This means that the effective attenuation of the dirty image is obtained by combinations of the fourth power of antenna voltage patterns each weighted by the number of antenna pairs in those antenna pair types.
+In an optimally weighted image, $\mathcal{W}^\textrm{G}_{(i)} \equiv \mathcal{W}^\textrm{I}_{(i)}$ and thus, equation~\ref{eqn:effective-weighting} reduces to:
+\begin{align}\label{eqn:effective-weighting-optimal}
+  \mathcal{W}_\textrm{eff}^\textrm{opt} &\approx \frac{1}{\Nant(\Nant-1)/2} \nonumber\\
+  &\quad \times \left[\frac{1}{2}\mathop{\sum_{i=1}^{N_\textrm{T}}\sum_{j=1}^{N_\textrm{T}}}_{i\ne j}n_in_j\,\left|\mathcal{W}^\textrm{I}_{(i)}(\hat{\mathbf{s}},f)\right|^2\left|\mathcal{W}^\textrm{I}_{(j)}(\hat{\mathbf{s}},f)\right|^2\right.\nonumber\\
+  &\qquad\qquad + \frac{1}{2}\sum_{i=1}^{N_\textrm{T}}n_i(n_i-1)\,\left|\mathcal{W}^\textrm{I}_{(i)}(\hat{\mathbf{s}},f)\right|^4\left.\vphantom{\sum_{\substack{ab\\a\ne b\\a\in i, b\in j}}}\right],
+\end{align}
+where the superscript `$\textrm{opt}$' denotes 'optimal'.
+
+For purposes of illustration, we consider a simple scenario in which there are two antenna types each containing roughly equal numbers of antennas with similar layouts but the analysis is performed under the wrong assumption that all antennas are identical and are of the first type.
+In an optimally weighted image, the optimal effective weighting given by equation~\ref{eqn:effective-weighting} will be:
+\begin{align}\label{eqn:effective-weighting-optimal-example}
+  \mathcal{W}_\textrm{eff}^\textrm{opt} &\approx \frac{1}{\Nant(\Nant-1)}\,\left[n_1n_2\,\left|\mathcal{W}^\textrm{I}_{(1)}(\hat{\mathbf{s}},f)\right|^2\,\left|\mathcal{W}^\textrm{I}_{(2)}(\hat{\mathbf{s}},f)\right|^2\right.\nonumber\\
+  &\qquad\qquad\qquad\qquad + n_1(n_1-1)\,\left|\mathcal{W}^\textrm{I}_{(1)}(\hat{\mathbf{s}},f)\right|^4 \nonumber\\
+  &\qquad\qquad\qquad\qquad \left. + n_2(n_2-1)\,\left|\mathcal{W}^\textrm{I}_{(2)}(\hat{\mathbf{s}},f)\right|^4\right].
+\end{align}
+On the other hand, $\mathcal{W}^\textrm{G}_{(1)}(\hat{\mathbf{s}},f) = \mathcal{W}^\textrm{I}_{(1)}(\hat{\mathbf{s}},f)$ and $\mathcal{W}^\textrm{G}_{(2)}(\hat{\mathbf{s}},f) = \mathcal{W}^\textrm{I}_{(1)}(\hat{\mathbf{s}},f)$ for an image obtained in the erroneous case in which all antennas are assumed to be identical.
+It will be sub-optimal whose effective weighting will be:
+\begin{align}\label{eqn:effective-weighting-optimal-example}
+  \mathcal{W}_\textrm{eff}^\textrm{sub} &\approx \frac{1}{\Nant(\Nant-1)}\,\left[\left\{n_1n_2\,\left|\mathcal{W}^\textrm{I}_{(1)}(\hat{\mathbf{s}},f)\right|^2 \right.\right. \nonumber\\
+  &\qquad\qquad\qquad\qquad\qquad \left.\times\mathcal{W}^\textrm{I}_{(1)}(\hat{\mathbf{s}},f)\mathcal{W}^{\textrm{I}\star}_{(2)}(\hat{\mathbf{s}},f)\right\}\nonumber\\
+  &\qquad\qquad\qquad\qquad + n_1(n_1-1)\,\left|\mathcal{W}^\textrm{I}_{(1)}(\hat{\mathbf{s}},f)\right|^4 \nonumber\\
+  &\qquad\qquad\qquad\qquad + n_2(n_2-1)\,\left|\mathcal{W}^\textrm{I}_{(1)}(\hat{\mathbf{s}},f)\right|^2\nonumber\\
+  &\qquad\qquad\qquad\qquad\qquad\qquad\qquad \left.\times\left|\mathcal{W}^\textrm{I}_{(2)}(\hat{\mathbf{s}},f)\right|^2\right].
+\end{align}
+Finally, since no {\it a priori} information will be available about the heterogeneous array in case of the erroneous assumption, the effective weighting in the image will be believed to be $\mathrm{W}_\textrm{eff}^\textrm{err} = \left|\mathcal{W}_{(1)}^\textrm{I}\right|^4$.
+
+Estimates of the dirty images with uniform, isotropic weighting in the optimal and erroneous cases will be obtained by dividing the respective weighted images by their assumed effective weights $\mathrm{W}_\textrm{eff}^\textrm{opt}$ and $\mathrm{W}_\textrm{eff}^\textrm{err}$ respectively as:
+\begin{align}
+  \widehat{\mathcal{I}}_\textrm{D}^\textrm{iso} &\approx \begin{cases}
+    \mathcal{I}_\textrm{D}^\textrm{iso} &\textrm{, \qquad optimal case} \\
+    \mathcal{I}_\textrm{D}^\textrm{iso}\,\frac{\mathcal{W}_\textrm{eff}^\textrm{sub}}{\mathcal{W}_\textrm{eff}^\textrm{err}} &\textrm{, \qquad erroneous case}
+  \end{cases}
+\end{align}
+
+If $\mathcal{I}_\textrm{M}(\hat{\mathbf{s}},f)$ is the sky model, then $\mathcal{I}_\textrm{D}^\textrm{iso}(\hat{\mathbf{s}},f)$ can be expressed as:
+\begin{align}\label{eqn:img-uncertainties}
+  \mathcal{I}_\textrm{D}^\textrm{iso}(\hat{\mathbf{s}},f) &= \mathcal{I}_\textrm{M}(\hat{\mathbf{s}},f) + \delta\mathcal{I}_\textrm{S}(\hat{\mathbf{s}},f) + \delta\mathcal{I}_\textrm{M}(\hat{\mathbf{s}},f),
+\end{align}
+where the second and third terms on the right hand side denote fluctuations due to sidelobes and those intrinsic to the model, respectively.
+We define the normalized estimates of the sky model $\widehat{\mathcal{I}}_\textrm{M}^\textrm{norm} \equiv \widehat{\mathcal{I}}_\textrm{D}^\textrm{iso} / \mathcal{I}_\textrm{M}$.
+Ideally, $\widehat{\mathcal{I}}_\textrm{M}^\textrm{norm}$ should be consistent with unity and not exceed beyond levels of uncertainty as determined by equation~\ref{eqn:img-uncertainties}.
+
+In our simulations, the antenna layout, duration of observation, center frequency and bandwidth used are the same as in \S\ref{sec:sim}.
+Antennas have square apertures.
+Those of the first and second types, which have roughly equal numbers of antennas with similar layouts, have 1.1~m and 6.6~m on their sides respectively.
+$\mathcal{I}_\textrm{M}(\hat{\mathbf{s}},f)$ consists of point sources of random flux densities brighter than 10~Jy along westward and south-westward directions as seen in Fig.~\ref{fig:nonphysical-image}, which shows an optimally weighted image, $\mathcal{I}^\prime(\hat{\mathbf{s}},f)$, obtained with EPIC using the MOFF algorithm after averaging over the entire bandwidth and duration of simulated observation.
 
 \begin{figure*}
-\subfloat[][Simulated Image]{\label{fig:nonphysical-image}\includegraphics[width=0.5\linewidth]{figure10a.eps}}
-\subfloat[][Normalized Radial Flux Density Profile]{\label{fig:norm-flux-density}\includegraphics[width=0.42\linewidth]{figure10b.eps}}
-\caption{.}
+\subfloat[][Optimally weighted EPIC image]{\label{fig:nonphysical-image}\includegraphics[width=0.5\linewidth]{figure10a.eps}}
+\subfloat[][Estimates of normalized radial flux density profile]{\label{fig:norm-flux-density}\includegraphics[width=0.42\linewidth]{figure10b.eps}}
+\caption{(a)~Optimally weighted image, $\mathcal{I}^\prime(\hat{\mathbf{s}},f)$, of simulated sky model obtained by EPIC using MOFF algorithm after averaging over entire bandwidth and duration of simulated observation.
+51 antennas in the inner core of the MWA layout were used with uniformly illuminated square apertures.
+Roughly half of them were squares with 1.1~m sides (first type) and the rest with 6.6~m (second type).
+Point sources were simulated along westward and south-westward directions with random flux densities brighter than 10~Jy.
+Black boxes around point sources are for purposes of collecting image statistics.
+(b)~Normalized flux density estimates, $\widehat{\mathcal{I}}_\textrm{M}^\textrm{norm}$, as a function of distance from image center.
+Bands indicate an approximate ``$1\sigma$'' uncertainty arising from sidelobe confusion and intrinsic randomness of electric fields simulated for point sources in the sky model.
+Gray and red bands correspond to optimal and erroneous cases respectively, where in the latter, all antennas were erroneously assumed to be identical and of the first type.
+The dashed and solid line styles correspond to point sources in the westward and south-westward directions respectively.
+In the optimal case, the normalized estimates are consistent with unity to within ``$1\sigma$'' in almost all regions of the image, whereas in the erroneous case, they are mis-estimated (underestimated) with high significance ($\gtrsim 5\sigma$) in most regions of the image.
+There is no significant difference in estimates between the two different directions of point sources studied.}
 \label{fig:versatility}
 \end{figure*}
+
+$\widehat{\mathcal{I}}_\textrm{D}^\textrm{iso}$ was estimated at the point source locations while the sidelobe levels were estimated from boxes three times bigger on the side than those shown using the `median absolute deviation' metric which is resistant to outliers (other point sources in the sky model in our case).
+The intrinsic fluctuations in the modeled point sources were obtained statistically over time.
+The net uncertainty was determined by adding the levels of sidelobe fluctuations and intrinsic source fluctuations in quadrature.
+The normalized estimates of the sky model, $\widehat{\mathcal{I}}_\textrm{M}^\textrm{norm}$, and the corresponding ``$1\sigma$'' error bars were obtained by dividing $\widehat{\mathcal{I}}_\textrm{D}^\textrm{iso}$ and the net uncertainty by the respective effective weights in either of the cases considered.
+
+Fig.~\ref{fig:norm-flux-density} shows $\widehat{\mathcal{I}}_\textrm{M}^\textrm{norm}$ including the uncertainty as a function of distance from the center of the image.
+The gray and red bands denote estimates from the optimal and the erroneous cases, respectively.
+The solid and dashed lines correspond to point sources in the south-westward and westward directions respectively.
+It can be clearly seen that in the erroneous case, the estimated values appear to be underestimated and deviate from the ideal value of unity with very high significance ($\gtrsim 5\sigma$) in most parts of the image.
+In complete contrast, the estimates in the optimal case in almost all parts of the image are consistent with unity to within $1\sigma$.
+The reason for underestimation is because nearly half the antennas are assumed to be smaller than they actually are and thus results in dividing by overestimated $\mathcal{W}_\textrm{eff}^\textrm{err}$.
+These trends do not seem to be affected by the placement of sources along westward and south-westward directions.
+Thus, we have demonstrated that images from EPIC can robustly image data from heterogeneous arrays while being mathematically optimal.
+Besides being sub-optimal, erroneous assumptions about the homogeneity of an array can lead to significant mis-estimates (in this case, an underestimate) of the sky model.
 
@@
 
-As radio astronomy is entering a new era, advances in instrumentation have to
-be accompanied by equal advances in processing techniques to manage 
-computational resources. Many future radio telescopes such as the SKA, HERA and 
-LWA are headed towards the large-N dense array layout model for which 
-computational cost from traditional FX/XF correlator-based architecture and 
-visibility-based imaging starts rising steeply. We have provided the first 
-software demonstration of a general purpose imaging algorithm using our generic 
-and efficient EPIC software that is designed to bring this cost down from 
-$\mathcal{O}(\Nant^2)$ to $\mathcal{O}(\Ngrid\log_2 \Ngrid)$. Under the class of 
-direct imaging 
-techniques, ours is one of the most generic -- neither does it place any 
-constraint on the array layout to be on a regular grid nor does it require the
-antenna array to be homogeneous.
+As radio astronomy is entering a new era, advances in instrumentation have to be accompanied by equal advances in processing techniques to manage computational resources.
+Many future radio telescopes such as the SKA, HERA and LWA are headed towards the large-N dense array layout model for which computational cost from traditional FX/XF correlator-based architecture and visibility-based imaging starts rising steeply.
+We have provided the first software demonstration of a general purpose imaging algorithm using our generic and efficient EPIC software that is designed to bring this cost down from $\mathcal{O}(\Nant^2)$ to $\mathcal{O}(\Ngrid\log_2 \Ngrid)$.
+Under the class of direct imaging techniques, ours is one of the most generic -- neither does it place any constraint on the array layout to be on a regular grid nor does it require the antenna array to be homogeneous.
+We have demonstrated its natural ability to robustly image data from heterogeneous arrays and shown that wrong assumptions about array homogeneity result in severe mis-estimation of the sky model.
 

### 2016-08-14 imaging_demo/ms.tex
Make minor revisions throughout
From 3f838e8 (mnras):
====================

In ms.tex,

(*) Make minor revisions throughout

In epic.bib,

(*) Add reference to HERA paper by DeBoer et al. (2016)

@@
 It takes advantage of the multiplication-convolution theorem of Fourier transforms.
-Not only does it bring down the cost for dense layouts to $\mathcal{O}(\Nant\log_2\Nant)$ but can also image from irregularly arranged heterogeneous antenna.
-We have demonstrated its versatility to image with heterogeneous arrays by showing that it robustly estimates the input sky model while wrong assumptions about array homogeneity lead to severe mis-estimates.
+Not only does it bring down the cost for dense layouts to $\mathcal{O}(\Nant\log_2\Nant)$ but can also image from irregular layouts and heterogeneous arrays of antennas.
+We demonstrate its versatility to image with heterogeneous arrays by showing that it robustly estimates the input sky model while wrong assumptions about array homogeneity lead to severe mis-estimates.
 EPIC is highly modular and parallelizable, implemented in object oriented Python, and publicly available.
@@
 
-Radio astronomy is entering an era in which interferometers of hundreds to
-thousands of individual antennas are needed to achieve desired survey speeds.
-Nowhere is this more apparent than at radio frequencies below 1.4 GHz. The study
-of the history of hydrogen gas throughout the universe's evolution is pushing
-technology development towards arrays of low-cost antennas with large fields of
-view and densely packed layouts. Similarly, the search for transient objects
-and regular monitoring of the time-dependent sky is driving instruments in the
-same direction with the added requirement of fast read-outs. A number of new 
-telescopes are being or were developed around the world based on this new 
-paradigm, including the Hydrogen Epoch of Reionization 
-Array\footnote{http://reionization.org} (HERA), the Murchison Widefield Array 
-(MWA;\citealt{tin13,bow13}), the Precision Array for Probing the Epoch of 
-Reionization (PAPER; \citealt{par10}), the LOw Frequency ARray 
-(LOFAR;\citealt{van13}), the Canadian Hydrogen Intensity Mapping Experiment 
-(CHIME,\citealt{ban14}), the Long Wavelength Array (LWA, \citealt{ell13}), and 
-the low frequency Square Kilometer Array (SKA1-Low \citealt{mel13}).
+Radio astronomy is entering an era in which interferometers of hundreds to thousands of individual antennas are needed to achieve desired survey speeds.
+Nowhere is this more apparent than at radio frequencies below 1.4 GHz.
+The study of the history of hydrogen gas throughout the universe's evolution is pushing technology development towards arrays of low-cost antennas with large fields of view and densely packed layouts.
+Similarly, the search for transient objects and regular monitoring of the time-dependent sky is driving instruments in the same direction with the added requirement of fast read-outs.
+A number of new telescopes are being or were developed around the world based on this new paradigm, including the Hydrogen Epoch of Reionization Array\footnote{http://reionization.org} (HERA; \citealt{deb16}), the Murchison Widefield Array (MWA; \citealt{tin13,bow13}), the Precision Array for Probing the Epoch of Reionization (PAPER; \citealt{par10}), the LOw Frequency ARray (LOFAR; \citealt{van13}), the Canadian Hydrogen Intensity Mapping Experiment (CHIME; \citealt{ban14}), the Long Wavelength Array (LWA; \citealt{ell13}), and the low frequency Square Kilometer Array (SKA1-Low; \citealt{mel13}).
 
@@
 $\Ngrid \log_2 \Ngrid$ computational scaling without placing any restriction on
-antenna placement, can accommodate non-identical beam patterns, and is provably 
+antenna placement, can accommodate non-identical antennas, and is provably 
 optimal \citep{mor11}. This algorithm uses the antenna beam patterns to 
@@
 We provide a brief summary of the mathematical equivalence of the MOFF and 
-FX correlators detailed in \citet{mor11}. We first relate the dirty image produced 
+FX correlators detailed in \citet{mor11}. We first relate the image produced 
 from visibilities to the electric fields of astrophysical sources, then show that 
@@
 \nonumber\\
-              &\qquad\qquad\qquad\qquad\qquad\qquad\qquad\quad + \left.\left. n^\star_b(f,t) \vphantom{\int}\right] \right\rangle_t \\
+              &\qquad\qquad\qquad\qquad\qquad\qquad\qquad\quad + \left.\left. n^\star_b(f,t) \vphantom{\int}\right] \right\rangle_t \nonumber\\
               &= \iint \mathcal{W}_a(\hat{\mathbf{s}},f,t) \mathcal{W}^\star_b(\hat{\mathbf{s'}},f,t) \left\langle \mathcal{E}(\hat{\mathbf{s}},f,t) \mathcal{E}^\star(\hat{\mathbf{s'}},f,t) \right\rangle_t \nonumber\\
@@
 where we have brought the time average into the integral under the assumption that the aperture illumination pattern does not change over the time-scale of the averaging and $\star$ denotes complex conjugation.
-{\bf The noise-like cross-terms ideally have zero mean}.
+The noise-like cross-terms ideally have zero mean.
 This expression can be further simplified with the sky brightness, $\mathcal{I}(\hat{\mathbf{s}},f,t)= \left\langle \mathcal{E}(\hat{\mathbf{s}},f,t)\,\mathcal{E}^\star(\hat{\mathbf{s'}},f,t) \right\rangle_t \delta(\hat{\mathbf{s}}-\hat{\mathbf{s'}})$, and defining the antenna-pair sky power response function (or the directional antenna-pair power pattern), $\mathcal{W}_{ab}(\hat{\mathbf{s}})~\equiv~\mathcal{W}_a(\hat{\mathbf{s}},f,t)~\mathcal{W}^\star_b(\hat{\mathbf{s}},f,t)$.
-The result is the visibility expressed in terms of the sky brightness, the power response, and uncorrelated noise terms which we group into $n_{ab}(f,t)$.
+The result is the visibility expressed in terms of the sky brightness, the power pattern, and uncorrelated noise terms which we group into $n_{ab}(f,t)$.
 \begin{align}\label{eqn:measurement-eqn-1}
@@
 \begin{align}\label{eqn:software-holography}
-  V_{ab}(f,t) &= \int W_{ab}(\mathbf{r}_{ab}^\prime-\mathbf{r},f,t) \nonumber\\ 
+  V_{ab}(f,t) &= \int W_{ab}(\mathbf{r}_{ab}-\mathbf{r},f,t) \nonumber\\ 
               &\qquad \times \left[\int I(\hat{\mathbf{s}},f,t)\,e^{-2\pi i f\mathbf{r}.\hat{\mathbf{s}}/c}\,\dif\Omega\right]\dif^2\mathbf{r}\,+\, n_{ab}(f,t),
@@
 
-We use the EPIC simulator to generate stochastic electric field samples from a 
-sky model consisting of 10 point sources of flux densities 10~Jy each at random 
-locations. In our simulations, we use 64 frequency channels each of width 
-$\Delta f = 40$~kHz. The number of time stamps integrated in one integration 
-cycle was kept at eight where each A/D timeseries is $1/\Delta f=25\,\mu$s long. 
-We use the MWA array layout \citep{bea12} for demonstration. Only the inner 51 
-tiles within a square bounding box of 150~m on each side were used. We assumed 
-all tiles are identical and have a square shaped electric field illumination 
-footprint 4.4~m on each side. Besides the stochastic sky noise present in the 
-simulated electric fields, no noise from the instrument is added.
+We use the EPIC simulator to generate stochastic electric field samples from a sky model consisting of 10 point sources of random flux densities $>10$~Jy each at random locations.
+In our simulations, we use 64 frequency channels each of width $\Delta f = 40$~kHz.
+The number of time stamps integrated in one integration cycle was kept at eight where each A/D timeseries is $1/\Delta f=25\,\mu$s long.
+We use the MWA array layout \citep{bea12} for demonstration.
+Only the inner 51 tiles within a square bounding box of 150~m on each side were used.
+We assumed all tiles are identical and have a square shaped electric field illumination footprint 4.4~m on each side.
+Besides the stochastic sky noise present in the simulated electric fields, no noise from the instrument is added.
 
@@
 
-Before the outputs can be compared, we describe the elimination of a minor 
-mathematical difference between the two techniques. The squaring operation under 
-MOFF imaging in the image plane introduces antenna auto-correlations around the 
-zero spacing in the $uv$-plane which are absent in traditional visibility-based 
-imaging. In order to facilitate a robust comparison between MOFF and 
-visibility-based imaging techniques, these auto-correlations are removed from 
-the MOFF algorithm output, which is otherwise not an essential part of the core 
-algorithm. We describe below how they are removed.
+Before the outputs can be compared, we describe the elimination of a distinct difference between the two techniques.
+The squaring operation under MOFF imaging in the image plane introduces antenna auto-correlations around the zero spacing in the $uv$-plane which are absent in traditional visibility-based imaging.
+In order to facilitate a robust comparison between MOFF and visibility-based imaging techniques, these auto-correlations are removed from the MOFF algorithm output, which is otherwise not an essential part of the core algorithm.
+We describe below how they are removed.
 
@@
 
-We inverse Fourier transform the squared images and beams back to the $uv$ 
-plane and subtract the estimated auto-correlation kernel scaled to the peak 
-value centred at the zero spacing pixel. The final averaged image is obtained 
-by Fourier transforming the $uv$ plane data and weights with the 
-auto-correlations subtracted to the image plane. These images are now comparable 
-to those obtained from visibility-based imaging. This step of removing 
-auto-correlations is optional and required to be performed only once per 
-integration time-scale and does not add significant cost to the full operation.
+We inverse Fourier transform the squared images and beams back to the $uv$ plane and subtract the estimated auto-correlation kernel scaled to the peak value centred at the zero spacing pixel.
+The final averaged image is obtained by Fourier transforming the $uv$ plane data and weights with the auto-correlations subtracted to the image plane.
+These images are now comparable to those obtained from visibility-based imaging.
+This step of removing auto-correlations needs to be performed only once per integration time-scale and does not add significant cost to the full operation.
 
@@
 
-Fig.~\ref{fig:MOFF-FX-image} shows the dirty images (top) and synthesized beams
-(bottom) obtained with antenna-based MOFF and FX visibility-based imaging 
-algorithms packaged in EPIC. The antenna auto-correlations that correspond to 
-zero spacing have been removed from the correlated weights and data in the 
-$uv$ plane, the MOFF image and the corresponding synthesized beam as 
-described in \S\ref{sec:rm-autocorr}. The sky positions of the simulated sources 
-are indicated by solid black circles. The reconstructed sky image has the 
-simulated sources at the expected sky positions in either case. Both algorithms 
-result in images and synthesized beams that are well matched with each other. 
-Their fluxes are modulated by a multiplicative power pattern corresponding to 
-that of a uniform square aperture.
+Fig.~\ref{fig:MOFF-FX-image} shows the weighted dirty images (top) and synthesized beams (bottom) obtained with antenna-based MOFF and FX visibility-based imaging algorithms packaged in EPIC.
+The antenna auto-correlations that correspond to zero spacing have been removed from the correlated weights and data in the $uv$ plane, the MOFF image and the corresponding synthesized beam as described in \S\ref{sec:rm-autocorr}.
+The reconstructed sky image has the simulated sources at the expected sky positions in either case.
+Both algorithms result in images and synthesized beams that are well matched with each other.
+As expected, their fluxes are attenuated by a weighting proportional to the square of the antenna power pattern corresponding to a uniform square aperture.
 
@@
   % \includegraphics[width=\columnwidth]{MOFF_FX_comparison_10_random_source_positions_4_iterations_test_aperture_zoomed.eps}
-  \caption{Dirty images (top) and synthesized beams (bottom) obtained from 
-    simulated data using EPIC implementation of antenna-based MOFF algorithm 
-    (left) and visibility-based imaging (right). The solid black circles in the
-    top panels indicate the simulated source positions. The antenna 
-    auto-correlations at zero-spacing have been removed from the MOFF images. 
-    The images in either case reconstruct the sources at the right locations 
-    with the fluxes expected after multiplication by the antenna power pattern. 
+  \caption{Weighted dirty images (top) and synthesized beams (bottom) obtained from simulated data using EPIC implementation of antenna-based MOFF algorithm (left) and visibility-based imaging (right).
+The antenna auto-correlations at zero-spacing have been removed from the MOFF images.
+The images in either case reconstruct the sources at the right locations with the fluxes attenuated expected after multiplication by the antenna power pattern.
     The synthesized beams from the two algorithms are well matched in size and
@@
 
+We examine in detail the respective synthesized beams in each case in Fig.~\ref{fig:MOFF-FX-psf-diff}.
+Slices at $m=0$ of the synthesized beams weighted by the antenna power pattern are shown for MOFF method using EPIC (solid black) and visibility-based (dashed gray) imaging.
+The two are found to match well.
+A magnified view shows that some differences at the level of $<0.5$\% exist in some regions.
+These are attributed to differences in the $uv$-plane antenna cross-correlation weights in the two methods which in turn arises due to the amount of coarseness in grid spacing as described below.
+
 \begin{figure}
   \includegraphics[width=\columnwidth]{figure6}
-  \caption{.}
-  \label{fig:f-engine}
+  \caption{Synthesized beams weighted by antenna power pattern sliced at $m=0$ obtained with MOFF algorithm using EPIC (solid black line) and visibility-based imaging (dashed gray line).
+The two appear almost identical.
+The inset provides a magnified view of differences in the synthesized beam slices between the two techniques at levels $\lesssim 0.5$\% relative to the peak.
+These are attributed to differences that arise in gridding and depends on the coarseness of the grid.}
+  \label{fig:MOFF-FX-psf-diff}
 \end{figure}
 
-Now we investigate the gridded cross-correlation weights in the $uv$-plane. 
-Fig.~\ref{fig:MOFF-FX-uvwts} shows the correlated weights obtained with MOFF 
-imaging (left) and visibility-based imaging (right). There are two notable 
-differences. The first is in the weights around zero spacing. Though both show a
-dominant void around zero-spacing, the void obtained with MOFF algorithm shows 
-many pixels with non-zero weights. In contrast, the zero-spacing void from 
-traditional imaging consists of predominantly zero-valued pixels. The second 
-difference is in the weights outside the zero-spacing void. In some regions they 
-are different from each other at the few percent level though the sum of weights 
-in these regions is the same. For instance, note the difference in weights at 
-$(u,v)\approx (2.5,4)$. The reason for these differences is described below.
-
-The gridding step in MOFF imaging samples the antenna footprint (either in 
-analytic or lookup table formats) at the grid locations. Coverage of grid pixels 
-by an antenna footprint may be $\sim 1$ pixel narrower particularly at the 
-edge of the footprint along one or both directions relative to that from another 
-identical antenna but with a fractionally different location relative to the 
-grid. This depends on the exact location of the centre of the antenna relative 
-to the grid and the coarseness of grid spacing. This first order loss of 
-precision of the sampled footprint propagates to second order ($\sim 2$ pixels) 
-upon correlation of the discretized weights. In other words, the correlated 
-weights may suffer further loss of precision in their sampled footprint after 
-correlation of two footprints each of which could be less precise to first 
-order. On the other hand, in visibility-based imaging, a directly sampled $uv$ 
-plane antenna power response (which is theoretically identical to the correlation 
-of individual antenna footprints) centred on a baseline has a loss in precision 
-at most to first order. Thus, although in the limit of infinitesimally small 
-grid spacing they should be identical, the coarseness of grid spacing introduces 
-subtle differences between the two.
-
-These differences which are dependent on the coarseness of grid spacing can be 
-mitigated by making the grid spacing finer at the expense of increased 
-computational cost. Residuals centred around zero spacing can also be lowered 
-by subtracting each auto-correlation of antenna weights separately by using the 
-shape and extent of the sampled footprint appropriate for that specific antenna 
-aperture. This is a general solution applicable even in case of heterogeneous 
-antenna arrays and is under active development for EPIC.
+The left panel of Fig.~\ref{fig:MOFF-FX-uvwts-diff} shows differences (in percentage relative to peak) in cross-correlated weights obtained with MOFF imaging in EPIC and visibility-based imaging.
+The maximum difference appears near the center of the $uv$-plane corresponding to antenna auto-correlations which have not been perfectly removed in the former.
+In other regions of the $uv$-plane, the differences are of the order of a few percent.
+The reason for these differences is described below.
 
@@
   % \includegraphics[width=\columnwidth]{MOFF_FX_comparison_uvwts_test_aperture_zero_spacing_removed.eps}
-  \caption{.}
-  \label{fig:MOFF-FX-uvwts}
+  \caption{{\it Left}: Differences (in \%) in cross-correlated weights in the $uv$ plane relative to the peak.
+The biggest difference ($\sim 20$\%) is found around the zero-spacing corresponding to auto-correlations of antenna weights due to gridding differences augmented by the relatively high number density of antenna auto-correlation footprints in that region.
+In most of the other regions, differences of the order of less than a few percent are seen.
+{\it Right}: Histogram (expressed as percentage) of the relative differences (in \%) between antenna cross-correlation weights shown on the left binned in intervals of 0.5\%.
+Over 70\% of the $uv$-cells do not differ by more than 0.5\% and over 90\% of $uv$-cells only differ by $<5$\% for the gridding coarseness used.}
+  \label{fig:MOFF-FX-uvwts-diff}
 \end{figure}
 
+The gridding step in MOFF imaging samples the antenna footprint (either in analytic or lookup table formats) at the grid locations.
+Coverage of grid pixels by an antenna footprint may be $\sim 1$ pixel narrower particularly at the edge of the footprint along one or both directions relative to that from another identical antenna but with a fractionally different location relative to the grid.
+This depends on the exact location of the centre of the antenna relative to the grid and the coarseness of grid spacing.
+This first order loss of precision of the sampled footprint propagates to second order ($\sim 2$ pixels) upon correlation of the discretized weights.
+In other words, the correlated weights may suffer further loss of precision in their sampled footprint after correlation of two footprints each of which could be less precise to first order.
+On the other hand, in visibility-based imaging, a directly sampled $uv$ plane antenna power response (which is theoretically identical to the correlation of individual antenna footprints) centred on a baseline has a loss in precision at most to first order.
+Thus, although in the limit of infinitesimally small grid spacing they should be identical, the coarseness of grid spacing introduces subtle differences between the two.
+
+These differences which are dependent on the coarseness of grid spacing can be mitigated by making the grid spacing finer at the expense of increased computational cost.
+Residuals centred around zero spacing can also be lowered by subtracting each auto-correlation of antenna weights separately by using the shape and extent of the sampled footprint appropriate for that specific antenna aperture.
+This is a general solution applicable even in case of heterogeneous antenna arrays and is under active development for EPIC.
+
 We study the effect of the differences in gridded weights on the image plane.
-Fig.~\ref{fig:psf-diff} shows the difference between the synthesized beams 
-obtained with the two methods. A difference map between the two synthesized 
-beams is shown on top. The amplitude of the difference appears to be 
-modulated by the directional power response of the antenna. At the bottom, 
-in radial bins, the rms of the synthesized beam (gray) and the rms of the 
-difference map (black) are plotted in percentage units relative to the peak 
-(to be read using the axis on the left side of the plot). The antenna power 
-pattern (red; to be read using the scale on the right) is plotted for reference.
+Fig.~\ref{fig:image-psf-diff} shows the difference between the synthesized beams obtained with the two methods.
+A difference map between the two synthesized beams is shown on the left.
+The amplitude of the difference appears to be modulated by the directional power response of the antenna.
+On the right, in radial bins, the rms of the synthesized beam (gray) and the rms of the difference map (black) are plotted in percentage units relative to the peak (to be read using the axis on the left side of the plot).
+The antenna power pattern (red; to be read using the scale on the right) is plotted for reference.
 
@@
     modulated by the power pattern of the antenna.}
-  \label{fig:psf-diff}
+  \label{fig:image-psf-diff}
 \end{figure}
@@
 implies that the beams are well matched in the central regions. In the outer 
-regions, their mismatch is comparable to the rms of synthesized beams. This 
-indicates the two synthesized beams are not completely randomly different from 
+regions, their mismatch is comparable to the rms of synthesized beams. These findings are consistent with Fig.~\ref{fig:MOFF-FX-psf-diff}. This 
+indicates the two synthesized beams are not randomly different from 
 each other in which case the rms of the difference would have been 
 $\approx \sqrt{2}$ higher than the rms of the each of the synthesized beams. 
-This indicates that while differences exist, large fractions of them are still 
+This means that while differences exist, large fractions of them are still 
 well matched to each other even out to the horizon. Thus the rounding errors in 
@@
 
-We corrected the cable delays, but otherwise assume the data is sufficiently 
-calibrated to image directly. A test of EPICal on this data will be conducted 
-in the future.
+We corrected the cable delays, but otherwise assume the data is sufficiently calibrated to image directly. A detailed demonstration of EPICal on this data is presented in \citet{bea16}.
 
@@
 We present a methodology to understand the effective angular weighting in the image obtained with EPIC from a heterogeneous array and compare it with that from standard interferometry that assumes a homogeneous array (all antennas are identical).
-We also demonstrate how the assumption in the latter results in a mis-estimation of recovered flux densities of sources.
+We also demonstrate how the assumption in the latter results in a mis-estimation of the recovered sky model.
 
@@
 \end{align}
-where, superscript `$\textrm{I}$' in the $\mathcal{W}$ term indicates the weighting is inherently introduced by the instrument during the measurement process. The dirty beam-weighted image is:
+where, superscript `$\textrm{I}$' in the $\mathcal{W}$ term indicates the weighting is inherently introduced by the instrument during the measurement process. The beam-weighted dirty image is:
 \begin{align}
@@
 
-While imaging, EPIC introduces a weighting to each antenna during gridding as given by equations~\ref{eqn:dirty-image-FX} and \ref{eqn:e-field-conv}, and equivalently, the weighted visibility is:
+While imaging, EPIC introduces a weighting to each antenna during gridding as given by equations~\ref{eqn:dirty-image-FX} and \ref{eqn:e-field-conv}, and equivalently, the weighted visibility from antenna pair $ab$ projected on the grid is:
 \begin{align}
@@
 \begin{align}
-  B^\textrm{G}_{ab}(\mathbf{r}-\mathbf{r}_{ab},f) &= \int W^\textrm{G}_a(\mathbf{r}^\prime-\mathbf{r}_a)\,W^{\textrm{G}\star}_b(\mathbf{r}^\prime+\mathbf{r}-\mathbf{r}_a)\,\dif^2\mathbf{r}^\prime,
-\end{align}
-and the superscript `$\textrm{G}$' denotes the weighting introduced in analysis during the gridding process.
+  B^\textrm{G}_{ab}(\mathbf{r}-\mathbf{r}_{ab},f) &= \int W^\textrm{G}_a(\mathbf{r}^\prime-\mathbf{r}_a)\,W^{\textrm{G}\star}_b(\mathbf{r}^\prime+\mathbf{r}-\mathbf{r}_b)\,\dif^2\mathbf{r}^\prime.
+\end{align}
+The superscript `$\textrm{G}$' denotes the weighting introduced in analysis during the gridding process.
 Optimal imaging requires $W^\textrm{G}_a = W^\textrm{I}_a$ \citep{mor09,mor11}.
-However, we keep the two superscripts to describe output images under the wrong assumption that the array is homogeneous when instead it is actually heterogeneous.
-
-Thus the sky response of the weighted visibility is:
+However, we keep the two superscripts separately to describe output images under the wrong assumption that the array is homogeneous when instead it is actually heterogeneous.
+
+The sky response of the weighted visibility is:
 \begin{align}
@@
                                               &\qquad\qquad \times e^{2\pi i f\mathbf{r}\cdot\,\hat{\mathbf{s}}/c}\,\dif^2\mathbf{r} \nonumber\\
-                                              &= \mathcal{W}^{\textrm{G}\star}_a(\hat{\mathbf{s}},f)\mathcal{W}^\textrm{G}_b(\hat{\mathbf{s}},f)\,V_{ab}\,e^{2\pi i f\mathbf{r}_{ab}\!\cdot\,\hat{\mathbf{s}}/c}
-\end{align}
-and denotes the fringe from the weighted visibility. The image output of EPIC is equivalent to summing these weighted fringes from all baselines:
+                                              &= \mathcal{W}^{\textrm{G}\star}_a(\hat{\mathbf{s}},f)\mathcal{W}^\textrm{G}_b(\hat{\mathbf{s}},f)\,V_{ab}\,e^{2\pi i f\mathbf{r}_{ab}\!\cdot\,\hat{\mathbf{s}}/c},
+\end{align}
+which denotes the fringe from the weighted visibility. The image output of EPIC is equivalent to averaging these weighted fringes from all antenna pairs:
 \begin{align}\label{eqn:wt-dirty-image-EPIC}
@@
 And under each antenna type $i$, there are $n_i$ antennas such that $\sum_i n_i = \Nant$.
-The total number of unique non-zero spacing antenna pairs is $\Nant(\Nant-1)/2$ and there are potentially $N_\textrm{T}(N_\textrm{T}+1)/2$ unique types of antenna pairs obtained by pairwise combination of $N_\textrm{T}$ antenna types.
+The total number of unique non-zero spacing antenna pairs is $\Nant(\Nant-1)/2$ and there are potentially $N_\textrm{T}(N_\textrm{T}+1)/2$ unique antenna pair types obtained by pairwise combination of the $N_\textrm{T}$ antenna types.
 Thus,
@@
 
-Since equation~\ref{eqn:wt-dirty-image-EPIC} is obtained by summing weighted fringes over all unique antenna pairs, it can be expressed as: 
+Since equation~\ref{eqn:wt-dirty-image-EPIC} is obtained by averaging weighted fringes over all unique antenna pairs, it can be expressed as: 
 \begin{align}\label{eqn:wt-dirty-image-EPIC-decomp}
@@
 
-For purposes of illustration, we consider a simple scenario in which there are two antenna types each containing roughly equal numbers of antennas with similar layouts but the analysis is performed under the wrong assumption that all antennas are identical and are of the first type.
+For purposes of illustration, we consider a simple scenario in which there are two antenna types each containing roughly equal numbers of antennas with similar layouts but the analysis is performed under the wrong assumption that all antennas are identical and are all of the first type.
 In an optimally weighted image, the optimal effective weighting given by equation~\ref{eqn:effective-weighting} will be:
@@
 The net uncertainty was determined by adding the levels of sidelobe fluctuations and intrinsic source fluctuations in quadrature.
-The normalized estimates of the sky model, $\widehat{\mathcal{I}}_\textrm{M}^\textrm{norm}$, and the corresponding ``$1\sigma$'' error bars were obtained by dividing $\widehat{\mathcal{I}}_\textrm{D}^\textrm{iso}$ and the net uncertainty by the respective effective weights in either of the cases considered.
+The normalized estimates of the sky model, $\widehat{\mathcal{I}}_\textrm{M}^\textrm{norm}$, and the corresponding ``$1\sigma$'' error bars were obtained by dividing $\widehat{\mathcal{I}}_\textrm{D}^\textrm{iso}$ and the net uncertainty by the respective effective weights in either case.
 
@@
 The solid and dashed lines correspond to point sources in the south-westward and westward directions respectively.
-It can be clearly seen that in the erroneous case, the estimated values appear to be underestimated and deviate from the ideal value of unity with very high significance ($\gtrsim 5\sigma$) in most parts of the image.
-In complete contrast, the estimates in the optimal case in almost all parts of the image are consistent with unity to within $1\sigma$.
+It can be clearly seen that the estimates in the optimal case in almost all parts of the image are consistent with unity to within $1\sigma$.
+In complete contrast, in the erroneous case, the estimated values appear to be underestimated and deviate from the ideal value of unity with very high significance ($\gtrsim 5\sigma$) in most parts of the image.
 The reason for underestimation is because nearly half the antennas are assumed to be smaller than they actually are and thus results in dividing by overestimated $\mathcal{W}_\textrm{eff}^\textrm{err}$.
 These trends do not seem to be affected by the placement of sources along westward and south-westward directions.
-Thus, we have demonstrated that images from EPIC can robustly image data from heterogeneous arrays while being mathematically optimal.
-Besides being sub-optimal, erroneous assumptions about the homogeneity of an array can lead to significant mis-estimates (in this case, an underestimate) of the sky model.
+
+This demonstrates that images from EPIC can robustly image data from heterogeneous arrays while being mathematically optimal.
+Besides being sub-optimal, erroneous assumptions about the homogeneity of an array can lead to significant and systematic mis-estimates (in this case, an underestimate) of the sky model.
 
@@
 stages even more expensive, the gridding convolution, squaring and time-averaging 
-take up only $\lesssim 20$, $\lesssim 20$ and $\lesssim 5$ per cent respectively 
+take up only $\lesssim 20$\%, $\lesssim 20$\% and $\lesssim 5$\% respectively 
 of the total processing time while the spatial Fourier transform takes up 
@@
 
-In the following discussions, we will assume that the computational cost for
-the MOFF imaging is determined by the spatial Fourier transform while  
-that for visibility-based imaging comes from the cross-correlations. However, 
-if non-linearities such as non-coplanarity of baselines \citep{cor08} and 
-wide-field phenomena like the {\it pitchfork} effect \citep{thy15a,thy15b} 
-are to be corrected for, the antenna-based illumination footprint can start 
-becoming less compact in the measurement plane and can result in a costlier 
-gridding process.
+In the following discussions, we will assume that the computational cost for the MOFF imaging is determined by the spatial Fourier transform while that for visibility-based imaging comes from the cross-correlations.
+However, if non-linearities such as non-coplanarity of baselines \citep{cor08} and wide-field phenomena like the {\it pitchfork} effect \citep{thy15a,thy15b} are to be corrected for, the antenna illumination footprint will become less compact in the measurement plane and can result in a costlier gridding process.
 
@@
 
-We now consider antenna array layouts described by three essential quantities 
-in radio interferometry, namely, maximum baseline length, number of antennas,
-and the size of each antenna.
-
-Fig.~\ref{fig:parameter-space-bll-nant-instruments} shows the boundaries where 
-the ratio of the number of computations required with visibility-based imaging 
-relative to MOFF imaging is unity. The different colored lines correspond to 
-different antenna sizes (cyan - 1~m$^2$, blue - 7~m$^2$, purple - 16~m$^2$, 
-green - 28~m$^2$, orange - 150~m$^2$, red - 740~m$^2$, gray - 5900~m$^2$). Solid 
-line style in each color denotes the maximum number of antennas with the 
-corresponding antenna size that can be densely packed inside various baseline 
-lengths. The region to the right of the solid lines for corresponding antenna 
-size represents a scenario of overlapping antennas that is physically 
-impossible. Dashed lines of each color denote the boundary to the left of which 
-visibility-based imaging is favoured for the corresponding antenna size. Region 
-inside the wedge enclosed by the solid and dashed lines favours using the MOFF 
-algorithm for the corresponding antenna size over visibility-based imaging.
-
-As antenna size increases the maximum number of antennas for a dense packing as 
-a function of baseline length decreases. Hence, the solid lines shift leftward 
-as antenna size increases. Similarly, with increase in antenna size, $\Ngrid$ 
-also decreases when field of view imaging is achieved with an increasing grid 
-spacing equal to antenna size and hence lowers the amount of computations 
-required with the MOFF algorithm. This shifts the dashed curves leftward.
-
-The different antennas are color coded by roughly the class of antenna size 
-they fall into. Thus symbols of one color falling into a wedge of the same color 
-indicates MOFF imaging is favoured for those telescopes and vice versa. For 
-e.g., MOFF imaging is favoured in HERA-331 and HERA-6769 because they are inside 
-the wedge but not so in cases of HERA-19 and HERA-37. A majority of the 
-next-generation radio telescopes, namely, HERA-331 and its future expanded 
-versions, SKA1-LC, SKA1-LCD, HIRAX, and CHIME will fall in the regime where MOFF 
-imaging will be desirable. LWA1 and LWA-OV are already very close to the dividing 
-line. Their hypothetical expansions,\footnote{LWA1-x2x1 and LWA1-x4x1.5, and 
-LWA-OVx2x1 and LWA-OVx4x1.5 denote two-fold and four-fold increase in number 
-of antennas within a core diameter that is 1 and 1.5 times the current size of 
-100~m and 200~m respectively for LWA1 and LWA-OV.} will be in the computational 
-regime favoured by MOFF. For a fixed baseline length, regions favouring the MOFF 
-algorithm tend to be towards large $\Nant$ indicating large-N dense array 
-layouts with smaller antenna elements are best suited for deploying EPIC.
+We now consider antenna array layouts described by three essential quantities in radio interferometry, namely, maximum baseline length, number of antennas, and the size of each antenna.
+Fig.~\ref{fig:parameter-space-bll-nant-instruments} shows the boundaries where the ratio of the number of computations required with visibility-based imaging relative to MOFF imaging is unity.
+The different colored lines correspond to different antenna sizes (cyan - 1~m$^2$, blue - 7~m$^2$, purple - 16~m$^2$, green - 28~m$^2$, orange - 150~m$^2$, red - 740~m$^2$, gray - 5900~m$^2$).
+Dashed lines of each color denote the boundary to the left of which the MOFF algorithm is favoured for the corresponding antenna size and vice versa for visibility-based imaging.
 
@@
   % \includegraphics[width=\columnwidth]{MOFF_FX_crossover_baseline_n-antennas_rho_fov_gridding_legended.eps}
-  \caption{Current and planned instruments in parameter space of
-    baseline length and number of antennas with MOFF and FX. Line styles of 
-    different colors denote different classes of antenna sizes (cyan - 1~m$^2$, 
-    blue - 7~m$^2$, purple - 16~m$^2$, green - 28~m$^2$, orange - 150~m$^2$, 
-    red - 740~m$^2$, gray - 5900~m$^2$). Solid line style in each color denotes 
-    the maximum number of antennas with the corresponding antenna size that can 
-    be packed inside various baseline lengths. The region to the right of the 
-    solid lines for corresponding antenna size is physically disallowed. Dashed
-    lines of each color denote the boundary to the left of which 
-    visibility-based imaging is favoured for the corresponding antenna size. 
-    Region inside the wedge enclosed by the solid and dashed lines favours using 
-    the MOFF algorithm for the corresponding antenna size. These wedges shift 
-    leftward with increasing antenna size. The different antennas are color 
-    coded by roughly the class of antenna size they fall into. Thus symbols of 
-    one color falling into a wedge of the same color indicates MOFF imaging is 
-    advantageous for those telescopes and vice versa. For example, MOFF imaging
-    is favoured in HERA-331 and HERA-6769 because they are inside the wedge 
-    but not so in cases of HERA-19 and HERA-37.}
+  \caption{Current and planned instruments in parameter space of baseline length and number of antennas with MOFF and FX.
+Lines of different colors denote different classes of antenna sizes (cyan - 1~m$^2$, blue - 7~m$^2$, purple - 16~m$^2$, green - 28~m$^2$, orange - 150~m$^2$, red - 740~m$^2$, gray - 5900~m$^2$).
+Lines of each color denote the boundary to the left of which MOFF algorithm is favoured for the corresponding antenna size.
+These lines shift rightward with increasing antenna size.
+The different antennas are color coded by roughly the class of antenna size they fall into.
+Thus symbols of one color falling to the left of a line of the same color indicate MOFF imaging is advantageous for those telescopes and vice versa.
+For example, MOFF imaging is favoured in HERA-331 and HERA-6769 because they lie to the left of the orange line but not so in cases of HERA-19 and HERA-37.}
   \label{fig:parameter-space-bll-nant-instruments}
 \end{figure}
+
+There is an upper limit set by the maximum number of antennas for each antenna size that can be densely packed inside various baseline lengths but is not shown on this figure to avoid crowding.
+We note that as antenna size increases the maximum number of antennas for a dense packing as a function of baseline length decreases.
+Hence, this upper limit shifts rightward as antenna size increases.
+Similarly, with increase in antenna size, $\Ngrid$ also decreases when field of view imaging is achieved with an increased grid spacing equal to antenna size and hence lowers the amount of computations required with the MOFF algorithm.
+This shifts the dashed lines rightward as well.
+
+The different antennas are color coded by roughly the class of antenna size they fall into.
+Thus symbols of one color that lie to the left of same colored line indicate MOFF imaging is favoured for those telescopes and vice versa.
+For e.g., MOFF imaging is favoured in HERA-331 and HERA-6769 because they lie to the left of the orange line but not so in cases of HERA-19 and HERA-37.
+A majority of the next-generation radio telescopes, namely, HERA-331 and its future expanded versions, SKA1-LC, SKA1-LCD, HIRAX, and CHIME will fall in the regime where MOFF imaging will be desirable.
+LWA1 and LWA-OV are already very close to the dividing line.
+Their hypothetical expansions,\footnote{LWA1-x2x1 and LWA1-x4x1.5, and LWA-OVx2x1 and LWA-OVx4x1.5 denote two-fold and four-fold increase in number of antennas within a core diameter that is 1 and 1.5 times the current size of 100~m and 200~m respectively for LWA1 and LWA-OV.} will be in the computational regime favoured by MOFF.
+For a fixed baseline length, regions favouring the MOFF algorithm tend to be towards large $\Nant$ indicating large-N dense array layouts with smaller antenna elements are best suited for deploying EPIC.
 
@@
 Under the class of direct imaging techniques, ours is one of the most generic -- neither does it place any constraint on the array layout to be on a regular grid nor does it require the antenna array to be homogeneous.
-We have demonstrated its natural ability to robustly image data from heterogeneous arrays and shown that wrong assumptions about array homogeneity result in severe mis-estimation of the sky model.
+We have demonstrated its natural ability to robustly image data from heterogeneous arrays and shown that wrong assumptions about array homogeneity result in severe and systematic mis-estimation of the sky model.
 
@@
 successfully tested on simulated MWA observations as well as real LWA 
-observations.
+observations from both imaging and calibration view points.
 
@@
 
-In the near future, we plan to upgrade our current Python implementation 
-of EPIC to a GPU-based pipeline in order to operate on real-time data and 
-develop a transient trigger and monitor backend. In the meanwhile, we plan 
-to demonstrate the capability of EPIC to calibrate and image from heterogeneous 
-arrays and incorporate corrections for non-coplanarity of baselines and 
-direction-dependence of calibration.
+In the near future, we plan to demonstrate its speed and precision by upgrading EPIC to a GPU-based implementation in order to operate on real-time data and develop a transient trigger and monitor backend.
+In the meanwhile, we plan to demonstrate imaging with non-coplanar arrays and direction-dependent calibration.
 
@@
 
-The antenna module is a fundamental building block upon which all the other 
-modules are built. There is one antenna module per antenna each having 
-attributes -- the propagated electric field time-series, $E(t)$, and spectrum 
-$\widetilde{E}(f)$ for both polarizations. The most important function inside 
-this module is the F-engine that Fourier transforms time-series electric field 
-data into spectra.
+The antenna module is a fundamental building block upon which all the other modules are built.
+There is one antenna module per antenna each having attributes -- the propagated electric field time-series, $\widetilde{E}(t)$, and spectrum $E(f)$ for both polarizations.
+The most important function inside this module is the F-engine that Fourier transforms time-series electric field data into spectra.
 
@@
 
-The interferometer module holds the attributes and functions pertaining to
-a pair of antennas and represents the cross-correlation information obtained
-from the pair. Its primary attributes are the two antenna modules. It also
-contains four cross-polarized visibility time-series (even for the FX 
-correlator for diagnostic purposes) and spectra.
-
-The critical component of the interferometer module is the X-engine. This is 
-essentially a software analog of hardware correlators of real telescope 
-systems. The X-engine can be toggled between two states of operation, 
-namely, the FX and XF modes. The FX mode obtains the electric field spectra, 
-$\widetilde{E}(f)$ from the individual antenna modules inside this module and 
-multiplies the two to obtain visibility spectra, $\widetilde{V}(f)$. On the 
-other hand, the XF mode cross-correlates the electric field time-series from 
-its Antenna modules to obtain the visibilities as a function of lags, 
-$\widetilde{V}_t(t)$, which is then Fourier transformed to obtain 
-$\widetilde{V}(f)$. Both modules can operate on dual-polarizations to obtain 
-all four cross-polarizations.
+The interferometer module holds the attributes and functions pertaining to a pair of antennas and represents the cross-correlation information obtained from the pair.
+Its primary attributes are the two antenna modules.
+It also contains four cross-polarized visibility time-series (even for the FX correlator for diagnostic purposes) and spectra.
+
+The critical component of the interferometer module is the X-engine.
+This is essentially a software analog of hardware correlators of real telescope systems.
+The X-engine can be toggled between two states of operation, namely, the FX and XF modes.
+The FX mode obtains the electric field spectra, $E(f)$ from the individual antenna modules inside this module and multiplies the two to obtain visibility spectra, $V(f)$.
+On the other hand, the XF mode cross-correlates the electric field time-series from its Antenna modules to obtain the visibilities as a function of lags, $V_t(\tau)$, which is then Fourier transformed to obtain $V(f)$.
+Both modules can operate on dual-polarizations to obtain all four cross-polarizations.
 
@@
 cross-polarization for the current data stream. Similar to the antenna module,
-it has an update function that can update the visibilities $\widetilde{V}_t(t)$ 
-or $\widetilde{V}(f)$ directly rather than through the electric fields of its 
+it has an update function that can update the visibilities $V_t(\tau)$ 
+or $V(f)$ directly rather than through the electric fields of its 
 component antennas. This functionality is to allow EPIC to operate while 

### 2016-08-15 imaging_demo/ms.tex
Update notation in block diagram figures
From 65210aa (mnras):
====================

(*) Update notation in block diagram figures

@@
 Not only does it bring down the cost for dense layouts to $\mathcal{O}(\Nant\log_2\Nant)$ but can also image from irregular layouts and heterogeneous arrays of antennas.
-We demonstrate its versatility to image with heterogeneous arrays by showing that it robustly estimates the input sky model while wrong assumptions about array homogeneity lead to severe mis-estimates.
+We demonstrate its versatility to image with heterogeneous arrays by showing that it robustly estimates the input sky model while wrong assumptions about array homogeneity lead to severe, systematic mis-estimates.
 EPIC is highly modular and parallelizable, implemented in object oriented Python, and publicly available.

### 2016-08-19 imaging_demo/ms.tex
Update abstract, Intro with Adam's suggestions
From 0c31757 (mnras):
====================

In ms.tex,

(*) Update abstract and Introduction with Adam's suggestions

--- 2016-08-14 imaging_demo/ms.tex
@@
 Not only does it bring down the cost for dense layouts to $\mathcal{O}(\Nant\log_2\Nant)$ but can also image from irregular layouts and heterogeneous arrays of antennas.
-We demonstrate its versatility to image with heterogeneous arrays by showing that it robustly estimates the input sky model while wrong assumptions about array homogeneity lead to severe mis-estimates.
 EPIC is highly modular and parallelizable, implemented in object oriented Python, and publicly available.
@@
 We have also validated our implementation on data observed with the Long Wavelength Array (LWA).
+We have demonstrated its versatility to image with heterogeneous arrays by showing that it robustly estimates the input sky model while wrong assumptions about array homogeneity lead to systematic mis-estimates.
 Antenna layouts with dense filling factors consisting of a large number of antennas such as LWA, the Square Kilometre Array, Hydrogen Epoch of Reionization Array, and Canadian Hydrogen Intensity Mapping Experiment will gain significant computational advantage by deploying EPIC.
@@
 Similarly, the search for transient objects and regular monitoring of the time-dependent sky is driving instruments in the same direction with the added requirement of fast read-outs.
-A number of new telescopes are being or were developed around the world based on this new paradigm, including the Hydrogen Epoch of Reionization Array\footnote{http://reionization.org} (HERA; \citealt{deb16}), the Murchison Widefield Array (MWA; \citealt{tin13,bow13}), the Precision Array for Probing the Epoch of Reionization (PAPER; \citealt{par10}), the LOw Frequency ARray (LOFAR; \citealt{van13}), the Canadian Hydrogen Intensity Mapping Experiment (CHIME; \citealt{ban14}), the Long Wavelength Array (LWA; \citealt{ell13}), and the low frequency Square Kilometer Array (SKA1-Low; \citealt{mel13}).
+A number of new telescopes are being or were developed around the world based on this new paradigm, including the Hydrogen Epoch of Reionization Array\footnote{http://reionization.org} (HERA; \citealt{deb16}), the Murchison Widefield Array (MWA; \citealt{tin13,bow13}), the Donald~C.~Backer Precision Array for Probing the Epoch of Reionization (PAPER; \citealt{par10}), the LOw Frequency ARray (LOFAR; \citealt{van13}), the Canadian Hydrogen Intensity Mapping Experiment (CHIME; \citealt{ban14}), the Long Wavelength Array (LWA; \citealt{ell13}), and the low frequency Square Kilometer Array (SKA1-Low; \citealt{mel13}).
 

### 2016-08-19 imaging_demo/ms.tex
Update sections 2 and 3 with Adam's suggestions
From 357d600 (Adam-v1):
======================

In ms.tex,

(*) Update sections 2 and 3 with Adam's suggestions

@@
 \begin{align}
-  E(\mathbf{r}, f, t) &= \int E(\hat{\mathbf{s}},f,t)\,e^{-2\pi i f\mathbf{r}\cdot\hat{\mathbf{s}}/c}\,\dif\Omega,
+  E(\mathbf{r}, f, t) &= \int \mathcal{E}(\hat{\mathbf{s}},f,t)\,e^{-2\pi i f\mathbf{r}\cdot\hat{\mathbf{s}}/c}\,\dif\Omega,
 \end{align}
@@
 The noise-like cross-terms ideally have zero mean.
-This expression can be further simplified with the sky brightness, $\mathcal{I}(\hat{\mathbf{s}},f,t)= \left\langle \mathcal{E}(\hat{\mathbf{s}},f,t)\,\mathcal{E}^\star(\hat{\mathbf{s'}},f,t) \right\rangle_t \delta(\hat{\mathbf{s}}-\hat{\mathbf{s'}})$, and defining the antenna-pair sky power response function (or the directional antenna-pair power pattern), $\mathcal{W}_{ab}(\hat{\mathbf{s}})~\equiv~\mathcal{W}_a(\hat{\mathbf{s}},f,t)~\mathcal{W}^\star_b(\hat{\mathbf{s}},f,t)$.
+This expression can be further simplified using the sky brightness, $\mathcal{I}(\hat{\mathbf{s}},f,t)$, as $\left\langle \mathcal{E}(\hat{\mathbf{s}},f,t)\,\mathcal{E}^\star(\hat{\mathbf{s'}},f,t) \right\rangle_t = \mathcal{I}(\hat{\mathbf{s}},f,t)\,\delta(\hat{\mathbf{s}}-\hat{\mathbf{s'}})$, and defining the antenna-pair sky power response function (or the directional antenna-pair power pattern), $\mathcal{W}_{ab}(\hat{\mathbf{s}})~\equiv~\mathcal{W}_a(\hat{\mathbf{s}},f,t)~\mathcal{W}^\star_b(\hat{\mathbf{s}},f,t)$.
 The result is the visibility expressed in terms of the sky brightness, the power pattern, and uncorrelated noise terms which we group into $n_{ab}(f,t)$.
@@
 where, the baseline coordinate $\mathbf{r}_{ab}\equiv\mathbf{r}_a-\mathbf{r}_b$ is the vector separation between the two antennas.
-This signifies that the visibility ($V_{ab}$) measured between a pair of antennas is obtained by the multiplying the sky brightness $\mathcal{I}(\hat{\mathbf{s}},f,t)$ by the antenna power response $\mathcal{W}_{ab}(\hat{\mathbf{s}},f,t)$ and Fourier transforming from the directional coordinates ($\hat{\mathbf{s}}$) to {\bf the measurement plane}, which are then sampled at the locations of the antenna spacings (or baselines), namely, $\mathbf{r}_{ab}$, and added to the noise $n_{ab}$.
+This signifies that the visibility ($V_{ab}$) measured between a pair of antennas is obtained by multiplying the sky brightness $\mathcal{I}(\hat{\mathbf{s}},f,t)$ by the antenna power response $\mathcal{W}_{ab}(\hat{\mathbf{s}},f,t)$ and Fourier transforming from the directional coordinates ($\hat{\mathbf{s}}$) to {\bf the measurement plane}, which are then sampled at the locations of the antenna spacings (or baselines), namely, $\mathbf{r}_{ab}$, and added to the noise $n_{ab}$.
 
@@
 \begin{align}
-  \boldsymbol{\mathcal{I}}^\prime(\hat{\mathbf{s}}) &= \left\langle \left|\,\mathbf{F}^\textrm{T}(\hat{\mathbf{s}},\mathbf{r})\,\mathbf{W}^\textrm{T}(\mathbf{r},\mathbf{a})\,\mathbf{N}^{-1}(\mathbf{a},\mathbf{a})\,\mathbf{E}(\mathbf{a})\,\right|^2\right\rangle_t.
+  \boldsymbol{\mathcal{I}}^\prime(\hat{\mathbf{s}}) &= \left\langle \left|\,\mathbf{F}^\textrm{T}(\hat{\mathbf{s}},\mathbf{r})\,\mathbf{W}^\textrm{T}(\mathbf{r},\mathbf{a})\,\widetilde{\mathbf{N}}(\mathbf{a},\mathbf{a})\,\mathbf{E}(\mathbf{a})\,\right|^2\right\rangle_t.
 \label{eqn:dirty-image-MOFF}
@@
 \begin{enumerate}
-\item The time-averaging cannot be performed on a stochastic measurement but
-  only on its statistical properties. In visibility-based imaging, the 
-  visibilities measured between antenna pairs represent spatial correlations 
-  which can be time-averaged followed by gridding and imaging. However, in MOFF 
-  imaging both antenna and gridded electric fields are stochastic and therefore 
-  must be imaged and squared before time-averaging. 
-\item In visibility-based imaging, electric fields measured by antennas are not 
-  correlated with themselves and hence lack zero spacing measurements. In 
-  contrast, in MOFF imaging, since the gridded electric fields are imaged and 
-  squared, they retain information from auto-correlated electric fields at zero 
-  spacing and thus yield the true total power of the imaged field.
+\item The time-averaging cannot be performed on a stochastic measurement but only on its statistical properties.
+In visibility-based imaging, the visibilities measured between antenna pairs represent spatial correlations which can be time-averaged followed by gridding and imaging.
+However, in MOFF imaging both antenna and gridded electric fields are stochastic and therefore must be imaged and squared before time-averaging.
+\item In visibility-based imaging, electric fields measured by antennas are not correlated with themselves and hence lack zero spacing measurements.
+In contrast, in MOFF imaging, since the gridded electric fields are imaged and squared, they retain information from auto-correlated electric fields at zero spacing {\bf but can be removed in EPIC.}
 \end{enumerate}
@@
 These calibrated electric fields are gridded using an antenna-based gridding convolution function after which it is spatially Fourier transformed and squared to obtain an image cube for every time stamp.
-These images are then time-averaged to obtain the accumulated image $\mathcal{I}^\prime(f)$ ($\boldsymbol{\mathcal{I}}(\hat{\mathbf{s}})$ in matrix notation).
+These images are then time-averaged to obtain the accumulated image $\mathcal{I}^\prime(f)$ ($\boldsymbol{\mathcal{I}^\prime}(\hat{\mathbf{s}})$ in matrix notation).
 

### 2016-08-20 imaging_demo/ms.tex
Simplify notation in section 5
From 60cde0b (Adam-v1):
======================

In ms.tex,

(*) Simplify notation in section 5

@@
 The noise-like cross-terms ideally have zero mean.
-This expression can be further simplified using the sky brightness, $\mathcal{I}(\hat{\mathbf{s}},f,t)$, as $\left\langle \mathcal{E}(\hat{\mathbf{s}},f,t)\,\mathcal{E}^\star(\hat{\mathbf{s'}},f,t) \right\rangle_t = \mathcal{I}(\hat{\mathbf{s}},f,t)\,\delta(\hat{\mathbf{s}}-\hat{\mathbf{s'}})$, and defining the antenna-pair sky power response function (or the directional antenna-pair power pattern), $\mathcal{W}_{ab}(\hat{\mathbf{s}})~\equiv~\mathcal{W}_a(\hat{\mathbf{s}},f,t)~\mathcal{W}^\star_b(\hat{\mathbf{s}},f,t)$.
+This expression can be further simplified using the sky brightness, $\mathcal{I}(\hat{\mathbf{s}},f,t)$, as $\left\langle \mathcal{E}(\hat{\mathbf{s}},f,t)\,\mathcal{E}^\star(\hat{\mathbf{s'}},f,t) \right\rangle_t = \mathcal{I}(\hat{\mathbf{s}},f,t)\,\delta(\hat{\mathbf{s}}-\hat{\mathbf{s'}})$, and defining the antenna-pair sky power response function (or the directional antenna-pair power pattern), $\mathcal{B}_{ab}(\hat{\mathbf{s}})~\equiv~\mathcal{W}_a(\hat{\mathbf{s}},f,t)~\mathcal{W}^\star_b(\hat{\mathbf{s}},f,t)$.
 The result is the visibility expressed in terms of the sky brightness, the power pattern, and uncorrelated noise terms which we group into $n_{ab}(f,t)$.
 \begin{align}\label{eqn:measurement-eqn-1}
-  V_{ab}(f,t) &= \int \mathcal{W}_{ab}(\hat{\mathbf{s}},f,t)\,\mathcal{I}(\hat{\mathbf{s}},f,t)\,e^{-2\pi i f\mathbf{r}_{ab}\!\cdot\,\hat{\mathbf{s}}/c}\,\dif\Omega \nonumber\\
+  V_{ab}(f,t) &= \int \mathcal{B}_{ab}(\hat{\mathbf{s}},f,t)\,\mathcal{I}(\hat{\mathbf{s}},f,t)\,e^{-2\pi i f\mathbf{r}_{ab}\!\cdot\,\hat{\mathbf{s}}/c}\,\dif\Omega \nonumber\\
   &\qquad\qquad + n_{ab}(f,t),
@@
 where, the baseline coordinate $\mathbf{r}_{ab}\equiv\mathbf{r}_a-\mathbf{r}_b$ is the vector separation between the two antennas.
-This signifies that the visibility ($V_{ab}$) measured between a pair of antennas is obtained by multiplying the sky brightness $\mathcal{I}(\hat{\mathbf{s}},f,t)$ by the antenna power response $\mathcal{W}_{ab}(\hat{\mathbf{s}},f,t)$ and Fourier transforming from the directional coordinates ($\hat{\mathbf{s}}$) to {\bf the measurement plane}, which are then sampled at the locations of the antenna spacings (or baselines), namely, $\mathbf{r}_{ab}$, and added to the noise $n_{ab}$.
+This signifies that the visibility ($V_{ab}$) measured between a pair of antennas is obtained by multiplying the sky brightness $\mathcal{I}(\hat{\mathbf{s}},f,t)$ by the antenna power response $\mathcal{B}_{ab}(\hat{\mathbf{s}},f,t)$ and Fourier transforming from the directional coordinates ($\hat{\mathbf{s}}$) to {\bf the measurement plane}, which are then sampled at the locations of the antenna spacings (or baselines), namely, $\mathbf{r}_{ab}$, and added to the noise $n_{ab}$.
 
@@
 \begin{align}\label{eqn:software-holography}
-  V_{ab}(f,t) &= \int W_{ab}(\mathbf{r}_{ab}-\mathbf{r},f,t) \nonumber\\ 
+  V_{ab}(f,t) &= \int B_{ab}(\mathbf{r}_{ab}-\mathbf{r},f,t) \nonumber\\ 
               &\qquad \times \left[\int I(\hat{\mathbf{s}},f,t)\,e^{-2\pi i f\mathbf{r}.\hat{\mathbf{s}}/c}\,\dif\Omega\right]\dif^2\mathbf{r}\,+\, n_{ab}(f,t),
 \end{align}
-where, $W_{ab}(\mathbf{r},f,t)$ denotes the {\bf power response of the antenna pair obtained in the measurement plane} by a spatial Fourier transform of $\mathcal{W}_{ab}(\hat{\mathbf{s}},f,t)$.
-Effectively, the multiplication in image space by $\mathcal{W}_{ab}(\hat{\mathbf{s}},f,t)$ has been replaced by a convolution with $W_{ab}(\mathbf{r},f,t)$ in {\bf measurement plane}.
+where, $B_{ab}(\mathbf{r},f,t)$ denotes the {\bf power response of the antenna pair obtained in the measurement plane} by a spatial Fourier transform of $\mathcal{B}_{ab}(\hat{\mathbf{s}},f,t)$.
+Effectively, the multiplication in image space by $\mathcal{B}_{ab}(\hat{\mathbf{s}},f,t)$ has been replaced by a convolution with $B_{ab}(\mathbf{r},f,t)$ in {\bf measurement plane}.
 This is the software holographic equivalent of a traditional FX correlator output.
@@
 Unlike the assumption frequently made in standard radio interferometry that all antennas are identical, EPIC can implement the most generic case where no such assumption can be made.
-In other words, EPIC does not assume $W_a(\mathbf{r}-\mathbf{r}_a)$ or $W_{ab}(\mathbf{r}-\mathbf{r}_{ab})$ for different antennas are identical.
+In other words, EPIC does not assume $W_a(\mathbf{r}-\mathbf{r}_a)$ or $B_{ab}(\mathbf{r}-\mathbf{r}_{ab})$ for different antennas are identical.
 However, any {\it a priori} knowledge available that certain sets of antennas have identical illumination patterns can be easily passed on to EPIC to avoid redundant computations and thus increase efficiency.
@@
 \begin{align}
-  \mathcal{I}_\textrm{D}(\hat{\mathbf{s}},f) &= \frac{1}{\Nant(\Nant-1)/2}\,\sum_{ab}\, V_{ab}(f)\,e^{2\pi i f\mathbf{r}_{ab}\!\cdot\,\hat{\mathbf{s}}/c}.
-\end{align}
+  \mathcal{I}_\textrm{D}(\hat{\mathbf{s}},f) &= \frac{1}{\Nant(\Nant-1)/2}\,\sum_{ab}\, V_{ab}(f)\,e^{2\pi i f\mathbf{r}_{ab}\!\cdot\,\hat{\mathbf{s}}/c},
+\end{align}
+{\bf where, we have assumed equal weighting for each antenna pair}.
 
@@
 \end{align}
-where, $\mathcal{I}_\textrm{D}^\textrm{iso}(\hat{\mathbf{s}},f)$ is the dirty image with no beam weighting (isotropic uniform weighting) determined only by the array layout and $|\mathcal{W}^\textrm{I}|^2$ is the directional power pattern of the antenna pair familiar in standard interferometry.
-
-While imaging, EPIC introduces a weighting to each antenna during gridding as given by equations~\ref{eqn:dirty-image-FX} and \ref{eqn:e-field-conv}, and equivalently, the weighted visibility from antenna pair $ab$ projected on the grid is:
+where, $\mathcal{I}_\textrm{D}^\textrm{iso}(\hat{\mathbf{s}},f)$ is the dirty image with no beam weighting (isotropic, uniform weighting {\bf of the sky}) determined only by the array layout and $|\mathcal{W}^\textrm{I}|^2$ is the directional power pattern of the antenna pair familiar in standard interferometry.
+
+{\bf In order to understand an image from a heterogeneous array, we start by looking at the contribution to the image from the visibility, $V_{ab}$, from each antenna pair.} While imaging, EPIC introduces a weighting to each antenna during gridding as given by equations~\ref{eqn:dirty-image-FX} and \ref{eqn:e-field-conv}, and equivalently, the weighted visibility from {\bf an} antenna pair $ab$ projected on the grid is:
 \begin{align}
@@
 
-We consider a heterogeneous array consisting of a total of $\Nant$ antennas under $N_\textrm{T}$ different types.
+{\bf For a heterogeneous array, it is necessary to keep track of antennas and antenna pairs of different types which will determine the final weighting in the synthesized image.} We consider a heterogeneous array consisting of a total of $\Nant$ antennas under $N_\textrm{T}$ different types.
 And under each antenna type $i$, there are $n_i$ antennas such that $\sum_i n_i = \Nant$.
 The total number of unique non-zero spacing antenna pairs is $\Nant(\Nant-1)/2$ and there are potentially $N_\textrm{T}(N_\textrm{T}+1)/2$ unique antenna pair types obtained by pairwise combination of the $N_\textrm{T}$ antenna types.
-Thus,
+Thus, the total number of antenna pairs,
 \begin{align}\label{eqn:total-baselines}
-  \frac{1}{2}\Nant(\Nant-1) &= \frac{1}{2}\,\mathop{\sum_{i=1}^{N_\textrm{T}}\sum_{j=1}^{N_\textrm{T}}}_{i\ne j} n_i\,n_j + \frac{1}{2}\,\sum_{i=1}^{N_\textrm{T}} n_i(n_i-1)
-\end{align}
-is obtained by counting antenna pairs from different antenna types (first term) and those from the same antenna type (second term).
+  \frac{1}{2}\Nant(\Nant-1) &= \frac{1}{2}\,\sum_{ij} n_{ij}, \\
+  \textrm{where}, \quad   n_{ij} &= \begin{cases}
+    n_i\,n_j & i\ne i \\
+    n_i(n_i-1) & i=j, 
+  \end{cases}
+  % \frac{1}{2}\Nant(\Nant-1) &= \frac{1}{2}\,\mathop{\sum_{i=1}^{N_\textrm{T}}\sum_{j=1}^{N_\textrm{T}}}_{i\ne j} n_i\,n_j + \frac{1}{2}\,\sum_{i=1}^{N_\textrm{T}} n_i(n_i-1) \nonumber\\
+\end{align}
+is obtained by counting antenna pairs, $n_{ij}$, in different antenna pair types, $ij$.
 
@@
 \begin{align}\label{eqn:wt-dirty-image-EPIC-decomp}
-  \mathcal{I}^\prime(\hat{\mathbf{s}},f) &= \frac{1}{\Nant(\Nant-1)/2}\,\left[\frac{1}{2}\mathop{\sum_{i=1}^{N_\textrm{T}}\sum_{j=1}^{N_\textrm{T}}}_{i\ne j}\sum_{\substack{ab\\a\ne b\\a\in i, b\in j}}\,\mathcal{I}^\prime_{ab}(\hat{\mathbf{s}},f)\right.
+  \mathcal{I}^\prime(\hat{\mathbf{s}},f) &= \frac{1/2}{\Nant(\Nant-1)/2}\,\sum_{ij}\sum_{\substack{ab\\a\ne b\\a\in i, b\in j}}\,\mathcal{I}^\prime_{ab}(\hat{\mathbf{s}},f)
+\end{align}
+% \begin{align}\label{eqn:wt-dirty-image-EPIC-decomp}
+%   \mathcal{I}^\prime(\hat{\mathbf{s}},f) &= \frac{1}{\Nant(\Nant-1)/2}\,\left[\frac{1}{2}\mathop{\sum_{i=1}^{N_\textrm{T}}\sum_{j=1}^{N_\textrm{T}}}_{i\ne j}\sum_{\substack{ab\\a\ne b\\a\in i, b\in j}}\,\mathcal{I}^\prime_{ab}(\hat{\mathbf{s}},f)\right.
 \nonumber\\
-  &\qquad\qquad\qquad\qquad\qquad\quad + \left.\frac{1}{2}\sum_{i=1}^{N_\textrm{T}}\sum_{\substack{ab\\a\ne b\\a,b\in i}}\,\mathcal{I}^\prime_{ab}(\hat{\mathbf{s}},f)\right]. 
-\end{align}
+%   &\qquad\qquad\qquad\qquad\qquad\quad + \left.\frac{1}{2}\sum_{i=1}^{N_\textrm{T}}\sum_{\substack{ab\\a\ne b\\a,b\in i}}\,\mathcal{I}^\prime_{ab}(\hat{\mathbf{s}},f)\right]. 
+% \end{align}
 Here, equation~\ref{eqn:wt-dirty-image-EPIC} has been decomposed as the sum of fringes over all unique antenna pairs in an antenna pair type (innermost summation) and subsequently summed over all antenna pair types.
 
-In order to simplify the analysis, we assume that the spatial distribution of antenna pairs in each antenna pair type are similar and thus result in a similar dirty image, $\mathcal{I}_\textrm{D}^\textrm{iso}(\hat{\mathbf{s}},f)$.
+{\bf For purposes of demonstration in this specific study, we make the following simplification.} We assume that the spatial distribution of antenna pairs in each antenna pair type are similar and thus result in a similar dirty image, $\mathcal{I}_\textrm{D}^\textrm{iso}(\hat{\mathbf{s}},f)$.
 This usually holds when $\Nant$ is large and antennas under each antenna type are chosen randomly such that differences in the point spread functions are insignificant.
@@
 \begin{align}\label{eqn:wt-dirty-image-EPIC-decomp-approx1}
-  \mathcal{I}^\prime(\hat{\mathbf{s}},f) &\approx \frac{\mathcal{I}_\textrm{D}^\textrm{iso}(\hat{\mathbf{s}},f)}{\Nant(\Nant-1)/2} \nonumber\\
-  &\,\times \left[\frac{1}{2}\mathop{\sum_{i=1}^{N_\textrm{T}}\sum_{j=1}^{N_\textrm{T}}}_{i\ne j}\sum_{\substack{ab\\a\ne b\\a\in i, b\in j}}\left\{\mathcal{W}^{\textrm{G}\star}_a(\hat{\mathbf{s}},f)\mathcal{W}^\textrm{G}_b(\hat{\mathbf{s}},f)\right.\right.\nonumber\\
-  &\qquad\qquad\qquad\qquad\qquad\qquad \left. \times \mathcal{W}^\textrm{I}_a(\hat{\mathbf{s}},f)\mathcal{W}^{\textrm{I}\star}_b(\hat{\mathbf{s}},f)\right\} \nonumber\\ 
-  &\qquad + \frac{1}{2}\sum_{i=1}^{N_\textrm{T}}\sum_{\substack{ab\\a\ne b\\a,b\in i}}\left\{\mathcal{W}^{\textrm{G}\star}_a(\hat{\mathbf{s}},f)\mathcal{W}^\textrm{G}_b(\hat{\mathbf{s}},f)\right.
+  \mathcal{I}^\prime(\hat{\mathbf{s}},f) &\approx \frac{\mathcal{I}_\textrm{D}^\textrm{iso}(\hat{\mathbf{s}},f)}{\Nant(\Nant-1)}\,\sum_{ij}\sum_{\substack{ab\\a\ne b\\a\in i, b\in j}} \left\{\mathcal{W}^{\textrm{G}\star}_a(\hat{\mathbf{s}},f)\mathcal{W}^\textrm{G}_b(\hat{\mathbf{s}},f)\right.\nonumber\\
+  &\qquad\qquad\qquad\qquad\qquad\qquad \left.\times\mathcal{W}^\textrm{I}_a(\hat{\mathbf{s}},f)\mathcal{W}^{\textrm{I}\star}_b(\hat{\mathbf{s}},f)\right\}
+  % &\,\times \left[\frac{1}{2}\mathop{\sum_{i=1}^{N_\textrm{T}}\sum_{j=1}^{N_\textrm{T}}}_{i\ne j}\sum_{\substack{ab\\a\ne b\\a\in i, b\in j}}\left\{\mathcal{W}^{\textrm{G}\star}_a(\hat{\mathbf{s}},f)\mathcal{W}^\textrm{G}_b(\hat{\mathbf{s}},f)\right.\right.\nonumber\\
+  % &\qquad\qquad\qquad\qquad\qquad\qquad \left. \times \mathcal{W}^\textrm{I}_a(\hat{\mathbf{s}},f)\mathcal{W}^{\textrm{I}\star}_b(\hat{\mathbf{s}},f)\right\} \nonumber\\ 
+  % &\qquad + \frac{1}{2}\sum_{i=1}^{N_\textrm{T}}\sum_{\substack{ab\\a\ne b\\a,b\in i}}\left\{\mathcal{W}^{\textrm{G}\star}_a(\hat{\mathbf{s}},f)\mathcal{W}^\textrm{G}_b(\hat{\mathbf{s}},f)\right.
 \nonumber\\
-  &\qquad\qquad\qquad\qquad\qquad\qquad \left.\left.\times\mathcal{W}^\textrm{I}_a(\hat{\mathbf{s}},f)\mathcal{W}^{\textrm{I}\star}_b(\hat{\mathbf{s}},f)\right\}\vphantom{\sum_{\substack{ab\\a\ne b\\a\in i, b\in j}}}\right],
-\end{align}
+  % &\qquad\qquad\qquad\qquad\qquad\qquad \left.\left.\times\mathcal{W}^\textrm{I}_a(\hat{\mathbf{s}},f)\mathcal{W}^{\textrm{I}\star}_b(\hat{\mathbf{s}},f)\right\}\vphantom{\sum_{\substack{ab\\a\ne b\\a\in i, b\in j}}}\right],
+\end{align}
+% \begin{align}\label{eqn:wt-dirty-image-EPIC-decomp-approx1}
+%   \mathcal{I}^\prime(\hat{\mathbf{s}},f) &\approx \frac{\mathcal{I}_\textrm{D}^\textrm{iso}(\hat{\mathbf{s}},f)}{\Nant(\Nant-1)/2} \nonumber\\
+%   &\,\times \left[\frac{1}{2}\mathop{\sum_{i=1}^{N_\textrm{T}}\sum_{j=1}^{N_\textrm{T}}}_{i\ne j}\sum_{\substack{ab\\a\ne b\\a\in i, b\in j}}\left\{\mathcal{W}^{\textrm{G}\star}_a(\hat{\mathbf{s}},f)\mathcal{W}^\textrm{G}_b(\hat{\mathbf{s}},f)\right.\right.\nonumber\\
+%   &\qquad\qquad\qquad\qquad\qquad\qquad \left. \times \mathcal{W}^\textrm{I}_a(\hat{\mathbf{s}},f)\mathcal{W}^{\textrm{I}\star}_b(\hat{\mathbf{s}},f)\right\} \nonumber\\ 
+%   &\qquad + \frac{1}{2}\sum_{i=1}^{N_\textrm{T}}\sum_{\substack{ab\\a\ne b\\a,b\in i}}\left\{\mathcal{W}^{\textrm{G}\star}_a(\hat{\mathbf{s}},f)\mathcal{W}^\textrm{G}_b(\hat{\mathbf{s}},f)\right.
+\nonumber\\
+%   &\qquad\qquad\qquad\qquad\qquad\qquad \left.\left.\times\mathcal{W}^\textrm{I}_a(\hat{\mathbf{s}},f)\mathcal{W}^{\textrm{I}\star}_b(\hat{\mathbf{s}},f)\right\}\vphantom{\sum_{\substack{ab\\a\ne b\\a\in i, b\in j}}}\right],
+% \end{align}
 where, $\mathcal{I}_\textrm{D}^\textrm{iso}(\hat{\mathbf{s}},f)$ is approximately the same for all terms and has been pulled outside the summations.
-For all antennas indexed by $a$ that belong to a particular type $i$, we replace $\mathcal{W}_a$ with $\mathcal{W}_{(i)}$ that arise from both instrumental (superscript `$I$') and gridding (superscript `$G$') origins.
+For all antennas indexed by $a$ that belong to a particular type $i$, we replace $\mathcal{W}_a$ with $\mathcal{W}_{(i)}$ {\bf and is applicable to those} that arise from both instrumental (superscript `$I$') and gridding (superscript `$G$') origins.
 Using arguments similar to those in equation~\ref{eqn:total-baselines}, equation~\ref{eqn:wt-dirty-image-EPIC-decomp-approx1} can be further simplified to:
 \begin{align}\label{eqn:wt-dirty-image-EPIC-decomp-approx2}
-  \mathcal{I}^\prime(\hat{\mathbf{s}},f) &\approx \frac{\mathcal{I}_\textrm{D}^\textrm{iso}(\hat{\mathbf{s}},f)}{\Nant(\Nant-1)/2} \nonumber\\
-  &\,\times \left[\frac{1}{2}\mathop{\sum_{i=1}^{N_\textrm{T}}\sum_{j=1}^{N_\textrm{T}}}_{i\ne j}n_in_j\left\{\mathcal{W}^{\textrm{G}\star}_{(i)}(\hat{\mathbf{s}},f)\mathcal{W}^\textrm{G}_{(j)}(\hat{\mathbf{s}},f)\right.\right.\nonumber\\
-  &\qquad\qquad\qquad\qquad\qquad\qquad \left. \times \mathcal{W}^\textrm{I}_{(i)}(\hat{\mathbf{s}},f)\mathcal{W}^{\textrm{I}\star}_{(j)}(\hat{\mathbf{s}},f)\right\} \nonumber\\ 
-  &\qquad + \frac{1}{2}\sum_{i=1}^{N_\textrm{T}}n_i(n_i-1)\,\left|\mathcal{W}^\textrm{G}_{(i)}(\hat{\mathbf{s}},f)\right|^2\,\left|\mathcal{W}^\textrm{I}_{(i)}(\hat{\mathbf{s}},f)\right|^2\left.\vphantom{\sum_{\substack{ab\\a\ne b\\a\in i, b\in j}}}\right].
-\end{align}
+  \mathcal{I}^\prime(\hat{\mathbf{s}},f) &\approx \frac{\mathcal{I}_\textrm{D}^\textrm{iso}(\hat{\mathbf{s}},f)}{\Nant(\Nant-1)}\,\sum_{ij}n_{ij}\left\{\mathcal{W}^{\textrm{G}\star}_{(i)}(\hat{\mathbf{s}},f)\mathcal{W}^\textrm{G}_{(j)}(\hat{\mathbf{s}},f)\right.\nonumber\\
+  &\qquad\qquad\qquad\qquad\qquad\qquad \left. \times \mathcal{W}^\textrm{I}_{(i)}(\hat{\mathbf{s}},f)\mathcal{W}^{\textrm{I}\star}_{(j)}(\hat{\mathbf{s}},f)\right\}.
+\end{align}
+% \begin{align}\label{eqn:wt-dirty-image-EPIC-decomp-approx2}
+%   \mathcal{I}^\prime(\hat{\mathbf{s}},f) &\approx \frac{\mathcal{I}_\textrm{D}^\textrm{iso}(\hat{\mathbf{s}},f)}{\Nant(\Nant-1)/2} \nonumber\\
+%   &\,\times \left[\frac{1}{2}\mathop{\sum_{i=1}^{N_\textrm{T}}\sum_{j=1}^{N_\textrm{T}}}_{i\ne j}n_in_j\left\{\mathcal{W}^{\textrm{G}\star}_{(i)}(\hat{\mathbf{s}},f)\mathcal{W}^\textrm{G}_{(j)}(\hat{\mathbf{s}},f)\right.\right.\nonumber\\
+%   &\qquad\qquad\qquad\qquad\qquad\qquad \left. \times \mathcal{W}^\textrm{I}_{(i)}(\hat{\mathbf{s}},f)\mathcal{W}^{\textrm{I}\star}_{(j)}(\hat{\mathbf{s}},f)\right\} \nonumber\\ 
+%   &\qquad + \frac{1}{2}\sum_{i=1}^{N_\textrm{T}}n_i(n_i-1)\,\left|\mathcal{W}^\textrm{G}_{(i)}(\hat{\mathbf{s}},f)\right|^2\,\left|\mathcal{W}^\textrm{I}_{(i)}(\hat{\mathbf{s}},f)\right|^2\left.\vphantom{\sum_{\substack{ab\\a\ne b\\a\in i, b\in j}}}\right].
+% \end{align}
 Then the effective attenuation of the dirty image due to instrumental and gridding weights is:
@@
 \begin{align}\label{eqn:effective-weighting-optimal}
-  \mathcal{W}_\textrm{eff}^\textrm{opt} &\approx \frac{1}{\Nant(\Nant-1)/2} \nonumber\\
-  &\quad \times \left[\frac{1}{2}\mathop{\sum_{i=1}^{N_\textrm{T}}\sum_{j=1}^{N_\textrm{T}}}_{i\ne j}n_in_j\,\left|\mathcal{W}^\textrm{I}_{(i)}(\hat{\mathbf{s}},f)\right|^2\left|\mathcal{W}^\textrm{I}_{(j)}(\hat{\mathbf{s}},f)\right|^2\right.\nonumber\\
-  &\qquad\qquad + \frac{1}{2}\sum_{i=1}^{N_\textrm{T}}n_i(n_i-1)\,\left|\mathcal{W}^\textrm{I}_{(i)}(\hat{\mathbf{s}},f)\right|^4\left.\vphantom{\sum_{\substack{ab\\a\ne b\\a\in i, b\in j}}}\right],
-\end{align}
+  \mathcal{W}_\textrm{eff}^\textrm{opt} &\approx \frac{1}{\Nant(\Nant-1)}\,\sum_{ij} n_{ij}\,\left|\mathcal{W}^\textrm{I}_{(i)}(\hat{\mathbf{s}},f)\right|^2\left|\mathcal{W}^\textrm{I}_{(j)}(\hat{\mathbf{s}},f)\right|^2,
+  % &\qquad\qquad + \frac{1}{2}\sum_{i=1}^{N_\textrm{T}}n_i(n_i-1)\,\left|\mathcal{W}^\textrm{I}_{(i)}(\hat{\mathbf{s}},f)\right|^4\left.\vphantom{\sum_{\substack{ab\\a\ne b\\a\in i, b\in j}}}\right],
+\end{align}
+% \begin{align}\label{eqn:effective-weighting-optimal}
+%   \mathcal{W}_\textrm{eff}^\textrm{opt} &\approx \frac{1}{\Nant(\Nant-1)/2} \nonumber\\
+%   &\quad \times \left[\frac{1}{2}\mathop{\sum_{i=1}^{N_\textrm{T}}\sum_{j=1}^{N_\textrm{T}}}_{i\ne j}n_in_j\,\left|\mathcal{W}^\textrm{I}_{(i)}(\hat{\mathbf{s}},f)\right|^2\left|\mathcal{W}^\textrm{I}_{(j)}(\hat{\mathbf{s}},f)\right|^2\right.\nonumber\\
+%   &\qquad\qquad + \frac{1}{2}\sum_{i=1}^{N_\textrm{T}}n_i(n_i-1)\,\left|\mathcal{W}^\textrm{I}_{(i)}(\hat{\mathbf{s}},f)\right|^4\left.\vphantom{\sum_{\substack{ab\\a\ne b\\a\in i, b\in j}}}\right],
+% \end{align}
 where the superscript `$\textrm{opt}$' denotes 'optimal'.

### 2016-08-20 imaging_demo/ms.tex
Simplify notation in section 5 further
From 8bce8b2 (Adam-v1):
======================

In ms.tex,

(*) Simplify notation in section 5 further

@@
 
-We present a methodology to understand the effective angular weighting in the image obtained with EPIC from a heterogeneous array and compare it with that from standard interferometry that assumes a homogeneous array (all antennas are identical).
+We present a methodology to understand the effective angular weighting in the image obtained with EPIC from a heterogeneous array and compare it with that from standard interferometry which assumes a homogeneous array (all antennas are identical).
 We also demonstrate how the assumption in the latter results in a mis-estimation of the recovered sky model.
@@
 
+{\bf We note that all quantities in the celestial plane (calligraphic fonts) are functions of position, $\hat{\mathbf{s}}$, and frequency, $f$.
+For convenience, we drop writing this dependence explicitly hereafter.}
+
 {\bf For a heterogeneous array, it is necessary to keep track of antennas and antenna pairs of different types which will determine the final weighting in the synthesized image.} We consider a heterogeneous array consisting of a total of $\Nant$ antennas under $N_\textrm{T}$ different types.
 And under each antenna type $i$, there are $n_i$ antennas such that $\sum_i n_i = \Nant$.
-The total number of unique non-zero spacing antenna pairs is $\Nant(\Nant-1)/2$ and there are potentially $N_\textrm{T}(N_\textrm{T}+1)/2$ unique antenna pair types obtained by pairwise combination of the $N_\textrm{T}$ antenna types.
+The total number of unique non-zero spacing antenna pairs is $\Nant(\Nant-1)/2$ and there are potentially up to $N_\textrm{T}(N_\textrm{T}+1)/2$ unique antenna pair types obtained by pairwise combination of the $N_\textrm{T}$ antenna types.
 Thus, the total number of antenna pairs,
@@
   \end{cases}
-  % \frac{1}{2}\Nant(\Nant-1) &= \frac{1}{2}\,\mathop{\sum_{i=1}^{N_\textrm{T}}\sum_{j=1}^{N_\textrm{T}}}_{i\ne j} n_i\,n_j + \frac{1}{2}\,\sum_{i=1}^{N_\textrm{T}} n_i(n_i-1) \nonumber\\
 \end{align}
@@
 \begin{align}\label{eqn:wt-dirty-image-EPIC-decomp}
-  \mathcal{I}^\prime(\hat{\mathbf{s}},f) &= \frac{1/2}{\Nant(\Nant-1)/2}\,\sum_{ij}\sum_{\substack{ab\\a\ne b\\a\in i, b\in j}}\,\mathcal{I}^\prime_{ab}(\hat{\mathbf{s}},f)
-\end{align}
-% \begin{align}\label{eqn:wt-dirty-image-EPIC-decomp}
-%   \mathcal{I}^\prime(\hat{\mathbf{s}},f) &= \frac{1}{\Nant(\Nant-1)/2}\,\left[\frac{1}{2}\mathop{\sum_{i=1}^{N_\textrm{T}}\sum_{j=1}^{N_\textrm{T}}}_{i\ne j}\sum_{\substack{ab\\a\ne b\\a\in i, b\in j}}\,\mathcal{I}^\prime_{ab}(\hat{\mathbf{s}},f)\right.
-\nonumber\\
-%   &\qquad\qquad\qquad\qquad\qquad\quad + \left.\frac{1}{2}\sum_{i=1}^{N_\textrm{T}}\sum_{\substack{ab\\a\ne b\\a,b\in i}}\,\mathcal{I}^\prime_{ab}(\hat{\mathbf{s}},f)\right]. 
-% \end{align}
+  \mathcal{I}^\prime &= \frac{1/2}{\Nant(\Nant-1)/2}\,\sum_{ij}\sum_{\substack{ab\\a\ne b\\a\in i, b\in j}}\,\mathcal{I}^\prime_{ab}
+\end{align}
 Here, equation~\ref{eqn:wt-dirty-image-EPIC} has been decomposed as the sum of fringes over all unique antenna pairs in an antenna pair type (innermost summation) and subsequently summed over all antenna pair types.
 
-{\bf For purposes of demonstration in this specific study, we make the following simplification.} We assume that the spatial distribution of antenna pairs in each antenna pair type are similar and thus result in a similar dirty image, $\mathcal{I}_\textrm{D}^\textrm{iso}(\hat{\mathbf{s}},f)$.
+{\bf For purposes of demonstration in this study, we make the following simplification.} We assume that the spatial distribution of antenna pairs in each antenna pair type are similar and thus result in a similar dirty image, $\mathcal{I}_\textrm{D}^\textrm{iso}$.
 This usually holds when $\Nant$ is large and antennas under each antenna type are chosen randomly such that differences in the point spread functions are insignificant.
@@
 \begin{align}\label{eqn:wt-dirty-image-EPIC-decomp-approx1}
-  \mathcal{I}^\prime(\hat{\mathbf{s}},f) &\approx \frac{\mathcal{I}_\textrm{D}^\textrm{iso}(\hat{\mathbf{s}},f)}{\Nant(\Nant-1)}\,\sum_{ij}\sum_{\substack{ab\\a\ne b\\a\in i, b\in j}} \left\{\mathcal{W}^{\textrm{G}\star}_a(\hat{\mathbf{s}},f)\mathcal{W}^\textrm{G}_b(\hat{\mathbf{s}},f)\right.\nonumber\\
-  &\qquad\qquad\qquad\qquad\qquad\qquad \left.\times\mathcal{W}^\textrm{I}_a(\hat{\mathbf{s}},f)\mathcal{W}^{\textrm{I}\star}_b(\hat{\mathbf{s}},f)\right\}
-  % &\,\times \left[\frac{1}{2}\mathop{\sum_{i=1}^{N_\textrm{T}}\sum_{j=1}^{N_\textrm{T}}}_{i\ne j}\sum_{\substack{ab\\a\ne b\\a\in i, b\in j}}\left\{\mathcal{W}^{\textrm{G}\star}_a(\hat{\mathbf{s}},f)\mathcal{W}^\textrm{G}_b(\hat{\mathbf{s}},f)\right.\right.\nonumber\\
-  % &\qquad\qquad\qquad\qquad\qquad\qquad \left. \times \mathcal{W}^\textrm{I}_a(\hat{\mathbf{s}},f)\mathcal{W}^{\textrm{I}\star}_b(\hat{\mathbf{s}},f)\right\} \nonumber\\ 
-  % &\qquad + \frac{1}{2}\sum_{i=1}^{N_\textrm{T}}\sum_{\substack{ab\\a\ne b\\a,b\in i}}\left\{\mathcal{W}^{\textrm{G}\star}_a(\hat{\mathbf{s}},f)\mathcal{W}^\textrm{G}_b(\hat{\mathbf{s}},f)\right.
-\nonumber\\
-  % &\qquad\qquad\qquad\qquad\qquad\qquad \left.\left.\times\mathcal{W}^\textrm{I}_a(\hat{\mathbf{s}},f)\mathcal{W}^{\textrm{I}\star}_b(\hat{\mathbf{s}},f)\right\}\vphantom{\sum_{\substack{ab\\a\ne b\\a\in i, b\in j}}}\right],
-\end{align}
-% \begin{align}\label{eqn:wt-dirty-image-EPIC-decomp-approx1}
-%   \mathcal{I}^\prime(\hat{\mathbf{s}},f) &\approx \frac{\mathcal{I}_\textrm{D}^\textrm{iso}(\hat{\mathbf{s}},f)}{\Nant(\Nant-1)/2} \nonumber\\
-%   &\,\times \left[\frac{1}{2}\mathop{\sum_{i=1}^{N_\textrm{T}}\sum_{j=1}^{N_\textrm{T}}}_{i\ne j}\sum_{\substack{ab\\a\ne b\\a\in i, b\in j}}\left\{\mathcal{W}^{\textrm{G}\star}_a(\hat{\mathbf{s}},f)\mathcal{W}^\textrm{G}_b(\hat{\mathbf{s}},f)\right.\right.\nonumber\\
-%   &\qquad\qquad\qquad\qquad\qquad\qquad \left. \times \mathcal{W}^\textrm{I}_a(\hat{\mathbf{s}},f)\mathcal{W}^{\textrm{I}\star}_b(\hat{\mathbf{s}},f)\right\} \nonumber\\ 
-%   &\qquad + \frac{1}{2}\sum_{i=1}^{N_\textrm{T}}\sum_{\substack{ab\\a\ne b\\a,b\in i}}\left\{\mathcal{W}^{\textrm{G}\star}_a(\hat{\mathbf{s}},f)\mathcal{W}^\textrm{G}_b(\hat{\mathbf{s}},f)\right.
-\nonumber\\
-%   &\qquad\qquad\qquad\qquad\qquad\qquad \left.\left.\times\mathcal{W}^\textrm{I}_a(\hat{\mathbf{s}},f)\mathcal{W}^{\textrm{I}\star}_b(\hat{\mathbf{s}},f)\right\}\vphantom{\sum_{\substack{ab\\a\ne b\\a\in i, b\in j}}}\right],
-% \end{align}
-where, $\mathcal{I}_\textrm{D}^\textrm{iso}(\hat{\mathbf{s}},f)$ is approximately the same for all terms and has been pulled outside the summations.
-For all antennas indexed by $a$ that belong to a particular type $i$, we replace $\mathcal{W}_a$ with $\mathcal{W}_{(i)}$ {\bf and is applicable to those} that arise from both instrumental (superscript `$I$') and gridding (superscript `$G$') origins.
+  \mathcal{I}^\prime &\approx \frac{\mathcal{I}_\textrm{D}^\textrm{iso}}{\Nant(\Nant-1)}\,\sum_{ij}\sum_{\substack{ab\\a\ne b\\a\in i, b\in j}} \mathcal{W}^{\textrm{G}\star}_a\,\mathcal{W}^\textrm{G}_b\,\mathcal{W}^\textrm{I}_a\,\mathcal{W}^{\textrm{I}\star}_b,
+\end{align}
+where, $\mathcal{I}_\textrm{D}^\textrm{iso}$ is approximately the same for all terms and has been pulled outside the summations.
+For all antennas indexed by $a$ that belong to a particular type $i$, we replace $\mathcal{W}_a$ with $\mathcal{W}_{(i)}$ {\bf and is applicable to those arising} from both instrumental (superscript `$I$') and gridding (superscript `$G$') origins.
 Using arguments similar to those in equation~\ref{eqn:total-baselines}, equation~\ref{eqn:wt-dirty-image-EPIC-decomp-approx1} can be further simplified to:
 \begin{align}\label{eqn:wt-dirty-image-EPIC-decomp-approx2}
-  \mathcal{I}^\prime(\hat{\mathbf{s}},f) &\approx \frac{\mathcal{I}_\textrm{D}^\textrm{iso}(\hat{\mathbf{s}},f)}{\Nant(\Nant-1)}\,\sum_{ij}n_{ij}\left\{\mathcal{W}^{\textrm{G}\star}_{(i)}(\hat{\mathbf{s}},f)\mathcal{W}^\textrm{G}_{(j)}(\hat{\mathbf{s}},f)\right.\nonumber\\
-  &\qquad\qquad\qquad\qquad\qquad\qquad \left. \times \mathcal{W}^\textrm{I}_{(i)}(\hat{\mathbf{s}},f)\mathcal{W}^{\textrm{I}\star}_{(j)}(\hat{\mathbf{s}},f)\right\}.
-\end{align}
-% \begin{align}\label{eqn:wt-dirty-image-EPIC-decomp-approx2}
-%   \mathcal{I}^\prime(\hat{\mathbf{s}},f) &\approx \frac{\mathcal{I}_\textrm{D}^\textrm{iso}(\hat{\mathbf{s}},f)}{\Nant(\Nant-1)/2} \nonumber\\
-%   &\,\times \left[\frac{1}{2}\mathop{\sum_{i=1}^{N_\textrm{T}}\sum_{j=1}^{N_\textrm{T}}}_{i\ne j}n_in_j\left\{\mathcal{W}^{\textrm{G}\star}_{(i)}(\hat{\mathbf{s}},f)\mathcal{W}^\textrm{G}_{(j)}(\hat{\mathbf{s}},f)\right.\right.\nonumber\\
-%   &\qquad\qquad\qquad\qquad\qquad\qquad \left. \times \mathcal{W}^\textrm{I}_{(i)}(\hat{\mathbf{s}},f)\mathcal{W}^{\textrm{I}\star}_{(j)}(\hat{\mathbf{s}},f)\right\} \nonumber\\ 
-%   &\qquad + \frac{1}{2}\sum_{i=1}^{N_\textrm{T}}n_i(n_i-1)\,\left|\mathcal{W}^\textrm{G}_{(i)}(\hat{\mathbf{s}},f)\right|^2\,\left|\mathcal{W}^\textrm{I}_{(i)}(\hat{\mathbf{s}},f)\right|^2\left.\vphantom{\sum_{\substack{ab\\a\ne b\\a\in i, b\in j}}}\right].
-% \end{align}
+  \mathcal{I}^\prime &\approx \frac{\mathcal{I}_\textrm{D}^\textrm{iso}}{\Nant(\Nant-1)}\,\sum_{ij}n_{ij}\,\mathcal{W}^{\textrm{G}\star}_{(i)}\,\mathcal{W}^\textrm{G}_{(j)}\,\mathcal{W}^\textrm{I}_{(i)}\,\mathcal{W}^{\textrm{I}\star}_{(j)}.
+\end{align}
 Then the effective attenuation of the dirty image due to instrumental and gridding weights is:
 \begin{align}\label{eqn:effective-weighting}
-  \mathcal{W}_\textrm{eff} &= \mathcal{I}^\prime(\hat{\mathbf{s}},f)\, / \,\mathcal{I}_\textrm{D}^\textrm{iso}(\hat{\mathbf{s}},f).
+  \mathcal{W}_\textrm{eff} &= \mathcal{I}^\prime\, / \,\mathcal{I}_\textrm{D}^\textrm{iso}.
 \end{align}
@@
 \begin{align}\label{eqn:effective-weighting-optimal}
-  \mathcal{W}_\textrm{eff}^\textrm{opt} &\approx \frac{1}{\Nant(\Nant-1)}\,\sum_{ij} n_{ij}\,\left|\mathcal{W}^\textrm{I}_{(i)}(\hat{\mathbf{s}},f)\right|^2\left|\mathcal{W}^\textrm{I}_{(j)}(\hat{\mathbf{s}},f)\right|^2,
-  % &\qquad\qquad + \frac{1}{2}\sum_{i=1}^{N_\textrm{T}}n_i(n_i-1)\,\left|\mathcal{W}^\textrm{I}_{(i)}(\hat{\mathbf{s}},f)\right|^4\left.\vphantom{\sum_{\substack{ab\\a\ne b\\a\in i, b\in j}}}\right],
-\end{align}
-% \begin{align}\label{eqn:effective-weighting-optimal}
-%   \mathcal{W}_\textrm{eff}^\textrm{opt} &\approx \frac{1}{\Nant(\Nant-1)/2} \nonumber\\
-%   &\quad \times \left[\frac{1}{2}\mathop{\sum_{i=1}^{N_\textrm{T}}\sum_{j=1}^{N_\textrm{T}}}_{i\ne j}n_in_j\,\left|\mathcal{W}^\textrm{I}_{(i)}(\hat{\mathbf{s}},f)\right|^2\left|\mathcal{W}^\textrm{I}_{(j)}(\hat{\mathbf{s}},f)\right|^2\right.\nonumber\\
-%   &\qquad\qquad + \frac{1}{2}\sum_{i=1}^{N_\textrm{T}}n_i(n_i-1)\,\left|\mathcal{W}^\textrm{I}_{(i)}(\hat{\mathbf{s}},f)\right|^4\left.\vphantom{\sum_{\substack{ab\\a\ne b\\a\in i, b\in j}}}\right],
-% \end{align}
-where the superscript `$\textrm{opt}$' denotes 'optimal'.
+  \mathcal{W}_\textrm{eff}^\textrm{opt} &\approx \frac{1}{\Nant(\Nant-1)}\,\sum_{ij} n_{ij}\,\left|\mathcal{W}^\textrm{I}_{(i)}\right|^2\left|\mathcal{W}^\textrm{I}_{(j)}\right|^2,
+\end{align}
+where the superscript `$\textrm{opt}$' denotes `optimal'.
 
 For purposes of illustration, we consider a simple scenario in which there are two antenna types each containing roughly equal numbers of antennas with similar layouts but the analysis is performed under the wrong assumption that all antennas are identical and are all of the first type.
-In an optimally weighted image, the optimal effective weighting given by equation~\ref{eqn:effective-weighting} will be:
+In an optimally weighted image, the optimal effective weighting given by equation~\ref{eqn:effective-weighting-optimal} will be:
 \begin{align}\label{eqn:effective-weighting-optimal-example}
-  \mathcal{W}_\textrm{eff}^\textrm{opt} &\approx \frac{1}{\Nant(\Nant-1)}\,\left[n_1n_2\,\left|\mathcal{W}^\textrm{I}_{(1)}(\hat{\mathbf{s}},f)\right|^2\,\left|\mathcal{W}^\textrm{I}_{(2)}(\hat{\mathbf{s}},f)\right|^2\right.\nonumber\\
-  &\qquad\qquad\qquad\qquad + n_1(n_1-1)\,\left|\mathcal{W}^\textrm{I}_{(1)}(\hat{\mathbf{s}},f)\right|^4 \nonumber\\
-  &\qquad\qquad\qquad\qquad \left. + n_2(n_2-1)\,\left|\mathcal{W}^\textrm{I}_{(2)}(\hat{\mathbf{s}},f)\right|^4\right].
-\end{align}
-On the other hand, $\mathcal{W}^\textrm{G}_{(1)}(\hat{\mathbf{s}},f) = \mathcal{W}^\textrm{I}_{(1)}(\hat{\mathbf{s}},f)$ and $\mathcal{W}^\textrm{G}_{(2)}(\hat{\mathbf{s}},f) = \mathcal{W}^\textrm{I}_{(1)}(\hat{\mathbf{s}},f)$ for an image obtained in the erroneous case in which all antennas are assumed to be identical.
-It will be sub-optimal whose effective weighting will be:
+  \mathcal{W}_\textrm{eff}^\textrm{opt} &\approx \frac{1}{\Nant(\Nant-1)}\,\left[n_1(n_1-1)\,\left|\mathcal{W}^\textrm{I}_{(1)}\right|^4 \right.\nonumber\\
+  &\qquad\qquad\qquad\qquad + n_2(n_2-1)\,\left|\mathcal{W}^\textrm{I}_{(2)}\right|^4 \nonumber\\
+  &\qquad\qquad\qquad\qquad \left. + n_1n_2\,\left|\mathcal{W}^\textrm{I}_{(1)}\right|^2\,\left|\mathcal{W}^\textrm{I}_{(2)}\right|^2\right].
+\end{align}
+On the other hand, $\mathcal{W}^\textrm{G}_{(1)} = \mathcal{W}^\textrm{I}_{(1)}$ and $\mathcal{W}^\textrm{G}_{(2)} = \mathcal{W}^\textrm{I}_{(1)}$ for an image obtained in the erroneous case in which all antennas are assumed to be identical.
+It will be sub-optimal with effective weighting:
 \begin{align}\label{eqn:effective-weighting-optimal-example}
-  \mathcal{W}_\textrm{eff}^\textrm{sub} &\approx \frac{1}{\Nant(\Nant-1)}\,\left[\left\{n_1n_2\,\left|\mathcal{W}^\textrm{I}_{(1)}(\hat{\mathbf{s}},f)\right|^2 \right.\right. \nonumber\\
-  &\qquad\qquad\qquad\qquad\qquad \left.\times\mathcal{W}^\textrm{I}_{(1)}(\hat{\mathbf{s}},f)\mathcal{W}^{\textrm{I}\star}_{(2)}(\hat{\mathbf{s}},f)\right\}\nonumber\\
-  &\qquad\qquad\qquad\qquad + n_1(n_1-1)\,\left|\mathcal{W}^\textrm{I}_{(1)}(\hat{\mathbf{s}},f)\right|^4 \nonumber\\
-  &\qquad\qquad\qquad\qquad + n_2(n_2-1)\,\left|\mathcal{W}^\textrm{I}_{(1)}(\hat{\mathbf{s}},f)\right|^2\nonumber\\
-  &\qquad\qquad\qquad\qquad\qquad\qquad\qquad \left.\times\left|\mathcal{W}^\textrm{I}_{(2)}(\hat{\mathbf{s}},f)\right|^2\right].
-\end{align}
-Finally, since no {\it a priori} information will be available about the heterogeneous array in case of the erroneous assumption, the effective weighting in the image will be believed to be $\mathrm{W}_\textrm{eff}^\textrm{err} = \left|\mathcal{W}_{(1)}^\textrm{I}\right|^4$.
-
-Estimates of the dirty images with uniform, isotropic weighting in the optimal and erroneous cases will be obtained by dividing the respective weighted images by their assumed effective weights $\mathrm{W}_\textrm{eff}^\textrm{opt}$ and $\mathrm{W}_\textrm{eff}^\textrm{err}$ respectively as:
+  \mathcal{W}_\textrm{eff}^\textrm{sub} &\approx \frac{1}{\Nant(\Nant-1)}\,\left[n_1(n_1-1)\,\left|\mathcal{W}^\textrm{I}_{(1)}\right|^4\right.\nonumber\\
+  &\qquad\qquad\qquad\qquad + n_2(n_2-1)\,\left|\mathcal{W}^\textrm{I}_{(1)}\right|^2\,\left|\mathcal{W}^\textrm{I}_{(2)}\right|^2 \nonumber\\
+  &\qquad\qquad\qquad\qquad \left. + n_1n_2\,\left|\mathcal{W}^\textrm{I}_{(1)}\right|^2 \,\mathcal{W}^\textrm{I}_{(1)}\,\mathcal{W}^{\textrm{I}\star}_{(2)}\right].
+\end{align}
+Finally, since no {\it a priori} information will be available about the heterogeneous array in case of the erroneous assumption, the effective weighting in the image will be believed to be $\mathcal{W}_\textrm{eff}^\textrm{err} = \left|\mathcal{W}_{(1)}^\textrm{I}\right|^4$.
+
+Estimates of the dirty images with uniform, isotropic weighting in the optimal and erroneous cases will be obtained by dividing the respective weighted images by their assumed effective weights $\mathcal{W}_\textrm{eff}^\textrm{opt}$ and $\mathcal{W}_\textrm{eff}^\textrm{err}$ respectively as:
 \begin{align}
@@
 
-If $\mathcal{I}_\textrm{M}(\hat{\mathbf{s}},f)$ is the sky model, then $\mathcal{I}_\textrm{D}^\textrm{iso}(\hat{\mathbf{s}},f)$ can be expressed as:
+If $\mathcal{I}_\textrm{M}$ is the sky model, then $\mathcal{I}_\textrm{D}^\textrm{iso}$ can be expressed as:
 \begin{align}\label{eqn:img-uncertainties}
-  \mathcal{I}_\textrm{D}^\textrm{iso}(\hat{\mathbf{s}},f) &= \mathcal{I}_\textrm{M}(\hat{\mathbf{s}},f) + \delta\mathcal{I}_\textrm{S}(\hat{\mathbf{s}},f) + \delta\mathcal{I}_\textrm{M}(\hat{\mathbf{s}},f),
+  \mathcal{I}_\textrm{D}^\textrm{iso} &= \mathcal{I}_\textrm{M} + \delta\mathcal{I}_\textrm{S} + \delta\mathcal{I}_\textrm{M},
 \end{align}
@@
 Those of the first and second types, which have roughly equal numbers of antennas with similar layouts, have 1.1~m and 6.6~m on their sides respectively.
-$\mathcal{I}_\textrm{M}(\hat{\mathbf{s}},f)$ consists of point sources of random flux densities brighter than 10~Jy along westward and south-westward directions as seen in Fig.~\ref{fig:nonphysical-image}, which shows an optimally weighted image, $\mathcal{I}^\prime(\hat{\mathbf{s}},f)$, obtained with EPIC using the MOFF algorithm after averaging over the entire bandwidth and duration of simulated observation.
+$\mathcal{I}_\textrm{M}$ consists of point sources of random flux densities brighter than 10~Jy along westward and south-westward directions as seen in Fig.~\ref{fig:nonphysical-image}, which shows an optimally weighted image, $\mathcal{I}^\prime$, obtained with EPIC using the MOFF algorithm after averaging over the entire bandwidth and duration of simulated observation.
 
@@
 \subfloat[][Estimates of normalized radial flux density profile]{\label{fig:norm-flux-density}\includegraphics[width=0.42\linewidth]{figure10b.eps}}
-\caption{(a)~Optimally weighted image, $\mathcal{I}^\prime(\hat{\mathbf{s}},f)$, of simulated sky model obtained by EPIC using MOFF algorithm after averaging over entire bandwidth and duration of simulated observation.
+\caption{(a)~Optimally weighted image, $\mathcal{I}^\prime$, of simulated sky model obtained by EPIC using MOFF algorithm after averaging over entire bandwidth and duration of simulated observation.
 51 antennas in the inner core of the MWA layout were used with uniformly illuminated square apertures.
@@
 Point sources were simulated along westward and south-westward directions with random flux densities brighter than 10~Jy.
-Black boxes around point sources are for purposes of collecting image statistics.
+Black boxes shown around point sources are for purposes of collecting image statistics.
 (b)~Normalized flux density estimates, $\widehat{\mathcal{I}}_\textrm{M}^\textrm{norm}$, as a function of distance from image center.

### 2016-08-20 imaging_demo/ms.tex
Mark changes with boldface
From 1d980f4 (mnras):
====================

In ms.tex,

(*) Use boldface to mark changes compared to previous submitted
version

@@
 We have also validated our implementation on data observed with the Long Wavelength Array (LWA).
-We have demonstrated its versatility to image with heterogeneous arrays by showing that it robustly estimates the input sky model while wrong assumptions about array homogeneity lead to systematic mis-estimates.
-Antenna layouts with dense filling factors consisting of a large number of antennas such as LWA, the Square Kilometre Array, Hydrogen Epoch of Reionization Array, and Canadian Hydrogen Intensity Mapping Experiment will gain significant computational advantage by deploying EPIC.
+{\bf We have demonstrated its versatility to image with heterogeneous arrays by showing that it robustly estimates the input sky model while wrong assumptions about array homogeneity lead to systematic mis-estimates.} Antenna layouts with dense filling factors consisting of a large number of antennas such as LWA, the Square Kilometre Array, Hydrogen Epoch of Reionization Array, and Canadian Hydrogen Intensity Mapping Experiment will gain significant computational advantage by deploying EPIC.
 Inherent availability of calibrated time-domain images on digitizer writeout time-scales and vastly lower I/O bandwidth relative to visibility-based systems will make it a prime candidate for transient searches of Fast Radio Bursts (FRB) as well as planetary and exoplanetary phenomena.
@@
 Similarly, the search for transient objects and regular monitoring of the time-dependent sky is driving instruments in the same direction with the added requirement of fast read-outs.
-A number of new telescopes are being or were developed around the world based on this new paradigm, including the Hydrogen Epoch of Reionization Array\footnote{http://reionization.org} (HERA; \citealt{deb16}), the Murchison Widefield Array (MWA; \citealt{tin13,bow13}), the Donald~C.~Backer Precision Array for Probing the Epoch of Reionization (PAPER; \citealt{par10}), the LOw Frequency ARray (LOFAR; \citealt{van13}), the Canadian Hydrogen Intensity Mapping Experiment (CHIME; \citealt{ban14}), the Long Wavelength Array (LWA; \citealt{ell13}), and the low frequency Square Kilometer Array (SKA1-Low; \citealt{mel13}).
+A number of new telescopes are being or were developed around the world based on this new paradigm, including the Hydrogen Epoch of Reionization Array\footnote{http://reionization.org} (HERA; \citealt{deb16}), the Murchison Widefield Array (MWA; \citealt{tin13,bow13}), the {\bf Donald~C.~Backer} Precision Array for Probing the Epoch of Reionization (PAPER; \citealt{par10}), the LOw Frequency ARray (LOFAR; \citealt{van13}), the Canadian Hydrogen Intensity Mapping Experiment (CHIME; \citealt{ban14}), the Long Wavelength Array (LWA; \citealt{ell13}), and the low frequency Square Kilometer Array (SKA1-Low; \citealt{mel13}).
 
@@
 We also demonstrate the performance with real-world data from the LWA.
-\S\ref{sec:versatility} demonstrates its ability to robustly image data from heterogeneous arrays.
-In \S\ref{sec:analysis}, we explore the scalability of the algorithm in the context of several array design choices.
+{\bf \S\ref{sec:versatility} demonstrates its ability to robustly image data from heterogeneous arrays.} In \S\ref{sec:analysis}, we explore the scalability of the algorithm in the context of several array design choices.
 We identify specific array design classes where the EPIC correlator -- is computationally more efficient; and in the field of transients, demands significantly lesser I/O bandwidth relative to visibility-based approaches.
@@
 where we have brought the time average into the integral under the assumption that the aperture illumination pattern does not change over the time-scale of the averaging and $\star$ denotes complex conjugation.
-The noise-like cross-terms ideally have zero mean.
-This expression can be further simplified using the sky brightness, $\mathcal{I}(\hat{\mathbf{s}},f,t)$, as $\left\langle \mathcal{E}(\hat{\mathbf{s}},f,t)\,\mathcal{E}^\star(\hat{\mathbf{s'}},f,t) \right\rangle_t = \mathcal{I}(\hat{\mathbf{s}},f,t)\,\delta(\hat{\mathbf{s}}-\hat{\mathbf{s'}})$, and defining the antenna-pair sky power response function (or the directional antenna-pair power pattern), $\mathcal{B}_{ab}(\hat{\mathbf{s}})~\equiv~\mathcal{W}_a(\hat{\mathbf{s}},f,t)~\mathcal{W}^\star_b(\hat{\mathbf{s}},f,t)$.
+{\bf The noise-like cross-terms ideally have zero mean.
+This expression can be further simplified using the sky brightness}, $\mathcal{I}(\hat{\mathbf{s}},f,t)$, as $\left\langle \mathcal{E}(\hat{\mathbf{s}},f,t)\,\mathcal{E}^\star(\hat{\mathbf{s'}},f,t) \right\rangle_t = \mathcal{I}(\hat{\mathbf{s}},f,t)\,\delta(\hat{\mathbf{s}}-\hat{\mathbf{s'}})$, and defining the antenna-pair sky power response function (or the directional antenna-pair power pattern), $\mathcal{B}_{ab}(\hat{\mathbf{s}})~\equiv~\mathcal{W}_a(\hat{\mathbf{s}},f,t)~\mathcal{W}^\star_b(\hat{\mathbf{s}},f,t)$.
 The result is the visibility expressed in terms of the sky brightness, the power pattern, and uncorrelated noise terms which we group into $n_{ab}(f,t)$.
@@
 where the measured visibilities are weighted by the inverse of the system noise, followed by a gridding process using the holographic antenna power response as the gridding kernel, followed by a Fourier transform to create an image $\boldsymbol{\mathcal{I}}^\prime(\hat{\mathbf{s}})$.
-This is the optimal estimate of the true image $\boldsymbol{\mathcal{I}}(\hat{\mathbf{s}})$ given the visibility measurements.
+This is the {\bf optimally weighted} estimate of the true image $\boldsymbol{\mathcal{I}}(\hat{\mathbf{s}})$ given the visibility measurements.
 
@@
 model of the sky. Since antennas are only correlated with calibrator pixels, the 
-computational complexity scales as $\sim N_\textrm{a} N_\textrm{c}$.
+computational complexity scales as $\sim N_\textrm{a} N_\textrm{c}$. {\bf The calibration performance of EPIC is explored for varying levels of completeness of sky model in \citet{bea16}.}
 
@@
 Both algorithms result in images and synthesized beams that are well matched with each other.
-As expected, their fluxes are attenuated by a weighting proportional to the square of the antenna power pattern corresponding to a uniform square aperture.
+{\bf As expected, their fluxes are attenuated by a weighting proportional to the square of the antenna power pattern corresponding to a uniform square aperture.}
 
@@
 The antenna auto-correlations at zero-spacing have been removed from the MOFF images.
-The images in either case reconstruct the sources at the right locations with the fluxes attenuated expected after multiplication by the antenna power pattern.
-    The synthesized beams from the two algorithms are well matched in size and
-    shape. The overall modulation by the power pattern is seen clearly in both
-    images.}
+The images in either case reconstruct the sources at the right locations with the fluxes attenuated expected after multiplication by the antenna power pattern {\bf squared}.
+The synthesized beams from the two algorithms are well matched in size and shape.
+The overall modulation by the power pattern is seen clearly in both images.}
   \label{fig:MOFF-FX-image}
@@
 
-We examine in detail the respective synthesized beams in each case in Fig.~\ref{fig:MOFF-FX-psf-diff}.
+{\bf We examine in detail the respective synthesized beams in each case in Fig.~\ref{fig:MOFF-FX-psf-diff}.
 Slices at $m=0$ of the synthesized beams weighted by the antenna power pattern are shown for MOFF method using EPIC (solid black) and visibility-based (dashed gray) imaging.
@@
 A magnified view shows that some differences at the level of $<0.5$\% exist in some regions.
-These are attributed to differences in the $uv$-plane antenna cross-correlation weights in the two methods which in turn arises due to the amount of coarseness in grid spacing as described below.
+These are attributed to differences in the $uv$-plane antenna cross-correlation weights in the two methods which in turn arises due to the amount of coarseness in grid spacing as described below.}
 
@@
   \includegraphics[width=\columnwidth]{figure6}
-  \caption{Synthesized beams weighted by antenna power pattern sliced at $m=0$ obtained with MOFF algorithm using EPIC (solid black line) and visibility-based imaging (dashed gray line).
+  \caption{{\bf Synthesized beams weighted by antenna power pattern sliced at $m=0$ obtained with MOFF algorithm using EPIC (solid black line) and visibility-based imaging (dashed gray line).
 The two appear almost identical.
 The inset provides a magnified view of differences in the synthesized beam slices between the two techniques at levels $\lesssim 0.5$\% relative to the peak.
-These are attributed to differences that arise in gridding and depends on the coarseness of the grid.}
+These are attributed to differences that arise in gridding and depends on the coarseness of the grid.}}
   \label{fig:MOFF-FX-psf-diff}
@@
 
-The left panel of Fig.~\ref{fig:MOFF-FX-uvwts-diff} shows differences (in percentage relative to peak) in cross-correlated weights obtained with MOFF imaging in EPIC and visibility-based imaging.
+{\bf The left panel of Fig.~\ref{fig:MOFF-FX-uvwts-diff} shows differences (in percentage relative to peak) in cross-correlated weights obtained with MOFF imaging in EPIC and visibility-based imaging.
 The maximum difference appears near the center of the $uv$-plane corresponding to antenna auto-correlations which have not been perfectly removed in the former.
 In other regions of the $uv$-plane, the differences are of the order of a few percent.
-The reason for these differences is described below.
+The reason for these differences is described below.}
 
@@
   % \includegraphics[width=\columnwidth]{MOFF_FX_comparison_uvwts_test_aperture_zero_spacing_removed.eps}
-  \caption{{\it Left}: Differences (in \%) in cross-correlated weights in the $uv$ plane relative to the peak.
+  \caption{{\bf {\it Left}: Differences (in \%) in cross-correlated weights in the $uv$ plane relative to the peak.
 The biggest difference ($\sim 20$\%) is found around the zero-spacing corresponding to auto-correlations of antenna weights due to gridding differences augmented by the relatively high number density of antenna auto-correlation footprints in that region.
@@
 {\it Right}: Histogram (expressed as percentage) of the relative differences (in \%) between antenna cross-correlation weights shown on the left binned in intervals of 0.5\%.
-Over 70\% of the $uv$-cells do not differ by more than 0.5\% and over 90\% of $uv$-cells only differ by $<5$\% for the gridding coarseness used.}
+Over 70\% of the $uv$-cells do not differ by more than 0.5\% and over 90\% of $uv$-cells only differ by $<5$\% for the gridding coarseness used.}}
   \label{fig:MOFF-FX-uvwts-diff}
@@
 Fig.~\ref{fig:image-psf-diff} shows the difference between the synthesized beams obtained with the two methods.
-A difference map between the two synthesized beams is shown on the left.
+A difference map between the two synthesized beams is shown on the {\bf left}.
 The amplitude of the difference appears to be modulated by the directional power response of the antenna.
-On the right, in radial bins, the rms of the synthesized beam (gray) and the rms of the difference map (black) are plotted in percentage units relative to the peak (to be read using the axis on the left side of the plot).
+{\bf On the right}, in radial bins, the rms of the synthesized beam (gray) and the rms of the difference map (black) are plotted in percentage units relative to the peak (to be read using the axis on the left side of the plot).
 The antenna power pattern (red; to be read using the scale on the right) is plotted for reference.
@@
 $\approx \sqrt{2}$ higher than the rms of the each of the synthesized beams. 
-This means that while differences exist, large fractions of them are still 
+This {\bf means} that while differences exist, large fractions of them are still 
 well matched to each other even out to the horizon. Thus the rounding errors in 
@@
 
-One of the advanced features of the MOFF algorithm and its implementation via EPIC is the ability to naturally account for heterogeneity of antennas while still producing images with maximal information \citep{mor11}.
+{\bf One of the advanced features of the MOFF algorithm and its implementation via EPIC is the ability to naturally account for heterogeneity of antennas while still producing images with maximal information \citep{mor11}.
 Unlike the assumption frequently made in standard radio interferometry that all antennas are identical, EPIC can implement the most generic case where no such assumption can be made.
@@
 This demonstrates that images from EPIC can robustly image data from heterogeneous arrays while being mathematically optimal.
-Besides being sub-optimal, erroneous assumptions about the homogeneity of an array can lead to significant and systematic mis-estimates (in this case, an underestimate) of the sky model.
+Besides being sub-optimal, erroneous assumptions about the homogeneity of an array can lead to significant and systematic mis-estimates (in this case, an underestimate) of the sky model.}
 
@@
 The different colored lines correspond to different antenna sizes (cyan - 1~m$^2$, blue - 7~m$^2$, purple - 16~m$^2$, green - 28~m$^2$, orange - 150~m$^2$, red - 740~m$^2$, gray - 5900~m$^2$).
-Dashed lines of each color denote the boundary to the left of which the MOFF algorithm is favoured for the corresponding antenna size and vice versa for visibility-based imaging.
+{\bf Dashed lines of each color denote the boundary to the left of which the MOFF algorithm is favoured for the corresponding antenna size and vice versa for visibility-based imaging.}
 
@@
 The different antennas are color coded by roughly the class of antenna size they fall into.
-Thus symbols of one color falling to the left of a line of the same color indicate MOFF imaging is advantageous for those telescopes and vice versa.
-For example, MOFF imaging is favoured in HERA-331 and HERA-6769 because they lie to the left of the orange line but not so in cases of HERA-19 and HERA-37.}
+{\bf Thus symbols of one color falling to the left of a line} of the same color indicate MOFF imaging is advantageous for those telescopes and vice versa.
+For example, MOFF imaging is favoured in HERA-331 and HERA-6769 because they {\bf lie to the left of the orange line} but not so in cases of HERA-19 and HERA-37.}
   \label{fig:parameter-space-bll-nant-instruments}
@@
 
-There is an upper limit set by the maximum number of antennas for each antenna size that can be densely packed inside various baseline lengths but is not shown on this figure to avoid crowding.
+{\bf There is an upper limit set by the maximum number of antennas for each antenna size that can be densely packed inside various baseline lengths but is not shown on this figure to avoid crowding.
 We note that as antenna size increases the maximum number of antennas for a dense packing as a function of baseline length decreases.
@@
 Similarly, with increase in antenna size, $\Ngrid$ also decreases when field of view imaging is achieved with an increased grid spacing equal to antenna size and hence lowers the amount of computations required with the MOFF algorithm.
-This shifts the dashed lines rightward as well.
-
-The different antennas are color coded by roughly the class of antenna size they fall into.
+This shifts the dashed lines rightward as well.}
+
+{\bf The different antennas are color coded by roughly the class of antenna size they fall into.
 Thus symbols of one color that lie to the left of same colored line indicate MOFF imaging is favoured for those telescopes and vice versa.
-For e.g., MOFF imaging is favoured in HERA-331 and HERA-6769 because they lie to the left of the orange line but not so in cases of HERA-19 and HERA-37.
+For e.g., MOFF imaging is favoured in HERA-331 and HERA-6769 because they lie to the left of the orange line but not so in cases of HERA-19 and HERA-37}.
 A majority of the next-generation radio telescopes, namely, HERA-331 and its future expanded versions, SKA1-LC, SKA1-LCD, HIRAX, and CHIME will fall in the regime where MOFF imaging will be desirable.
@@
 Under the class of direct imaging techniques, ours is one of the most generic -- neither does it place any constraint on the array layout to be on a regular grid nor does it require the antenna array to be homogeneous.
-We have demonstrated its natural ability to robustly image data from heterogeneous arrays and shown that wrong assumptions about array homogeneity result in severe and systematic mis-estimation of the sky model.
+{\bf We have demonstrated its natural ability to robustly image data from heterogeneous arrays and shown that wrong assumptions about array homogeneity could result in significant and systematic mis-estimation of the sky model.}
 
@@
 successfully tested on simulated MWA observations as well as real LWA 
-observations from both imaging and calibration view points.
+observations {\bf from both imaging and calibration view points}.
 
@@
 
-In the near future, we plan to demonstrate its speed and precision by upgrading EPIC to a GPU-based implementation in order to operate on real-time data and develop a transient trigger and monitor backend.
-In the meanwhile, we plan to demonstrate imaging with non-coplanar arrays and direction-dependent calibration.
+In the near future, we plan to {\bf demonstrate speed and precision by upgrading EPIC to a GPU-based implementation} in order to operate on real-time data and develop a transient trigger and monitor backend.
+{\bf In the meanwhile, we plan to demonstrate imaging with non-coplanar arrays and direction-dependent calibration.}
 

### 2016-08-22 imaging_demo/ms.tex
Minor changes
From 5c6f2b0 (mnras):
====================

In ms.tex,

(*) Minor changes throughout

In epic.bib,

(*) Add references to holography papers

@@
 This is the software holographic equivalent of a traditional FX correlator output.
+{\bf For context, the term ``holography'' is derived from holographic measurements of the complex antenna aperture illumination pattern \citep{ben76,sco77}.
+In the present context, ``software holography'' refers to the application of this antenna aperture illumination pattern in software for imaging.}
 
@@
 
-We have implemented the MOFF imaging technique in our ``E-field Parallel Imaging
-Correlator'' -- a highly parallelized Object Oriented Python 
-package,\footnote{The E-field Parallel Imaging Correlator (EPIC) package can be 
-accessed at https://github.com/nithyanandan/EPIC} now publicly available. Besides 
-implementing the MOFF imaging algorithm it also includes visibility-based imaging 
-using the software holography technique and a simulator for generating electric 
-fields from a sky model.
-
-EPIC can accept dual-polarization inputs and produce images of all four 
-instrumental cross-polarizations. Currently two data input formats exist for 
-reading in the 
-electric field time samples measured by the antennas -- simulated electric 
-fields based on a sky model using the simulator packaged with EPIC; and LWA 
-data. Efforts to build interfaces for data from other telescopes are underway.
+We have implemented the MOFF imaging technique in our ``E-field Parallel Imaging Correlator'' -- a highly parallelized Object Oriented Python package,\footnote{The E-field Parallel Imaging Correlator (EPIC) package can be accessed at https://github.com/nithyanandan/EPIC} now publicly available.
+Besides implementing the MOFF imaging algorithm it also includes {\bf a simulator for generating electric fields from a sky model and a visibility-based imaging using the software holography technique.}
+
+EPIC can accept dual-polarization inputs and produce images of all four instrumental cross-polarizations.
+Currently two data input formats exist for reading in the electric field time samples measured by the antennas -- simulated electric fields based on a sky model using the simulator packaged with EPIC; and LWA data.
+Efforts to build interfaces for data from other telescopes are underway.
 
@@
   \textrm{where}, \quad   n_{ij} &= \begin{cases}
-    n_i\,n_j & i\ne i \\
-    n_i(n_i-1) & i=j, 
+    n_i\,n_j, & i\ne i \\
+    n_i(n_i-1), & i=j, 
   \end{cases}
@@
 \end{align}
-Here, equation~\ref{eqn:wt-dirty-image-EPIC} has been decomposed as the sum of fringes over all unique antenna pairs in an antenna pair type (innermost summation) and subsequently summed over all antenna pair types.
-
-{\bf For purposes of demonstration in this study, we make the following simplification.} We assume that the spatial distribution of antenna pairs in each antenna pair type are similar and thus result in a similar dirty image, $\mathcal{I}_\textrm{D}^\textrm{iso}$.
+Here, equation~\ref{eqn:wt-dirty-image-EPIC} has been decomposed as the sum of fringes over all unique antenna pairs in an antenna pair type (inner sum) and subsequently summed over all antenna pair types (outer sum).
+
+For purposes of demonstration in this study, we make the following simplification.
+We assume that the spatial distribution of antenna pairs in each antenna pair type are similar and thus result in a similar dirty image, $\mathcal{I}_\textrm{D}^\textrm{iso}$.
 This usually holds when $\Nant$ is large and antennas under each antenna type are chosen randomly such that differences in the point spread functions are insignificant.
@@
 It will be sub-optimal with effective weighting:
-\begin{align}\label{eqn:effective-weighting-optimal-example}
+\begin{align}\label{eqn:effective-weighting-suboptimal-example}
   \mathcal{W}_\textrm{eff}^\textrm{sub} &\approx \frac{1}{\Nant(\Nant-1)}\,\left[n_1(n_1-1)\,\left|\mathcal{W}^\textrm{I}_{(1)}\right|^4\right.\nonumber\\
@@
 \subfloat[][Estimates of normalized radial flux density profile]{\label{fig:norm-flux-density}\includegraphics[width=0.42\linewidth]{figure10b.eps}}
-\caption{(a)~Optimally weighted image, $\mathcal{I}^\prime$, of simulated sky model obtained by EPIC using MOFF algorithm after averaging over entire bandwidth and duration of simulated observation.
+\caption{{\bf (a)~Optimally weighted image, $\mathcal{I}^\prime$, of simulated sky model obtained by EPIC using MOFF algorithm after averaging over entire bandwidth and duration of simulated observation.
 51 antennas in the inner core of the MWA layout were used with uniformly illuminated square apertures.
@@
 In the optimal case, the normalized estimates are consistent with unity to within ``$1\sigma$'' in almost all regions of the image, whereas in the erroneous case, they are mis-estimated (underestimated) with high significance ($\gtrsim 5\sigma$) in most regions of the image.
-There is no significant difference in estimates between the two different directions of point sources studied.}
+There is no significant difference in estimates between the two different directions of point sources studied.}}
 \label{fig:versatility}
@@
     \hline
-    Telescope & Core & Number & Antenna & Frequency \\
-              & size & of Antennas & size & \\
-              & $b_\textrm{max}$ (in m) & $\Nant$ & $A_\textrm{a}$ (in m$^2$) & $f_0$ (in MHz) \\
+    Telescope & Core & Number of & Antenna & Frequency \\
+              & size & Antennas & size & \\
+              & $b_\textrm{max}$ (m) & $\Nant$ & $A_\textrm{a}$ (m$^2$) & $f_0$ (MHz) \\
     \hline

### 2016-08-24 imaging_demo/ms.tex
Shorten section 5, add appendix for detailed math
From 601fdcb (Judd-v1):
======================

In ms.tex,

(*) Shorten section 5 and add appendix for detailed math

@@
 A high level software architecture of the EPIC package is described in the 
-appendix \S\ref{sec:software-modules} for the interested reader.
+appendix~\ref{sec:software-modules} for the interested reader.
 
@@
 {\bf One of the advanced features of the MOFF algorithm and its implementation via EPIC is the ability to naturally account for heterogeneity of antennas while still producing images with maximal information \citep{mor11}.
-Unlike the assumption frequently made in standard radio interferometry that all antennas are identical, EPIC can implement the most generic case where no such assumption can be made.
+Unlike the assumption frequently made in existing direct imaging techniques that all antennas are identical, EPIC can implement the most generic case where no such assumption can be made.
 In other words, EPIC does not assume $W_a(\mathbf{r}-\mathbf{r}_a)$ or $B_{ab}(\mathbf{r}-\mathbf{r}_{ab})$ for different antennas are identical.
 However, any {\it a priori} knowledge available that certain sets of antennas have identical illumination patterns can be easily passed on to EPIC to avoid redundant computations and thus increase efficiency.
-
-We present a methodology to understand the effective angular weighting in the image obtained with EPIC from a heterogeneous array and compare it with that from standard interferometry which assumes a homogeneous array (all antennas are identical).
-We also demonstrate how the assumption in the latter results in a mis-estimation of the recovered sky model.
-
-Dropping the noise term and time dependence, the visibility measured by an antenna pair can be written from equation~\ref{eqn:measurement-eqn-1} as:
-\begin{align}\label{eqn:measurement-eqn-2}
-  V_{ab}(f) &= \int \mathcal{W}^\textrm{I}_a(\hat{\mathbf{s}},f)\mathcal{W}^{\textrm{I}\star}_b(\hat{\mathbf{s}},f)\,\mathcal{I}(\hat{\mathbf{s}},f)\,e^{-2\pi i f\mathbf{r}_{ab}\!\cdot\,\hat{\mathbf{s}}/c}\,\dif\Omega,
-\end{align}
-where, superscript `$\textrm{I}$' in the $\mathcal{W}$ term indicates the weighting is inherently introduced by the instrument during the measurement process. The beam-weighted dirty image is:
-\begin{align}
-  \mathcal{I}_\textrm{D}(\hat{\mathbf{s}},f) &= \frac{1}{\Nant(\Nant-1)/2}\,\sum_{ab}\, V_{ab}(f)\,e^{2\pi i f\mathbf{r}_{ab}\!\cdot\,\hat{\mathbf{s}}/c},
-\end{align}
-{\bf where, we have assumed equal weighting for each antenna pair}.
-
-When the array is homogeneous, $\mathcal{W}^\textrm{I} \equiv \mathcal{W}^\textrm{I}_a \equiv \mathcal{W}^\textrm{I}_b$ and the above equation reduces to:
-\begin{align}\label{eqn:dirty-image-homogeneous}
-  \mathcal{I}_\textrm{D}(\hat{\mathbf{s}},f) &= |\mathcal{W}^\textrm{I}(\hat{\mathbf{s}},f)|^2\,\mathcal{I}_\textrm{D}^\textrm{iso}(\hat{\mathbf{s}},f),
-\end{align}
-where, $\mathcal{I}_\textrm{D}^\textrm{iso}(\hat{\mathbf{s}},f)$ is the dirty image with no beam weighting (isotropic, uniform weighting {\bf of the sky}) determined only by the array layout and $|\mathcal{W}^\textrm{I}|^2$ is the directional power pattern of the antenna pair familiar in standard interferometry.
-
-{\bf In order to understand an image from a heterogeneous array, we start by looking at the contribution to the image from the visibility, $V_{ab}$, from each antenna pair.} While imaging, EPIC introduces a weighting to each antenna during gridding as given by equations~\ref{eqn:dirty-image-FX} and \ref{eqn:e-field-conv}, and equivalently, the weighted visibility from {\bf an} antenna pair $ab$ projected on the grid is:
+Here we present key highlights from the generic methodology detailed in appendix~\ref{sec:math-versatility} to understand the effective angular weighting in the image obtained with EPIC from a heterogeneous array.
+
+We look at the contribution to the image from the visibility, $V_{ab}$, from each antenna pair.
+While imaging, EPIC introduces a weighting to each antenna during gridding as given by equations~\ref{eqn:dirty-image-FX} and \ref{eqn:e-field-conv}, and equivalently, the weighted visibility from {\bf an} antenna pair $ab$ projected on the grid is:
 \begin{align}
@@
 \end{align}
-where, $B^\textrm{G}_{ab}(\mathbf{r}-\mathbf{r}_{ab},f)$ is obtained by the spatial cross-correlation of weighting kernels associated with the individual antennas using the definition:
-\begin{align}
-  B^\textrm{G}_{ab}(\mathbf{r}-\mathbf{r}_{ab},f) &= \int W^\textrm{G}_a(\mathbf{r}^\prime-\mathbf{r}_a)\,W^{\textrm{G}\star}_b(\mathbf{r}^\prime+\mathbf{r}-\mathbf{r}_b)\,\dif^2\mathbf{r}^\prime.
-\end{align}
+where, $B^\textrm{G}_{ab}(\mathbf{r}-\mathbf{r}_{ab},f)$ is obtained by the spatial cross-correlation of weighting kernels associated with the individual antennas.
 The superscript `$\textrm{G}$' denotes the weighting introduced in analysis during the gridding process.
+We will use superscript `I' to denote the inherent weighting applied by the instrument during the measurement process in obtaining $V_{ab}$ from the true sky model.
 Optimal imaging requires $W^\textrm{G}_a = W^\textrm{I}_a$ \citep{mor09,mor11}.
-However, we keep the two superscripts separately to describe output images under the wrong assumption that the array is homogeneous when instead it is actually heterogeneous.
-
-The sky response of the weighted visibility is:
-\begin{align}
-  \mathcal{I}^\prime_{ab}(\hat{\mathbf{s}},f) &= \int V^\prime_{ab}(\mathbf{r},f)\,e^{2\pi i f\mathbf{r}\cdot\,\hat{\mathbf{s}}/c}\,\dif^2\mathbf{r} \nonumber\\
-                                              &= \int B^{\textrm{G}\star}_{ab}(\mathbf{r}-\mathbf{r}_{ab},f)\,V_{ab}\,e^{2\pi i f\mathbf{r}\cdot\,\hat{\mathbf{s}}/c}\,\dif^2\mathbf{r} \nonumber\\
-                                              &= V_{ab}\,\int \left[\int W^{\textrm{G}\star}_a(\mathbf{r}^\prime-\mathbf{r}_a)\,W^\textrm{G}_b(\mathbf{r}^\prime+\mathbf{r}-\mathbf{r}_a)\,\dif^2\mathbf{r}^\prime\right] \nonumber\\
-                                              &\qquad\qquad \times e^{2\pi i f\mathbf{r}\cdot\,\hat{\mathbf{s}}/c}\,\dif^2\mathbf{r} \nonumber\\
-                                              &= \mathcal{W}^{\textrm{G}\star}_a(\hat{\mathbf{s}},f)\mathcal{W}^\textrm{G}_b(\hat{\mathbf{s}},f)\,V_{ab}\,e^{2\pi i f\mathbf{r}_{ab}\!\cdot\,\hat{\mathbf{s}}/c},
-\end{align}
-which denotes the fringe from the weighted visibility. The image output of EPIC is equivalent to averaging these weighted fringes from all antenna pairs:
+However, we keep the two superscripts separately to describe output images under the wrong assumption that the array is homogeneous when instead it is actually not.
+
+The sky response of the weighted visibility, referred to as {\it fringes}, is:
+\begin{align}
+  \mathcal{I}^\prime_{ab}(\hat{\mathbf{s}},f) &= \mathcal{W}^{\textrm{G}\star}_a(\hat{\mathbf{s}},f)\mathcal{W}^\textrm{G}_b(\hat{\mathbf{s}},f)\,V_{ab}\,e^{2\pi i f\mathbf{r}_{ab}\!\cdot\,\hat{\mathbf{s}}/c}.
+\end{align}
+Since all quantities in the celestial plane (calligraphic fonts) are functions of position, $\hat{\mathbf{s}}$, and frequency, $f$, we drop writing this dependence explicitly hereafter.
+
+The image output of EPIC is the average of these weighted fringes from all antenna pairs:
 \begin{align}\label{eqn:wt-dirty-image-EPIC}
-  \mathcal{I}^\prime(\hat{\mathbf{s}},f) &= \frac{1}{\Nant(\Nant-1)/2}\,\sum_{\substack{ab\\a\ne b}}\,\mathcal{I}^\prime_{ab}(\hat{\mathbf{s}},f) \nonumber\\
-                                         &= \frac{1}{\Nant(\Nant-1)/2}\,\sum_{\substack{ab\\a\ne b}}\,\left[\mathcal{W}^{\textrm{G}\star}_a(\hat{\mathbf{s}},f)\mathcal{W}^\textrm{G}_b(\hat{\mathbf{s}},f)\,V_{ab}\right.
-\nonumber\\
-  &\qquad\qquad\qquad\qquad\qquad\qquad \times \left. e^{2\pi i f\mathbf{r}_{ab}\!\cdot\,\hat{\mathbf{s}}/c}\right].
-\end{align}
-
-If all antennas are identically weighted ($W^\textrm{G} \equiv W^\textrm{G}_a \equiv W^\textrm{G}_b$), using equation~\ref{eqn:dirty-image-homogeneous}, equation~\ref{eqn:wt-dirty-image-EPIC} reduces to:
+  \mathcal{I}^\prime &= \frac{1}{\Nant(\Nant-1)}\,\sum_{\substack{ab\\a\ne b}}\mathcal{I}^\prime_{ab}
+\end{align}
+
+We define the effective attenuation of the dirty image due to instrumental and gridding weights as:
+\begin{align}\label{eqn:effective-weighting}
+  \mathcal{W}_\textrm{eff} &= \mathcal{I}^\prime\, / \,\mathcal{I}_\textrm{D}^\textrm{iso},
+\end{align}
+where, $\mathcal{I}_\textrm{D}^\textrm{iso}$ is the dirty image with no beam weighting (isotropic, uniform weighting of the sky) determined only by the array layout.
+
+For a homogeneous array, $W^\textrm{G} \equiv W^\textrm{G}_a \equiv W^\textrm{G}_b$. Thus,  
 \begin{align}\label{eqn:wt-dirty-image-homogeneous}
-  \mathcal{I}^\prime(\hat{\mathbf{s}},f) &= \frac{\left|\mathcal{W}^\textrm{G}(\hat{\mathbf{s}},f)\right|^2}{\Nant(\Nant-1)/2}\,\sum_{\substack{ab\\a\ne b}}\,V_{ab}\,e^{2\pi i f\mathbf{r}_{ab}\!\cdot\,\hat{\mathbf{s}}/c} \nonumber\\
-  &= \left|\mathcal{W}^\textrm{G}(\hat{\mathbf{s}},f)\right|^2\,\mathcal{I}_\textrm{D}(\hat{\mathbf{s}},f) \nonumber\\
-  &= \left|\mathcal{W}^\textrm{G}(\hat{\mathbf{s}},f)\right|^2\,\left|\mathcal{W}^\textrm{I}(\hat{\mathbf{s}},f)\right|^2\,\mathcal{I}_\textrm{D}^\textrm{iso}(\hat{\mathbf{s}},f).
-\end{align}
-Thus, the dirty image, which is already attenuated by the instrumental power pattern, gets further attenuated by the power pattern introduced in the gridding step in EPIC imager.
-This is consistent with \citet{mor09}.
-
-{\bf We note that all quantities in the celestial plane (calligraphic fonts) are functions of position, $\hat{\mathbf{s}}$, and frequency, $f$.
-For convenience, we drop writing this dependence explicitly hereafter.}
-
-{\bf For a heterogeneous array, it is necessary to keep track of antennas and antenna pairs of different types which will determine the final weighting in the synthesized image.} We consider a heterogeneous array consisting of a total of $\Nant$ antennas under $N_\textrm{T}$ different types.
-And under each antenna type $i$, there are $n_i$ antennas such that $\sum_i n_i = \Nant$.
-The total number of unique non-zero spacing antenna pairs is $\Nant(\Nant-1)/2$ and there are potentially up to $N_\textrm{T}(N_\textrm{T}+1)/2$ unique antenna pair types obtained by pairwise combination of the $N_\textrm{T}$ antenna types.
-Thus, the total number of antenna pairs,
-\begin{align}\label{eqn:total-baselines}
-  \frac{1}{2}\Nant(\Nant-1) &= \frac{1}{2}\,\sum_{ij} n_{ij}, \\
-  \textrm{where}, \quad   n_{ij} &= \begin{cases}
-    n_i\,n_j, & i\ne i \\
-    n_i(n_i-1), & i=j, 
-  \end{cases}
-\end{align}
-is obtained by counting antenna pairs, $n_{ij}$, in different antenna pair types, $ij$.
-
-Since equation~\ref{eqn:wt-dirty-image-EPIC} is obtained by averaging weighted fringes over all unique antenna pairs, it can be expressed as: 
+  \mathcal{I}^\prime &= \left|\mathcal{W}^\textrm{G}\right|^2\,\left|\mathcal{W}^\textrm{I}\right|^2\,\mathcal{I}_\textrm{D}^\textrm{iso}.
+\end{align}
+Consistent with \citet{mor09}, the dirty image, which is already attenuated by the instrumental power pattern inherent in the measurement process, gets further attenuated by the power pattern introduced in the gridding step in EPIC imager.
+
+For a heterogeneous array, 
 \begin{align}\label{eqn:wt-dirty-image-EPIC-decomp}
-  \mathcal{I}^\prime &= \frac{1/2}{\Nant(\Nant-1)/2}\,\sum_{ij}\sum_{\substack{ab\\a\ne b\\a\in i, b\in j}}\,\mathcal{I}^\prime_{ab}
-\end{align}
-Here, equation~\ref{eqn:wt-dirty-image-EPIC} has been decomposed as the sum of fringes over all unique antenna pairs in an antenna pair type (inner sum) and subsequently summed over all antenna pair types (outer sum).
-
-For purposes of demonstration in this study, we make the following simplification.
-We assume that the spatial distribution of antenna pairs in each antenna pair type are similar and thus result in a similar dirty image, $\mathcal{I}_\textrm{D}^\textrm{iso}$.
-This usually holds when $\Nant$ is large and antennas under each antenna type are chosen randomly such that differences in the point spread functions are insignificant.
-Under this assumption, equation~\ref{eqn:wt-dirty-image-EPIC-decomp} can be written as:
+  \mathcal{I}^\prime &= \frac{1}{\Nant(\Nant-1)}\,\sum_{ij}\sum_{\substack{ab\\a\ne b\\a\in i, b\in j}}\,\mathcal{I}^\prime_{ab},
+\end{align}
+where, $ab$ indexes pairs of antennas $a$ and $b$, $ij$ indexes types of pairs of antennas, with $i$ and $j$ indexing antenna types.
+Thus, $\mathcal{I}^\prime$ is written as the sum of fringes over all unique antenna pairs in an antenna pair type (inner sum) and subsequently summed over all antenna pair types (outer sum).
+
+For purposes of demonstration in this study, we assume that the spatial distribution of antenna pairs in each antenna pair type is similar and thus results in a similar dirty image, $\mathcal{I}_\textrm{D}^\textrm{iso}$.
+This usually holds when antennas under each antenna type are chosen randomly or identically such that differences in the synthesized beams are insignificant.
+Then, equation~\ref{eqn:wt-dirty-image-EPIC-decomp} can be approximated as:
 \begin{align}\label{eqn:wt-dirty-image-EPIC-decomp-approx1}
@@
 \end{align}
-where, $\mathcal{I}_\textrm{D}^\textrm{iso}$ is approximately the same for all terms and has been pulled outside the summations.
+where, $\mathcal{I}_\textrm{D}^\textrm{iso}$ is the same for all terms and has been pulled out of the summations.
 For all antennas indexed by $a$ that belong to a particular type $i$, we replace $\mathcal{W}_a$ with $\mathcal{W}_{(i)}$ {\bf and is applicable to those arising} from both instrumental (superscript `$I$') and gridding (superscript `$G$') origins.
-Using arguments similar to those in equation~\ref{eqn:total-baselines}, equation~\ref{eqn:wt-dirty-image-EPIC-decomp-approx1} can be further simplified to:
+If $n_{ij}$ is the number of antenna pairs in the antenna pair type $ij$, $\mathcal{W}_\textrm{eff}$ can be simplified as:
 \begin{align}\label{eqn:wt-dirty-image-EPIC-decomp-approx2}
-  \mathcal{I}^\prime &\approx \frac{\mathcal{I}_\textrm{D}^\textrm{iso}}{\Nant(\Nant-1)}\,\sum_{ij}n_{ij}\,\mathcal{W}^{\textrm{G}\star}_{(i)}\,\mathcal{W}^\textrm{G}_{(j)}\,\mathcal{W}^\textrm{I}_{(i)}\,\mathcal{W}^{\textrm{I}\star}_{(j)}.
-\end{align}
-Then the effective attenuation of the dirty image due to instrumental and gridding weights is:
-\begin{align}\label{eqn:effective-weighting}
-  \mathcal{W}_\textrm{eff} &= \mathcal{I}^\prime\, / \,\mathcal{I}_\textrm{D}^\textrm{iso}.
-\end{align}
-This means that the effective attenuation of the dirty image is obtained by combinations of the fourth power of antenna voltage patterns each weighted by the number of antenna pairs in those antenna pair types.
-In an optimally weighted image, $\mathcal{W}^\textrm{G}_{(i)} \equiv \mathcal{W}^\textrm{I}_{(i)}$ and thus, equation~\ref{eqn:effective-weighting} reduces to:
+  \mathcal{W}_\textrm{eff} &\approx \frac{1}{\Nant(\Nant-1)}\,\sum_{ij}n_{ij}\,\mathcal{W}^{\textrm{G}\star}_{(i)}\,\mathcal{W}^\textrm{G}_{(j)}\,\mathcal{W}^\textrm{I}_{(i)}\,\mathcal{W}^{\textrm{I}\star}_{(j)}.
+\end{align}
+In an optimally weighted image, $\mathcal{W}^\textrm{G}_{(i)} \equiv \mathcal{W}^\textrm{I}_{(i)}$. Thus,
 \begin{align}\label{eqn:effective-weighting-optimal}
@@
 \end{align}
-where the superscript `$\textrm{opt}$' denotes `optimal'.
+where, superscript `$\textrm{opt}$' denotes `optimal'.
 
@@
 
+\section{Mathematical Framework for Imaging with Heterogeneous Arrays}\label{sec:math-versatility}
+
+{\bf We present a methodology to understand the effective angular weighting in the image obtained with EPIC from a heterogeneous array and compare it with that from existing imaging applications which assume a homogeneous array (all antennas are identical).
+% We also demonstrate how the assumption in the latter results in a mis-estimation of the recovered sky model.
+
+Dropping the noise term and time dependence, the visibility measured by an antenna pair can be written from equation~\ref{eqn:measurement-eqn-1} as:
+\begin{align}\label{eqn:apndx-measurement-eqn-2}
+  V_{ab}(f) &= \int \mathcal{W}^\textrm{I}_a(\hat{\mathbf{s}},f)\mathcal{W}^{\textrm{I}\star}_b(\hat{\mathbf{s}},f)\,\mathcal{I}(\hat{\mathbf{s}},f)\,e^{-2\pi i f\mathbf{r}_{ab}\!\cdot\,\hat{\mathbf{s}}/c}\,\dif\Omega,
+\end{align}
+where, superscript `$\textrm{I}$' in the $\mathcal{W}$ term indicates the weighting is inherently introduced by the instrument during the measurement process. The beam-weighted dirty image is:
+\begin{align}
+  \mathcal{I}_\textrm{D}(\hat{\mathbf{s}},f) &= \frac{1}{\Nant(\Nant-1)}\,\sum_{ab}\, V_{ab}(f)\,e^{2\pi i f\mathbf{r}_{ab}\!\cdot\,\hat{\mathbf{s}}/c},
+\end{align}
+where, $N_A$ is the number of antennas in the array and we have assumed equal weighting for each antenna pair.
+
+When the array is homogeneous, $\mathcal{W}^\textrm{I} \equiv \mathcal{W}^\textrm{I}_a \equiv \mathcal{W}^\textrm{I}_b$, and the above equation reduces to:
+\begin{align}\label{eqn:apndx-dirty-image-homogeneous}
+  \mathcal{I}_\textrm{D}(\hat{\mathbf{s}},f) &= |\mathcal{W}^\textrm{I}(\hat{\mathbf{s}},f)|^2\,\mathcal{I}_\textrm{D}^\textrm{iso}(\hat{\mathbf{s}},f),
+\end{align}
+where, $\mathcal{I}_\textrm{D}^\textrm{iso}(\hat{\mathbf{s}},f)$ is the dirty image with no beam weighting (isotropic, uniform weighting of the sky) determined only by the array layout and $|\mathcal{W}^\textrm{I}|^2$ is the directional power pattern of the antenna pair familiar in standard interferometry.
+
+In order to understand an image from a heterogeneous array, we start by looking at the contribution to the image from the visibility, $V_{ab}$, from each antenna pair.
+While imaging, EPIC introduces a weighting to each antenna during gridding as given by equations~\ref{eqn:dirty-image-FX} and \ref{eqn:e-field-conv}, and equivalently, the weighted visibility from an antenna pair $ab$ projected on the grid is:
+\begin{align}
+  V^\prime_{ab}(\mathbf{r},f) &= B^{\textrm{G}\star}_{ab}(\mathbf{r}-\mathbf{r}_{ab},f)\,V_{ab},
+\end{align}
+where, $B^\textrm{G}_{ab}(\mathbf{r}-\mathbf{r}_{ab},f)$ is obtained by the spatial cross-correlation of weighting kernels associated with the individual antennas using the definition:
+\begin{align}
+  B^\textrm{G}_{ab}(\mathbf{r}-\mathbf{r}_{ab},f) &= \int \left[ W^{\textrm{G}\star}_b(\mathbf{r}^\prime+\mathbf{r}-\mathbf{r}_b,f)\right. \nonumber\\
+  &\qquad\qquad\qquad \left.\times\, W^\textrm{G}_a(\mathbf{r}^\prime-\mathbf{r}_a,f)\right]\,\dif^2\mathbf{r}^\prime.
+\end{align}
+The superscript `$\textrm{G}$' denotes the weighting introduced in analysis during the gridding process.
+Optimal imaging requires $W^\textrm{G}_a = W^\textrm{I}_a$ \citep{mor09,mor11}.
+However, we keep the two superscripts separately to describe output images under the wrong assumption that the array is homogeneous when instead it is actually heterogeneous.
+
+The sky response of the weighted visibility is:
+\begin{align}
+  \mathcal{I}^\prime_{ab}(\hat{\mathbf{s}},f) &= \int V^\prime_{ab}(\mathbf{r},f)\,e^{2\pi i f\mathbf{r}\cdot\,\hat{\mathbf{s}}/c}\,\dif^2\mathbf{r} \nonumber\\
+                                              &= \int B^{\textrm{G}\star}_{ab}(\mathbf{r}-\mathbf{r}_{ab},f)\,V_{ab}\,e^{2\pi i f\mathbf{r}\cdot\,\hat{\mathbf{s}}/c}\,\dif^2\mathbf{r} \nonumber\\
+                                              &= V_{ab}\,\int \left[\int W^\textrm{G}_b(\mathbf{r}^\prime+\mathbf{r}-\mathbf{r}_a,f)\right. \nonumber\\
+                                              &\qquad\qquad\,\,\, \left.\times\, W^{\textrm{G}\star}_a(\mathbf{r}^\prime-\mathbf{r}_a,f)\,\dif^2\mathbf{r}^\prime\right]\,e^{2\pi i f\mathbf{r}\cdot\,\hat{\mathbf{s}}/c}\,\dif^2\mathbf{r} \nonumber\\
+                                              % &\qquad\qquad \times e^{2\pi i f\mathbf{r}\cdot\,\hat{\mathbf{s}}/c}\,\dif^2\mathbf{r} \nonumber\\
+                                              &= \mathcal{W}^{\textrm{G}\star}_a(\hat{\mathbf{s}},f)\mathcal{W}^\textrm{G}_b(\hat{\mathbf{s}},f)\,V_{ab}\,e^{2\pi i f\mathbf{r}_{ab}\!\cdot\,\hat{\mathbf{s}}/c},
+\end{align}
+which denotes the fringe from the weighted visibility. The image output of EPIC is equivalent to averaging these weighted fringes from all antenna pairs:
+\begin{align}\label{eqn:apndx-wt-dirty-image-EPIC}
+  \mathcal{I}^\prime(\hat{\mathbf{s}},f) &= \frac{1}{\Nant(\Nant-1)}\,\sum_{\substack{ab\\a\ne b}}\,\mathcal{I}^\prime_{ab}(\hat{\mathbf{s}},f) \nonumber\\
+                                         &= \frac{1}{\Nant(\Nant-1)}\,\sum_{\substack{ab\\a\ne b}}\,\left[\mathcal{W}^{\textrm{G}\star}_a(\hat{\mathbf{s}},f)\mathcal{W}^\textrm{G}_b(\hat{\mathbf{s}},f)\right.
+\nonumber\\
+  &\qquad\qquad\qquad\qquad\qquad\qquad \times \left. V_{ab}\,e^{2\pi i f\mathbf{r}_{ab}\!\cdot\,\hat{\mathbf{s}}/c}\right].
+\end{align}
+
+If all antennas are identically weighted ($W^\textrm{G} \equiv W^\textrm{G}_a \equiv W^\textrm{G}_b$), using equation~\ref{eqn:apndx-dirty-image-homogeneous}, equation~\ref{eqn:apndx-wt-dirty-image-EPIC} reduces to:
+\begin{align}\label{eqn:apndx-wt-dirty-image-homogeneous}
+  \mathcal{I}^\prime(\hat{\mathbf{s}},f) &= \frac{\left|\mathcal{W}^\textrm{G}(\hat{\mathbf{s}},f)\right|^2}{\Nant(\Nant-1)}\,\sum_{\substack{ab\\a\ne b}}\,V_{ab}\,e^{2\pi i f\mathbf{r}_{ab}\!\cdot\,\hat{\mathbf{s}}/c} \nonumber\\
+  &= \left|\mathcal{W}^\textrm{G}(\hat{\mathbf{s}},f)\right|^2\,\mathcal{I}_\textrm{D}(\hat{\mathbf{s}},f) \nonumber\\
+  &= \left|\mathcal{W}^\textrm{G}(\hat{\mathbf{s}},f)\right|^2\,\left|\mathcal{W}^\textrm{I}(\hat{\mathbf{s}},f)\right|^2\,\mathcal{I}_\textrm{D}^\textrm{iso}(\hat{\mathbf{s}},f).
+\end{align}
+Thus, the dirty image, which is already attenuated by the instrumental power pattern, gets further attenuated by the power pattern introduced in the gridding step in EPIC imager.
+This is consistent with \citet{mor09}.
+
+We note that all quantities in the celestial plane (calligraphic fonts) are functions of position, $\hat{\mathbf{s}}$, and frequency, $f$.
+For convenience, we drop writing this dependence explicitly hereafter.
+
+For a heterogeneous array, it is necessary to keep track of antennas and antenna pairs of different types which will determine the final weighting in the synthesized image.
+We consider a heterogeneous array consisting of a total of $\Nant$ antennas under $N_\textrm{T}$ different types.
+And under each antenna type $i$, there are $n_i$ antennas such that $\sum_i n_i = \Nant$.
+The total number of unique non-zero spacing antenna pairs is $\Nant(\Nant-1)/2$ and there are potentially up to $N_\textrm{T}(N_\textrm{T}+1)/2$ unique antenna pair types obtained by pairwise combination of the $N_\textrm{T}$ antenna types.
+The total number of antenna pairs,
+\begin{align}\label{eqn:apndx-total-baselines}
+  \Nant(\Nant-1) &= \sum_{ij} n_{ij}, \\
+  \textrm{where}, \quad   n_{ij} &= \begin{cases}
+    n_i\,n_j, & i\ne j \\
+    n_i(n_i-1), & i=j, 
+  \end{cases}
+\end{align}
+is obtained by counting antenna pairs, $n_{ij}$, in different antenna pair types, $ij$.
+
+Since equation~\ref{eqn:apndx-wt-dirty-image-EPIC} is obtained by averaging weighted fringes over all antenna pairs, it can be expressed as: 
+\begin{align}\label{eqn:apndx-wt-dirty-image-EPIC-decomp}
+  \mathcal{I}^\prime &= \frac{1}{\Nant(\Nant-1)}\,\sum_{ij}\sum_{\substack{ab\\a\ne b\\a\in i, b\in j}}\,\mathcal{I}^\prime_{ab}
+\end{align}
+Here, equation~\ref{eqn:wt-dirty-image-EPIC} has been re-written as the sum of fringes over all antenna pairs in an antenna pair type (inner sum) and subsequently summed over all antenna pair types (outer sum).
+The effective attenuation of the dirty image due to instrumental and gridding weights is:
+\begin{align}\label{eqn:apndx-effective-weighting}
+  \mathcal{W}_\textrm{eff} &= \mathcal{I}^\prime\, / \,\mathcal{I}_\textrm{D}^\textrm{iso}.
+\end{align}
+
+We can simplify further if we assume that the spatial distribution of antenna pairs in each antenna pair type are similar and thus result in a similar dirty image, $\mathcal{I}_\textrm{D}^\textrm{iso}$.
+This usually holds when $\Nant$ is large and antennas under each antenna type are chosen randomly such that differences in the point spread functions are insignificant.
+Under this assumption, equation~\ref{eqn:apndx-wt-dirty-image-EPIC-decomp} can be written as:
+\begin{align}\label{eqn:apndx-t-dirty-image-EPIC-decomp-approx1}
+  \mathcal{I}^\prime &\approx \frac{\mathcal{I}_\textrm{D}^\textrm{iso}}{\Nant(\Nant-1)}\,\sum_{ij}\sum_{\substack{ab\\a\ne b\\a\in i, b\in j}} \mathcal{W}^{\textrm{G}\star}_a\,\mathcal{W}^\textrm{G}_b\,\mathcal{W}^\textrm{I}_a\,\mathcal{W}^{\textrm{I}\star}_b,
+\end{align}
+where, $\mathcal{I}_\textrm{D}^\textrm{iso}$ is the same for all terms and has been pulled outside the summations.
+For all antennas indexed by $a$ that belong to a particular type $i$, we replace $\mathcal{W}_a$ with $\mathcal{W}_{(i)}$ {\bf and is applicable to those arising} from both instrumental (superscript `$I$') and gridding (superscript `$G$') origins.
+Using arguments similar to those in equation~\ref{eqn:apndx-total-baselines}, equation~\ref{eqn:apndx-wt-dirty-image-EPIC-decomp-approx1} can be further simplified to:
+\begin{align}\label{eqn:apndx-wt-dirty-image-EPIC-decomp-approx2}
+  \mathcal{I}^\prime &\approx \frac{\mathcal{I}_\textrm{D}^\textrm{iso}}{\Nant(\Nant-1)}\,\sum_{ij}n_{ij}\,\mathcal{W}^{\textrm{G}\star}_{(i)}\,\mathcal{W}^\textrm{G}_{(j)}\,\mathcal{W}^\textrm{I}_{(i)}\,\mathcal{W}^{\textrm{I}\star}_{(j)}.
+\end{align}
+
+This means $\mathcal{W}_\textrm{eff}$ can be expressed as fourth power combinations of antenna voltage patterns each weighted by the number of antenna pairs in those antenna pair types:
+\begin{align}
+  \mathcal{W}_\textrm{eff}^\textrm{opt} &\approx \frac{1}{\Nant(\Nant-1)}\,\sum_{ij}n_{ij}\,\mathcal{W}^{\textrm{G}\star}_{(i)}\,\mathcal{W}^\textrm{G}_{(j)}\,\mathcal{W}^\textrm{I}_{(i)}\,\mathcal{W}^{\textrm{I}\star}_{(j)}.
+\end{align}
+In an optimally weighted image, $\mathcal{W}^\textrm{G}_{(i)} \equiv \mathcal{W}^\textrm{I}_{(i)}$. Then, equation~\ref{eqn:apndx-effective-weighting} reduces to:
+\begin{align}\label{eqn:apndx-effective-weighting-optimal}
+  \mathcal{W}_\textrm{eff}^\textrm{opt} &\approx \frac{1}{\Nant(\Nant-1)}\,\sum_{ij} n_{ij}\,\left|\mathcal{W}^\textrm{I}_{(i)}\right|^2\left|\mathcal{W}^\textrm{I}_{(j)}\right|^2,
+\end{align}
+where the superscript `$\textrm{opt}$' denotes `optimal'.}
+
+% For purposes of illustration, we consider a simple scenario in which there are two antenna types each containing roughly equal numbers of antennas with similar layouts but the analysis is performed under the wrong assumption that all antennas are identical and are all of the first type.
+In an optimally weighted image, the optimal effective weighting given by equation~\ref{eqn:effective-weighting-optimal} will be:
+% \begin{align}\label{eqn:effective-weighting-optimal-example}
+%   \mathcal{W}_\textrm{eff}^\textrm{opt} &\approx \frac{1}{\Nant(\Nant-1)}\,\left[n_1(n_1-1)\,\left|\mathcal{W}^\textrm{I}_{(1)}\right|^4 \right.\nonumber\\
+%   &\qquad\qquad\qquad\qquad + n_2(n_2-1)\,\left|\mathcal{W}^\textrm{I}_{(2)}\right|^4 \nonumber\\
+%   &\qquad\qquad\qquad\qquad \left. + n_1n_2\,\left|\mathcal{W}^\textrm{I}_{(1)}\right|^2\,\left|\mathcal{W}^\textrm{I}_{(2)}\right|^2\right].
+% \end{align}
+% On the other hand, $\mathcal{W}^\textrm{G}_{(1)} = \mathcal{W}^\textrm{I}_{(1)}$ and $\mathcal{W}^\textrm{G}_{(2)} = \mathcal{W}^\textrm{I}_{(1)}$ for an image obtained in the erroneous case in which all antennas are assumed to be identical.
+It will be sub-optimal with effective weighting:
+% \begin{align}\label{eqn:effective-weighting-suboptimal-example}
+%   \mathcal{W}_\textrm{eff}^\textrm{sub} &\approx \frac{1}{\Nant(\Nant-1)}\,\left[n_1(n_1-1)\,\left|\mathcal{W}^\textrm{I}_{(1)}\right|^4\right.\nonumber\\
+%   &\qquad\qquad\qquad\qquad + n_2(n_2-1)\,\left|\mathcal{W}^\textrm{I}_{(1)}\right|^2\,\left|\mathcal{W}^\textrm{I}_{(2)}\right|^2 \nonumber\\
+%   &\qquad\qquad\qquad\qquad \left. + n_1n_2\,\left|\mathcal{W}^\textrm{I}_{(1)}\right|^2 \,\mathcal{W}^\textrm{I}_{(1)}\,\mathcal{W}^{\textrm{I}\star}_{(2)}\right].
+% \end{align}
+% Finally, since no {\it a priori} information will be available about the heterogeneous array in case of the erroneous assumption, the effective weighting in the image will be believed to be $\mathcal{W}_\textrm{eff}^\textrm{err} = \left|\mathcal{W}_{(1)}^\textrm{I}\right|^4$.
+
+% Estimates of the dirty images with uniform, isotropic weighting in the optimal and erroneous cases will be obtained by dividing the respective weighted images by their assumed effective weights $\mathcal{W}_\textrm{eff}^\textrm{opt}$ and $\mathcal{W}_\textrm{eff}^\textrm{err}$ respectively as:
+% \begin{align}
+%   \widehat{\mathcal{I}}_\textrm{D}^\textrm{iso} &\approx \begin{cases}
+%     \mathcal{I}_\textrm{D}^\textrm{iso} &\textrm{, \qquad optimal case} \\
+%     \mathcal{I}_\textrm{D}^\textrm{iso}\,\frac{\mathcal{W}_\textrm{eff}^\textrm{sub}}{\mathcal{W}_\textrm{eff}^\textrm{err}} &\textrm{, \qquad erroneous case}
+%   \end{cases}
+% \end{align}
+
+% If $\mathcal{I}_\textrm{M}$ is the sky model, then $\mathcal{I}_\textrm{D}^\textrm{iso}$ can be expressed as:
+% \begin{align}\label{eqn:img-uncertainties}
+%   \mathcal{I}_\textrm{D}^\textrm{iso} &= \mathcal{I}_\textrm{M} + \delta\mathcal{I}_\textrm{S} + \delta\mathcal{I}_\textrm{M},
+% \end{align}
+% where the second and third terms on the right hand side denote fluctuations due to sidelobes and those intrinsic to the model, respectively.
+We define the normalized estimates of the sky model $\widehat{\mathcal{I}}_\textrm{M}^\textrm{norm} \equiv \widehat{\mathcal{I}}_\textrm{D}^\textrm{iso} / \mathcal{I}_\textrm{M}$.
+Ideally, $\widehat{\mathcal{I}}_\textrm{M}^\textrm{norm}$ should be consistent with unity and not exceed beyond levels of uncertainty as determined by equation~\ref{eqn:img-uncertainties}.}
+
 % \section{Some extra material}

### 2016-08-24 imaging_demo/ms.tex
Make corrections suggested by Judd
From e1be335 (Judd-v1):
======================

In ms.tex,

(*) Make corrections suggested by Judd

@@
 We have also validated our implementation on data observed with the Long Wavelength Array (LWA).
-{\bf We have demonstrated its versatility to image with heterogeneous arrays by showing that it robustly estimates the input sky model while wrong assumptions about array homogeneity lead to systematic mis-estimates.} Antenna layouts with dense filling factors consisting of a large number of antennas such as LWA, the Square Kilometre Array, Hydrogen Epoch of Reionization Array, and Canadian Hydrogen Intensity Mapping Experiment will gain significant computational advantage by deploying EPIC.
-Inherent availability of calibrated time-domain images on digitizer writeout time-scales and vastly lower I/O bandwidth relative to visibility-based systems will make it a prime candidate for transient searches of Fast Radio Bursts (FRB) as well as planetary and exoplanetary phenomena.
+{\bf We show that it robustly estimates the input sky model for heterogeneous arrays.} Antenna layouts with dense filling factors consisting of a large number of antennas such as LWA, the Square Kilometre Array, Hydrogen Epoch of Reionization Array, and Canadian Hydrogen Intensity Mapping Experiment will gain significant computational advantage by deploying {\bf an optimized version of} EPIC.
+{\bf The algorithm is a strong candidate for instruments targeting transient searches of Fast Radio Bursts (FRB) as well as planetary and exoplanetary phenomena due to the availability of high-speed calibrated time-domain images and low output bandwidth relative to visibility-based systems.}
 \end{abstract}
@@
 Similarly, the search for transient objects and regular monitoring of the time-dependent sky is driving instruments in the same direction with the added requirement of fast read-outs.
-A number of new telescopes are being or were developed around the world based on this new paradigm, including the Hydrogen Epoch of Reionization Array\footnote{http://reionization.org} (HERA; \citealt{deb16}), the Murchison Widefield Array (MWA; \citealt{tin13,bow13}), the {\bf Donald~C.~Backer} Precision Array for Probing the Epoch of Reionization (PAPER; \citealt{par10}), the LOw Frequency ARray (LOFAR; \citealt{van13}), the Canadian Hydrogen Intensity Mapping Experiment (CHIME; \citealt{ban14}), the Long Wavelength Array (LWA; \citealt{ell13}), and the low frequency Square Kilometer Array (SKA1-Low; \citealt{mel13}).
+A number of new telescopes around the world {\bf are} based on this new paradigm, including the Hydrogen Epoch of Reionization Array\footnote{http://reionization.org} (HERA; \citealt{deb16}), the Murchison Widefield Array (MWA; \citealt{tin13,bow13}), the {\bf Donald~C.~Backer} Precision Array for Probing the Epoch of Reionization (PAPER; \citealt{par10}), the LOw Frequency ARray (LOFAR; \citealt{van13}), the Canadian Hydrogen Intensity Mapping Experiment (CHIME; \citealt{ban14}), the Long Wavelength Array (LWA; \citealt{ell13}), and the low frequency Square Kilometer Array (SKA1-Low; \citealt{mel13}).
 
@@
 grid points in the Fourier transform \citep[e.g.][]{mor11,teg09,teg10}. This 
-architecture is often referred to as a ``direct imaging'' correlator because it 
+architecture is often referred to as a ``direct imaging correlator'' because it 
 eliminates the intermediate cross-correlation data products of the FX and XF 
@@
 of pulsar timing \citep{oto94, dai00}. However, each of these examples make
-assumptions about the redundancy of the array layout, and require the collecting
+assumptions about the redundancy of the array layout, and require that the collecting
 elements are identical. On the other hand, the MOFF algorithm achieves the same
@@
 holography/A-transpose fashion \citep{mor09,bha08,teg97b} before performing the
-spatial FFT. This process has been shown to theoretically produce a data product
+spatial FFT. This process has been shown to mathematically produce a data product
 identical to images produced from traditional visibility-based techniques.
@@
 code. EPIC is a highly parallel, object oriented Python package that primarily 
-implements the MOFF imaging algorithm besides emulating real-life telescopes and 
+implements the MOFF imaging algorithm {\bf as well as} emulating real-life telescopes and 
 FX/XF correlators in software, and includes a visibility-based imaging technique 
@@
 {\bf \S\ref{sec:versatility} demonstrates its ability to robustly image data from heterogeneous arrays.} In \S\ref{sec:analysis}, we explore the scalability of the algorithm in the context of several array design choices.
-We identify specific array design classes where the EPIC correlator -- is computationally more efficient; and in the field of transients, demands significantly lesser I/O bandwidth relative to visibility-based approaches.
+We identify specific array design classes where the EPIC correlator is computationally more efficient, and in the field of transients, demands significantly lesser I/O bandwidth relative to visibility-based approaches.
 We conclude and discuss future research prospects in \S\ref{sec:conclusions}.
@@
 where, $B_{ab}(\mathbf{r},f,t)$ denotes the {\bf power response of the antenna pair obtained in the measurement plane} by a spatial Fourier transform of $\mathcal{B}_{ab}(\hat{\mathbf{s}},f,t)$.
-Effectively, the multiplication in image space by $\mathcal{B}_{ab}(\hat{\mathbf{s}},f,t)$ has been replaced by a convolution with $B_{ab}(\mathbf{r},f,t)$ in {\bf measurement plane}.
+Effectively, the multiplication in image space by $\mathcal{B}_{ab}(\hat{\mathbf{s}},f,t)$ has been replaced by a convolution with $B_{ab}(\mathbf{r},f,t)$ in {\bf the measurement plane}.
 This is the software holographic equivalent of a traditional FX correlator output.
@@
 \item In visibility-based imaging, electric fields measured by antennas are not correlated with themselves and hence lack zero spacing measurements.
-In contrast, in MOFF imaging, since the gridded electric fields are imaged and squared, they retain information from auto-correlated electric fields at zero spacing {\bf but can be removed in EPIC.}
+In contrast, in MOFF imaging, since the gridded electric fields are imaged and squared, they retain information from auto-correlated electric fields at zero spacing.
+{\bf However, we will show in \S\ref{sec:rm-autocorr} that they can be removed by fitting and subtracting the zero-spacing contribution during imaging.}
 \end{enumerate}
@@
 We have implemented the MOFF imaging technique in our ``E-field Parallel Imaging Correlator'' -- a highly parallelized Object Oriented Python package,\footnote{The E-field Parallel Imaging Correlator (EPIC) package can be accessed at https://github.com/nithyanandan/EPIC} now publicly available.
-Besides implementing the MOFF imaging algorithm it also includes {\bf a simulator for generating electric fields from a sky model and a visibility-based imaging using the software holography technique.}
+Besides implementing the MOFF imaging algorithm it also includes {\bf a simulator for generating electric fields from a sky model and a visibility-based imaging routine using the software holography technique.}
 
 EPIC can accept dual-polarization inputs and produce images of all four instrumental cross-polarizations.
-Currently two data input formats exist for reading in the electric field time samples measured by the antennas -- simulated electric fields based on a sky model using the simulator packaged with EPIC; and LWA data.
+Currently, two data input formats exist for reading in the electric field time samples measured by the antennas -- simulated electric fields based on a sky model using the simulator packaged with EPIC; and LWA data.
 Efforts to build interfaces for data from other telescopes are underway.
@@
   % \includegraphics[width=\columnwidth]{MOFF_flowchart.eps}
-  \caption{A flowchart of MOFF imaging in EPIC.
+  \caption{Flowchart of MOFF imaging in EPIC.
 The propagated electric fields shown on the left are measured as time-series $\widetilde{E}_1(t)\ldots \widetilde{E}_\textrm{N}(t)$ by the antennas $\textrm{A}_1\ldots \textrm{A}_\textrm{N}$ which are then Fourier transformed by the F-engine to produce electric field spectra $E_1(f)\ldots E_\textrm{N}(f)$.
@@
 They are calibrated and time-averaged to obtain $\langle V_\textrm{ab}(f)\rangle$ which are then gridded and imaged to obtain the image $\mathcal{I}^\prime(f)$.
-The $\mathcal{I}^\prime(f)$ obtained from both techniques are theoretically identical as explained in \S\ref{sec:math}.
+The $\mathcal{I}^\prime(f)$ obtained from both techniques are mathematically identical as explained in \S\ref{sec:math}.
 
@@
   % \includegraphics[width=\columnwidth]{FX_flowchart.eps}
-  \caption{A flowchart of visibility-based software holographic imaging in EPIC.
+  \caption{Flowchart of visibility-based software holographic imaging in EPIC.
 The FX process flow shares the F-engine with the MOFF process.
@@
 
-Here we discuss the components of these architectures in detail.
+{\bf A high level software architecture of the EPIC package is described in the 
+appendix~\ref{sec:software-modules} for the interested reader.} Here we discuss the components of these architectures in detail.
 
@@
 \noindent A grid is generated on the coordinate system in which antenna 
-locations are specified with a grid spacing. The grid spacing can be controlled 
+locations are specified. The grid spacing can be controlled 
 by the user. By default, it is set to be $\le\lambda/2$ even at the 
@@
 
-A high level software architecture of the EPIC package is described in the 
-appendix~\ref{sec:software-modules} for the interested reader.
-
 \section{Verification}\label{sec:verify}
@@
   % \includegraphics[width=\columnwidth]{autocorr_uvwts_pbeam.png}
-  \caption{The auto-correlation of weights of a square shaped antenna aperture
+  \caption{Auto-correlation of weights of a square shaped antenna aperture
     in the $uv$ plane (left) and the corresponding directional antenna power 
@@
 In other words, the correlated weights may suffer further loss of precision in their sampled footprint after correlation of two footprints each of which could be less precise to first order.
-On the other hand, in visibility-based imaging, a directly sampled $uv$ plane antenna power response (which is theoretically identical to the correlation of individual antenna footprints) centred on a baseline has a loss in precision at most to first order.
+On the other hand, in visibility-based imaging, a directly sampled $uv$ plane antenna power response (which is mathematically identical to the correlation of individual antenna footprints) centred on a baseline has a loss in precision at most to first order.
 Thus, although in the limit of infinitesimally small grid spacing they should be identical, the coarseness of grid spacing introduces subtle differences between the two.
@@
 {\bf One of the advanced features of the MOFF algorithm and its implementation via EPIC is the ability to naturally account for heterogeneity of antennas while still producing images with maximal information \citep{mor11}.
-Unlike the assumption frequently made in existing direct imaging techniques that all antennas are identical, EPIC can implement the most generic case where no such assumption can be made.
-In other words, EPIC does not assume $W_a(\mathbf{r}-\mathbf{r}_a)$ or $B_{ab}(\mathbf{r}-\mathbf{r}_{ab})$ for different antennas are identical.
+Unlike the assumption frequently made in existing direct imaging techniques that all antennas are identical, EPIC does not assume $W_a(\mathbf{r}-\mathbf{r}_a)$ or $B_{ab}(\mathbf{r}-\mathbf{r}_{ab})$ for different antennas are identical.
 However, any {\it a priori} knowledge available that certain sets of antennas have identical illumination patterns can be easily passed on to EPIC to avoid redundant computations and thus increase efficiency.
@@
 Optimal imaging requires $W^\textrm{G}_a = W^\textrm{I}_a$ \citep{mor09,mor11}.
-However, we keep the two superscripts separately to describe output images under the wrong assumption that the array is homogeneous when instead it is actually not.
+However, we keep the two superscripts separate here to describe output images under the wrong assumption that the array is homogeneous when instead it is actually not.
 
@@
 Thus symbols of one color that lie to the left of same colored line indicate MOFF imaging is favoured for those telescopes and vice versa.
-For e.g., MOFF imaging is favoured in HERA-331 and HERA-6769 because they lie to the left of the orange line but not so in cases of HERA-19 and HERA-37}.
+For example, MOFF imaging is favoured in HERA-331 and HERA-6769 because they lie to the left of the orange line but not so in cases of HERA-19 and HERA-37}.
 A majority of the next-generation radio telescopes, namely, HERA-331 and its future expanded versions, SKA1-LC, SKA1-LCD, HIRAX, and CHIME will fall in the regime where MOFF imaging will be desirable.

### 2016-08-24 imaging_demo/ms.tex
Minor changes throughout
From b52cb67 (Judd-v1):
======================

In ms.tex,

(*) Minor changes throughout

@@
 Not only does it bring down the cost for dense layouts to $\mathcal{O}(\Nant\log_2\Nant)$ but can also image from irregular layouts and heterogeneous arrays of antennas.
-EPIC is highly modular and parallelizable, implemented in object oriented Python, and publicly available.
-We have verified the images produced to be equivalent to those produced using traditional techniques to within a precision set by gridding coarseness.
+EPIC is highly modular, parallelizable, implemented in object-oriented Python, and publicly available.
+We have verified the images produced to be equivalent to those from traditional techniques to within a precision set by gridding coarseness.
 We have also validated our implementation on data observed with the Long Wavelength Array (LWA).
@@
 
-Direct imaging correlators have begun to be explored on deployed arrays including
-the Basic Element for SKA Training II (BEST-2) array \citep{fos14}, the Omniscope
-\citep{zhe14}, and an earlier incarnation at higher frequencies with the intent
-of pulsar timing \citep{oto94, dai00}. However, each of these examples make
-assumptions about the redundancy of the array layout, and require that the collecting
-elements are identical. On the other hand, the MOFF algorithm achieves the same
-$\Ngrid \log_2 \Ngrid$ computational scaling without placing any restriction on
-antenna placement, can accommodate non-identical antennas, and is provably 
-optimal \citep{mor11}. This algorithm uses the antenna beam patterns to 
-grid the electric field measurements to a regular grid in the software 
-holography/A-transpose fashion \citep{mor09,bha08,teg97b} before performing the
-spatial FFT. This process has been shown to mathematically produce a data product
-identical to images produced from traditional visibility-based techniques.
+Direct imaging correlators have begun to be explored on deployed arrays including the Basic Element for SKA Training II (BEST-2) array \citep{fos14}, the MIT EoR (MITEoR) experiment \citep{zhe14} using the {\it Omniscope} \citep{teg09,teg10}, and an earlier incarnation at higher frequencies with the intent of pulsar timing \citep{oto94,dai00}.
+However, each of these examples make assumptions about the redundancy of the array layout, and require that the collecting elements are identical.
+On the other hand, the MOFF algorithm achieves the same $\Ngrid \log_2 \Ngrid$ computational scaling without placing any restriction on antenna placement, can accommodate non-identical antennas, and is provably optimal \citep{mor11}.
+This algorithm uses the antenna beam patterns to grid the electric field measurements to a regular grid in the software holography/A-transpose fashion \citep{mor09,bha08,teg97b} before performing the spatial FFT.
+This process has been shown to mathematically produce a data product identical to images produced from traditional visibility-based techniques.
 
@@
 {\bf One of the advanced features of the MOFF algorithm and its implementation via EPIC is the ability to naturally account for heterogeneity of antennas while still producing images with maximal information \citep{mor11}.
-Unlike the assumption frequently made in existing direct imaging techniques that all antennas are identical, EPIC does not assume $W_a(\mathbf{r}-\mathbf{r}_a)$ or $B_{ab}(\mathbf{r}-\mathbf{r}_{ab})$ for different antennas are identical.
+Unlike the assumption frequently made in existing direct imaging techniques that all antennas are identical \citep{oto94,dai00,teg09,teg10,fos14,zhe14}, EPIC does not assume $W_a(\mathbf{r}-\mathbf{r}_a)$ or $B_{ab}(\mathbf{r}-\mathbf{r}_{ab})$ for different antennas are identical.
 However, any {\it a priori} knowledge available that certain sets of antennas have identical illumination patterns can be easily passed on to EPIC to avoid redundant computations and thus increase efficiency.
@@
 Optimal imaging requires $W^\textrm{G}_a = W^\textrm{I}_a$ \citep{mor09,mor11}.
-However, we keep the two superscripts separate here to describe output images under the wrong assumption that the array is homogeneous when instead it is actually not.
+However, we keep the two superscripts separate here to describe output images when the array is wrongly assumed to be homogeneous.
 
@@
 \begin{align}
-  \mathcal{I}^\prime_{ab}(\hat{\mathbf{s}},f) &= \mathcal{W}^{\textrm{G}\star}_a(\hat{\mathbf{s}},f)\mathcal{W}^\textrm{G}_b(\hat{\mathbf{s}},f)\,V_{ab}\,e^{2\pi i f\mathbf{r}_{ab}\!\cdot\,\hat{\mathbf{s}}/c}.
+  \mathcal{I}^\prime_{ab}(\hat{\mathbf{s}},f) &= \mathcal{W}^{\textrm{G}\star}_a(\hat{\mathbf{s}},f)\,\mathcal{W}^\textrm{G}_b(\hat{\mathbf{s}},f)\,V_{ab}\,e^{2\pi i f\mathbf{r}_{ab}\!\cdot\,\hat{\mathbf{s}}/c}.
 \end{align}
@@
 \begin{align}\label{eqn:wt-dirty-image-EPIC}
-  \mathcal{I}^\prime &= \frac{1}{\Nant(\Nant-1)}\,\sum_{\substack{ab\\a\ne b}}\mathcal{I}^\prime_{ab}
+  \mathcal{I}^\prime &= \frac{1}{\Nant(\Nant-1)}\,\sum_{\substack{ab\\a\ne b}}\,\mathcal{I}^\prime_{ab}.
 \end{align}
@@
 \end{align}
-where, $\mathcal{I}_\textrm{D}^\textrm{iso}$ is the dirty image with no beam weighting (isotropic, uniform weighting of the sky) determined only by the array layout.
+where, $\mathcal{I}_\textrm{D}^\textrm{iso}$ is the dirty image with no beam weighting (isotropic, uniform weighting of the sky).
 
@@
 where, $ab$ indexes pairs of antennas $a$ and $b$, $ij$ indexes types of pairs of antennas, with $i$ and $j$ indexing antenna types.
-Thus, $\mathcal{I}^\prime$ is written as the sum of fringes over all unique antenna pairs in an antenna pair type (inner sum) and subsequently summed over all antenna pair types (outer sum).
+Thus, equation~\ref{eqn:wt-dirty-image-homogeneous} is re-written as the sum of fringes over all unique antenna pairs in an antenna pair type (inner sum) and subsequently summed over all antenna pair types (outer sum).
 
@@
 On the other hand, $\mathcal{W}^\textrm{G}_{(1)} = \mathcal{W}^\textrm{I}_{(1)}$ and $\mathcal{W}^\textrm{G}_{(2)} = \mathcal{W}^\textrm{I}_{(1)}$ for an image obtained in the erroneous case in which all antennas are assumed to be identical.
-It will be sub-optimal with effective weighting:
+It will be sub-optimal.
+From equation~\ref{eqn:wt-dirty-image-EPIC-decomp-approx2}, the effective weighting will be:
 \begin{align}\label{eqn:effective-weighting-suboptimal-example}
@@
 Optimal imaging requires $W^\textrm{G}_a = W^\textrm{I}_a$ \citep{mor09,mor11}.
-However, we keep the two superscripts separately to describe output images under the wrong assumption that the array is homogeneous when instead it is actually heterogeneous.
+However, we keep the two superscripts separate here to describe output images made assuming wrongly that the array is homogeneous.
 
@@
 We can simplify further if we assume that the spatial distribution of antenna pairs in each antenna pair type are similar and thus result in a similar dirty image, $\mathcal{I}_\textrm{D}^\textrm{iso}$.
-This usually holds when $\Nant$ is large and antennas under each antenna type are chosen randomly such that differences in the point spread functions are insignificant.
-Under this assumption, equation~\ref{eqn:apndx-wt-dirty-image-EPIC-decomp} can be written as:
+This is valid when antennas under each antenna type are chosen such that differences in the point spread functions are insignificant.
+Then, equation~\ref{eqn:apndx-wt-dirty-image-EPIC-decomp} can be written as:
 \begin{align}\label{eqn:apndx-t-dirty-image-EPIC-decomp-approx1}

### 2016-08-24 imaging_demo/ms.tex
Add description to solid angle element
From 8c6110f (mnras):
====================

In ms.tex,

(*) Add description to solid angle element

@@
 \end{align}
-where, $\mathbf{r}$ denotes the observer's location, {\bf $f$ is the frequency of radiation, $c$ is the speed of light, $t$ denotes time} and $E(\mathbf{r}, f, t)$ is the propagated electric field.
+where, $\mathbf{r}$ denotes the observer's location, {\bf $f$ is the frequency of radiation, $c$ is the speed of light, $t$ denotes time, $\dif\Omega$ is the infinitesimal solid angle element on the celestial sphere whose normal is the unit vector $\hat{\mathbf{s}}$}, and $E(\mathbf{r}, f, t)$ is the propagated electric field.
 Thus the propagated electric field is a linear superposition of the electric fields emanating from astronomical sources with appropriate complex phases.
@@
 \end{align}
-Here, equation~\ref{eqn:wt-dirty-image-EPIC} has been re-written as the sum of fringes over all antenna pairs in an antenna pair type (inner sum) and subsequently summed over all antenna pair types (outer sum).
+Here, equation~\ref{eqn:apndx-wt-dirty-image-EPIC} has been re-written as the sum of fringes over all antenna pairs in an antenna pair type (inner sum) and subsequently summed over all antenna pair types (outer sum).
 The effective attenuation of the dirty image due to instrumental and gridding weights is:
@@
 
-% For purposes of illustration, we consider a simple scenario in which there are two antenna types each containing roughly equal numbers of antennas with similar layouts but the analysis is performed under the wrong assumption that all antennas are identical and are all of the first type.
-In an optimally weighted image, the optimal effective weighting given by equation~\ref{eqn:effective-weighting-optimal} will be:
-% \begin{align}\label{eqn:effective-weighting-optimal-example}
-%   \mathcal{W}_\textrm{eff}^\textrm{opt} &\approx \frac{1}{\Nant(\Nant-1)}\,\left[n_1(n_1-1)\,\left|\mathcal{W}^\textrm{I}_{(1)}\right|^4 \right.\nonumber\\
-%   &\qquad\qquad\qquad\qquad + n_2(n_2-1)\,\left|\mathcal{W}^\textrm{I}_{(2)}\right|^4 \nonumber\\
-%   &\qquad\qquad\qquad\qquad \left. + n_1n_2\,\left|\mathcal{W}^\textrm{I}_{(1)}\right|^2\,\left|\mathcal{W}^\textrm{I}_{(2)}\right|^2\right].
-% \end{align}
-% On the other hand, $\mathcal{W}^\textrm{G}_{(1)} = \mathcal{W}^\textrm{I}_{(1)}$ and $\mathcal{W}^\textrm{G}_{(2)} = \mathcal{W}^\textrm{I}_{(1)}$ for an image obtained in the erroneous case in which all antennas are assumed to be identical.
-It will be sub-optimal with effective weighting:
-% \begin{align}\label{eqn:effective-weighting-suboptimal-example}
-%   \mathcal{W}_\textrm{eff}^\textrm{sub} &\approx \frac{1}{\Nant(\Nant-1)}\,\left[n_1(n_1-1)\,\left|\mathcal{W}^\textrm{I}_{(1)}\right|^4\right.\nonumber\\
-%   &\qquad\qquad\qquad\qquad + n_2(n_2-1)\,\left|\mathcal{W}^\textrm{I}_{(1)}\right|^2\,\left|\mathcal{W}^\textrm{I}_{(2)}\right|^2 \nonumber\\
-%   &\qquad\qquad\qquad\qquad \left. + n_1n_2\,\left|\mathcal{W}^\textrm{I}_{(1)}\right|^2 \,\mathcal{W}^\textrm{I}_{(1)}\,\mathcal{W}^{\textrm{I}\star}_{(2)}\right].
-% \end{align}
-% Finally, since no {\it a priori} information will be available about the heterogeneous array in case of the erroneous assumption, the effective weighting in the image will be believed to be $\mathcal{W}_\textrm{eff}^\textrm{err} = \left|\mathcal{W}_{(1)}^\textrm{I}\right|^4$.
-
-% Estimates of the dirty images with uniform, isotropic weighting in the optimal and erroneous cases will be obtained by dividing the respective weighted images by their assumed effective weights $\mathcal{W}_\textrm{eff}^\textrm{opt}$ and $\mathcal{W}_\textrm{eff}^\textrm{err}$ respectively as:
-% \begin{align}
-%   \widehat{\mathcal{I}}_\textrm{D}^\textrm{iso} &\approx \begin{cases}
-%     \mathcal{I}_\textrm{D}^\textrm{iso} &\textrm{, \qquad optimal case} \\
-%     \mathcal{I}_\textrm{D}^\textrm{iso}\,\frac{\mathcal{W}_\textrm{eff}^\textrm{sub}}{\mathcal{W}_\textrm{eff}^\textrm{err}} &\textrm{, \qquad erroneous case}
-%   \end{cases}
-% \end{align}
-
-% If $\mathcal{I}_\textrm{M}$ is the sky model, then $\mathcal{I}_\textrm{D}^\textrm{iso}$ can be expressed as:
-% \begin{align}\label{eqn:img-uncertainties}
-%   \mathcal{I}_\textrm{D}^\textrm{iso} &= \mathcal{I}_\textrm{M} + \delta\mathcal{I}_\textrm{S} + \delta\mathcal{I}_\textrm{M},
-% \end{align}
-% where the second and third terms on the right hand side denote fluctuations due to sidelobes and those intrinsic to the model, respectively.
-We define the normalized estimates of the sky model $\widehat{\mathcal{I}}_\textrm{M}^\textrm{norm} \equiv \widehat{\mathcal{I}}_\textrm{D}^\textrm{iso} / \mathcal{I}_\textrm{M}$.
-Ideally, $\widehat{\mathcal{I}}_\textrm{M}^\textrm{norm}$ should be consistent with unity and not exceed beyond levels of uncertainty as determined by equation~\ref{eqn:img-uncertainties}.}
-
 % \section{Some extra material}

### 2016-09-17 imaging_demo/ms.tex
Remove boldface fonts highlighting changes
From 9767443 (mnras):
====================

In ms.tex,

(*) Remove boldface fonts that were highlighting changes

@@
 We have also validated our implementation on data observed with the Long Wavelength Array (LWA).
-{\bf We show that it robustly estimates the input sky model for heterogeneous arrays.} Antenna layouts with dense filling factors consisting of a large number of antennas such as LWA, the Square Kilometre Array, Hydrogen Epoch of Reionization Array, and Canadian Hydrogen Intensity Mapping Experiment will gain significant computational advantage by deploying {\bf an optimized version of} EPIC.
-{\bf The algorithm is a strong candidate for instruments targeting transient searches of Fast Radio Bursts (FRB) as well as planetary and exoplanetary phenomena due to the availability of high-speed calibrated time-domain images and low output bandwidth relative to visibility-based systems.}
+We show that it robustly estimates the input sky model for heterogeneous arrays.
+Antenna layouts with dense filling factors consisting of a large number of antennas such as LWA, the Square Kilometre Array, Hydrogen Epoch of Reionization Array, and Canadian Hydrogen Intensity Mapping Experiment will gain significant computational advantage by deploying an optimized version of EPIC.
+The algorithm is a strong candidate for instruments targeting transient searches of Fast Radio Bursts (FRB) as well as planetary and exoplanetary phenomena due to the availability of high-speed calibrated time-domain images and low output bandwidth relative to visibility-based systems.
 \end{abstract}
@@
 Similarly, the search for transient objects and regular monitoring of the time-dependent sky is driving instruments in the same direction with the added requirement of fast read-outs.
-A number of new telescopes around the world {\bf are} based on this new paradigm, including the Hydrogen Epoch of Reionization Array\footnote{http://reionization.org} (HERA; \citealt{deb16}), the Murchison Widefield Array (MWA; \citealt{tin13,bow13}), the {\bf Donald~C.~Backer} Precision Array for Probing the Epoch of Reionization (PAPER; \citealt{par10}), the LOw Frequency ARray (LOFAR; \citealt{van13}), the Canadian Hydrogen Intensity Mapping Experiment (CHIME; \citealt{ban14}), the Long Wavelength Array (LWA; \citealt{ell13}), and the low frequency Square Kilometer Array (SKA1-Low; \citealt{mel13}).
+A number of new telescopes around the world are based on this new paradigm, including the Hydrogen Epoch of Reionization Array\footnote{http://reionization.org} (HERA; \citealt{deb16}), the Murchison Widefield Array (MWA; \citealt{tin13,bow13}), the Donald~C.~Backer Precision Array for Probing the Epoch of Reionization (PAPER; \citealt{par10}), the LOw Frequency ARray (LOFAR; \citealt{van13}), the Canadian Hydrogen Intensity Mapping Experiment (CHIME; \citealt{ban14}), the Long Wavelength Array (LWA; \citealt{ell13}), and the low frequency Square Kilometer Array (SKA1-Low; \citealt{mel13}).
 
@@
 code. EPIC is a highly parallel, object oriented Python package that primarily 
-implements the MOFF imaging algorithm {\bf as well as} emulating real-life telescopes and 
+implements the MOFF imaging algorithm as well as emulating real-life telescopes and 
 FX/XF correlators in software, and includes a visibility-based imaging technique 
@@
 We also demonstrate the performance with real-world data from the LWA.
-{\bf \S\ref{sec:versatility} demonstrates its ability to robustly image data from heterogeneous arrays.} In \S\ref{sec:analysis}, we explore the scalability of the algorithm in the context of several array design choices.
+\S\ref{sec:versatility} demonstrates its ability to robustly image data from heterogeneous arrays.
+In \S\ref{sec:analysis}, we explore the scalability of the algorithm in the context of several array design choices.
 We identify specific array design classes where the EPIC correlator is computationally more efficient, and in the field of transients, demands significantly lesser I/O bandwidth relative to visibility-based approaches.
@@
 \end{align}
-where, $\mathbf{r}$ denotes the observer's location, {\bf $f$ is the frequency of radiation, $c$ is the speed of light, $t$ denotes time, $\dif\Omega$ is the infinitesimal solid angle element on the celestial sphere whose normal is the unit vector $\hat{\mathbf{s}}$}, and $E(\mathbf{r}, f, t)$ is the propagated electric field.
+where, $\mathbf{r}$ denotes the observer's location, $f$ is the frequency of radiation, $c$ is the speed of light, $t$ denotes time, $\dif\Omega$ is the infinitesimal solid angle element on the celestial sphere whose normal is the unit vector $\hat{\mathbf{s}}$, and $E(\mathbf{r}, f, t)$ is the propagated electric field.
 Thus the propagated electric field is a linear superposition of the electric fields emanating from astronomical sources with appropriate complex phases.
-{\bf Ignoring wide-field effects}, it can be {\bf simplified as} a Fourier transform of the electric fields in the sky coordinates.
-
-An antenna, {\bf located at $\mathbf{r}_a$ (indexed by $a$)}, measures a phased sum of these propagated electric fields over its effective collecting area with an additive receiver noise:
+Ignoring wide-field effects, it can be simplified as a Fourier transform of the electric fields in the sky coordinates.
+
+An antenna, located at $\mathbf{r}_a$ (indexed by $a$), measures a phased sum of these propagated electric fields over its effective collecting area with an additive receiver noise:
 \begin{align}\label{eqn:measured-E-field}
@@
 \end{align}
-where, {\bf $W_a(\mathbf{r},f,t)$ is the aperture electric field illumination pattern of the antenna and its Fourier transform, $\mathcal{W}_a(\hat{\mathbf{s}},f,t)$, is the directional antenna voltage response at a given frequency and time}.
+where, $W_a(\mathbf{r},f,t)$ is the aperture electric field illumination pattern of the antenna and its Fourier transform, $\mathcal{W}_a(\hat{\mathbf{s}},f,t)$, is the directional antenna voltage response at a given frequency and time.
 
@@
 where we have brought the time average into the integral under the assumption that the aperture illumination pattern does not change over the time-scale of the averaging and $\star$ denotes complex conjugation.
-{\bf The noise-like cross-terms ideally have zero mean.
-This expression can be further simplified using the sky brightness}, $\mathcal{I}(\hat{\mathbf{s}},f,t)$, as $\left\langle \mathcal{E}(\hat{\mathbf{s}},f,t)\,\mathcal{E}^\star(\hat{\mathbf{s'}},f,t) \right\rangle_t = \mathcal{I}(\hat{\mathbf{s}},f,t)\,\delta(\hat{\mathbf{s}}-\hat{\mathbf{s'}})$, and defining the antenna-pair sky power response function (or the directional antenna-pair power pattern), $\mathcal{B}_{ab}(\hat{\mathbf{s}})~\equiv~\mathcal{W}_a(\hat{\mathbf{s}},f,t)~\mathcal{W}^\star_b(\hat{\mathbf{s}},f,t)$.
+The noise-like cross-terms ideally have zero mean.
+This expression can be further simplified using the sky brightness, $\mathcal{I}(\hat{\mathbf{s}},f,t)$, as $\left\langle \mathcal{E}(\hat{\mathbf{s}},f,t)\,\mathcal{E}^\star(\hat{\mathbf{s'}},f,t) \right\rangle_t = \mathcal{I}(\hat{\mathbf{s}},f,t)\,\delta(\hat{\mathbf{s}}-\hat{\mathbf{s'}})$, and defining the antenna-pair sky power response function (or the directional antenna-pair power pattern), $\mathcal{B}_{ab}(\hat{\mathbf{s}})~\equiv~\mathcal{W}_a(\hat{\mathbf{s}},f,t)~\mathcal{W}^\star_b(\hat{\mathbf{s}},f,t)$.
 The result is the visibility expressed in terms of the sky brightness, the power pattern, and uncorrelated noise terms which we group into $n_{ab}(f,t)$.
@@
 where, the baseline coordinate $\mathbf{r}_{ab}\equiv\mathbf{r}_a-\mathbf{r}_b$ is the vector separation between the two antennas.
-This signifies that the visibility ($V_{ab}$) measured between a pair of antennas is obtained by multiplying the sky brightness $\mathcal{I}(\hat{\mathbf{s}},f,t)$ by the antenna power response $\mathcal{B}_{ab}(\hat{\mathbf{s}},f,t)$ and Fourier transforming from the directional coordinates ($\hat{\mathbf{s}}$) to {\bf the measurement plane}, which are then sampled at the locations of the antenna spacings (or baselines), namely, $\mathbf{r}_{ab}$, and added to the noise $n_{ab}$.
+This signifies that the visibility ($V_{ab}$) measured between a pair of antennas is obtained by multiplying the sky brightness $\mathcal{I}(\hat{\mathbf{s}},f,t)$ by the antenna power response $\mathcal{B}_{ab}(\hat{\mathbf{s}},f,t)$ and Fourier transforming from the directional coordinates ($\hat{\mathbf{s}}$) to the measurement plane, which are then sampled at the locations of the antenna spacings (or baselines), namely, $\mathbf{r}_{ab}$, and added to the noise $n_{ab}$.
 
@@
 \end{align}
-where, $B_{ab}(\mathbf{r},f,t)$ denotes the {\bf power response of the antenna pair obtained in the measurement plane} by a spatial Fourier transform of $\mathcal{B}_{ab}(\hat{\mathbf{s}},f,t)$.
-Effectively, the multiplication in image space by $\mathcal{B}_{ab}(\hat{\mathbf{s}},f,t)$ has been replaced by a convolution with $B_{ab}(\mathbf{r},f,t)$ in {\bf the measurement plane}.
+where, $B_{ab}(\mathbf{r},f,t)$ denotes the power response of the antenna pair obtained in the measurement plane by a spatial Fourier transform of $\mathcal{B}_{ab}(\hat{\mathbf{s}},f,t)$.
+Effectively, the multiplication in image space by $\mathcal{B}_{ab}(\hat{\mathbf{s}},f,t)$ has been replaced by a convolution with $B_{ab}(\mathbf{r},f,t)$ in the measurement plane.
 This is the software holographic equivalent of a traditional FX correlator output.
-{\bf For context, the term ``holography'' is derived from holographic measurements of the complex antenna aperture illumination pattern \citep{ben76,sco77}.
-In the present context, ``software holography'' refers to the application of this antenna aperture illumination pattern in software for imaging.}
+For context, the term ``holography'' is derived from holographic measurements of the complex antenna aperture illumination pattern \citep{ben76,sco77}.
+In the present context, ``software holography'' refers to the application of this antenna aperture illumination pattern in software for imaging.
 
@@
 where the measured visibilities are weighted by the inverse of the system noise, followed by a gridding process using the holographic antenna power response as the gridding kernel, followed by a Fourier transform to create an image $\boldsymbol{\mathcal{I}}^\prime(\hat{\mathbf{s}})$.
-This is the {\bf optimally weighted} estimate of the true image $\boldsymbol{\mathcal{I}}(\hat{\mathbf{s}})$ given the visibility measurements.
+This is the optimally weighted estimate of the true image $\boldsymbol{\mathcal{I}}(\hat{\mathbf{s}})$ given the visibility measurements.
 
@@
 In contrast, in MOFF imaging, since the gridded electric fields are imaged and squared, they retain information from auto-correlated electric fields at zero spacing.
-{\bf However, we will show in \S\ref{sec:rm-autocorr} that they can be removed by fitting and subtracting the zero-spacing contribution during imaging.}
+However, we will show in \S\ref{sec:rm-autocorr} that they can be removed by fitting and subtracting the zero-spacing contribution during imaging.
 \end{enumerate}
@@
 We have implemented the MOFF imaging technique in our ``E-field Parallel Imaging Correlator'' -- a highly parallelized Object Oriented Python package,\footnote{The E-field Parallel Imaging Correlator (EPIC) package can be accessed at https://github.com/nithyanandan/EPIC} now publicly available.
-Besides implementing the MOFF imaging algorithm it also includes {\bf a simulator for generating electric fields from a sky model and a visibility-based imaging routine using the software holography technique.}
+Besides implementing the MOFF imaging algorithm it also includes a simulator for generating electric fields from a sky model and a visibility-based imaging routine using the software holography technique.
 
@@
 
-{\bf A high level software architecture of the EPIC package is described in the 
-appendix~\ref{sec:software-modules} for the interested reader.} Here we discuss the components of these architectures in detail.
+A high level software architecture of the EPIC package is described in the 
+appendix~\ref{sec:software-modules} for the interested reader. Here we discuss the components of these architectures in detail.
 
@@
 model of the sky. Since antennas are only correlated with calibrator pixels, the 
-computational complexity scales as $\sim N_\textrm{a} N_\textrm{c}$. {\bf The calibration performance of EPIC is explored for varying levels of completeness of sky model in \citet{bea16}.}
+computational complexity scales as $\sim N_\textrm{a} N_\textrm{c}$. The calibration performance of EPIC is explored for varying levels of completeness of sky model in \citet{bea16}.
 
@@
 Both algorithms result in images and synthesized beams that are well matched with each other.
-{\bf As expected, their fluxes are attenuated by a weighting proportional to the square of the antenna power pattern corresponding to a uniform square aperture.}
+As expected, their fluxes are attenuated by a weighting proportional to the square of the antenna power pattern corresponding to a uniform square aperture.
 
@@
 The antenna auto-correlations at zero-spacing have been removed from the MOFF images.
-The images in either case reconstruct the sources at the right locations with the fluxes attenuated expected after multiplication by the antenna power pattern {\bf squared}.
+The images in either case reconstruct the sources at the right locations with the fluxes attenuated expected after multiplication by the antenna power pattern squared.
 The synthesized beams from the two algorithms are well matched in size and shape.
@@
 
-{\bf We examine in detail the respective synthesized beams in each case in Fig.~\ref{fig:MOFF-FX-psf-diff}.
+We examine in detail the respective synthesized beams in each case in Fig.~\ref{fig:MOFF-FX-psf-diff}.
 Slices at $m=0$ of the synthesized beams weighted by the antenna power pattern are shown for MOFF method using EPIC (solid black) and visibility-based (dashed gray) imaging.
@@
 A magnified view shows that some differences at the level of $<0.5$\% exist in some regions.
-These are attributed to differences in the $uv$-plane antenna cross-correlation weights in the two methods which in turn arises due to the amount of coarseness in grid spacing as described below.}
+These are attributed to differences in the $uv$-plane antenna cross-correlation weights in the two methods which in turn arises due to the amount of coarseness in grid spacing as described below.
 
@@
   \includegraphics[width=\columnwidth]{figure6}
-  \caption{{\bf Synthesized beams weighted by antenna power pattern sliced at $m=0$ obtained with MOFF algorithm using EPIC (solid black line) and visibility-based imaging (dashed gray line).
+  \caption{Synthesized beams weighted by antenna power pattern sliced at $m=0$ obtained with MOFF algorithm using EPIC (solid black line) and visibility-based imaging (dashed gray line).
 The two appear almost identical.
 The inset provides a magnified view of differences in the synthesized beam slices between the two techniques at levels $\lesssim 0.5$\% relative to the peak.
-These are attributed to differences that arise in gridding and depends on the coarseness of the grid.}}
+These are attributed to differences that arise in gridding and depends on the coarseness of the grid.}
   \label{fig:MOFF-FX-psf-diff}
@@
 
-{\bf The left panel of Fig.~\ref{fig:MOFF-FX-uvwts-diff} shows differences (in percentage relative to peak) in cross-correlated weights obtained with MOFF imaging in EPIC and visibility-based imaging.
+The left panel of Fig.~\ref{fig:MOFF-FX-uvwts-diff} shows differences (in percentage relative to peak) in cross-correlated weights obtained with MOFF imaging in EPIC and visibility-based imaging.
 The maximum difference appears near the center of the $uv$-plane corresponding to antenna auto-correlations which have not been perfectly removed in the former.
 In other regions of the $uv$-plane, the differences are of the order of a few percent.
-The reason for these differences is described below.}
+The reason for these differences is described below.
 
@@
   % \includegraphics[width=\columnwidth]{MOFF_FX_comparison_uvwts_test_aperture_zero_spacing_removed.eps}
-  \caption{{\bf {\it Left}: Differences (in \%) in cross-correlated weights in the $uv$ plane relative to the peak.
+  \caption{{\it Left}: Differences (in \%) in cross-correlated weights in the $uv$ plane relative to the peak.
 The biggest difference ($\sim 20$\%) is found around the zero-spacing corresponding to auto-correlations of antenna weights due to gridding differences augmented by the relatively high number density of antenna auto-correlation footprints in that region.
@@
 {\it Right}: Histogram (expressed as percentage) of the relative differences (in \%) between antenna cross-correlation weights shown on the left binned in intervals of 0.5\%.
-Over 70\% of the $uv$-cells do not differ by more than 0.5\% and over 90\% of $uv$-cells only differ by $<5$\% for the gridding coarseness used.}}
+Over 70\% of the $uv$-cells do not differ by more than 0.5\% and over 90\% of $uv$-cells only differ by $<5$\% for the gridding coarseness used.}
   \label{fig:MOFF-FX-uvwts-diff}
@@
 Fig.~\ref{fig:image-psf-diff} shows the difference between the synthesized beams obtained with the two methods.
-A difference map between the two synthesized beams is shown on the {\bf left}.
+A difference map between the two synthesized beams is shown on the left.
 The amplitude of the difference appears to be modulated by the directional power response of the antenna.
-{\bf On the right}, in radial bins, the rms of the synthesized beam (gray) and the rms of the difference map (black) are plotted in percentage units relative to the peak (to be read using the axis on the left side of the plot).
+On the right, in radial bins, the rms of the synthesized beam (gray) and the rms of the difference map (black) are plotted in percentage units relative to the peak (to be read using the axis on the left side of the plot).
 The antenna power pattern (red; to be read using the scale on the right) is plotted for reference.
@@
 $\approx \sqrt{2}$ higher than the rms of the each of the synthesized beams. 
-This {\bf means} that while differences exist, large fractions of them are still 
+This means that while differences exist, large fractions of them are still 
 well matched to each other even out to the horizon. Thus the rounding errors in 
@@
 
-{\bf One of the advanced features of the MOFF algorithm and its implementation via EPIC is the ability to naturally account for heterogeneity of antennas while still producing images with maximal information \citep{mor11}.
+One of the advanced features of the MOFF algorithm and its implementation via EPIC is the ability to naturally account for heterogeneity of antennas while still producing images with maximal information \citep{mor11}.
 Unlike the assumption frequently made in existing direct imaging techniques that all antennas are identical \citep{oto94,dai00,teg09,teg10,fos14,zhe14}, EPIC does not assume $W_a(\mathbf{r}-\mathbf{r}_a)$ or $B_{ab}(\mathbf{r}-\mathbf{r}_{ab})$ for different antennas are identical.
@@
 We look at the contribution to the image from the visibility, $V_{ab}$, from each antenna pair.
-While imaging, EPIC introduces a weighting to each antenna during gridding as given by equations~\ref{eqn:dirty-image-FX} and \ref{eqn:e-field-conv}, and equivalently, the weighted visibility from {\bf an} antenna pair $ab$ projected on the grid is:
+While imaging, EPIC introduces a weighting to each antenna during gridding as given by equations~\ref{eqn:dirty-image-FX} and \ref{eqn:e-field-conv}, and equivalently, the weighted visibility from an antenna pair $ab$ projected on the grid is:
 \begin{align}
@@
 where, $\mathcal{I}_\textrm{D}^\textrm{iso}$ is the same for all terms and has been pulled out of the summations.
-For all antennas indexed by $a$ that belong to a particular type $i$, we replace $\mathcal{W}_a$ with $\mathcal{W}_{(i)}$ {\bf and is applicable to those arising} from both instrumental (superscript `$I$') and gridding (superscript `$G$') origins.
+For all antennas indexed by $a$ that belong to a particular type $i$, we replace $\mathcal{W}_a$ with $\mathcal{W}_{(i)}$ and is applicable to those arising from both instrumental (superscript `$I$') and gridding (superscript `$G$') origins.
 If $n_{ij}$ is the number of antenna pairs in the antenna pair type $ij$, $\mathcal{W}_\textrm{eff}$ can be simplified as:
@@
 \subfloat[][Estimates of normalized radial flux density profile]{\label{fig:norm-flux-density}\includegraphics[width=0.42\linewidth]{figure10b.eps}}
-\caption{{\bf (a)~Optimally weighted image, $\mathcal{I}^\prime$, of simulated sky model obtained by EPIC using MOFF algorithm after averaging over entire bandwidth and duration of simulated observation.
+\caption{(a)~Optimally weighted image, $\mathcal{I}^\prime$, of simulated sky model obtained by EPIC using MOFF algorithm after averaging over entire bandwidth and duration of simulated observation.
 51 antennas in the inner core of the MWA layout were used with uniformly illuminated square apertures.
@@
 In the optimal case, the normalized estimates are consistent with unity to within ``$1\sigma$'' in almost all regions of the image, whereas in the erroneous case, they are mis-estimated (underestimated) with high significance ($\gtrsim 5\sigma$) in most regions of the image.
-There is no significant difference in estimates between the two different directions of point sources studied.}}
+There is no significant difference in estimates between the two different directions of point sources studied.}
 \label{fig:versatility}
@@
 This demonstrates that images from EPIC can robustly image data from heterogeneous arrays while being mathematically optimal.
-Besides being sub-optimal, erroneous assumptions about the homogeneity of an array can lead to significant and systematic mis-estimates (in this case, an underestimate) of the sky model.}
+Besides being sub-optimal, erroneous assumptions about the homogeneity of an array can lead to significant and systematic mis-estimates (in this case, an underestimate) of the sky model.
 
@@
 The different colored lines correspond to different antenna sizes (cyan - 1~m$^2$, blue - 7~m$^2$, purple - 16~m$^2$, green - 28~m$^2$, orange - 150~m$^2$, red - 740~m$^2$, gray - 5900~m$^2$).
-{\bf Dashed lines of each color denote the boundary to the left of which the MOFF algorithm is favoured for the corresponding antenna size and vice versa for visibility-based imaging.}
+Dashed lines of each color denote the boundary to the left of which the MOFF algorithm is favoured for the corresponding antenna size and vice versa for visibility-based imaging.
 
@@
 The different antennas are color coded by roughly the class of antenna size they fall into.
-{\bf Thus symbols of one color falling to the left of a line} of the same color indicate MOFF imaging is advantageous for those telescopes and vice versa.
-For example, MOFF imaging is favoured in HERA-331 and HERA-6769 because they {\bf lie to the left of the orange line} but not so in cases of HERA-19 and HERA-37.}
+Thus symbols of one color falling to the left of a line of the same color indicate MOFF imaging is advantageous for those telescopes and vice versa.
+For example, MOFF imaging is favoured in HERA-331 and HERA-6769 because they lie to the left of the orange line but not so in cases of HERA-19 and HERA-37.}
   \label{fig:parameter-space-bll-nant-instruments}
@@
 
-{\bf There is an upper limit set by the maximum number of antennas for each antenna size that can be densely packed inside various baseline lengths but is not shown on this figure to avoid crowding.
+There is an upper limit set by the maximum number of antennas for each antenna size that can be densely packed inside various baseline lengths but is not shown on this figure to avoid crowding.
 We note that as antenna size increases the maximum number of antennas for a dense packing as a function of baseline length decreases.
@@
 Similarly, with increase in antenna size, $\Ngrid$ also decreases when field of view imaging is achieved with an increased grid spacing equal to antenna size and hence lowers the amount of computations required with the MOFF algorithm.
-This shifts the dashed lines rightward as well.}
-
-{\bf The different antennas are color coded by roughly the class of antenna size they fall into.
+This shifts the dashed lines rightward as well.
+
+The different antennas are color coded by roughly the class of antenna size they fall into.
 Thus symbols of one color that lie to the left of same colored line indicate MOFF imaging is favoured for those telescopes and vice versa.
-For example, MOFF imaging is favoured in HERA-331 and HERA-6769 because they lie to the left of the orange line but not so in cases of HERA-19 and HERA-37}.
+For example, MOFF imaging is favoured in HERA-331 and HERA-6769 because they lie to the left of the orange line but not so in cases of HERA-19 and HERA-37.
 A majority of the next-generation radio telescopes, namely, HERA-331 and its future expanded versions, SKA1-LC, SKA1-LCD, HIRAX, and CHIME will fall in the regime where MOFF imaging will be desirable.
@@
 Under the class of direct imaging techniques, ours is one of the most generic -- neither does it place any constraint on the array layout to be on a regular grid nor does it require the antenna array to be homogeneous.
-{\bf We have demonstrated its natural ability to robustly image data from heterogeneous arrays and shown that wrong assumptions about array homogeneity could result in significant and systematic mis-estimation of the sky model.}
+We have demonstrated its natural ability to robustly image data from heterogeneous arrays and shown that wrong assumptions about array homogeneity could result in significant and systematic mis-estimation of the sky model.
 
@@
 successfully tested on simulated MWA observations as well as real LWA 
-observations {\bf from both imaging and calibration view points}.
+observations from both imaging and calibration viewpoints.
 
@@
 
-In the near future, we plan to {\bf demonstrate speed and precision by upgrading EPIC to a GPU-based implementation} in order to operate on real-time data and develop a transient trigger and monitor backend.
-{\bf In the meanwhile, we plan to demonstrate imaging with non-coplanar arrays and direction-dependent calibration.}
+In the near future, we plan to demonstrate speed and precision by upgrading EPIC to a GPU-based implementation in order to operate on real-time data and develop a transient trigger and monitor backend.
+In the meanwhile, we plan to demonstrate imaging with non-coplanar arrays and direction-dependent calibration.
 
@@
 
-{\bf We present a methodology to understand the effective angular weighting in the image obtained with EPIC from a heterogeneous array and compare it with that from existing imaging applications which assume a homogeneous array (all antennas are identical).
+We present a methodology to understand the effective angular weighting in the image obtained with EPIC from a heterogeneous array and compare it with that from existing imaging applications which assume a homogeneous array (all antennas are identical).
 % We also demonstrate how the assumption in the latter results in a mis-estimation of the recovered sky model.
@@
 Then, equation~\ref{eqn:apndx-wt-dirty-image-EPIC-decomp} can be written as:
-\begin{align}\label{eqn:apndx-t-dirty-image-EPIC-decomp-approx1}
+\begin{align}\label{eqn:apndx-wt-dirty-image-EPIC-decomp-approx1}
   \mathcal{I}^\prime &\approx \frac{\mathcal{I}_\textrm{D}^\textrm{iso}}{\Nant(\Nant-1)}\,\sum_{ij}\sum_{\substack{ab\\a\ne b\\a\in i, b\in j}} \mathcal{W}^{\textrm{G}\star}_a\,\mathcal{W}^\textrm{G}_b\,\mathcal{W}^\textrm{I}_a\,\mathcal{W}^{\textrm{I}\star}_b,
@@
 where, $\mathcal{I}_\textrm{D}^\textrm{iso}$ is the same for all terms and has been pulled outside the summations.
-For all antennas indexed by $a$ that belong to a particular type $i$, we replace $\mathcal{W}_a$ with $\mathcal{W}_{(i)}$ {\bf and is applicable to those arising} from both instrumental (superscript `$I$') and gridding (superscript `$G$') origins.
+For all antennas indexed by $a$ that belong to a particular type $i$, we replace $\mathcal{W}_a$ with $\mathcal{W}_{(i)}$ and is applicable to those arising from both instrumental (superscript `$I$') and gridding (superscript `$G$') origins.
 Using arguments similar to those in equation~\ref{eqn:apndx-total-baselines}, equation~\ref{eqn:apndx-wt-dirty-image-EPIC-decomp-approx1} can be further simplified to:
@@
 \end{align}
-where the superscript `$\textrm{opt}$' denotes `optimal'.}
+where the superscript `$\textrm{opt}$' denotes `optimal'.
 

### 2016-12-18 imaging_demo/ms.tex
Start addressing second round of referee report
From 25eeda9 (mnras):
====================

(*) Update figures to have uniform properties as per second round of
referee report
Give heterogeneous array imaging analysis context
From 1c821b1 (mnras):
====================

In epic.bib,

(*) Add Iguchi et al. (2009) reference

In ms.tex,

(*) Add a couple of paragraphs in section 5 to set up the context of
heterogeneous array imaging analysis

@@
   \includegraphics[width=\columnwidth]{figure6}
+  % \includegraphics[width=\columnwidth]{MOFF_FX_synthesized_beam_slices}
   \caption{Synthesized beams weighted by antenna power pattern sliced at $m=0$ obtained with MOFF algorithm using EPIC (solid black line) and visibility-based imaging (dashed gray line).
@@
   \includegraphics[width=\columnwidth]{figure7}
-  % \includegraphics[width=\columnwidth]{MOFF_FX_comparison_uvwts_test_aperture_zero_spacing_removed}
-  % \includegraphics[width=\columnwidth]{MOFF_FX_comparison_uvwts_test_aperture_zero_spacing_removed.eps}
+  % \includegraphics[width=\columnwidth]{MOFF_FX_fracdiff_uvwts_test_aperture_zero_spacing_removed}
+  % \includegraphics[width=\columnwidth]{MOFF_FX_fracdiff_uvwts_test_aperture_zero_spacing_removed.eps}
   \caption{{\it Left}: Differences (in \%) in cross-correlated weights in the $uv$ plane relative to the peak.
@@
 Here we present key highlights from the generic methodology detailed in appendix~\ref{sec:math-versatility} to understand the effective angular weighting in the image obtained with EPIC from a heterogeneous array.
+
+{\bf Heterogeneity in antenna arrays occurs primarily in two scenarios.
+The first is when the array is intended to be identical but differences in aperture and electrical properties arise due to deviations during the manufacturing process or from random deviant behaviour such as in dipoles of a phased array.
+The second scenario is when the array is designed to be heterogeneous to gain specific advantages such as sampling a wider range of spatial frequencies and decoupling the sky brightness distribution from the antenna power pattern.
+This has motivated a number of heterogeneous arrays such as the Combined Array for Research in Millimeter-wave Astronomy (CARMA; \citealt{woo04,wri10}), the Atacama Large Millimeter/submillimeter Array (ALMA; \citealt{igu09}), and the Very Long Baseline Array (VLBA).
+In practice, almost all arrays are heterogeneous albeit by varying degrees.
+The following analysis applies to both of the aforementioned scenarios although for purposes of clear demonstration, we will choose an example that is closer to the second scenario.}
 
@@
 This demonstrates that images from EPIC can robustly image data from heterogeneous arrays while being mathematically optimal.
-Besides being sub-optimal, erroneous assumptions about the homogeneity of an array can lead to significant and systematic mis-estimates (in this case, an underestimate) of the sky model.
+Besides being sub-optimal, erroneous assumptions about the homogeneity of an array can lead to significant and systematic mis-estimates ({\bf an underestimate in the above example}) of the sky model.
+
+{\bf In the above example, we chose a scenario wherein the antennas in the array have large differences between them to clearly demonstrate the functionality in EPIC to image data from heterogeneous arrays.
+The analysis, however, is applicable to any heterogeneous array.
+In cases where the differences are smaller, the fidelity with which the sky model is recovered will still be degraded, although to a correspondingly smaller extent.
+To robustly quantify this degradation as a function of deviation from homogeneity will require image deconvolution and will be investigated in the future as it is currently beyond the scope of this paper.}
 
@@
   \includegraphics[width=\columnwidth]{figure12}
-  % \includegraphics[width=\columnwidth]{MOFF_FX_crossover_baseline_n-antennas_rho_fov_gridding_legended}
-  % \includegraphics[width=\columnwidth]{MOFF_FX_crossover_baseline_n-antennas_rho_fov_gridding_legended.eps}
+  % \includegraphics[width=\columnwidth]{MOFF_FX_crossover_baseline_n-antennas_rho_fov_gridding_legended_v2}
+  % \includegraphics[width=\columnwidth]{MOFF_FX_crossover_baseline_n-antennas_rho_fov_gridding_legended_v2.eps}
   \caption{Current and planned instruments in parameter space of baseline length and number of antennas with MOFF and FX.

### 2016-12-20 imaging_demo/ms.tex
Modify figures 1 and A1
From b76b8b9 (mnras):
====================

(*) Modify figures 1 and A1

In ms.tex,

(*) Minor changes in text

@@
 {\bf Heterogeneity in antenna arrays occurs primarily in two scenarios.
-The first is when the array is intended to be identical but differences in aperture and electrical properties arise due to deviations during the manufacturing process or from random deviant behaviour such as in dipoles of a phased array.
+The first is when the array is intended to be identical but differences in aperture and electrical properties arise due to deviations during the manufacturing process or from random deviant behaviour such as differences in gains between dipoles in a phased array.
+In practice, due to this reason almost all arrays are heterogeneous albeit by varying degrees.
 The second scenario is when the array is designed to be heterogeneous to gain specific advantages such as sampling a wider range of spatial frequencies and decoupling the sky brightness distribution from the antenna power pattern.
 This has motivated a number of heterogeneous arrays such as the Combined Array for Research in Millimeter-wave Astronomy (CARMA; \citealt{woo04,wri10}), the Atacama Large Millimeter/submillimeter Array (ALMA; \citealt{igu09}), and the Very Long Baseline Array (VLBA).
-In practice, almost all arrays are heterogeneous albeit by varying degrees.
-The following analysis applies to both of the aforementioned scenarios although for purposes of clear demonstration, we will choose an example that is closer to the second scenario.}
+The following analysis applies to both of the aforementioned scenarios although for purposes of clear demonstration, we will choose an example with large differences across antennas, making it closer to the second scenario.}
 
@@
 
-{\bf In the above example, we chose a scenario wherein the antennas in the array have large differences between them to clearly demonstrate the functionality in EPIC to image data from heterogeneous arrays.
+% {\bf In the above example, we chose a scenario wherein the antennas in the array have large differences between them to clearly demonstrate the functionality in EPIC to image data from heterogeneous arrays.
 The analysis, however, is applicable to any heterogeneous array.
@@
   \caption{Current and planned instruments in parameter space of
-    number of complex multiplies and adds with MOFF and FX. The solid line
+    number of complex {\bf multiplications and additions}. The solid line
     is the boundary at which the number of operations with MOFF and 
@@
   % \includegraphics[width=\columnwidth]{MOFF_FX_crossover_baseline_n-antennas_rho_fov_gridding_legended_v2.eps}
-  \caption{Current and planned instruments in parameter space of baseline length and number of antennas with MOFF and FX.
+  \caption{Current and planned instruments in parameter space of baseline length and number of antennas.
 Lines of different colors denote different classes of antenna sizes (cyan - 1~m$^2$, blue - 7~m$^2$, purple - 16~m$^2$, green - 28~m$^2$, orange - 150~m$^2$, red - 740~m$^2$, gray - 5900~m$^2$).

### 2016-12-20 imaging_demo/ms.tex
Adhere to British spelling
From 0236a60 (mnras):
====================

In ms.tex,

(*) Remove comment on all arrays being heterogeneous
(*) Adhere to British spelling for colour, neighbour, centre, and
grey

@@
 be specified either by a functional form or as a table of values against 
-locations around the antennas. A nearest neighbor mapping from all antenna 
+locations around the antennas. A nearest neighbour mapping from all antenna 
 footprints to grid locations is created using an efficient k-d tree algorithm 
@@
     of unity yielding a peak response of unity in the antenna's directional
-    power pattern on the sky. The color scale for the directional power 
+    power pattern on the sky. The colour scale for the directional power 
     pattern is logarithmic. The black circle indicates the sky horizon and
@@
 We examine in detail the respective synthesized beams in each case in Fig.~\ref{fig:MOFF-FX-psf-diff}.
-Slices at $m=0$ of the synthesized beams weighted by the antenna power pattern are shown for MOFF method using EPIC (solid black) and visibility-based (dashed gray) imaging.
+Slices at $m=0$ of the synthesized beams weighted by the antenna power pattern are shown for MOFF method using EPIC (solid black) and visibility-based (dashed grey) imaging.
 The two are found to match well.
@@
   % \includegraphics[width=\columnwidth]{MOFF_FX_synthesized_beam_slices}
-  \caption{Synthesized beams weighted by antenna power pattern sliced at $m=0$ obtained with MOFF algorithm using EPIC (solid black line) and visibility-based imaging (dashed gray line).
+  \caption{Synthesized beams weighted by antenna power pattern sliced at $m=0$ obtained with MOFF algorithm using EPIC (solid black line) and visibility-based imaging (dashed grey line).
 The two appear almost identical.
@@
 The left panel of Fig.~\ref{fig:MOFF-FX-uvwts-diff} shows differences (in percentage relative to peak) in cross-correlated weights obtained with MOFF imaging in EPIC and visibility-based imaging.
-The maximum difference appears near the center of the $uv$-plane corresponding to antenna auto-correlations which have not been perfectly removed in the former.
+The maximum difference appears near the centre of the $uv$-plane corresponding to antenna auto-correlations which have not been perfectly removed in the former.
 In other regions of the $uv$-plane, the differences are of the order of a few percent.
@@
 The amplitude of the difference appears to be modulated by the directional power response of the antenna.
-On the right, in radial bins, the rms of the synthesized beam (gray) and the rms of the difference map (black) are plotted in percentage units relative to the peak (to be read using the axis on the left side of the plot).
+On the right, in radial bins, the rms of the synthesized beam (grey) and the rms of the difference map (black) are plotted in percentage units relative to the peak (to be read using the axis on the left side of the plot).
 The antenna power pattern (red; to be read using the scale on the right) is plotted for reference.
@@
     their differences (right). The radial variations of rms of the synthesized 
-    beam (gray) and that of the rms of the difference (black) are shown as 
+    beam (grey) and that of the rms of the difference (black) are shown as 
     percentage of the peak synthesized beam. Radially averaged directional 
@@
 The first is when the array is intended to be identical but differences in aperture and electrical properties arise due to deviations during the manufacturing process or from random deviant behaviour such as differences in gains between dipoles in a phased array.
-In practice, due to this reason almost all arrays are heterogeneous albeit by varying degrees.
+% In practice, due to this reason almost all arrays are heterogeneous albeit by varying degrees.
   The second scenario is when the array is designed to be heterogeneous to gain specific advantages such as sampling a wider range of spatial frequencies and decoupling the sky brightness distribution from the antenna power pattern.
@@
 
-In our simulations, the antenna layout, duration of observation, center frequency and bandwidth used are the same as in \S\ref{sec:sim}.
+In our simulations, the antenna layout, duration of observation, centre frequency and bandwidth used are the same as in \S\ref{sec:sim}.
 Antennas have square apertures.
@@
 Black boxes shown around point sources are for purposes of collecting image statistics.
-(b)~Normalized flux density estimates, $\widehat{\mathcal{I}}_\textrm{M}^\textrm{norm}$, as a function of distance from image center.
+(b)~Normalized flux density estimates, $\widehat{\mathcal{I}}_\textrm{M}^\textrm{norm}$, as a function of distance from image centre.
 Bands indicate an approximate ``$1\sigma$'' uncertainty arising from sidelobe confusion and intrinsic randomness of electric fields simulated for point sources in the sky model.
-Gray and red bands correspond to optimal and erroneous cases respectively, where in the latter, all antennas were erroneously assumed to be identical and of the first type.
+Grey and red bands correspond to optimal and erroneous cases respectively, where in the latter, all antennas were erroneously assumed to be identical and of the first type.
 The dashed and solid line styles correspond to point sources in the westward and south-westward directions respectively.
@@
 
-Fig.~\ref{fig:norm-flux-density} shows $\widehat{\mathcal{I}}_\textrm{M}^\textrm{norm}$ including the uncertainty as a function of distance from the center of the image.
-The gray and red bands denote estimates from the optimal and the erroneous cases, respectively.
+Fig.~\ref{fig:norm-flux-density} shows $\widehat{\mathcal{I}}_\textrm{M}^\textrm{norm}$ including the uncertainty as a function of distance from the centre of the image.
+The grey and red bands denote estimates from the optimal and the erroneous cases, respectively.
 The solid and dashed lines correspond to point sources in the south-westward and westward directions respectively.
@@
 similarly a hypothetical trajectory for the MWA with more tiles added inside the 
-same core diameter. The gray shaded area is for a projected LWA expansion and is 
+same core diameter. The grey shaded area is for a projected LWA expansion and is 
 also predominantly in the region favouring MOFF imaging. It is bounded by the LWA1 
@@
     similarly shows hypothetical expanded MWA layouts with more tiles added in 
-    the same core. The gray shaded area denotes the projected trajectory of the 
+    the same core. The grey shaded area denotes the projected trajectory of the 
     LWA bounded by LWA1 (left edge), LWA-OV (right edge), current layout 
@@
 Fig.~\ref{fig:parameter-space-bll-nant-instruments} shows the boundaries where the ratio of the number of computations required with visibility-based imaging relative to MOFF imaging is unity.
-The different colored lines correspond to different antenna sizes (cyan - 1~m$^2$, blue - 7~m$^2$, purple - 16~m$^2$, green - 28~m$^2$, orange - 150~m$^2$, red - 740~m$^2$, gray - 5900~m$^2$).
-Dashed lines of each color denote the boundary to the left of which the MOFF algorithm is favoured for the corresponding antenna size and vice versa for visibility-based imaging.
+The different coloured lines correspond to different antenna sizes (cyan - 1~m$^2$, blue - 7~m$^2$, purple - 16~m$^2$, green - 28~m$^2$, orange - 150~m$^2$, red - 740~m$^2$, grey - 5900~m$^2$).
+Dashed lines of each colour denote the boundary to the left of which the MOFF algorithm is favoured for the corresponding antenna size and vice versa for visibility-based imaging.
 
@@
   \caption{Current and planned instruments in parameter space of baseline length and number of antennas.
-Lines of different colors denote different classes of antenna sizes (cyan - 1~m$^2$, blue - 7~m$^2$, purple - 16~m$^2$, green - 28~m$^2$, orange - 150~m$^2$, red - 740~m$^2$, gray - 5900~m$^2$).
-Lines of each color denote the boundary to the left of which MOFF algorithm is favoured for the corresponding antenna size.
+Lines of different colours denote different classes of antenna sizes (cyan - 1~m$^2$, blue - 7~m$^2$, purple - 16~m$^2$, green - 28~m$^2$, orange - 150~m$^2$, red - 740~m$^2$, grey - 5900~m$^2$).
+Lines of each colour denote the boundary to the left of which MOFF algorithm is favoured for the corresponding antenna size.
 These lines shift rightward with increasing antenna size.
-The different antennas are color coded by roughly the class of antenna size they fall into.
-Thus symbols of one color falling to the left of a line of the same color indicate MOFF imaging is advantageous for those telescopes and vice versa.
+The different antennas are colour coded by roughly the class of antenna size they fall into.
+Thus symbols of one colour falling to the left of a line of the same colour indicate MOFF imaging is advantageous for those telescopes and vice versa.
 For example, MOFF imaging is favoured in HERA-331 and HERA-6769 because they lie to the left of the orange line but not so in cases of HERA-19 and HERA-37.}
@@
 
-The different antennas are color coded by roughly the class of antenna size they fall into.
-Thus symbols of one color that lie to the left of same colored line indicate MOFF imaging is favoured for those telescopes and vice versa.
+The different antennas are colour coded by roughly the class of antenna size they fall into.
+Thus symbols of one colour that lie to the left of same coloured line indicate MOFF imaging is favoured for those telescopes and vice versa.
 For example, MOFF imaging is favoured in HERA-331 and HERA-6769 because they lie to the left of the orange line but not so in cases of HERA-19 and HERA-37.

### 2017-01-06 imaging_demo/ms.tex
Add keyword 'telescopes', remove unused comments
From 01314a1 (mnras):
====================

In ms.tex,

(*) Add 'telescopes' to list of keywords
(*) Remove unused comments
Emphasize heterogeneous array imaging framework
From e50a08f (mnras):
====================

In ms.tex,

(*) Emphasize the extra message that a mathematical framework is
provided for imaging with heterogeneous arrays in abstract, intro
and conclusions
(*) Remove erroneous superscript in the estimated weights in equation
B15 in appendix B
(*) Strikethrough the sentence on "taking advantage of the
multiplication-convolution th

@@
 \usepackage[caption=false]{subfig} % For subfloats
+\usepackage[normalem]{ulem} % For strikethrough text
 
@@
 Through our generic and efficient E-field Parallel Imaging Correlator (EPIC), we present the first software demonstration of a generalized direct imaging algorithm, namely, the Modular Optimal Frequency Fourier (MOFF) imager.
-It takes advantage of the multiplication-convolution theorem of Fourier transforms.
-Not only does it bring down the cost for dense layouts to $\mathcal{O}(\Nant\log_2\Nant)$ but can also image from irregular layouts and heterogeneous arrays of antennas.
+\sout{It takes advantage of the multiplication-convolution theorem of Fourier transforms.} Not only does it bring down the cost for dense layouts to $\mathcal{O}(\Nant\log_2\Nant)$ but can also image from irregular layouts and heterogeneous arrays of antennas.
 EPIC is highly modular, parallelizable, implemented in object-oriented Python, and publicly available.
@@
 We have also validated our implementation on data observed with the Long Wavelength Array (LWA).
-We show that it robustly estimates the input sky model for heterogeneous arrays.
+{\bf We provide a detailed framework for imaging with heterogeneous arrays and show that EPIC} robustly estimates the input sky model for {\bf such} arrays.
 Antenna layouts with dense filling factors consisting of a large number of antennas such as LWA, the Square Kilometre Array, Hydrogen Epoch of Reionization Array, and Canadian Hydrogen Intensity Mapping Experiment will gain significant computational advantage by deploying an optimized version of EPIC.
@@
 \begin{keywords}
-instrumentation: interferometers -- techniques: image processing -- techniques: interferometric
+instrumentation: interferometers -- techniques: image processing -- techniques: interferometric -- telescopes
 \end{keywords}
@@
 
-This paradigm shift requires a fundamentally new approach to the design of
-digital correlators \citep{lon00}. Modern correlators calculate the cross-power
-correlation between all antenna pairs in many narrow frequencies, forming
-\emph{visibilities}, the fundamental measurement of traditional radio
-interferometers. The computational requirements for a modern FX correlator scale
-with the number of antenna pairs, or the square of the number of antennas $\sim
-\Nant^2$ \citep{bun04}. For this reason traditional correlators have difficulty
-scaling to thousands of antennas. As an example, the full HERA correlator for
-352 dishes with 200 MHz of bandwidth requires 212 trillion complex multiplies
-and adds per second (TMACS). Future arrays with thousands of collecting elements
-will require orders of magnitude more computation, making the correlator the
-dominant cost.
-
-For certain classes of radio arrays there is an alternative to the FX correlator
-that can lower the computational burden by directly performing a spatial Fast
-Fourier Transform \citep[FFT;][]{coo65} on the electric fields measured by each 
-antenna in the array at each time step, removing the cross-correlation step. 
-This relieves the computational scaling from the harsh $\Nant^2$ to the more 
-gentle envelope of $\sim\Ngrid\log_2\Ngrid$, where $\Ngrid$ is the number of 
-grid points in the Fourier transform \citep[e.g.][]{mor11,teg09,teg10}. This 
-architecture is often referred to as a ``direct imaging correlator'' because it 
-eliminates the intermediate cross-correlation data products of the FX and XF 
-correlators, but instead directly forms images from the electric field 
-measurements.
+This paradigm shift requires a fundamentally new approach to the design of digital correlators \citep{lon00}.
+Modern correlators calculate the cross-power correlation between all antenna pairs in many narrow frequencies, forming \emph{visibilities}, the fundamental measurement of traditional radio interferometers.
+The computational requirements for a modern FX correlator scale with the number of antenna pairs, or the square of the number of antennas $\sim \Nant^2$ \citep{bun04}.
+For this reason traditional correlators have difficulty scaling to thousands of antennas.
+As an example, the full HERA correlator for 352 dishes with 200 MHz of bandwidth requires 212 trillion complex multiply-accumulates per second (TMACS).
+Future arrays with thousands of collecting elements will require orders of magnitude more computation, making the correlator the dominant cost.
+
+For certain classes of radio arrays there is an alternative to the FX correlator that can lower the computational burden by directly performing a spatial Fast Fourier Transform \citep[FFT;][]{coo65} on the electric fields measured by each antenna in the array at each time step, removing the cross-correlation step.
+This relieves the computational scaling from the harsh $\Nant^2$ to the more gentle envelope of $\sim\Ngrid\log_2\Ngrid$, where $\Ngrid$ is the number of grid points in the Fourier transform \citep[e.g.][]{mor11,teg09,teg10}.
+This architecture is often referred to as a ``direct imaging correlator'' because it eliminates the intermediate cross-correlation data products of the FX and XF correlators, but instead directly forms images from the electric field measurements.
 
@@
 
-Here we present the first software implementation of the MOFF correlator, and
-announce the public release of the E-field Parallel Imaging Correlator (EPIC)
-code. EPIC is a highly parallel, object oriented Python package that primarily 
-implements the MOFF imaging algorithm as well as emulating real-life telescopes and 
-FX/XF correlators in software, and includes a visibility-based imaging technique 
-for reference. It is intended to provide a development platform to test different 
-imaging approaches, characterize scaling relations and serve as a stepping stone 
-for real-life GPU/FPGA-based implementation on telescopes.
+Here we present the first software implementation of the MOFF correlator, and announce the public release of the E-field Parallel Imaging Correlator (EPIC) code.
+EPIC is a highly parallel, object-oriented Python package that primarily implements the MOFF imaging algorithm, \sout{as well as emulating} {\bf emulates} real-life telescopes and FX/XF correlators in software, and includes a visibility-based imaging technique for reference.
+It is intended to provide a development platform to test different imaging approaches, characterize scaling relations and serve as a stepping stone for real-life GPU/FPGA-based implementation on telescopes.
 
@@
 We also demonstrate the performance with real-world data from the LWA.
-\S\ref{sec:versatility} demonstrates its ability to robustly image data from heterogeneous arrays.
+{\bf In \S\ref{sec:versatility}, we present a mathematical framework for imaging with heterogeneous arrays and demonstrate the ability of EPIC to robustly image data from such arrays}.
 In \S\ref{sec:analysis}, we explore the scalability of the algorithm in the context of several array design choices.
@@
 
-We provide a brief summary of the mathematical equivalence of the MOFF and 
-FX correlators detailed in \citet{mor11}. We first relate the image produced 
-from visibilities to the electric fields of astrophysical sources, then show that 
-operations can be reordered to produce the same images at a lower 
-computational cost.
+We provide a brief summary of the mathematical equivalence of the MOFF and FX correlators detailed in \citet{mor11}.
+We first relate the image produced from visibilities to the electric fields of astrophysical sources, then show that operations can be reordered to produce the same images at a lower computational cost.
 
@@
 \begin{align}
-  E(\mathbf{r}, f, t) &= \int \mathcal{E}(\hat{\mathbf{s}},f,t)\,e^{-2\pi i f\mathbf{r}\cdot\hat{\mathbf{s}}/c}\,\dif\Omega,
-\end{align}
-where, $\mathbf{r}$ denotes the observer's location, $f$ is the frequency of radiation, $c$ is the speed of light, $t$ denotes time, $\dif\Omega$ is the infinitesimal solid angle element on the celestial sphere whose normal is the unit vector $\hat{\mathbf{s}}$, and $E(\mathbf{r}, f, t)$ is the propagated electric field.
+  E(\mathbf{r}, f, t) &= \int \mathcal{E}(\hat{\mathbf{s}},f,t)\,e^{-i 2\pi f\mathbf{r}\cdot\hat{\mathbf{s}}/c}\,\dif\Omega,
+\end{align}
+where, $i=\sqrt{-1}$, $\mathbf{r}$ denotes the observer's location, $f$ is the frequency of radiation, $c$ is the speed of light, $t$ denotes time, $\dif\Omega$ is the infinitesimal solid angle element on the celestial sphere whose normal is the unit vector $\hat{\mathbf{s}}$, and $E(\mathbf{r}, f, t)$ is the propagated electric field.
 Thus the propagated electric field is a linear superposition of the electric fields emanating from astronomical sources with appropriate complex phases.
@@
            &= \int W_a(\mathbf{r}-\mathbf{r}_a,f,t) \nonumber \\
-           &\quad \times \left[ \int \mathcal{E}(\hat{\mathbf{s}},f,t)\,e^{-2\pi i f\mathbf{r}\cdot\hat{\mathbf{s}}/c}\,\dif\Omega \right] \dif^2\mathbf{r} + n_a(f,t) \\
-           &= \int \mathcal{W}_a(\hat{\mathbf{s}},f,t)\,\mathcal{E}(\hat{\mathbf{s}},f,t)\,e^{-2\pi i f\mathbf{r}_a\!\cdot\,\hat{\mathbf{s}}/c}\,\dif\Omega + n_a(f,t)
+           &\quad \times \left[ \int \mathcal{E}(\hat{\mathbf{s}},f,t)\,e^{-i 2\pi f\mathbf{r}\cdot\hat{\mathbf{s}}/c}\,\dif\Omega \right] \dif^2\mathbf{r} + n_a(f,t) \\
+           &= \int \mathcal{W}_a(\hat{\mathbf{s}},f,t)\,\mathcal{E}(\hat{\mathbf{s}},f,t)\,e^{-i 2\pi f\mathbf{r}_a\!\cdot\,\hat{\mathbf{s}}/c}\,\dif\Omega + n_a(f,t)
 \end{align}
@@
   V_{ab}(f,t) &= \left\langle E_a(f,t)E_b^\star(f,t) \right\rangle_t \label{eqn:cc-vis}\hfill\\
-              &= \left\langle \left[ \int \mathcal{W}_a(\hat{\mathbf{s}},f,t)\,\mathcal{E}(\hat{\mathbf{s}},f,t)\,e^{-2\pi i f\mathbf{r}_a\!\cdot\,\hat{\mathbf{s}}/c}\,\dif\Omega\right.\right.
+              &= \left\langle \left[ \int \mathcal{W}_a(\hat{\mathbf{s}},f,t)\,\mathcal{E}(\hat{\mathbf{s}},f,t)\,e^{-i 2\pi f\mathbf{r}_a\!\cdot\,\hat{\mathbf{s}}/c}\,\dif\Omega\right.\right.
 \nonumber\\
               &\qquad\qquad\qquad\qquad\qquad\qquad\qquad\quad + \left. n_a(f,t) \vphantom{\int}\right] \nonumber\\
-              &\quad \times \left[ \int \mathcal{W}^\star_b(\hat{\mathbf{s'}},f,t)\,\mathcal{E}^\star (\hat{\mathbf{s'}},f,t)\,e^{2\pi i f\mathbf{r}_b\!\cdot\,\hat{\mathbf{s'}}/c}\,\dif\Omega^\prime \right.
+              &\quad \times \left[ \int \mathcal{W}^\star_b(\hat{\mathbf{s'}},f,t)\,\mathcal{E}^\star (\hat{\mathbf{s'}},f,t)\,e^{i 2\pi f\mathbf{r}_b\!\cdot\,\hat{\mathbf{s'}}/c}\,\dif\Omega^\prime \right.
 \nonumber\\
@@
               &= \iint \mathcal{W}_a(\hat{\mathbf{s}},f,t) \mathcal{W}^\star_b(\hat{\mathbf{s'}},f,t) \left\langle \mathcal{E}(\hat{\mathbf{s}},f,t) \mathcal{E}^\star(\hat{\mathbf{s'}},f,t) \right\rangle_t \nonumber\\
-              &\qquad\qquad\qquad\qquad \times e^{-2\pi i f(\mathbf{r}_a\!\cdot\,\hat{\mathbf{s}}-\mathbf{r}_b\!\cdot\,\hat{\mathbf{s'}})/c}\,\dif\Omega\,\dif\Omega^\prime \nonumber\\
+              &\qquad\qquad\qquad\qquad \times e^{-i 2\pi f(\mathbf{r}_a\!\cdot\,\hat{\mathbf{s}}-\mathbf{r}_b\!\cdot\,\hat{\mathbf{s'}})/c}\,\dif\Omega\,\dif\Omega^\prime \nonumber\\
               &\qquad +\textrm{noise-like cross-terms.}
@@
 \begin{align}\label{eqn:measurement-eqn-1}
-  V_{ab}(f,t) &= \int \mathcal{B}_{ab}(\hat{\mathbf{s}},f,t)\,\mathcal{I}(\hat{\mathbf{s}},f,t)\,e^{-2\pi i f\mathbf{r}_{ab}\!\cdot\,\hat{\mathbf{s}}/c}\,\dif\Omega \nonumber\\
+  V_{ab}(f,t) &= \int \mathcal{B}_{ab}(\hat{\mathbf{s}},f,t)\,\mathcal{I}(\hat{\mathbf{s}},f,t)\,e^{-i 2\pi f\mathbf{r}_{ab}\!\cdot\,\hat{\mathbf{s}}/c}\,\dif\Omega \nonumber\\
   &\qquad\qquad + n_{ab}(f,t),
@@
   V_{ab}(f,t) &= \int B_{ab}(\mathbf{r}_{ab}-\mathbf{r},f,t) \nonumber\\ 
-              &\qquad \times \left[\int I(\hat{\mathbf{s}},f,t)\,e^{-2\pi i f\mathbf{r}.\hat{\mathbf{s}}/c}\,\dif\Omega\right]\dif^2\mathbf{r}\,+\, n_{ab}(f,t),
+              &\qquad \times \left[\int I(\hat{\mathbf{s}},f,t)\,e^{-i 2\pi f\mathbf{r}.\hat{\mathbf{s}}/c}\,\dif\Omega\right]\dif^2\mathbf{r}\,+\, n_{ab}(f,t),
 \end{align}
@@
 
-A high level software architecture of the EPIC package is described in the 
-appendix~\ref{sec:software-modules} for the interested reader. Here we discuss the components of these architectures in detail.
+A high level software architecture of the EPIC package is described in the appendix~\ref{sec:software-modules} for the interested reader.
+Here we discuss the components of these architectures in detail.
 
@@
 \par\medskip
-\noindent A grid is generated on the coordinate system in which antenna 
-locations are specified. The grid spacing can be controlled 
-by the user. By default, it is set to be $\le\lambda/2$ even at the 
-lowest wavelength to ensure there is no aliasing even from regions of the sky 
-far away from the field of view. The number of locations on the grid is 
-restricted to be a power of 2 for efficient use of FFT.
-
-The gridding kernel in the simplest case is given by the antenna aperture
-illumination function, $W(\mathbf{r}-\mathbf{r}_a)$, which can
-be specified either by a functional form or as a table of values against 
-locations around the antennas. A nearest neighbour mapping from all antenna 
-footprints to grid locations is created using an efficient k-d tree algorithm 
-\citep{man99}. There is no restriction here that the aperture illumination 
-function has to be identical across antennas.
-
-In the most general case, this gridding kernel could contain information on
-$w$-projection effects, and other time-dependent ionospheric effects. For a
-stationary antenna array in the absence of any time-dependent effects, this
-mapping must only be determined once in the antenna array coordinate frame. The
-antenna-to-grid mapping matrix, $\mathbf{M}(\mathbf{r},\mathbf{a})$ is described
-as a transformation matrix from the space of measured electric fields by the 
-antennas ($\mathbf{a}$) to the antenna array grid denoted by the coordinate 
-$\mathbf{r}$. Since each antenna occupies a footprint typically the size of its 
-aperture, $\mathbf{M}(\mathbf{r},\mathbf{a})$, which is generally of size
-$\Ngrid\times \Nant$, reduces to a sparse block-diagonal matrix
-with only $\Nant$ blocks and roughly $N_\textrm{k}$ non-zero entries per
-block, where $N_\textrm{k}$ is the number of grid points that fall inside an
-antenna's footprint. This sparse matrix is stored in a Compressed Sparse Row 
-(CSR) format. Fig.~\ref{fig:a2g-mapping} illustrates the antenna-to-grid mapping
-matrix and the grid containing the mapped aperture footprints of the antennas.
+\noindent A grid is generated on the coordinate system in which antenna locations are specified.
+The grid spacing can be controlled by the user.
+By default, it is set to be $\le\lambda/2$ even at the lowest wavelength to ensure there is no aliasing even from regions of the sky far away from the field of view.
+The number of locations on the grid is restricted to be a power of 2 for efficient use of FFT.
+
+The gridding kernel in the simplest case is given by the antenna aperture illumination function, $W(\mathbf{r}-\mathbf{r}_a)$, which can be specified either by a functional form or as a table of values against locations around the antennas.
+A nearest neighbour mapping from all antenna footprints to grid locations is created using an efficient k-d tree algorithm \citep{man99}.
+There is no restriction here that the aperture illumination function has to be identical across antennas.
+
+In the most general case, this gridding kernel could contain information on $w$-projection effects, and other time-dependent ionospheric effects.
+For a stationary antenna array in the absence of any time-dependent effects, this mapping must only be determined once in the antenna array coordinate frame.
+The antenna-to-grid mapping matrix, $\mathbf{M}(\mathbf{r},\mathbf{a})$ is described as a transformation matrix from the space of measured electric fields by the antennas ($\mathbf{a}$) to the antenna array grid denoted by the coordinate $\mathbf{r}$.
+Since each antenna occupies a footprint typically the size of its aperture, $\mathbf{M}(\mathbf{r},\mathbf{a})$, which is generally of size $\Ngrid\times \Nant$, reduces to a sparse block-diagonal matrix with only $\Nant$ blocks and roughly $N_\textrm{k}$ non-zero entries per block, where $N_\textrm{k}$ is the number of grid points that fall inside an antenna's footprint.
+This sparse matrix is stored in a Compressed Sparse Row (CSR) format.
+Fig.~\ref{fig:a2g-mapping} illustrates the antenna-to-grid mapping matrix and the grid containing the mapped aperture footprints of the antennas.
 
@@
   % \includegraphics[width=\columnwidth]{a2g_mapping.eps}
-  \caption{Block diagram of an antenna-to-grid mapping. A sparse block-diagonal
-    matrix of total size $\Ngrid\times \Nant$ is created where each
-    block contains roughly the number of pixels covered by the respective kernel.
-    The antenna aperture illumination kernels do not have to be identical to each
-    other. A discrete set of arbitrarily placed antennas are now placed on to a
-    regular grid.}
+  \caption{Block diagram of an antenna-to-grid mapping.
+A sparse block-diagonal     matrix of total size $\Ngrid\times \Nant$ is created where each block contains roughly the number of pixels covered by the respective kernel.
+The antenna aperture illumination kernels do not have to be identical to each other.
+A discrete set of arbitrarily placed antennas are now placed on to a regular grid.}
   \label{fig:a2g-mapping}
@@
 \par\medskip
-\noindent This module is common to the MOFF and visibility-based imaging 
-techniques. Time samples of electric fields measured by the antenna and 
-digitized by the A/D converter is Fourier transformed to generate electric 
-field spectra. This step can be parallelized across antennas%  as shown in 
-% Fig.~\ref{fig:f-engine}
-. The output is then fed to either MOFF or 
-visibility-based imaging pipelines.
-
-% \begin{figure}
-%   \includegraphics[width=\columnwidth]{figure4}
-%   % \includegraphics[width=\columnwidth]{parallel_F_engine}
-%   % \includegraphics[width=\columnwidth]{parallel_F_engine.eps}
-%   \caption{Block diagram of a F-engine. The electric field data streams from
-%     antennas are Fourier transformed in parallel to generate electric field
-%     spectra.}
-%   \label{fig:f-engine}
-% \end{figure}
+\noindent This module is common to the MOFF and visibility-based imaging techniques.
+Time samples of electric fields measured by the antenna and digitized by the A/D converter is Fourier transformed to generate electric field spectra.
+This step can be parallelized across antennas.
+The output is then fed to either MOFF or visibility-based imaging pipelines.
 
@@
 \par\medskip
-\noindent Calibration of direct imaging correlators remains a challenge. Contrary
-to the FX data flow, direct imagers mix the signals from all antennas before
-averaging and writing to disk. It is therefore essential to apply gain solutions 
-before the gridding step. Previous efforts have resorted to applying FX-generated
-calibration solutions \citep{zhe14,fos14}, or integrating a dedicated FX 
-correlator which periodically forms the full visibility matrix 
-\citep{wij09,dev09}.
-
-In a companion paper \citep{bea16}, we demonstrate a novel calibration 
-technique (EPICal) which leverages the data products formed by direct imaging 
-correlators to estimate antenna complex gains. This method correlates the antenna 
-electric field signals with an image pixel from the output of the correlator in 
-the feedback calibration fashion outlined in \citealt{mor11} (illustrated in 
-Fig.~\ref{fig:MOFF-flowchart} by the arrow leading from the imager to the 
-calibration block). Furthermore it allows for arbitrarily complex sky models, 
-and following the MOFF algorithm places no restriction on array layout, and 
-accounts for non-identical antenna beam patterns. Direction dependent calibration 
-can be achieved by correlating antenna signals with output pixels in the 
-direction of $N_\textrm{c}$ calibration sources, then fitting for a functional 
-model of the sky. Since antennas are only correlated with calibrator pixels, the 
-computational complexity scales as $\sim N_\textrm{a} N_\textrm{c}$. The calibration performance of EPIC is explored for varying levels of completeness of sky model in \citet{bea16}.
-
-The calibration module included in EPIC allows for application of pre-determined 
-calibration solutions, or can solve for the complex gains using the EPICal 
-algorithm.
+\noindent Calibration of direct imaging correlators remains a challenge.
+Contrary to the FX data flow, direct imagers mix the signals from all antennas before averaging and writing to disk.
+It is therefore essential to apply gain solutions before the gridding step.
+Previous efforts have resorted to applying FX-generated calibration solutions \citep{zhe14,fos14}, or integrating a dedicated FX correlator which periodically forms the full visibility matrix \citep{wij09,dev09}.
+
+In a companion paper \citep{bea16}, we demonstrate a novel calibration technique (EPICal) which leverages the data products formed by direct imaging correlators to estimate antenna complex gains.
+This method correlates the antenna electric field signals with an image pixel from the output of the correlator in the feedback calibration fashion outlined in \citealt{mor11} (illustrated in Fig.~\ref{fig:MOFF-flowchart} by the arrow leading from the imager to the calibration block).
+Furthermore it allows for arbitrarily complex sky models, and following the MOFF algorithm places no restriction on array layout, and accounts for non-identical antenna beam patterns.
+Direction dependent calibration can be achieved by correlating antenna signals with output pixels in the direction of $N_\textrm{c}$ calibration sources, then fitting for a functional model of the sky.
+Since antennas are only correlated with calibrator pixels, the computational complexity scales as $\sim N_\textrm{a} N_\textrm{c}$.
+The calibration performance of EPIC is explored for varying levels of completeness of sky model in \citet{bea16}.
+
+The calibration module included in EPIC allows for application of pre-determined calibration solutions, or can solve for the complex gains using the EPICal algorithm.
 
@@
 \par\medskip
-\noindent Before the spatial Fourier transform, the gridded electric fields are 
-padded with zeros in order to match the grid size and angular size of each image 
-pixel that would have been obtained with the software holography output from an 
-FX correlator.
-
-In MOFF imaging, these are spatially Fourier transformed followed by a squaring
-operation at every time stamp for every frequency channel. In visibility-based 
-imaging, the spatial Fourier transform is performed only once per integration 
-time-scale and does not include a squaring operation.
+\noindent Before the spatial Fourier transform, the gridded electric fields are padded with zeros in order to match the grid size and angular size of each image pixel that would have been obtained with the software holography output from an FX correlator.
+
+In MOFF imaging, these are spatially Fourier transformed followed by a squaring operation at every time stamp for every frequency channel.
+In visibility-based imaging, the spatial Fourier transform is performed only once per integration time-scale and does not include a squaring operation.
 
@@
 \par\medskip
-\noindent In MOFF imaging, the measured antenna electric fields and the 
-corresponding holographic electric field images are zero-mean stochastic 
-quantities. Hence, they cannot be time-averaged to reduce noise. The statistical 
-quantity stable with time in this case are the square of the holographic 
-electric field images. Thus, squared images have to be formed at every instant 
-of time before averaging as indicated in equation~\ref{eqn:dirty-image-MOFF}.
-
-In contrast, visibilities measured by an antenna are statistically stable within
-an integration time interval. Hence, they are averaged after calibration as shown
-in equation~\ref{eqn:cc-vis}. It is advantageous to average them in visibilities
-before imaging because the repeated cost of spatial FFT can be avoided. Since 
-this averaging has been performed already on the visibilities over an integration 
-time-scale, the imaging step has to be performed only once per integration cycle.
+\noindent In MOFF imaging, the measured antenna electric fields and the corresponding holographic electric field images are zero-mean stochastic quantities.
+Hence, they cannot be time-averaged to reduce noise.
+The statistical quantity stable with time in this case are the square of the holographic electric field images.
+Thus, squared images have to be formed at every instant of time before averaging as indicated in equation~\ref{eqn:dirty-image-MOFF}.
+
+In contrast, visibilities measured by an antenna are statistically stable within an integration time interval.
+Hence, they are averaged after calibration as shown in equation~\ref{eqn:cc-vis}.
+It is advantageous to average them in visibilities before imaging because the repeated cost of spatial FFT can be avoided.
+Since this averaging has been performed already on the visibilities over an integration time-scale, the imaging step has to be performed only once per integration cycle.
 
@@
 
-In order to verify the accuracy of the EPIC code, we characterize the images 
-produced through simulations. We simulate electric field streams from a model 
-sky and process the data through both the MOFF and a visibility-based imaging 
-algorithm. We then compare the output images to demonstrate their equivalence.
+In order to verify the accuracy of the EPIC code, we characterize the images produced through simulations.
+We simulate electric field streams from a model sky and process the data through both the MOFF and a visibility-based imaging algorithm.
+We then compare the output images to demonstrate their equivalence.
 
@@
 
-The shape and extent of these auto-correlations can be estimated from the 
-antenna aperture illumination pattern. The aperture illumination patterns
-are already available from the gridding step. Fig.~\ref{fig:autocorr_wts_PB} 
-shows the estimated weights from antenna auto-correlations in the $uv$-plane 
-(left) and the corresponding response in the image plane (right). The latter 
-is simply the directional antenna power response.
+The shape and extent of these auto-correlations can be estimated from the antenna aperture illumination pattern.
+The aperture illumination patterns are already available from the gridding step.
+Fig.~\ref{fig:autocorr_wts_PB} shows the estimated weights from antenna auto-correlations in the $uv$-plane (left) and the corresponding response in the image plane (right).
+The latter is simply the directional antenna power response.
 
@@
   % \includegraphics[width=\columnwidth]{autocorr_uvwts_pbeam.png}
-  \caption{Auto-correlation of weights of a square shaped antenna aperture
-    in the $uv$ plane (left) and the corresponding directional antenna power 
-    response on the sky (right) in coordinates specified by direction cosines. 
-    The antenna auto-correlation weights are normalized to a sum
-    of unity yielding a peak response of unity in the antenna's directional
-    power pattern on the sky. The colour scale for the directional power 
-    pattern is logarithmic. The black circle indicates the sky horizon and
-    values beyond it are not physical and hence ignored.}
+  \caption{Auto-correlation of weights of a square shaped antenna aperture in the $uv$ plane (left) and the corresponding directional antenna power response on the sky (right) in coordinates specified by direction cosines.
+The antenna auto-correlation weights are normalized to a sum of unity yielding a peak response of unity in the antenna's directional power pattern on the sky.
+The colour scale for the directional power pattern is logarithmic.
+The black circle indicates the sky horizon and values beyond it are not physical and hence ignored.}
   \label{fig:autocorr_wts_PB}
@@
 
-We investigate the two imaging algorithms for differences from the point of 
-view of the quality of their outputs. We begin by comparing the images produced 
-with the two approaches.
+We investigate the two imaging algorithms for differences from the point of view of the quality of their outputs. We begin by comparing the images produced with the two approaches.
 
@@
   % \includegraphics[width=\columnwidth]{diff_psf_MOFF-FX_test_aperture.eps}
-  \caption{Map of difference between the synthesized beams obtained with the 
-    two methods (left) and radial statistics of the synthesized beams and 
-    their differences (right). The radial variations of rms of the synthesized 
-    beam (grey) and that of the rms of the difference (black) are shown as 
-    percentage of the peak synthesized beam. Radially averaged directional 
-    antenna power response in absolute scale is shown in red and is to be read 
-    with the scale on the right side of the axis. The maximum difference is of 
-    the order of a few percent. Amplitude of the rms of the difference is 
-    modulated by the power pattern of the antenna.}
+  \caption{Map of difference between the synthesized beams obtained with the two methods (left) and radial statistics of the synthesized beams and their differences (right).
+The radial variations of rms of the synthesized beam (grey) and that of the rms of the difference (black) are shown as percentage of the peak synthesized beam.
+Radially averaged directional antenna power response in absolute scale is shown in red and is to be read with the scale on the right side of the axis.
+The maximum difference is of the order of a few percent.
+Amplitude of the rms of the difference is modulated by the power pattern of the antenna.}
   \label{fig:image-psf-diff}
@@
 
-The synthesized beam rms is proportional to the antenna power pattern as 
-expected from a point spread function uncorrected for the antenna power pattern. 
-The rms of differenced synthesized beams is also modulated by the antenna power 
-pattern. The rms of the difference is definitely lesser than the rms of the 
-synthesized beam in the central regions up to $(l^2+m^2)^{1/2}\lesssim 0.3$. This 
-implies that the beams are well matched in the central regions. In the outer 
-regions, their mismatch is comparable to the rms of synthesized beams. These findings are consistent with Fig.~\ref{fig:MOFF-FX-psf-diff}. This 
-indicates the two synthesized beams are not randomly different from 
-each other in which case the rms of the difference would have been 
-$\approx \sqrt{2}$ higher than the rms of the each of the synthesized beams. 
-This means that while differences exist, large fractions of them are still 
-well matched to each other even out to the horizon. Thus the rounding errors in 
-gridding do not affect the statistics of the images or the synthesized beams.
+The synthesized beam rms is proportional to the antenna power pattern as expected from a point spread function uncorrected for the antenna power pattern.
+The rms of differenced synthesized beams is also modulated by the antenna power pattern.
+The rms of the difference is definitely lesser than the rms of the synthesized beam in the central regions up to $(l^2+m^2)^{1/2}\lesssim 0.3$.
+This implies that the beams are well matched in the central regions.
+In the outer regions, their mismatch is comparable to the rms of synthesized beams.
+These findings are consistent with Fig.~\ref{fig:MOFF-FX-psf-diff}.
+This indicates the two synthesized beams are not randomly different from each other in which case the rms of the difference would have been $\approx \sqrt{2}$ higher than the rms of the each of the synthesized beams.
+This means that while differences exist, large fractions of them are still well matched to each other even out to the horizon.
+Thus the rounding errors in gridding do not affect the statistics of the images or the synthesized beams.
 
@@
 
-Here we demonstrate our software using narrow band data from the LWA station 
-in New Mexico. This data is in LWA narrow-band transient buffer (TBN) format 
-from 255 antennas within roughly a diameter of 100~m. The data is centred at a 
-frequency of 74.03~MHz, with a sample rate (equal to the bandwidth) of 100~kHz 
-with 512 complex time samples per antenna in a A/D writeout time-scale of 
-5.12~ms, a frequency resolution of 195.3125~Hz and dual polarization. There are 
-391 such writeouts (or time stamps; each contains 512 time samples at 100~kHz 
-sampling) yielding a total duration of 2~s.
+Here we demonstrate our software using narrow band data from the LWA station in New Mexico.
+This data is in LWA narrow-band transient buffer (TBN) format from 255 antennas within roughly a diameter of 100~m.
+The data is centred at a frequency of 74.03~MHz, with a sample rate (equal to the bandwidth) of 100~kHz with 512 complex time samples per antenna in a A/D writeout time-scale of 5.12~ms, a frequency resolution of 195.3125~Hz and dual polarization.
+There are 391 such writeouts (or time stamps; each contains 512 time samples at 100~kHz sampling) yielding a total duration of 2~s.
 
@@
 
-Fig.~\ref{fig:LWA-image} shows the image produced with MOFF imaging 
-packaged in EPIC after averaging over $\approx$~20~ms (four writeouts) of data 
-and the inner $\approx 80$ per cent of bandwidth (roughly 80~kHz). The image is 
-shown in direction cosine coordinates -- $l$ along the horizontal axis and $m$ 
-along the vertical axis. The flux scale is arbitrary. Even in this 
-proof-of-concept demonstration, we see Cyg~A and Cas~A prominently as annotated, 
-thus validating the functionality of EPIC.
+Fig.~\ref{fig:LWA-image} shows the image produced with MOFF imaging packaged in EPIC after averaging over $\approx$~20~ms (four writeouts) of data and the inner $\approx 80$ per cent of bandwidth (roughly 80~kHz).
+The image is shown in direction cosine coordinates -- $l$ along the horizontal axis and $m$ along the vertical axis.
+The flux scale is arbitrary.
+Even in this proof-of-concept demonstration, we see Cyg~A and Cas~A prominently as annotated, thus validating the functionality of EPIC.
 
@@
   % \includegraphics[width=\columnwidth]{LWA_MOFF_bandavg_image_4_iterations_analytic_aperture.eps}
-  \caption{Image from LWA TBN data obtained with MOFF imaging using EPIC 
-    package after averaging over 20~ms and $\approx 80$~kHz. The x- and y-axes
-    denote direction cosines $l$ and $m$ respectively. The antenna voltages
-    are compensated for their respective delays. The flux scale is arbitrary.
+  \caption{Image from LWA TBN data obtained with MOFF imaging using EPIC package after averaging over 20~ms and $\approx 80$~kHz.
+The x- and y-axes denote direction cosines $l$ and $m$ respectively.
+The antenna voltages are compensated for their respective delays.
+The flux scale is arbitrary.
 Locations of Cyg~A and Cas~A are annotated.}
@@
 
-{\bf Heterogeneity in antenna arrays occurs primarily in two scenarios.
+Heterogeneity in antenna arrays occurs primarily in two scenarios.
 The first is when the array is intended to be identical but differences in aperture and electrical properties arise due to deviations during the manufacturing process or from random deviant behaviour such as differences in gains between dipoles in a phased array.
-% In practice, due to this reason almost all arrays are heterogeneous albeit by varying degrees.
 The second scenario is when the array is designed to be heterogeneous to gain specific advantages such as sampling a wider range of spatial frequencies and decoupling the sky brightness distribution from the antenna power pattern.
 This has motivated a number of heterogeneous arrays such as the Combined Array for Research in Millimeter-wave Astronomy (CARMA; \citealt{woo04,wri10}), the Atacama Large Millimeter/submillimeter Array (ALMA; \citealt{igu09}), and the Very Long Baseline Array (VLBA).
-The following analysis applies to both of the aforementioned scenarios although for purposes of clear demonstration, we will choose an example with large differences across antennas, making it closer to the second scenario.}
+The following analysis applies to both of the aforementioned scenarios although for purposes of clear demonstration, we will choose an example with large differences across antennas, making it closer to the second scenario.
 
@@
 \begin{align}
-  \mathcal{I}^\prime_{ab}(\hat{\mathbf{s}},f) &= \mathcal{W}^{\textrm{G}\star}_a(\hat{\mathbf{s}},f)\,\mathcal{W}^\textrm{G}_b(\hat{\mathbf{s}},f)\,V_{ab}\,e^{2\pi i f\mathbf{r}_{ab}\!\cdot\,\hat{\mathbf{s}}/c}.
+  \mathcal{I}^\prime_{ab}(\hat{\mathbf{s}},f) &= \mathcal{W}^{\textrm{G}\star}_a(\hat{\mathbf{s}},f)\,\mathcal{W}^\textrm{G}_b(\hat{\mathbf{s}},f)\,V_{ab}\,e^{i 2\pi f\mathbf{r}_{ab}\!\cdot\,\hat{\mathbf{s}}/c}.
 \end{align}
@@
 This demonstrates that images from EPIC can robustly image data from heterogeneous arrays while being mathematically optimal.
-Besides being sub-optimal, erroneous assumptions about the homogeneity of an array can lead to significant and systematic mis-estimates ({\bf an underestimate in the above example}) of the sky model.
-
-% {\bf In the above example, we chose a scenario wherein the antennas in the array have large differences between them to clearly demonstrate the functionality in EPIC to image data from heterogeneous arrays.
-The analysis, however, is applicable to any heterogeneous array.
-In cases where the differences are smaller, the fidelity with which the sky model is recovered will still be degraded, although to a correspondingly smaller extent.
-To robustly quantify this degradation as a function of deviation from homogeneity will require image deconvolution and will be investigated in the future as it is currently beyond the scope of this paper.}
+Besides being sub-optimal, erroneous assumptions about the homogeneity of an array can lead to significant and systematic mis-estimates (an underestimate in the above example) of the sky model.
 
@@
 
-We now investigate the feasibility of implementing the EPIC imager on current
-and future radio telescopes.
+We now investigate the feasibility of implementing the EPIC imager on current and future radio telescopes.
 
@@
 
-We have profiled the core routines of EPIC line-by-line for various ranges of 
-parameters such as antenna filling fraction, maximum baseline length, bandwidth 
-and frequency resolution, integration time-scale, etc. for HERA antenna layouts 
-which are highly compact. However, we note that in general, the hardware and 
-optimization of routines in place will determine the relative speeds of the 
-different stages in the pipeline.
-
-Of all steps in the MOFF pipeline that are repeated for every writeout from 
-the F-engine, the most expensive step even for dense HERA layouts is found to be 
-the spatial two-dimensional FFT in the imaging stage relative to applying the 
-sparse matrix gridding convolution, squaring or time-averaging. For instance, 
-even in the conservative dense array layout scenario that makes these other 
-stages even more expensive, the gridding convolution, squaring and time-averaging 
-take up only $\lesssim 20$\%, $\lesssim 20$\% and $\lesssim 5$\% respectively 
-of the total processing time while the spatial Fourier transform takes up 
-$\gtrsim 55$ per cent of the total time. With sparser arrays the gridding process 
-will be be even faster.
-
-In visibility-based imaging, the predominant computational cost is at the 
-X-engine requiring $\Nant(\Nant-1)/2$ complex multiplications per channel at 
-every A/D writeout time-scale.
+We have profiled the core routines of EPIC line-by-line for various ranges of parameters such as antenna filling fraction, maximum baseline length, bandwidth and frequency resolution, integration time-scale, etc.
+for HERA antenna layouts which are highly compact.
+However, we note that in general, the hardware and optimization of routines in place will determine the relative speeds of the different stages in the pipeline.
+
+Of all the steps in the MOFF pipeline that are repeated for every writeout from the F-engine, the most expensive step even for dense HERA layouts is found to be the spatial two-dimensional FFT in the imaging stage relative to applying the sparse matrix gridding convolution, squaring or time-averaging.
+For instance, even in the conservative dense array layout scenario that makes these other stages even more expensive, the gridding convolution, squaring and time-averaging take up only $\lesssim 20$\%, $\lesssim 20$\% and $\lesssim 5$\% respectively of the total processing time while the spatial Fourier transform takes up $\gtrsim 55$ per cent of the total time.
+With sparser arrays the gridding process will be be even faster.
+
+In visibility-based imaging, the predominant computational cost is at the X-engine requiring $\Nant(\Nant-1)/2$ complex multiplications per channel at every A/D writeout time-scale.
 
@@
 
-The number of complex multiplications and additions in the spatial Fourier 
-transform implemented via FFT is $\approx \beta\,(4\Ngrid)\log_2(4\Ngrid)$ 
-where $\Ngrid$ is the number of pixels on the grid, the factor 4 accounts for 
-increase in number of pixels as a result of zero-padding before spatial Fourier 
-transform, and $\beta$ is a constant that depends on the implementation of 
-twiddle FFT algorithms \citep{bri74}. In our study, we set $\beta=5$, a 
-value\footnote{http://www.fftw.org/speed/method.html} much more conservative 
-than was indicated in \citet{mor11}. We set the number of complex 
-multiplications in the X-engine in visibility-based imaging to 
-$\Nant(\Nant-1)/2$.
-
-We consider a variety of current and planned radio telescopes. Their antenna 
-layouts are summarized in Table~\ref{tab:antenna-layouts}. The size of the 
-layout gives the maximum baseline $b_\textrm{max}$. The grid spacing is 
-determined by the science goals of the experiment in general. For our 
-purpose, we assume a typical requirement that only the field of view of the 
-antenna is to be imaged. This sets the grid spacing to be equal to the size 
-of the antenna, $A_\textrm{a}$. Hence, 
-$\Ngrid\simeq b_\textrm{max}^2/A_\textrm{a}$.
+The number of complex multiplications and additions in the spatial Fourier transform implemented via FFT is $\approx \beta\,(4\Ngrid)\log_2(4\Ngrid)$ where $\Ngrid$ is the number of pixels on the grid, the factor 4 accounts for increase in number of pixels as a result of zero-padding before spatial Fourier transform, and $\beta$ is a constant that depends on the implementation of twiddle FFT algorithms \citep{bri74}.
+In our study, we set $\beta=5$, a value\footnote{http://www.fftw.org/speed/method.html} much more conservative than was indicated in \citet{mor11}.
+We set the number of complex multiplications in the X-engine in visibility-based imaging to $\Nant(\Nant-1)/2$.
+
+We consider a variety of current and planned radio telescopes.
+Their antenna layouts are summarized in Table~\ref{tab:antenna-layouts}.
+The size of the layout gives the maximum baseline $b_\textrm{max}$.
+The grid spacing is determined by the science goals of the experiment in general.
+For our purpose, we assume a typical requirement that only the field of view of the antenna is to be imaged.
+This sets the grid spacing to be equal to the size of the antenna, $A_\textrm{a}$.
+Hence, $\Ngrid\simeq b_\textrm{max}^2/A_\textrm{a}$.
 
@@
 
-Fig.~\ref{fig:parameter-space-computations-instruments} shows the number of 
-complex operations per frequency channel per cross-polarization per integration 
-time-scale. Telescopes that fall to the left of the solid line indicate MOFF 
-imaging is computationally more efficient than visibility-based imaging. All HERA 
-layouts except HERA-19 and HERA-37 are in a region of parameter space 
-where MOFF imaging holds the advantage. The dashed line showing future trajectory 
-of HERA like systems will be clearly favoured by MOFF imaging. The dotted line is 
-similarly a hypothetical trajectory for the MWA with more tiles added inside the 
-same core diameter. The grey shaded area is for a projected LWA expansion and is 
-also predominantly in the region favouring MOFF imaging. It is bounded by the LWA1 
-and LWA-OV on the left and right respectively. The current (see 
-Table~\ref{tab:antenna-layouts}) and a hypothetical expanded layout with a 
-four-fold increase in number of elements over a 50 per cent increase in 
-$b_\textrm{max}$ provide the bounds at the bottom and top respectively. Current 
-instruments such as MWA and LOFAR lie in parameter space favouring 
-visibility-based imaging.
+Fig.~\ref{fig:parameter-space-computations-instruments} shows the number of complex operations per frequency channel per cross-polarization per integration time-scale.
+Telescopes that fall to the left of the solid line indicate MOFF imaging is computationally more efficient than visibility-based imaging.
+All HERA layouts except HERA-19 and HERA-37 are in a region of parameter space where MOFF imaging holds the advantage.
+The dashed line showing future trajectory of HERA like systems will be clearly favoured by MOFF imaging.
+The dotted line is similarly a hypothetical trajectory for the MWA with more tiles added inside the same core diameter.
+The grey shaded area is for a projected LWA expansion and is also predominantly in the region favouring MOFF imaging.
+It is bounded by the LWA1 and LWA-OV on the left and right respectively.
+The current (see Table~\ref{tab:antenna-layouts}) and a hypothetical expanded layout with a four-fold increase in number of elements over a 50 per cent increase in $b_\textrm{max}$ provide the bounds at the bottom and top respectively.
+Current instruments such as MWA and LOFAR lie in parameter space favouring visibility-based imaging.
 
@@
   % \includegraphics[width=\columnwidth]{MOFF_FX_computations_fov_gridding_annotated.eps}
-  \caption{Current and planned instruments in parameter space of
-    number of complex {\bf multiplications and additions}. The solid line
-    is the boundary at which the number of operations with MOFF and 
-    visibility-based imaging are equal. MOFF imaging is more efficient for 
-    telescopes occupying the left of this line and vice versa. CHIME, HIRAX, 
-    SKA1-LC, SKA1-LCD and all the HERA layouts except HERA-19 and HERA-37 lie 
-    in the parameter space favoured by MOFF imaging. The dashed line shows the 
-    projected trajectory of hypothetical expanded HERA layouts. The dotted line 
-    similarly shows hypothetical expanded MWA layouts with more tiles added in 
-    the same core. The grey shaded area denotes the projected trajectory of the 
-    LWA bounded by LWA1 (left edge), LWA-OV (right edge), current layout 
-    (bottom) and a four-fold increase in the number of elements within a 50 per 
-    cent increase in the core size (top). Current instruments such as MWA and 
-    LOFAR fall in a region favoured by visibility-based imaging.}
+  \caption{Current and planned instruments in parameter space of number of complex multiplications and additions.
+The solid line is the boundary at which the number of operations with MOFF and visibility-based imaging are equal.
+MOFF imaging is more efficient for telescopes occupying the left of this line and vice versa.
+CHIME, HIRAX, SKA1-LC, SKA1-LCD and all the HERA layouts except HERA-19 and HERA-37 lie in the parameter space favoured by MOFF imaging.
+The dashed line shows the projected trajectory of hypothetical expanded HERA layouts.
+The dotted line similarly shows hypothetical expanded MWA layouts with more tiles added in the same core.
+The grey shaded area denotes the projected trajectory of the LWA bounded by LWA1 (left edge), LWA-OV (right edge), current layout (bottom) and a four-fold increase in the number of elements within a 50 per cent increase in the core size (top).
+Current instruments such as MWA and LOFAR fall in a region favoured by visibility-based imaging.}
   \label{fig:parameter-space-computations-instruments}
@@
 
-We elaborate on the I/O data rates required with the MOFF and visibility-based
-algorithms. This is particularly relevant in the context of radio transient 
-detection.
-
-Implementation of the MOFF algorithm with EPIC yields calibrated images on
-time-scales of the output generated by the digitizer and is set by the inverse
-of the frequency channel width. These calibrated images are accumulated and 
-averaged to a certain time-scale depending on science or hardware requirements,
-or when the sky has rotated significantly, whichever is lesser. In 
-visibility-based algorithms, the visibilities are accumulated and averaged to 
-this time-scale before images are produced. Thus the data throughput (in samples 
-per second) per cross-polarization with MOFF and X-engine outputs are: 
+We elaborate on the I/O data rates required with the MOFF and visibility-based algorithms. This is particularly relevant in the context of radio transient detection.
+
+Implementation of the MOFF algorithm with EPIC yields calibrated images on time-scales of the output generated by the digitizer and is set by the inverse of the frequency channel width.
+These calibrated images are accumulated and averaged to a certain time-scale depending on science or hardware requirements, or when the sky has rotated significantly, whichever is lesser.
+In visibility-based algorithms, the visibilities are accumulated and averaged to this time-scale before images are produced.
+Thus the data throughput (in samples per second) per cross-polarization with MOFF and X-engine outputs are:
 \begin{align}
@@
 \end{align}
-where, the factor 4 in the expression for $r_\textrm{MOFF}$ accounts for imaging
-after zero-padding the gridded electric fields, the leading factor of 2 in the
-expression for $r_\textrm{X}$ accounts for the real and imaginary parts of the
-complex visibilities, $\Delta B$ is the bandwidth, $\Delta f$ is the frequency 
-resolution, and $\Delta t$ is the time-scale over which the transient phenomenon
-is sampled and the data (images or visibilities) are averaged to.
-
-Though a full understanding of the FRB phenomena is yet to emerge, there are 
-indications the time-scales of FRB objects are $\Delta t \sim 1$--10~ms 
-\citep{tho13}. For a telescope like HERA, $\Delta B \simeq 100$~MHz, 
-$\Delta f \simeq 100$~kHz. For HERA-331, $\Nant=331$ and with a grid 
-spacing to image the field of view, $\Ngrid \simeq 441$ or 
-$\Ngrid \simeq 1024$ if $\Ngrid$ is preferred as a power of 2 in either 
-direction. Using 8 bytes for each floating point sample in the MOFF images and 
-4 bytes each for real and imaginary parts of visibility samples, the throughputs 
-are $r_\textrm{MOFF} \lesssim 3$~GB~s$^{-1}$ and 
-$r_\textrm{X} \simeq 41$~GB~s$^{-1}$. For HERA-37, 
-$r_\textrm{MOFF} \lesssim 190$~MB~s$^{-1}$ and 
-$r_\textrm{X} \simeq 0.5$~GB~s$^{-1}$. In such a case, The X-engine throughput 
-corresponds to an extreme rate of $\simeq 1.8$~TB an hour per 
-cross-polarization. Conversely, for the same data throughput, the MOFF algorithm 
-can sample even shorter time-scales.
-
-Table~\ref{tab:data-rates} shows these data rates for some of the current and 
-planned telescopes for $\Delta t=10$~ms. In almost all cases listed, even with 
-conservative estimates, the MOFF algorithm provides very economic data 
-throughput for a majority of next generation radio telescopes with a dense 
-layout. The most significant advantage is that calibrated images are also 
-available at no extra cost.
+where, the factor 4 in the expression for $r_\textrm{MOFF}$ accounts for imaging after zero-padding the gridded electric fields, the leading factor of 2 in the expression for $r_\textrm{X}$ accounts for the real and imaginary parts of the complex visibilities, $\Delta B$ is the bandwidth, $\Delta f$ is the frequency resolution, and $\Delta t$ is the time-scale over which the transient phenomenon is sampled and the data (images or visibilities) are averaged to.
+
+Though a full understanding of the FRB phenomena is yet to emerge, there are indications the time-scales of FRB objects are $\Delta t \sim 1$--10~ms \citep{tho13}.
+For a telescope like HERA, $\Delta B \simeq 100$~MHz, $\Delta f \simeq 100$~kHz.
+For HERA-331, $\Nant=331$ and with a grid spacing to image the field of view, $\Ngrid \simeq 441$ or $\Ngrid \simeq 1024$ if $\Ngrid$ is preferred as a power of 2 in either direction.
+Using 8 bytes for each floating point sample in the MOFF images and 4 bytes each for real and imaginary parts of visibility samples, the throughputs are $r_\textrm{MOFF} \lesssim 3$~GB~s$^{-1}$ and $r_\textrm{X} \simeq 41$~GB~s$^{-1}$.
+For HERA-37, $r_\textrm{MOFF} \lesssim 190$~MB~s$^{-1}$ and $r_\textrm{X} \simeq 0.5$~GB~s$^{-1}$.
+In such a case, The X-engine throughput corresponds to an extreme rate of $\simeq 1.8$~TB an hour per cross-polarization.
+Conversely, for the same data throughput, the MOFF algorithm can sample even shorter time-scales.
+
+Table~\ref{tab:data-rates} shows these data rates for some of the current and planned telescopes for $\Delta t=10$~ms.
+In almost all cases listed, even with conservative estimates, the MOFF algorithm provides very economic data throughput for a majority of next generation radio telescopes with a dense layout.
+The most significant advantage is that calibrated images are also available at no extra cost.
 
@@
   \centering
-  \caption{Data throughput per cross-polarization for various telescopes with 
-    MOFF and X-engine outputs on time-scales of $\Delta t=10$~ms assuming 
-    $\Delta B=100$~MHz and $\Delta f=100$~kHz.}
+  \caption{Data throughput per cross-polarization for various telescopes with MOFF and X-engine outputs on time-scales of $\Delta t=10$~ms assuming $\Delta B=100$~MHz and $\Delta f=100$~kHz.}
   \label{tab:data-rates}
@@
     \item[a] Antenna layouts are listed in Table~\ref{tab:antenna-layouts}.
-    \item[b] $\Ngrid$ is usually greater than true value because of 
-      rounding to the next power of 2 in either direction. Thus 
-      $r_\textrm{MOFF}$ is usually lesser than the conservative values 
-      listed here.
-    \item[c] We assume 8 bytes for each real sample from MOFF images, and
-      4 bytes each for real and imaginary parts of visibility samples.
+    \item[b] $\Ngrid$ is usually greater than true value because of rounding to the next power of 2 in either direction.
+Thus $r_\textrm{MOFF}$ is usually lesser than the conservative values listed here.
+    \item[c] We assume 8 bytes for each real sample from MOFF images, and 4 bytes each for real and imaginary parts of visibility samples.
   \end{tablenotes}
@@
 Under the class of direct imaging techniques, ours is one of the most generic -- neither does it place any constraint on the array layout to be on a regular grid nor does it require the antenna array to be homogeneous.
-We have demonstrated its natural ability to robustly image data from heterogeneous arrays and shown that wrong assumptions about array homogeneity could result in significant and systematic mis-estimation of the sky model.
-
-Our EPIC package, now publicly available, written in object oriented Python, is 
-highly modularized and parallelizable. It includes an implementation of the 
-MOFF algorithm in addition to visibility-based software holography imaging and 
-a data simulator for sky models. It is designed to provide a development 
-platform to compare different imaging approaches and serve as a stepping stone 
-for real-life GPU/FPGA-based implementation on telescopes. It has been 
-successfully tested on simulated MWA observations as well as real LWA 
-observations from both imaging and calibration viewpoints.
-
-The MOFF algorithm packaged with EPIC is already found to be most suitable 
-for many present and planned radio telescopes such as the LWA, HERA, CHIME,
-HIRAX and SKA. In general, MOFF is most suited to operate in the region of 
-parameter space characterized by dense packing of a large number of antennas 
-especially when consisting of a large number of small antenna elements.
-
-It is seen to have significant savings in data throughput relative to a 
-X-engine based pipeline. A unique advantage is the instantaneous availability 
-of calibrated time-domain images at no extra cost. Hence, it is a compelling 
-candidate for time-domain radio astronomy, e.g. search for and monitoring of 
-transients. Potentially, it could allow on-chip processing thus lowering even 
-further the already relatively low I/O bandwidth shown in 
-Table~\ref{tab:data-rates} except when a transient event is detected. Transient 
-detection pipelines at the backend of EPIC can be fine-tuned to target fast 
-transients such as the Fast Radio Bursts \citep[FRB;][]{tho13} on millisecond 
-time-scales at GHz frequencies or slow transients from planetary and exoplanetary 
-origins at frequencies around 100~MHz.
-
-Thus, EPIC with the MOFF algorithm packaged is uniquely poised to offer a
-substantial advantage to imaging with large-N dense arrays typical of 
-next-generation radio telescopes as well as push the frontiers of 
-time-domain astronomy to fill gaps in understanding the science behind 
-phenomena responsible for extreme transient events in the Universe.
+{\bf We have provided a detailed mathematical framework for imaging with heterogeneous arrays, demonstrated the natural ability of EPIC to robustly image data from such arrays, and shown that wrong assumptions about array homogeneity could result in significant and systematic mis-estimation of the sky model}.
+
+Our EPIC package, now publicly available, written in object-oriented Python, is highly modularized and parallelizable.
+It includes an implementation of the MOFF algorithm in addition to visibility-based software holography imaging and a data simulator for sky models.
+It is designed to provide a development platform to compare different imaging approaches and serve as a stepping stone for real-life GPU/FPGA-based implementation on telescopes.
+It has been successfully tested on simulated MWA observations as well as real LWA observations from both imaging and calibration viewpoints.
+
+The MOFF algorithm packaged with EPIC is already found to be most suitable for many present and planned radio telescopes such as the LWA, HERA, CHIME, HIRAX and SKA.
+In general, MOFF is most suited to operate in the region of parameter space characterized by dense packing of a large number of antennas especially when consisting of a large number of small antenna elements.
+
+It is seen to have significant savings in data throughput relative to a X-engine based pipeline.
+A unique advantage is the instantaneous availability of calibrated time-domain images at no extra cost.
+Hence, it is a compelling candidate for time-domain radio astronomy, e.g.
+search for and monitoring of transients.
+Potentially, it could allow on-chip processing thus lowering even further the already relatively low I/O bandwidth shown in Table~\ref{tab:data-rates} except when a transient event is detected.
+Transient detection pipelines at the backend of EPIC can be fine-tuned to target fast transients such as the Fast Radio Bursts \citep[FRB;][]{tho13} on millisecond time-scales at GHz frequencies or slow transients from planetary and exoplanetary origins at frequencies around 100~MHz.
+
+Thus, EPIC with the MOFF algorithm packaged is uniquely poised to offer a substantial advantage to imaging with large-N dense arrays typical of next-generation radio telescopes as well as push the frontiers of time-domain astronomy to fill gaps in understanding the science behind phenomena responsible for extreme transient events in the Universe.
 
@@
 
-We thank Larry D'Addario, Gregg Hallinan, Joseph Lazio and Harish Vedantham for 
-their valuable inputs, and Greg Taylor for providing us with LWA data. This work 
-has been supported by the National Science Foundation through award AST-1206552. 
-Construction of the LWA has been supported by the Office of Naval Research under 
-Contract N00014-07-C-0147. Support for operations and continuing development of 
-the LWA1 is provided by the National Science Foundation under grant AST-1139974 
-of the University Radio Observatory program.
+We thank Larry D'Addario, Gregg Hallinan, Joseph Lazio and Harish Vedantham for their valuable inputs, and Greg Taylor for providing us with LWA data.
+This work has been supported by the National Science Foundation through award AST-1206552.
+Construction of the LWA has been supported by the Office of Naval Research under Contract N00014-07-C-0147.
+Support for operations and continuing development of the LWA1 is provided by the National Science Foundation under grant AST-1139974 of the University Radio Observatory program.
 
@@
 
-EPIC is built using object oriented programming in Python and is built on
-carefully crafted modules which closely represent real-life entities in radio 
-interferometer arrays and observations. The essential modules along with their
-key attributes and methods are illustrated in Fig.~\ref{fig:software-modules}.
+EPIC is built using object-oriented programming in Python and is built on carefully crafted modules which closely represent real-life entities in radio interferometer arrays and observations.
+The essential modules along with their key attributes and methods are illustrated in Fig.~\ref{fig:software-modules}.
 These modules are described below.
@@
   % \includegraphics[width=\linewidth]{EPIC-modules.eps}
-  \caption{Software architecture of EPIC with core modules, their essential
-    attributes and functions. The antenna module forms the fundamental building 
-    block. It consists of electric field time-series and spectra and the 
-    F-engine that performs a temporal FFT to obtain electric field spectra from 
-    the time-series. The interferometer module is made of a pair of antenna 
-    modules. Its main function is the X-engine (FX or XF) to produce visibility 
-    spectra. The antenna array module is made of all individual antenna modules
-    as its components and contains collective properties about the antenna 
-    subsystems. Its core function is the creation of antenna-to-grid mapping, 
-    gridded aperture weights and electric fields. The interferometer array 
-    module is very similar in principle to the antenna array module except it 
-    operates on cross-correlations and produces gridded visibilities. The image 
-    module takes gridded electric fields or visibilities and performs a 
-    two-dimensional spatial FFT (and squares the intermediate image in case 
-    of the former) to produce output images. Broadly, the MOFF algorithm is 
-    implemented by modules below the horizontal dashed line while the 
-    visibility-based imaging uses modules above the line. The exact processing 
-    pathway implementing the MOFF algorithm is shown in bolded modules.}
+  \caption{Software architecture of EPIC with core modules, their essential attributes and functions.
+The antenna module forms the fundamental building block.
+It consists of electric field time-series and spectra and the F-engine that performs a temporal FFT to obtain electric field spectra from the time-series.
+The interferometer module is made of a pair of antenna modules.
+Its main function is the X-engine (FX or XF) to produce visibility spectra.
+The antenna array module is made of all individual antenna modules as its components and contains collective properties about the antenna subsystems.
+Its core function is the creation of antenna-to-grid mapping, gridded aperture weights and electric fields.
+The interferometer array module is very similar in principle to the antenna array module except it operates on cross-correlations and produces gridded visibilities.
+The image module takes gridded electric fields or visibilities and performs a two-dimensional spatial FFT (and squares the intermediate image in case of the former) to produce output images.
+Broadly, the MOFF algorithm is implemented by modules below the horizontal dashed line while the visibility-based imaging uses modules above the line.
+The exact processing pathway implementing the MOFF algorithm is shown in bolded modules.}
   \label{fig:software-modules}
@@
 
-The other function (not shown in the figure) is to update the data as new data 
-streams in. This can also be parallelized. Another important attribute consists 
-of antenna flags (not shown in the figure) for each polarization appropriate 
-for the data stream being held by the module.
+The other function (not shown in the figure) is to update the data as new data streams in.
+This can also be parallelized.
+Another important attribute consists of antenna flags (not shown in the figure) for each polarization appropriate for the data stream being held by the module.
 
@@
 
-The other attributes (not shown in the figure) are the flags applicable for each
-cross-polarization for the current data stream. Similar to the antenna module,
-it has an update function that can update the visibilities $V_t(\tau)$ 
-or $V(f)$ directly rather than through the electric fields of its 
-component antennas. This functionality is to allow EPIC to operate while 
-attached to the backend of traditional correlator systems. This feature is not 
-utilized for purposes of this paper.
-
-This module forms the fundamental unit for the interferometer array module 
-(to be discussed below) and in general for visibility-based correlator and 
-imaging systems.
+The other attributes (not shown in the figure) are the flags applicable for each cross-polarization for the current data stream.
+Similar to the antenna module, it has an update function that can update the visibilities $V_t(\tau)$ or $V(f)$ directly rather than through the electric fields of its component antennas.
+This functionality is to allow EPIC to operate while attached to the backend of traditional correlator systems.
+This feature is not utilized for purposes of this paper.
+
+This module forms the fundamental unit for the interferometer array module (to be discussed below) and in general for visibility-based correlator and imaging systems.
 
@@
 
-The antenna array module consists of all the antenna modules as its attributes
-and represents the collective properties of its component antennas. By virtue of 
-holding each antenna data independently in their respective modules, the F-engine 
-for the entire array can be distributed to the F-engines of the component 
-antenna modules thus achieving a highly parallelized F-engine while emulating 
-real telescope systems.
-
-The primary attributes held by this module are the antenna aperture 
-illumination weights and electric fields projected on the grid using the 
-gridding convolution method described above and implemented by the gridding 
-function in this module. Significant parts of the antenna-to-grid mapping and 
-gridding convolution are parallelizable across antennas and frequencies.
-
-Individual antenna flags are carried over as additional weights to be applied
-to the gridded aperture illumination and electric fields. A series of data 
-streams can be stacked up to take advantage of the array optimization available 
-in Python. This module is also equipped to manage dual-polarization.
+The antenna array module consists of all the antenna modules as its attributes and represents the collective properties of its component antennas.
+By virtue of holding each antenna data independently in their respective modules, the F-engine for the entire array can be distributed to the F-engines of the component antenna modules thus achieving a highly parallelized F-engine while emulating real telescope systems.
+
+The primary attributes held by this module are the antenna aperture illumination weights and electric fields projected on the grid using the gridding convolution method described above and implemented by the gridding function in this module.
+Significant parts of the antenna-to-grid mapping and gridding convolution are parallelizable across antennas and frequencies.
+
+Individual antenna flags are carried over as additional weights to be applied to the gridded aperture illumination and electric fields.
+A series of data streams can be stacked up to take advantage of the array optimization available in Python.
+This module is also equipped to manage dual-polarization.
 
@@
 
-Similar to the antenna array module, the interferometer array module consists of
-individual interferometer modules. It can parallelize the correlator operations
-by distributing the X-operation over the X-engines of its component 
-interferometer modules. The interferometer-to-grid mapping and gridding 
-convolution are very similar in nature to that of the antenna array module. 
-Flag-based grid weights, stacking and ability to handle all four 
-cross-polarizations are built into this module.
+Similar to the antenna array module, the interferometer array module consists of individual interferometer modules.
+It can parallelize the correlator operations by distributing the X-operation over the X-engines of its component interferometer modules.
+The interferometer-to-grid mapping and gridding convolution are very similar in nature to that of the antenna array module.
+Flag-based grid weights, stacking and ability to handle all four cross-polarizations are built into this module.
 
@@
 
-The image module is built as a general purpose module that can switch between 
-operating on gridded electric fields or visibilities. At its heart, it consists
-of a two-dimensional spatial FFT where the padding can be specified by the user 
-to control the resolution in the output images. In case of MOFF imaging, there 
-is an additional step of squaring the holographic electric field images.
-
-Besides its core functions of spatial Fourier transform and squaring, it can 
-stack, accumulate and average images, and optionally remove the antenna
-auto-correlations centred around the zero-spacing pixel in the $uv$ plane. 
-It also handles all four cross-polarization products. Currently, it supports 
-writing data out in standard FITS format.
+The image module is built as a general purpose module that can switch between operating on gridded electric fields or visibilities.
+At its heart, it consists of a two-dimensional spatial FFT where the padding can be specified by the user to control the resolution in the output images.
+In case of MOFF imaging, there is an additional step of squaring the holographic electric field images.
+
+Besides its core functions of spatial Fourier transform and squaring, it can stack, accumulate and average images, and optionally remove the antenna auto-correlations centred around the zero-spacing pixel in the $uv$ plane.
+It also handles all four cross-polarization products.
+Currently, it supports writing data out in standard FITS format.
 
@@
 \begin{align}\label{eqn:apndx-measurement-eqn-2}
-  V_{ab}(f) &= \int \mathcal{W}^\textrm{I}_a(\hat{\mathbf{s}},f)\mathcal{W}^{\textrm{I}\star}_b(\hat{\mathbf{s}},f)\,\mathcal{I}(\hat{\mathbf{s}},f)\,e^{-2\pi i f\mathbf{r}_{ab}\!\cdot\,\hat{\mathbf{s}}/c}\,\dif\Omega,
+  V_{ab}(f) &= \int \mathcal{W}^\textrm{I}_a(\hat{\mathbf{s}},f)\mathcal{W}^{\textrm{I}\star}_b(\hat{\mathbf{s}},f)\,\mathcal{I}(\hat{\mathbf{s}},f)\,e^{-i 2\pi f\mathbf{r}_{ab}\!\cdot\,\hat{\mathbf{s}}/c}\,\dif\Omega,
 \end{align}
@@
 \begin{align}
-  \mathcal{I}_\textrm{D}(\hat{\mathbf{s}},f) &= \frac{1}{\Nant(\Nant-1)}\,\sum_{ab}\, V_{ab}(f)\,e^{2\pi i f\mathbf{r}_{ab}\!\cdot\,\hat{\mathbf{s}}/c},
+  \mathcal{I}_\textrm{D}(\hat{\mathbf{s}},f) &= \frac{1}{\Nant(\Nant-1)}\,\sum_{ab}\, V_{ab}(f)\,e^{i 2\pi f\mathbf{r}_{ab}\!\cdot\,\hat{\mathbf{s}}/c},
 \end{align}
@@
 \begin{align}
-  \mathcal{I}^\prime_{ab}(\hat{\mathbf{s}},f) &= \int V^\prime_{ab}(\mathbf{r},f)\,e^{2\pi i f\mathbf{r}\cdot\,\hat{\mathbf{s}}/c}\,\dif^2\mathbf{r} \nonumber\\
-                                              &= \int B^{\textrm{G}\star}_{ab}(\mathbf{r}-\mathbf{r}_{ab},f)\,V_{ab}\,e^{2\pi i f\mathbf{r}\cdot\,\hat{\mathbf{s}}/c}\,\dif^2\mathbf{r} \nonumber\\
+  \mathcal{I}^\prime_{ab}(\hat{\mathbf{s}},f) &= \int V^\prime_{ab}(\mathbf{r},f)\,e^{i 2\pi f\mathbf{r}\cdot\,\hat{\mathbf{s}}/c}\,\dif^2\mathbf{r} \nonumber\\
+                                              &= \int B^{\textrm{G}\star}_{ab}(\mathbf{r}-\mathbf{r}_{ab},f)\,V_{ab}\,e^{i 2\pi f\mathbf{r}\cdot\,\hat{\mathbf{s}}/c}\,\dif^2\mathbf{r} \nonumber\\
                                               &= V_{ab}\,\int \left[\int W^\textrm{G}_b(\mathbf{r}^\prime+\mathbf{r}-\mathbf{r}_a,f)\right. \nonumber\\
-                                              &\qquad\qquad\,\,\, \left.\times\, W^{\textrm{G}\star}_a(\mathbf{r}^\prime-\mathbf{r}_a,f)\,\dif^2\mathbf{r}^\prime\right]\,e^{2\pi i f\mathbf{r}\cdot\,\hat{\mathbf{s}}/c}\,\dif^2\mathbf{r} \nonumber\\
-                                              % &\qquad\qquad \times e^{2\pi i f\mathbf{r}\cdot\,\hat{\mathbf{s}}/c}\,\dif^2\mathbf{r} \nonumber\\
-                                              &= \mathcal{W}^{\textrm{G}\star}_a(\hat{\mathbf{s}},f)\mathcal{W}^\textrm{G}_b(\hat{\mathbf{s}},f)\,V_{ab}\,e^{2\pi i f\mathbf{r}_{ab}\!\cdot\,\hat{\mathbf{s}}/c},
+                                              &\qquad\qquad\,\,\, \left.\times\, W^{\textrm{G}\star}_a(\mathbf{r}^\prime-\mathbf{r}_a,f)\,\dif^2\mathbf{r}^\prime\right]\,e^{i 2\pi f\mathbf{r}\cdot\,\hat{\mathbf{s}}/c}\,\dif^2\mathbf{r} \nonumber\\
+                                              % &\qquad\qquad \times e^{i 2\pi f\mathbf{r}\cdot\,\hat{\mathbf{s}}/c}\,\dif^2\mathbf{r} \nonumber\\
+                                              &= \mathcal{W}^{\textrm{G}\star}_a(\hat{\mathbf{s}},f)\mathcal{W}^\textrm{G}_b(\hat{\mathbf{s}},f)\,V_{ab}\,e^{i 2\pi f\mathbf{r}_{ab}\!\cdot\,\hat{\mathbf{s}}/c},
 \end{align}
@@
 \nonumber\\
-  &\qquad\qquad\qquad\qquad\qquad\qquad \times \left. V_{ab}\,e^{2\pi i f\mathbf{r}_{ab}\!\cdot\,\hat{\mathbf{s}}/c}\right].
+  &\qquad\qquad\qquad\qquad\qquad\qquad \times \left. V_{ab}\,e^{i 2\pi f\mathbf{r}_{ab}\!\cdot\,\hat{\mathbf{s}}/c}\right].
 \end{align}
@@
 \begin{align}\label{eqn:apndx-wt-dirty-image-homogeneous}
-  \mathcal{I}^\prime(\hat{\mathbf{s}},f) &= \frac{\left|\mathcal{W}^\textrm{G}(\hat{\mathbf{s}},f)\right|^2}{\Nant(\Nant-1)}\,\sum_{\substack{ab\\a\ne b}}\,V_{ab}\,e^{2\pi i f\mathbf{r}_{ab}\!\cdot\,\hat{\mathbf{s}}/c} \nonumber\\
+  \mathcal{I}^\prime(\hat{\mathbf{s}},f) &= \frac{\left|\mathcal{W}^\textrm{G}(\hat{\mathbf{s}},f)\right|^2}{\Nant(\Nant-1)}\,\sum_{\substack{ab\\a\ne b}}\,V_{ab}\,e^{i 2\pi f\mathbf{r}_{ab}\!\cdot\,\hat{\mathbf{s}}/c} \nonumber\\
   &= \left|\mathcal{W}^\textrm{G}(\hat{\mathbf{s}},f)\right|^2\,\mathcal{I}_\textrm{D}(\hat{\mathbf{s}},f) \nonumber\\
@@
 
-This means $\mathcal{W}_\textrm{eff}$ can be expressed as fourth power combinations of antenna voltage patterns each weighted by the number of antenna pairs in those antenna pair types:
-\begin{align}
-  \mathcal{W}_\textrm{eff}^\textrm{opt} &\approx \frac{1}{\Nant(\Nant-1)}\,\sum_{ij}n_{ij}\,\mathcal{W}^{\textrm{G}\star}_{(i)}\,\mathcal{W}^\textrm{G}_{(j)}\,\mathcal{W}^\textrm{I}_{(i)}\,\mathcal{W}^{\textrm{I}\star}_{(j)}.
+This means $\mathcal{W}_\textrm{eff}$ can be expressed as fourth-power combinations of antenna voltage patterns each weighted by the number of antenna pairs in those antenna pair types:
+\begin{align}
+  \mathcal{W}_\textrm{eff} &\approx \frac{1}{\Nant(\Nant-1)}\,\sum_{ij}n_{ij}\,\mathcal{W}^{\textrm{G}\star}_{(i)}\,\mathcal{W}^\textrm{G}_{(j)}\,\mathcal{W}^\textrm{I}_{(i)}\,\mathcal{W}^{\textrm{I}\star}_{(j)}.
 \end{align}

### 2017-01-06 imaging_demo/ms.tex
Clarify notation and possible confusion
From 0e5007d (mnras):
====================

In ms.tex,

(*) Change indexing of antenna types from i and j to p and q
respectively
(*) Clarify indexing notation involving 'ab' and 'pq'
(*) Clarify quantity denoted by n is different in section 2 compared
to that in sections 5 and appendix B

@@
 \begin{align}\label{eqn:wt-dirty-image-EPIC-decomp}
-  \mathcal{I}^\prime &= \frac{1}{\Nant(\Nant-1)}\,\sum_{ij}\sum_{\substack{ab\\a\ne b\\a\in i, b\in j}}\,\mathcal{I}^\prime_{ab},
-\end{align}
-where, $ab$ indexes pairs of antennas $a$ and $b$, $ij$ indexes types of pairs of antennas, with $i$ and $j$ indexing antenna types.
+  \mathcal{I}^\prime &= \frac{1}{\Nant(\Nant-1)}\,\sum_{pq}\sum_{\substack{ab\\a\ne b\\a\in p, b\in q}}\,\mathcal{I}^\prime_{ab},
+\end{align}
+where, $ab$ indexes pairs of antennas $a$ and $b$, {\bf $pq$ indexes the different types of antenna pairs, with $p$ and $q$ indexing antenna types (note that $ab$ and $pq$ are simply indices in our notation and are not to be interpreted as products of the individual indices)}.
 Thus, equation~\ref{eqn:wt-dirty-image-homogeneous} is re-written as the sum of fringes over all unique antenna pairs in an antenna pair type (inner sum) and subsequently summed over all antenna pair types (outer sum).
@@
 \begin{align}\label{eqn:wt-dirty-image-EPIC-decomp-approx1}
-  \mathcal{I}^\prime &\approx \frac{\mathcal{I}_\textrm{D}^\textrm{iso}}{\Nant(\Nant-1)}\,\sum_{ij}\sum_{\substack{ab\\a\ne b\\a\in i, b\in j}} \mathcal{W}^{\textrm{G}\star}_a\,\mathcal{W}^\textrm{G}_b\,\mathcal{W}^\textrm{I}_a\,\mathcal{W}^{\textrm{I}\star}_b,
+  \mathcal{I}^\prime &\approx \frac{\mathcal{I}_\textrm{D}^\textrm{iso}}{\Nant(\Nant-1)}\,\sum_{pq}\sum_{\substack{ab\\a\ne b\\a\in p, b\in q}} \mathcal{W}^{\textrm{G}\star}_a\,\mathcal{W}^\textrm{G}_b\,\mathcal{W}^\textrm{I}_a\,\mathcal{W}^{\textrm{I}\star}_b,
 \end{align}
 where, $\mathcal{I}_\textrm{D}^\textrm{iso}$ is the same for all terms and has been pulled out of the summations.
-For all antennas indexed by $a$ that belong to a particular type $i$, we replace $\mathcal{W}_a$ with $\mathcal{W}_{(i)}$ and is applicable to those arising from both instrumental (superscript `$I$') and gridding (superscript `$G$') origins.
-If $n_{ij}$ is the number of antenna pairs in the antenna pair type $ij$, $\mathcal{W}_\textrm{eff}$ can be simplified as:
+For all antennas indexed by $a$ that belong to a particular type $p$, we replace $\mathcal{W}_a$ with $\mathcal{W}_{(p)}$ and is applicable to those arising from both instrumental (superscript `$I$') and gridding (superscript `$G$') origins.
+If $n_{pq}$ is the number of antenna pairs in the antenna pair type $pq$ {\bf (unrelated to the noise quantity also denoted by $n$ in \S\ref{sec:math})}, $\mathcal{W}_\textrm{eff}$ can be simplified as:
 \begin{align}\label{eqn:wt-dirty-image-EPIC-decomp-approx2}
-  \mathcal{W}_\textrm{eff} &\approx \frac{1}{\Nant(\Nant-1)}\,\sum_{ij}n_{ij}\,\mathcal{W}^{\textrm{G}\star}_{(i)}\,\mathcal{W}^\textrm{G}_{(j)}\,\mathcal{W}^\textrm{I}_{(i)}\,\mathcal{W}^{\textrm{I}\star}_{(j)}.
-\end{align}
-In an optimally weighted image, $\mathcal{W}^\textrm{G}_{(i)} \equiv \mathcal{W}^\textrm{I}_{(i)}$. Thus,
+  \mathcal{W}_\textrm{eff} &\approx \frac{1}{\Nant(\Nant-1)}\,\sum_{pq}n_{pq}\,\mathcal{W}^{\textrm{G}\star}_{(p)}\,\mathcal{W}^\textrm{G}_{(q)}\,\mathcal{W}^\textrm{I}_{(p)}\,\mathcal{W}^{\textrm{I}\star}_{(q)}.
+\end{align}
+In an optimally weighted image, $\mathcal{W}^\textrm{G}_{(p)} \equiv \mathcal{W}^\textrm{I}_{(q)}$. Thus,
 \begin{align}\label{eqn:effective-weighting-optimal}
-  \mathcal{W}_\textrm{eff}^\textrm{opt} &\approx \frac{1}{\Nant(\Nant-1)}\,\sum_{ij} n_{ij}\,\left|\mathcal{W}^\textrm{I}_{(i)}\right|^2\left|\mathcal{W}^\textrm{I}_{(j)}\right|^2,
+  \mathcal{W}_\textrm{eff}^\textrm{opt} &\approx \frac{1}{\Nant(\Nant-1)}\,\sum_{pq} n_{pq}\,\left|\mathcal{W}^\textrm{I}_{(p)}\right|^2\left|\mathcal{W}^\textrm{I}_{(q)}\right|^2,
 \end{align}
@@
 We consider a heterogeneous array consisting of a total of $\Nant$ antennas under $N_\textrm{T}$ different types.
-And under each antenna type $i$, there are $n_i$ antennas such that $\sum_i n_i = \Nant$.
+And under each antenna type $p$, there are $n_p$ antennas such that $\sum_p n_p = \Nant$ {\bf (unrelated to the noise quantity also denoted by $n$ in \S\ref{sec:math})}.
 The total number of unique non-zero spacing antenna pairs is $\Nant(\Nant-1)/2$ and there are potentially up to $N_\textrm{T}(N_\textrm{T}+1)/2$ unique antenna pair types obtained by pairwise combination of the $N_\textrm{T}$ antenna types.
@@
 \begin{align}\label{eqn:apndx-total-baselines}
-  \Nant(\Nant-1) &= \sum_{ij} n_{ij}, \\
-  \textrm{where}, \quad   n_{ij} &= \begin{cases}
-    n_i\,n_j, & i\ne j \\
-    n_i(n_i-1), & i=j, 
+  \Nant(\Nant-1) &= \sum_{pq} n_{pq}, \\
+  \textrm{where}, \quad   n_{pq} &= \begin{cases}
+    n_p\,n_q, & p\ne q \\
+    n_p(n_p-1), & p=q, 
   \end{cases}
 \end{align}
-is obtained by counting antenna pairs, $n_{ij}$, in different antenna pair types, $ij$.
+is obtained by counting antenna pairs, $n_{pq}$, in different antenna pair types, $pq$.
+{\bf Note that $ab$ and $pq$ act as simple indices into pairs formed from individual indices and not their product.}
 
@@
 \begin{align}\label{eqn:apndx-wt-dirty-image-EPIC-decomp}
-  \mathcal{I}^\prime &= \frac{1}{\Nant(\Nant-1)}\,\sum_{ij}\sum_{\substack{ab\\a\ne b\\a\in i, b\in j}}\,\mathcal{I}^\prime_{ab}
+  \mathcal{I}^\prime &= \frac{1}{\Nant(\Nant-1)}\,\sum_{pq}\sum_{\substack{ab\\a\ne b\\a\in p, b\in q}}\,\mathcal{I}^\prime_{ab}
 \end{align}
@@
 \begin{align}\label{eqn:apndx-wt-dirty-image-EPIC-decomp-approx1}
-  \mathcal{I}^\prime &\approx \frac{\mathcal{I}_\textrm{D}^\textrm{iso}}{\Nant(\Nant-1)}\,\sum_{ij}\sum_{\substack{ab\\a\ne b\\a\in i, b\in j}} \mathcal{W}^{\textrm{G}\star}_a\,\mathcal{W}^\textrm{G}_b\,\mathcal{W}^\textrm{I}_a\,\mathcal{W}^{\textrm{I}\star}_b,
+  \mathcal{I}^\prime &\approx \frac{\mathcal{I}_\textrm{D}^\textrm{iso}}{\Nant(\Nant-1)}\,\sum_{pq}\sum_{\substack{ab\\a\ne b\\a\in p, b\in q}} \mathcal{W}^{\textrm{G}\star}_a\,\mathcal{W}^\textrm{G}_b\,\mathcal{W}^\textrm{I}_a\,\mathcal{W}^{\textrm{I}\star}_b,
 \end{align}
 where, $\mathcal{I}_\textrm{D}^\textrm{iso}$ is the same for all terms and has been pulled outside the summations.
-For all antennas indexed by $a$ that belong to a particular type $i$, we replace $\mathcal{W}_a$ with $\mathcal{W}_{(i)}$ and is applicable to those arising from both instrumental (superscript `$I$') and gridding (superscript `$G$') origins.
+For all antennas indexed by $a$ that belong to a particular type $p$, we replace $\mathcal{W}_a$ with $\mathcal{W}_{(p)}$ and is applicable to those arising from both instrumental (superscript `$I$') and gridding (superscript `$G$') origins.
 Using arguments similar to those in equation~\ref{eqn:apndx-total-baselines}, equation~\ref{eqn:apndx-wt-dirty-image-EPIC-decomp-approx1} can be further simplified to:
 \begin{align}\label{eqn:apndx-wt-dirty-image-EPIC-decomp-approx2}
-  \mathcal{I}^\prime &\approx \frac{\mathcal{I}_\textrm{D}^\textrm{iso}}{\Nant(\Nant-1)}\,\sum_{ij}n_{ij}\,\mathcal{W}^{\textrm{G}\star}_{(i)}\,\mathcal{W}^\textrm{G}_{(j)}\,\mathcal{W}^\textrm{I}_{(i)}\,\mathcal{W}^{\textrm{I}\star}_{(j)}.
+  \mathcal{I}^\prime &\approx \frac{\mathcal{I}_\textrm{D}^\textrm{iso}}{\Nant(\Nant-1)}\,\sum_{pq}n_{pq}\,\mathcal{W}^{\textrm{G}\star}_{(p)}\,\mathcal{W}^\textrm{G}_{(q)}\,\mathcal{W}^\textrm{I}_{(p)}\,\mathcal{W}^{\textrm{I}\star}_{(q)}.
 \end{align}
@@
 \begin{align}
-  \mathcal{W}_\textrm{eff} &\approx \frac{1}{\Nant(\Nant-1)}\,\sum_{ij}n_{ij}\,\mathcal{W}^{\textrm{G}\star}_{(i)}\,\mathcal{W}^\textrm{G}_{(j)}\,\mathcal{W}^\textrm{I}_{(i)}\,\mathcal{W}^{\textrm{I}\star}_{(j)}.
-\end{align}
-In an optimally weighted image, $\mathcal{W}^\textrm{G}_{(i)} \equiv \mathcal{W}^\textrm{I}_{(i)}$. Then, equation~\ref{eqn:apndx-effective-weighting} reduces to:
+  \mathcal{W}_\textrm{eff} &\approx \frac{1}{\Nant(\Nant-1)}\,\sum_{pq}n_{pq}\,\mathcal{W}^{\textrm{G}\star}_{(p)}\,\mathcal{W}^\textrm{G}_{(q)}\,\mathcal{W}^\textrm{I}_{(p)}\,\mathcal{W}^{\textrm{I}\star}_{(q)}.
+\end{align}
+In an optimally weighted image, $\mathcal{W}^\textrm{G}_{(p)} \equiv \mathcal{W}^\textrm{I}_{(q)}$. Then, equation~\ref{eqn:apndx-effective-weighting} reduces to:
 \begin{align}\label{eqn:apndx-effective-weighting-optimal}
-  \mathcal{W}_\textrm{eff}^\textrm{opt} &\approx \frac{1}{\Nant(\Nant-1)}\,\sum_{ij} n_{ij}\,\left|\mathcal{W}^\textrm{I}_{(i)}\right|^2\left|\mathcal{W}^\textrm{I}_{(j)}\right|^2,
+  \mathcal{W}_\textrm{eff}^\textrm{opt} &\approx \frac{1}{\Nant(\Nant-1)}\,\sum_{pq} n_{pq}\,\left|\mathcal{W}^\textrm{I}_{(p)}\right|^2\left|\mathcal{W}^\textrm{I}_{(q)}\right|^2,
 \end{align}

### 2017-01-07 imaging_demo/ms.tex
Minor updates throughout
From b5e83c5 (mnras):
====================

In ms.tex,

(*) Specify data used is from LWA1
(*) Minor changes throughout

@@
 We have verified the images produced to be equivalent to those from traditional techniques to within a precision set by gridding coarseness.
-We have also validated our implementation on data observed with the Long Wavelength Array (LWA).
+We have also validated our implementation on data observed with the Long Wavelength Array (LWA1).
 {\bf We provide a detailed framework for imaging with heterogeneous arrays and show that EPIC} robustly estimates the input sky model for {\bf such} arrays.
@@
 For this reason traditional correlators have difficulty scaling to thousands of antennas.
-As an example, the full HERA correlator for 352 dishes with 200 MHz of bandwidth requires 212 trillion complex multiply-accumulates per second (TMACS).
+For example, the full HERA correlator for 352 dishes with 200 MHz of bandwidth requires 212 trillion complex multiply-accumulates per second (TMACS).
 Future arrays with thousands of collecting elements will require orders of magnitude more computation, making the correlator the dominant cost.
@@
 We then verify the output data quality from our code in \S\ref{sec:verify} by presenting simulated images from both the EPIC correlator and comparing to a simulated FX correlator.
-We also demonstrate the performance with real-world data from the LWA.
-{\bf In \S\ref{sec:versatility}, we present a mathematical framework for imaging with heterogeneous arrays and demonstrate the ability of EPIC to robustly image data from such arrays}.
+We also demonstrate the performance with real-world data from the LWA1.
+{\bf In \S\ref{sec:versatility}, we present a mathematical framework for imaging with heterogeneous arrays and demonstrate the ability of EPIC} to robustly image data from such arrays.
 In \S\ref{sec:analysis}, we explore the scalability of the algorithm in the context of several array design choices.
@@
 where, the baseline coordinate $\mathbf{r}_{ab}\equiv\mathbf{r}_a-\mathbf{r}_b$ is the vector separation between the two antennas.
-This signifies that the visibility ($V_{ab}$) measured between a pair of antennas is obtained by multiplying the sky brightness $\mathcal{I}(\hat{\mathbf{s}},f,t)$ by the antenna power response $\mathcal{B}_{ab}(\hat{\mathbf{s}},f,t)$ and Fourier transforming from the directional coordinates ($\hat{\mathbf{s}}$) to the measurement plane, which are then sampled at the locations of the antenna spacings (or baselines), namely, $\mathbf{r}_{ab}$, and added to the noise $n_{ab}$.
+This signifies that the visibility ($V_{ab}$) measured between a pair of antennas is obtained by multiplying the sky brightness $\mathcal{I}(\hat{\mathbf{s}},f,t)$ by the antenna power response $\mathcal{B}_{ab}(\hat{\mathbf{s}},f,t)$ and Fourier-transforming from the directional coordinates ($\hat{\mathbf{s}}$) to the measurement plane, which are then sampled at the locations of the antenna spacings (or baselines), namely, $\mathbf{r}_{ab}$, and added to the noise $n_{ab}$.
 
@@
 \end{align}
-where the sky brightness $\boldsymbol{\mathcal{I}}(\hat{\mathbf{s}})$ is Fourier transformed using $\mathbf{F}(\mathbf{u},\hat{\mathbf{s}})$ and the resultant spatial coherence function is weighted and summed using the antenna power response, $\mathbf{B}(\mathbf{v},\mathbf{u})$ in $uv$-space sampled at the baseline location to obtain the measured visibilities:
+where the sky brightness $\boldsymbol{\mathcal{I}}(\hat{\mathbf{s}})$ is Fourier-transformed using $\mathbf{F}(\mathbf{u},\hat{\mathbf{s}})$ and the resultant spatial coherence function is weighted and summed using the antenna power response, $\mathbf{B}(\mathbf{v},\mathbf{u})$ in $uv$-space sampled at the baseline location to obtain the measured visibilities:
 \begin{align}
@@
 
-We can then use the multiplication-convolution theorem to move the convolution in Equation~\ref{eqn:e-field-conv} to a square after the Fourier transform in equation~\ref{eqn:dirty-image-FX}.
+We can then use the multiplication-convolution theorem to move the convolution in Equation~\ref{eqn:e-field-conv} to a \sout{square} {\bf product} after the Fourier transform in equation~\ref{eqn:dirty-image-FX}.
 \begin{align}
@@
 The term inside the angular brackets before squaring has a very similar form as that in equation~\ref{eqn:dirty-image-FX}.
-It signifies that the measured antenna electric fields are weighted by the antenna noise, weighted and gridded by the antenna aperture kernel, Fourier transformed and finally squared to obtain the same image estimate that would have been obtained using equation~\ref{eqn:dirty-image-FX}.
+It signifies that the measured antenna electric fields are weighted by the antenna noise, weighted and gridded by the antenna aperture kernel, Fourier-transformed and finally squared to obtain the same image estimate that would have been obtained using equation~\ref{eqn:dirty-image-FX}.
 
@@
 EPIC can accept dual-polarization inputs and produce images of all four instrumental cross-polarizations.
-Currently, two data input formats exist for reading in the electric field time samples measured by the antennas -- simulated electric fields based on a sky model using the simulator packaged with EPIC; and LWA data.
+Currently, two data input formats exist for reading in the electric field time samples measured by the antennas -- simulated electric fields based on a sky model using the simulator packaged with EPIC; and LWA1 data.
 Efforts to build interfaces for data from other telescopes are underway.
@@
 Each of the complex antenna gains are calibrated to correct the corresponding electric field spectra.
-These calibrated electric fields are gridded using an antenna-based gridding convolution function after which it is spatially Fourier transformed and squared to obtain an image cube for every time stamp.
+These calibrated electric fields are gridded using an antenna-based gridding convolution function after which it is spatially Fourier-transformed and squared to obtain an image cube for every time stamp.
 These images are then time-averaged to obtain the accumulated image $\mathcal{I}^\prime(f)$ ($\boldsymbol{\mathcal{I}^\prime}(\hat{\mathbf{s}})$ in matrix notation).
@@
   \caption{Flowchart of MOFF imaging in EPIC.
-The propagated electric fields shown on the left are measured as time-series $\widetilde{E}_1(t)\ldots \widetilde{E}_\textrm{N}(t)$ by the antennas $\textrm{A}_1\ldots \textrm{A}_\textrm{N}$ which are then Fourier transformed by the F-engine to produce electric field spectra $E_1(f)\ldots E_\textrm{N}(f)$.
+The propagated electric fields shown on the left are measured as time-series $\widetilde{E}_1(t)\ldots \widetilde{E}_\textrm{N}(t)$ by the antennas $\textrm{A}_1\ldots \textrm{A}_\textrm{N}$ which are then Fourier-transformed by the F-engine to produce electric field spectra $E_1(f)\ldots E_\textrm{N}(f)$.
 They are calibrated and gridded.
@@
 Following the F-engine, the electric fields pass through the X-engine to produce visibilities $V_\textrm{ab}(f,t)$ which are calibrated and time-averaged.
-Then they are gridded to obtain the gridded visibilities $V_\textrm{g}(f)$ which are then Fourier transformed to obtain the image, $\mathcal{I}^\prime(f)$.}
+Then they are gridded to obtain the gridded visibilities $V_\textrm{g}(f)$ which are then Fourier-transformed to obtain the image, $\mathcal{I}^\prime(f)$.}
   \label{fig:FX-flowchart}
@@
 \noindent This module is common to the MOFF and visibility-based imaging techniques.
-Time samples of electric fields measured by the antenna and digitized by the A/D converter is Fourier transformed to generate electric field spectra.
+Time samples of electric fields measured by the antenna and digitized by the A/D converter is Fourier-transformed to generate electric field spectra.
 This step can be parallelized across antennas.
@@
 This method correlates the antenna electric field signals with an image pixel from the output of the correlator in the feedback calibration fashion outlined in \citealt{mor11} (illustrated in Fig.~\ref{fig:MOFF-flowchart} by the arrow leading from the imager to the calibration block).
-Furthermore it allows for arbitrarily complex sky models, and following the MOFF algorithm places no restriction on array layout, and accounts for non-identical antenna beam patterns.
-Direction dependent calibration can be achieved by correlating antenna signals with output pixels in the direction of $N_\textrm{c}$ calibration sources, then fitting for a functional model of the sky.
+Furthermore, it allows for arbitrarily complex sky models, \sout{and following the MOFF algorithm} places no restrictions on array layout, and accounts for non-identical antenna beam patterns.
+Direction-dependent calibration can be achieved by correlating antenna signals with output pixels in the direction of $N_\textrm{c}$ calibration sources, then fitting for a functional model of the sky.
 Since antennas are only correlated with calibrator pixels, the computational complexity scales as $\sim N_\textrm{a} N_\textrm{c}$.
@@
 where, $\mathbf{I}(\mathbf{a})$ is a row of ones denoting coverage of the grid by kernel footprints of antennas.
-This is achieved by efficient multiplication with the sparse matrix created in the antenna-to-grid mapping process using the sparse matrices module in Python SciPy package.
+This is achieved by efficient multiplication with the sparse matrix created in the antenna-to-grid mapping process using the sparse matrix module in Python SciPy package.
 Unless $\mathbf{W}(\mathbf{r})$ includes time-dependent effects of the ionosphere or the instrument, it needs to be computed just once for the entire observation.
@@
 
-In MOFF imaging, these are spatially Fourier transformed followed by a squaring operation at every time stamp for every frequency channel.
+In MOFF imaging, these are spatially Fourier-transformed followed by squaring at every time stamp for every frequency channel.
 In visibility-based imaging, the spatial Fourier transform is performed only once per integration time-scale and does not include a squaring operation.
@@
 Hence, they cannot be time-averaged to reduce noise.
-The statistical quantity stable with time in this case are the square of the holographic electric field images.
+The statistical quantity stable with time in this case is the square of the holographic electric field image.
 Thus, squared images have to be formed at every instant of time before averaging as indicated in equation~\ref{eqn:dirty-image-MOFF}.
@@
 In contrast, visibilities measured by an antenna are statistically stable within an integration time interval.
-Hence, they are averaged after calibration as shown in equation~\ref{eqn:cc-vis}.
+Hence, they are averaged after {\bf (and sometimes before)} calibration as shown in equation~\ref{eqn:cc-vis}.
 It is advantageous to average them in visibilities before imaging because the repeated cost of spatial FFT can be avoided.
@@
 In our simulations, we use 64 frequency channels each of width $\Delta f = 40$~kHz.
-The number of time stamps integrated in one integration cycle was kept at eight where each A/D timeseries is $1/\Delta f=25\,\mu$s long.
+The number of time stamps integrated in one integration cycle was kept at eight where {\bf the duration of} each \sout{A/D} timeseries is $1/\Delta f=25\,\mu$s long.
 We use the MWA array layout \citep{bea12} for demonstration.
 Only the inner 51 tiles within a square bounding box of 150~m on each side were used.
-We assumed all tiles are identical and have a square shaped electric field illumination footprint 4.4~m on each side.
+We assumed all tiles are identical and have a square-shaped electric field illumination footprint 4.4~m on each side.
 Besides the stochastic sky noise present in the simulated electric fields, no noise from the instrument is added.
@@
 Before the outputs can be compared, we describe the elimination of a distinct difference between the two techniques.
-The squaring operation under MOFF imaging in the image plane introduces antenna auto-correlations around the zero spacing in the $uv$-plane which are absent in traditional visibility-based imaging.
+The squaring operation under MOFF imaging in the image plane introduces antenna auto-correlations around the zero-spacing in the $uv$-plane which are absent in traditional visibility-based imaging.
 In order to facilitate a robust comparison between MOFF and visibility-based imaging techniques, these auto-correlations are removed from the MOFF algorithm output, which is otherwise not an essential part of the core algorithm.
@@
 
-We inverse Fourier transform the squared images and beams back to the $uv$ plane and subtract the estimated auto-correlation kernel scaled to the peak value centred at the zero spacing pixel.
-The final averaged image is obtained by Fourier transforming the $uv$ plane data and weights with the auto-correlations subtracted to the image plane.
+We perform the inverse Fourier transform of the squared images and beams back to the $uv$ plane and subtract the estimated auto-correlation kernel scaled to the peak value centred at the zero-spacing pixel.
+The final averaged image is obtained by Fourier-transforming the $uv$ plane data and weights with the auto-correlations subtracted to the image plane.
 These images are now comparable to those obtained from visibility-based imaging.
@@
 Fig.~\ref{fig:MOFF-FX-image} shows the weighted dirty images (top) and synthesized beams (bottom) obtained with antenna-based MOFF and FX visibility-based imaging algorithms packaged in EPIC.
-The antenna auto-correlations that correspond to zero spacing have been removed from the correlated weights and data in the $uv$ plane, the MOFF image and the corresponding synthesized beam as described in \S\ref{sec:rm-autocorr}.
+The antenna auto-correlations that correspond to zero-spacing have been removed from the correlated weights and data in the $uv$ plane, the MOFF image and the corresponding synthesized beam as described in \S\ref{sec:rm-autocorr}.
 The reconstructed sky image has the simulated sources at the expected sky positions in either case.
@@
 These differences which are dependent on the coarseness of grid spacing can be mitigated by making the grid spacing finer at the expense of increased computational cost.
-Residuals centred around zero spacing can also be lowered by subtracting each auto-correlation of antenna weights separately by using the shape and extent of the sampled footprint appropriate for that specific antenna aperture.
+Residuals centred around zero-spacing can also be lowered by subtracting each auto-correlation of antenna weights separately by using the shape and extent of the sampled footprint appropriate for that specific antenna aperture.
 This is a general solution applicable even in case of heterogeneous antenna arrays and is under active development for EPIC.
@@
 
-\subsection{Application to LWA data}\label{sec:LWA-data}
-
-Here we demonstrate our software using narrow band data from the LWA station in New Mexico.
-This data is in LWA narrow-band transient buffer (TBN) format from 255 antennas within roughly a diameter of 100~m.
+\subsection{Application to LWA1 data}\label{sec:LWA-data}
+
+Here we demonstrate our software using narrow band data from the LWA1 station in New Mexico.
+This data is in LWA1 narrow-band transient buffer (TBN) format from 255 antennas within roughly a diameter of 100~m.
 The data is centred at a frequency of 74.03~MHz, with a sample rate (equal to the bandwidth) of 100~kHz with 512 complex time samples per antenna in a A/D writeout time-scale of 5.12~ms, a frequency resolution of 195.3125~Hz and dual polarization.
@@
 
-Fig.~\ref{fig:LWA-image} shows the image produced with MOFF imaging packaged in EPIC after averaging over $\approx$~20~ms (four writeouts) of data and the inner $\approx 80$ per cent of bandwidth (roughly 80~kHz).
+Fig.~\ref{fig:LWA-image} shows the image produced with MOFF imaging packaged in EPIC after averaging over $\approx$~20~ms (four writeouts) of data and the inner $\approx 80$\% of bandwidth (roughly 80~kHz).
 The image is shown in direction cosine coordinates -- $l$ along the horizontal axis and $m$ along the vertical axis.
@@
   % \includegraphics[width=\columnwidth]{LWA_MOFF_bandavg_image_4_iterations_analytic_aperture.eps}
-  \caption{Image from LWA TBN data obtained with MOFF imaging using EPIC package after averaging over 20~ms and $\approx 80$~kHz.
+  \caption{Image from LWA1 TBN data obtained with MOFF imaging using EPIC package after averaging over 20~ms and $\approx 80$~kHz.
 The x- and y-axes denote direction cosines $l$ and $m$ respectively.
@@
 Of all the steps in the MOFF pipeline that are repeated for every writeout from the F-engine, the most expensive step even for dense HERA layouts is found to be the spatial two-dimensional FFT in the imaging stage relative to applying the sparse matrix gridding convolution, squaring or time-averaging.
-For instance, even in the conservative dense array layout scenario that makes these other stages even more expensive, the gridding convolution, squaring and time-averaging take up only $\lesssim 20$\%, $\lesssim 20$\% and $\lesssim 5$\% respectively of the total processing time while the spatial Fourier transform takes up $\gtrsim 55$ per cent of the total time.
+For instance, even in the conservative dense array layout scenario that makes these other stages even more expensive, the gridding convolution, squaring and time-averaging take up only $\lesssim 20$\%, $\lesssim 20$\% and $\lesssim 5$\% respectively of the total processing time while the spatial Fourier transform takes up $\gtrsim 55$\% of the total time.
 With sparser arrays the gridding process will be be even faster.
@@
 
-The number of complex multiplications and additions in the spatial Fourier transform implemented via FFT is $\approx \beta\,(4\Ngrid)\log_2(4\Ngrid)$ where $\Ngrid$ is the number of pixels on the grid, the factor 4 accounts for increase in number of pixels as a result of zero-padding before spatial Fourier transform, and $\beta$ is a constant that depends on the implementation of twiddle FFT algorithms \citep{bri74}.
+The number of complex multiplications and \sout{additions} {\bf accumulations} in the spatial Fourier transform implemented via FFT is $\approx \beta\,(4\Ngrid)\log_2(4\Ngrid)$ where $\Ngrid$ is the number of pixels on the grid, the factor 4 accounts for the increase in the number of pixels as a result of zero-padding before the spatial Fourier transform, and $\beta$ is a constant that depends on the implementation of twiddle FFT algorithms \citep{bri74}.
 In our study, we set $\beta=5$, a value\footnote{http://www.fftw.org/speed/method.html} much more conservative than was indicated in \citet{mor11}.
@@
     \item[a] MWA-N denotes N tiles in the specified core diameter
-    \item[b] LC and HC denotes low band and high band stations inside the 
-      specified core diameter 
+    \item[b] LC and HC denotes low band and high band stations inside the specified core diameter 
     \item[c] Owens Valley LWA
-    \item[d] Hypothetically chosen to have a total collecting area of 
-      1~km$^2$
-    \item[e] This is the number of beamformed stations expected to be in the 
-      core, roughly three-fourths of the total number
-    \item[f] All dipoles inside the core are used as independent elements 
-      without station beamforming
+    \item[d] Hypothetically chosen to have a total collecting area of 1~km$^2$
+    \item[e] This is the number of beamformed stations expected to be in the core, roughly three-fourths of the total number
+    \item[f] All dipoles inside the core are used as independent elements without station beamforming
     \item[g] Hydrogen Intensity mapping and Real-time Analysis eXperiment
@@
 All HERA layouts except HERA-19 and HERA-37 are in a region of parameter space where MOFF imaging holds the advantage.
-The dashed line showing future trajectory of HERA like systems will be clearly favoured by MOFF imaging.
+The dashed line showing future trajectory of HERA-like systems will be clearly favoured by MOFF imaging.
 The dotted line is similarly a hypothetical trajectory for the MWA with more tiles added inside the same core diameter.
@@
 It is bounded by the LWA1 and LWA-OV on the left and right respectively.
-The current (see Table~\ref{tab:antenna-layouts}) and a hypothetical expanded layout with a four-fold increase in number of elements over a 50 per cent increase in $b_\textrm{max}$ provide the bounds at the bottom and top respectively.
-Current instruments such as MWA and LOFAR lie in parameter space favouring visibility-based imaging.
+The current (see Table~\ref{tab:antenna-layouts}) and a hypothetical expanded layout with a four-fold increase in number of elements over a 50\% increase in $b_\textrm{max}$ provide the bounds at the bottom and top respectively.
+Current instruments such as MWA and LOFAR lie in parameter space favouring visibility-based imaging {\bf due to the sparseness of their layouts}.
 
@@
 The dotted line similarly shows hypothetical expanded MWA layouts with more tiles added in the same core.
-The grey shaded area denotes the projected trajectory of the LWA bounded by LWA1 (left edge), LWA-OV (right edge), current layout (bottom) and a four-fold increase in the number of elements within a 50 per cent increase in the core size (top).
-Current instruments such as MWA and LOFAR fall in a region favoured by visibility-based imaging.}
+The grey shaded area denotes the projected trajectory of the LWA bounded by LWA1 (left edge), LWA-OV (right edge), current layout (bottom) and a four-fold increase in the number of elements within a 50\% increase in the core size (top).
+Current instruments such as MWA and LOFAR fall in a region favoured by visibility-based imaging {\bf due to sparse layouts}.}
   \label{fig:parameter-space-computations-instruments}
@@
 Under the class of direct imaging techniques, ours is one of the most generic -- neither does it place any constraint on the array layout to be on a regular grid nor does it require the antenna array to be homogeneous.
-{\bf We have provided a detailed mathematical framework for imaging with heterogeneous arrays, demonstrated the natural ability of EPIC to robustly image data from such arrays, and shown that wrong assumptions about array homogeneity could result in significant and systematic mis-estimation of the sky model}.
+{\bf We have provided a detailed mathematical framework for imaging with heterogeneous arrays, demonstrated the natural ability of EPIC} to robustly image data from such arrays, and shown that wrong assumptions about array homogeneity could result in significant and systematic mis-estimation of the sky model.
 
@@
 It is designed to provide a development platform to compare different imaging approaches and serve as a stepping stone for real-life GPU/FPGA-based implementation on telescopes.
-It has been successfully tested on simulated MWA observations as well as real LWA observations from both imaging and calibration viewpoints.
+It has been successfully tested on simulated MWA observations as well as real LWA1 observations from both imaging and calibration viewpoints.
 
@@
 
-We thank Larry D'Addario, Gregg Hallinan, Joseph Lazio and Harish Vedantham for their valuable inputs, and Greg Taylor for providing us with LWA data.
+We thank Larry D'Addario, Gregg Hallinan, Joseph Lazio and Harish Vedantham for their valuable inputs, and Greg Taylor for providing us with LWA1 data.
 This work has been supported by the National Science Foundation through award AST-1206552.
@@
 There is one antenna module per antenna each having attributes -- the propagated electric field time-series, $\widetilde{E}(t)$, and spectrum $E(f)$ for both polarizations.
-The most important function inside this module is the F-engine that Fourier transforms time-series electric field data into spectra.
+The most important function inside this module is the F-engine that Fourier-transforms time-series electric field data into spectra.
 
@@
 The FX mode obtains the electric field spectra, $E(f)$ from the individual antenna modules inside this module and multiplies the two to obtain visibility spectra, $V(f)$.
-On the other hand, the XF mode cross-correlates the electric field time-series from its Antenna modules to obtain the visibilities as a function of lags, $V_t(\tau)$, which is then Fourier transformed to obtain $V(f)$.
+On the other hand, the XF mode cross-correlates the electric field time-series from its Antenna modules to obtain the visibilities as a function of lags, $V_t(\tau)$, which is then Fourier-transformed to obtain $V(f)$.
 Both modules can operate on dual-polarizations to obtain all four cross-polarizations.

### 2017-01-10 imaging_demo/ms.tex
Remove strikethroughs and bold fonts
From c6789d1 (mnras):
====================

In ms.tex,

(*) Remove strikethroughs and bold fonts

@@
 \usepackage[caption=false]{subfig} % For subfloats
-\usepackage[normalem]{ulem} % For strikethrough text
 
@@
 Through our generic and efficient E-field Parallel Imaging Correlator (EPIC), we present the first software demonstration of a generalized direct imaging algorithm, namely, the Modular Optimal Frequency Fourier (MOFF) imager.
-\sout{It takes advantage of the multiplication-convolution theorem of Fourier transforms.} Not only does it bring down the cost for dense layouts to $\mathcal{O}(\Nant\log_2\Nant)$ but can also image from irregular layouts and heterogeneous arrays of antennas.
+Not only does it bring down the cost for dense layouts to $\mathcal{O}(\Nant\log_2\Nant)$ but can also image from irregular layouts and heterogeneous arrays of antennas.
 EPIC is highly modular, parallelizable, implemented in object-oriented Python, and publicly available.
@@
 We have also validated our implementation on data observed with the Long Wavelength Array (LWA1).
-{\bf We provide a detailed framework for imaging with heterogeneous arrays and show that EPIC} robustly estimates the input sky model for {\bf such} arrays.
+We provide a detailed framework for imaging with heterogeneous arrays and show that EPIC robustly estimates the input sky model for such arrays.
 Antenna layouts with dense filling factors consisting of a large number of antennas such as LWA, the Square Kilometre Array, Hydrogen Epoch of Reionization Array, and Canadian Hydrogen Intensity Mapping Experiment will gain significant computational advantage by deploying an optimized version of EPIC.
@@
 Here we present the first software implementation of the MOFF correlator, and announce the public release of the E-field Parallel Imaging Correlator (EPIC) code.
-EPIC is a highly parallel, object-oriented Python package that primarily implements the MOFF imaging algorithm, \sout{as well as emulating} {\bf emulates} real-life telescopes and FX/XF correlators in software, and includes a visibility-based imaging technique for reference.
+EPIC is a highly parallel, object-oriented Python package that primarily implements the MOFF imaging algorithm, emulates real-life telescopes and FX/XF correlators in software, and includes a visibility-based imaging technique for reference.
 It is intended to provide a development platform to test different imaging approaches, characterize scaling relations and serve as a stepping stone for real-life GPU/FPGA-based implementation on telescopes.
@@
 We also demonstrate the performance with real-world data from the LWA1.
-{\bf In \S\ref{sec:versatility}, we present a mathematical framework for imaging with heterogeneous arrays and demonstrate the ability of EPIC} to robustly image data from such arrays.
+In \S\ref{sec:versatility}, we present a mathematical framework for imaging with heterogeneous arrays and demonstrate the ability of EPIC to robustly image data from such arrays.
 In \S\ref{sec:analysis}, we explore the scalability of the algorithm in the context of several array design choices.
@@
 
-We can then use the multiplication-convolution theorem to move the convolution in Equation~\ref{eqn:e-field-conv} to a \sout{square} {\bf product} after the Fourier transform in equation~\ref{eqn:dirty-image-FX}.
+We can then use the multiplication-convolution theorem to move the convolution in Equation~\ref{eqn:e-field-conv} to a product after the Fourier transform in equation~\ref{eqn:dirty-image-FX}.
 \begin{align}
@@
 This method correlates the antenna electric field signals with an image pixel from the output of the correlator in the feedback calibration fashion outlined in \citealt{mor11} (illustrated in Fig.~\ref{fig:MOFF-flowchart} by the arrow leading from the imager to the calibration block).
-Furthermore, it allows for arbitrarily complex sky models, \sout{and following the MOFF algorithm} places no restrictions on array layout, and accounts for non-identical antenna beam patterns.
+Furthermore, it allows for arbitrarily complex sky models, places no restrictions on array layout, and accounts for non-identical antenna beam patterns.
 Direction-dependent calibration can be achieved by correlating antenna signals with output pixels in the direction of $N_\textrm{c}$ calibration sources, then fitting for a functional model of the sky.
@@
 In contrast, visibilities measured by an antenna are statistically stable within an integration time interval.
-Hence, they are averaged after {\bf (and sometimes before)} calibration as shown in equation~\ref{eqn:cc-vis}.
+Hence, they are averaged after (and sometimes before) calibration as shown in equation~\ref{eqn:cc-vis}.
 It is advantageous to average them in visibilities before imaging because the repeated cost of spatial FFT can be avoided.
@@
 In our simulations, we use 64 frequency channels each of width $\Delta f = 40$~kHz.
-The number of time stamps integrated in one integration cycle was kept at eight where {\bf the duration of} each \sout{A/D} timeseries is $1/\Delta f=25\,\mu$s long.
+The number of time stamps integrated in one integration cycle was kept at eight where the duration of each time-series is $1/\Delta f=25\,\mu$s long.
 We use the MWA array layout \citep{bea12} for demonstration.
@@
 \end{align}
-where, $ab$ indexes pairs of antennas $a$ and $b$, {\bf $pq$ indexes the different types of antenna pairs, with $p$ and $q$ indexing antenna types (note that $ab$ and $pq$ are simply indices in our notation and are not to be interpreted as products of the individual indices)}.
+where, $ab$ indexes pairs of antennas $a$ and $b$, $pq$ indexes the different types of antenna pairs, with $p$ and $q$ indexing antenna types (note that $ab$ and $pq$ are simply indices in our notation and are not to be interpreted as products of the individual indices).
 Thus, equation~\ref{eqn:wt-dirty-image-homogeneous} is re-written as the sum of fringes over all unique antenna pairs in an antenna pair type (inner sum) and subsequently summed over all antenna pair types (outer sum).
@@
 For all antennas indexed by $a$ that belong to a particular type $p$, we replace $\mathcal{W}_a$ with $\mathcal{W}_{(p)}$ and is applicable to those arising from both instrumental (superscript `$I$') and gridding (superscript `$G$') origins.
-If $n_{pq}$ is the number of antenna pairs in the antenna pair type $pq$ {\bf (unrelated to the noise quantity also denoted by $n$ in \S\ref{sec:math})}, $\mathcal{W}_\textrm{eff}$ can be simplified as:
+If $n_{pq}$ is the number of antenna pairs in the antenna pair type $pq$ (unrelated to the noise quantity also denoted by $n$ in \S\ref{sec:math}), $\mathcal{W}_\textrm{eff}$ can be simplified as:
 \begin{align}\label{eqn:wt-dirty-image-EPIC-decomp-approx2}
@@
 
-The number of complex multiplications and \sout{additions} {\bf accumulations} in the spatial Fourier transform implemented via FFT is $\approx \beta\,(4\Ngrid)\log_2(4\Ngrid)$ where $\Ngrid$ is the number of pixels on the grid, the factor 4 accounts for the increase in the number of pixels as a result of zero-padding before the spatial Fourier transform, and $\beta$ is a constant that depends on the implementation of twiddle FFT algorithms \citep{bri74}.
+The number of complex multiplications and accumulations in the spatial Fourier transform implemented via FFT is $\approx \beta\,(4\Ngrid)\log_2(4\Ngrid)$ where $\Ngrid$ is the number of pixels on the grid, the factor 4 accounts for the increase in the number of pixels as a result of zero-padding before the spatial Fourier transform, and $\beta$ is a constant that depends on the implementation of twiddle FFT algorithms \citep{bri74}.
 In our study, we set $\beta=5$, a value\footnote{http://www.fftw.org/speed/method.html} much more conservative than was indicated in \citet{mor11}.
@@
 The current (see Table~\ref{tab:antenna-layouts}) and a hypothetical expanded layout with a four-fold increase in number of elements over a 50\% increase in $b_\textrm{max}$ provide the bounds at the bottom and top respectively.
-Current instruments such as MWA and LOFAR lie in parameter space favouring visibility-based imaging {\bf due to the sparseness of their layouts}.
+Current instruments such as MWA and LOFAR lie in parameter space favouring visibility-based imaging due to the sparseness of their layouts.
 
@@
 The grey shaded area denotes the projected trajectory of the LWA bounded by LWA1 (left edge), LWA-OV (right edge), current layout (bottom) and a four-fold increase in the number of elements within a 50\% increase in the core size (top).
-Current instruments such as MWA and LOFAR fall in a region favoured by visibility-based imaging {\bf due to sparse layouts}.}
+Current instruments such as MWA and LOFAR fall in a region favoured by visibility-based imaging due to sparse layouts.}
   \label{fig:parameter-space-computations-instruments}
@@
 Under the class of direct imaging techniques, ours is one of the most generic -- neither does it place any constraint on the array layout to be on a regular grid nor does it require the antenna array to be homogeneous.
-{\bf We have provided a detailed mathematical framework for imaging with heterogeneous arrays, demonstrated the natural ability of EPIC} to robustly image data from such arrays, and shown that wrong assumptions about array homogeneity could result in significant and systematic mis-estimation of the sky model.
+We have provided a detailed mathematical framework for imaging with heterogeneous arrays, demonstrated the natural ability of EPIC to robustly image data from such arrays, and shown that wrong assumptions about array homogeneity could result in significant and systematic mis-estimation of the sky model.
 
@@
 We consider a heterogeneous array consisting of a total of $\Nant$ antennas under $N_\textrm{T}$ different types.
-And under each antenna type $p$, there are $n_p$ antennas such that $\sum_p n_p = \Nant$ {\bf (unrelated to the noise quantity also denoted by $n$ in \S\ref{sec:math})}.
+And under each antenna type $p$, there are $n_p$ antennas such that $\sum_p n_p = \Nant$ (unrelated to the noise quantity also denoted by $n$ in \S\ref{sec:math}).
 The total number of unique non-zero spacing antenna pairs is $\Nant(\Nant-1)/2$ and there are potentially up to $N_\textrm{T}(N_\textrm{T}+1)/2$ unique antenna pair types obtained by pairwise combination of the $N_\textrm{T}$ antenna types.
@@
 \end{align}
-is obtained by counting antenna pairs, $n_{pq}$, in different antenna pair types, $pq$.
-{\bf Note that $ab$ and $pq$ act as simple indices into pairs formed from individual indices and not their product.}
+is obtained by counting antenna pairs, $n_{pq}$, in different antenna pair types, $pq$. Note that $ab$ and $pq$ act as simple indices into pairs formed from individual indices and not their product.
 

### 2017-02-16 cal_demo/ms.tex
Somewhere along the way a conjugate was dropped.
unbolded all changes from v2 to prep for v3

@@
 need for a real-time calibration method for direct imaging radio astronomy correlators. Direct 
-imaging involves a spatial fast Fourier transform of antenna \textbf{signals}, alleviating 
-\textbf{an 
+imaging involves a spatial fast Fourier transform of antenna signals, alleviating 
+an 
 $\mathcal{O}(\Nant^2)$ computational bottleneck typical in radio correlators, and yielding a more gentle $\mathcal{O}(\Nant \log_2 
-\Nant)$ scaling.} \textbf{This} can save orders of magnitude in computation cost for next generation arrays 
-consisting of hundreds to thousands of antennas. However, because \textbf{antenna} signals are mixed in the 
-correlator, gain correction must be applied \textbf{prior to imaging, rather than on visibilities post-correlation.} We develop the EPICal algorithm 
+\Nant)$ scaling. This can save orders of magnitude in computation cost for next generation arrays 
+consisting of hundreds to thousands of antennas. However, because antenna signals are mixed in the 
+correlator, gain correction must be applied prior to imaging, rather than on visibilities post-correlation. We develop the EPICal algorithm 
 to form gain solutions in real time without ever forming visibilities. This method scales as the 
 number of antennas, and produces results comparable to those from visibilities. 
-\textbf{We use simulations to demonstrate the EPICal technique and study the noise properties of
+We use simulations to demonstrate the EPICal technique and study the noise properties of
 our gain solutions, showing they are similar to visibility based solutions in realistic 
@@
 improvement compared to uncalibrated images, showing this algorithm is a promising
-solution for next generation instruments.}
+solution for next generation instruments.
 \end{abstract}
@@
 
-\textbf{The most common radio correlator designs of today, the FX and lag correlators \citep{rom99}},
-cross-multiply the \textbf{signals} from all pairs of \textbf{antennas.
-This} computation scales as the number of antennas squared, $\mathcal{O}(\Nant^2)$ 
+The most common radio correlator designs of today, the FX and lag correlators \citep{rom99},
+cross-multiply the signals from all pairs of antennas.
+This computation scales as the number of antennas squared, $\mathcal{O}(\Nant^2)$ 
 \citep{bun04}. As the number of elements in future arrays grows, the computational cost will 
@@
  generation cosmology experiments, this scaling is a large improvement over the $\Nant^2$ 
- scaling of traditional methods. Furthermore, because images are generated \textbf{in real-time}, the native 
- output bandwidth will be lowered \textbf{in cases where $\Ng < \Nant^2$}, and has the potential to be 
- lowered even further with \textbf{real-time} transient processing.
+ scaling of traditional methods. Furthermore, because images are generated in real-time, the native 
+ output bandwidth will be lowered in cases where $\Ng < \Nant^2$, and has the potential to be 
+ lowered even further with real-time transient processing.
 
@@
 A challenge common to all direct imaging algorithms is calibration of the antenna gains. With a 
-traditional \textbf{cross correlator}, visibilities are written to disk and used to calibrate 
-\textbf{after observing and before further processing} such as imaging. However, a direct imaging correlator mixes 
+traditional cross correlator, visibilities are written to disk and used to calibrate 
+after observing and before further processing such as imaging. However, a direct imaging correlator mixes 
 the signals from all antennas before averaging and writing to disk, making calibration a 
@@
 then demonstrate the algorithm in simulations in \S \ref{sec:sim}, 
-\textbf{and discuss noise trends in \S \ref{sec:noise}. We then apply the algorithm to a sample LWA 
-data set in \S \ref{sec:data}}.
+and discuss noise trends in \S \ref{sec:noise}. We then apply the algorithm to a sample LWA 
+data set in \S \ref{sec:data}.
 Finally we conclude and discuss potential extensions to the 
@@
 \caption{Data flow for the MOFF algorithm, reproduced from \citealt{thy15c}. The time ordered 
-electric field \textbf{measurements} are collected by the antenna elements then frequency transformed in the 
+electric field measurements are collected by the antenna elements then frequency transformed in the 
 F-engine. The calibration is applied, then the electric field measurements are gridded using the 
-\textbf{antenna aperture illumination patterns}. The imager performs a spatial 2D FFT on the gridded data before 
+antenna aperture illumination patterns. The imager performs a spatial 2D FFT on the gridded data before 
 squaring and averaging. The accumulated images are then written to disk.}
@@
 %%% New text
-\textbf{We begin by considering the electric field vector incident on the ground as a function of time, 
+We begin by considering the electric field vector incident on the ground as a function of time, 
 $\mathbf{E}(\mathbf{r},t)$. The position vector, $\mathbf{r}$, is measured from some arbitrary 
@@
 functions. We retain the time dependence because in practice each Fourier component will only be 
-measured over discrete sampling intervals. }
-
-\textbf{We will assume all antennas lie in a plane, 
+measured over discrete sampling intervals.
+
+We will assume all antennas lie in a plane, 
 reducing the problem to two dimensions. While the MOFF algorithm in principle can account for 
 non-coplanar arrays through the W-projection algorithm \citep{cor08}, the EPIC code does not 
-currently support this so we leave it to future work.}
-
-\textbf{Next we treat the electric field as a scalar quantity, ignoring polarization effects. This allows us 
+currently support this so we leave it to future work.
+
+Next we treat the electric field as a scalar quantity, ignoring polarization effects. This allows us 
 to express the electric field as a scalar, $E(\mathbf{r},f,t)$, and treat subsequent field 
@@
 require a sky model expressed in the instrumental polarization basis. We can therefore treat 
-each instrument polarization independently.}
-
-\textbf{Finally we assume the sources of the electric field are far away, allowing us to describe the 
+each instrument polarization independently.
+
+Finally we assume the sources of the electric field are far away, allowing us to describe the 
 electric fields propagating from sky sources as originating from an imaginary celestial sphere 
@@
 By making the latter assumptions we have left treatment of the ionosphere and terrestrial 
-emission such as local interference or thermal radiation from the ground for future work. }
-
-\textbf{Next we turn to the measurement performed by an antenna with frequency dependent far-field radiation pattern 
+emission such as local interference or thermal radiation from the ground for future work.
+
+Next we turn to the measurement performed by an antenna with frequency dependent far-field radiation pattern 
 $\mathcal{W}_a(\hat{\mathbf{s}},f)$, where the subscript $a$ denotes a specific antenna. We 
@@
 concepts of the primary beam (the product of two antenna radiation patterns) and the array
-point spread function (square of the Fourier transform of the array layout).}
-
-\textbf{The first step of the MOFF algorithm is to digitize the antenna signals and perform an FFT to 
+point spread function (square of the Fourier transform of the array layout).
+
+The first step of the MOFF algorithm is to digitize the antenna signals and perform an FFT to 
 channelize the time-ordered data. This is effectively estimating the quasi-monochromatic 
@@
 correlator which performs the frequency transform (F-engine) before cross multiplying 
-(X-engine). }
-
-Thus far we have \textbf{considered the perfectly measured electric field distribution according to 
-the antenna far-field radiation patterns}. Here we introduce a multiplicative complex gain as 
-well as an additive noise term which corrupt the measured signals \textbf{at each of the antennas}.
+(X-engine).
+
+Thus far we have considered the perfectly measured electric field distribution according to 
+the antenna far-field radiation patterns. Here we introduce a multiplicative complex gain as 
+well as an additive noise term which corrupt the measured signals at each of the antennas.
 \begin{equation}\label{eq:apply_gain}
@@
 \end{equation}
-\textbf{We have carried the frequency dependence to this point to make clear the complex gains are
-frequency dependent. However,} each subsequent step treats each frequency 
+We have carried the frequency dependence to this point to make clear the complex gains are
+frequency dependent. However, each subsequent step treats each frequency 
 channel independently, so we will drop the $f$ to simplify notation. 
-\textbf{Furthermore, while antenna gains may vary with time, we will assume they are constant over
-the period of our calibration, and thus drop their time dependence from the notation.}
-
-The MOFF algorithm next calls for a calibration \textbf{of the electric field measurements}. The goal of our new calibration method will be 
-to form an estimate of the gains, $g'_a$, \textbf{where the prime represents an estimate}. For now we will assume we have formed an estimate 
+Furthermore, while antenna gains may vary with time, we will assume they are constant over
+the period of our calibration, and thus drop their time dependence from the notation.
+
+The MOFF algorithm next calls for a calibration of the electric field measurements. The goal of our new calibration method will be 
+to form an estimate of the gains, $g'_a$, where the prime represents an estimate. For now we will assume we have formed an estimate 
 to proceed with the MOFF pipeline. We correct the incoming electric field data stream using our 
@@
 
-The next step is to grid the \textbf{calibrated electric field measurements to a finite set of regularly spaced 
+The next step is to grid the calibrated electric field measurements to a finite set of regularly spaced 
 positions on the plane of the ground, $\mathbf{r}_j$.
 The extent of the grid is chosen to encompass the footprint of the antenna array.
-MOFF uses the antenna aperture illumination patterns as the} gridding 
+MOFF uses the antenna aperture illumination patterns as the gridding 
 kernel according to the software holography/A-transpose technique \citep{mor09,bha08}. 
-\textbf{Mathematically we can express the gridding operation as}
+Mathematically we can express the gridding operation as
 \begin{equation}\label{eq:gridding}
@@
 \end{equation}
-\textbf{Formally the value at each pixel is calculated by summing the antenna
+Formally the value at each pixel is calculated by summing the antenna
 measurements over all antennas, each weighted by its respective aperture illumination
@@
 will only contain contribution from one antenna, allowing equation~\ref{eq:gridding}
-to be calculated through a sparse matrix multiplication.}
+to be calculated through a sparse matrix multiplication.
 This gridding 
@@
 After gridding, the imaging portion of the MOFF performs a 2D spatial FFT to form 
-instantaneous electric field images of the sky \textbf{at fixed locations $\s_k$.
+instantaneous electric field images of the sky at fixed locations $\s_k$.
 The FFT is possible because we have gridded to regularly spaced samples in a 2D plane.
@@
 Being direction cosines, grid locations with ${l^2+m^2>1}$ are non-physical and should
-be ignored.}
+be ignored.
 
@@
 \end{align}
-The sum over antennas uses the \textbf{antenna aperture illumination pattern} to grid the calibrated electric field \textbf{measurements} onto 
-regular gridpoints, $\mathbf{r}_j$. The \textbf{sum over $j$} denotes the 2D FFT to sky coordinates, resulting in an 
+The sum over antennas uses the antenna aperture illumination pattern to grid the calibrated electric field measurements onto 
+regular gridpoints, $\mathbf{r}_j$. The sum over $j$ denotes the 2D FFT to sky coordinates, resulting in an 
 estimate for the instantaneous electric field image. 
-\textbf{The sums over $a$ and $j$ mix the antenna signals together, which is the primary 
+The sums over $a$ and $j$ mix the antenna signals together, which is the primary 
 reason calibration must be performed beforehand where individual antennas retain their 
-identities.}
-
-We can simplify \textbf{the above} expression by 
-exchanging the sums \textbf{and use the DFT to approximate the Fourier relationship between
-the antenna aperture illumination pattern and the far-field radiation pattern.}
+identities.
+
+We can simplify the above expression by 
+exchanging the sums and use the DFT to approximate the Fourier relationship between
+the antenna aperture illumination pattern and the far-field radiation pattern.
 \begin{align}\label{eq:epix}
@@
 \end{align}
-In this form we see that the effect of gridding with the antenna \textbf{aperture illumination pattern} is to attenuate 
-the image by a factor of the \textbf{far-field radiation response, $\mathcal{W}_a(\s)$.}
-\textbf{This added factor of the instrument response is sometimes referred to as the
-\emph{holographic frame} in the literature \citep[e.g.][]{mor09, sul12}.}
-
-Finally, the images are squared and averaged in time to form \textbf{dirty images, 
-$I'(\hat{\mathbf{s}}_k)$, then} written 
+In this form we see that the effect of gridding with the antenna aperture illumination pattern is to attenuate 
+the image by a factor of the far-field radiation response, $\mathcal{W}_a(\s)$.
+This added factor of the instrument response is sometimes referred to as the
+\emph{holographic frame} in the literature \citep[e.g.][]{mor09, sul12}.
+
+Finally, the images are squared and averaged in time to form dirty images, 
+$I'(\hat{\mathbf{s}}_k)$, then written 
 to disk. While operations have been reordered, the resulting image is equivalent to one 
@@
 antennas to a spatial FFT, the computational cost is significantly lowered for certain classes of 
-compact arrays. The output bandwidth is also greatly reduced \textbf{in many scenarios} as visibilities are \textbf{replaced with} 
-gridded images. \textbf{These scalings were explored in \citealt{thy15c}.}
+compact arrays. The output bandwidth is also greatly reduced in many scenarios as visibilities are replaced with
+gridded images. These scalings were explored in \citealt{thy15c}.
 
@@
 signal from all antennas coherently, and thus requires the calibration to be applied in real time 
-\textbf{before imaging}. Furthermore, visibilities are never formed, which are 
+before imaging. Furthermore, visibilities are never formed, which are 
 traditionally the basic measurement used to form calibration solutions. Here we derive an 
@@
 form an estimate which will be applied to the subsequent stream of data. 
-\textbf{We assume that the changes in gains are negligible over adjacent intervals, which 
+We assume that the changes in gains are negligible over adjacent intervals, which 
 in some cases may limit the length of data streams.
-Because data from one interval is used to update the gain solutions for the next,} the 
+Because data from one interval is used to update the gain solutions for the next, the 
 process of finding a solution is an iterative one. We will use parenthetical superscripts to 
@@
 As a starting point, we consider the feedback calibration outlined in \citealt{mor11}. There it 
-was suggested to form a correlation of the uncalibrated antenna measurements with the \textbf{complex conjugate of} an image 
+was suggested to form a correlation of the uncalibrated antenna measurements with the complex conjugate of an image 
 pixel from the output of the correlator. This is a statistically stationary quantity, which can be 
 related to the sum of visibilities involving the antenna used in the correlation -- exactly the sum 
-needed to calibrate \textbf{in the case of a single point source sky}. However, we aim for a more 
+needed to calibrate in the case of a single point source sky. However, we aim for a more 
 generalized solution for arbitrarily complex sky models. We therefore study the full expression 
@@
 \begin{equation}\label{eq:Kna_def}
-\Kna \equiv \left<E_a(t) \mathcal{E}'(\spix,t)\right>_{t^{(n)}},
+\Kna \equiv \left<E_a(t) \mathcal{E}'^*(\spix,t)\right>_{t^{(n)}},
 \end{equation}
@@
 calibration loop,
-\textbf{and we use $t^{(n)}$ outside the brackets on the right-hand side to indicate all quantities
-within the brackets are drawn from time samples in the $n^\mathrm{th}$ calibration loop.}
-\textbf{The superscript $^*$ indicates complex conjugation,}
+and we use $t^{(n)}$ outside the brackets on the right-hand side to indicate all quantities
+within the brackets are drawn from time samples in the $n^\mathrm{th}$ calibration loop.
+The superscript $^*$ indicates complex conjugation,
 and \spix\, is the pixel center nearest a bright calibrator of interest. The 
 following will hold for any chosen pixel, \spix, though it is advantageous to choose a pixel which 
-contains a bright source to achieve a high signal to noise \textbf{ratio}.
+contains a bright source to achieve a high signal to noise ratio.
 
 Plugging equation \ref{eq:epix} into equation \ref{eq:Kna_def}, 
-\textbf{and substituting the $g^{(n)}_b$ for $g'_b$ to indicate the specific loop number,}
+and substituting the $g^{(n)}_b$ for $g'_b$ to indicate the specific loop number,
 we find,
@@
 where in the second step we group time-dependent terms, 
-and in the third \textbf{step we expanded the product within the time average.
+and in the third step we expanded the product within the time average.
 Because the noise does not correlate with the ideal antenna measurements we can
@@
 is zero mean. Then by defining an ideal visibility as
-$V_{ab}^I \equiv \left<E^I_a(t)E^{*I}_b(t)\right>_t$, we can simplify to}
+$V_{ab}^I \equiv \left<E^I_a(t)E^{*I}_b(t)\right>_t$, we can simplify to
 \begin{align}\label{eq:Kna}
@@
 \end{align}
-\textbf{where $\delta_{ab}$ is the Kronecker-delta function. To be clear, visibilities are never directly 
+where $\delta_{ab}$ is the Kronecker-delta function. To be clear, visibilities are never directly 
 formed from the data. Rather it is convenient to express $\Kna$ in terms of the ideal 
-visibilities, which we can model when solving for the gains below.}
-
-\textbf{We will ultimately wish to model the right side of equation~\ref{eq:Kna} in order to solve for the
+visibilities, which we can model when solving for the gains below.
+
+We will ultimately wish to model the right side of equation~\ref{eq:Kna} in order to solve for the
 antenna gains. However, the self-correlated noise, $\left<|n_a(t)|^2\right>$, can often be 
@@
 uncalibrated antenna electric field measurements to remove the corresponding
-terms from equation~\ref{eq:Kna}:}
+terms from equation~\ref{eq:Kna}:
 \begin{align}\label{eq:Cna}
@@
 \end{align}
-\textbf{Due to the above subtraction, the sum is now over $b \ne a$, explicitly removing the
-dependence on the self-correlated antenna signals.}
+Due to the above subtraction, the sum is now over $b \ne a$, explicitly removing the
+dependence on the self-correlated antenna signals.
 
@@
 the right hand side of equation~\ref{eq:Cna} by assuming our current gain estimates are 
-approximately correct, $g^{(n)}_b\approx g_b$, and replacing the \textbf{ideal} visibilities with a set of 
+approximately correct, $g^{(n)}_b\approx g_b$, and replacing the ideal visibilities with a set of 
 model visibilities, $V^M_{ab}$. These model visibilities rely on knowledge of the sky, and should 
-be precomputed \textbf{before observing}. We can then solve equation~\ref{eq:Cna} for $g_a$ to achieve an 
+be precomputed before observing. We can then solve equation~\ref{eq:Cna} for $g_a$ to achieve an 
 updated estimate.
@@
 This equation is our prescription for estimating the antenna gains of a direct imaging array. The 
-\textbf{real-time} computation complexity scales as $\mathcal{O}(\Nant)$ as we form 
-\textbf{an antenna-pixel correlation,} $\Cna$, for each 
+real-time computation complexity scales as $\mathcal{O}(\Nant)$ as we form 
+an antenna-pixel correlation, $\Cna$, for each 
 antenna.
@@
 We show schematically the process of calibrating a direct imaging correlator in 
-Fig.~\ref{fig:schematic}. Computationally expensive steps that must be performed \textbf{in near real-time} 
+Fig.~\ref{fig:schematic}. Computationally expensive steps that must be performed in near real-time 
 are shown inside the gray box. The uncalibrated antenna signals are tapped out after the 
 F-engine and correlated against the output image pixel of interest (i.e., equation~\ref{eq:Cna} is 
-\textbf{computed in near real-time}). The correlated values are then \textbf{used to estimate the gains}
+computed in near real-time). The correlated values are then used to estimate the gains
 using equation~\ref{eq:cal_solution}, and additional fitting if desired 
-\textbf{(see Section \ref{sec:discussion}).}
-\textbf{This step does not have
+(see Section \ref{sec:discussion}).
+This step does not have
 the stringent cadence requirement of the steps in the gray box because the 
 antenna-pixel correlations have been 
-averaged in time.}
+averaged in time.
 The gains are then passed back 
@@
 from the (unsquared) image is tapped out and correlated against the input antenna electric field 
-\textbf{measurements} to form $\Cna$ coefficients (equation~\ref{eq:Cna}). These coefficients are then 
-\textbf{used to update the gain estimates using} equations~\ref{eq:cal_solution}~and~\ref{eq:damping},
-and \textbf{any additional fitting (Section \ref{sec:discussion}).}
+measurements to form $\Cna$ coefficients (equation~\ref{eq:Cna}). These coefficients are then 
+used to update the gain estimates using equations~\ref{eq:cal_solution}~and~\ref{eq:damping},
+and any additional fitting (Section \ref{sec:discussion}).
 The gray box shows operations which must be done at high speed (before averaging in time),
-\textbf{while the} white boxes show operations which can be performed \textbf{at lower speeds because
-they involve time averaged quantities, and need to be updated on slower timescales.}
+while the white boxes show operations which can be performed at lower speeds because
+they involve time averaged quantities, and need to be updated on slower timescales.
 }
@@
 
-While testing we found equation~\ref{eq:cal_solution} resulted in oscillatory gain solutions \textbf{over time,} 
+While testing we found equation~\ref{eq:cal_solution} resulted in oscillatory gain solutions over time, 
 as is often the case in iterative methods. To mitigate this we 
@@
  Once the loop converges the damped version is 
-essentially a weighted average over the past several iterations, giving \textbf{the solutions} a longer effective 
+essentially a weighted average over the past several iterations, giving the solutions a longer effective 
 integration time.
@@
 We conclude this section by connecting our calibration expression to that found in a visibility 
-framework. In the limit of a single bright calibrating source \textbf{at zenith,} we can greatly 
-simplify equations~\ref{eq:Cna} and~\ref{eq:cal_solution}. We will assume the \textbf{radiation patterns} are 
+framework. In the limit of a single bright calibrating source at zenith, we can greatly 
+simplify equations~\ref{eq:Cna} and~\ref{eq:cal_solution}. We will assume the radiation patterns are 
 normalized such that $\mathcal{W}(0)=1$. We can further drop the exponential phase terms 
-because $\spix$ \textbf{is perpendicular to our planar antenna array.} We then absorb the true gains into the true visibilities in 
+because $\spix$ is perpendicular to our planar antenna array. We then absorb the true gains into the true visibilities in 
 equation~\ref{eq:Cna} to express as a sum of measured, uncalibrated, visibilities.
@@
 The ability to recover the equivalent expression despite not actually forming the visibilities is a 
-result of the fact that only sums over visibilities come into the \textbf{visibility-based} solution, as was described in 
+result of the fact that only sums over visibilities come into the visibility-based solution, as was described in 
 \citealt{mor11}. We have confirmed the limiting case equivalence here, and will explore the 
-more general case in more detail in \textbf{the following sections.}
+more general case in more detail in the following sections.
 
@@
 created for each antenna with random phase and amplitude, which is used to corrupt the 
-simulated data stream, then we attempt to \textbf{estimate} the gains using our calibration routine. The 
+simulated data stream, then we attempt to estimate the gains using our calibration routine. The 
 simulation software used is included in the EPIC package.
 
-\textbf{Our simulated antenna 
-array consists of} the inner 51 antennas of the MWA layout \citep{bea12}, within a bounding 
+Our simulated antenna 
+array consists of the inner 51 antennas of the MWA layout \citep{bea12}, within a bounding 
 box of 150~m. 
-\textbf{The physical parameters of our simulation can be arbitrarily scaled to other cases. However, 
+The physical parameters of our simulation can be arbitrarily scaled to other cases. However, 
 we provide typical units for an MWA observation as an example. 
-The antenna aperture illumination pattern} used is a 4.4~m square tophat; \textbf{a very rough 
-approximation to the square shaped tiles of the MWA.} Because our algorithm treats frequency 
-channels independently, we simulate only one channel. For context we treat this as \textbf{a channel
+The antenna aperture illumination pattern used is a 4.4~m square tophat; a very rough 
+approximation to the square shaped tiles of the MWA. Because our algorithm treats frequency 
+channels independently, we simulate only one channel. For context we treat this as a channel
 at 150~MHz with 40 kHz bandwidth, and adopt a sampling period of 25~$\mu$s. The simulated signal 
 consists of 10 random point sources with flux densitities 0.5~Jy~$\lesssim S \lesssim$~1~Jy 
-within the main lobe of the primary beam.}
+within the main lobe of the primary beam.
 
 For our unknown gains, we create a set of random complex numbers where the amplitude is 
-drawn from \textbf{a gaussian distribution centered around unity with width 0.25, and truncated at zero. The gain phases are drawn from a uniform distribution in the range $[-\pi, \pi)$. }
+drawn from a gaussian distribution centered around unity with width 0.25, and truncated at zero. The gain phases are drawn from a uniform distribution in the range $[-\pi, \pi)$. 
 These are our ``true gains", and we apply them to the frequency
-domain simulated antenna electric field \textbf{measurements} as in equation~\ref{eq:apply_gain}. Our analysis is 
+domain simulated antenna electric field measurements as in equation~\ref{eq:apply_gain}. Our analysis is 
 blind to these values until the end of the process to check accuracy. The gain estimates are 
@@
 
-We next process and image \caliter~time \textbf{samples} (10~ms). We also form the \textbf{antenna-pixel} correlations, 
+We next process and image \caliter~time samples (10~ms). We also form the antenna-pixel correlations, 
 \Cna[0], used in our calibration loop. The pixel used for the correlation is the source with the 
 largest apparent flux (intrinsic flux attenuated by the primary beam). These correlation values 
-are used to update the gain estimates. \textbf{We create perfect model 
+are used to update the gain estimates. We create perfect model 
 visibilities of our 10 simulated sources to be used in equation~\ref{eq:damping}. 
-The updated gain estimates are} used to calibrate the following 
-\caliter~time samples. Through \textbf{trial and error} we found a damping factor of $\damp=0.35$ 
+The updated gain estimates are used to calibrate the following 
+\caliter~time samples. Through trial and error we found a damping factor of $\damp=0.35$ 
 resulted in the quickest convergence in this simulation.
 
-The calibration loop continues by updating the gain estimates every \caliter~time \textbf{samples.} The 
-\textbf{phase errors} of our gain estimates are shown in Fig.~\ref{fig:sim_phase} for 20 such iterations. 
+The calibration loop continues by updating the gain estimates every \caliter~time samples. The 
+phase errors of our gain estimates are shown in Fig.~\ref{fig:sim_phase} for 20 such iterations. 
 The phase error plotted is the phase relative to the true gain for each antenna (various colored 
@@
 error. The other 50 antennas are shown to have error spanning the range $[-\pi,\pi)$ initially, and after about 
-10 iterations lock into a solution, settling down to noise levels around iteration \textbf{12 (0.12~s)}. We 
+10 iterations lock into a solution, settling down to noise levels around iteration 12 (0.12~s). We 
 stop the simulation when the updated gains trace the thermal noise of the simulated sources, 
@@
 gains were initialized with random phases, but the calibration loop was able to recover the 
-correct phases after about \textbf{12} iterations. Each line represents an antenna in our 51 MWA 
+correct phases after about 12 iterations. Each line represents an antenna in our 51 MWA 
 antenna sample.
@@
 g_a\right|$, which places all antennas on the same scale. We can see the amplitudes 
-converge toward their true values around the same time as the phases (iteration \textbf{$\sim$12}). At 
-the beginning of calibration we can see the \textbf{importance} of the damping factor. At $n=0$, a couple of 
+converge toward their true values around the same time as the phases (iteration $\sim$12). At 
+the beginning of calibration we can see the importance of the damping factor. At $n=0$, a couple of 
 gains are shown to have abnormally high amplitude estimates, notably one about 3.3 times its 
@@
 \includegraphics[width=\columnwidth]{fig4.pdf}
-\caption{\textbf{Same as Fig.~\ref{fig:sim_phase}, but for gain amplitudes.
-Each} line represents an antenna in the 51 MWA antenna sample. 
-\textbf{The relative amplitude plotted is the magnitude of the estimated gains divided by the true gains.}
-After about \textbf{12} iterations we see the 
+\caption{Same as Fig.~\ref{fig:sim_phase}, but for gain amplitudes.
+Each line represents an antenna in the 51 MWA antenna sample. 
+The relative amplitude plotted is the magnitude of the estimated gains divided by the true gains.
+After about 12 iterations we see the 
 calibration loop has settled around the correct values, with only noise remaining.}
@@
 fluctuate coherently. This is due to the stochastic fluctuations of the sources 
-themselves, \textbf{which are simulated as zero mean random processes with variance proportional
-to their flux density.}
+themselves, which are simulated as zero mean random processes with variance proportional
+to their flux density.
 For this simulation we restricted each calibration 
@@
 Images created at the beginning of calibration and at the end are shown in 
-Fig.~\ref{fig:sim_images}. Each image is obtained over a 400 time sample \textbf{(10 ms)} integration, corresponding to 
-all snapshot images created with a given set of gain estimates. The \textbf{left} panel shows the image 
+Fig.~\ref{fig:sim_images}. Each image is obtained over a 400 time sample (10 ms) integration, corresponding to 
+all snapshot images created with a given set of gain estimates. The left panel shows the image 
 produced with our initialized unity gains. Because the phases are completely random, the 
 image is essentially noise with the primary beam evident. After 20 calibration iterations, the image is far 
-more clear, shown in the \textbf{middle} panel. 
-\textbf{The ten sources are visible, along with rumble throughout the image. By comparing with
+more clear, shown in the middle panel. 
+The ten sources are visible, along with rumble throughout the image. By comparing with
 a perfectly calibrated image (right panel), we see that the rumble and the shapes of the
 point sources are matched between the two images. We therefore conclude that the 
-artifacts visible are dominated by the point spread function, rather than calibration errors.}
+artifacts visible are dominated by the point spread function, rather than calibration errors.
 
@@
 \caption{Images formed during simulated calibration. \emph{Left:} An image generated from a 
-single frequency channel and \textbf{400 time sample (10 ms)} integration after the gain estimates are randomized. As 
+single frequency channel and 400 time sample (10 ms) integration after the gain estimates are randomized. As 
 expected with random phases, the image is completely noisy, with the shape of the primary 
 beam evident. \emph{Middle:} An image formed after calibration, again with a single frequency 
-channel and \textbf{400 time sample} integration. \textbf{Now the simulated point sources are visible in the image. 
+channel and 400 time sample integration. Now the simulated point sources are visible in the image. 
 \emph{Right:} An image generated from the same simulated data as the middle image, but 
 with perfect calibration correction. Many of the features seen in the middle image are also
-seen here, confirming that they are point spread function artifacts. The ten simulated (and modeled) point sources are highlighted with red circles.}
+seen here, confirming that they are point spread function artifacts. The ten simulated (and modeled) point sources are highlighted with red circles.
 \label{fig:sim_images}
@@
 \section{Noise trends}\label{sec:noise}
-\textbf{Next we} study the effect of noise in our system, and the consequences 
+Next we study the effect of noise in our system, and the consequences 
 of an incomplete sky model. We run a suite of simulations varying the receiver noise and 
 integration times, while forming antenna-pixel correlations, $\Cna$, for EPICal gain solutions, 
-and simultaneously forming visibilities from the same streams \textbf{of measured electric fields} to find visibility-based 
+and simultaneously forming visibilities from the same streams of measured electric fields to find visibility-based 
 gain solutions for comparison.
@@
 these sources randomly, but kept them fixed for each run. 
-\textbf{We simulate the same antenna layout as in Section~\ref{sec:sim}. 
-However, we increase the number of frequency channels simulated to 64, while keeping the flux of all sources constant across frequency.}
+We simulate the same antenna layout as in Section~\ref{sec:sim}. 
+However, we increase the number of frequency channels simulated to 64, while keeping the flux of all sources constant across frequency.
 Because our calibration loop treats 
@@
 To simulate receiver noise, we add a gaussian distributed complex random number to each 
-antenna electric field measurement at each time \textbf{sample} according to 
+antenna electric field measurement at each time sample according to 
 equation~\ref{eq:apply_gain}. The level of the receiver noise, $\sigma_r = \left<\left|
 n_a(f,t)\right|^2\right>$, is varied in different simulation runs. We include two limiting 
-cases, \textbf{one} where the receiver noise power is subdominant to the sky power ($\sigma_r = 
-10.0$~Jy), and \textbf{the other} where the receiver \textbf{noise} dominates the total power ($\sigma_r=100.0$~Jy).
+cases, one where the receiver noise power is subdominant to the sky power ($\sigma_r = 
+10.0$~Jy), and the other where the receiver noise dominates the total power ($\sigma_r=100.0$~Jy).
 
@@
 calibration loop are created using all point sources in the simulated sky. For the single point 
-source model, we only \textbf{include the bright calibrator source in the model visibilities.}
+source model, we only include the bright calibrator source in the model visibilities.
 While we vary the model used for calibration, the simulated ``true sky'' remains fixed with all 50 
@@
 iterations of the calibration loop. 
-\textbf{This allows us to bypass the initial convergence time and quickly study the effect of the
-noise over many simuations.}
-We \textbf{again} adopt a damping factor $\damp=0.35$. Because EPICal 
+This allows us to bypass the initial convergence time and quickly study the effect of the
+noise over many simuations.
+We again adopt a damping factor $\damp=0.35$. Because EPICal 
 updates the gain estimate at each calibration loop, \tcal, but retains a memory of previous 
@@
 
-We simultaneously form simulated visibilities, $V_{ab}$, by correlating all pairs of antenna \textbf{electric field 
-measurements} (including the receiver noise). The \textbf{electric field measurements are correlated} for a duration equal to 
+We simultaneously form simulated visibilities, $V_{ab}$, by correlating all pairs of antenna electric field 
+measurements (including the receiver noise). The electric field measurements are correlated for a duration equal to 
 the effective integration time of the EPICal loop for comparison. We then find the 
@@
 
-The results of our simulations \textbf{for all four calibration methods are shown in Fig.~\ref{fig:errors}. }
+The results of our simulations for all four calibration methods are shown in Fig.~\ref{fig:errors}. 
 In the case of the full sky model, we see the errors trend 
@@
 sky model. However, we see that with this perfect model the errors do behave like noise
-\textbf{in the sense that they integrate down with longer integrations with the expected $t^{-1/2}$ 
-trend. }
-\textbf{With longer integration time (or larger damping factor), EPICal can achieve the same 
-level of gain error as visibility based solutions.}
+in the sense that they integrate down with longer integrations with the expected $t^{-1/2}$ 
+trend. 
+With longer integration time (or larger damping factor), EPICal can achieve the same 
+level of gain error as visibility based solutions.
 
@@
 average EPICal errors are 5\% lower in this regime.
-\textbf{The results from the high and low receiver noise regimes are similar, with both approaching the same floor in the single source model case.}
+The results from the high and low receiver noise regimes are similar, with both approaching the same floor in the single source model case.
 }
@@
 
-The exact ratio of \textbf{noise in gains from EPICal and the visibility-based estimator}
+The exact ratio of noise in gains from EPICal and the visibility-based estimator
 depends on the fraction of 
@@
 
-\textbf{In a more realistic situation, the observer will not have a perfectly complete model of the sky.
+In a more realistic situation, the observer will not have a perfectly complete model of the sky.
 A common method is to instead model the brightest sources in the field,
-or even a single dominant source.}
+or even a single dominant source.
 When using the single point source model, all calibration methods in Fig.~\ref{fig:errors} trend 
@@
 EPICal underperformed with the full sky model: EPICal only used a single pixel in the sky to 
-form its solutions. When using the full sky model, it was \textbf{under utilizing} the information in the 
+form its solutions. When using the full sky model, it was under utilizing the information in the 
 rest of the sky. But when the sky model only contains the bright point source, the single pixel 
@@
 the pixels with incomplete sky model. 
-\textbf{In contrast, the visibility-based calibration method 
+In contrast, the visibility-based calibration method 
 does not have any preference to a certain direction, and
@@
 of the instrument).
-This results in a more harsh penalty for missing sources in the model.}
+This results in a more harsh penalty for missing sources in the model.
 In the regime typical of HERA discussed above, EPICal's 
@@
 
-\textbf{Comparing the two regimes of receiver noise levels (red versus black lines in 
+Comparing the two regimes of receiver noise levels (red versus black lines in 
 Fig.~\ref{fig:errors}), we see that the gains errors behave as one might expect given our
@@
 because the errors are dominated by the thermal noise. But as integration time increases,
-they approach the same floor due to the same level of power from unmodeled sources.}
+they approach the same floor due to the same level of power from unmodeled sources.
 
@@
 We next demonstrate our calibration algorithm using an observation from the LWA station in 
-New Mexico \textbf{\citep{ell13}.} The data is from the LWA narrow-band transient buffer (TBN), with time ordered 
-voltage data from the 255 core antennas within a \textbf{110~m x 100~m ellipse. }
+New Mexico \citep{ell13}. The data is from the LWA narrow-band transient buffer (TBN), with time ordered 
+voltage data from the 255 core antennas within a 110~m x 100~m ellipse. 
 The central frequency is 74.03 
-MHz, with a bandwidth of 100 kHz and \textbf{sampling period} of 5.12 ms (frequency channel 
+MHz, with a bandwidth of 100 kHz and sampling period of 5.12 ms (frequency channel 
 resolution of 195.3125 Hz). For this demonstration we limit ourselves to a single polarization.
@@
 We proceed by forming model visibilities. We model only two bright objects as point sources: 
-Cyg A with flux \textbf{20,539 Jy, and Cas A with flux 19,328 Jy \citep{lan12}. }
+Cyg A with flux 20,539 Jy, and Cas A with flux 19,328 Jy \citep{lan12}. 
 Because the raw data is attenuated by the primary beam of the instrument, we also account for 
@@
 converge quickly we initialized our gain estimates at this level. 
-We also found a boost in signal to noise is achieved easily by \textbf{assuming the gains are
-constant across a range of frequencies, and averaging solutions across channels.}
-Here we average solutions across \textbf{300} channels, or about \textbf{58.6 kHz.} 
+We also found a boost in signal to noise is achieved easily by assuming the gains are
+constant across a range of frequencies, and averaging solutions across channels.
+Here we average solutions across 300 channels, or about 58.6 kHz. 
 With a fractional bandwidth $B/f_0 = 7.9 \times 10^{-4} \ll 1$, we assume a smooth bandpass 
@@
 A damping factor of $\gamma = 0.7$ was adopted.
-\textbf{The large damping factor, relative to what was used in simulations, was necessary due to
+The large damping factor, relative to what was used in simulations, was necessary due to
 the larger noise levels in the LWA data compared to the ideal simulations. This yields
 a larger effective integration time, and thus lower noise on each calibration solution
-(Eq.~\ref{eq:teff})}
+(Eq.~\ref{eq:teff})
 
@@
 and 244 were omitted.} produced unstable gain solutions, and in fact corrupted the entire array. 
-We therefore \textbf{omitted these antennas from our analysis,} resulting in a total of 248 antennas to 
+We therefore omitted these antennas from our analysis, resulting in a total of 248 antennas to 
 calibrate. In each calibration loop, we form $C_{i,\hat{\boldsymbol{s}}_0}, i=1,2,\ldots N_a$ and 
-update our gain estimates over 10 \textbf{samples} (51.2 ms). We iterate the loop 30 times for a 
+update our gain estimates over 10 samples (51.2 ms). We iterate the loop 30 times for a 
 total of 1.536 seconds of data processed. The results of this calibration experiment are shown in Figs.~
@@
 iterations, again with each colored line representing a different antenna. Given the quality of 
-uncalibrated image demonstrated in \cite{thy15c}, \textbf{we had anticipated that the phase
-solutions would not exhibit much variation between antennas. Hence the relatively large variation was unexpected.}
+uncalibrated image demonstrated in \cite{thy15c}, we had anticipated that the phase
+solutions would not exhibit much variation between antennas. Hence the relatively large variation was unexpected.
 However, the phases are relatively flat after about 15 iterations 
@@
 not surprising given the fairly good quality image produced from uncalibrated data.
-\textbf{For plotting clarity, we unwrapped phases resulting in phases that appear to exceed 
-$\pm \pi$.}
+For plotting clarity, we unwrapped phases resulting in phases that appear to exceed 
+$\pm \pi$.
 }
@@
 
-\textbf{To ensure the level of noise in our solutions is reasonable we compared to the fractional errors
+To ensure the level of noise in our solutions is reasonable we compared to the fractional errors
 from our simulations shown in Fig.~\ref{fig:errors}. We first use the Global Sky Model
@@
 should therefore expect our fractional gain noise levels to be at best around 0.1, the level 
-to which the simulations asymptotically approached. }
-
-\textbf{To estimate the fractional gain noise of our solutions we calculate the standard deviation of the gain
+to which the simulations asymptotically approached.
+
+To estimate the fractional gain noise of our solutions we calculate the standard deviation of the gain
 estimates after iteration 15 for each antenna separately. We then divide these standard
@@
 ($\sigma_g \approx 0.29$). Upon closer inspection the gain amplitudes of these outliers had
-converged later in the calibration loop, after iteration 15.}
+converged later in the calibration loop, after iteration 15.
 
@@
 \includegraphics[width=\columnwidth]{fig8.pdf}
-\caption{\textbf{Same as Fig.~\ref{fig:data_phase}, but for gain amplitudes.}
+\caption{Same as Fig.~\ref{fig:data_phase}, but for gain amplitudes.
 The 
@@
 Figure~\ref{fig:data_images} shows the improvement in the images due to our calibration. The 
-left panel shows the uncalibrated image integrated over 51.2 ms, \textbf{58.6 kHz.} Cyg~A is prominent 
+left panel shows the uncalibrated image integrated over 51.2 ms, 58.6 kHz. Cyg~A is prominent 
 near the center of the image, and Cas~A is also clearly visible in the upper right. The middle 
 panel shows the image produced after calibration with identical integration time and bandwidth. 
-\textbf{We show each of these images with a color scale which saturates at half the
+We show each of these images with a color scale which saturates at half the
 maximum value in the respective image data. We find that this choice allows us to
 qualitatively compare the quality of the images, independent of the absolute scale which
-changed during the calibration.}
-The sidelobes throughout the \textbf{calibrated} image are significantly suppressed, and the galactic plane is 
+changed during the calibration.
+The sidelobes throughout the calibrated image are significantly suppressed, and the galactic plane is 
 much more evident, despite only modeling Cyg~A and Cas~A. We also note that the feature 
@@
 from the GSM. 
-\textbf{For reference we show the GSM in the right panel of 
+For reference we show the GSM in the right panel of 
 Fig.~\ref{fig:data_images}, convolved by the LWA point spread function and weighted
-by  two factors of the primary beam to match the holographic frame of the data images.}
+by  two factors of the primary beam to match the holographic frame of the data images.
 
@@
 \caption{Images produced before (left) and after (middle) calibrating LWA data. These images 
-were produced with \textbf{58.6 kHz} bandwidth and 51.2 ms integration. 
-\textbf{The color scales are saturated at half the peak of the respective images.}
+were produced with 58.6 kHz bandwidth and 51.2 ms integration. 
+The color scales are saturated at half the peak of the respective images.
 The calibrated image shows 
-significant reduction in \textbf{rumble} throughout, while retaining the prominent Cyg~A and 
+significant reduction in rumble throughout, while retaining the prominent Cyg~A and 
 Cas~A sources. The galactic plane is also substantially more evident after calibration. For 
-reference, the GSM is shown \textbf{in the 
+reference, the GSM is shown in the 
 right panel convolved with the LWA point spread function and weighted by two factors
-of the primary beam to place it in the same holographic frame as the data images.} While much of the GSM is visible in the calibrated image, we 
+of the primary beam to place it in the same holographic frame as the data images. While much of the GSM is visible in the calibrated image, we 
 note that the model used to calibrate was only the two bright sources, Cyg~A and Cas~A.
@@
 peak of the image over the noise level. We estimate the noise level as the median of the 
-absolute deviation of the image. With this metric we find the \textbf{uncalibrated 
-and} calibrated images to have \textbf{dynamic ranges of 80.4 and 133.1, respectively, a 65\% 
-improvement.}
+absolute deviation of the image. With this metric we find the uncalibrated 
+and calibrated images to have dynamic ranges of 80.4 and 133.1, respectively, a 65\% 
+improvement.
 
@@
 ``estimate gains" box in Fig.~\ref{fig:schematic}, which can be performed at a much 
-\textbf{slower rate} than the actual antenna-pixel correlations.
+slower rate than the actual antenna-pixel correlations.
 
@@
 Additionally, with knowledge of 
-\textbf{any filters applied to the data stream in the receiver chain,}
+any filters applied to the data stream in the receiver chain,
 we could include channels closer to the edge of the band.
@@
 both compact and diffuse sources over a large patch of sky. 
-\textbf{As with any sky-based calibration scheme, EPICal requires that the sky model be attenuated
-by the primary beam,} which 
+As with any sky-based calibration scheme, EPICal requires that the sky model be attenuated
+by the primary beam, which 
 can be difficult to measure at the precision necessary \citep[e.g.][]{neb15,vir14,thy15b}. 
-However, the direct imaging correlator provides exactly the \textbf{product of the sky and beam} necessary, and 
-can be iterated over to improve the images and gain solutions, \textbf{leaving final beam
-correction to post-processing as in traditional processing.}
+However, the direct imaging correlator provides exactly the product of the sky and beam necessary, and 
+can be iterated over to improve the images and gain solutions, leaving final beam
+correction to post-processing as in traditional processing.
 
@@
 increase the scope. With next generation instruments in the planning and development stages, 
-EPICal is poised to \textbf{enable new design spaces}. The software is integrated into the EPIC 
-package and freely available to use in simulations or \textbf{post-processing} of data. Work is 
+EPICal is poised to enable new design spaces. The software is integrated into the EPIC 
+package and freely available to use in simulations or post-processing of data. Work is 
 underway to port the code to real-time GPU systems for deployment.
